### 2014-10-11 Vragenlijst1_Sophie_Marien.tex
tekst toegevoegd

\documentclass[a4paper, 11pt]{article}
\setlength{\topmargin}{-.5in}
\setlength{\textheight}{9in}
\setlength{\oddsidemargin}{.100in}
\setlength{\textwidth}{6.25in}
\usepackage [dutch] {babel}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{fullpage}
\usepackage{listings}
\usepackage{color}
\usepackage{soul}
\usepackage{tikz}
\usepackage{hyperref}
\usepackage{parskip}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\lstset{frame=tb,
  language=Java,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=true
  tabsize=3
}

\newcommand{\todo}[1]{{\huge \textcolor{green}{#1}}\\}
\newcommand{\comment}[1]{\textcolor{red}{#1}\\}
\newcommand{\viscomment}[1]{\textcolor{red}{#1}}

\begin{document}

\begin{titlepage}
\title{Vragenlijst thesis evaluatie}
\author{Sophie Marien}
\end{titlepage}

\maketitle

\section{Wat is de doelstelling van je thesis? (beschrijf deze doelstelling met eigen woorden in min 5, max 10 lijnen)}
Mijn thesis gaat over gametheorie en cybersecurity. Gametheorie is de studie van een spel waarbij er een interactie is tussen twee of meerdere spelers en de interactie die zal gebeuren afhangt van wat elke speler zal doen. Elke speler heeft een level van geluk voor de verschillende acties die gedaan kunnen worden. Met gametheory kunnen er security problemen gemodelleerd worden en kan er gekeken worden wat de best volgende zet is als defender van een systeem dat beschermd moet worden.

\section{Wat zijn de belangrijkste problemen die je moet oplossen? (max 5 lijnen)}
Een interessante link vinden tussen gametheory en cybersecurity die goed genoeg is om een thesis over te schrijven.
\section{Wat heb je tot op heden gedaan? (literatuur (wat precies?), leren werken met tools (welke?), analyse (van wat?), al iets ontworpen en/of geïmplementeerd (wat precies?), ...)}
Vooral literatuur doorgenomen en over gametheory gelezen en ingewerkt. Momenteel ben ik begonnen aan de coursera cursus over gametheory om dit topic volledig door en door te kennen. Via de infosessie heb ik ook leren werken met EndNote om een lijst bij te kunnen houden van al de literatuur die ik al heb doorgenomen. De literatuur die ik heb doorgenomen waren kleine papers over gametheory en cybersecurity en een grotere paper over FlipIT.
\section{Wat heb je gepland om te doen tijdens de volgende 4 weken?}
Volgende weken heb ik ingepland om mijn verloren uren van de eerste twee weken in te halen. Ik ga vooral nog verder lezen over gametheory en de lessen verder volgen van de coursera course. Ook moet ik mijn hypothese kunnen opstellen en duidelijk afbakenen. Daarna kan ik me dieper inwerken. 
\section{Tegen wanneer plan je om je thesis in te dienen (juni, september, januari)?}
Ik plan mijn thesis tegen Juni in te dienen.
\section{Stuur je ISP naar je begeleider en je promotor, en geef aan hoeveel examens je in de januari-zittijd en in de juni-zittijd zult hebben.}
In Januari zal ik drie examens hebben en in Juni zittijd maar een die ook twee weken na de thesis deadline zal vallen.
\section{Evalueer de tijd die je in je thesis gestoken hebt in de voorbije 4 weken: veel te weinig / te weinig / voldoende / ruim voldoende / meer dan verwacht.}
Voor de voorbije 4 (3?) weken heb ik er veel te weinig tijd ingestoken. Ik zit nog maar aan 25 uur. Ik ben twee weken in het buitenland geweest in de eerste twee weken van het academie jaar en ik was me ervan bewust dat ik al deze tijd terug zal moeten inhalen. De komende weken ga ik proberen om deze uren in te halen.
\section{Evalueer de vorderingen van je thesis t.o.v. het bereiken van de vooropgestelde doelstelling: ver achter op het schema / achter op het schema / goed op schema / vooruit op het schema / ver vooruit op het schema.}
Omdat ik wist dat ik de eerste weken niet aan mijn thesis zou kunnen werken zit ik op schema, maar dat betekent niet dat ik nu al goed bezig ben. 
\section{Stuur de ingevulde timesheet mee als bijlage.}
Ik kon de eerste vergadering niet komen en weet dus niet hoe ik de timesheet moet invullen. Ik heb een eigen timesheet aangemaakt om mijn uren bij te houden.
Voorlopig kom ik dus aan 24.15 uur voor deze week (weekend zit hier nog niet bij in).
\end{document}

### 2015-01-07 Thesisfollowup/ThesisFollowUpSophieMarien.tex
thesis followup

\documentclass[a4paper, 11pt]{article}
\setlength{\topmargin}{-.5in}
\setlength{\textheight}{9in}
\setlength{\oddsidemargin}{.100in}
\setlength{\textwidth}{6.25in}
\usepackage [dutch] {babel}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{fullpage}
\usepackage{listings}
\usepackage{color}
\usepackage{soul}
\usepackage{tikz}
\usepackage{hyperref}
\usepackage{parskip}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\lstset{frame=tb,
  language=Java,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=true
  tabsize=3
}

\newcommand{\todo}[1]{{\huge \textcolor{green}{#1}}\\}
\newcommand{\comment}[1]{\textcolor{red}{#1}\\}
\newcommand{\viscomment}[1]{\textcolor{red}{#1}}
\newcommand{\flip}[1] {\textcolor{black}{#1}}

\begin{document}

\begin{titlepage}
\title{Thesis Follow Up}
\author{Sophie Marien}
\date{} 
\end{titlepage}

\maketitle

\section{Geef een uitgewerkt werkplan voor je thesis-activiteiten tussen nu en half juni (dan moet je je masterproef indienen). Dit werkplan moet PER TWEE WEKEN aangeven wat je precies gaat doen voor je masterproef. Maak het heel CONCREET: in het geval van implementatie, implementatie van welk deel precies, in het geval van het schrijven van tekst, het schrijven van welk hoofdstuk precies, e.d. Stel dit werkplan zelfstandig op, en bespreek het daarna met je begeleider, voordat je op deze mail antwoordt.}

Uitgewerkt plan steekt bij in de bijlage. In de timesheet file staat naast de werkuren aan wat ik ga werken.

\section{Schrijf een stuk tekst van minstens 5 pagina's waarin je de probleemstelling van je masterproef formuleert, gevolgd door de doelstellingen. Je verwerkt hier onder meer elementen van je literatuurstudie in.}

Ik heb afgesproken dat ik het stuk tekst tegen dit weekend uitschrijf. Hier is al een kleine inleiding.

Virussen zijn een groot probleem voor bedrijfsnetwerken. Ze kunnen gevoelige informatie verzamelen of een bedrijfsnetwerk platleggen. Gegeven de grote kost gebonden aan schade door malware, is het vinden van de juiste verdegingingsstrategie belangrijk. Het aanvallen en verdedigen van een bedrijfsnetwerk kan gezien worden als een spel, waarbij de verdediger en de aanvaller elk proberen de beste strategie te vinden.  Ik ga dus toelichten hoe het spel van aanvallen en verdedigen kan gedefinieerd worden als een variatie op het spel FlipIt. Dit laat toe om te onderzoeken wat de verschillende strategieën zijn van de netwerkbeheerder enerzijds en van de aanvaller die virussen zendt, anderzijds. Bedoeling is om in een volgende stap het spel verder te analyseren met behulp van Game Theory om te bepalen welke de dominerende strategieën zijn en of er zich Nash equilibria voordoen.

\section{Evalueer de tijd die je in je thesis gestoken hebt in de voorbije 4 weken: veel te weinig / te weinig / voldoende / ruim voldoende / meer dan verwacht.}

Ik heb voldoende tijd in mijn thesis gestoken. Ik heb de timesheet aangepast naar mijn planning. Ik heb weinig examens en ben een week eerder klaar dan de lesvrije week dus dan kan ik nog verder werken aan mijn thesis. Voor deze komende weken heb ik minder uren ingepland omdat ik van andere vakken projecten moet afwerken en presentaties moet geven. De laatste week heb ik minder kunnen werken dan dat ik graag had willen doen omwille van een kaakgewrichtsontsteking die momenteel nog steeds niet over is. De volgende weken ga ik proberen meer te werken aan mijn thesis dan dat ik gepland heb, om mijn uren in te halen.

\section{Evalueer de vorderingen van je thesis t.o.v. het bereiken van de vooropgestelde doelstelling: ver achter op het schema / achter op het schema / goed op schema / vooruit op het schema / ver vooruit op het schema.}

Ik sta op schema volgens het schema dat ik zelf gemaakt heb. Ik heb mijn schema voor de komende weken aangepast.

\section{Stuur de ingevulde timesheet mee als bijlage}
\end{document}

### 2015-01-07 LatexVerslag/ThesisNotities.tex
thesis followup

\documentclass[a4paper, 11pt]{article}
\setlength{\topmargin}{-.5in}
\setlength{\textheight}{9in}
\setlength{\oddsidemargin}{.100in}
\setlength{\textwidth}{6.25in}
\usepackage [dutch] {babel}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{fullpage}
\usepackage{listings}
\usepackage{color}
\usepackage{soul}
\usepackage{tikz}
\usepackage{hyperref}
\usepackage{parskip}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\lstset{frame=tb,
  language=Java,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=true
  tabsize=3
}

\newcommand{\todo}[1]{{\huge \textcolor{green}{#1}}\\}
\newcommand{\comment}[1]{\textcolor{red}{#1}\\}
\newcommand{\viscomment}[1]{\textcolor{red}{#1}}
\newcommand{\flip}[1] {\textcolor{black}{#1}}

\begin{document}

\begin{titlepage}
\title{Vragen OSS}
\author{Sophie Marien}
\end{titlepage}

\maketitle
\tableofcontents
\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%																					%%%%%%
%%%%%					FLipIt with Multiple resources									%%%%%%
%%%%%																					%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Abstract}
\label{Section:Abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%																					%%%%%%
%%%%%					FLipIt with Multiple resources									%%%%%%
%%%%%																					%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{Section:Introduction}
important for economics, psygology, computer science, ..
How should these interactions be structured
"Game is an interaction between two or more people, where the outcome of the interactions depends on what everybody does, and everybody has different levels of happiness for the different outcomes"

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%																					%%%%%%
%%%%%					Intoruction to FlipIt 											%%%%%%
%%%%%																					%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{FLipIt introduction}

In this section, we introduce the game \flip{FlipIt}. \flip{FlipIt} is a two-players game with a shared (single) resource that the players want to control as long as possible. The shared resource can be a password, a network or a secret key depending on the setting being modeled. In the rest of the paper we will call the players the Attacker and the Defender. To get the control over the resource, players can flip the resource at any given time. Each move will imply a certain cost. The unique feature of \flip{FlipIt} is that the move will happen in a stealthy way, meaning that the other player has no clue that the other player has flipped the resource. For instance, the defender will not find out if the resource has already been compromised by the attacker, but he can only potentially know it after he flips the resource himself. The goal of the player is to maximize the time that he or she has control over the resource while minimizing total cost of the moves. Players won't move to frequently. A move can also result in a "wasted move", in this paper called flop. It could be that the resource was for example already under control of the defender. If the defender moves when he or she has already the control over the resource, he or she would have a wasted move because it would not result in a change of ownership.

Because the players move in a sthealty way, there are different types of feedback that a player can get while moving:
\begin{itemize}
\item Non-adaptive (NA): The player does not receive any feedback while flipping.
\item Last move (LM): The player finds out the exact time the opponent played the last time.
\item Full History (FH): The player finds out the complete history of the opponents move.
\end{itemize}
The game can be "uitgebreid" by the amount of information that a player receives. It can also be possible for a player to get information at the start of the game. Both interesting cases are:
\begin{itemize}
\item Rate-of-play (RP: The player finds out the exact rate of play of the opponent.
\item Knowledge-of-strategy (KS): The player finds out the complete information of the strategy that the opponent is playing.
\end{itemize}

In our "aanname" will the strategy of both players be non-adaptive. Non of the players has information of the strategy of the opponent.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%																					%%%%%%
%%%%%					FLipIt with Multiple resources									%%%%%%
%%%%%																					%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Changing FlipIt to multiple resources}
\label{Section:FLipItmulti}
In this section we will discuss the changes made to \flip{FlipIt} to make a model to model multiple resources. 
\flip{FlipIt} is a framework for a single shared resource and two players, the defender and the attacker. 
\end{document}

### 2015-01-07 doc/template/chap-1.tex
meer papers

%\chapter{Intoduction to GameTheory}
%\label{cha:1}
\documentclass[10pt]{article}
\begin{document}

In this chapter an introduction to gametheory will be given with the formulas that will be used troughout this paper. We start with the basics of gametheory. People that have a background in gametheory can skip this chapter.

\section{Virusses}

Many network security threats today are spread over the Internet. The most common include:

Viruses, worms, and Trojan horses
Spyware and adware
Zero-day attacks, also called zero-hour attacks
Hacker attacks
Denial of service attacks
Data interception and theft
Identity theft

%http://www.ists.dartmouth.edu/library/258.pdf Email Virus Propagation Modeling and Analysis
%Cliff C. Zou∗, Don Towsley†, Weibo Gong∗
%∗Department of Electrical & Computer Engineering
%†Department of Computer Science
%Univ. Massachusetts, Amherst
%Technical Report: TR-CSE-03-04

Computer virus through mail. 
Though virus spreading through email is an old technique, it is still effective and is widely used by
current viruses and worms. Sending viruses through email has some advantages that are attractive to
virus writers:
 Sending viruses through email does not require any security holes in computer operating systems
or software.
 Almost everyone who uses computers uses email service.
 A large number of users have little knowledge of email viruses and trust most email they receive,
especially email from their friends [28][29].
 Email are private properties like post office letters. Thus correspondent laws or policies are required
to permit checking email content for detecting viruses before end users receive email [18].

Send a email with malicious attachment. Only again infected if attachment again opened. Thus this is the action of attacking every neighbour node + also can attack again the node where the virus was coming from.
There are also email viruses were the malicious program is hidden in the txt and the attachment does not need to be opened.

%http://www.cisco.com/web/offer/gist_ty2_asset/Cisco_2014_ASR.pdf p49
%http://repo.hackerzvoice.net/depot_madchat/vxdevl/papers/avers/2004-35.pdf
%http://www.mcafee.com/us/resources/white-papers/foundstone/wp-managing-malware-outbreak.pdf

\subsection{What are my topics}
\begin{itemize}
\item Security, Costs, Cybersecurity
\item Viruses, kinds
\item Gametheory
\item Flip-it
\item Flip-it multiple resources
\item 
\end{itemize}

\subsection{Malware}
%Does a company network faces lot of malware? what is the cost ?
Relevant researches:
\begin{itemize}
%
\item 
\end{itemize}

\section{Conclusion}
The final section of the chapter gives an overview of the important results
of this chapter. This implies that the introductory chapter and the
concluding chapter don't need a conclusion.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "thesis"
%%% End:

\end{document}

### 2015-01-25 doc/template/chap-3.tex
Literatuurstudie + eerste schrijfwerk over flipit

%\chapter{Intoduction to GameTheory}
%\label{cha:1}
\documentclass[10pt]{article}
\begin{document}

\section{Literatuurstudie}
\begin{description}
\item Distributed Worm Simulation with a REalistic Internet 2005  \\
Modeling of congestions of network through worm propagation. Mathematical model focussing on the underlying network infrastructure.(diff no game theory) 
\item Of threats and Costs: A Game-theoretic approach to security risk managment 2013\\
Model network security of networks with a non-cooperative node through game theory. Attacker knows the defense strategies and the defender has knowledge of the possible attacks. Each actor considers the actions of the other before deciding to strive to optimize their own utility. (diff not stealthy)
\item Game theory meet network security and privacy (2013) \\
Chapter 3 addresses several games in game theory for modeling network security.
\item Game theoretic approach for cost-benefit analysis of malware proliferation prevention (..)
Introduces SIS and SIR together with 'patch', 'removal' and 'patch and removal'.
\end{description}

\subsection{What can be done in further research}
\begin{itemize}
\item Looking for the dynamics of the spread of the virus/worm limited by the bandwidth of the network links, BPG routing failure with high volume scan traffic
\end{itemize}
\section{Conclusion}

\section{Why Game Theory to model security problem}
Actors in a security protocol must follow the systems and some arbitrarily actors that are malisious and do not follow the protocol. [Bridging GameTheory and Cryptography]. Gametheoretic approach proposes a model where all the actors act with self-interest.

\section{..}

Flip-it. Some authors have written other papers about flipit. One of them is the [Game theoretic approach for cost-benefit analysis of malware proliferation prevention].

Company networks are targeted. It costs a lot. They want to defend their company networks. No loss of data, integrity and confidentiality.  Many ways to attack a company. Virusses, trojans, worms, DOS, .. Hard to protect against every attacker.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "thesis"
%%% End:

\end{document}

### 2015-01-25 doc/template/chap-4.tex
Literatuurstudie + eerste schrijfwerk over flipit

--- 2015-01-25 doc/template/chap-3.tex
@@
 
-\section{Literatuurstudie}
-\begin{description}
-\item Distributed Worm Simulation with a REalistic Internet 2005  \\
-Modeling of congestions of network through worm propagation. Mathematical model focussing on the underlying network infrastructure.(diff no game theory) 
-\item Of threats and Costs: A Game-theoretic approach to security risk managment 2013\\
-Model network security of networks with a non-cooperative node through game theory.
-Attacker knows the defense strategies and the defender has knowledge of the possible attacks.
-Each actor considers the actions of the other before deciding to strive to optimize their own utility.
-(diff not stealthy)
-\item Game theory meet network security and privacy (2013) \\
-Chapter 3 addresses several games in game theory for modeling network security.
-\item Game theoretic approach for cost-benefit analysis of malware proliferation prevention (..)
-Introduces SIS and SIR together with 'patch', 'removal' and 'patch and removal'.
-\end{description}
+\section{Write down the settings of the game}
 
-\subsection{What can be done in further research}
-\begin{itemize}
-\item Looking for the dynamics of the spread of the virus/worm limited by the bandwidth of the network links, BPG routing failure with high volume scan traffic
-\end{itemize}
-\section{Conclusion}
-
-\section{Why Game Theory to model security problem}
-Actors in a security protocol must follow the systems and some arbitrarily actors that are malisious and do not follow the protocol.
-[Bridging GameTheory and Cryptography].
-Gametheoretic approach proposes a model where all the actors act with self-interest.
-
-\section{..}
-
-Flip-it. Some authors have written other papers about flipit. One of them is the [Game theoretic approach for cost-benefit analysis of malware proliferation prevention].
-
-Company networks are targeted.
-It costs a lot.
-They want to defend their company networks.
-No loss of data, integrity and confidentiality.
-Many ways to attack a company.
-Virusses, trojans, worms, DOS, ..
-Hard to protect against every attacker.
+What is different from Flip-It.
 

### 2015-02-13 doc/template/chap-3.tex
commit

@@
 
+\section{Gametheory}
+
+Game theory is the study between selfinterested agents and the actions they perform.
+self interested agents means not that they want to harm eachother or that they only care about themselfes.
+Agent has a description of the world he likes.
+A utility function is a mapping of the world that the agent likes to real numbers.
+
 \section{Literatuurstudie}
+
+Difference with FlipThem: subpart of nodes for control and strategy difference: dependant of grade of the nodes, instead of just periodic.
+- Literatuurstudie
+- Flipit
+- Game Motivation
+- Formal definition
+
 \begin{description}
-\item Distributed Worm Simulation with a REalistic Internet 2005  \\
+\item Distributed Worm Simulation with a Realistic Internet 2005  \\
 Modeling of congestions of network through worm propagation. Mathematical model focussing on the underlying network infrastructure.(diff no game theory) 

### 2015-02-14 doc/template/chap-1.tex
meer papers
referenties bibtex aanvullen

@@
-%\chapter{Intoduction to GameTheory}
-%\label{cha:1}
-\documentclass[10pt]{article}
-\begin{document}
+\chapter{Intoduction to GameTheory}
+\label{cha:1}
+%\documentclass[10pt]{article}
+%\begin{document}
 
-In this chapter an introduction to gametheory will be given with the formulas that will be used troughout this paper.
+In this chapter an introduction to gametheory will be given with the formulas that will be used throughout this paper.
 We start with the basics of gametheory.
 People that have a background in gametheory can skip this chapter.
+\section{Intro to Gametheory}
+
+Game Theory is the study of interactions between independent and self-interested agents.
+It is a set of tools to analyze situations in which the optimal action for an agent depends on anticipations that form over another agent decision \todo{citatie vanuit wikipedia}.
+An agent can be either an individual, an animal or an enterprise.
+Game theory can be applied on various domains such as politics, economics, biology, sociology.
+In game theory studies the strategic interaction between players in a game.
+A game is played by a number of players and each of these players has actions.
+These actions are represented by a number indicating their preference.
 
@@
 \begin{itemize}
-%
-\item 
+%http://ants.iis.sinica.edu.tw/3BkMJ9lTeWXTSrrvNoKNFDxRm3zFwRR/17/04483668.pdf
+\item How Viruses and worm can be detected. Difference between UDP en TCP worm propagation
 \end{itemize}
@@
 
-\end{document}
+%\end{document}

### 2015-02-15 doc/template/chap-1.tex
intro gametheory

--- 2015-01-07 doc/template/chap-1.tex
@@
-%\chapter{Intoduction to GameTheory}
-%\label{cha:1}
-\documentclass[10pt]{article}
-\begin{document}
+\chapter{Intoduction to GameTheory}
+\label{cha:1}
+%\documentclass[10pt]{article}
+%\begin{document}
 
+\section{Intro Game Theory}
+
+%begin over dat gametheorie handig is in de economie
+In the following paragraph an introduction to game theory is given based on the work of \todo{werk van Essentials of Game Theory} and \todo{Coursera game theory}.
+Game theory studies the interaction between independent and self-interested agents.
+For this reason it is very important for economics and also for politics, biology, computer science, philosophy and a variety of other disciplines.
+It is a mathematical way of modelling the interactions between two or more agents where the outcome depends on what everybody does and how it should be structured to lead to good outcomes.
+\\
+%Every agent has different levels of happiness for the different outcomes.
+A self-interested independent agent does not necessarily mean that they want to harm other agents or that they only care about themselves.
+Each agent has preferences about the states of the world he likes.
+These preferences are mapped to natural numbers and are called the utility function.
+The numbers are interpreted as a mathematical measure to tell you how much an agent likes or dislikes the states of the world.
+It also explains the impact of uncertainty.
+When an agent is uncertain about a distribution of outcomes, his utility will describe the expected value of the utility function with respect to the probability of the distribution of the outcomes.
+For example: with 0.7 probability it will be 7 degrees outside and 0.3 probability it will be 10 degrees.
+The agent can have a different opinion about that distribution versus another distribution.
+(\todo{uitleggen aan de hand van een voorbeeld}).\\
+
+A game in game theory consists of multiple agents and every agent has a set of actions that he can play. 
+%self interested meaing
+
+%utility function meaning
+
+%Cooperative and non cooperative games
+
+% example of a game
+
+\section{Other}
 In this chapter an introduction to gametheory will be given with the formulas that will be used troughout this paper.

### 2015-02-15 doc/template/chap-1.tex
tadaa

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%			Introduction Chapter 1				%%%%%%
+%%%%%												%%%%%%
+%%%%%												%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+In the following paragraph an introduction to game theory is given based on the work of \todo{werk van Essentials of Game Theory} and \todo{Coursera game theory}.
+Voor een meer gedetailleerdere intro ook dit book.
+%------------------------------------------------%
+%            Intro Game Theory 					 %
+%------------------------------------------------%
 \section{Intro Game Theory}
+\label{Cha:1:Intro.Game.Theory}
 
 %begin over dat gametheorie handig is in de economie
-In the following paragraph an introduction to game theory is given based on the work of \todo{werk van Essentials of Game Theory} and \todo{Coursera game theory}.
+
 Game theory studies the interaction between independent and self-interested agents.
@@
 %Every agent has different levels of happiness for the different outcomes.
+%self interested meaning
 A self-interested independent agent does not necessarily mean that they want to harm other agents or that they only care about themselves. 
+%utility function meaning
 Each agent has preferences about the states of the world he likes.
@@
 (\todo{uitleggen aan de hand van een voorbeeld}).\\
+%Cooperative and non cooperative games
+In a decision game theoretic approach an agent will try to act in such a way to maximise his expected or average utility function.
+It becomes more complicated when two or more agents want to maximise their utility and whose actions can affect each other utilities.
+This kind of games are referred to as non cooperative game theory, where the basic modelling unit is the group of agents.
+The individualistic approach, where the basic modelling is only one agent, is referred as cooperative game theory.
+
+In the following list a couple of terms that will be used throughout the paper.
+\begin{description}
+\item \textit{Players}: players are referred as the ones who are the decision makers. It can be a person, a company or an animal.
+\item \textit{Actions}: actions are what the player can do. 
+\item \textit{Outcomes}:  
+\item \textit{Utility function}: the utility function is the mapping of the level of happiness of an agent about the state of the world to natural numbers.
+\item \textit{Strategies}: A strategy is the combination of different actions. A pure strategy is only one action.
+\end{description}
 
 A game in game theory consists of multiple agents and every agent has a set of actions that he can play.
-%self interested meaing
 
-%utility function meaning
-
-%Cooperative and non cooperative games
-
-% example of a game
+% --------------- example of a game -----------------%
 
@@
 
+%------------------------------------------------%
+%            Intro about virusses				 %
+%------------------------------------------------%
 \section{Virusses}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%			Conclusion Chapter 1				%%%%%%
+%%%%%												%%%%%%
+%%%%%												%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusion}
+\label{Cha:1:Conclusion}
 The final section of the chapter gives an overview of the important results

### 2015-02-15 doc/template/chap-1.tex
aanvulling intro gametheory

@@
 Game theory studies the interaction between independent and self-interested agents.
+It is a mathematical way of modelling the interactions between two or more agents where the outcomes depend on what everybody does and how it should be structured to lead to good outcomes.
 For this reason it is very important for economics and also for politics, biology, computer science, philosophy and a variety of other disciplines.
-It is a mathematical way of modelling the interactions between two or more agents where the outcome depends on what everybody does and how it should be structured to lead to good outcomes.
 \\
@@
 %self interested meaning
-A self-interested independent agent does not necessarily mean that they want to harm other agents or that they only care about themselves. 
+
+One of the assumptions underlying game theory is that the players of the game, the agents, are independent and self-interested.
+This does not necessarily mean that they want to harm other agents or that they only care about themselves.
 %utility function meaning 
-Each agent has preferences about the states of the world he likes.
+Instead it means that each agent has preferences about the states of the world he likes.
 These preferences are mapped to natural numbers and are called the utility function.
 The numbers are interpreted as a mathematical measure to tell you how much an agent likes or dislikes the states of the world.
+\\
 It also explains the impact of uncertainty.
@@
 The individualistic approach, where the basic modelling is only one agent, is referred as cooperative game theory.
+
+There are two standard representations for games. The first one is the Normal Form. The second one is the Extensive Form.
 

### 2015-02-15 doc/template/chap-2.tex
aanvulling intro gametheory

\chapter{The FlipIt game}
\label{cha:2}

\section{Extensions on FlipIt}

There a various possible ways to extend \flip{FlipIt}. For instance Laszka et al. extended the basic \flip{FlipIt} game to multiple resources. The incentive is that for compromising a system in a real case it needs more than just taking over one resource. An example is gaining access to a system and breaking the password. The model is called FlipThem \cite{FlipThem}. To ways of flipping the resources are used: the AND and the OR control model. In the AND model the attacker only controls the system if he controls all the resources of the system, whereas in the OR model the attacker only needs to compromise one resource to be in control of the entire system. The difference with FlipThem and this paper is that we introduce a Graph Model in the beginning.
Another extention on FlipIt is done by Pham [\todo{citatie needed voor Are We Compromised?}]. Beside the action Flip their is another action Test. The basic idea is to test with an extra action if the resource has been compromised or not. This action involves also an extra cost. This model is useful if somebody wants to know for example if his password has been compromised or wants to assess the periodic security of a system.

In this section, we introduce the game \flip{FlipIt} \cite{FlipIt}. \flip{FlipIt} is a game introduced by .. .. and Rivest. First we explain the framework of FlipIt and after that the formulas and aannames that we will make for the game for during the whole paper.

\section{The First Topic of this Chapter}
\flip{FlipIt} is a two-players game with a shared (single) resource that the players want to control as long as possible. The shared resource can be a password, a network or a secret key depending on the setting being modeled. In the rest of the paper we will call the players the Attacker and the Defender. To get the control over the resource, players can flip the resource at any given time. Each move will imply a certain cost. The unique feature of \flip{FlipIt} is that the move will happen in a stealthy way, meaning that the other player has no clue that the other player has flipped the resource. For instance, the defender will not find out if the resource has already been compromised by the attacker, but he can only potentially know it after he flips the resource himself. The goal of the player is to maximize the time that he or she has control over the resource while minimizing total cost of the moves. Players won't move to frequently. A move can also result in a "wasted move", in this paper called flop. It could be that the resource was for example already under control of the defender. If the defender moves when he or she has already the control over the resource, he or she would have a wasted move because it would not result in a change of ownership.

Because the players move in a sthealty way, there are different types of feedback that a player can get while moving:
\begin{itemize}
\item Non-adaptive (NA): The player does not receive any feedback while flipping.
\item Last move (LM): The player finds out the exact time the opponent played the last time.
\item Full History (FH): The player finds out the complete history of the opponents move.
\end{itemize}
The game can be extended by the amount of information that a player receives. It can also be possible for a player to get information at the start of the game. Both interesting cases are:
\begin{itemize}
\item Rate-of-play (RP: The player finds out the exact rate of play of the opponent.
\item Knowledge-of-strategy (KS): The player finds out the complete information of the strategy that the opponent is playing.
\end{itemize}

In our assumption will the strategy of both players be non-adaptive. Non of the players has information of the strategy of the opponent.

\subsection{An item}
A master thesis is never an isolated work. This means that your text must
contain references. On-line documents\cite{FlipThem} as well as
books\cite{DefendingAgainstUnknownEnemy} can be referenced. \cite{Craig2005}.

\section{Figures}
Figures are used to add illustrations to the text. The \fref{fig:logo} shows
the KU~Leuven logo as an illustration.
\begin{figure}
  \centering
  \includegraphics{logokul}
  \caption{The KU~Leuven logo.}
  \label{fig:logo}
\end{figure}

\section{Tables}
Tables are used to present data neatly arranged. A table is normally
not a spreadsheet! Compare \tref{tab:wrong} en \tref{tab:ok}: which table do
you prefer?

\begin{table}
  \centering
  \begin{tabular}{||l|lr||} \hline
    gnats     & gram      & \$13.65 \\ \cline{2-3}
              & each      & .01 \\ \hline
    gnu       & stuffed   & 92.50 \\ \cline{1-1} \cline{3-3}
    emu       &           & 33.33 \\ \hline
    armadillo & frozen    & 8.99 \\ \hline
  \end{tabular}
  \caption{A table with the wrong layout.}
  \label{tab:wrong}
\end{table}

\begin{table}
  \centering
  \begin{tabular}{@{}llr@{}} \toprule
    \multicolumn{2}{c}{Item} \\ \cmidrule(r){1-2}
    Animal    & Description & Price (\$)\\ \midrule
    Gnat      & per gram    & 13.65 \\
              & each        & 0.01 \\
    Gnu       & stuffed     & 92.50 \\
    Emu       & stuffed     & 33.33 \\
    Armadillo & frozen      & 8.99 \\ \bottomrule
  \end{tabular}
  \caption{A table with the correct layout.}
  \label{tab:ok}
\end{table}

\section{Lorem Ipsum}
This section is added to check headers and footers. So this chapter must at
least contain three pages. To make sure that we get the required amount,
the \textsf{lipsum} package isn't used but the text is put directly in the
text.

\section{Conclusion}
The final section of the chapter gives an overview of the important results
of this chapter. This implies that the introductory chapter and the
concluding chapter don't need a conclusion.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "thesis"
%%% End:

### 2015-02-16 doc/template/chap-2.tex
aanvullingen aan intro flipit

@@
 and Rivest.
-First we explain the framework of FlipIt and after that the formulas and aannames that we will make for the game for during the whole paper.
+First we explain the framework of FlipIt and after that the formulas and assumptions that we will make for the game for during the whole paper.
 
@@
 \flip{FlipIt} is a two-players game with a shared (single) resource that the players want to control as long as possible.
-The shared resource can be a password, a network or a secret key depending on the setting being modeled.
+The shared resource can be a password, a network or a secret key depending on the setting being modelled.
 In the rest of the paper we will call the players the Attacker and the Defender.
@@
 Players won't move to frequently.
-A move can also result in a "wasted move", in this paper called flop.
-It could be that the resource was for example already under control of the defender.
-If the defender moves when he or she has already the control over the resource, he or she would have a wasted move because it would not result in a change of ownership.
+A move can also result in a "wasted move", called a flop.
+It may happen that the resource was already under control by the defender.
+If the defender moves when he or she has already control over the resource, he or she would have wasted move since it does not result in a change of ownership.
 
-Because the players move in a sthealty way, there are different types of feedback that a player can get while moving:
+Because the players move in a stealthy way, there are different types of feedback that a player can get while moving:
 \begin{itemize}
@@
 
-In our assumption will the strategy of both players be non-adaptive. Non of the players has information of the strategy of the opponent.
+In our assumption the strategy of both players will be non-adaptive. None of the players has information of the strategy of the opponent.
 

### 2015-02-18 doc/template/chap-1.tex
aanvullingen aan intro flipit
virussen updata

--- 2015-02-15 doc/template/chap-1.tex
@@
 
-\section{Intro Game Theory}
-
-%begin over dat gametheorie handig is in de economie
-In the following paragraph an introduction to game theory is given based on the work of \todo{werk van Essentials of Game Theory} and \todo{Coursera game theory}.
-Game theory studies the interaction between independent and self-interested agents.
-For this reason it is very important for economics and also for politics, biology, computer science, philosophy and a variety of other disciplines.
-It is a mathematical way of modelling the interactions between two or more agents where the outcome depends on what everybody does and how it should be structured to lead to good outcomes.
-\\
-%Every agent has different levels of happiness for the different outcomes.
-A self-interested independent agent does not necessarily mean that they want to harm other agents or that they only care about themselves.
-Each agent has preferences about the states of the world he likes.
-These preferences are mapped to natural numbers and are called the utility function.
-The numbers are interpreted as a mathematical measure to tell you how much an agent likes or dislikes the states of the world.
-It also explains the impact of uncertainty.
-When an agent is uncertain about a distribution of outcomes, his utility will describe the expected value of the utility function with respect to the probability of the distribution of the outcomes.
-For example: with 0.7 probability it will be 7 degrees outside and 0.3 probability it will be 10 degrees.
-The agent can have a different opinion about that distribution versus another distribution.
-(\todo{uitleggen aan de hand van een voorbeeld}).\\
-
-A game in game theory consists of multiple agents and every agent has a set of actions that he can play. 
-%self interested meaing
-
-%utility function meaning
-
-%Cooperative and non cooperative games
-
-% example of a game
-
-\section{Other}
-In this chapter an introduction to gametheory will be given with the formulas that will be used troughout this paper.
+In this chapter an introduction to gametheory will be given with the formulas that will be used throughout this paper.
 We start with the basics of gametheory.
 People that have a background in gametheory can skip this chapter.
+\section{Intro to Gametheory}
+
+Game Theory is the study of interactions between independent and self-interested agents.
+It is a set of tools to analyze situations in which the optimal action for an agent depends on anticipations that form over another agent decision \todo{citatie vanuit wikipedia}.
+An agent can be either an individual, an animal or an enterprise.
+Game theory can be applied on various domains such as politics, economics, biology, sociology.
+In game theory studies the strategic interaction between players in a game.
+A game is played by a number of players and each of these players has actions.
+These actions are represented by a number indicating their preference.
+
+Ways of defending a network:
+\begin{itemize}
+\item Self-defending networks: The next generation of network security
+\item Honeynet games: a game theoretic approach to defending network monitors
+
+\end{itemize}
 

### 2015-02-18 doc/template/chap-2.tex
virussen updata

--- 2015-02-15 doc/template/chap-2.tex
@@
 This model is useful if somebody wants to know for example if his password has been compromised or wants to assess the periodic security of a system.
+In [\todo{citatie van Mitigatie blabla en nog is mitigatie blabla}] \cite{MitigationNonTargeted} Laszka et al.
+they also consider non targeted attacks by non-strategic players and .
 
@@
 
-\subsection{An item}
-A master thesis is never an isolated work. This means that your text must
-contain references. On-line documents\cite{FlipThem} as well as
-books\cite{DefendingAgainstUnknownEnemy} can be referenced. \cite{Craig2005}.
-
-\section{Figures}
-Figures are used to add illustrations to the text. The \fref{fig:logo} shows
-the KU~Leuven logo as an illustration.
-\begin{figure}
-  \centering
-  \includegraphics{logokul}
-  \caption{The KU~Leuven logo.}
-  \label{fig:logo}
-\end{figure}
-
-\section{Tables}
-Tables are used to present data neatly arranged. A table is normally
-not a spreadsheet! Compare \tref{tab:wrong} en \tref{tab:ok}: which table do
-you prefer?
-
-\begin{table}
-  \centering
-  \begin{tabular}{||l|lr||} \hline
-    gnats     & gram      & \$13.65 \\ \cline{2-3}
-              & each      & .01 \\ \hline
-    gnu       & stuffed   & 92.50 \\ \cline{1-1} \cline{3-3}
-    emu       &           & 33.33 \\ \hline
-    armadillo & frozen    & 8.99 \\ \hline
-  \end{tabular}
-  \caption{A table with the wrong layout.}
-  \label{tab:wrong}
-\end{table}
-
-\begin{table}
-  \centering
-  \begin{tabular}{@{}llr@{}} \toprule
-    \multicolumn{2}{c}{Item} \\ \cmidrule(r){1-2}
-    Animal    & Description & Price (\$)\\ \midrule
-    Gnat      & per gram    & 13.65 \\
-              & each        & 0.01 \\
-    Gnu       & stuffed     & 92.50 \\
-    Emu       & stuffed     & 33.33 \\
-    Armadillo & frozen      & 8.99 \\ \bottomrule
-  \end{tabular}
-  \caption{A table with the correct layout.}
-  \label{tab:ok}
-\end{table}
-
-\section{Lorem Ipsum}
-This section is added to check headers and footers. So this chapter must at
-least contain three pages. To make sure that we get the required amount,
-the \textsf{lipsum} package isn't used but the text is put directly in the
-text.
-
-\section{Conclusion}
-The final section of the chapter gives an overview of the important results
-of this chapter. This implies that the introductory chapter and the
-concluding chapter don't need a conclusion.
-
 %%% Local Variables: 

### 2015-02-18 doc/template/conclusion.tex

\chapter{Conclusion}
\label{cha:conclusion}
The final chapter contains the overall conclusion. It also contains
suggestions for future work and industrial applications.

\section{trala}
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "thesis"
%%% End:

### 2015-02-18 doc/template/intro.tex

\chapter{Introduction}
\label{cha:intro}
The first contains a general introduction to the work. The goals are
defined and the modus operandi is explained.

\section{chap}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "thesis"
%%% End:

### 2015-02-18 doc/template/chap-5.tex

--- 2015-02-15 doc/template/chap-1.tex
@@
 In the following paragraph an introduction to game theory is given based on the work of \todo{werk van Essentials of Game Theory} and \todo{Coursera game theory}.
-Voor een meer gedetailleerdere intro ook dit book.
+Voor een meer gedetailleerdere intro ook dit boek.
 %------------------------------------------------%
@@
 
+%Nash equilibrium
+
 % --------------- example of a game -----------------%
@@
 \section{Virusses}
+Ways of defending a network:
+\begin{itemize}
+\item Self-defending networks: The next generation of network security
+\item Honeynet games: a game theoretic approach to defending network monitors
 
+\end{itemize}
 Many network security threats today are spread over the Internet. The most common include:

### 2015-02-21 doc/template/chap-1.tex
juiste tekst comitten
juiste tekst
tekst
Proberen
gedaan
bib probleem proberen op te lossen

--- 2015-02-15 doc/template/chap-1.tex
@@
 
-In the following paragraph an introduction to game theory is given based on the work of \todo{werk van Essentials of Game Theory} and \todo{Coursera game theory}.
-Voor een meer gedetailleerdere intro ook dit book.
+In the following paragraph an introduction to game theory is given based on the work of \cite{leyton2008essentials} and \cite{Coursera}.
+For a more detailed and full introduction to game theory, the reader is referred to \cite{leyton2008essentials}.
 %------------------------------------------------%
@@
 %begin over dat gametheorie handig is in de economie
-
 Game theory studies the interaction between independent and self-interested agents.
@@
 When an agent is uncertain about a distribution of outcomes, his utility will describe the expected value of the utility function with respect to the probability of the distribution of the outcomes.
-For example: with 0.7 probability it will be 7 degrees outside and 0.3 probability it will be 10 degrees.
+For example: with 0.7 probability it will be 7 degrees outside and with 0.3 probability it will be 10 degrees.
 The agent can have a different opinion about that distribution versus another distribution.
@@
 
-In the following list a couple of terms that will be used throughout the paper.
+In the following lists a couple of terms that will be used throughout the paper.
 \begin{description}
@@
 
+%Nash equilibrium
+
 % --------------- example of a game -----------------%
-
-\section{Other}
-In this chapter an introduction to gametheory will be given with the formulas that will be used troughout this paper.
-We start with the basics of gametheory.
-People that have a background in gametheory can skip this chapter.
 
@@
 
-\subsection{What are my topics}
-\begin{itemize}
-\item Security, Costs, Cybersecurity
-\item Viruses, kinds
-\item Gametheory
-\item Flip-it
-\item Flip-it multiple resources
-\item 
-\end{itemize}
-
 \subsection{Malware}

### 2015-02-21 doc/template/chap-2.tex
juiste tekst comitten
bib probleem proberen op te lossen

--- 2015-02-16 doc/template/chap-2.tex
@@
 The model is called FlipThem \cite{FlipThem}.
-To ways of flipping the resources are used: the AND and the OR control model.
+Two ways of flipping the resources are used: the AND and the OR control model.
 In the AND model the attacker only controls the system if he controls all the resources of the system, whereas in the OR model the attacker only needs to compromise one resource to be in control of the entire system.
-The difference with FlipThem and this paper is that we introduce a Graph Model in the beginning.
-Another extention on FlipIt is done by Pham [\todo{citatie needed voor Are We Compromised?}].
+The difference with FlipThem and this paper is that we introduce a Graph Model in the beginning.\\
+Another extension on FlipIt is done by Pham [\todo{citatie needed voor Are We Compromised?}].
 Beside the action Flip their is another action Test.
@@
 This model is useful if somebody wants to know for example if his password has been compromised or wants to assess the periodic security of a system.
+In \citep{MitigationCovert} \cite{MitigationNonTargeted} Laszka et al.
+they also consider non targeted attacks by non-strategic players and \todo{verder aanvullen}.
 
@@
 
-\subsection{An item}
-A master thesis is never an isolated work. This means that your text must
-contain references. On-line documents\cite{FlipThem} as well as
-books\cite{DefendingAgainstUnknownEnemy} can be referenced. \cite{Craig2005}.
-
 \section{Figures}
-Figures are used to add illustrations to the text. The \fref{fig:logo} shows
-the KU~Leuven logo as an illustration.
-\begin{figure}
+\begin{figure}[hbtp]
 \centering
-  \includegraphics{logokul}
-  \caption{The KU~Leuven logo.}
-  \label{fig:logo}
+\includegraphics[scale=1]{Images/FlipItDefault.jpg}
+\caption{The FlipIt game where both players are playing periodically}
+\label{fig:FLipItDefault}
 \end{figure}
-
-\section{Tables}
-Tables are used to present data neatly arranged. A table is normally
-not a spreadsheet! Compare \tref{tab:wrong} en \tref{tab:ok}: which table do
-you prefer?
-
-\begin{table}
-  \centering
-  \begin{tabular}{||l|lr||} \hline
-    gnats     & gram      & \$13.65 \\ \cline{2-3}
-              & each      & .01 \\ \hline
-    gnu       & stuffed   & 92.50 \\ \cline{1-1} \cline{3-3}
-    emu       &           & 33.33 \\ \hline
-    armadillo & frozen    & 8.99 \\ \hline
-  \end{tabular}
-  \caption{A table with the wrong layout.}
-  \label{tab:wrong}
-\end{table}
-
-\begin{table}
-  \centering
-  \begin{tabular}{@{}llr@{}} \toprule
-    \multicolumn{2}{c}{Item} \\ \cmidrule(r){1-2}
-    Animal    & Description & Price (\$)\\ \midrule
-    Gnat      & per gram    & 13.65 \\
-              & each        & 0.01 \\
-    Gnu       & stuffed     & 92.50 \\
-    Emu       & stuffed     & 33.33 \\
-    Armadillo & frozen      & 8.99 \\ \bottomrule
-  \end{tabular}
-  \caption{A table with the correct layout.}
-  \label{tab:ok}
-\end{table}
-
-\section{Lorem Ipsum}
-This section is added to check headers and footers. So this chapter must at
-least contain three pages. To make sure that we get the required amount,
-the \textsf{lipsum} package isn't used but the text is put directly in the
-text.
+\todo{verwijzen naar de figuur \ref{fig:FLipItDefault}}
 

### 2015-02-21 doc/template/chap-5.tex
bib probleem proberen op te lossen

@@
 
-In the following paragraph an introduction to game theory is given based on the work of \todo{werk van Essentials of Game Theory} and \todo{Coursera game theory}.
-Voor een meer gedetailleerdere intro ook dit boek.
+In the following paragraph an introduction to game theory is given based on the work of \citep{leyton2008essentials} and \citep{Coursera}.
+For a more detailed and full introduction to game theory, the reader is referred to \citep{leyton2008essentials}.
 %------------------------------------------------%
@@
 
-\section{Other}
-In this chapter an introduction to gametheory will be given with the formulas that will be used troughout this paper.
-We start with the basics of gametheory.
-People that have a background in gametheory can skip this chapter.
-
 %------------------------------------------------%

### 2015-02-22 doc/template/chap-4.tex
Extra shizzle

@@
 
-What is different from Flip-It.
+We model the network through an undirected Graph $G = < V, E> $ where V denotes the number of resources in the network and E the connection between two resources. 
+$N_{0}$ denotes the initially infected resources at the beginning of the virus propagation. 
+$P_{x}(R_{n},t|R_{0},r,t_{0})$ denotes the chance that resource $R_{n}$ is infected at time $t$ after dropping virus number x on to resource $R_{0}$ at $t_{0}$ with rate $r$. \\
 
+$S(R_{n},R_{0})$ denotes the shortest path from the infected resource $R_{0}$ to resource $R_{n}$.
+It gives back a value with the distance measured with how many resources are in between including the end resource.
+\\
+
+So the chance that a resource is infected after time t is the chance that the resource is infected by all the previous infections and that the defender has not flipped the resource.
 %%% Local Variables: 

### 2015-02-22 doc/template/chap-4.tex
matrixen committen

--- 2015-01-25 doc/template/chap-4.tex
@@
 
-What is different from Flip-It.
+%$N_{0}$ denotes the initially infected resources at the beginning of the virus propagation. 
+%$P_{x}(R_{n},t|R_{0},r,t_{0})$ denotes the chance that resource $R_{n}$ is infected at time $t$ after dropping virus number x on to resource $R_{0}$ at $t_{0}$ with rate $r$. \\
+
+%$S(R_{n},R_{0})$ denotes the shortest path from the infected resource $R_{0}$ to resource $R_{n}$.
+It gives back a value with the distance measured with how many resources are in between including the end resource.
+\\
+
+%So the chance that a resource is infected after time t is the chance that the resource is infected by all the previous infections and that the defender has not flipped the resource.
+source: \url{http://en.wikipedia.org/wiki/Adjacency_matrix} \\
+We model the network through an undirected Graph $G = < V, E> $ where $|V|$ denotes the number of resources in the network and $|E|$ the number of connections.
+We can convert this to a adjacent matrix where we can represent which vertices of the graph are neighbours of other vertices.
+\\
+For our graph we have an $|V| \times |V|$ matrix with on every entry $a_{ij}$ a 1 as value if there is a connection between node $V_{i}$ and $V_{j}$ and with zeros its diagonal.
+Because our graph is undirected we have a symmetric matrix.
+
+\textit{"If \textit{A} is the adjacency matrix of the directed or undirected graph \textit{G}, then the matrix $A^{n}$ (i.e., the matrix product of n copies of \textit{A}) has an interesting interpretation: the entry in row i and column j gives the number of (directed or undirected) walks of length n from vertex i to vertex j.
+If n is the smallest nonnegative integer, such that for all i ,j , the (i,j)-entry of $A^{n} > 0$, then n is the distance between vertex i and vertex j."} [Wikipedia]
 
@@
 %%% End: 
+\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2.8cm,
+                    semithick]
+  \tikzstyle{every state}=[fill=blue,draw=none,text=white]
 
+  \node[initial,state] (A)                    {$N_1$};
+  \node[state]         (B) [above right of=A] {$N_2$};
+  \node[state]         (D) [below right of=A] {$N_3$};
+  \node[state]         (C) [below right of=B] {$N_4$};
+  \node[state]         (E) [below right of=C] {$N_5$};
+  \node[state]		   (F) [above right of=C] {$N_6$};
+
+  \path (A) edge              node {} (B)
+            edge              node {} (D)
+        (B) edge              node {} (A)
+        	edge			  node {} (C)
+        (C) edge              node {} (B)
+            edge 			  node {} (D)
+            edge			  node {} (E)
+            edge			  node {} (F)
+        (D) edge 			  node {} (C)
+            edge              node {} (A)
+        (E) edge 			  node {} (C)
+    	(F)	edge			  node {} (C);
+\end{tikzpicture}
+\\
+%\[
+%\begin{bmatrix}
+%    x_{11}       & x_{12} & x_{13} & \dots & x_{1n} \\
+%    x_{21}       & x_{22} & x_{23} & \dots & x_{2n} \\
+%    \hdotsfor{5} \\
+%    x_{d1}       & x_{d2} & x_{d3} & \dots & x_{dn}
+%\end{bmatrix}
+%=
+%\begin{bmatrix}
+%    x_{11} & x_{12} & x_{13} & \dots  & x_{1n} \\
+%    x_{21} & x_{22} & x_{23} & \dots  & x_{2n} \\
+%    \vdots & \vdots & \vdots & \ddots & \vdots \\
+%    x_{d1} & x_{d2} & x_{d3} & \dots  & x_{dn}
+%\end{bmatrix}
+%\] 
+The adjacent matrix becomes this matrix $A$: \\
+
+$
+\bordermatrix{
+         & N_1		& N_2	& N_3	& N_4 	& N_5	&N_6     \cr
+    N_1   & 0		& 1		& 1		& 0		& 0		& 0	     \cr
+    N_2   & 1		& 0		& 0		& 1		& 0		& 0	     \cr
+    N_3   & 1		& 0		& 0		& 1		& 0		& 0	     \cr
+    N_4   & 0		& 1		& 1		& 0		& 1		& 1	     \cr
+	N_5   & 0		& 0		& 0		& 1		& 0		& 0	     \cr
+	N_6   & 0		& 0		& 0		& 1		& 0		& 0	     \cr
+}$
+\\
+
+Matrix $A \times A = A^{2}$ becomes the matrix with the number of paths with 2 steps from $N_{i}$ to $N_{j}$:\\
+
+$
+\bordermatrix{
+         & N_1		& N_2	& N_3	& N_4 	& N_5	&N_6     \cr
+    N_1   & 2		& 0		& 0		& 2		& 0		& 0	     \cr
+    N_2   & 0		& 2		& 2		& 0		& 1		& 1	     \cr
+    N_3   & 0		& 2		& 2		& 0		& 1		& 1	     \cr
+    N_4   & 2		& 0		& 0		& 4		& 0		& 0	     \cr
+	N_5   & 0		& 1		& 1		& 0		& 1		& 1	     \cr
+	N_6   & 0		& 1		& 1		& 0		& 1		& 1	     \cr
+}$
+\\
+
+Matrix $A^{2} \times A = A^{3}$ becomes the matrix with the number of paths with 3 steps from $N_{i}$ to $N_{j}$:\\
+
+$
+\bordermatrix{
+         & N_1		& N_2	& N_3	& N_4 	& N_5	&N_6     \cr
+    N_1   & 0		& 4		& 4		& 0		& 2		& 2	     \cr
+    N_2   & 4		& 0		& 0		& 6		& 0		& 0	     \cr
+    N_3   & 4		& 0		& 0		& 6		& 0		& 0	     \cr
+    N_4   & 0		& 6		& 6		& 0		& 4		& 4	     \cr
+	N_5   & 2		& 0		& 0		& 4		& 0		& 0	     \cr
+	N_6   & 2		& 0		& 0		& 4		& 0		& 0	     \cr
+}$ 
+\\
+
+So for $A^{N}$ every $a_{ij}$ entry gives the number of paths with N steps from $N_{i}$ to $N_{j}$.\\
+
+With this knowledge we can calculate in how many steps a node is infected.
+$A$ calculates which nodes are infected after 1 step, $A^{N}$ calculates which nodes are infected in N steps..
+So if we want to know how many nodes are infected after 3 steps we have to add every matrix $(A + A^{2} + A^{3}) $ and see which entry is a non zero entry.
 \end{document}

### 2015-02-23 doc/template/chap-4.tex
Matrixen vervolg commit

@@
 %\] 
-The adjacent matrix becomes this matrix $A$: \\
+The adjacent matrix becomes this matrix $[A]$: \\
 
@@
 
-Matrix $A \times A = A^{2}$ becomes the matrix with the number of paths with 2 steps from $N_{i}$ to $N_{j}$:\\
+Matrix $A \times A = A^{2}$ becomes the matrix with the number of paths with 2 steps from $N_{i}$ to $N_{j}$: We denote this matrix as matrix \textit{[B]}\\
 
@@
 
-Matrix $A^{2} \times A = A^{3}$ becomes the matrix with the number of paths with 3 steps from $N_{i}$ to $N_{j}$:\\
+Matrix $A^{2} \times A = A^{3}$ becomes the matrix with the number of paths with 3 steps from $N_{i}$ to $N_{j}$: We denote this matrix as matrix \textit{[C]}\\
 
@@
 So if we want to know how many nodes are infected after 3 steps we have to add every matrix $(A + A^{2} + A^{3}) $ and see which entry is a non zero entry.
+
 \end{document}

### 2015-02-27 Thesisfollowup/ThesisFollowUpSophieMarien.tex
Game formal definition

@@
 
-\section{Geef een uitgewerkt werkplan voor je thesis-activiteiten tussen nu en half juni (dan moet je je masterproef indienen).
-Dit werkplan moet PER TWEE WEKEN aangeven wat je precies gaat doen voor je masterproef.
-Maak het heel CONCREET: in het geval van implementatie, implementatie van welk deel precies, in het geval van het schrijven van tekst, het schrijven van welk hoofdstuk precies, e.d.
-Stel dit werkplan zelfstandig op, en bespreek het daarna met je begeleider, voordat je op deze mail antwoordt.}
-
-Uitgewerkt plan steekt bij in de bijlage. In de timesheet file staat naast de werkuren aan wat ik ga werken.
-
-\section{Schrijf een stuk tekst van minstens 5 pagina's waarin je de probleemstelling van je masterproef formuleert, gevolgd door de doelstellingen.
-Je verwerkt hier onder meer elementen van je literatuurstudie in.}
-
-Ik heb afgesproken dat ik het stuk tekst tegen dit weekend uitschrijf. Hier is al een kleine inleiding.
-
-Virussen zijn een groot probleem voor bedrijfsnetwerken.
-Ze kunnen gevoelige informatie verzamelen of een bedrijfsnetwerk platleggen.
-Gegeven de grote kost gebonden aan schade door malware, is het vinden van de juiste verdegingingsstrategie belangrijk.
-Het aanvallen en verdedigen van een bedrijfsnetwerk kan gezien worden als een spel, waarbij de verdediger en de aanvaller elk proberen de beste strategie te vinden.
-Ik ga dus toelichten hoe het spel van aanvallen en verdedigen kan gedefinieerd worden als een variatie op het spel FlipIt.
-Dit laat toe om te onderzoeken wat de verschillende strategieën zijn van de netwerkbeheerder enerzijds en van de aanvaller die virussen zendt, anderzijds.
-Bedoeling is om in een volgende stap het spel verder te analyseren met behulp van Game Theory om te bepalen welke de dominerende strategieën zijn en of er zich Nash equilibria voordoen.
-
+\section{Vergelijk wat je gedaan hebt de vorige weken met het opgestelde werkplan. Hoever sta je?}
+Het waren examens vorige weken en er was een lesvrije week, dus veel vooruitgang heb ik niet geboekt in vergelijking met andere weken.
+Ik zat ook een beetje vast op een gegeven moment maar dan heb ik verder gewerkt aan de literatuurstudie en al beginnen te schrijven aan men thesis.
+\section{Verfijn je werkplan (pas het eventueel aan) zodat dit werkplan PER WEEK, heel CONCREET opgeeft wat je van plan bent te doen.}
+Dit werkplan heb ik concreter gemaakt in mijn timesheet. Dus daar kan het concretere werkplan teruggevonden worden
+\section{Geef een outline van je thesistekst (Titels van hoofdstukken met per hoofdstuk 2 tot 5 regels over de inhoud). Dit mag je natuurlijk opstellen in overleg met je thesisbegeleider.}
+\begin{description}
+\item ABSTRACT
+\item INTRODUCTION\\
+\begin{itemize}
+\item motivation game
+\item contributions and results
+\item lessons derived
+\end{itemize}
+\item CYBER SECURITY AND GAME THEORY
+\item PROBLEM DEFINITION
+\item MY GAME
+\begin{itemize}
+\item flipIt 
+\item game motivation
+\item examples
+\item formal definitions
+\item Default game is FlipIt
+\item Multiple resources nash?
+\item write out gain for different graphs and parameters
+\end{itemize}
+\item CONCLUDING REMARKS
+\item GENERAL VARIATIONS AND EXTENSIONS
+\item APPENDIX
+\item REFERENCES
+\end{description}
 \section{Evalueer de tijd die je in je thesis gestoken hebt in de voorbije 4 weken: veel te weinig / te weinig / voldoende / ruim voldoende / meer dan verwacht.}
-
-Ik heb voldoende tijd in mijn thesis gestoken.
-Ik heb de timesheet aangepast naar mijn planning.
-Ik heb weinig examens en ben een week eerder klaar dan de lesvrije week dus dan kan ik nog verder werken aan mijn thesis.
-Voor deze komende weken heb ik minder uren ingepland omdat ik van andere vakken projecten moet afwerken en presentaties moet geven.
-De laatste week heb ik minder kunnen werken dan dat ik graag had willen doen omwille van een kaakgewrichtsontsteking die momenteel nog steeds niet over is.
-De volgende weken ga ik proberen meer te werken aan mijn thesis dan dat ik gepland heb, om mijn uren in te halen.
+Voldoende. Ik ben maar 5 uur achtergeraakt op het schema in de week voor de lesvrije week maar toen ben ik ook ziek geweest.
 
@@
 het bereiken van de vooropgestelde doelstelling: ver achter op het schema / achter op het schema / goed op schema / vooruit op het schema / ver vooruit op het schema.}
+Beetje achter op het schema, maar die achterstand kan snel ingehaald worden.
+\section{Stuur de ingevulde timesheet mee als bijlage}
 
-Ik sta op schema volgens het schema dat ik zelf gemaakt heb. Ik heb mijn schema voor de komende weken aangepast.
-
-\section{Stuur de ingevulde timesheet mee als bijlage}
 \end{document}

### 2015-02-27 doc/template/chap-2.tex
Game formal definition

@@
 \centering
-\includegraphics[scale=1]{Images/FlipItDefault.jpg}
+\includegraphics[scale=0.25]{Images/FlipItDefault.jpg}
 \caption{The FlipIt game where both players are playing periodically}
@@
 \todo{verwijzen naar de figuur \ref{fig:FLipItDefault}}
+
+\subsection{Strategies}
+
+ In this subsection we go through the strategies used in FlipIt and the most important results. 
+ \begin{table}
+ \centering
+ \begin{tabular}{ l | c  }
+  \textbf{Categories} & \textbf{Classes of Strategies} \\
+  \hline Non-adaptive (NA) & Exponential \\
+  & Periodic \\
+  & Renewal \\
+  & General non-adaptive \\
+  \hline Adaptive (AD) & Last move (LM) \\
+  & Full History (FH) \\  
+\end{tabular}
+ \caption{Classes of strategies in FlipIt}
+ \label{table:Strategies}
+ \end{table}
+
+There are two different kinds of strategies, the \textit{non-adaptive strategies} and the \textit{renewal strategies}.
+If there is no need for feedback for both of the players, we say that we have a non-adaptive strategy.
+Because the player does not receive any feedback during the game it will play in the same manner against every opponent.
+They are not dependent on the opponents movements.
+This means that they can already generate the time sequence for all the moves in advance.
+But they can depend on some randomness because the non-adaptive strategies can be randomised.
+In this paper we will focus in the beginning on the non-adaptive strategies.
+Reasons behind this that a player (defender or attacker) rarely knows what the strategies are of his opponent.
+[If the attacker wants to move stealthily, it might have limited attack options FLIPTHEM].
+\todo{nog redenen zoeken}\\
+A renewal strategy is a non-adaptive strategy where the time intervals between two consecutive moves are generated by a renewal process. \\
+
+ \begin{description}
+ \item Periodic
+ \item Non-Arithmetic Renewal
+ \item Exponential
+ \end{description}
+
+\section{Formal definition Game}
+
+In this section we provide the formal definition of the game and the notation that we will use throughout the paper.
+
+\begin{description}
+\item \textit{Players}  There are two players in the game, one is the defender and the other one is the attacker. They are respectively identified by 0 and 1.
+\item \textit{Time}  The game starts at $t=0$ and continuous indefinitely as $t \rightarrow \infty$. The game is a continuous game.
+\item \textit{Game State} There is also a time-dependent variable that represents the state of the game.
+$C=C(t)$ is either 0 if the Game is under control by the defender and 1 if the Game is under control by the attacker.
+We can also define the state of each resource by $C^{A}$ and $C^{D}$.
+If $C^{A} = 1$ then this means that the attacker has control over the resource, and 0 otherwise.
+For $C^{D}$ it is visa versa, $C^{D} = 1 - C^{A}$.
+\item \textit{Graph} We represent the company network as a Graph $G = < V,E>$.
+G is an ordered pair where V denotes the set of resources or nodes in the network and E denotes the set of connections or links, which are a two-element subset of V.
+We use the notations resources and nodes interleaving in this paper.\\
+We have N resources in the network. $N \in $  \todo{aanvullen}. This means we can denote the resources by:
+\begin{center}
+$V \in {V_{0}, V_{1}, V_{2}, ... , V_{N} }$
+\end{center}
+The set E of connections indicates if there is a link between two resources.
+We see the links as bidirectional so the total graph is undirected.
+If there is a link between resource $V_{n}$ and $V_{n+1}$ then there is also a link between $V_{n+1}$ and $V_{n}$.
+
+\item \textit{Moves} Both players can make a move in the game.
+Moves done in a finite numbers of time in any finite time interval.
+Both players can play at any time they want, they can also play at the same time.
+If this happens the one that has control over the resource will keep having control over the resource.
+This makes the game fully symmetric \todo{beter uitleggen}. The sequence of move times are denoted by the following infinite sequence:
+\begin{center}
+$t=t_{1},t_{2},t_{3},..$
+\end{center}
+Two move times can be the same because we allow players to move at the same time.
+We can also denote the infinite sequence of times when player \textit{i}  moves. We write this as :
+\begin{center}
+$t=t_{i,1},t_{i,2},t_{i,3},..$ with \textit{i} $ \in $ $\lbrace 0,1 \rbrace$
+\end{center}
+The sequences $t_{1}$ and $t_{0}$ are disjoint subsets of the sequent t. 
+We can also denote who made the \textit{k}th move by defining a sequence \textit{p} that denotes the sequence of who played:
+\begin{center}
+$p=p_{1},p_{2},p_{3}, .. $ with $p_{k}$ $\in$ $\lbrace 0,1 \rbrace$
+\end{center}
+
+\item \textit{Number of moves}  $n_{i}(t$ denotes the number of moves made by player \textit{i} up to and including time t. This means that 
+\begin{center}
+$n(t)=n_{1}(t) + n_{0}(t)$
+\end{center}
+is the sum of the number of moves made by the defender and the attacker up to and including time t.
+
+\item \textit{Average move rate} We denote $\alpha_{i}(t)$ as the average move rate by player i:
+\begin{center}
+$\alpha_{i}(t) = n_{i}(t)/t$ with $t > 0$ and \textit{i} $ \in $ $\lbrace 0,1 \rbrace$
+\end{center}
+
+\item \textit{Period} We can also define the period in terms of the average move rate:
+\begin{center}
+$\delta_{i}=1/\alpha_{i}$
+\end{center}
+
+\item \textit{Cost} The cost is an important property of the game.
+In FlipIt for every player the cost of a move is denoted by $k_{i}$.
+These costs can be very different for every player.
+In this game we denote the players flipping cost for resource $V_{N}$ by $c_{i}^{V_{N}}$.
+\\
+For the defender the cost will be either the cost of flipping every resource or the cost of flipping a subgroup of the resources.\\
+For the attacker the cost will be the cost of dropping a virus on a node. The spreading of the virus will not imply an extra cost.
+
+\item \textit{Utility} In FlipIt the Gain definition is the utility function.
+The Gain denotes the total time a player i has gained control over a resource.
+\todo{nu gain van een resource, moet voor verschillende resources zijn}
+The Gain $G_{i}$ denotes players i total gain of a game, wich is the total time the player has gained control over a subset of resources thus controlling the game.
+If we sum up the total Gain of the attacker and the defender we end up with the time:
+\begin{center}
+$G_{1}(t) + G_{0}(t) = t$
+\end{center}
+\item \textit{Average gain rate} The average gain rate for player i is defined as
+\begin{center}
+$\gamma_{i}(t)= G_{i}(t)/t$
+\end{center}
+
+\subsection{Our Game parameters}
+\begin{description}
+\item \textit{Graph Matrix} We represent the graph of the network through a matrix $ A = |V| \times |V|$.
+The (i,j)-entry of the matrix A will have a 1 if there is a connection between node $V_{i}$ and node $V_{j}$.
+If we are working with an undirected graph the matrix will be symmetric.
+\item \textit{Attack Vector} We denote $X = 1 \times |V|$ as the attack vector.  
+\item \textit{Reset vector} The reset vector will make sure that the right entries in the matrix become zero.
+\item \textit{Cummulative Matrix} 
+\item \textit{State Matrix} The State matrix $T(t) = |V| \times |V| $ will keep at every time t the state of the game.
+\end{description}
+
+\end{description}
 

### 2015-02-27 doc/template/chap-2.tex
Formal definition extra

@@
 \begin{description}
+% ---- PLAYERS ---- %
 \item \textit{Players}  There are two players in the game, one is the defender and the other one is the attacker. They are respectively identified by 0 and 1.
+
+% ---- TIME ---- %
 \item \textit{Time}  The game starts at $t=0$ and continuous indefinitely as $t \rightarrow \infty$. The game is a continuous game.
-\item \textit{Game State} There is also a time-dependent variable that represents the state of the game.
-$C=C(t)$ is either 0 if the Game is under control by the defender and 1 if the Game is under control by the attacker.
-We can also define the state of each resource by $C^{A}$ and $C^{D}$.
-If $C^{A} = 1$ then this means that the attacker has control over the resource, and 0 otherwise.
-For $C^{D}$ it is visa versa, $C^{D} = 1 - C^{A}$.
+
+% ---- GRAPH ---- %
 \item \textit{Graph} We represent the company network as a Graph $G = < V,E>$.
@@
 
+% ---- GAME STATE ---- %
+\item \textit{Game State} There is also a time-dependent variable that represents the state of the game.
+$C=C(t)$ is either 0 if the game is under control by the defender and 1 if the Game is under control by the attacker.
+\\
+We start at $t=0$ with the defender who has control over the game.
+We do this because we assume that the defender will only put the network online without having a virus or worm in it.
+The Attacker can gain control over the game when it compromises a subset \textit{s} of the resources.
+The subset \textit{s} is a minimum of 1 resource and a maximum of all the resources N.
+\\
+\todo{deze variabele nodig ja of nee ?
+JA} We can also define the state of each resource by $C^{A}_{N}$ and $C^{D}_{N}$.
+If $C^{A}_{N} = 1$ then this means that the attacker has control over the resource, and 0 otherwise.
+For $C^{D}_{N}$ it is visa versa, $C^{D}_{N} = 1 - C^{A}_{N}$.\\
+
+% ---- MOVES ---- %
 \item \textit{Moves} Both players can make a move in the game.
@@
 
+% ---- NUMVER OF MOVES ---- %
 \item \textit{Number of moves}  $n_{i}(t$ denotes the number of moves made by player \textit{i} up to and including time t. This means that 
@@
 
+% ---- AVERAGE MOVE RATE ---- %
 \item \textit{Average move rate} We denote $\alpha_{i}(t)$ as the average move rate by player i:
@@
 
+% ---- PERIOD ---- %
 \item \textit{Period} We can also define the period in terms of the average move rate:
@@
 
+% ---- WHO PLAYED LAST ---- %
+\item \textit{Who played last} We know who played last by taking the modulo with the period.
+$Z_{i}$ represents the time since the last flip of player i.
+We can also denote the time since the last flip of player i on resource r by $Z_{i}^{N}$.
+
+% ---- COST ---- %
 \item \textit{Cost} The cost is an important property of the game.
@@
 
+% ---- UTILITY/ GAIN ---- %
 \item \textit{Utility} In FlipIt the Gain definition is the utility function.
@@
 \end{center}
+
+% ---- AVERAGE GAIM RATE ---- %
 \item \textit{Average gain rate} The average gain rate for player i is defined as

### 2015-02-27 doc/template/chap-2.tex
map weggedaan

@@
 The difference with FlipThem and this paper is that we introduce a Graph Model in the beginning.\\
-Another extension on FlipIt is done by Pham [\todo{citatie needed voor Are We Compromised?}].
+Another extension on FlipIt is done by Pham\cite{GameTheorApprCostBenefitAnalyses} [\todo{citatie needed voor Are We Compromised?}].
 Beside the action Flip their is another action Test.
@@
 This model is useful if somebody wants to know for example if his password has been compromised or wants to assess the periodic security of a system.
-In \citep{MitigationCovert} \cite{MitigationNonTargeted} Laszka et al.
+In \cite{MitigationCovert} \cite{MitigationNonTargeted} Laszka et al.
 they also consider non targeted attacks by non-strategic players and \todo{verder aanvullen}.
@@
 We can also denote the time since the last flip of player i on resource r by $Z_{i}^{N}$.
+For a non adaptive game, period deterministic: At time $t=n$ is $Z_{i} = n mod i$.  \todo{er kan nog steeds tegelijk geflipt zijn maar dan hebben ze wel geflipt}.
 

### 2015-02-27 Thesisfollowup/ThesisFollowUpSophieMarien.tex
Follow up

@@
 
-\section{Vergelijk wat je gedaan hebt de vorige weken met het opgestelde werkplan. Hoever sta je?}
-Het waren examens vorige weken en er was een lesvrije week, dus veel vooruitgang heb ik niet geboekt in vergelijking met andere weken.
-Ik zat ook een beetje vast op een gegeven moment maar dan heb ik verder gewerkt aan de literatuurstudie en al beginnen te schrijven aan men thesis.
-\section{Verfijn je werkplan (pas het eventueel aan) zodat dit werkplan PER WEEK, heel CONCREET opgeeft wat je van plan bent te doen.}
-Dit werkplan heb ik concreter gemaakt in mijn timesheet. Dus daar kan het concretere werkplan teruggevonden worden
-\section{Geef een outline van je thesistekst (Titels van hoofdstukken met per hoofdstuk 2 tot 5 regels over de inhoud). Dit mag je natuurlijk opstellen in overleg met je thesisbegeleider.}
-\begin{description}
-\item ABSTRACT
-\item INTRODUCTION\\
-\begin{itemize}
-\item motivation game
-\item contributions and results
-\item lessons derived
-\end{itemize}
-\item CYBER SECURITY AND GAME THEORY
-\item PROBLEM DEFINITION
-\item MY GAME
-\begin{itemize}
-\item flipIt 
-\item game motivation
-\item examples
-\item formal definitions
-\item Default game is FlipIt
-\item Multiple resources nash?
-\item write out gain for different graphs and parameters
-\end{itemize}
-\item CONCLUDING REMARKS
-\item GENERAL VARIATIONS AND EXTENSIONS
-\item APPENDIX
-\item REFERENCES
-\end{description}
+\section{Vergelijk wat je gedaan hebt met het opgestelde werkplan. Hoever sta je? Moet het werkplan aangepast worden?}
+\section{Schrijf een eerste versie van een hoofstuk over een afgewerkte bijdrage van je masterproef.}
+\section{Geef een korte inhoud van het tweede artikel dat je gaat schrijven. (Dit is niet van toepassing voor de studenten van de Master Toegepaste Informatica)}
+\section{Beschrijf in 10 lijnen de demo die je bij je verdediging zult geven.}
+\section{Wat is er afgesproken met je begeleider en/of promotor om op te leveren als major milestone die het praktische werk van je thesis afsluit?}
 \section{Evalueer de tijd die je in je thesis gestoken hebt in de voorbije 4 weken: veel te weinig / te weinig / voldoende / ruim voldoende / meer dan verwacht.}
-Voldoende. Ik ben maar 5 uur achtergeraakt op het schema in de week voor de lesvrije week maar toen ben ik ook ziek geweest.
-
 \section{Evalueer de vorderingen van je thesis t.o.v.
 het bereiken van de vooropgestelde doelstelling: ver achter op het schema / achter op het schema / goed op schema / vooruit op het schema / ver vooruit op het schema.}
-Beetje achter op het schema, maar die achterstand kan snel ingehaald worden.
-\section{Stuur de ingevulde timesheet mee als bijlage}
+\section{Stuur de ingevulde timesheet mee als bijlage.}
 

### 2015-02-27 doc/template/chap-2.tex
Follow up

@@
 they also consider non targeted attacks by non-strategic players and \todo{verder aanvullen}.
+\\
 
@@
 \begin{itemize}
-\item Non-adaptive (NA): The player does not receive any feedback while flipping.
-\item Last move (LM): The player finds out the exact time the opponent played the last time.
-\item Full History (FH): The player finds out the complete history of the opponents move.
+\item Non-adaptive (NA): The player does not receive any feedback during the game while flipping.
+\item Last move (LM): When a player flips it will find out the exact time that the opponent played the last time.
+\item Full History (FH): When a player flips it will find out the whole history of the opponents move.
 \end{itemize}
@@
 In this paper we will focus in the beginning on the non-adaptive strategies.
-Reasons behind this that a player (defender or attacker) rarely knows what the strategies are of his opponent.
+Reasons behind this is that a player (defender or attacker) rarely knows what the strategies are of his opponent.
 [If the attacker wants to move stealthily, it might have limited attack options FLIPTHEM].
@@
 \todo{nu gain van een resource, moet voor verschillende resources zijn}
-The Gain $G_{i}$ denotes players i total gain of a game, wich is the total time the player has gained control over a subset of resources thus controlling the game.
+The Gain $G_{i}$ denotes players \textit{i} total gain of a game, which is the total time the player has gained control over a subset of resources thus controlling the game.
+This is denoted by the following:\\
+\begin{center}
+$G_{i}(t) = \int^{t}_{0} C_{i}(x) dx$ 
+\end{center}
  If we sum up the total Gain of the attacker and the defender we end up with the time:
@@
 \item \textit{Attack Vector} We denote $X = 1 \times |V|$ as the attack vector.
-\item \textit{Reset vector} The reset vector will make sure that the right entries in the matrix become zero.
-\item \textit{Cummulative Matrix} 
-\item \textit{State Matrix} The State matrix $T(t) = |V| \times |V| $ will keep at every time t the state of the game.
+It will be a vector with only zeros.
+The attacker will place a virus on a node V.
+This will be denoted by the Vth entry in the vector that is changed by a 1.
+\item \textit{Reset vector} The reset vector will make sure that the right entries in the matrix become zero. If the defender flips every node every time it flips then the attack vector will be 0.
+\item \textit{Cummulative Matrix} This matrix will keep record of the propagation of the virus through the network.
+\item \textit{State Matrix} The State matrix $T(t) = 1 \times |V| $ will keep at every time t the state of the game and denote which node at time t is infected with the virus.
+At time $t=0$ the State Matrix will be the null matrix.
 \end{description}
-
+De eerste infectie is de attack vector * Graph matrix . 
 \end{description}

### 2015-03-01 doc/template/chap-2.tex
sort of games

@@
 
+\subsection{Different kind of games}
+De virus has different kind of ways of making his way through the network.
+The virus can be dropped on a node and than spread itself out to all the neighbours.
+When a neighbour node is infected it will infect all its neighbours too, so the initial node is back infected.
+Then their are two options, either the initial node stays infected and does not infect all it neighbours again, or it does.
+\\
+Another propagation method is that the virus works as a token. It will propagate to only one neighbour and continue to spread. 
+%waarom kiezen voor een methode
+\begin{itemize}
+\item A virus can use the mail system to spread itself. This will use then every contact in the contact list so it will spread itself to every neighbour
+\item If the virus does not want to get caught it can use the method of a path, infecting only one neighbour.
+\item 
+\end{itemize}
 \section{Conclusion}

### 2015-03-03 doc/template/chap-2.tex
Spel uitschrijven, verschillende propagaties

--- 2015-02-27 doc/template/chap-2.tex
@@
 
+\subsection{Different kind of games}
+A virus has different kind of ways of making his way through a company network.
+We will describe the different ways of how the virus can propagate.
+For start we will say that the virus or worm will be dropped on Node i and that it has k numbers of neighbours.
+\begin{enumerate}
+\item Node i is infected and will spread the virus or worm to every k neighbours and will stop infecting the neighbours in the next step
+\item Node i is infected and will spread the virus or worm to every k neighbours and will keep on spreading the virus in every next step
+\item Node i is infected and will spread the virus to only one of the k neighbours and will stop infecting another one in the next step
+\item Node i is infected and will spread the virus to only one of the k neighbours and in the next step it will infect another one of the k neighbours 
+\end{enumerate}
+
+In the game that will be modelled in the paper we will use the settings of the 1 spreading method.
+We will not use method 2 because this kind of propagation will float the network.
+Because we use the settings of a mail system and contact in a mailling list the method of 3 and 4 are not used.
+\\
+In the first method the node that has been infected can be again infected.
+If one of the neighbours infect the node again the node will infect his neighbours again.
+By using this spreading method we have three states of a node.
+An infected state, a clean state and a spreading state.
+And infected state means that the node is infected and will not spread itself again to its neighbours, a clean state means that there is no virus or worm on the node and a spreading state means that the node is infected and that it will spread itself in the next step to its neighbours.
+We can argument this kind of propagation through a mail worm. \todo{voorbeeld geven van zo een worm}
+%Another propagation method is that the virus works as a token. It will propagate to only one neighbour and continue to spread.
+
 \section{Conclusion}

### 2015-03-03 doc/template/chap-2.tex
update tekst spel

@@
  \end{description}
+
+ \subsection{Actions of the attacker}
+A virus has different kind of ways of making his way through a company network.
+We will describe the different ways of how the virus can propagate.
+For start we will say that the virus or worm will be dropped on Node i and that it has k numbers of neighbours.
+\begin{enumerate}
+\item Node i is infected and will spread the virus or worm to every k neighbours and will stop infecting the neighbours in the next step
+\item Node i is infected and will spread the virus or worm to every k neighbours and will keep on spreading the virus in every next step
+\item Node i is infected and will spread the virus to only one of the k neighbours and will stop infecting another one in the next step
+\item Node i is infected and will spread the virus to only one of the k neighbours and in the next step it will infect another one of the k neighbours 
+\end{enumerate}
+
+In the game that will be modelled in the paper we will use the settings of the 1 spreading method.
+We will not use method 2 because this kind of propagation will float the network.
+Because we use the settings of a mail system and contact in a mailing list the method of 3 and 4 are not used.
+\\
+In the first method the node that has been infected can be again infected.
+If one of the neighbours infect the node again the node will infect his neighbours again.
+By using this spreading method we have three states of a node.
+An infected state, a clean state and a spreading state.
+And infected state means that the node is infected and will not spread itself again to its neighbours, a clean state means that there is no virus or worm on the node and a spreading state means that the node is infected and that it will spread itself in the next step to its neighbours.
+We can argument this kind of propagation through a mail worm. \todo{voorbeeld geven van zo een worm}
+%Another propagation method is that the virus works as a token. It will propagate to only one neighbour and continue to spread.
+
+We will model two kind of attacks of the attacker:
+\begin{enumerate}
+\item The attacker drops the virus on a random node on the network
+\item The attacker drops the virus on a targeted node on the network
+\end{enumerate}
+The attacker in this game will put a virus or worm on one of the nodes in the network.
+This will happen at random.
+The attacker does not know on which node the virus will be dropped.
+We will use this randomness because \todo{feit uit security rapport symantec} most virusses are spread via a usb stick or a shared resource.
+If we use this spreading method where we have a targeted attack the attacker will have more information about the network.
+\\
+
+The attacker can choose at which rate it will drop a virus on one of the nodes on the network.
+The cost of dropping a virus will be the same.
+It will not increase.
+If it will increase this means that the attacker will eventually drop out of the game because it becomes to expensive.\\
+The attacker is in control over the game if it manages to infect a subset of all the resources of the company network.
+
+\subsection{Actions of the defender}
+The attacker wants to protect all the nodes of his network.
+It can do so by getting back control over the resources.
+We will assume that the defender of the network has knowledge over his own network.
+Which is convenient in the real world because a company has to know how his infrastructure looks like.\\
+
+The defender has two possible ways of defending its network:
+\begin{enumerate}
+\item The defender flips all the nodes of his network
+\item The defender will flip a subset of the nodes of his network
+\end{enumerate}
 
@@
 
-\subsection{Our Game parameters}
+\subsection{Formal definition}
 \begin{description}
@@
 
-\subsection{Different kind of games}
-A virus has different kind of ways of making his way through a company network.
-We will describe the different ways of how the virus can propagate.
-For start we will say that the virus or worm will be dropped on Node i and that it has k numbers of neighbours.
-\begin{enumerate}
-\item Node i is infected and will spread the virus or worm to every k neighbours and will stop infecting the neighbours in the next step
-\item Node i is infected and will spread the virus or worm to every k neighbours and will keep on spreading the virus in every next step
-\item Node i is infected and will spread the virus to only one of the k neighbours and will stop infecting another one in the next step
-\item Node i is infected and will spread the virus to only one of the k neighbours and in the next step it will infect another one of the k neighbours 
-\end{enumerate}
-
-In the game that will be modelled in the paper we will use the settings of the 1 spreading method.
-We will not use method 2 because this kind of propagation will float the network.
-Because we use the settings of a mail system and contact in a mailling list the method of 3 and 4 are not used.
-\\
-In the first method the node that has been infected can be again infected.
-If one of the neighbours infect the node again the node will infect his neighbours again.
-By using this spreading method we have three states of a node.
-An infected state, a clean state and a spreading state.
-And infected state means that the node is infected and will not spread itself again to its neighbours, a clean state means that there is no virus or worm on the node and a spreading state means that the node is infected and that it will spread itself in the next step to its neighbours.
-We can argument this kind of propagation through a mail worm. \todo{voorbeeld geven van zo een worm}
-%Another propagation method is that the virus works as a token. It will propagate to only one neighbour and continue to spread.
-
 \section{Conclusion}

### 2015-03-14 doc/template/chap-2.tex
commit fouten verbeteren

@@
 The shared resource can be a password, a network or a secret key depending on the setting being modelled.
-In the rest of the paper we will call the players the Attacker and the Defender.
-To get the control over the resource, players can flip the resource at any given time.
-Each move will imply a certain cost.
-The unique feature of \flip{FlipIt} is that the move will happen in a stealthy way, meaning that the other player has no clue that the other player has flipped the resource.
+In the rest of the paper we will call the two players the Attacker and the Defender.
+To get control over the resource, the players can flip the resource at any given time.
+A flip will be regarded as a move from a player.
+Each move will imply a certain cost and the cost can vary for each player.
+Both players will try to minimize their cost.
+By adding a cost will prevent players to move to frequently.
+The unique feature of \flip{FlipIt} is that every move will happen in a stealthy way, meaning that the other player has no clue that the other player has flipped the resource.
 For instance, the defender will not find out if the resource has already been compromised by the attacker, but he can only potentially know it after he flips the resource himself.
 The goal of the player is to maximize the time that he or she has control over the resource while minimizing total cost of the moves.
-Players won't move to frequently.
 A move can also result in a "wasted move", called a flop.
@@
 \begin{itemize}
-\item Rate-of-play (RP: The player finds out the exact rate of play of the opponent.
+\item Rate-of-play (RP): The player finds out the exact rate of play of the opponent.
 \item Knowledge-of-strategy (KS): The player finds out the complete information of the strategy that the opponent is playing.
@@
 \item Node i is infected and will spread the virus or worm to every k neighbours and will stop infecting the neighbours in the next step
-\item Node i is infected and will spread the virus or worm to every k neighbours and will keep on spreading the virus in every next step
-\item Node i is infected and will spread the virus to only one of the k neighbours and will stop infecting another one in the next step
+\item Node i is infected and will spread the virus or worm to every k neighbours and will keep on spreading the virus to the same neighbours in every next step
+\item Node i is infected and will spread the virus to only one of the k neighbours and will stop infecting another neighbour in the next step
 \item Node i is infected and will spread the virus to only one of the k neighbours and in the next step it will infect another one of the k neighbours 
@@
 
-In the game that will be modelled in the paper we will use the settings of the 1 spreading method.
+In the game that will be modelled in the paper we will use the settings of the first spreading method.
 We will not use method 2 because this kind of propagation will float the network.
@@
 In the first method the node that has been infected can be again infected.
-If one of the neighbours infect the node again the node will infect his neighbours again.
-By using this spreading method we have three states of a node.
-An infected state, a clean state and a spreading state.
-And infected state means that the node is infected and will not spread itself again to its neighbours, a clean state means that there is no virus or worm on the node and a spreading state means that the node is infected and that it will spread itself in the next step to its neighbours.
+If one of the neighbours infects the node again the node will infect his neighbours again.
+By using this spreading method we have three distinct states in which a node can be situated.
+An \textit{infected state}, a \textit{clean state} and a \textit{spreading state}.
+An infected state means that the node is infected and will not spread the virus to its neighbours, a clean state means that the node is not infected on that moment and a spreading state means that the node is infected and that it will spread the virus or worm to its neighbours in the next step.
 We can argument this kind of propagation through a mail worm. \todo{voorbeeld geven van zo een worm}
@@
 
-We will model two kind of attacks of the attacker:
+The Attacker itself has two different ways of attacking the company network.
+It will only infected one node of the network and will wait for the virus to spread itself through the network.
+We will model two ways of attacks of an Attacker:
 \begin{enumerate}
@@
 The attacker in this game will put a virus or worm on one of the nodes in the network.
-This will happen at random.
-The attacker does not know on which node the virus will be dropped.
-We will use this randomness because \todo{feit uit security rapport symantec} most virusses are spread via a usb stick or a shared resource.
+(This will happen at random.) The attacker does not know on which node the virus will be dropped.
+We will use this randomness because \todo{feit uit security rapport symantec} most viruses are spread via a usb stick or a shared resource.
 If we use this spreading method where we have a targeted attack the attacker will have more information about the network.
@@
 
+The cost of flipping all the nods of the network will be greater than the cost of flipping a subset of nodes.
+We make this assumption because otherwise it will be beneficial for the defender to always flip all the nodes in the network.\\
+
+We will also make the assumption that as a defender flips a node the node can get infected again.
+A flip will not be  correlated to a patch but to a clean-up.
+\todo{waarom geen patch, wormen kunnen veranderen gaandeweg}
+\todo{andere mogelijkheid:} Another setting of the game can be that the flip of the defender is equal to a patch and that the resource cannot be infected any more.
+But with this case we deviate from the flipIt game, because the attacker cannot flip the resource any more.
+Unless we work with different virusses every time the attacker flips.
+We start with the less complex game of flipping is equal to a clean-up.
+
+\subsection{Strategies of both players}
+We explained what the actions of each player are.
+
 \section{Formal definition Game}
 
-In this section we provide the formal definition of the game and the notation that we will use throughout the paper.
+In this section we provide a formal definition of the game and the notation that we will use throughout the paper.
 

### 2015-03-29 doc/template/chap-1.tex
thesis followup

--- 2015-02-18 doc/template/chap-1.tex
@@
-\chapter{Intoduction to GameTheory}
-\label{cha:1}
-%\documentclass[10pt]{article}
-%\begin{document}
+%\chapter{Intoduction to GameTheory}
+%\label{cha:1}
+\documentclass[10pt]{article}
+\begin{document}
 
-In this chapter an introduction to gametheory will be given with the formulas that will be used throughout this paper.
+In this chapter an introduction to gametheory will be given with the formulas that will be used troughout this paper.
 We start with the basics of gametheory.
 People that have a background in gametheory can skip this chapter.
-\section{Intro to Gametheory}
-
-Game Theory is the study of interactions between independent and self-interested agents.
-It is a set of tools to analyze situations in which the optimal action for an agent depends on anticipations that form over another agent decision \todo{citatie vanuit wikipedia}.
-An agent can be either an individual, an animal or an enterprise.
-Game theory can be applied on various domains such as politics, economics, biology, sociology.
-In game theory studies the strategic interaction between players in a game.
-A game is played by a number of players and each of these players has actions.
-These actions are represented by a number indicating their preference.
-
-Ways of defending a network:
-\begin{itemize}
-\item Self-defending networks: The next generation of network security
-\item Honeynet games: a game theoretic approach to defending network monitors
-
-\end{itemize}
 
@@
 %http://ants.iis.sinica.edu.tw/3BkMJ9lTeWXTSrrvNoKNFDxRm3zFwRR/17/04483668.pdf
-\item How Viruses and worm can be detected. Difference between UDP en TCP worm propagation
+\item How Viruses and worm can be detected. Difference between UDP en TCP worm propagation. Difference for the propagation speed. 
 \end{itemize}
+
+Purpose thesis: model a worm propagation with adaptations of Flip-It.
+Flip-It cannot address the evaluation of individual nodes.
+Flip-It with multiple resources has not addressed the fact that a virus does not need to compromise the whole network.
+A subpart of the network can already cause problems for the company.
+Data leackages.
+So the virus can propagate
 
@@
 
-%\end{document}
+\end{document}

### 2015-03-29 Thesisfollowup/ThesisFollowUpSophieMarien.tex

@@
 \section{Schrijf een eerste versie van een hoofstuk over een afgewerkte bijdrage van je masterproef.}
+Dit is meegegeven in bijlage
 \section{Geef een korte inhoud van het tweede artikel dat je gaat schrijven. (Dit is niet van toepassing voor de studenten van de Master Toegepaste Informatica)}
+Het tweede artikel gaat gaan over mijn implementatie van FlipIt op een bedrijfsnetwerk.
+Eerst bespreek ik hoe ik FlipIt aanpas.
+Daarna stel ik een methode voor via matrices om de propagatie van een virus te beschrijven.
+De methode die gebruikt wordt is gelijkaardig aan die van Google Page Ranking.
+Via deze functie kan ik de totale controle berekenen van een speler over een spel.
+In het artikel leg ik de methode uit aan de hand van een default spel.
 \section{Beschrijf in 10 lijnen de demo die je bij je verdediging zult geven.}
+Ik heb een theoretische thesis dus kan ik moeilijk een demo geven. Ik zal enkele voorbeelden uitwerken om mijn werkingsmethode uit te leggen.
 \section{Wat is er afgesproken met je begeleider en/of promotor om op te leveren als major milestone die het praktische werk van je thesis afsluit?}
+
 \section{Evalueer de tijd die je in je thesis gestoken hebt in de voorbije 4 weken: veel te weinig / te weinig / voldoende / ruim voldoende / meer dan verwacht.}
+Ik heb voldoende tijd in mijn thesis gestoken. 
 \section{Evalueer de vorderingen van je thesis t.o.v.

### 2015-03-29 doc/template/chap-2.tex

--- 2015-03-03 doc/template/chap-2.tex
@@
 The attacker does not know on which node the virus will be dropped.
-We will use this randomness because \todo{feit uit security rapport symantec} most virusses are spread via a usb stick or a shared resource.
+We will use this randomness because \todo{feit uit security rapport symantec} most viruses are spread via a usb stick or a shared resource.
 If we use this spreading method where we have a targeted attack the attacker will have more information about the network.

### 2015-03-29 doc/template/chap-5.tex

@@
 \section{Virusses}
+
+Stealth
+Regin's developers put considerable effort into making it highly inconspicuous.
+Its low key nature means it can potentially be used in espionage campaigns lasting several years.
+Even when its presence is detected, it is very difficult to ascertain what it is doing.
+Symantec was only able to analyze the payloads after it decrypted sample files.
+
+It has several ''stealth'' features.
+These include anti-forensics capabilities, a custom-built encrypted virtual file system (EVFS), and alternative encryption in the form of a variant of RC5, which isn't commonly used.
+Regin uses multiple sophisticated means to covertly communicate with the attacker including via ICMP/ping, embedding commands in HTTP cookies, and custom TCP and UDP protocols
+%http://www.symantec.com/connect/blogs/regin-top-tier-espionage-tool-enables-stealthy-surveillance
 Ways of defending a network:

### 2015-03-29 doc/template/intro.tex

@@
 defined and the modus operandi is explained.
+\todo{bib referenties in orde brengen}
+\section{Introduction}
 
-\section{chap}
+Security is an important asset in the computer science.
+Technology is growing fast and so are the malicious people.
+Defending a network of a company is not very easy.
+It makes use of firewalls, routers, IDS systems, virus scans, and so on.
+Protecting a company network is difficult job.
+They are often the victim of targetted attacks.
+In a security report of 2014, \todo{verwijzing naar report}, state that 80\% of the companies are the vitims of targeted attacks.
+Many corporate networks have to continuously defend themselves against outside invaders such as viruses and worms.
+The network administrator will try to keep the network to malware-free as possible.
+If still there is an intruder managed to penetrate the network then the network manager this intruder trying to get out as quickly as possible.
+This is not always easy.
+Especially when the intruders secretly sneak and then spread rapidly.
+In this paper we will work further on the work made by Marten van Dijk, Ari Juels, Alina Oprea and Ronals L.
+Rivest \todo{verwijzing naar FlipIT} who wrote a report on the Game FlipIt.
+FlipIt is a the game of '' Stealthy Takeovers''.
+It models a game by means of two players, the attacker and the defender.
+Both can gain control over a single shared resource by flipping it.
+The most important property of the game is that the flipping happens stealthy.
+This means that the players have no clue about when the other player moves.
+The goal of the game is to maximise the time the player controls the resource minus the average cost of the flipping.
+
+\subsection{Motivation of the game}
+%waarom FlipIt gebruiken en niet iets anders.
+\subsection{Contributions and results}
+\subsection{Conclusions}
+The "I love you" virus is an example of a virus that spreads quickly.
+This virus propagates via mail systems.
+If someone opens an email with "I love you" virus in annex this virus spreads itself by sending a mail itself to everyone in your contact list.
+So the virus can multiply rapidly and eventually a business network shut down by the heavy traffic.
+In this example, there is a need human interaction to spread the virus to do.
+If no one opens the virus can not spread the mail.
+Unfortunately, there are viruses that can spread without human interaction.
+These viruses are referred to as worms.
+A worm is also a computer program that replicates itself to spread to other computers so.
+Via a computer network, copies of the worm forwarded without an intermediary is used for.
+The worm will use vulnerabilities to infect other computers.
+Most worms are designed to spread out and just try not to make any changes to the systems that they pass.
+These worms can still inflict damage by increased network traffic they generate.
+Worms that contain Harm damage a program to install a backdoor or a rootkit on the infected computers.
+Backdoors and rootkits ensure that future use can be made of the infected computers.
+The Stuxnetworm is a very famous worm.
+Initially this worm spread via infected USB sticks and from then it could spread through the Internet to other computers.
+The purpose of the Stuxnetworm was broken to run the centrifuges in nuclear reactors.
+Many reactors have been infected.
+From the standpoint of the defender, it is very important to respond as quickly as possible so that the worm can not spread quickly.
 

### 2015-03-31 Thesisfollowup/IEEE.tex
IEEE first stuff

\documentclass[a4paper, 11pt]{article}
\setlength{\topmargin}{-.5in}
\setlength{\textheight}{9in}
\setlength{\oddsidemargin}{.100in}
\setlength{\textwidth}{6.25in}
\usepackage [dutch] {babel}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{fullpage}
\usepackage{listings}
\usepackage{color}
\usepackage{soul}
\usepackage{tikz}
\usepackage{hyperref}
\usepackage{parskip}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\lstset{frame=tb,
  language=Java,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=true
  tabsize=3
}

\newcommand{\todo}[1]{{\huge \textcolor{green}{#1}}\\}
\newcommand{\comment}[1]{\textcolor{red}{#1}\\}
\newcommand{\viscomment}[1]{\textcolor{red}{#1}}
\newcommand{\flip}[1] {\textcolor{black}{#1}}

\begin{document}

\begin{titlepage}
\title{IEEE: artikel}
\author{Sophie Marien}
\date{} 
\end{titlepage}

\maketitle
\section*{Abstract}
Recently the attacks on Belgacom and other high profile targeted attacks shows us that even the most secure companies can still be compromised. It also shows that these attacks are not immediately detected. FlipIt is a framework that can model these stealthy takeovers and is proposed by a group of researchers at RSA. It is a 2-players game composed of a single attacker, a single defender and a singe shared resource. These players will try to gain control over the shared resource and they do this in a stealthy way. In this paper we try to adapt FlipIt in a way that we can use it to model the game of defending a company network. This network will consist of multiple shared resources instead of one. Through analytic results we can provide useful information for the defender to defend his network.

\section{REst}

 Not so long ago we discovered some high profile targeted attacks and the NSA who infiltrated in a lot of systems without anyone noticing it. This means that even the most secure company can still be compromised without even noticing it. Such an attack can gain access to vulnerable data, credentials, cryptographic keys and many more. APT's are..

 The problem that we want to address in this article is how to react as a defender if our network system can be compromised by an attacker in a stealthy way. A research group at the RSA proposed a game to model the scenario where the compromising the system is done stealthy and not immediately noticed by the system manager. In this FlipIt game we have 2 players, the attacker and the defender, and a shared resource.
 Both players can gain at any time control of the shared resource, but this will always happen in a secretly manner. None of the players can observe who is in control of the resource. Every unit of time that a player is in control it gets utility. A player can 'flip' a resource which means that he is taking control over the resource. Each time the player flips the resource it has to pay a fixed cost.
 In this paper we want to model the problem of a corporte network that is under attack daily by virusses and worms. To do so we have to adapt the basic FlipIt game. Instead of one shared resource we have multiple resources that represent the nodes or devices in a corporate network. Each of these nodes are connected to other nodes by a graph. \comment{email systeem }.

 The actions of the defender are. The actions of the attacker are. When system compromised

 \subsection{Basic model}
 In our basic model the defender can only 'flip' all the nodes in the network at the same time. The attacker will always drop his virus on the most \comment{woord vinden} node. The virus will then spread itself one time to all of his neighbours. Once infected again the virus will again spread itself to all of his neighbours. The network is compromised by the attacker when all of the nodes of the network are infected.

 We can model the spreading of the virus through matrices. Google uses the same technique for page ranking. First the graph of the network is defined. Once the graph is known we can construct the first matrix for the calculation. If we have n nodes in our graph, the matrix will be of size nxn. The columns and rows correspond with the nodes. Every $n_{ij}$ entry of the matrix will be set on 1 if there is a connection between node i and node j. The diagonal of the matrix will also be set on 1.

 Now we know how we can model the spreading of the virus. If we want to set-up a gain function we have to know how ling it takes for the virus to compromise the whole system. This will be equal to the shortest path from the start infected node to the node that is the furthest away from this node. With this information we can calculate after how many steps this node is infected through our matrix calculation. We will end up with kind of a delayed FlipIT
\end{document}

### 2015-04-03 Thesisfollowup/ThesisFollowUpSophieMarien.tex

@@
 \section{Vergelijk wat je gedaan hebt met het opgestelde werkplan. Hoever sta je? Moet het werkplan aangepast worden?}
+Ik moet een tandje bijsteken en al verder beginnen schrijven. Ik was vergeten in mijn werkplan bij te steken dat ik een week op Athens ging gaan. 
 \section{Schrijf een eerste versie van een hoofstuk over een afgewerkte bijdrage van je masterproef.}
-Dit is meegegeven in bijlage
+Dit is meegegeven in bijlage. Ik heb ineens heel mijn voorlopige thesistekst in bijlage gestoken. 
 \section{Geef een korte inhoud van het tweede artikel dat je gaat schrijven. (Dit is niet van toepassing voor de studenten van de Master Toegepaste Informatica)}
@@
 In het artikel leg ik de methode uit aan de hand van een default spel.
+Ook geef ik een eerste setting van een spel en de bijhorende benefit formules waaruit ik al dan niet een equilibrium kan bepalen.
 \section{Beschrijf in 10 lijnen de demo die je bij je verdediging zult geven.}
@@
 \section{Wat is er afgesproken met je begeleider en/of promotor om op te leveren als major milestone die het praktische werk van je thesis afsluit?}
-
+Dit is nog niet afgesproken.
 \section{Evalueer de tijd die je in je thesis gestoken hebt in de voorbije 4 weken: veel te weinig / te weinig / voldoende / ruim voldoende / meer dan verwacht.}
-Ik heb voldoende tijd in mijn thesis gestoken. 
+De voorbije weken heb ik minder dan voldoende aan mijn thesis gewerkt. Ik liep vast op een aantal zaken en ik ben een week op Athens cursus geweest. 
 \section{Evalueer de vorderingen van je thesis t.o.v.
 het bereiken van de vooropgestelde doelstelling: ver achter op het schema / achter op het schema / goed op schema / vooruit op het schema / ver vooruit op het schema.}
+Ik loop een beetje achter op schema want ik had toch graag nu al een aantal formuleringen bepaald en een model uitgewerkt.
 \section{Stuur de ingevulde timesheet mee als bijlage.}

### 2015-04-03 doc/template/chap-3.tex

--- 2015-01-25 doc/template/chap-3.tex
@@
 
+\section{Gametheory}
+
+Game theory is the study between self interested agents and the actions they perform.
+self interested agents means not that they want to harm each other or that they only care about themselves.
+Agent has a description of the world he likes.
+A utility function is a mapping of the world that the agent likes to real numbers.
+
 \section{Literatuurstudie}
+
+Difference with FlipThem: sub part of nodes for control and strategy difference: dependant of grade of the nodes, instead of just periodic.
+- Literatuurstudie
+- Flipit
+- Game Motivation
+- Formal definition
+
 \begin{description}
-\item Distributed Worm Simulation with a REalistic Internet 2005  \\
-Modeling of congestions of network through worm propagation. Mathematical model focussing on the underlying network infrastructure.(diff no game theory) 
-\item Of threats and Costs: A Game-theoretic approach to security risk managment 2013\\
+\item Distributed Worm Simulation with a Realistic Internet 2005  \\
+Modelling of congestions of network through worm propagation. Mathematical model focussing on the underlying network infrastructure.(diff no game theory) 
+\item Of threats and Costs: A Game-theoretic approach to security risk management 2013\\
 Model network security of networks with a non-cooperative node through game theory.
-Attacker knows the defense strategies and the defender has knowledge of the possible attacks.
+Attacker knows the defence strategies and the defender has knowledge of the possible attacks.
 Each actor considers the actions of the other before deciding to strive to optimize their own utility.
@@
 \item Game theory meet network security and privacy (2013) \\
-Chapter 3 addresses several games in game theory for modeling network security.
+Chapter 3 addresses several games in game theory for modelling network security.
 \item Game theoretic approach for cost-benefit analysis of malware proliferation prevention (..)
@@
 \section{Why Game Theory to model security problem}
-Actors in a security protocol must follow the systems and some arbitrarily actors that are malisious and do not follow the protocol.
-[Bridging GameTheory and Cryptography].
-Gametheoretic approach proposes a model where all the actors act with self-interest.
+Actors in a security protocol must follow the systems and some arbitrarily actors that are malicious and do not follow the protocol.
+[Bridging Game Theory and Cryptography].
+Game theoretic approach proposes a model where all the actors act with self-interest.
 
@@
 Many ways to attack a company.
-Virusses, trojans, worms, DOS, ..
+Viruses, Trojans, worms, DOS, ..
 Hard to protect against every attacker.

### 2015-04-03 doc/template/intro.tex

@@
 
-Security is an important asset in the computer science.
-Technology is growing fast and so are the malicious people.
-Defending a network of a company is not very easy.
-It makes use of firewalls, routers, IDS systems, virus scans, and so on.
-Protecting a company network is difficult job.
-They are often the victim of targetted attacks.
-In a security report of 2014, \todo{verwijzing naar report}, state that 80\% of the companies are the vitims of targeted attacks.
-Many corporate networks have to continuously defend themselves against outside invaders such as viruses and worms.
-The network administrator will try to keep the network to malware-free as possible.
-If still there is an intruder managed to penetrate the network then the network manager this intruder trying to get out as quickly as possible.
+Security is an important asset in Computer Science.
+Defending a network of a company is not an easy job.
+To prevent intruders it can make use of firewalls, routers, IDS systems, virus scans, and other defence mechanisms.
+Unfortunately technology is growing fast and attacks are getting more sophisticated and the causes of these attacks can be very different.
+\todo{iets tussen nog} Companies are often the victim of targeted attacks.
+In a security report of 2014, \todo{verwijzing naar report}, states that 80\% of the companies are the victims of targeted attacks.
+Many companies don't see themselves as a target, but sometimes they might be collateral, the target on the way to the real target.
+This means that everybody can be a target.
+Corporate networks should continuously defend themselves against outside invaders such as viruses and worms.
+By doing so the network administrator can keep the network as malware-free as possible.
+If there is  an intruder managed to penetrate the network then the network manager this intruder trying to get out as quickly as possible.
 This is not always easy.
@@
 
+\section{introduction number 2}
+
+(We live in an era) In this era where digitalization becomes prominent in every aspect of our lives, where technology is growing fast and where business are always under attack, security becomes an issue of increasing complexity.
+Since 2009, the number of reported security attacks has increased 66\%, year over year.
+\todo{security report van pwc}.
+These numbers only represent the attacks that are detected.
+In 2014 117,339 attacks where coming in daily.
+Many of those attacks have a different cause.
+Some of them can be benign, others can be harmful.
+Many companies are unaware of all the attacks.
+Some of them think that they are not a target, but they might be a target on the way to a real target.
+Recently there where some high profiled targeted attacks which have been revealed.
+(Belgacom).
+Targeted attacks are ...
+The \textit{Kill Chain} is a concept by Lockheed Martin Corporation, explained in the whitepaper \todo{withepaper toevoegen}.
+It explains the different phases of a typical attack from the view of an attacker.
+It also outlines the typical attacker activities on the right.
+This model is very useful to define the different moments of the life cycle of an attack and when a company should act to defend itself.
+In this paper we would like to prevent the viruses of spreading into the network system of a company.
+This means that we have to act in phase Installation, Command and Control and Action on Objectives of the kill chain.
+
+  Security is an important asset in Computer Science.
+Defending a network of a company is not an easy job.
+To prevent intruders it can make use of firewalls, routers, IDS systems, virus scans, and other defence mechanisms.
+Unfortunately technology is growing fast and attacks are getting more sophisticated and the causes of these attacks can be very different.
+\todo{iets tussen nog} Companies are often the victim of targeted attacks.
+In a security report of 2014, \todo{verwijzing naar report}, states that 80\% of the companies are the victims of targeted attacks.
+Many companies don't see themselves as a target, but sometimes they might be collateral, the target on the way to the real target.
+This means that everybody can be a target.
+Corporate networks should continuously defend themselves against outside invaders and targeted attacks.
+Researchers have already investigated the situations through the FlipIt game in which a system is continuously compromised by an attacker through targeted attacks.
+FlipIt is a the game of '' Stealthy Takeovers''.
+It models a game by means of two players, the attacker and the defender.
+Both can gain control over a single shared resource by flipping it.
+The most important property of the game is that the flipping happens stealthy.
+This means that the players have no clue about when the other player moves and has control over the shared resource.
+The goal of the game is to maximise the time the player controls the resource minus the average cost of the number of flipping.
+In this paper we model a company network through multiple shared resources and a flip from the attacker that drops a virus that will spread itself autonomously. We show that ...
+%http://www.pwc.com/gx/en/consulting-services/information-security-survey/index.jhtml
 %%% Local Variables: 

### 2015-04-03 Thesisfollowup/IEEE.tex
IEEE artikel aangevuld

@@
 \usepackage{parskip}
+\usepackage{multicol}
 
@@
 
- Not so long ago we discovered some high profile targeted attacks and the NSA who infiltrated in a lot of systems without anyone noticing it.
+ Not so long ago we discovered some high profile targeted attacks that infiltrated in a lot of systems without anyone noticing it.
+Belgacom [] has been the victim of a cyber security attack that has been uncovered for almost 2 years.
+The communication infrastructure of the Telecom company was infected with highly sophisticated malware making it possible to eavesdrop remotely.
 This means that even the most secure company can still be compromised without even noticing it.
 Such an attack can gain access to vulnerable data, credentials, cryptographic keys and many more.
-APT's are..
+These kind of attacks that target organisations in a stealthy way and stay undetected for a long period are called APT (Advanced Persistent Threats).
 
+\comment{uitleggen waarom we gametheorie gaan gebruiken om het spel te modelleren}
  The problem that we want to address in this article is how to react as a defender if our network system can be compromised by an attacker in a stealthy way.
-A research group at the RSA proposed a game to model the scenario where the compromising the system is done stealthy and not immediately noticed by the system manager.
-In this FlipIt game we have 2 players, the attacker and the defender, and a shared resource.
+A research group at the RSA proposed a game to model the scenario where compromising the system is done stealthy and not immediately noticed by the system manager.
+In this FlipIt [1] game we have 2 players, the attacker and the defender, and a shared resource.
  Both players can gain at any time control of the shared resource, but this will always happen in a secretly manner.
@@
 Each time the player flips the resource it has to pay a fixed cost.
- In this paper we want to model the problem of a corporte network that is under attack daily by virusses and worms.
+ In this paper we want to model the problem of a corporate network that is under attack daily by viruses and worms.
 To do so we have to adapt the basic FlipIt game.
@@
 We will end up with kind of a delayed FlipIT
+
+\section{References}
+\begin{itemize}
+\item \textit{FlipIt: Game of stealthy takeovers}
+\end{itemize}
 \end{document}

### 2015-04-10 Thesisfollowup/IEEE.tex
commit

@@
  Not so long ago we discovered some high profile targeted attacks that infiltrated in a lot of systems without anyone noticing it.
-Belgacom [] has been the victim of a cyber security attack that has been uncovered for almost 2 years.
+Belgacom \comment{[]} has been the victim of a cyber security attack that has been uncovered for almost 2 years.
 The communication infrastructure of the Telecom company was infected with highly sophisticated malware making it possible to eavesdrop remotely.
@@
 A research group at the RSA proposed a game to model the scenario where compromising the system is done stealthy and not immediately noticed by the system manager.
-In this FlipIt [1] game we have 2 players, the attacker and the defender, and a shared resource.
+In this FlipIt \comment{[1]} game we have 2 players, the attacker and the defender, and a shared resource.
  Both players can gain at any time control of the shared resource, but this will always happen in a secretly manner.
@@
 Each of these nodes are connected to other nodes by a graph.
-\comment{email systeem }.
+The connections represent the logical communication connections like a person in a contact list of an email application.
+The second adaptation is that when the attacker flips, it will drop a virus on that node.
+This virus will spread autonomously.
+ In the following section we will describe the actions that the attacker and the defender can play.
 
+ \subsection{actions of the defender}
+ Dit is een generiek probleem. Het komt terug. 
  The actions of the defender are. The actions of the attacker are. When system compromised

### 2015-04-12 Thesisfollowup/IEEE.tex
extra schrijven over nash

@@
 These kind of attacks that target organisations in a stealthy way and stay undetected for a long period are called APT (Advanced Persistent Threats).
+A good definition of an ATP is given by Bruce Schneier:
+ “A conventional hacker or criminal isn't interested in any particular target.
+He wants a thousand credit card numbers for fraud, or to break into an account and turn it into a zombie, or whatever.
+Security against this sort of attacker is relative; as long as you're more secure than almost everyone else, the attackers will go after other people, not you.
+An APT is different; it's an attacker who – for whatever reason – wants to attack you.
+Against this sort of attacker, the absolute level of your security is what's important.
+It doesn't matter how secure you are compared to your peers; all that matters is whether you're secure enough to keep him out.” - Bruce Schneier: APT is a Useful Buzzword.
 

### 2015-04-12 doc/template/chap-5.tex
extra schrijven over nash

@@
 
-In the following paragraph an introduction to game theory is given based on the work of \citep{leyton2008essentials} and \citep{Coursera}.
-For a more detailed and full introduction to game theory, the reader is referred to \citep{leyton2008essentials}.
+In the following paragraph an introduction to game theory is given based on the work of 
+{leyton2008essentials} 
+and 
+{Coursera}. 
+For a more detailed and full introduction to game theory, the reader is referred to 
+{leyton2008essentials}. 
 %------------------------------------------------%
@@
 (\todo{uitleggen aan de hand van een voorbeeld}).\\
+\todo{players rationeel en max outcomes}
 %Cooperative and non cooperative games
-In a decision game theoretic approach an agent will try to act in such a way to maximise his expected or average utility function.
+In a Decision Game Theoretic Approach an agent will try to act in such a way to maximise his expected or average utility function.
 It becomes more complicated when two or more agents want to maximise their utility and whose actions can affect each other utilities.
@@
 
+\todo{strategien en acties definieren}
+
 %Nash equilibrium
+
+In game theory, the Nash equilibrium is a solution concept of a non-cooperative game involving two or more players, in which each player is assumed to know the equilibrium strategies of the other players, and no player has anything to gain by changing only their own strategy.[1] If each player has chosen a strategy and no player can benefit by changing strategies while the other players keep theirs unchanged, then the current set of strategy choices and the corresponding payoffs constitutes a Nash equilibrium.
+The reality of the Nash equilibrium of a game can be tested using experimental economics method.
+
+In de speltheorie, een deelgebied van de wiskunde, is een Nash-evenwicht een oplossingsconcept voor een niet-coooperatief spel, waar twee of meer spelers aan meedoen.
+In een Nash-evenwicht wordt elke speler geacht de evenwichtsstrategieeen van de andere spelers te kennen en heeft geen van de spelers er voordeel bij om zijn of haar strategie eenzijdig te wijzigen.
+Als elke speler een strategie heeft gekozen en geen enkele speler kan profiteren door zijn strategie te veranderen, terwijl de andere spelers dat ook niet doen, dan vormt de huidige verzameling van strategiekeuzes plus de bijbehorende uitbetalingen een Nash-evenwicht.
+
+Een Nash-evenwicht gaat uit van een spel, waarin iedere speler een strategie heeft.
+Die strategie geeft precies aan wat een speler in de verschillende fases van een spel doet.
+Een strategie kan zowel een pure strategie als een gemengde strategie zijn.
+De verzameling van strategieeen van alle spelers die meedoen aan een bepaald spel noemt men een strategieprofiel.
+In de speltheorie is een Nash-evenwicht een strategieprofiel waarbij het voor geen enkele speler voordelig is daarvan af te wijken, als de andere spelers dat ook niet doen.
+
+Het Nash-evenwichtsconcept is een begrip dat vooral toepassing vindt in de economie.
 
@@
 
+Spy vs Spy: Aldrich Ames was a CIA Counter-Intelligence officer.
+He was also a spy feeding valuable intelligence to the Soviets and compromising US intelligence operations in the Soviet Union.
+He operated for 9 years before the CIA recognized that they had a spy and began an investigation and determined that he was the leak.
+This strategic situation is the same one faced by computer networks, drug cartels, intelligence agencies and guerrilla networks.
+
+All such organisations have a reasonable expectation that trusted personal/systems will eventually be recruited/captured by enemy organisations.
+Therefore such organisations must consume valuable resources to discover such betrayals and thereby regain secrecy.
+The question is then given the possible threats how often and at what cost should they spend resources on investigations/spy hunts/virus scans.
+This is where flipIt comes in.
+
+FLIPIT: The Game of “Stealthy Takeover:” FlipIt was created to model these sorts of strategic situations and to study the best courses of action.
+Specifically flipIt was motivated by the recent interest in and success of Advanced Persistent Threats, or APT.
+
+The basic idea is that given the current experience that perfect protection of trusted resources is unattainable, lets think about how we can optimally manage compromises of the our most trusted systems.
+
+Rules
+
+Two players, player X (blue) and player Y (red) attempt to maintain control over a shared resource.
+At anytime in the game each player is allowed to play ‘flip’.
+The only way a player can learn the state of the game (who is in control) is when they play 'flip’.
+If a player is in control of the resource and they play 'flip’ they remain in control of the resource.
+If a player is not in control of the resource and they play 'flip’ they gain control of the resource.
+Players gain points for the length of time they control the resource.
+Players lose points every time they play flip.
+This reflects the situation that the CIA is placed in with regard to moles/enemy spies.
+They don't know if they have been compromised.
+They can perform an investigation and determine if they have been compromised, also catching the spy in the act, but this action is very expensive.
+That is, the CIA has to trade off between remaining “mole free” (a good) and investigations (an expense).
+
+Winning: How do you win a fair game of flipIt against intelligent adaptive human adversaries? I'm not sure.
+
+In the real world what is the best move given that the other players can secretly capture/corrupt your most trusted personal/systems? Rives suggests in his talk that you:
+Be prepared to deal with repeated total failure (loss of control).
+Play fast! Aim to make opponent drop out!
+Arrange game so that your moves cost much less than your opponent's!
 %http://www.cisco.com/web/offer/gist_ty2_asset/Cisco_2014_ASR.pdf p49

### 2015-04-16 doc/template/chap-5.tex
vanalles

@@
 
-In game theory, the Nash equilibrium is a solution concept of a non-cooperative game involving two or more players, in which each player is assumed to know the equilibrium strategies of the other players, and no player has anything to gain by changing only their own strategy.[1] If each player has chosen a strategy and no player can benefit by changing strategies while the other players keep theirs unchanged, then the current set of strategy choices and the corresponding payoffs constitutes a Nash equilibrium.
+%John Nash speelde ook een grote rol in de geschiedenis van de speltheorie.
+Hij is een van de wiskundigen geweest die speltheorie geformaliseerd heeft.
+Het Nash evenwicht werd naar hem vernoemd.
+Een Nash evenwicht wordt gezien als een evenwicht tussen beide spelers zodat ze allebei de beste tactiek kiezen en niet meer veranderen als de andere van tactiek veranderen.
+John Nash breide de theorie over het Nash evenwicht in een paper nog uit met gemengde strategieën.
+In 1994 kreeg John Nash samen met twee andere wiskundigen gespecialiseerd op het vlak van speltheorie de Nobelprijs voor de economie op basis van hun prestaties in de niet-coöperatieve speltheorie.
+.
+Over John Nash is een prachtige film
+\todo{Best response ook uitleggen?}
+\todo{Voorbeeld ook nog uitleggen?}
+One of the solution concepts in Game Theory for non-cooperative games is a Nash Equilibrium.
+A Nash Equilibrium is a subset of outcomes that can be interesting to analyse a game.
+For a Nash Equilibrium each player has a consist list of actions and each player's action maximizes his or her payoff given the actions of the other players.
+Nobody has the incentive to change his or her action if an equilibrium profile is played.
+In general we can say that a Nash Equilibrium is a stable strategy profile: each player is considered to know the equilibrium strategies of the other players and no player would want to change his own strategy if he knows the strategies of the other players.
+%In game theory, the Nash equilibrium is a solution concept of a non-cooperative game involving two or more players, in which each player is assumed to know the equilibrium strategies of the other players, and no player has anything to gain by changing only their own strategy.[1] If each player has chosen a strategy and no player can benefit by changing strategies while the other players keep theirs unchanged, then the current set of strategy choices and the corresponding payoffs constitutes a Nash equilibrium.
 The reality of the Nash equilibrium of a game can be tested using experimental economics method.
 
-In de speltheorie, een deelgebied van de wiskunde, is een Nash-evenwicht een oplossingsconcept voor een niet-coooperatief spel, waar twee of meer spelers aan meedoen.
+%In de speltheorie, een deelgebied van de wiskunde, is een Nash-evenwicht een oplossingsconcept voor een niet-coooperatief spel, waar twee of meer spelers aan meedoen.
 In een Nash-evenwicht wordt elke speler geacht de evenwichtsstrategieeen van de andere spelers te kennen en heeft geen van de spelers er voordeel bij om zijn of haar strategie eenzijdig te wijzigen.
@@
 
-Een Nash-evenwicht gaat uit van een spel, waarin iedere speler een strategie heeft.
+%Een Nash-evenwicht gaat uit van een spel, waarin iedere speler een strategie heeft.
 Die strategie geeft precies aan wat een speler in de verschillende fases van een spel doet.
@@
 
-Het Nash-evenwichtsconcept is een begrip dat vooral toepassing vindt in de economie.
+%Het Nash-evenwichtsconcept is een begrip dat vooral toepassing vindt in de economie.
 

### 2015-04-17 doc/template/chap-6.tex
Formule neerschrijven

\chapter{Intro to virus}
\label{cha:6}
%\documentclass[10pt]{article}\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%			Introduction Chapter 6			%%%%%%
%%%%%												%%%%%%
%%%%%												%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{tekst}

In this section we are going to elaborate how we are going to model a Flipit game with multiple resources and a virus that propagates and infects the resources. We come up with a formula for the normal FlipIt and then reform it to a FlipIt game with a virus.

\subsection{FlipIt with a virus}
In the previous version the FlipIt game is already been explained. In this section we will introduce the virus propagation. An attacker will drop a virus on one of the resources that is available. The virus will then spread itself to the neighbour resources. The attacker will only gain control over the whole network, in general the game, when it has infected a certain amount of resources. We will call this amount for now ''d''. If we want to measure how many time it takes for the virus to infect ''d'' resources, we have to calculate the shortest path to the ''d'' th node.  If we know how much time it takes to control ''d'' resources, we can model a normal FlipIt game but with a delay. The model will not be completely a FlipIt game with a delay, because if the delay is bigger than the period of the attacker, the attacker will gain no control. If it would be with the delay the attacker would gain control after the defender flipt again. \todo{laten zien met een figuur}. In the next section we are going to describe what the setting of the game is.

\subsection{chap}
The setting of the game that we are going to play is one with multiple resources. When the defender flips it will always flip all the resources. The attacker will flip the node in the graph that can infect all the nodes in the shortest time possible. The attacker will gain the control over the resources when all the resources are infected. So ''d'' will be the shortest path to the furthest node. We will model ''d'' in time units. \todo{moet misschien niet ? }.
\begin{figure}[hbtp]
\caption{Difference in a FlipIt game between delay caused by a virus and a delay for the Attacker}
\centering
\includegraphics[scale=0.3]{Images/diffVirusDelay.jpg}
\end{figure}

If \textit{i} is an Irrational number:
$ i \neq \dfrac{a}{b}$ with $b \in Z, a \in N$
Because we cannot write \textit{i} in a fraction, this means that we won't have a cycle. If we would have a cycle that means that we do have a number that divides \textit{i}. If we don't have a cycle it goes on forever. Meaning that it goes on to infinity. This also means that no number will be repeated two times. If it does that means that their is repetition, meaning again that their is a cycle. We can conclude that if we have no cycle and no number will be repeated twice, that it will enumerate every number between 0 and the biggest interval. \todo{interval definieren}
Even though time is continious, irrational numbers are not considered. We kunnen continue tijd voldoende dicht benaderen door enkel rationele getallen in beschouwing te nemen. Vlakbij elk Irrationaal getal ligt immers een Rationaal getal.

A $\delta$ that is irrational will have no cycle.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "thesis"
%%% End:

### 2015-04-17 Thesisfollowup/thesisfollowup.tex
latex file compileert niet met chap 5

--- 2015-02-27 Thesisfollowup/ThesisFollowUpSophieMarien.tex
@@
 \begin{titlepage}
-\title{Thesis Follow Up}
+\title{Thesis followup}
 \author{Sophie Marien}
-\date{} 
 \end{titlepage}
@@
 
-\section{Vergelijk wat je gedaan hebt met het opgestelde werkplan. Hoever sta je? Moet het werkplan aangepast worden?}
-\section{Schrijf een eerste versie van een hoofstuk over een afgewerkte bijdrage van je masterproef.}
-\section{Geef een korte inhoud van het tweede artikel dat je gaat schrijven. (Dit is niet van toepassing voor de studenten van de Master Toegepaste Informatica)}
-\section{Beschrijf in 10 lijnen de demo die je bij je verdediging zult geven.}
-\section{Wat is er afgesproken met je begeleider en/of promotor om op te leveren als major milestone die het praktische werk van je thesis afsluit?}
+\section{Wat zijn de belangrijkste elementen uit het besluit dat je gaat schrijven in je tekst?}
+Dat zal in de komende week duidelijk worden als ik kan aantonen dat mijn intuitie voor een bepaalde formule juist was. Want hiermee kan ik dan mijn FlipIt spel berekenen met een virus.
+\section{Verfijn de outline van je thesistekst. Geef per hoofdstuk de subsecties, en per subsectie het kernidee en een schatting van het aantal bladzijdes van die subsectie.}
+
 \section{Evalueer de tijd die je in je thesis gestoken hebt in de voorbije 4 weken: veel te weinig / te weinig / voldoende / ruim voldoende / meer dan verwacht.}
+De voorbije 4 weken heb ik voldoende gewerkt aan mijn thesis. Ik had veel tijd en heb die ook ten volle benuttigd voor mijn thesis. 
 \section{Evalueer de vorderingen van je thesis t.o.v.
 het bereiken van de vooropgestelde doelstelling: ver achter op het schema / achter op het schema / goed op schema / vooruit op het schema / ver vooruit op het schema.}
+Ik ben verder vooruit gekomen. Als ik dit kan aanhouden zit ik goed.
 \section{Stuur de ingevulde timesheet mee als bijlage.}

### 2015-04-17 doc/template/chap-5.tex
chap 5

@@
 
-FLIPIT: The Game of “Stealthy Takeover:” FlipIt was created to model these sorts of strategic situations and to study the best courses of action.
+FLIPIT: The Game of ''Stealthy Takeover:'' FlipIt was created to model these sorts of strategic situations and to study the best courses of action.
 Specifically flipIt was motivated by the recent interest in and success of Advanced Persistent Threats, or APT.
@@
 They can perform an investigation and determine if they have been compromised, also catching the spy in the act, but this action is very expensive.
-That is, the CIA has to trade off between remaining “mole free” (a good) and investigations (an expense).
+That is, the CIA has to trade off between remaining''mole free'' (a good) and investigations (an expense).
 

### 2015-04-17 doc/template/chap-5.tex
Latex werkt weer !!!

@@
 Two players, player X (blue) and player Y (red) attempt to maintain control over a shared resource.
-At anytime in the game each player is allowed to play ‘flip’.
-The only way a player can learn the state of the game (who is in control) is when they play 'flip’.
-If a player is in control of the resource and they play 'flip’ they remain in control of the resource.
-If a player is not in control of the resource and they play 'flip’ they gain control of the resource.
+At anytime in the game each player is allowed to play 'flip'.
+The only way a player can learn the state of the game (who is in control) is when they play flip.
+If a player is in control of the resource and they play flip they remain in control of the resource.
+If a player is not in control of the resource and they play flip they gain control of the resource.
 Players gain points for the length of time they control the resource.

### 2015-04-21 doc/template/chap-7.tex
Verder werken op APTs

\chapter{APT}
\label{cha:7}
%\documentclass[10pt]{article}\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%			Introduction Chapter 7			%%%%%%
%%%%%												%%%%%%
%%%%%												%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Advanced Persistent Threats}

A targeted attack follows most of the time a serie of stages to attack its victim. This pattern of stages is also know as the Kill Chain, first mentioned by .. []. An APT will not always follow exact each step of this chain but it will give a good guideline of how an APT works. 
\begin{enumerate}
\item \textbf{Reconnaissance}: During the first step of the Kill Chain an attacker will look for information to find an interesting victim. This information can be emailaddresses, IP addresses, conference information, anything that is available about the victim.
\item \textbf{Weaponization}: In the second stage the attacker will use an exploit and add a malicious playload to be send to the victim. 
\item \textbf{Delivery}: The attacker will deliver his malicious code to the victim through different kins of intrusion methods. This can include email, usb stick, cd's, web, applications or other means.
\item \textbf{Exploitation}:The attacker executes the exploit, which is only relevant if the attacker uses an exploit.
\item \textbf{Installation}: The malware will be installed on the asset. This is only relevant if the attacker uses malware as a part of the attack.
\item \textbf{Command and Control}: The attacker will set up a command and control channel for remote manipulation of the victim.
\item \textbf{Actions on Objectives}: With ''hands on keyboard'' access, intruders accomplish their original goal. 
\end{enumerate}
\todo{nog uitbreiden, toevoegen that attackers will stay unnoticed for as long as possible or leave unnoticed with sensitive information}

### 2015-04-23 doc/template/chap-6.tex
APT's
tekst bijwerken

@@
 In this section we are going to elaborate how we are going to model a Flipit game with multiple resources and a virus that propagates and infects the resources.
-We come up with a formula for the normal FlipIt and then reform it to a FlipIt game with a virus.
+We come up with a formula for the normal FlipIt (normal as in specific parameters and no normalising over the first interval) and then reform it to a FlipIt game with a virus.
 
@@
 If it would be with the delay the attacker would gain control after the defender flipt again.
-\todo{laten zien met een figuur}.
-In the next section we are going to describe what the setting of the game is.
+This is explained in figure \ref{fig:virusflip}.
+\todo{laten zien met een figuur} In the next section we are going to describe what the setting of the game is.
 
@@
 \centering
-\includegraphics[scale=0.3]{Images/diffVirusDelay.jpg}
+\includegraphics[scale=1]{Images/Flipvirus}
+\label{fig:virusflip}
 \end{figure}
 
-If \textit{i} is an Irrational number:
-$ i \neq \dfrac{a}{b}$ with $b \in Z, a \in N$
+\subsection{define formula}
+Their is a definition given by the writers of the paper FlipIt, but we want to add the property of a virus to the game so we are trying to find a formula that defines a game by counting the amount of time one of the players has control.
+\\
+\begin{figure}[hbtp]
+\caption{Defining unit of control}
+\centering
+\includegraphics[scale=1]{Images/FlipSpel.png}
+\end{figure}
+
+%\begin{equation}\label{first}
+%n = \delta_{1} mod \delta_{0}
+%\end{equation}
+%
+%\begin{equation}\label{first}
+%\Delta A = [( \delta_{0} - n + 1 ) * \delta_{1}] mod \delta_{1}
+%\end{equation}
+%
+%\begin{equation}\label{first}
+%\sum_{i=0}^{\delta_{1}} \lbrace [( \delta_{0} - i + 1 ) * \delta_{1}] mod \delta_{1} \rbrace
+%\end{equation}
+%\todo{formule met i nakijken}
+
+for $delta_{1} > delta_{0}$ (The defender moves faster than the attacker.) \\
+
+We will call every continuous time step that one of the players has control over the resource a unit of control. 
+Next we will divide our time line of our FlipIt game into different intervals of size $delta_{1}$.
+So every time the attacker takes control we have the start of a new interval.
+Because the defender will move faster than the attacker it will at least move 1 time during the interval.
+Because the attacker only moves at the start of the interval we can say that the defender will always end as being in control of the resource.
+This brings us to the next forumula to calculate the length of a unit of control of the attacker.
+For every real number $delta_{1}$ and $delta_{0}$ and every n $\in$ of the natural numbers:
+\begin{equation}\label{first}
+\Delta A = [( 1- n  ) \times \delta_{1}] mod \delta_{0}
+\end{equation}
+where n is the number of the unit of control that you want to calculate of the attacker.\\
+
+In this formula we multiply the number of unit of control that we want with the period of Attacker ( $\delta_{1}$).
+The 1 - n is when we count beginning from 1.
+If we start counting starting at 0 we leave the 1 and the formula becomes:
+\begin{equation}\label{first}
+\Delta A = [( - n  ) \times \delta_{1}] mod \delta_{0}
+\end{equation}
+
+We know that each interval ends with the control for the defender.
+This means that we only need to know how long the defender had control during that interval and take the rest.
+The rest will be the amount of time the attacker has control in that interval.
+We take the rest by doing the $mod \delta_{0} $
+
+This means that if we want to calculate the gain of the attacker we need to calculate the time the attacker has control over the total amount of time that has passed by.
+For $\delta_{0}$ and $\delta_{1}$ $\in$ Rational numbers we can see that we have a cycle.
+A pattern that comes back over and over again.
+That is when the amount of time is a multiple of $\delta_{0}$ and $\delta_{1}$ or the largest common multiplier (lcm).
+At this point the Attacker and the Defender move at the same time what brings us back to the beginning.
+So to calculate the gain of $\delta_{0}$ and $\delta_{1}$ $\in$ Rational numbers we need to calculate the amount of control units of the attacker that go into the length of time units equal to the lcm of  $\delta_{0}$ and $\delta_{1}$.
+After this calculation we divide it by the lcm of $\delta_{0} $ and $ \delta_{1}$, which is the total amount of time and the amount of time for one cycle.
+This gives us the following formula:
+\begin{equation}\label{first}
+a = \dfrac{\delta_{0} }{lcm(\delta_{0},\delta_{1})} 
+\end{equation}
+\begin{equation}\label{first}
+\dfrac{\sum_{i=0}^{a} \lbrace [( 1 - i ) \times \delta_{1}] mod \delta_{0} \rbrace \rbrace }{lcm(\delta_{0},\delta_{1})} 
+\end{equation}
+
+We can also define a formula without the greatest common divider. Every $\delta_{0}$ and $\delta_{1}$ have to be written in a fraction:
+\begin{equation}\label{first}
+\delta_{0}=\dfrac{a}{b} ~~~~and~~~~\delta_{1}=\dfrac{c}{d}
+\end{equation}
+If  $\delta_{0}$ or $\delta_{1}$ is a Geheel getal then b or d will be 1. The formula for the gain becomes different:\\
+
+\todo{formule zoeken zonder lcm en gcd}
+
+If $\delta_{0}$ and/or $\delta_{1}$ is an Irrational number:
+An irrational number $ i \neq \dfrac{a}{b}$ with $b \in Z, a \in N$
 Because we cannot write \textit{i} in a fraction, this means that we won't have a cycle.
@@
 If it does that means that their is repetition, meaning again that their is a cycle.
-We can conclude that if we have no cycle and no number will be repeated twice, that it will enumerate every number between 0 and the biggest interval.
+We can conclude that if we have no cycle and no number will be repeated twice, that it will enumerate every number between 0 and the biggest interval (which is $\delta_{0}$).
 \todo{interval definieren}
-Even though time is continious, irrational numbers are not considered.
-We kunnen continue tijd voldoende dicht benaderen door enkel rationele getallen in beschouwing te nemen.
-Vlakbij elk Irrationaal getal ligt immers een Rationaal getal.
+\textit{The reals are uncountable; that is: while both the set of all natural numbers and the set of all real numbers are infinite sets, there can be no one-to-one function from the real numbers to the natural numbers} [WikiPedia: real numbers] If they are uncountable that means that we cannot calculate the sum of all the numbers between 0 and the biggest interval.
+This is proved by the Cantor diagonalisation argument.
+Uncountable does not mean that we cannot order it.
+The Field of the real numbers is ordered.
 
-A $\delta$ that is irrational will have no cycle.
+What we can do is take the limit, count as many control units of time of the attacker and divide it by the greatest amount of time.
+We can see that this eventually will result to the solution given by the writers of FlipIt.
+[r/2].
+Example delta1 Pi and delta0 1.
+Grafiek voor maken.
+
+\subsection{Formula with a virus propagation}
+Now we can define how we can use the previous formula to calculate the benefit of the attacker with a virus propagation.
+As mentioned before we have a parameter \textit{d} that defines the virus propagation.
+It will take an amount of time d before the attacker gains control over all the resources.
+We know how to calculate each unit of control of the attacker.
+If it takes d time before it can take control we have to subtract d form each unit of control.
+It can be that the unit of control is less than d.
+This means that we will have a negative number of time.
+In this case this means that the defender has flipped all the resources before the attacker good gain all the control.
+So if we want to calculate the benefit we can only take unit of control that are bigger than 0.
+So the formula becomes:
+
+\begin{equation}\label{first}
+\dfrac{\sum_{i=0}^{\delta_{0}} \lbrace [( 1 - i ) \times \delta_{1}] mod \delta_{0} - d \rbrace  > 0 \rbrace }{delta_{0} \times delta_{1}} 
+\end{equation}
+
+\subsection{Random phase}
+For now we assumed that the first move of both players started at phase $t=0$.
+In the FlipIt game the first move is chosen uniformly over the interval [0,$\delta$].
+We will call this first move the phase move and denote it by $T_{1}$ for the attacker and $T_{0}$ for the defender.
+We will have to integrate these two phases into the formula.  
+%\begin{equation}      
+%\boxed{\eta \leq C(\delta(\eta) +\Lambda_M(0,\delta))}
+%\end{equation}
+%
+%\begin{equation}\label{first}
+%a=b+c
+%\end{equation}
+%
+%\begin{subequations}\label{grp}
+%\begin{align}
+%a&=b+c\label{second}\\
+%d&=e+f+g\label{third}\\
+%h&=i+j\label{fourth}
+%\end{align}
+%\end{subequations}
 

### 2015-04-24 doc/template/chap-6.tex
Virus propagatie

@@
 
-\section{tekst}
-
-In this section we are going to elaborate how we are going to model a Flipit game with multiple resources and a virus that propagates and infects the resources.
+\section{Introduction}
+This section gives a formal definition of the FlipIt game with a virus propagation.
+First we derive a formula for a FlipIt game without a virus.
+After that we introduce a modification to this formula to achieve an adapted formula for a FlipIt game with a virus propagation.
+%In this section we are going to elaborate how we are going to model a Flipit game with multiple resources and a virus that propagates and infects the resources.
 We come up with a formula for the normal FlipIt (normal as in specific parameters and no normalising over the first interval) and then reform it to a FlipIt game with a virus.
 
-\subsection{FlipIt with a virus}
-In the previous version the FlipIt game is already been explained.
-In this section we will introduce the virus propagation.
-An attacker will drop a virus on one of the resources that is available.
+\subsection{FlipIt with a virus propagation}
+In the previous section the FlipIt game is already been explained.
+In this section we will introduce a FlipIt game with a virus propagation.
+A FlipIt game with a virus propagation is a game where the attacker will drop a virus on one of the resources that is available.
 The virus will then spread itself to the neighbour resources.
-The attacker will only gain control over the whole network, in general the game, when it has infected a certain amount of resources.
-We will call this amount for now ''d''.
-If we want to measure how many time it takes for the virus to infect ''d'' resources, we have to calculate the shortest path to the ''d'' th node.
-If we know how much time it takes to control ''d'' resources, we can model a normal FlipIt game but with a delay.
-The model will not be completely a FlipIt game with a delay, because if the delay is bigger than the period of the attacker, the attacker will gain no control.
-If it would be with the delay the attacker would gain control after the defender flipt again.
+The attacker will only gain control over the whole network, in general the game, when it has infected all the resources.
+All the resources will not be infected immediately when the attacker has dropped its virus.
+It will take some time before the virus is spread and has infected everything.
+The time that it takes for the virus to infect every resource will be denoted as parameter \textit{d}.
+If we want to measure how long it takes for the virus to infect all resources, we have to calculate the shortest path to the farthest node.
+This can be measured by a method that we will explain in section [].
+With this parameter we can compose a formula to calculate the gain of a FlipIt game with a virus propagation.
+\\ \todo{uitleggen waarom we een gewone flippit kunnen nemen met 1 resource}
+
+The gain of a player is defined as the total amount of time that a player has owned the resource over the amount of time that has passed since the beginning of the game.
+If the attacker attacks with the virus it will cause a delay of length \textit{d}.
+This means that the gain of the player from the normal game has to be abstracted with a delay everytime the attacker moves.
+This game, where the attacker drops a virus, cannot be modelled completely by a FlipIt game with a delay.
+This is because if the delay is bigger than the period of the attacker, the attacker will gain no control.
+If it would be with the delay caused by a phase bigger than zero the attacker would gain control after the defender flipt again.
 This is explained in figure \ref{fig:virusflip}.
-\todo{laten zien met een figuur} In the next section we are going to describe what the setting of the game is.
+In the next section a formal definition
 
@@
 \begin{figure}[hbtp]
-\caption{Difference in a FlipIt game between delay caused by a virus and a delay for the Attacker}
+\caption{Difference in a FlipIt game between delay caused by a virus and a phase bigger than zero for the Attacker}
 \centering

### 2015-04-25 doc/template/chap-6.tex
tekst uitgebreid tot stuk phases

@@
 If the attacker attacks with the virus it will cause a delay of length \textit{d}.
-This means that the gain of the player from the normal game has to be abstracted with a delay everytime the attacker moves.
+This means that the gain of the player from the normal game has to be abstracted with a delay every time the attacker moves.
 This game, where the attacker drops a virus, cannot be modelled completely by a FlipIt game with a delay.
@@
 This is explained in figure \ref{fig:virusflip}.
-In the next section a formal definition
+In the next section a formal definition ..
+\todo{verder uitleggen nog}
 
-\subsection{chap}
-The setting of the game that we are going to play is one with multiple resources.
-When the defender flips it will always flip all the resources.
-The attacker will flip the node in the graph that can infect all the nodes in the shortest time possible.
-The attacker will gain the control over the resources when all the resources are infected.
-So ''d'' will be the shortest path to the furthest node.
-We will model ''d'' in time units.
-\todo{moet misschien niet ?
-}.
 \begin{figure}[hbtp]
@@
 \subsection{define formula}
-Their is a definition given by the writers of the paper FlipIt, but we want to add the property of a virus to the game so we are trying to find a formula that defines a game by counting the amount of time one of the players has control.
+The game FlipIt with a virus propagation is a two-player game with multiple resources.
+The multiple resources represent the nodes in a network.
+One of the players, the defender, will try to defend his network.
+The defender will do this by flipping all the nodes of the network in every move he plays.
+The attacker, the other player, will try to infect all the nodes in the network.
+The attacker will do this by flipping the node in the graph that can infect all the nodes in the shortest time possible.
+The attacker will gain the control over the network when all the resources are infected.
+So parameter \textit{d} will be the time of the shortest path from the start node to the furthest node.
 \\
+
+Their is a definition given for the gain of a player \textit{i} by the writers of the paper FlipIt, but we want to add the property of a virus propagation to the game, hence parameter \textit{d}, so we are trying to find another formula that defines a game by counting the amount of time one of the players has control.
+\\
+
+First a list of notations that will be used throughout the formal definition (see figure \ref{fig:notations} for a graphic representation of some of the notations):
+\begin{description}
+\item $\delta_{0}$: This is the period of the defender. This denotes the length of the interval between two consecutive moves of the defender. 
+\item $\delta_{1}$: This is the period of the attacker. This denotes the length of the interval between two consecutive moves of the attacker.
+\item \textit{$T_{0}$}: This denotes the phase of the attacker that was random and uniform chosen over the interval [0,$\delta_{0}$].
+\item \textit{$T_{1}$}: This denotes the phase of the defender that was chosen uniformly at random in interval [0,$\delta_{0}$].
+\item \textit{Unit of control}: Every continuous time step that one of the players has control over the resource is defined as a unit of control. 
+\item \textit{n}: n is the n'th interval of the attacker, starting from interval 1.
+\item $\Delta A$ : This is a function that denotes the length of a unit of control of the attacker in the n'th interval of the attacker.
+\item \textit{lcm(a,b)}: The lcm of a and b is the least common multiple of a and b.
+\item \textit{gcd(a,b)}: The gcd of a and b is the greatest common divider of a and b.
+\end{description}
 \begin{figure}[hbtp]
@@
 \includegraphics[scale=1]{Images/FlipSpel.png}
+\label{fig:notations}
 \end{figure}
@@
 
-for $delta_{1} > delta_{0}$ (The defender moves faster than the attacker.) \\
+We \todo{We mag gebruikt worden in een wetenschappelijke tekst zolang de focus blijft op het werk en niet op de schrijver, FlipIt schrijvers gebruiken ook veel de we vorm} start by computing the gain of the attacker in a periodic game without phases.
+After that we introduce the phases.
+Next we compute the benefit of a player i in function of the gain.
+As last we adapt the computed formula for the gain and the benefit to formulate a new formula for a FlipIt game with a virus propagation.
+\todo{sectie over phases nog uitleggen}
+\\
+\subsubsection{Computing the benefit for an attacker of a normal FlipIt game}
+\textbf{Case  1:} \\
+For $delta_{1} > delta_{0}$ (The defender moves faster than the attacker.) \\
 
-We will call every continuous time step that one of the players has control over the resource a unit of control. 
-Next we will divide our time line of our FlipIt game into different intervals of size $delta_{1}$.
-So every time the attacker takes control we have the start of a new interval.
-Because the defender will move faster than the attacker it will at least move 1 time during the interval.
-Because the attacker only moves at the start of the interval we can say that the defender will always end as being in control of the resource.
-This brings us to the next forumula to calculate the length of a unit of control of the attacker.
-For every real number $delta_{1}$ and $delta_{0}$ and every n $\in$ of the natural numbers:
+We consider a game in which both of the players start with a phase $T_{0}$ and phase $T_{1}$ equal to zero.
+Both players start their first move at $t=0$.
+As previously stated (in the formal definition of the game and the introduction of different notations used throughout the paper), the defender has control in the beginning of the game at $t=0$.
+If the two players move at the same time during the game, the moves cancel and no change of state happens.
+\\ \todo{eerder gedefinieerd dat elk spel begint op t=0}
+
+To compute the gain formula for the attacker, we need to calculate the amount of time that the attacker has control over the game from the start of the game up to time t.
+This can be done by computing all the units of control of the defender up to time t \todo{formule is niet in staat om dat te doen als functie van t} and summarize it.
+\\
+The formula that we are going to compute will not be in function of the time but in function of the intervals of the attacker.
+By doing this we always have a whole unit of control.
+If time is used the last unit of control can be shortened.
+Because the game goes on indefinitely and because it is easier to compute a formula in function of the intervals of the attacker, the gain formula will be in function of the intervals of the attacker.
+\todo{kan misschien beter uitgelegd worden}
+
+For that reason we divide our time line of our FlipIt game into different intervals of size $delta_{1}$.
+Therefore every time the attacker moves we have the start of a new interval.
+Considering that the defender will move faster than the attacker, he or she will at least move one time during the interval of the attacker.
+Because the attacker only moves at the start of his or her interval we can say that the defender will always end as being in control of the resource.
+To calculate how long the unit of control of the attacker is in every interval, we only need to know how long the defender had control during that interval and substract it from the length of the interval of the attacker.
+\todo{blabla en dan hebben we de volgende formule}
+This brings us to the next forumula to calculate the length of a unit of control in the n'th interval of the attacker. 
+For every real number $delta_{1}$ and $delta_{0}$ and every n $\in$ of the natural numbers (including 0 in the set of natural numbers) :
 \begin{equation}\label{first}
@@
 \end{equation}
-where n is the number of the unit of control that you want to calculate of the attacker.\\
+where n is the number of the n'th interval of the attacker where the length of the unit of control of the attacker is calculated.\\
 
-In this formula we multiply the number of unit of control that we want with the period of Attacker ( $\delta_{1}$).
-The 1 - n is when we count beginning from 1.
-If we start counting starting at 0 we leave the 1 and the formula becomes:
+%In this formula we multiply the number of unit of control that we want with the period of Attacker ( $\delta_{1}$). 
+The 1 - n is when we count beginning from 1. If we start counting starting at 0 we leave the 1 and the formula becomes:
 \begin{equation}\label{first}
@@
 
-We know that each interval ends with the control for the defender.
+%We know that each interval ends with the control for the defender.
 This means that we only need to know how long the defender had control during that interval and take the rest.
@@
 We take the rest by doing the $mod \delta_{0} $
+
+For phases ..
 

### 2015-04-25 doc/template/chap-6.tex
Update tekst en symbolen

@@
 \item $\Delta A$ : This is a function that denotes the length of a unit of control of the attacker in the n'th interval of the attacker.
-\item \textit{lcm(a,b)}: The lcm of a and b is the least common multiple of a and b.
-\item \textit{gcd(a,b)}: The gcd of a and b is the greatest common divider of a and b.
+\item \textit{lcm(a,b)}: The \textit{lcm} of a and b is the least common multiple of a and b.
+\item \textit{gcd(a,b)}: The \textit{gcd} of a and b is the greatest common divider of a and b.
 \end{description}
@@
 \\
-\subsubsection{Computing the benefit for an attacker of a normal FlipIt game}
+\subsubsection{Computing the gain for an attacker of a normal FlipIt game}
 \textbf{Case  1:} \\
-For $delta_{1} > delta_{0}$ (The defender moves faster than the attacker.) \\
+For $\delta_{1} > \delta_{0}$ (The defender moves faster than the attacker.) \\
 
@@
 
-For that reason we divide our time line of our FlipIt game into different intervals of size $delta_{1}$.
+For that reason we divide our time line of our FlipIt game into different intervals of size $\delta_{1}$.
 Therefore every time the attacker moves we have the start of a new interval.
@@
 This brings us to the next forumula to calculate the length of a unit of control in the n'th interval of the attacker. 
-For every real number $delta_{1}$ and $delta_{0}$ and every n $\in$ of the natural numbers (including 0 in the set of natural numbers) :
+For every real $\delta_{1}$ and $\delta_{0}$ $\in$ \(\mathbb{R}\) and every n $\in$ \(\mathbb{N}\) (including 0 in the set of natural numbers) :
 \begin{equation}\label{first}
-\Delta A = [( 1- n  ) \times \delta_{1}] mod \delta_{0}
+\Delta A = [( 1- n  ) \cdot \delta_{1}] mod \delta_{0}
 \end{equation}
@@
 \begin{equation}\label{first}
-\Delta A = [( - n  ) \times \delta_{1}] mod \delta_{0}
+\Delta A = [( - n  ) \cdot \delta_{1}] mod \delta_{0}
 \end{equation}
@@
 
-For phases ..
+For phases .. \\
 
-This means that if we want to calculate the gain of the attacker we need to calculate the time the attacker has control over the total amount of time that has passed by.
-For $\delta_{0}$ and $\delta_{1}$ $\in$ Rational numbers we can see that we have a cycle.
+\textbf{Case 2}:
+Alles omdraaien \todo{beter beschrijven}
+\subsubsection{Computing the average benefit for a normal FlipIt game}
+
+The average benefit of a player \textit{i} is denoted by $\beta_(i)(t) = \gamma_(i)(t) - k_{i} / \delta_{i}$, which is equal to the fraction of time that the resource has been owned by player \textit{i}, minus the cost rate for moving.
+For now we consider the cost rate equal to 0.
+$\gamma_{i}(t) = G_{i}(t)/t$ is the average gain rate of player \textit{i} which is defined as the fraction of time that player \textit{i} has control over the resource up to time t.
+As stated before the formula that will be computed is in function of $\delta_{1}$.
+So parameter t will be in this case a multiple of the period of the attacker, hence a multiple of $\delta_{1}$.
+In the rest of the paper the average benefit will be addressed as the benefit of the game for player \textit{i}.
+\\
+
+To be able to calculate the benefit of the attacker, we need to calculate the average gain rate of the attacker.
+We will look at this for two cases.
+One case where $\delta_{0}$ and $\delta_{1}$ $\in$ \(\mathbb{Q}\) and the other one where $\delta_{0}$ and $\delta_{1}$ $\in$ \(\mathbb{I}.\) \\
+
+\textbf{Rational numbers (\(\mathbb{Q}\)):} For $\delta_{0}$ and $\delta_{1}$ $\in$ \(\mathbb{Q}\), a repetition of the game will occur.
+As every rational number is any number that can be expressed as the fraction p/q with p and q $\in$ \(\mathbb{Z}\) (integers), with the denominator q not equal to zero, it is possible to find the \textit{lcm} of $\delta_{0}$ and $\delta_{1}$.
+The \textit{lcm} is defined for all rational numbers as: $lcm(\dfrac{a}{b},\dfrac{c}{d})= \dfrac{a \cdot c}{gcd(b,d)} with  $ [\todo{referentie}].
+When \textit{t} is equal to the \textit{lcm} of $\delta_{0}$ and $\delta_{1}$, both players will move again at the same time and this can be mapped to the beginning of the game.
+Because we stated that at the end of the interval of the attacker, the defender is in control and because if two players move at the same time the moves are cancelled, we can map this to the beginning of the game.
+Since the game goes on infinite, to calculate the average gain of the attacker, it is sufficient to calculate the average gain of the attacker only during this repetition equal to the \textit{lcm} of $\delta_{0}$ and $\delta_{1}$.
+%This means that if we want to calculate the gain of the attacker we need to calculate the time the attacker has control over the total amount of time that has passed by.
+%For $\delta_{0}$ and $\delta_{1}$ $\in$ \(\mathbb{Q}\) we can see that we have a cycle.
 A pattern that comes back over and over again.
-That is when the amount of time is a multiple of $\delta_{0}$ and $\delta_{1}$ or the largest common multiplier (lcm).
+That is when the amount of time is a multiple of $\delta_{0}$ and $\delta_{1}$ or the largest common multiplier (\textit{lcm}).
 At this point the Attacker and the Defender move at the same time what brings us back to the beginning.
-So to calculate the gain of $\delta_{0}$ and $\delta_{1}$ $\in$ Rational numbers we need to calculate the amount of control units of the attacker that go into the length of time units equal to the lcm of  $\delta_{0}$ and $\delta_{1}$.
-After this calculation we divide it by the lcm of $\delta_{0} $ and $ \delta_{1}$, which is the total amount of time and the amount of time for one cycle.
-This gives us the following formula:
+So to calculate the gain of $\delta_{0}$ and $\delta_{1}$ $\in$ \(\mathbb{Q}\) we need to calculate the amount of control units of the attacker that go into the length of time units equal to the \textit{lcm} of  $\delta_{0}$ and $\delta_{1}$.
+This will be equal to the amount of time that the period of the attacker goes into the \textit{lcm}.
+We denote this by parameter \textit{p} and define it as follow:
 \begin{equation}\label{first}
-a = \dfrac{\delta_{0} }{lcm(\delta_{0},\delta_{1})} 
+p = \dfrac{lcm(\delta_{0},\delta_{1})}{\delta_{1} } 
 \end{equation}
+After this calculation we summarize our units of control in function of \textit{p} and divide it by the \textit{lcm} of $\delta_{0} $ and $ \delta_{1}$, which is the total amount of time for one cycle.
+This gives us the following formula of the benefit of the attacker with a cost rate equal to zero:
+
 \begin{equation}\label{first}
-\dfrac{\sum_{i=0}^{a} \lbrace [( 1 - i ) \times \delta_{1}] mod \delta_{0} \rbrace \rbrace }{lcm(\delta_{0},\delta_{1})} 
+\beta_{1} = \dfrac{\sum_{i=0}^{p} \lbrace [( 1 - i ) \cdot \delta_{1}] mod \delta_{0} \rbrace }{lcm(\delta_{0},\delta_{1})} 
 \end{equation}
 
-We can also define a formula without the greatest common divider. Every $\delta_{0}$ and $\delta_{1}$ have to be written in a fraction:
-\begin{equation}\label{first}
-\delta_{0}=\dfrac{a}{b} ~~~~and~~~~\delta_{1}=\dfrac{c}{d}
-\end{equation}
-If  $\delta_{0}$ or $\delta_{1}$ is a Geheel getal then b or d will be 1. The formula for the gain becomes different:\\
+%We can also define a formula without the greatest common divider. Every $\delta_{0}$ and $\delta_{1}$ have to be written in a fraction:
+%\begin{equation}\label{first}
+%\delta_{0}=\dfrac{a}{b} ~~~~and~~~~\delta_{1}=\dfrac{c}{d}
+%\end{equation}
+%If  $\delta_{0}$ or $\delta_{1}$ is a Geheel getal then b or d will be 1. The formula for the gain becomes different:\\
+%
+%\todo{formule zoeken zonder lcm en gcd}
 
-\todo{formule zoeken zonder lcm en gcd}
-
-If $\delta_{0}$ and/or $\delta_{1}$ is an Irrational number:
-An irrational number $ i \neq \dfrac{a}{b}$ with $b \in Z, a \in N$
-Because we cannot write \textit{i} in a fraction, this means that we won't have a cycle.
-If we would have a cycle that means that we do have a number that divides \textit{i}.
-If we don't have a cycle it goes on forever.
-Meaning that it goes on to infinity.
-This also means that no number will be repeated two times.
+\textbf{Irrational numbers (\(\mathbb{I}\)):} If $\delta_{0}$ and/or $\delta_{1}$ $\in$ \(\mathbb{I}\):
+An irrational number $ i \neq \dfrac{a}{b}$ with $b \in$ \(\mathbb{Z}\) , a $\in$ \(\mathbb{N}\).
+Because we cannot write \textit{i} as a fraction, this means that we cannot compute a common multiplier of $\delta_{0}$ and $\delta_{1}$.
+If their is no common multiplier the attacker and the defender won't move at one point on the same time, meaning that this does not result in a cycle.
+If we would have a cycle that means that there exists a number \textit{x} that can be divided by $\delta_{0}$ and $\delta_{1}$.
+At $t=x$ both of the players will move at the same time, which is not possible because then their would be a cycle.
+Since their is no cycle it also means that no unit of control will be repeated two times.
+Every unit of control will have a distinct length.
 If it does that means that their is repetition, meaning again that their is a cycle.
-We can conclude that if we have no cycle and no number will be repeated twice, that it will enumerate every number between 0 and the biggest interval (which is $\delta_{0}$).
-\todo{interval definieren}
+\todo{nog verder uitleggen} We can conclude that if we have no cycle and no number will be repeated twice, that it will enumerate every number between 0 and the biggest interval (which is $\delta_{0}$).
 \textit{The reals are uncountable; that is: while both the set of all natural numbers and the set of all real numbers are infinite sets, there can be no one-to-one function from the real numbers to the natural numbers} [WikiPedia: real numbers] If they are uncountable that means that we cannot calculate the sum of all the numbers between 0 and the biggest interval.
@@
 \begin{equation}\label{first}
-\dfrac{\sum_{i=0}^{\delta_{0}} \lbrace [( 1 - i ) \times \delta_{1}] mod \delta_{0} - d \rbrace  > 0 \rbrace }{delta_{0} \times delta_{1}} 
+\dfrac{\sum_{i=0}^{\delta_{0}} \lbrace [( 1 - i ) \cdot \delta_{1}] mod \delta_{0} - d \rbrace  > 0 \rbrace }{delta_{0} \cdot delta_{1}} 
 \end{equation}

### 2015-04-26 doc/template/chap-6.tex
opmerkingen van Jeroen

@@
-\chapter{Intro to virus}
+\chapter{FlipIt with a virus propagation }
 \label{cha:6}
@@
 
+\todo{resources vs nodes}
 \section{Introduction}
@@
 
-\subsection{FlipIt with a virus propagation}
-In the previous section the FlipIt game is already been explained.
-In this section we will introduce a FlipIt game with a virus propagation.
-A FlipIt game with a virus propagation is a game where the attacker will drop a virus on one of the resources that is available.
-The virus will then spread itself to the neighbour resources.
-The attacker will only gain control over the whole network, in general the game, when it has infected all the resources.
-All the resources will not be infected immediately when the attacker has dropped its virus.
-It will take some time before the virus is spread and has infected everything.
+\subsection{FlipIt vs FlipIt with virus propagation}
+Section [] explained the FlipIt game.
+This section will expands the FlipIt game to a FlipIt game with virus propagation.
+A FlipIt game with virus propagation is a game with multiple nodes where the attacker will drop a virus on one of the nodes.
+The virus will then spread itself to the neighbour nodes trying to infect every node.
+The attacker will only gain control over the whole network -or the game in general- when it has infected all the nodes.
+
+After dropping a virus on the first resource, it takes a while for the virus to infect the entire network.
 The time that it takes for the virus to infect every resource will be denoted as parameter \textit{d}.
-If we want to measure how long it takes for the virus to infect all resources, we have to calculate the shortest path to the farthest node.
+If we want to measure how long it takes for the virus to infect all the resources, we have to calculate the shortest path from the first infected node to the farthest node.
 This can be measured by a method that we will explain in section [].
-With this parameter we can compose a formula to calculate the gain of a FlipIt game with a virus propagation.
+Using this parameter, the gain of a FlipIt game will be calculated in section [].
 \\ \todo{uitleggen waarom we een gewone flippit kunnen nemen met 1 resource}
+\todo{a virus propagation}
+The gain of a player is defined as the total amount of time that a player has owned the resource since the start of the game.
+In the context of a FlipIt game with virus propagation, the whole network is seen as one resource.
+This resource is owned by the attacked if he has control over the entire network.
+If the defender owns at least one node, he owns the resource.\\
 
-The gain of a player is defined as the total amount of time that a player has owned the resource over the amount of time that has passed since the beginning of the game.
-If the attacker attacks with the virus it will cause a delay of length \textit{d}.
-This means that the gain of the player from the normal game has to be abstracted with a delay every time the attacker moves.
+%If the attacker attacks with the virus, the propagation will cause a delay of length \textit{d} before the attacker gains control over the whole network.
+This means that the gain of the attacker from the normal game has to be substracted with a delay every time the attacker moves.
 This game, where the attacker drops a virus, cannot be modelled completely by a FlipIt game with a delay.
-This is because if the delay is bigger than the period of the attacker, the attacker will gain no control.
+It may happen that the delay \textit{d} is bigger than the initial amount of control of the attacker.
+In this case the attacker will gain no control.
 If it would be with the delay caused by a phase bigger than zero the attacker would gain control after the defender flipt again.
@@
 In the next section a formal definition ..
-\todo{verder uitleggen nog}
+\todo{verder uitleggen nog.
+Phases zijn vooraf uitgelegd.}
 
@@
 \subsection{define formula}
-The game FlipIt with a virus propagation is a two-player game with multiple resources.
+%The game FlipIt with a virus propagation is a two-player game with multiple resources.
 The multiple resources represent the nodes in a network.
@@
 
-Their is a definition given for the gain of a player \textit{i} by the writers of the paper FlipIt, but we want to add the property of a virus propagation to the game, hence parameter \textit{d}, so we are trying to find another formula that defines a game by counting the amount of time one of the players has control.
+\todo{deze sectie herschrijven}
+There is a definition given for the gain of a player \textit{i} by the writers of the paper FlipIt, but we want to add the property of a virus propagation to the game, hence parameter \textit{d}, so we are trying to find another formula that defines a game by counting the amount of time one of the players has control.
 \\
+\todo{we hebben een periodisch spel, simpelste spel.}
+\todo{ergens vermelden waarom we deze case bespreken en niet waarbij de aanvaller niet periodisch speelt}
 
+\todo{zeggen dat we eerst de functie voorstellen van de normale FlipIt}
 First a list of notations that will be used throughout the formal definition (see figure \ref{fig:notations} for a graphic representation of some of the notations):
@@
 \item $\delta_{1}$: This is the period of the attacker. This denotes the length of the interval between two consecutive moves of the attacker.
-\item \textit{$T_{0}$}: This denotes the phase of the attacker that was random and uniform chosen over the interval [0,$\delta_{0}$].
-\item \textit{$T_{1}$}: This denotes the phase of the defender that was chosen uniformly at random in interval [0,$\delta_{0}$].
-\item \textit{Unit of control}: Every continuous time step that one of the players has control over the resource is defined as a unit of control. 
+\item \textit{$T_{0}$}: This denotes the phase of the defender that was random and uniform chosen over the interval [0,$\delta_{0}$].
+\item \textit{$T_{1}$}: This denotes the phase of the attacker that was chosen uniformly at random in interval [0,$\delta_{1}$].
+\item \textit{Unit of control}: Defined as the period between gaining control and losing control over the resource.  
 \item \textit{n}: n is the n'th interval of the attacker, starting from interval 1.
@@
 
+\todo{hier staat al benefit dat nog niet is uitgelegd}
 We \todo{We mag gebruikt worden in een wetenschappelijke tekst zolang de focus blijft op het werk en niet op de schrijver, FlipIt schrijvers gebruiken ook veel de we vorm} start by computing the gain of the attacker in a periodic game without phases.
 After that we introduce the phases.
-Next we compute the benefit of a player i in function of the gain.
-As last we adapt the computed formula for the gain and the benefit to formulate a new formula for a FlipIt game with a virus propagation.
-\todo{sectie over phases nog uitleggen}
+Next we compute the benefit of the attacker in function of the gain.
+Finally we adapt our gain and benefit calculation to include virus propagation.
 \\
@@
 To compute the gain formula for the attacker, we need to calculate the amount of time that the attacker has control over the game from the start of the game up to time t.
-This can be done by computing all the units of control of the defender up to time t \todo{formule is niet in staat om dat te doen als functie van t} and summarize it.
+This can be done by computing all the units of control of the attacker up to time t \todo{formule is niet in staat om dat te doen als functie van t} and summing them.
 \\
@@
 Because the attacker only moves at the start of his or her interval we can say that the defender will always end as being in control of the resource.
+\todo{tekening met voorbeeld}.
 To calculate how long the unit of control of the attacker is in every interval, we only need to know how long the defender had control during that interval and substract it from the length of the interval of the attacker.
@@
 This brings us to the next forumula to calculate the length of a unit of control in the n'th interval of the attacker. 
+\todo{definieer dat delta groter als nul is en niet negatief}
 For every real $\delta_{1}$ and $\delta_{0}$ $\in$ \(\mathbb{R}\) and every n $\in$ \(\mathbb{N}\) (including 0 in the set of natural numbers) :
@@
 
-The average benefit of a player \textit{i} is denoted by $\beta_(i)(t) = \gamma_(i)(t) - k_{i} / \delta_{i}$, which is equal to the fraction of time that the resource has been owned by player \textit{i}, minus the cost rate for moving.
+The average benefit of a player \textit{i} is denoted by $\beta_{i}(t) = \gamma_{i}(t) - k_{i} / \delta_{i}$, which is equal to the fraction of time that the resource has been owned by player \textit{i}, minus the cost rate for moving.
 For now we consider the cost rate equal to 0.
@@
 Now we can define how we can use the previous formula to calculate the benefit of the attacker with a virus propagation.
-As mentioned before we have a parameter \textit{d} that defines the virus propagation.
-It will take an amount of time d before the attacker gains control over all the resources.
-We know how to calculate each unit of control of the attacker.
-If it takes d time before it can take control we have to subtract d form each unit of control.
-It can be that the unit of control is less than d.
-This means that we will have a negative number of time.
-In this case this means that the defender has flipped all the resources before the attacker good gain all the control.
-So if we want to calculate the benefit we can only take unit of control that are bigger than 0.
+As mentioned before parameter \textit{d} defines the virus propagation.
+It will take an amount of time \textit{d} before the attacker gains control over all the resources.
+In the previous section we defined a formula to calculate each unit of control of the attacker.
+If the virus propagation takes \textit{d} time before every resource is infected then this d has to be subtracted from each unit of control.
+(see figure \ref{fig:virusflip}).
+It may happen that the unit of control is less than \textit{d}.
+The result of the substration will be a negative number in time.
+In this case this means that the defender has flipped all the resources before the attacker could gain control over all the resources.
+To calculate the benefit only the units of control bigger than 0 have to be summarized.
 So the formula becomes:
@@
 \begin{equation}\label{first}
-\dfrac{\sum_{i=0}^{\delta_{0}} \lbrace [( 1 - i ) \cdot \delta_{1}] mod \delta_{0} - d \rbrace  > 0 \rbrace }{delta_{0} \cdot delta_{1}} 
+Gain_{1} = \dfrac{\sum_{i=0}^{\delta_{0}} \lbrace [( 1 - i ) \cdot \delta_{1}] mod \delta_{0} - d \rbrace  > 0 \rbrace }{delta_{0} \cdot delta_{1}} 
 \end{equation}
 
-\subsection{Random phase}
-For now we assumed that the first move of both players started at phase $t=0$.
+%\subsection{Random phase}
+%For now we assumed that the first move of both players started at phase $t=0$.
 In the FlipIt game the first move is chosen uniformly over the interval [0,$\delta$].
 We will call this first move the phase move and denote it by $T_{1}$ for the attacker and $T_{0}$ for the defender.
-We will have to integrate these two phases into the formula.  
+%We will have to integrate these two phases into the formula.  
 %\begin{equation}      

### 2015-04-26 doc/template/chap-8.tex
aanpassing van hoofdstuk

--- 2015-04-26 doc/template/chap-6.tex
@@
 \section{Introduction}
-This section gives a formal definition of the FlipIt game with a virus propagation.
+
+%This section gives a formal definition of the FlipIt game with a virus propagation.
 First we derive a formula for a FlipIt game without a virus.
 After that we introduce a modification to this formula to achieve an adapted formula for a FlipIt game with a virus propagation.
-%In this section we are going to elaborate how we are going to model a Flipit game with multiple resources and a virus that propagates and infects the resources.
-We come up with a formula for the normal FlipIt (normal as in specific parameters and no normalising over the first interval) and then reform it to a FlipIt game with a virus.
+In this section we are going to elaborate how we are going to model a FlipIt game with multiple resources and a virus that propagates and infects the resources.
+First we derive a formula for a FlipIt game without a virus.
+After that we introduce a modification to this formula to achieve an adapted formula for a FlipIt game with a virus propagation.
+% We come up with a formula for the normal FlipIt (normal as in specific parameters and no normalising over the first interval) and then reform it to a FlipIt game with a virus.
 
@@
 Section [] explained the FlipIt game.
-This section will expands the FlipIt game to a FlipIt game with virus propagation.
-A FlipIt game with virus propagation is a game with multiple nodes where the attacker will drop a virus on one of the nodes.
-The virus will then spread itself to the neighbour nodes trying to infect every node.
-The attacker will only gain control over the whole network -or the game in general- when it has infected all the nodes.
+This section will adapt a FlipIt game to a FlipIt game with virus propagation.
+We consider the non-adaptive continuous FlipIt game where both players play a periodic strategy with a random phase.
+De adaptation will be a game where the single resource consist of a computer network with multiple nodes.
+One of the players, the defender, will try to defend his network.
+The defender will do this by flipping all the nodes of the network in every move he plays.
+The attacker, the other player, will try to infect all the nodes in the network.
+The attacker will do this by flipping the node in the graph that can infect all the nodes in the shortest time possible.
+The attacker will only gain the control over the resource when he infects all the nodes of the network.
+The defender in contrast will gain control over the whole network when it has control over at least one node in the network.
+\\
+\todo{ergens vermelden waarom we deze case bespreken en niet waarbij de aanvaller niet periodisch speelt}
 
 After dropping a virus on the first resource, it takes a while for the virus to infect the entire network.
-The time that it takes for the virus to infect every resource will be denoted as parameter \textit{d}.
-If we want to measure how long it takes for the virus to infect all the resources, we have to calculate the shortest path from the first infected node to the farthest node.
+The time that it takes for the virus to infect every node will be denoted as parameter \textit{d}.
+If we want to measure how long it takes for the virus to infect all the nodes in the network, we have to calculate the shortest path from the first infected node to the farthest node.
 This can be measured by a method that we will explain in section [].
-Using this parameter, the gain of a FlipIt game will be calculated in section [].
-\\ \todo{uitleggen waarom we een gewone flippit kunnen nemen met 1 resource}
-\todo{a virus propagation}
-The gain of a player is defined as the total amount of time that a player has owned the resource since the start of the game.
-In the context of a FlipIt game with virus propagation, the whole network is seen as one resource.
-This resource is owned by the attacked if he has control over the entire network.
-If the defender owns at least one node, he owns the resource.\\
+Using this parameter, a FlipIt game with virus propagation can be modelled.
+\\
 
@@
 Phases zijn vooraf uitgelegd.}
-
-\begin{figure}[hbtp]
-\caption{Difference in a FlipIt game between delay caused by a virus and a phase bigger than zero for the Attacker}
-\centering
-\includegraphics[scale=1]{Images/Flipvirus}
-\label{fig:virusflip}
-\end{figure}
 
@@
 \todo{deze sectie herschrijven}
-There is a definition given for the gain of a player \textit{i} by the writers of the paper FlipIt, but we want to add the property of a virus propagation to the game, hence parameter \textit{d}, so we are trying to find another formula that defines a game by counting the amount of time one of the players has control.
+There is a definition given for the gain of a player \textit{i} by the writers of the paper FlipIt [], but we want to add the property of a virus propagation to the game, hence parameter \textit{d}, so we are trying to find another formula that defines a game by counting the amount of time one of the players has control.
 \\
 \todo{we hebben een periodisch spel, simpelste spel.}
-\todo{ergens vermelden waarom we deze case bespreken en niet waarbij de aanvaller niet periodisch speelt}
 
@@
 \begin{description}
-\item $\delta_{0}$: This is the period of the defender. This denotes the length of the interval between two consecutive moves of the defender. 
+\item $\delta_{0}$: Positive and non zero number. This is the period of the defender. This denotes the length of the interval between two consecutive moves of the defender. 
 \item $\delta_{1}$: This is the period of the attacker. This denotes the length of the interval between two consecutive moves of the attacker.
@@
 \item $\Delta A$ : This is a function that denotes the length of a unit of control of the attacker in the n'th interval of the attacker.
+\item \textit{d}: Amount of time that the virus needs to infect every node in the network.
 \item \textit{lcm(a,b)}: The \textit{lcm} of a and b is the least common multiple of a and b.
 \item \textit{gcd(a,b)}: The \textit{gcd} of a and b is the greatest common divider of a and b.
+\item \textit{Gain}:  The gain of a player is defined as the total amount of time that a player has owned the resource since the start of the game.
+In the context of a FlipIt game with virus propagation, the whole network is seen as one resource.
+This resource is owned by the attacked if he has control over the entire network.
+If the defender owns at least one node, he owns the resource.\\
+\item \textit{Average Gain}:  $\gamma_{i}(t) = G_{i}(t)/t$ is the average gain rate of player \textit{i} which is defined as the fraction of time that player \textit{i} has control over the resource up to time t.
+\item \textit{Benefit}: The average benefit of a player \textit{i} is denoted by $\beta_{i}(t) = \gamma_{i}(t) - k_{i} / \delta_{i}$, which is equal to the fraction of time that the resource has been owned by player \textit{i}, minus the cost rate for moving.
+For now we consider the cost rate equal to 0.
+In the rest of the paper the average benefit will be addressed as the benefit of the game for player \textit{i}.
 \end{description}
@@
 
-\todo{hier staat al benefit dat nog niet is uitgelegd}
-We \todo{We mag gebruikt worden in een wetenschappelijke tekst zolang de focus blijft op het werk en niet op de schrijver, FlipIt schrijvers gebruiken ook veel de we vorm} start by computing the gain of the attacker in a periodic game without phases.
+\todo{We mag gebruikt worden in een wetenschappelijke tekst zolang de focus blijft op het werk en niet op de schrijver, FlipIt schrijvers gebruiken ook veel de we vorm} 
+We start by computing the gain of the attacker in a periodic game without phases.
+Two cases are considered: case 1 where the defender moves as least as fast as the attacker and case 2 where the attacker plays as least as fast as the defender.
 After that we introduce the phases.
-Next we compute the benefit of the attacker in function of the gain.
-Finally we adapt our gain and benefit calculation to include virus propagation.
+Next we adapt our gain calculation to include virus propagation.
+Finally we compute the benefit of the attacker in function of the gain.
 \\
@@
 \todo{tekening met voorbeeld}.
-To calculate how long the unit of control of the attacker is in every interval, we only need to know how long the defender had control during that interval and substract it from the length of the interval of the attacker.
+To calculate how long the unit of control of the attacker is in the n'th interval, we only need to know how long the attacker has control over the resource before the defender moves in that interval.
+First we need to calculate the start time of the n'th interval which will be a multiple of the period of the attacker.
+Once we know this we need to know what the difference is between this amount of time and the amount of time when the defender will flip.
+To do this we take the modulo of the negative of the amount of time of the attacker with the period of the defender.
+By doing so we calculate how much it takes to add to this time to compleet to a time that is a multiplier of the period of the defender.
+See figure \ref{fig:modulo} to see a graphic representation of this modulo.
 \todo{blabla en dan hebben we de volgende formule}
+\begin{figure}[hbtp]
+\caption{Taking the modulo of a negative number}
+\centering
+\includegraphics[scale=0.6]{Images/Modulo.png}
+\label{fig:modulo}
+\end{figure}
+
+\todo{modulo uitleggen en dan verder door tot deze formule}
 This brings us to the next forumula to calculate the length of a unit of control in the n'th interval of the attacker. 
 \todo{definieer dat delta groter als nul is en niet negatief}
-For every real $\delta_{1}$ and $\delta_{0}$ $\in$ \(\mathbb{R}\) and every n $\in$ \(\mathbb{N}\) (including 0 in the set of natural numbers) :
+For every positive and non zero real $\delta_{1}$ and $\delta_{0}$ $\in$ \(\mathbb{R}\) and every n $\in$ \(\mathbb{N}\) (including 0 in the set of natural numbers) :
 \begin{equation}\label{first}
@@
 Alles omdraaien \todo{beter beschrijven}
-\subsubsection{Computing the average benefit for a normal FlipIt game}
 
-The average benefit of a player \textit{i} is denoted by $\beta_{i}(t) = \gamma_{i}(t) - k_{i} / \delta_{i}$, which is equal to the fraction of time that the resource has been owned by player \textit{i}, minus the cost rate for moving.
-For now we consider the cost rate equal to 0.
-$\gamma_{i}(t) = G_{i}(t)/t$ is the average gain rate of player \textit{i} which is defined as the fraction of time that player \textit{i} has control over the resource up to time t.
-As stated before the formula that will be computed is in function of $\delta_{1}$.
-So parameter t will be in this case a multiple of the period of the attacker, hence a multiple of $\delta_{1}$.
-In the rest of the paper the average benefit will be addressed as the benefit of the game for player \textit{i}.
-\\
+\subsection{Formula with a virus propagation}
+Now we can define how we can use the previous formula to calculate the gain of the attacker in a FlipIt game with virus propagation.
+As mentioned before parameter \textit{d} defines the virus propagation.
+It will take an amount of time \textit{d} before the attacker gains control over all the resources.
+In the previous section we defined a formula to calculate each unit of control of the attacker.
+If the virus propagation takes \textit{d} time before every resource is infected then this d has to be subtracted from each unit of control.
+(see figure \ref{fig:virusflip}).
+It may happen that the unit of control is less than \textit{d}.
+The result of the substration will be a negative number in time.
+In this case this means that the defender has flipped all the resources before the attacker could gain control over all the resources.
+To calculate the gain only the units of control bigger than 0 have to be summarized.
+So the formula becomes:
+
+\begin{equation}\label{first}
+Gain_{1} = \sum_{i=0}^{\delta_{0}} \lbrace [( 1 - i ) \cdot \delta_{1}] mod \delta_{0} - d \rbrace  > 0 \rbrace 
+\end{equation}
+
+\begin{figure}[hbtp]
+\caption{Difference in a FlipIt game between delay caused by a virus and a phase bigger than zero for the Attacker}
+\centering
+\includegraphics[scale=1]{Images/Flipvirus}
+\label{fig:virusflip}
+\end{figure}
+
+\subsubsection{Computing the benefit of a FLipIt game with virus propagation}
 
@@
 
-\subsection{Formula with a virus propagation}
-Now we can define how we can use the previous formula to calculate the benefit of the attacker with a virus propagation.
-As mentioned before parameter \textit{d} defines the virus propagation.
-It will take an amount of time \textit{d} before the attacker gains control over all the resources.
-In the previous section we defined a formula to calculate each unit of control of the attacker.
-If the virus propagation takes \textit{d} time before every resource is infected then this d has to be subtracted from each unit of control.
-(see figure \ref{fig:virusflip}).
-It may happen that the unit of control is less than \textit{d}.
-The result of the substration will be a negative number in time.
-In this case this means that the defender has flipped all the resources before the attacker could gain control over all the resources.
-To calculate the benefit only the units of control bigger than 0 have to be summarized.
-So the formula becomes:
-
-\begin{equation}\label{first}
-Gain_{1} = \dfrac{\sum_{i=0}^{\delta_{0}} \lbrace [( 1 - i ) \cdot \delta_{1}] mod \delta_{0} - d \rbrace  > 0 \rbrace }{delta_{0} \cdot delta_{1}} 
-\end{equation}
-
 %\subsection{Random phase}

### 2015-04-27 doc/template/chap-8.tex
Eddit voor de versie met de verbeteringen
vqnqlles

@@
-\chapter{FlipIt with a virus propagation }
+\chapter{FlipIt with virus propagation }
 \label{cha:6}
@@
 
-\todo{resources vs nodes}
-\section{Introduction}
+%\todo{resources vs nodes}
+\section{FlipIt vs FlipIt with virus propagation}
 
-%This section gives a formal definition of the FlipIt game with a virus propagation.
-First we derive a formula for a FlipIt game without a virus.
-After that we introduce a modification to this formula to achieve an adapted formula for a FlipIt game with a virus propagation.
-In this section we are going to elaborate how we are going to model a FlipIt game with multiple resources and a virus that propagates and infects the resources.
-First we derive a formula for a FlipIt game without a virus.
-After that we introduce a modification to this formula to achieve an adapted formula for a FlipIt game with a virus propagation.
-% We come up with a formula for the normal FlipIt (normal as in specific parameters and no normalising over the first interval) and then reform it to a FlipIt game with a virus.
+This chapter explains how to model a FlipIt game with a virus that propagates and infects the nodes in a network. First, this section explains the difference between FlipIt with and without a virus. 
+Then, section \ref{ch:8:GainFlip} derives the formula to calculate the gain for a FlipIt game without a virus.
+After that section \ref{ch:8:GainVirus} introduces a modification to this formula to achieve an adapted gain formula for a FlipIt game with virus propagation.
+This allows to derive the benefit formula.\\
 
-\subsection{FlipIt vs FlipIt with virus propagation}
-Section [] explained the FlipIt game.
-This section will adapt a FlipIt game to a FlipIt game with virus propagation.
-We consider the non-adaptive continuous FlipIt game where both players play a periodic strategy with a random phase.
-De adaptation will be a game where the single resource consist of a computer network with multiple nodes.
-One of the players, the defender, will try to defend his network.
-The defender will do this by flipping all the nodes of the network in every move he plays.
-The attacker, the other player, will try to infect all the nodes in the network.
-The attacker will do this by flipping the node in the graph that can infect all the nodes in the shortest time possible.
-The attacker will only gain the control over the resource when he infects all the nodes of the network.
-The defender in contrast will gain control over the whole network when it has control over at least one node in the network.
-\\
-\todo{ergens vermelden waarom we deze case bespreken en niet waarbij de aanvaller niet periodisch speelt}
+\todo{check reff}
+In chapter 2 the FlipIt game was explained.
+This chapter starts from the specific case of a non-adaptive continuous FlipIt game where both players play a periodic strategy with a random phase.
+This choice is motivated by the assumption that in the practical situation of most organisations, the defence strategy is to periodically defend the network.
+This corresponds to a periodic defender strategy.
+To simplify the analysis in a first time, a periodic attacker strategy is assumed as well.
+Further research can investigate the effect of relaxing this assumption.\\
 
+A FlipIt game consists of a single resource. To represent the security problem, the game now defines its single resource as a computer network with multiple
+nodes. One of the players, the defender, will try to defend his network. The defender
+will do this by flipping all the nodes of the network (i.e. the entire resource) in every move he plays. The
+attacker, the other player, will try to infect all the nodes in the network. The attacker
+will do this by flipping the node in the graph that can infect all the nodes in the
+shortest time possible.
 After dropping a virus on the first resource, it takes a while for the virus to infect the entire network.
-The time that it takes for the virus to infect every node will be denoted as parameter \textit{d}.
-If we want to measure how long it takes for the virus to infect all the nodes in the network, we have to calculate the shortest path from the first infected node to the farthest node.
-This can be measured by a method that we will explain in section [].
-Using this parameter, a FlipIt game with virus propagation can be modelled.
+However, since the original FlipIt game works with a single resource that is always flipped entirely, the assumption is made that the attacker is considered to have gained the control over the resource only when all the nodes of the network have been infected, i.e.
+the entire resource has been flipped.
+
+\section{Gain formula for a FlipIt game without virus propagation}
+\label{ch:8:GainFlip}
+In their FlipIt paper [], Marten van Dijk and Ari Juels and Alina Oprea and Ronald L.
+Rivest, give a definition of the gain of a player \textit{i} .
+This definition is however not easy to adapt to the situation with virus propagation.
+Therefore, this section presents an alternative formula in that defines a game by quantifying the amount of time each player has control.
 \\
 
-%If the attacker attacks with the virus, the propagation will cause a delay of length \textit{d} before the attacker gains control over the whole network.
-This means that the gain of the attacker from the normal game has to be substracted with a delay every time the attacker moves.
-This game, where the attacker drops a virus, cannot be modelled completely by a FlipIt game with a delay.
-It may happen that the delay \textit{d} is bigger than the initial amount of control of the attacker.
-In this case the attacker will gain no control.
-If it would be with the delay caused by a phase bigger than zero the attacker would gain control after the defender flipt again.
-This is explained in figure \ref{fig:virusflip}.
-In the next section a formal definition ..
-\todo{verder uitleggen nog.
-Phases zijn vooraf uitgelegd.}
+The following notations will be used throughout the formal definition (see figure \ref{fig:notations} for a graphic representation of some of the notations):
 
-\subsection{define formula}
-%The game FlipIt with a virus propagation is a two-player game with multiple resources.
-The multiple resources represent the nodes in a network.
-One of the players, the defender, will try to defend his network.
-The defender will do this by flipping all the nodes of the network in every move he plays.
-The attacker, the other player, will try to infect all the nodes in the network.
-The attacker will do this by flipping the node in the graph that can infect all the nodes in the shortest time possible.
-The attacker will gain the control over the network when all the resources are infected.
-So parameter \textit{d} will be the time of the shortest path from the start node to the furthest node.
-\\
-
-\todo{deze sectie herschrijven}
-There is a definition given for the gain of a player \textit{i} by the writers of the paper FlipIt [], but we want to add the property of a virus propagation to the game, hence parameter \textit{d}, so we are trying to find another formula that defines a game by counting the amount of time one of the players has control.
-\\
-\todo{we hebben een periodisch spel, simpelste spel.}
-
-\todo{zeggen dat we eerst de functie voorstellen van de normale FlipIt}
-First a list of notations that will be used throughout the formal definition (see figure \ref{fig:notations} for a graphic representation of some of the notations):
 \begin{description}
-\item $\delta_{0}$: Positive and non zero number. This is the period of the defender. This denotes the length of the interval between two consecutive moves of the defender. 
-\item $\delta_{1}$: This is the period of the attacker. This denotes the length of the interval between two consecutive moves of the attacker.
-\item \textit{$T_{0}$}: This denotes the phase of the defender that was random and uniform chosen over the interval [0,$\delta_{0}$].
-\item \textit{$T_{1}$}: This denotes the phase of the attacker that was chosen uniformly at random in interval [0,$\delta_{1}$].
-\item \textit{Unit of control}: Defined as the period between gaining control and losing control over the resource.  
-\item \textit{n}: n is the n'th interval of the attacker, starting from interval 1.
-\item $\Delta A$ : This is a function that denotes the length of a unit of control of the attacker in the n'th interval of the attacker.
-\item \textit{d}: Amount of time that the virus needs to infect every node in the network.
-\item \textit{lcm(a,b)}: The \textit{lcm} of a and b is the least common multiple of a and b.
-\item \textit{gcd(a,b)}: The \textit{gcd} of a and b is the greatest common divider of a and b.
-\item \textit{Gain}:  The gain of a player is defined as the total amount of time that a player has owned the resource since the start of the game.
+\item $\delta_{D}$: This is the period of the defender. This denotes the length of the interval between two consecutive moves of the defender. ($\delta_{D} > 0$)
+\item $\delta_{A}$: This is the period of the attacker. This denotes the length of the interval between two consecutive moves of the attacker. ($\delta_{A} > 0$)
+\item \textit{$T_{D}$}: This denotes the phase of the defender that was chosen randomly and uniformly over the interval [0,$\delta_{D}$].
+\item \textit{$T_{A}$}: This denotes the phase of the attacker that was chosen randomly and uniformly over the interval [0,$\delta_{A}$].
+\item \textit{Unit of control}: Defined as the period between gaining (full) control and losing control over the resource.  
+\item $n_{D}$: n is the n'th interval of the defender if $\delta_{D} \geq \delta_{A}$ , starting from interval 0.
+\item $n_{A}$: n is the n'th interval of the attacker if $\delta_{A} \geq \delta_{D}$ , starting from interval 0.
+\item $\Delta Unit_{D}(n)$ : This is a function that denotes the length of a unit of control of the defender in the n'th interval of the defender, if $\delta_{D} \geq \delta_{A}$.
+\item $\Delta Unit_{A}(n)$ : This is a function that denotes the length of a unit of control of the attacker in the n'th interval of the attacker, if $\delta_{A} \geq \delta_{D}$.
+\item \textit{lcm(a,b)}: The least common multiple of a and b.
+\item \textit{gcd(a,b)}: The greatest common divider of a and b.
+\item $G_{i}(t)$: Gain of player \textit{i} at time t.
+The gain of a player is defined as the total amount of time that a player has owned the resource since the start of the game up to time t.
 In the context of a FlipIt game with virus propagation, the whole network is seen as one resource.
-This resource is owned by the attacked if he has control over the entire network.
-If the defender owns at least one node, he owns the resource.\\
+This resource is owned by the attacker if he has control over the entire network.\\
 \item \textit{Average Gain}:  $\gamma_{i}(t) = G_{i}(t)/t$ is the average gain rate of player \textit{i} which is defined as the fraction of time that player \textit{i} has control over the resource up to time t.
@@
 For now we consider the cost rate equal to 0.
-In the rest of the paper the average benefit will be addressed as the benefit of the game for player \textit{i}.
+In the rest of the paper the 'benefit' of the game for player \textit{i} will be used as shorthand for 'average benefit'.
+
+%%------------------------------------------------------------%%
+%% Figuur
+%%------------------------------------------------------------%%
 \end{description}
@@
 
-%\begin{equation}\label{first}
-%n = \delta_{1} mod \delta_{0}
-%\end{equation}
-%
-%\begin{equation}\label{first}
-%\Delta A = [( \delta_{0} - n + 1 ) * \delta_{1}] mod \delta_{1}
-%\end{equation}
-%
-%\begin{equation}\label{first}
-%\sum_{i=0}^{\delta_{1}} \lbrace [( \delta_{0} - i + 1 ) * \delta_{1}] mod \delta_{1} \rbrace
-%\end{equation}
-%\todo{formule met i nakijken}
+%To compute the gain of the attacker 
+To compute the gain of a player in a periodic game without phases, two cases are considered: case 1 where the defender moves as least as fast as the attacker and case 2 where the attacker plays as least as fast as the defender.
+Next, the formula is enriched by introducing the phases.
 
-\todo{We mag gebruikt worden in een wetenschappelijke tekst zolang de focus blijft op het werk en niet op de schrijver, FlipIt schrijvers gebruiken ook veel de we vorm} 
-We start by computing the gain of the attacker in a periodic game without phases.
-Two cases are considered: case 1 where the defender moves as least as fast as the attacker and case 2 where the attacker plays as least as fast as the defender.
-After that we introduce the phases.
-Next we adapt our gain calculation to include virus propagation.
-Finally we compute the benefit of the attacker in function of the gain.
-\\
 \subsubsection{Computing the gain for an attacker of a normal FlipIt game}
-\textbf{Case  1:} \\
-For $\delta_{1} > \delta_{0}$ (The defender moves faster than the attacker.) \\
-
-We consider a game in which both of the players start with a phase $T_{0}$ and phase $T_{1}$ equal to zero.
+Consider a game without phases, so in which both players start with a phase $T_{D}$ and phase $T_{A}$ equal to zero.
 Both players start their first move at $t=0$.
 As previously stated (in the formal definition of the game and the introduction of different notations used throughout the paper), the defender has control in the beginning of the game at $t=0$.
-If the two players move at the same time during the game, the moves cancel and no change of state happens.
-\\ \todo{eerder gedefinieerd dat elk spel begint op t=0}
+For the remainder of the game, if the two players move at the same time during the game, the moves cancel each other out and no change of state happens.
+\\
 
-To compute the gain formula for the attacker, we need to calculate the amount of time that the attacker has control over the game from the start of the game up to time t.
-This can be done by computing all the units of control of the attacker up to time t \todo{formule is niet in staat om dat te doen als functie van t} and summing them.
-\\
-The formula that we are going to compute will not be in function of the time but in function of the intervals of the attacker.
-By doing this we always have a whole unit of control.
-If time is used the last unit of control can be shortened.
-Because the game goes on indefinitely and because it is easier to compute a formula in function of the intervals of the attacker, the gain formula will be in function of the intervals of the attacker.
-\todo{kan misschien beter uitgelegd worden}
+\textbf{Case  1:} \\
+$\delta_{A} \geq \delta_{D}$ (The defender moves at least as fast as the attacker.) \\
 
-For that reason we divide our time line of our FlipIt game into different intervals of size $\delta_{1}$.
-Therefore every time the attacker moves we have the start of a new interval.
-Considering that the defender will move faster than the attacker, he or she will at least move one time during the interval of the attacker.
+To compute the gain formula for the attacker,the amount of time that the attacker has control over the game from the start of the game up to time t has to be calculated.
+This can be done by computing the sum of all the units of control of the attacker up to time t.
+
+To calculate a single unit of control of the attacker, the time line of the FlipIt game is divided into intervals of size $\delta_{A}$.
+Every time the attacker moves we have the start of a new interval, with the attacker being in control, unless there is a simultaneous move with the defender.
+Considering that the defender moves at least as fast as the attacker, he or she will at least move one time during the interval of the attacker.
 Because the attacker only moves at the start of his or her interval we can say that the defender will always end as being in control of the resource.
-\todo{tekening met voorbeeld}.
+\todo{tekening met voorbeeld}. \\
+
 To calculate how long the unit of control of the attacker is in the n'th interval, we only need to know how long the attacker has control over the resource before the defender moves in that interval.
-First we need to calculate the start time of the n'th interval which will be a multiple of the period of the attacker.
-Once we know this we need to know what the difference is between this amount of time and the amount of time when the defender will flip.
-To do this we take the modulo of the negative of the amount of time of the attacker with the period of the defender.
-By doing so we calculate how much it takes to add to this time to compleet to a time that is a multiplier of the period of the defender.
-See figure \ref{fig:modulo} to see a graphic representation of this modulo.
-\todo{blabla en dan hebben we de volgende formule}
+The start time of the n'th interval will be a multiple of the period of the attacker.
+Once the attacker has played, the time he can stay in control until the next move of the defender, depends on the time elapsed since the last time the defender played in the previous interval ( the (n-1)th interval).
+
+The time the defender plays in an interval is a multiple of its period.
+Once the defender has played for the last time in the (n-1)'th interval, the remaining time until the attacker will play can be calculated as the remainder of $n\cdot \delta_{A}$ divided by $\delta_{D}$ or $n \cdot \delta_{a}~modulo~ \delta_{D}$.
+The time the attacker will stay in control is then $\delta_{D}- n \cdot \delta_{A}~ modulo ~ \delta_{D}$, which can also be calculated as $( -n \cdot \delta_{A}) ~modulo~ \delta_{D}$.
+
+Figure \ref{fig:modulo} illustrates this graphically, for $\delta_{A}= 4$, $\delta_{D}=3$ and the $n=2$ interval.
+We see that in interval 1, the defender will stay $8 ~modulo~ 3 = 2$ in control, and so, in interval 2, the attacker will stay $1 = 3 - 2 = -(2*4) ~modulo~ 3$ in control.
+
+%%------------------------------------------------------------%%
+%% Figuur
+%%------------------------------------------------------------%%
+
 \begin{figure}[hbtp]
@@
 \centering
-\includegraphics[scale=0.6]{Images/Modulo.png}
+\includegraphics[scale=0.4]{Images/modulotek}
+\includegraphics[scale=0.4]{Images/Modulosimpel}
 \label{fig:modulo}
@@
 
-\todo{modulo uitleggen en dan verder door tot deze formule}
-This brings us to the next forumula to calculate the length of a unit of control in the n'th interval of the attacker. 
-\todo{definieer dat delta groter als nul is en niet negatief}
-For every positive and non zero real $\delta_{1}$ and $\delta_{0}$ $\in$ \(\mathbb{R}\) and every n $\in$ \(\mathbb{N}\) (including 0 in the set of natural numbers) :
+This brings us to the next formula to calculate the length of a unit of control in the n'th interval of the attacker.
+
+For every positive and non zero real $\delta_{A}$ and $\delta_{D}$ $\in$ \(\mathbb{R}\) and every n $\in$ \(\mathbb{N}\) (including 0 in the set of natural numbers) :
 \begin{equation}\label{first}
-\Delta A = [( 1- n  ) \cdot \delta_{1}] mod \delta_{0}
+\Delta Unit_{A}(n_{A}) =  [( - n_{A}  ) \cdot \delta_{A}] mod \delta_{D}
 \end{equation}
-where n is the number of the n'th interval of the attacker where the length of the unit of control of the attacker is calculated.\\
+where $n_{A}$ is the number of the n'th interval of the attacker starting form interval 0 where the length of the unit of control of the attacker is calculated.\\
 
-%In this formula we multiply the number of unit of control that we want with the period of Attacker ( $\delta_{1}$). 
-The 1 - n is when we count beginning from 1. If we start counting starting at 0 we leave the 1 and the formula becomes:
+The length of a unit of control of the defender is the remainder of the interval after the attacker loses control over the resource when the defender plays.
+This can be defines as follows for the n'th interval of the attacker:
 \begin{equation}\label{first}
-\Delta A = [( - n  ) \cdot \delta_{1}] mod \delta_{0}
+\Delta Unit_{D}(n_{A}) = \delta_{A} - [( - n_{A}  ) \cdot \delta_{A}] mod \delta_{D}
 \end{equation}
+
+An example: Figure \ref{fig:modulo} shows a FlipIt game were the period of the attacker is Pi and the period of the defender is 1. ..\todo{aanvullen}.
+On figure \ref{fig:unitofcontrolformula}
+
+%%------------------------------------------------------------%%
+%% Figuur
+%%------------------------------------------------------------%%
+\begin{figure}[hbtp]
+\caption{Example for calculating the control unit in interval 3 for a FlipIt game with period defender = 1 and period attacker = Pi}
+\centering
+\includegraphics[scale=0.6]{Images/VoorbeeldUnit.png}
+\label{fig:unitofcontrolformula}
+\end{figure}
+
+The gain formula can be calculated by taking all the units of control of the player up to an amount of p intervals of the attacker.
+
+\begin{equation}\label{first}
+Gain_{A} = \sum_{i=0}^{p} \lbrace [( - i ) \cdot \delta_{A}] mod \delta_{D}  \rbrace \rbrace 
+\end{equation}
+
+where \textit{p} is the number of units of control that have to be summed.
+
+The gain of the defender is the sum of the units of control of the defender up to the same amount of p intervals of the attacker:
+\begin{equation}\label{first}
+Gain_{D} = \sum_{i=0}^{p} \lbrace \delta_{A} - [( - i ) \cdot \delta_{A}] mod \delta_{D} \rbrace  \rbrace  
+\end{equation}
+\begin{equation}\label{first}
+Gain_{D} = \sum_{i=0}^{p} \lbrace \delta_{A} \cdot i \rbrace - \sum_{i=0}^{p} \lbrace [( - i ) \cdot \delta_{A}] mod \delta_{D} \rbrace \rbrace  \\
+\end{equation}
+\begin{equation}\label{first}
+Gain_{D} =\delta_{A} \cdot p - \sum_{i=0}^{p} \lbrace [( - i ) \cdot \delta_{A}] mod \delta_{D} \rbrace  \rbrace  \\
+\end{equation}
+
+Note: The gain formula is not in function of time t but the amount of p intervals of the attacker.
+This approach is chosen because it will result in whole units of control.
+It is possible to make a gain formula using the time, but this will result in a much more complicated function.
 
@@
 
-\textbf{Case 2}:
-Alles omdraaien \todo{beter beschrijven}
+\textbf{Case  2:} \\
+$\delta_{D} \geq \delta_{A}$ (The attacker moves at least as fast as the defender.) \\
 
-\subsection{Formula with a virus propagation}
-Now we can define how we can use the previous formula to calculate the gain of the attacker in a FlipIt game with virus propagation.
-As mentioned before parameter \textit{d} defines the virus propagation.
-It will take an amount of time \textit{d} before the attacker gains control over all the resources.
-In the previous section we defined a formula to calculate each unit of control of the attacker.
-If the virus propagation takes \textit{d} time before every resource is infected then this d has to be subtracted from each unit of control.
-(see figure \ref{fig:virusflip}).
+For this case we use the same approach as in case 1 but with a small difference.
+To compute the unit of control of both players we divide the time line of the FlipIt game into intervals of size $\delta_{D}$.
+The defender moves at the start of each interval, the end of the interval is the beginning of the next interval.
+Considering that the attacker will move at least as fast as the defender, he or she will move at least one time during the interval of the defender.
+Because the defender only moves in the beginning of each interval, the attacker will end as being in control of the resource.
+
+If the unit of control of the defender need to be calculated, we only need to know how long it takes for the attacker to move in the interval.
+This can be done in the same way as in case 1 by taking the modulo of the negative of the beginning of the interval.
+The big difference with case 1 is when the length of the unit of control in the 0'th interval is calculated.
+Because the defender always has control in the beginning of the game, the first interval is computed in a different way.
+The unit of control of the defender in the 0'th interval is equal to the length of the period of the attacker, since from that moment the attacker takes control.
+
+For every positive and non zero real $\delta_{A}$ and $\delta_{D}$ $\in$ \(\mathbb{R}\) and every n $\in$ \(\mathbb{N}\) (including 0 in the set of natural numbers) :\\
+%\begin{subequations}\label{grp}
+%\begin{align}
+%a&=b+c\label{second}\\
+%d&=e+f+g\label{third}\\
+%h&=i+j\label{fourth}
+%\end{align}
+%\end{subequations}
+
+\textit{for $n_{A} = 0$}
+\begin{equation}\label{first}
+\Delta Unit_{D}(n_{D}) = \delta_{A} \\
+\end{equation}
+\\ 
+\textit{for $n_{A} > 0$}
+\begin{equation}\label{first}
+\Delta Unit_{D}(n_{D}) =  [( - n_{D}  ) \cdot \delta_{D}] mod \delta_{A}\\
+\end{equation}
+where $n_{D}$ is the number of the n'th interval of the defender starting form interval 0 where the length of the unit of control of the defender is calculated.\\
+
+The length of a unit of control of the attacker is the remainder of the interval after the defender loses control over the resource when the attacker plays.
+This can be defines as follows for the n'th interval of the defender:
+\begin{equation}\label{first}
+\Delta Unit_{A}(n_{D}) = \delta_{D} - \Delta Unit_{D}(n_{D})
+\end{equation}
+
+The gain formula for $\delta_{D} \geq \delta_{A}$ is calculated in the same manner as case 1:
+
+The gain of the defender is the sum of the units of control of the defender up to the same amount of p intervals of the attacker:
+\begin{equation}\label{first}
+Gain_{D} =\sum_{i=1}^{p} \lbrace [( - i ) \cdot \delta_{D}] mod \delta_{A} \rbrace  \rbrace ~~with~~ i=0 ~~ Gain_{D}=\delta_{A}\\ 
+\end{equation}
+
+and for the attacker:
+\begin{equation}\label{first}
+Gain_{A} =\delta_{D} \cdot p - \sum_{i=1}^{p} \lbrace [( - i ) \cdot \delta_{D}] mod \delta_{A} \rbrace  \rbrace ~~with~~ i=0 ~~ Gain_{A}=\delta_{D} - \delta_{A}\\ 
+\end{equation}
+
+\section{Gain and benefit formula for a FlipIt game with virus propagation}
+\label{ch:8:GainVirus}
+The formulas from the previous section can now be adapted to calculate the gain of both players in a FlipIt game with virus propagation.
+As mentioned before, the attacker will try to infect all the nodes in the network.
+He will do this by flipping the node in the graph that can infect all the nodes in the shortest time possible.
+After dropping a virus on the first node, it takes a while for the virus to infect the entire network.
+The time that it takes for the virus to infect every node will be denoted as parameter \textit{d}.
+If we want to measure how long it takes for the virus to infect all the nodes in the network, we have to calculate the shortest path from the
+first infected node to the farthest node.
+This can be measured by a method explained in section [matrix berekeningen].
+Assume that an attacker attacks at time t, then only at time t + \textit{d} he gains control over the entire network.
+If the defender flips the network before the period d has elapsed (so, somewhere between t and t+ \textit{d}), then the attacker will never gain control over the entire network.
+Using this parameter \textit{d}, a FlipIt game with virus propagation can be modelled.
+
+The previous section defined a formula to calculate each unit of control of the attacker for two cases.
+If the virus propagation takes d time before every resource is infected then this d has to be subtracted from each unit of control.
+(see figure 7.4 \ref{fig:virusflip}).
 It may happen that the unit of control is less than \textit{d}.
-The result of the substration will be a negative number in time.
-In this case this means that the defender has flipped all the resources before the attacker could gain control over all the resources.
-To calculate the gain only the units of control bigger than 0 have to be summarized.
+In that case, the result of the substration will be a negative number, meaning that the defender has flipped all the resources before the attacker could gain control over all the resources.
+To calculate the gain only the units of control bigger than 0 have to be summed.
 So the formula becomes:
 
+%%------------------------------------------------------------%%
+%% Formule
+%%------------------------------------------------------------%%
+
+For $\delta_{A} \geq \delta_{D}$ :
+
 \begin{equation}\label{first}
-Gain_{1} = \sum_{i=0}^{\delta_{0}} \lbrace [( 1 - i ) \cdot \delta_{1}] mod \delta_{0} - d \rbrace  > 0 \rbrace 
+Gain_{A} = \sum_{i=0}^{p} \lbrace [( - i ) \cdot \delta_{A}] mod \delta_{D} - d \rbrace  > 0 \rbrace 
 \end{equation}
 
+where \textit{p} is the number of units of control that have to be summed.
+
+The gain of the defender is equal to the amount of time that the attacker is not in control of the resource. So the formula for the defender becomes:
+\begin{equation}\label{first}
+Gain_{D} = p \cdot \delta_{A}  -  \sum_{i=0}^{p} \lbrace [( - i ) \cdot \delta_{A}] mod \delta_{D} - d \rbrace  > 0 \rbrace 
+\end{equation}
+
+For $\delta_{D} \geq \delta_{A}$ :
+
+%%------------------------------------------------------------%%
+%% Figuur
+%%------------------------------------------------------------%%
 \begin{figure}[hbtp]
@@
 
-To be able to calculate the benefit of the attacker, we need to calculate the average gain rate of the attacker.
-We will look at this for two cases.
-One case where $\delta_{0}$ and $\delta_{1}$ $\in$ \(\mathbb{Q}\) and the other one where $\delta_{0}$ and $\delta_{1}$ $\in$ \(\mathbb{I}.\) \\
+Calculating the benefit of the attacker, requires calculating the average gain rate of the attacker.
+To compute the benefit the value of parameter \textit{p} needs to be determined.
+Two cases can be considered: one case where $\delta_{D}$ and $\delta_{A}$ $\in$ \(\mathbb{Q}\) and the other one where $\delta_{D}$ and $\delta_{A}$ $\in$ \(\mathbb{I}.\) \\
 
-\textbf{Rational numbers (\(\mathbb{Q}\)):} For $\delta_{0}$ and $\delta_{1}$ $\in$ \(\mathbb{Q}\), a repetition of the game will occur.
-As every rational number is any number that can be expressed as the fraction p/q with p and q $\in$ \(\mathbb{Z}\) (integers), with the denominator q not equal to zero, it is possible to find the \textit{lcm} of $\delta_{0}$ and $\delta_{1}$.
+\textbf{Rational numbers (\(\mathbb{Q}\)):} When $\delta_{D}$ and $\delta_{A}$ are rational numbers, after a number of intervals (namely their least common multiple), the same pattern of intervals will be repeated over and over again.
+Why?
+A rational number is a number that can be expressed as the fraction p/q with p and q $\in$ \(\mathbb{Z}\) (integers), with the denominator q not equal to zero, it is possible to find the \textit{lcm} of $\delta_{D}$ and $\delta_{A}$.
 The \textit{lcm} is defined for all rational numbers as: $lcm(\dfrac{a}{b},\dfrac{c}{d})= \dfrac{a \cdot c}{gcd(b,d)} with  $ [\todo{referentie}].
-When \textit{t} is equal to the \textit{lcm} of $\delta_{0}$ and $\delta_{1}$, both players will move again at the same time and this can be mapped to the beginning of the game.
-Because we stated that at the end of the interval of the attacker, the defender is in control and because if two players move at the same time the moves are cancelled, we can map this to the beginning of the game.
-Since the game goes on infinite, to calculate the average gain of the attacker, it is sufficient to calculate the average gain of the attacker only during this repetition equal to the \textit{lcm} of $\delta_{0}$ and $\delta_{1}$.
-%This means that if we want to calculate the gain of the attacker we need to calculate the time the attacker has control over the total amount of time that has passed by.
-%For $\delta_{0}$ and $\delta_{1}$ $\in$ \(\mathbb{Q}\) we can see that we have a cycle.
-A pattern that comes back over and over again.
-That is when the amount of time is a multiple of $\delta_{0}$ and $\delta_{1}$ or the largest common multiplier (\textit{lcm}).
-At this point the Attacker and the Defender move at the same time what brings us back to the beginning.
-So to calculate the gain of $\delta_{0}$ and $\delta_{1}$ $\in$ \(\mathbb{Q}\) we need to calculate the amount of control units of the attacker that go into the length of time units equal to the \textit{lcm} of  $\delta_{0}$ and $\delta_{1}$.
-This will be equal to the amount of time that the period of the attacker goes into the \textit{lcm}.
-We denote this by parameter \textit{p} and define it as follow:
+When \textit{t} is equal to the \textit{lcm} of $\delta_{D}$ and $\delta_{A}$, both players will move again at the same time and this can be mapped to the beginning of the game.
+Because we stated that at the end of the interval of the attacker, the defender is in control and because if two players move at the same time the moves cancel each other out, we can map this to the beginning of the game.
+Since the game goes on infinitely, to calculate the average gain of the attacker, it is sufficient to calculate the average gain of the attacker only during a period of time equal to the \textit{lcm} of $\delta_{D}$ and $\delta_{A}$.
+Since \textit{lcm} is a multiple of $\delta_{D}$ and $\delta_{A}$, there is a number \textit{p} so that \textit{lcm} $=$p $\cdot \delta_{A}$, meaning that the attacker will have played \textit{p} times.
+\textit{p} can be defined as follows:
+
+%%------------------------------------------------------------%%
+%% Formule
+%%------------------------------------------------------------%%
 \begin{equation}\label{first}
-p = \dfrac{lcm(\delta_{0},\delta_{1})}{\delta_{1} } 
-\end{equation}
-After this calculation we summarize our units of control in function of \textit{p} and divide it by the \textit{lcm} of $\delta_{0} $ and $ \delta_{1}$, which is the total amount of time for one cycle.
-This gives us the following formula of the benefit of the attacker with a cost rate equal to zero:
-
-\begin{equation}\label{first}
-\beta_{1} = \dfrac{\sum_{i=0}^{p} \lbrace [( 1 - i ) \cdot \delta_{1}] mod \delta_{0} \rbrace }{lcm(\delta_{0},\delta_{1})} 
+p = \dfrac{lcm(\delta_{D},\delta_{A})}{\delta_{A} } 
 \end{equation}
 
-%We can also define a formula without the greatest common divider. Every $\delta_{0}$ and $\delta_{1}$ have to be written in a fraction:
-%\begin{equation}\label{first}
-%\delta_{0}=\dfrac{a}{b} ~~~~and~~~~\delta_{1}=\dfrac{c}{d}
-%\end{equation}
-%If  $\delta_{0}$ or $\delta_{1}$ is a Geheel getal then b or d will be 1. The formula for the gain becomes different:\\
-%
-%\todo{formule zoeken zonder lcm en gcd}
+This results in the following formula for the benefit of the attacker with a cost rate equal to zero:
+
+%%------------------------------------------------------------%%
+%% Formule
+%%------------------------------------------------------------%%
+\begin{equation}\label{first}
+\beta_{A} = \dfrac{\sum_{i=0}^{p} \lbrace [[(  - i ) \cdot \delta_{A}] mod \delta_{D} - d] > 0 \rbrace }{lcm(\delta_{D},\delta_{A})} 
+\end{equation}
+
+As stated before [], the benefit of the attacker and the benefit of the defender add up to 1 ($\beta_{A} + \beta_{B} = 1$). The benefit of the defender can be writen as follows:
+\begin{equation}\label{first}
+\beta_{D} = 1 - \beta_{A}
+\end{equation}
 
@@
 An irrational number $ i \neq \dfrac{a}{b}$ with $b \in$ \(\mathbb{Z}\) , a $\in$ \(\mathbb{N}\).
-Because we cannot write \textit{i} as a fraction, this means that we cannot compute a common multiplier of $\delta_{0}$ and $\delta_{1}$.
+
+Two cases can be distinguished.
+(A) $\dfrac{\delta_{D}}{ \delta_{A}}$ is a rational number a/b with a <= b. In that case, after b intervals, the pattern will repeat itself.
+
+Since either $\delta_{D}$ and $\delta_{A}$ cannot be written as a fraction, the least common multiplier cannot be calculated.
+Moreover, there will be no repeating pattern because we cannot write $\delta_{D}$ and/or $\delta_{A}$ as a fraction, this means that we cannot compute a common multiplier of $\delta_{D}$ and $\delta_{A}$.
 If their is no common multiplier the attacker and the defender won't move at one point on the same time, meaning that this does not result in a cycle.
-If we would have a cycle that means that there exists a number \textit{x} that can be divided by $\delta_{0}$ and $\delta_{1}$.
+If we would have a cycle that means that there exists a number \textit{x} that can be divided by $\delta_{D}$ and $\delta_{A}$.
 At $t=x$ both of the players will move at the same time, which is not possible because then their would be a cycle.
@@
 If it does that means that their is repetition, meaning again that their is a cycle.
-\todo{nog verder uitleggen} We can conclude that if we have no cycle and no number will be repeated twice, that it will enumerate every number between 0 and the biggest interval (which is $\delta_{0}$).
+\todo{nog verder uitleggen} We can conclude that if we have no cycle and no number will be repeated twice, that it will enumerate every number between 0 and the biggest interval (which is $\delta_{D}$).
 \textit{The reals are uncountable; that is: while both the set of all natural numbers and the set of all real numbers are infinite sets, there can be no one-to-one function from the real numbers to the natural numbers} [WikiPedia: real numbers] If they are uncountable that means that we cannot calculate the sum of all the numbers between 0 and the biggest interval.
@@
 
-What we can do is take the limit, count as many control units of time of the attacker and divide it by the greatest amount of time.
+%What we can do is take the limit, count as many control units of time of the attacker and divide it by the greatest amount of time.
 We can see that this eventually will result to the solution given by the writers of FlipIt.

### 2015-04-27 doc/template/chap-8.tex
tekst zo goed als klqqr

@@
 shortest time possible.
-After dropping a virus on the first resource, it takes a while for the virus to infect the entire network.
+After dropping a virus on the first node, it takes a while for the virus to infect the entire network.
 However, since the original FlipIt game works with a single resource that is always flipped entirely, the assumption is made that the attacker is considered to have gained the control over the resource only when all the nodes of the network have been infected, i.e.
@@
 This definition is however not easy to adapt to the situation with virus propagation.
-Therefore, this section presents an alternative formula in that defines a game by quantifying the amount of time each player has control.
+Therefore, this section presents an alternative formula that defines a game by quantifying the amount of time each player has control.
 \\
@@
 \item \textit{Unit of control}: Defined as the period between gaining (full) control and losing control over the resource.  
-\item $n_{D}$: n is the n'th interval of the defender if $\delta_{D} \geq \delta_{A}$ , starting from interval 0.
-\item $n_{A}$: n is the n'th interval of the attacker if $\delta_{A} \geq \delta_{D}$ , starting from interval 0.
-\item $\Delta Unit_{D}(n)$ : This is a function that denotes the length of a unit of control of the defender in the n'th interval of the defender, if $\delta_{D} \geq \delta_{A}$.
-\item $\Delta Unit_{A}(n)$ : This is a function that denotes the length of a unit of control of the attacker in the n'th interval of the attacker, if $\delta_{A} \geq \delta_{D}$.
+\item $n_{D}$: The n'th interval of the defender if $\delta_{D} \geq \delta_{A}$ , starting from interval 0.
+\item $n_{A}$: The n'th interval of the attacker if $\delta_{A} \geq \delta_{D}$ , starting from interval 0.
+\item $\Delta Unit_{D}(n)$ : This is a function that denotes the length of a unit of control of the defender in the n'th interval of the attacker or the defender depending on who is playing faster.
+\item $\Delta Unit_{A}(n)$ : This is a function that denotes the length of a unit of control of the attacker in the n'th interval of the attacker or the defender depending on who is playing faster.
 \item \textit{lcm(a,b)}: The least common multiple of a and b.
@@
 \begin{figure}[hbtp]
-\caption{Defining unit of control}
+\caption{Graphic representation of some of the notations}
 \centering
-\includegraphics[scale=1]{Images/FlipSpel.png}
+\includegraphics[scale=0.8]{Images/notqties}
 \label{fig:notations}
@@
 
-The gain formula can be calculated by taking all the units of control of the player up to an amount of p intervals of the attacker.
+The gain formula can be calculated by taking all the units of control of the player up to an amount of p intervals of the attacker. The gain formula for the attacker is stated as follows:
 
@@
 The length of a unit of control of the attacker is the remainder of the interval after the defender loses control over the resource when the attacker plays.
-This can be defines as follows for the n'th interval of the defender:
+This can be defined as follows for the n'th interval of the defender:
 \begin{equation}\label{first}
@@
 \begin{equation}\label{first}
-Gain_{D} =\sum_{i=1}^{p} \lbrace [( - i ) \cdot \delta_{D}] mod \delta_{A} \rbrace  \rbrace ~~with~~ i=0 ~~ Gain_{D}=\delta_{A}\\ 
+Gain_{D} =\sum_{i=1}^{p} \lbrace [( - i ) \cdot \delta_{D}] mod \delta_{A} \rbrace   ~~with~~ i=0 ~~ Gain_{D}=\delta_{A}\\ 
 \end{equation}
@@
 \begin{equation}\label{first}
-Gain_{A} =\delta_{D} \cdot p - \sum_{i=1}^{p} \lbrace [( - i ) \cdot \delta_{D}] mod \delta_{A} \rbrace  \rbrace ~~with~~ i=0 ~~ Gain_{A}=\delta_{D} - \delta_{A}\\ 
+Gain_{A} =\delta_{D} \cdot p - \sum_{i=1}^{p} \lbrace [( - i ) \cdot \delta_{D}] mod \delta_{A} \rbrace   ~~with~~ i=0 ~~ Gain_{A}=\delta_{D} - \delta_{A}\\ 
 \end{equation}
@@
 
-The previous section defined a formula to calculate each unit of control of the attacker for two cases.
-If the virus propagation takes d time before every resource is infected then this d has to be subtracted from each unit of control.
+The previous section defined a formula to calculate each unit of control of the attacker and the defender for two cases.
+If the virus propagation takes d time before every resource is infected then this \textit{d} has to be subtracted from each unit of control.
 (see figure 7.4 \ref{fig:virusflip}).
@@
 To calculate the gain only the units of control bigger than 0 have to be summed.
-So the formula becomes:
+So the formula becomes: \\
 
@@
 For $\delta_{D} \geq \delta_{A}$ :
+\begin{subequations}\label{grp}
+\begin{align}
+Gain_{A} =\delta_{D} \cdot p - \sum_{i=1}^{p} \lbrace [( - i ) \cdot \delta_{D}] mod \delta_{A} -d \rbrace > 0 \rbrace \\
+with~~ i=0 ~~ Gain_{A}=\delta_{D} -( \delta_{A} - d) > 0\\ 
+\end{align}
+\end{subequations}
+
+and for the defender:
+\begin{subequations}\label{grp}
+\begin{align}
+Gain_{D} =\sum_{i=1}^{p} \lbrace [( - i ) \cdot \delta_{D}] mod \delta_{A} - d \rbrace > 0  \rbrace \\
+with~~ i=0 ~~ Gain_{D}=\delta_{A} - d > 0\\ 
+\end{align}
+\end{subequations}
 
@@
 
-Calculating the benefit of the attacker, requires calculating the average gain rate of the attacker.
+Calculating the benefit of both players, requires calculating the average gain rate of both players.
 To compute the benefit the value of parameter \textit{p} needs to be determined.
-Two cases can be considered: one case where $\delta_{D}$ and $\delta_{A}$ $\in$ \(\mathbb{Q}\) and the other one where $\delta_{D}$ and $\delta_{A}$ $\in$ \(\mathbb{I}.\) \\
+Two cases can be considered: one case where $\delta_{D}$ and $\delta_{A}$ $\in$ \(\mathbb{Q}\) and the other one where $\delta_{D}$ and $\delta_{A}$ $\in$ \(\mathbb{I}.\) In both cases we first calculate the benefit of the attacker in case that the defender moves at least as fast as the attacker.
+The benefit of the defender will be BenD = 1 - BenA.
+The benefit of both players for the case where the defender moves as least as fast as the defender is done in a similar way.
+\\
 
@@
 
-Since either $\delta_{D}$ and $\delta_{A}$ cannot be written as a fraction, the least common multiplier cannot be calculated.
-Moreover, there will be no repeating pattern because we cannot write $\delta_{D}$ and/or $\delta_{A}$ as a fraction, this means that we cannot compute a common multiplier of $\delta_{D}$ and $\delta_{A}$.
-If their is no common multiplier the attacker and the defender won't move at one point on the same time, meaning that this does not result in a cycle.
+(B)If either $\delta_{D}$ and $\delta_{A}$ cannot be written as a fraction, and they are no multiple of each other, the least common multiplier cannot be calculated.
+Moreover, there will be no repeating pattern.
+If both players move at one point in the game at the same moment, this point of time has to be a multiple of the period of the attacker and a multiple of the period of the defender.
+But because there is no least common multiple, no such point of time exists during the game.
+If both players never play at the same moment, it is not possible to have a repeated pattern because no mapping to the beginning of the game can occur.
+Additionly two unit of controls with the same length cannot exist.
+This would mean that the game has a repeated pattern, which is not possible.
+\\
+The game will go on forever, if no repeating pattern occurs and it would keep on generating units of control with different lengths.
+This implies that if the game goes on forever, every length between 0 and the smallest interval (which is $\delta_{D}$ ) will be generated.
+%If their is no common multiplier the attacker and the defender won't move at one point on the same time, meaning that this does not result in a cycle.
 If we would have a cycle that means that there exists a number \textit{x} that can be divided by $\delta_{D}$ and $\delta_{A}$.
@@
 If it does that means that their is repetition, meaning again that their is a cycle.
-\todo{nog verder uitleggen} We can conclude that if we have no cycle and no number will be repeated twice, that it will enumerate every number between 0 and the biggest interval (which is $\delta_{D}$).
+\todo{nog verder uitleggen} We can conclude that if we have no cycle and no number will be repeated twice, that it will enumerate every number between 0 and the biggest interval (which is $\delta_{D}$)
+To calculate the benefit we want to summarize the unit of controls up to a number of interval \textit{p}.
+Considering that the game goes on forever without repetition we cannot rely on the fact that the benefit can also be calculated only during the repetition.
+Calculating the benefit of a game without repetition would imply that all the unit of control to infinite have to be calculated.
+This implicates that all the numbers between 0 and   $\delta_{D}$ have to be summed but this is impossible.
 \textit{The reals are uncountable; that is: while both the set of all natural numbers and the set of all real numbers are infinite sets, there can be no one-to-one function from the real numbers to the natural numbers} [WikiPedia: real numbers] If they are uncountable that means that we cannot calculate the sum of all the numbers between 0 and the biggest interval.
@@
 The Field of the real numbers is ordered.
+\\
+
+\todo{nog mooie tekst schrijven in engels: hier essentie}
+We kunnen de benefit wel benaderen door een zo groot mogelijke som te nemen van de unit of controls.
+Uit deze benadering is af te leiden waar de verhouding naartoe zou gaan als de limiet zou genomen worden.
+-> laten zien met een voorbeeld van Pi en 1.
 

### 2015-04-29 doc/template/chap-8.tex
veranderingen van assistenten

@@
 After that section \ref{ch:8:GainVirus} introduces a modification to this formula to achieve an adapted gain formula for a FlipIt game with virus propagation.
-This allows to derive the benefit formula.\\
+This allows us to derive the benefit formula.\\
 
@@
 \label{ch:8:GainFlip}
-In their FlipIt paper [], Marten van Dijk and Ari Juels and Alina Oprea and Ronald L.
-Rivest, give a definition of the gain of a player \textit{i} .
-This definition is however not easy to adapt to the situation with virus propagation.
+In their FlipIt paper [], Marten van Dijk et al.
+, give a definition of the gain of a player \textit{i} .
+This definition is however, in the best of our knowledge, not easy to adapt to the situation with virus propagation.
 Therefore, this section presents an alternative formula that defines a game by quantifying the amount of time each player has control.
@@
 \begin{description}
-\item $\delta_{D}$: This is the period of the defender. This denotes the length of the interval between two consecutive moves of the defender. ($\delta_{D} > 0$)
-\item $\delta_{A}$: This is the period of the attacker. This denotes the length of the interval between two consecutive moves of the attacker. ($\delta_{A} > 0$)
+\item $\delta_{D}$: This is the period of the defender. This denotes the length of the interval between two consecutive moves of the defender ($\delta_{D} > 0$).
+\item $\delta_{A}$: This is the period of the attacker. This denotes the length of the interval between two consecutive moves of the attacker ($\delta_{A} > 0$).
 \item \textit{$T_{D}$}: This denotes the phase of the defender that was chosen randomly and uniformly over the interval [0,$\delta_{D}$].
@@
 \item \textit{Unit of control}: Defined as the period between gaining (full) control and losing control over the resource.  
-\item $n_{D}$: The n'th interval of the defender if $\delta_{D} \geq \delta_{A}$ , starting from interval 0.
-\item $n_{A}$: The n'th interval of the attacker if $\delta_{A} \geq \delta_{D}$ , starting from interval 0.
+\item $n_{D}$: The n'th interval of the defender, starting from interval 0.
+\item $n_{A}$: The n'th interval of the attacker, starting from interval 0.
 \item $\Delta Unit_{D}(n)$ : This is a function that denotes the length of a unit of control of the defender in the n'th interval of the attacker or the defender depending on who is playing faster.
 \item $\Delta Unit_{A}(n)$ : This is a function that denotes the length of a unit of control of the attacker in the n'th interval of the attacker or the defender depending on who is playing faster.
-\item \textit{lcm(a,b)}: The least common multiple of a and b.
-\item \textit{gcd(a,b)}: The greatest common divider of a and b.
+\item \textit{lcm(a,b)}: The least common multiple of \textit{a} and \textit{b}.
+\item \textit{gcd(a,b)}: The greatest common divider of \textit{a} and \textit{b}.
 \item $G_{i}(t)$: Gain of player \textit{i} at time t.
@@
 %To compute the gain of the attacker 
-To compute the gain of a player in a periodic game without phases, two cases are considered: case 1 where the defender moves as least as fast as the attacker and case 2 where the attacker plays as least as fast as the defender.
+To compute the gain of a player in a periodic game without phases, two cases are considered: case 1 where the defender moves at least as fast as the attacker and case 2 where the attacker plays at least as fast as the defender.
 Next, the formula is enriched by introducing the phases.
@@
 
-To compute the gain formula for the attacker,the amount of time that the attacker has control over the game from the start of the game up to time t has to be calculated.
-This can be done by computing the sum of all the units of control of the attacker up to time t.
+To compute the gain formula for the attacker,the amount of time that the attacker has control over the resource from the start of the game up to time \textit{t} has to be calculated.
+This can be done by computing the sum of all the units of control of the attacker up to time \textit{t}.
 
@@
 Considering that the defender moves at least as fast as the attacker, he or she will at least move one time during the interval of the attacker.
-Because the attacker only moves at the start of his or her interval we can say that the defender will always end as being in control of the resource.
+Because the attacker only moves at the start of his or her interval we can say that the defender will always end as being in control of the resource at the end of an attacker's interval.
 \todo{tekening met voorbeeld}. \\
 
-To calculate how long the unit of control of the attacker is in the n'th interval, we only need to know how long the attacker has control over the resource before the defender moves in that interval.
-The start time of the n'th interval will be a multiple of the period of the attacker.
-Once the attacker has played, the time he can stay in control until the next move of the defender, depends on the time elapsed since the last time the defender played in the previous interval ( the (n-1)th interval).
+To calculate how long the unit of control of the attacker is in the \textit{n}'th interval, we only need to know how long the attacker has control over the resource before the defender moves in that interval.
+The start time of the \textit{n}'th interval will be a multiple of the period of the attacker.
+Once the attacker has played, the time he can stay in control until the next move of the defender, depends on the time elapsed since the last time the defender played in the previous interval ( the (\textit{n}-1)th interval).
 
-The time the defender plays in an interval is a multiple of its period.
-Once the defender has played for the last time in the (n-1)'th interval, the remaining time until the attacker will play can be calculated as the remainder of $n\cdot \delta_{A}$ divided by $\delta_{D}$ or $n \cdot \delta_{a}~modulo~ \delta_{D}$.
-The time the attacker will stay in control is then $\delta_{D}- n \cdot \delta_{A}~ modulo ~ \delta_{D}$, which can also be calculated as $( -n \cdot \delta_{A}) ~modulo~ \delta_{D}$.
+The time at which the defender plays in an interval is a multiple of its period.
+Once the defender has played for the last time in the (\textit{n}-1)'th interval, the remaining time until the attacker will play can be calculated as $n \cdot \delta_{a}~modulo~ \delta_{D}$ which is equal to the remainder of $n\cdot \delta_{A}$ divided by $\delta_{D}$.
+[referentie naar matworks: $http://nl.mathworks.com/help/matlab/ref/mod.html$ ]The time the attacker will stay in control is then $\delta_{D}- n \cdot \delta_{A}~ modulo ~ \delta_{D}$, which can also be calculated as $( -n \cdot \delta_{A}) ~modulo~ \delta_{D}$.
 
@@
 
-This brings us to the next formula to calculate the length of a unit of control in the n'th interval of the attacker.
+This brings us to the next formula to calculate the length of a unit of control in the \textit{n}'th interval of the attacker.
 
-For every positive and non zero real $\delta_{A}$ and $\delta_{D}$ $\in$ \(\mathbb{R}\) and every n $\in$ \(\mathbb{N}\) (including 0 in the set of natural numbers) :
+For every positive and non zero real $\delta_{A}$ and $\delta_{D}$ $\in$ \(\mathbb{R}\) and every \textit{n} $\in$ \(\mathbb{N}\) (including 0 in the set of natural numbers) :
 \begin{equation}\label{first}
@@
 \end{equation}
-where $n_{A}$ is the number of the n'th interval of the attacker starting form interval 0 where the length of the unit of control of the attacker is calculated.\\
+where $n_{A}$ is the number of the \textit{n}'th interval of the attacker starting from interval 0 where the length of the unit of control of the attacker is calculated.\\
 
 The length of a unit of control of the defender is the remainder of the interval after the attacker loses control over the resource when the defender plays.
-This can be defines as follows for the n'th interval of the attacker:
+This can be defines as follows for the \textit{n}'th interval of the attacker:
 \begin{equation}\label{first}
@@
 
-An example: Figure \ref{fig:modulo} shows a FlipIt game were the period of the attacker is Pi and the period of the defender is 1. ..\todo{aanvullen}.
+An example: Figure \ref{fig:modulo} shows a FlipIt game were the period of the attacker is $\pi$ and the period of the defender is 1. ..\todo{aanvullen}.
 On figure \ref{fig:unitofcontrolformula}
@@
 \begin{figure}[hbtp]
-\caption{Example for calculating the control unit in interval 3 for a FlipIt game with period defender = 1 and period attacker = Pi}
+\caption{Example for calculating the control unit in interval 3 for a FlipIt game with period defender = 1 and period attacker = pi}
 \centering
@@
 
-The gain formula can be calculated by taking all the units of control of the player up to an amount of p intervals of the attacker. The gain formula for the attacker is stated as follows:
+The gain formula can be calculated by taking all the units of control of the player up to an amount of \textit{p} intervals of the attacker. The gain formula for the attacker is stated as follows:
 
@@
 
-The gain of the defender is the sum of the units of control of the defender up to the same amount of p intervals of the attacker:
+The gain of the defender is the sum of the units of control of the defender up to the same amount of \textit{p} intervals of the attacker:
 \begin{equation}\label{first}
@@
 
-Note: The gain formula is not in function of time t but the amount of p intervals of the attacker.
+Note: The gain formula is not in function of time \textit{t }but the amount of \textit{p} intervals of the attacker.
 This approach is chosen because it will result in whole units of control.
@@
 \end{equation}
-where $n_{D}$ is the number of the n'th interval of the defender starting form interval 0 where the length of the unit of control of the defender is calculated.\\
+where $n_{D}$ is the number of the \textit{n}'th interval of the defender starting form interval 0 where the length of the unit of control of the defender is calculated.\\
 
@@
 This can be measured by a method explained in section [matrix berekeningen].
-Assume that an attacker attacks at time t, then only at time t + \textit{d} he gains control over the entire network.
-If the defender flips the network before the period d has elapsed (so, somewhere between t and t+ \textit{d}), then the attacker will never gain control over the entire network.
+Assume that an attacker attacks at time \textit{t}, then only at time $t+d$ he gains control over the entire network.
+If the defender flips the network before the period d has elapsed (so, somewhere between \textit{t} and $t+d$), then the attacker will never gain control over the entire network.
 Using this parameter \textit{d}, a FlipIt game with virus propagation can be modelled.
@@
 The previous section defined a formula to calculate each unit of control of the attacker and the defender for two cases.
-If the virus propagation takes d time before every resource is infected then this \textit{d} has to be subtracted from each unit of control.
+If the virus propagation takes $d$ time before every resource is infected then this \textit{d} has to be subtracted from each unit of control.
 (see figure 7.4 \ref{fig:virusflip}).
@@
 
-\subsubsection{Computing the benefit of a FLipIt game with virus propagation}
+\subsubsection{Computing the benefit of a FlipIt game with virus propagation}
 
@@
 A rational number is a number that can be expressed as the fraction p/q with p and q $\in$ \(\mathbb{Z}\) (integers), with the denominator q not equal to zero, it is possible to find the \textit{lcm} of $\delta_{D}$ and $\delta_{A}$.
-The \textit{lcm} is defined for all rational numbers as: $lcm(\dfrac{a}{b},\dfrac{c}{d})= \dfrac{a \cdot c}{gcd(b,d)} with  $ [\todo{referentie}].
+The \textit{lcm} is defined for all rational numbers as: $lcm(\dfrac{a}{b},\dfrac{c}{d})= \dfrac{lcm( a , c)}{gcd(b,d)} with  $ [\todo{referentie}].
 When \textit{t} is equal to the \textit{lcm} of $\delta_{D}$ and $\delta_{A}$, both players will move again at the same time and this can be mapped to the beginning of the game.
@@
 \begin{equation}\label{first}
-\beta_{A} = \dfrac{\sum_{i=0}^{p} \lbrace [[(  - i ) \cdot \delta_{A}] mod \delta_{D} - d] > 0 \rbrace }{lcm(\delta_{D},\delta_{A})} 
+\beta_{A} = \dfrac{\sum_{i=0}^{p}  [[(  - i ) \cdot \delta_{A}] mod \delta_{D} - d] > 0 }{lcm(\delta_{D},\delta_{A})} 
 \end{equation}
@@
 
-\textbf{Irrational numbers (\(\mathbb{I}\)):} If $\delta_{0}$ and/or $\delta_{1}$ $\in$ \(\mathbb{I}\):
+\textbf{Irrational numbers (\(\mathbb{I}\)):} If $\delta_{D}$ and/or $\delta_{A}$ $\in$ \(\mathbb{I}\):
 An irrational number $ i \neq \dfrac{a}{b}$ with $b \in$ \(\mathbb{Z}\) , a $\in$ \(\mathbb{N}\).

### 2015-05-04 doc/template/chap-9.tex
nieuwe formule

\chapter{Formula}
\label{cha:9}
%\documentclass[10pt]{article}
%\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%			Introduction Chapter 1				%%%%%%
%%%%%												%%%%%%
%%%%%												%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%------------------------------------------------%
%            Intro Game Theory 					 %
%------------------------------------------------%
\section{something}
\label{Cha:9:inrto}

Periodic Game with delay for the attacker:

\textbf{Case 1:} $\delta_{D} \leq \delta_{A} $(The defender plays at least as fast as the attacker.) \\

Let $r = \dfrac{\delta_{D}}{ \delta_{A} }$. The intervals between two consecutive defender's moves have length $\delta_{D}$. Consider a given defender move interval. The probability over the attacker's phase selection that the attacker moves in this interval is r. Given that the attacker moves within the interval, he moves exactly once within the interval (since $\delta_{D} \leq \delta_{A} $) and his move is distributed uniformly at random. \\

The expected period of attacker control within the interval would be r/2, without considering the delay. \\

However, because of the delay, the maximal time of control is reduced to $\delta_{D}-d$. If we consider a duration of $\delta_{D} \cdot \delta_{A}$ the attacker will play $\delta_{D}$ times. If the attacker plays soon enough it will get a gain of $\dfrac{\delta_{D}-d}{2}$ in $\delta_{D}-d$ of the cases. In \textit{d} cases it will receive a gain of zero. This is the case were the duration of the delay causes the defender to play before the attacker can get control over the resource. 
So the gain of the attacker can be expressed as follows:
\begin{equation}\label{first}
Gain =  \dfrac{\delta_{D}-d}{2} \cdot (\delta_{D}-d) + 0 \cdot d = \dfrac{\delta_{D}-d}{2} \cdot (\delta_{D}-d)
\end{equation}

The benefit of the attacker can be expressed as follows

\begin{equation}\label{first}
\beta_{A}(\alpha_{D},\alpha_{A}) = \dfrac { (\delta_{D}-d) ^{2}} {2 \cdot \delta_{D}  \delta_{A}} + k_{A} \cdot \alpha_{A}
\end{equation}
\begin{equation}\label{first}
\beta_{A}(\alpha_{D},\alpha_{A}) = \dfrac { \delta_{D}} {2 \cdot \delta_{A}} + k_{A} \cdot \alpha_{A} + \dfrac{d}{\delta_{A}} + \dfrac{d^{2}}{2 \cdot \delta_{A} \delta_{D}}
\end{equation}
\\

The benefit of the defender is then:

\begin{equation}\label{first}
\beta_{D}(\alpha_{D},\alpha_{A}) = 1 - \dfrac { (\delta_{D}-d) ^{2}} {2 \cdot \delta_{D} \cdot \delta_{A}} + k_{D} \cdot \alpha_{D}
\end{equation}
\begin{equation}\label{first}
\beta_{D}(\alpha_{D},\alpha_{A}) = 1 - \dfrac { \delta_{D}} {2 \cdot \delta_{A}} + k_{D} \cdot \alpha_{D} + \dfrac{d}{\delta_{A}} + \dfrac{d^{2}}{2 \cdot \delta_{A} \delta_{D}}
\end{equation}

\textbf{Case 2:} $\delta_{A} \leq \delta_{D} $ (The attacker plays at least as fast as the defender.) \\

Let $r = \dfrac{\delta_{D}}{ \delta_{A} }$. The intervals between two consecutive attacker's moves have length $\delta_{A}$. Consider a given attackers move interval. The probability over the attacker's phase selection that the defender moves in this interval is $\dfrac{\delta_{D}}{ \delta_{A} } = (1/r)$. Given that the defender moves within the interval, he moves exactly once within the interval (since$\delta_{A} \leq \delta_{D} $) and his move is distributed uniformly at random. \\

If we consider a duration of $\delta_{A} \cdot \delta_{D} $ there is a probability of $\dfrac{\delta_{A} } {\delta_{D}} $ that the defender moves within the interval of the attacker. The defender will then receive an average gain of $\dfrac{\delta_{A}} {2} $. There is $1- \dfrac{\delta_{A} } {\delta_{D}} $ probability that the defender will not move in the interval of the attacker and so the defender will receive no gain. The benefit can be expressed as follows when the  defender plays $\delta_{D}$ times during a duration of $\delta_{A} \cdot \delta_{D}$:
\begin{equation}\label{first}
\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac { 1} {\delta_{A} \delta_{D}} \cdot \delta_{D} \cdot [ \dfrac{\delta_{A}}{\delta_{D}} \cdot \dfrac{\delta_{A}}{2}+[ 1-\dfrac{\delta_{A}}{ \delta_{D}}] \cdot 0 ]
\end{equation}
\begin{equation}\label{first}
\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac {\delta_{A} }{2 \cdot \delta_{D}} $$same as the FlipIt solution$$
\end{equation}

However, because of the delay, the maximal time of control of the defender is increased by d. In other words, the defender has some benefit time of d before the attacker really gains control over the resource, meaning that the attacker gains control only after $\delta_{A}+d$ instead of after $\delta_{A}$. So, when the defender plays, with a probability of $\dfrac{\delta_{A} } {\delta_{D}} $, the expected gain of the defender's control in this interval would be more than half of the period $\delta_{A} $: it is $\dfrac{\delta_{A} + d}{2}$. There is $1- \dfrac{\delta_{A} } {\delta_{D}} $ probability that the defender will not move in the interval of the attacker but because of the delay the defender will receive a gain of d. So the benefit of the defender can be expressed as:
\begin{equation}\label{first}
\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac { 1} {\delta_{A} \delta_{D}} \cdot \delta_{D} \cdot [ \dfrac{\delta_{A} }{\delta_{D}} \cdot \dfrac{\delta_{A} + d}{2}+[ 1-\dfrac{\delta_{A}}{ \delta_{D}}] \cdot d ]
\end{equation}
\begin{equation}\label{first}
\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac {\delta_{A} - d}{2 \cdot \delta_{D}} + \dfrac{d}{\delta_{A}} 
\end{equation}

%nce the defender has played, it takes 1 time until the attacker plays and the defender loses control again. Therefere, the expected period of attacker control within the interval would be 1/2.1/0, without considering the delay.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "thesis"
%%% End:

%\end{document}

### 2015-05-05 doc/template/chap-9.tex
nieuwe formule
case 1 opnieuw geschreven

@@
 %------------------------------------------------%
+
+\section*{Playing periodically with virus propagation}
+
+This chapter explains how to model a FlipIt game with a virus propagation that infects a network.
+The first section explains the difference between a normal FlipIt game and a FlipIt game with virus propagation.
+The next section derives a formula to calculate the benefit for a FlipIt game with a virus propagation.
+In the last section we calculate the Nash equilibrium for the benefit formula.
+
+\section{Explaining difference between FlipIt with and without virus propagation}
+zelfde als in vorige chapter:
+
+In chapter 2 the FlipIt game was explained.
+This chapter starts from the specific case of a non-adaptive continuous FlipIt game where both players play a periodic strategy with a random phase.
+This choice is motivated by the assumption that in the practical situation of most organisations, the defence strategy is to periodically defend the network.
+This corresponds to a periodic defender strategy.
+To simplify the analysis in a first time, a periodic attacker strategy is assumed as well.
+Further research can investigate the effect of relaxing this assumption.\\
+
+A FlipIt game consists of a single resource. To represent the security problem, the game now defines its single resource as a computer network with multiple
+nodes. One of the players, the defender, will try to defend his network. The defender
+will do this by flipping all the nodes of the network (i.e. the entire resource) in every move he plays. The
+attacker, the other player, will try to infect all the nodes in the network. The attacker
+will do this by flipping the node in the graph that can infect all the nodes in the
+shortest time possible.
+After dropping a virus on the first node, it takes a while for the virus to infect the entire network.
+However, since the original FlipIt game works with a single resource that is always flipped entirely, the assumption is made that the attacker is considered to have gained the control over the resource only when all the nodes of the network have been infected, i.e.
+the entire resource has been flipped.
+
+After dropping a virus on the first resource, it takes a while for the virus to infect
+the entire network. The time that it takes for the virus to infect every node will be
+denoted as parameter d. If we want to measure how long it takes for the virus to
+infect all the nodes in the network, we have to calculate the shortest path from the
+first infected node to the farthest node. This can be measured by a method that we
+will explain in section [].
+Assume that an attacker attacks at time t, then only at time t + d he gains control over the entire network.
+If the defender flips the network before the period d has elapsed (so, somewhere between t and t+ d), then the attacker will never gain control over the entire network.
+Using this parameter d, a FlipIt game with virus propagation
+can be modelled.
+
+\section{Benefit for FlipIt game with virus propagation}
+
+Periodic Game with delay for the attacker:
+
+\textbf{Case 1:} $\delta_{D} \leq \delta_{A} $(The defender plays at least as fast as the attacker.) \\
+
+Let $r = \dfrac{\delta_{D}}{ \delta_{A} }$.
+The intervals between two consecutive defender's moves have length $\delta_{D}$.
+Consider a given defender move interval.
+The probability over the attacker's phase selection that the attacker moves in this interval is r.
+Given that the attacker moves within the interval, he moves exactly once within the interval (since $\delta_{D} \leq \delta_{A} $) and his move is distributed uniformly at random.
+\\
+
+The expected period of attacker control within the interval would be r/2, without considering the delay. \\
+
+However, because of the delay, the maximal time of control is reduced to $\delta_{D}-d$.
+There is a probability of \textit{r} that the attacker will move in the interval of the defender.
+The attacker has to play soon enough to gain control, meaning that the attacker has to play during the period of $\delta_{D}-d$ during the interval of the defender.
+There is $\dfrac{\delta_{D}-d}{\delta_{D}}$ probability that the attacker will move soon enough which gives the attacker a gain of $\dfrac{\delta_{D}-d}{2}$.
+If the attacker moves after the period of $\delta_{D}-d$, the gain of the attacker will be zero.
+The average gain rate of the attacker can be expressed as follows if we look at one interval of the defender:
+\begin{equation}\label{first}
+\beta_{A}(\alpha_{D},\alpha_{A}) = \dfrac {1}{\delta_{D}} [ \dfrac{\delta_{D}}{\delta_{A}} \cdot \dfrac{\delta_{D}-d}{\delta_{D}} \cdot \dfrac{\delta_{D}-d}{2} + \dfrac{\delta_{D}}{\delta_{A}} \cdot \dfrac{d}{\delta_{D}} \cdot 0 ]
+\end{equation}
+
+To complete the formula to derive the benefit function, the cost of moving is added. In the second formula we can see the formula of the original FlipIt game.
+\begin{equation}\label{first}
+\beta_{A}(\alpha_{D},\alpha_{A}) = \dfrac { (\delta_{D}-d) ^{2}} {2 \cdot \delta_{D}  \delta_{A}} - k_{A} \alpha_{A}
+\end{equation}
+\begin{equation}\label{first}
+\beta_{A}(\alpha_{D},\alpha_{A}) = \dfrac { \delta_{D}} {2 \cdot \delta_{A}} - k_{A} \alpha_{A} + \dfrac{d}{\delta_{A}} + \dfrac{d^{2}}{2 \cdot \delta_{A} \delta_{D}}
+\end{equation}
+
+ The benefit of the defender is expressed as follows:
+ \begin{equation}\label{first}
+\beta_{D}(\alpha_{D},\alpha_{A}) = 1 - \dfrac { (\delta_{D}-d) ^{2}} {2 \cdot \delta_{D}  \delta_{A}} - k_{D} \alpha_{D}
+\end{equation}
+
 \section{something}
@@
 \begin{equation}\label{first}
-\beta_{D}(\alpha_{D},\alpha_{A}) = 1 - \dfrac { \delta_{D}} {2 \cdot \delta_{A}} + k_{D} \cdot \alpha_{D} + \dfrac{d}{\delta_{A}} + \dfrac{d^{2}}{2 \cdot \delta_{A} \delta_{D}}
+\beta_{D}(\alpha_{D},\alpha_{A}) = 1 - \dfrac { \delta_{D}} {2 \cdot \delta_{A}} + k_{D} \cdot \alpha_{D} - \dfrac{d}{\delta_{A}}- \dfrac{d^{2}}{2 \cdot \delta_{A} \delta_{D}}
 \end{equation}
@@
 \begin{equation}\label{first}
-\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac { 1} {\delta_{A} \delta_{D}} \cdot \delta_{D} \cdot [ \dfrac{\delta_{A}}{\delta_{D}} \cdot \dfrac{\delta_{A}}{2}+[ 1-\dfrac{\delta_{A}}{ \delta_{D}}] \cdot 0 ]
+\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac { 1} {\delta_{A} \delta_{D}} \cdot \delta_{D} \cdot [ \dfrac{\delta_{A}}{\delta_{D}} \cdot \dfrac{\delta_{A}}{2}+[ 1-\dfrac{\delta_{A}}{ \delta_{D}}] \cdot 0 ] + k_{D} \cdot \alpha_{D}
 \end{equation}
 \begin{equation}\label{first}
-\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac {\delta_{A} }{2 \cdot \delta_{D}} $$same as the FlipIt solution$$
+\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac {\delta_{A} }{2 \cdot \delta_{D}} + k_{D} \cdot \alpha_{D} $$same as the FlipIt solution$$
 \end{equation}
@@
 \begin{equation}\label{first}
-\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac { 1} {\delta_{A} \delta_{D}} \cdot \delta_{D} \cdot [ \dfrac{\delta_{A} }{\delta_{D}} \cdot \dfrac{\delta_{A} + d}{2}+[ 1-\dfrac{\delta_{A}}{ \delta_{D}}] \cdot d ]
+\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac { 1} {\delta_{A} \delta_{D}} \cdot \delta_{D} \cdot [ \dfrac{\delta_{A} }{\delta_{D}} \cdot \dfrac{\delta_{A} + d}{2}+[ 1-\dfrac{\delta_{A}}{ \delta_{D}}] \cdot d ] + k_{D} \cdot \alpha_{D}
 \end{equation}
 \begin{equation}\label{first}
-\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac {\delta_{A} - d}{2 \cdot \delta_{D}} + \dfrac{d}{\delta_{A}} 
+\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac {\delta_{A} - d}{2 \cdot \delta_{D}} + \dfrac{d}{\delta_{A}}  + k_{D} \cdot \alpha_{D}
+\end{equation}
+
+The benefit of the attacker is expressed as follows:
+\begin{equation}\label{first}
+\beta_{A}(\alpha_{D},\alpha_{A}) = 1- [ \dfrac {\delta_{A} - d}{2 \cdot \delta_{D}} + \dfrac{d}{\delta_{A}}  ] + k_{A} \cdot \alpha_{A}
 \end{equation}

### 2015-05-06 doc/template/chap-2.tex
alles committen

--- 2015-03-14 doc/template/chap-2.tex
@@
 
-\section{Extensions on FlipIt}
+In this chapter, we introduce the game \flip{FlipIt} \cite{FlipIt}.
+\flip{FlipIt} is a game introduced by ..
+et al.
+First we explain the framework of FlipIt and it's important results.
+In the next section the formulas and assumptions are made that will be used throughout the paper.
+To understand how to model a FlipIt game with virus propagation it is important to get familiar with the concepts of the normal FlipIt game and it's notations.
 
-There a various possible ways to extend \flip{FlipIt}.
-For instance Laszka et al.
-extended the basic \flip{FlipIt} game to multiple resources.
-The incentive is that for compromising a system in a real case it needs more than just taking over one resource.
-An example is gaining access to a system and breaking the password.
-The model is called FlipThem \cite{FlipThem}.
-Two ways of flipping the resources are used: the AND and the OR control model.
-In the AND model the attacker only controls the system if he controls all the resources of the system, whereas in the OR model the attacker only needs to compromise one resource to be in control of the entire system.
-The difference with FlipThem and this paper is that we introduce a Graph Model in the beginning.\\
-Another extension on FlipIt is done by Pham\cite{GameTheorApprCostBenefitAnalyses} [\todo{citatie needed voor Are We Compromised?}].
-Beside the action Flip their is another action Test.
-The basic idea is to test with an extra action if the resource has been compromised or not.
-This action involves also an extra cost.
-This model is useful if somebody wants to know for example if his password has been compromised or wants to assess the periodic security of a system.
-In \cite{MitigationCovert} \cite{MitigationNonTargeted} Laszka et al.
-they also consider non targeted attacks by non-strategic players and \todo{verder aanvullen}.
+\section{The FlipIt game}
+\flip{FlipIt} is a two-players game with a shared (single) resource that the players want to control as long as possible.
+The shared resource can be a password, a network or a secret key depending on the setting being modelled.
+In the rest of the paper we will call the two players the attacker, denoted by the subscript \textit{A} and the defender, denoted by subscript \textit{D}.
+
+The game begins at $t=0$ and continuous indefinitely ($t \rightarrow \infty $).
+The time in the game can be viewed as being continuous, but a discrete time can also be viewed.
+To get control over the resource, the players \textit{i} can flip the resource at any given time.
+A flip will be regarded as a move from a player \textit{i}.
+Each move will imply a certain cost $k_{i}$ and the cost can vary for each player.
+Both players will try to minimize their cost.
+By adding a cost, it will prevent players to move to frequently.
 \\
 
-In this section, we introduce the game \flip{FlipIt} \cite{FlipIt}.
-\flip{FlipIt} is a game introduced by ..
-..
-and Rivest.
-First we explain the framework of FlipIt and after that the formulas and assumptions that we will make for the game for during the whole paper.
-
-\section{The First Topic of this Chapter}
-\flip{FlipIt} is a two-players game with a shared (single) resource that the players want to control as long as possible.
-The shared resource can be a password, a network or a secret key depending on the setting being modelled.
-In the rest of the paper we will call the two players the Attacker and the Defender.
-To get control over the resource, the players can flip the resource at any given time.
-A flip will be regarded as a move from a player.
-Each move will imply a certain cost and the cost can vary for each player.
-Both players will try to minimize their cost.
-By adding a cost will prevent players to move to frequently.
-The unique feature of \flip{FlipIt} is that every move will happen in a stealthy way, meaning that the other player has no clue that the other player has flipped the resource.
+The unique feature of \flip{FlipIt} is that every move will happen in a stealthy way, meaning that the player has no clue that (his adversary) the other player has flipped the resource.
 For instance, the defender will not find out if the resource has already been compromised by the attacker, but he can only potentially know it after he flips the resource himself.
@@
 It may happen that the resource was already under control by the defender.
-If the defender moves when he or she has already control over the resource, he or she would have wasted move since it does not result in a change of ownership.
+If the defender moves when he or she has already control over the resource, he or she would have wasted a move since it does not result in a change of ownership and a cost is involved.
+\\
+
+The state of the resource is denoted as a time independent variable $C=C_{i}(t)$. 
+$C_{D}(t)$ is either 1 if the game is under control by the defender and 0 if the game is under control by the attacker. For $C^{A}(t)$ it is visa versa, $C^{A}(t)= 1 - C^{D}(t)$.\\ \\
+The game starts with defender being in control of the game, $C_{D}(0)= 1$.
+
+The players receive a benefit equal to the time of units that they were in possession of the resource minus the cost of making their moves. The cost of a player \textit{i} is denoted by $k_{i}$. 
+The total gain of player \textit{i} is equal to the total amount of time that a player \textit{i} has owned the resource from the beginning of the game up to time \textit{t}.
+It is expressed as follows:
+\begin{equation}\label{first}
+G_{i}(t) = integraal [0][t] C_{i}(x) dx.
+\end{equation}
+The average gain of player \textit{i} is defined as:
+\begin{equation}\label{first}
+\gamma_{i}(t) = G_{i}(t)/t.
+\end{equation}
+
+Let $\beta_{i}(t)$ denote player's \textit{i} average benefit upto time \textit{t}:
+\begin{equation}\label{first}
+\beta_{i}(t) = \gamma_{i}(t) - k_{i}\alpha_{i}.
+\end{equation}
+This is equal to the fraction of time the resource has been owned by player \textit{i}, minus the cost of making the moves.
+~$ \alpha_{i}$ defines the average move rate by player \textit{i} up to time \textit{t}.
+\\
 
@@
 
-\section{Figures}
 \begin{figure}[hbtp]
@@
 
+ \chapter{FlipIt game with virus propagation}
  \subsection{Actions of the attacker}
@@
 
+\section{Extensions on FlipIt}
+
+There a various possible ways to extend \flip{FlipIt}.
+For instance Laszka et al.
+extended the basic \flip{FlipIt} game to multiple resources.
+The incentive is that for compromising a system in a real case it needs more than just taking over one resource.
+An example is gaining access to a system and breaking the password.
+The model is called FlipThem \cite{FlipThem}.
+Two ways of flipping the resources are used: the AND and the OR control model.
+In the AND model the attacker only controls the system if he controls all the resources of the system, whereas in the OR model the attacker only needs to compromise one resource to be in control of the entire system.
+The difference with FlipThem and this paper is that we introduce a Graph Model in the beginning.\\
+Another extension on FlipIt is done by Pham\cite{GameTheorApprCostBenefitAnalyses} [\todo{citatie needed voor Are We Compromised?}].
+Beside the action Flip their is another action Test.
+The basic idea is to test with an extra action if the resource has been compromised or not.
+This action involves also an extra cost.
+This model is useful if somebody wants to know for example if his password has been compromised or wants to assess the periodic security of a system.
+In \cite{MitigationCovert} \cite{MitigationNonTargeted} Laszka et al.
+they also consider non targeted attacks by non-strategic players and \todo{verder aanvullen}.
+\\
+
 %%% Local Variables: 

### 2015-05-06 doc/template/chap-2.tex
ales even updaten, related work

@@
 \\
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[scale=0.8]{Images/DefFlipit}
+\caption{A representation of a FlipIt game where both players are playing periodically and discrete.
+Every move or "flip" is indicated by a blue or orange circle.
+The blue and orange rectangles represent the amount of time one of the players is in control of the resource.}
+\label{fig:FLipItDefault}
+\end{figure}
+\todo{verwijzen naar de figuur \ref{fig:FLipItDefault}}
 
@@
 
-In our assumption the strategy of both players will be non-adaptive. None of the players has information of the strategy of the opponent.
-
-\begin{figure}[hbtp]
-\centering
-\includegraphics[scale=0.25]{Images/FlipItDefault.jpg}
-\caption{The FlipIt game where both players are playing periodically}
-\label{fig:FLipItDefault}
-\end{figure}
-\todo{verwijzen naar de figuur \ref{fig:FLipItDefault}}
+In our analyses of the FlipIt game with a virus propagation in section [], we assume the strategy of both players to be non-adaptive.
+None of the players has information of the strategy of the opponent.
+The defender will never know when the attacker will attack the network with a virus.
+Conversely, the attacker does not know how often the defender defends his network.
 
@@
 
- In this subsection we go through the strategies used in FlipIt and the most important results. 
+ In this subsection we elaborate about the strategy of FlipIt . For the other strategies of FlipIt we refer the reader to [the paper of FlipIt]
  \begin{table}
@@
 
-\section{Extensions on FlipIt}
-
-There a various possible ways to extend \flip{FlipIt}.
-For instance Laszka et al.
-extended the basic \flip{FlipIt} game to multiple resources.
-The incentive is that for compromising a system in a real case it needs more than just taking over one resource.
-An example is gaining access to a system and breaking the password.
-The model is called FlipThem \cite{FlipThem}.
-Two ways of flipping the resources are used: the AND and the OR control model.
-In the AND model the attacker only controls the system if he controls all the resources of the system, whereas in the OR model the attacker only needs to compromise one resource to be in control of the entire system.
-The difference with FlipThem and this paper is that we introduce a Graph Model in the beginning.\\
-Another extension on FlipIt is done by Pham\cite{GameTheorApprCostBenefitAnalyses} [\todo{citatie needed voor Are We Compromised?}].
-Beside the action Flip their is another action Test.
-The basic idea is to test with an extra action if the resource has been compromised or not.
-This action involves also an extra cost.
-This model is useful if somebody wants to know for example if his password has been compromised or wants to assess the periodic security of a system.
-In \cite{MitigationCovert} \cite{MitigationNonTargeted} Laszka et al.
-they also consider non targeted attacks by non-strategic players and \todo{verder aanvullen}.
-\\
-
 %%% Local Variables: 

### 2015-05-06 doc/template/chap-3.tex
alles committen
ales even updaten, related work

@@
-%\chapter{Intoduction to GameTheory}
-%\label{cha:1}
-\documentclass[10pt]{article}
-\begin{document}
+\chapter{Related work}
+\label{cha:1}
 
-\section{Gametheory}
-
-Game theory is the study between self interested agents and the actions they perform.
-self interested agents means not that they want to harm each other or that they only care about themselves.
-Agent has a description of the world he likes.
-A utility function is a mapping of the world that the agent likes to real numbers.
-
-\section{Literatuurstudie}
+\section{Related work}
 
@@
 - Formal definition
+
+\section{Extensions on FlipIt}
+
+In this section we discuss the extentions that can be made on the FlipIt game.
+
+There a various possible ways to extend \flip{FlipIt}. 
+Laszka et al.
+made a lot of additions and extensions on the original game of FlipIt.
+For instance Laszka et al.
+extended the basic \flip{FlipIt} game to multiple resources.
+The incentive is that for compromising a system in a real case it needs more than just taking over just one resource.
+An example is that one resource can be gaining access to a system and breaking the password of the system is another resource.
+The model is called FlipThem \cite{FlipThem}.
+They use two ways to flip the multiple resources: the AND and the OR control model.
+In the AND model the attacker only controls the system if he controls all the resources of the system, whereas in the OR model the attacker only needs to compromise one resource to be in control of the entire system.
+%The difference with FlipThem and this paper is that we introduce a Graph Model in the beginning.\\
+Another addition of Laszka et al.
+to the game of FlipIt is extending the game to also consider non-targeted attacks by non-strategic players.
+In this game the defender tries to maintain control over the resource that is subjected to both targeted and non-targeted attacks.
+
+A last important addition from Laszka et al.
+to the game of FlipIt is   The moves made by the attacker are still covert but the moves made by the defender are known to the attacker.
+This means that the attacker can base his attacks on the defender's moves.
+Both the targeted and non-targeted attacks don't succeed immidiatly.
+For the targeted attack the time till it succeeds is given by an exponential distributed random variable with a known rate.
+The non-targeted attacks are modelled as a single attacker and the time till it succeeds is given by a Poisson process.\todo{misschien meer uitleggen}.
+The conclusion of this paper is that the optimal strategy for the defender is moving periodically.
+
+on FlipIt is done by Pham\cite{GameTheorApprCostBenefitAnalyses} [\todo{citatie needed voor Are We Compromised?}].
+Beside the action Flip their is another action Test.
+The basic idea is to test with an extra action if the resource has been compromised or not.
+This action involves also an extra cost.
+This model is useful if somebody wants to know for example if his or her password has been compromised or wants to assess the periodic security of a system.
+In \cite{MitigationCovert} \cite{MitigationNonTargeted} Laszka et al.
+they also consider non targeted attacks by non-strategic players and \todo{verder aanvullen}.
+\\
 
@@
 %%% End:
-
-\end{document}

### 2015-05-07 doc/template/chap-10.tex
extra toevoegingen

--- 2015-05-06 doc/template/chap-2.tex
@@
-\chapter{The FlipIt game}
-\label{cha:2}
+\chapter{FlipIt game with virus propagation}
+\label{cha:1}
+%\documentclass[10pt]{article}
+%\begin{document}
 
-In this chapter, we introduce the game \flip{FlipIt} \cite{FlipIt}.
-\flip{FlipIt} is a game introduced by ..
-et al.
-First we explain the framework of FlipIt and it's important results.
-In the next section the formulas and assumptions are made that will be used throughout the paper.
-To understand how to model a FlipIt game with virus propagation it is important to get familiar with the concepts of the normal FlipIt game and it's notations.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%			Introduction Chapter 1				%%%%%%
+%%%%%												%%%%%%
+%%%%%												%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\section{The FlipIt game}
-\flip{FlipIt} is a two-players game with a shared (single) resource that the players want to control as long as possible.
-The shared resource can be a password, a network or a secret key depending on the setting being modelled.
-In the rest of the paper we will call the two players the attacker, denoted by the subscript \textit{A} and the defender, denoted by subscript \textit{D}.
-
-The game begins at $t=0$ and continuous indefinitely ($t \rightarrow \infty $).
-The time in the game can be viewed as being continuous, but a discrete time can also be viewed.
-To get control over the resource, the players \textit{i} can flip the resource at any given time.
-A flip will be regarded as a move from a player \textit{i}.
-Each move will imply a certain cost $k_{i}$ and the cost can vary for each player.
-Both players will try to minimize their cost.
-By adding a cost, it will prevent players to move to frequently.
-\\
-
-The unique feature of \flip{FlipIt} is that every move will happen in a stealthy way, meaning that the player has no clue that (his adversary) the other player has flipped the resource.
-For instance, the defender will not find out if the resource has already been compromised by the attacker, but he can only potentially know it after he flips the resource himself.
-The goal of the player is to maximize the time that he or she has control over the resource while minimizing total cost of the moves.
-A move can also result in a "wasted move", called a flop.
-It may happen that the resource was already under control by the defender.
-If the defender moves when he or she has already control over the resource, he or she would have wasted a move since it does not result in a change of ownership and a cost is involved.
-\\
-
-\begin{figure}[hbtp]
-\centering
-\includegraphics[scale=0.8]{Images/DefFlipit}
-\caption{A representation of a FlipIt game where both players are playing periodically and discrete.
-Every move or "flip" is indicated by a blue or orange circle.
-The blue and orange rectangles represent the amount of time one of the players is in control of the resource.}
-\label{fig:FLipItDefault}
-\end{figure}
-\todo{verwijzen naar de figuur \ref{fig:FLipItDefault}}
-
-The state of the resource is denoted as a time independent variable $C=C_{i}(t)$. 
-$C_{D}(t)$ is either 1 if the game is under control by the defender and 0 if the game is under control by the attacker. For $C^{A}(t)$ it is visa versa, $C^{A}(t)= 1 - C^{D}(t)$.\\ \\
-The game starts with defender being in control of the game, $C_{D}(0)= 1$.
-
-The players receive a benefit equal to the time of units that they were in possession of the resource minus the cost of making their moves. The cost of a player \textit{i} is denoted by $k_{i}$. 
-The total gain of player \textit{i} is equal to the total amount of time that a player \textit{i} has owned the resource from the beginning of the game up to time \textit{t}.
-It is expressed as follows:
-\begin{equation}\label{first}
-G_{i}(t) = integraal [0][t] C_{i}(x) dx.
-\end{equation}
-The average gain of player \textit{i} is defined as:
-\begin{equation}\label{first}
-\gamma_{i}(t) = G_{i}(t)/t.
-\end{equation}
-
-Let $\beta_{i}(t)$ denote player's \textit{i} average benefit upto time \textit{t}:
-\begin{equation}\label{first}
-\beta_{i}(t) = \gamma_{i}(t) - k_{i}\alpha_{i}.
-\end{equation}
-This is equal to the fraction of time the resource has been owned by player \textit{i}, minus the cost of making the moves.
-~$ \alpha_{i}$ defines the average move rate by player \textit{i} up to time \textit{t}.
-\\
-
-Because the players move in a stealthy way, there are different types of feedback that a player can get while moving:
-\begin{itemize}
-\item Non-adaptive (NA): The player does not receive any feedback during the game while flipping.
-\item Last move (LM): When a player flips it will find out the exact time that the opponent played the last time.
-\item Full History (FH): When a player flips it will find out the whole history of the opponents move.
-\end{itemize}
-The game can be extended by the amount of information that a player receives. It can also be possible for a player to get information at the start of the game. Both interesting cases are:
-\begin{itemize}
-\item Rate-of-play (RP): The player finds out the exact rate of play of the opponent.
-\item Knowledge-of-strategy (KS): The player finds out the complete information of the strategy that the opponent is playing.
-\end{itemize}
-
-In our analyses of the FlipIt game with a virus propagation in section [], we assume the strategy of both players to be non-adaptive.
-None of the players has information of the strategy of the opponent.
-The defender will never know when the attacker will attack the network with a virus.
-Conversely, the attacker does not know how often the defender defends his network.
-
-\subsection{Strategies}
-
- In this subsection we elaborate about the strategy of FlipIt . For the other strategies of FlipIt we refer the reader to [the paper of FlipIt]
- \begin{table}
- \centering
- \begin{tabular}{ l | c  }
-  \textbf{Categories} & \textbf{Classes of Strategies} \\
-  \hline Non-adaptive (NA) & Exponential \\
-  & Periodic \\
-  & Renewal \\
-  & General non-adaptive \\
-  \hline Adaptive (AD) & Last move (LM) \\
-  & Full History (FH) \\  
-\end{tabular}
- \caption{Classes of strategies in FlipIt}
- \label{table:Strategies}
- \end{table}
-
-There are two different kinds of strategies, the \textit{non-adaptive strategies} and the \textit{renewal strategies}.
-If there is no need for feedback for both of the players, we say that we have a non-adaptive strategy.
-Because the player does not receive any feedback during the game it will play in the same manner against every opponent.
-They are not dependent on the opponents movements.
-This means that they can already generate the time sequence for all the moves in advance.
-But they can depend on some randomness because the non-adaptive strategies can be randomised.
-In this paper we will focus in the beginning on the non-adaptive strategies.
-Reasons behind this is that a player (defender or attacker) rarely knows what the strategies are of his opponent.
-[If the attacker wants to move stealthily, it might have limited attack options FLIPTHEM].
-\todo{nog redenen zoeken}\\
-A renewal strategy is a non-adaptive strategy where the time intervals between two consecutive moves are generated by a renewal process. \\
-
- \begin{description}
- \item Periodic
- \item Non-Arithmetic Renewal
- \item Exponential
- \end{description}
-
- \chapter{FlipIt game with virus propagation}
  \subsection{Actions of the attacker}
@@
 
-%%% Local Variables: 
-%%% mode: latex
-%%% TeX-master: "thesis"
-%%% End:
+%
+%In the following paragraph an introduction to game theory is given based on the work of \cite{leyton2008essentials} and \cite{Coursera}.
+For a more detailed and full introduction to game theory, the reader is referred to \cite{leyton2008essentials}.
+%%------------------------------------------------%
+%%            Intro Game Theory 					 %
+%%------------------------------------------------%
+%\section{Intro Game Theory}
+%\label{Cha:1:Intro.Game.Theory}
+%
+%%begin over dat gametheorie handig is in de economie
+%Game theory studies the interaction between independent and self-interested agents.
+It is a mathematical way of modelling the interactions between two or more agents where the outcomes depend on what everybody does and how it should be structured to lead to good outcomes.
+For this reason it is very important for economics and also for politics, biology, computer science, philosophy and a variety of other disciplines.
+\\
+%%Every agent has different levels of happiness for the different outcomes.
+%%self interested meaning
+%
+%One of the assumptions underlying game theory is that the players of the game, the agents, are independent and self-interested.
+This does not necessarily mean that they want to harm other agents or that they only care about themselves.
+%%utility function meaning 
+%Instead it means that each agent has preferences about the states of the world he likes.
+These preferences are mapped to natural numbers and are called the utility function.
+The numbers are interpreted as a mathematical measure to tell you how much an agent likes or dislikes the states of the world.
+\\
+%It also explains the impact of uncertainty.
+When an agent is uncertain about a distribution of outcomes, his utility will describe the expected value of the utility function with respect to the probability of the distribution of the outcomes.
+For example: with 0.7 probability it will be 7 degrees outside and with 0.3 probability it will be 10 degrees.
+The agent can have a different opinion about that distribution versus another distribution.
+(\todo{uitleggen aan de hand van een voorbeeld}).\\
+%%Cooperative and non cooperative games
+%In a decision game theoretic approach an agent will try to act in such a way to maximise his expected or average utility function.
+It becomes more complicated when two or more agents want to maximise their utility and whose actions can affect each other utilities.
+This kind of games are referred to as non cooperative game theory, where the basic modelling unit is the group of agents.
+The individualistic approach, where the basic modelling is only one agent, is referred as cooperative game theory.
+%
+%There are two standard representations for games. The first one is the Normal Form. The second one is the Extensive Form.
+%
+%In the following lists a couple of terms that will be used throughout the paper.
+%\begin{description}
+%\item \textit{Players}: players are referred as the ones who are the decision makers. It can be a person, a company or an animal.
+%\item \textit{Actions}: actions are what the player can do. 
+%\item \textit{Outcomes}:  
+%\item \textit{Utility function}: the utility function is the mapping of the level of happiness of an agent about the state of the world to natural numbers.
+%\item \textit{Strategies}: A strategy is the combination of different actions. A pure strategy is only one action.
+%\end{description}
+%
+%A game in game theory consists of multiple agents and every agent has a set of actions that he can play. 
+%
+%
+%
+%%Nash equilibrium
+%
+%
+%
+%% --------------- example of a game -----------------%
+%
+%
+%%------------------------------------------------%
+%%            Intro about virusses				 %
+%%------------------------------------------------%
+%\section{Virusses}
+%
+%Many network security threats today are spread over the Internet. The most common include:
+%
+%Viruses, worms, and Trojan horses
+%Spyware and adware
+%Zero-day attacks, also called zero-hour attacks
+%Hacker attacks
+%Denial of service attacks
+%Data interception and theft
+%Identity theft
+%
+%%http://www.ists.dartmouth.edu/library/258.pdf Email Virus Propagation Modeling and Analysis
+%%Cliff C. Zou∗, Don Towsley†, Weibo Gong∗
+%%∗Department of Electrical & Computer Engineering
+%%†Department of Computer Science
+%%Univ. Massachusetts, Amherst
+%%Technical Report: TR-CSE-03-04
+%
+%Computer virus through mail. 
+%Though virus spreading through email is an old technique, it is still effective and is widely used by
+%current viruses and worms. Sending viruses through email has some advantages that are attractive to
+%virus writers:
+% Sending viruses through email does not require any security holes in computer operating systems
+%or software.
+% Almost everyone who uses computers uses email service.
+% A large number of users have little knowledge of email viruses and trust most email they receive,
+%especially email from their friends [28][29].
+% Email are private properties like post office letters. Thus correspondent laws or policies are required
+%to permit checking email content for detecting viruses before end users receive email [18].
+%
+%Send a email with malicious attachment.
+Only again infected if attachment again opened.
+Thus this is the action of attacking every neighbour node + also can attack again the node where the virus was coming from.
+%There are also email viruses were the malicious program is hidden in the txt and the attachment does not need to be opened. 
+%
+%%http://www.cisco.com/web/offer/gist_ty2_asset/Cisco_2014_ASR.pdf p49
+%%http://repo.hackerzvoice.net/depot_madchat/vxdevl/papers/avers/2004-35.pdf
+%%http://www.mcafee.com/us/resources/white-papers/foundstone/wp-managing-malware-outbreak.pdf
+%
+%
+%
+%\subsection{Malware}
+%%Does a company network faces lot of malware? what is the cost ?
+%Relevant researches:
+%\begin{itemize}
+%%http://ants.iis.sinica.edu.tw/3BkMJ9lTeWXTSrrvNoKNFDxRm3zFwRR/17/04483668.pdf
+%\item How Viruses and worm can be detected. Difference between UDP en TCP worm propagation. Difference for the propagation speed. 
+%\end{itemize}
+%
+%
+%
+%Purpose thesis: model a worm propagation with adaptations of Flip-It.
+Flip-It cannot address the evaluation of individual nodes.
+Flip-It with multiple resources has not addressed the fact that a virus does not need to compromise the whole network.
+A subpart of the network can already cause problems for the company.
+Data leackages.
+%So the virus can propagate 
+%
+%
+%=======
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%			Conclusion Chapter 1				%%%%%%
+%%%%%%												%%%%%%
+%%%%%%												%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%\section{Conclusion}
+%\label{Cha:1:Conclusion}
+%The final section of the chapter gives an overview of the important results
+%of this chapter. This implies that the introductory chapter and the
+%concluding chapter don't need a conclusion.
+%
+%
+%%%% Local Variables: 
+%%%% mode: latex
+%%%% TeX-master: "thesis"
+%%%% End:
+
+%\end{document}

### 2015-05-07 doc/template/chap-2.tex
extra toevoegingen

@@
 
- \chapter{FlipIt game with virus propagation}
- \subsection{Actions of the attacker}
-A virus has different kind of ways of making his way through a company network.
-We will describe the different ways of how the virus can propagate.
-For start we will say that the virus or worm will be dropped on Node i and that it has k numbers of neighbours.
-\begin{enumerate}
-\item Node i is infected and will spread the virus or worm to every k neighbours and will stop infecting the neighbours in the next step
-\item Node i is infected and will spread the virus or worm to every k neighbours and will keep on spreading the virus to the same neighbours in every next step
-\item Node i is infected and will spread the virus to only one of the k neighbours and will stop infecting another neighbour in the next step
-\item Node i is infected and will spread the virus to only one of the k neighbours and in the next step it will infect another one of the k neighbours 
-\end{enumerate}
-
-In the game that will be modelled in the paper we will use the settings of the first spreading method.
-We will not use method 2 because this kind of propagation will float the network.
-Because we use the settings of a mail system and contact in a mailing list the method of 3 and 4 are not used.
-\\
-In the first method the node that has been infected can be again infected.
-If one of the neighbours infects the node again the node will infect his neighbours again.
-By using this spreading method we have three distinct states in which a node can be situated.
-An \textit{infected state}, a \textit{clean state} and a \textit{spreading state}.
-An infected state means that the node is infected and will not spread the virus to its neighbours, a clean state means that the node is not infected on that moment and a spreading state means that the node is infected and that it will spread the virus or worm to its neighbours in the next step.
-We can argument this kind of propagation through a mail worm. \todo{voorbeeld geven van zo een worm}
-%Another propagation method is that the virus works as a token. It will propagate to only one neighbour and continue to spread.
-
-The Attacker itself has two different ways of attacking the company network.
-It will only infected one node of the network and will wait for the virus to spread itself through the network.
-We will model two ways of attacks of an Attacker:
-\begin{enumerate}
-\item The attacker drops the virus on a random node on the network
-\item The attacker drops the virus on a targeted node on the network
-\end{enumerate}
-The attacker in this game will put a virus or worm on one of the nodes in the network.
-(This will happen at random.) The attacker does not know on which node the virus will be dropped.
-We will use this randomness because \todo{feit uit security rapport symantec} most viruses are spread via a usb stick or a shared resource.
-If we use this spreading method where we have a targeted attack the attacker will have more information about the network.
-\\
-
-The attacker can choose at which rate it will drop a virus on one of the nodes on the network.
-The cost of dropping a virus will be the same.
-It will not increase.
-If it will increase this means that the attacker will eventually drop out of the game because it becomes to expensive.\\
-The attacker is in control over the game if it manages to infect a subset of all the resources of the company network.
-
-\subsection{Actions of the defender}
-The attacker wants to protect all the nodes of his network.
-It can do so by getting back control over the resources.
-We will assume that the defender of the network has knowledge over his own network.
-Which is convenient in the real world because a company has to know how his infrastructure looks like.\\
-
-The defender has two possible ways of defending its network:
-\begin{enumerate}
-\item The defender flips all the nodes of his network
-\item The defender will flip a subset of the nodes of his network
-\end{enumerate}
-
-The cost of flipping all the nods of the network will be greater than the cost of flipping a subset of nodes.
-We make this assumption because otherwise it will be beneficial for the defender to always flip all the nodes in the network.\\
-
-We will also make the assumption that as a defender flips a node the node can get infected again.
-A flip will not be  correlated to a patch but to a clean-up.
-\todo{waarom geen patch, wormen kunnen veranderen gaandeweg}
-\todo{andere mogelijkheid:} Another setting of the game can be that the flip of the defender is equal to a patch and that the resource cannot be infected any more.
-But with this case we deviate from the flipIt game, because the attacker cannot flip the resource any more.
-Unless we work with different virusses every time the attacker flips.
-We start with the less complex game of flipping is equal to a clean-up.
-
-\subsection{Strategies of both players}
-We explained what the actions of each player are.
-
-\section{Formal definition Game}
-
-In this section we provide a formal definition of the game and the notation that we will use throughout the paper.
-
-\begin{description}
-% ---- PLAYERS ---- %
-\item \textit{Players}  There are two players in the game, one is the defender and the other one is the attacker. They are respectively identified by 0 and 1.
-
-% ---- TIME ---- %
-\item \textit{Time}  The game starts at $t=0$ and continuous indefinitely as $t \rightarrow \infty$. The game is a continuous game.
-
-% ---- GRAPH ---- %
-\item \textit{Graph} We represent the company network as a Graph $G = < V,E>$.
-G is an ordered pair where V denotes the set of resources or nodes in the network and E denotes the set of connections or links, which are a two-element subset of V.
-We use the notations resources and nodes interleaving in this paper.\\
-We have N resources in the network. $N \in $  \todo{aanvullen}. This means we can denote the resources by:
-\begin{center}
-$V \in {V_{0}, V_{1}, V_{2}, ... , V_{N} }$
-\end{center}
-The set E of connections indicates if there is a link between two resources.
-We see the links as bidirectional so the total graph is undirected.
-If there is a link between resource $V_{n}$ and $V_{n+1}$ then there is also a link between $V_{n+1}$ and $V_{n}$.
-
-% ---- GAME STATE ---- %
-\item \textit{Game State} There is also a time-dependent variable that represents the state of the game.
-$C=C(t)$ is either 0 if the game is under control by the defender and 1 if the Game is under control by the attacker.
-\\
-We start at $t=0$ with the defender who has control over the game.
-We do this because we assume that the defender will only put the network online without having a virus or worm in it.
-The Attacker can gain control over the game when it compromises a subset \textit{s} of the resources.
-The subset \textit{s} is a minimum of 1 resource and a maximum of all the resources N.
-\\
-\todo{deze variabele nodig ja of nee ?
-JA} We can also define the state of each resource by $C^{A}_{N}$ and $C^{D}_{N}$.
-If $C^{A}_{N} = 1$ then this means that the attacker has control over the resource, and 0 otherwise.
-For $C^{D}_{N}$ it is visa versa, $C^{D}_{N} = 1 - C^{A}_{N}$.\\
-
-% ---- MOVES ---- %
-\item \textit{Moves} Both players can make a move in the game.
-Moves done in a finite numbers of time in any finite time interval.
-Both players can play at any time they want, they can also play at the same time.
-If this happens the one that has control over the resource will keep having control over the resource.
-This makes the game fully symmetric \todo{beter uitleggen}. The sequence of move times are denoted by the following infinite sequence:
-\begin{center}
-$t=t_{1},t_{2},t_{3},..$
-\end{center}
-Two move times can be the same because we allow players to move at the same time.
-We can also denote the infinite sequence of times when player \textit{i}  moves. We write this as :
-\begin{center}
-$t=t_{i,1},t_{i,2},t_{i,3},..$ with \textit{i} $ \in $ $\lbrace 0,1 \rbrace$
-\end{center}
-The sequences $t_{1}$ and $t_{0}$ are disjoint subsets of the sequent t. 
-We can also denote who made the \textit{k}th move by defining a sequence \textit{p} that denotes the sequence of who played:
-\begin{center}
-$p=p_{1},p_{2},p_{3}, .. $ with $p_{k}$ $\in$ $\lbrace 0,1 \rbrace$
-\end{center}
-
-% ---- NUMVER OF MOVES ---- %
-\item \textit{Number of moves}  $n_{i}(t$ denotes the number of moves made by player \textit{i} up to and including time t. This means that 
-\begin{center}
-$n(t)=n_{1}(t) + n_{0}(t)$
-\end{center}
-is the sum of the number of moves made by the defender and the attacker up to and including time t.
-
-% ---- AVERAGE MOVE RATE ---- %
-\item \textit{Average move rate} We denote $\alpha_{i}(t)$ as the average move rate by player i:
-\begin{center}
-$\alpha_{i}(t) = n_{i}(t)/t$ with $t > 0$ and \textit{i} $ \in $ $\lbrace 0,1 \rbrace$
-\end{center}
-
-% ---- PERIOD ---- %
-\item \textit{Period} We can also define the period in terms of the average move rate:
-\begin{center}
-$\delta_{i}=1/\alpha_{i}$
-\end{center}
-
-% ---- WHO PLAYED LAST ---- %
-\item \textit{Who played last} We know who played last by taking the modulo with the period.
-$Z_{i}$ represents the time since the last flip of player i.
-We can also denote the time since the last flip of player i on resource r by $Z_{i}^{N}$.
-For a non adaptive game, period deterministic: At time $t=n$ is $Z_{i} = n mod i$.  \todo{er kan nog steeds tegelijk geflipt zijn maar dan hebben ze wel geflipt}.
-
-% ---- COST ---- %
-\item \textit{Cost} The cost is an important property of the game.
-In FlipIt for every player the cost of a move is denoted by $k_{i}$.
-These costs can be very different for every player.
-In this game we denote the players flipping cost for resource $V_{N}$ by $c_{i}^{V_{N}}$.
-\\
-For the defender the cost will be either the cost of flipping every resource or the cost of flipping a subgroup of the resources.\\
-For the attacker the cost will be the cost of dropping a virus on a node. The spreading of the virus will not imply an extra cost.
-
-% ---- UTILITY/ GAIN ---- %
-\item \textit{Utility} In FlipIt the Gain definition is the utility function.
-The Gain denotes the total time a player i has gained control over a resource.
-\todo{nu gain van een resource, moet voor verschillende resources zijn}
-The Gain $G_{i}$ denotes players \textit{i} total gain of a game, which is the total time the player has gained control over a subset of resources thus controlling the game.
-This is denoted by the following:\\
-\begin{center}
-$G_{i}(t) = \int^{t}_{0} C_{i}(x) dx$ 
-\end{center}
- If we sum up the total Gain of the attacker and the defender we end up with the time:
-\begin{center}
-$G_{1}(t) + G_{0}(t) = t$
-\end{center}
-
-% ---- AVERAGE GAIM RATE ---- %
-\item \textit{Average gain rate} The average gain rate for player i is defined as
-\begin{center}
-$\gamma_{i}(t)= G_{i}(t)/t$
-\end{center}
-
-\subsection{Formal definition}
-\begin{description}
-\item \textit{Graph Matrix} We represent the graph of the network through a matrix $ A = |V| \times |V|$.
-The (i,j)-entry of the matrix A will have a 1 if there is a connection between node $V_{i}$ and node $V_{j}$.
-If we are working with an undirected graph the matrix will be symmetric.
-\item \textit{Attack Vector} We denote $X = 1 \times |V|$ as the attack vector.
-It will be a vector with only zeros.
-The attacker will place a virus on a node V.
-This will be denoted by the Vth entry in the vector that is changed by a 1.
-\item \textit{Reset vector} The reset vector will make sure that the right entries in the matrix become zero. If the defender flips every node every time it flips then the attack vector will be 0.
-\item \textit{Cummulative Matrix} This matrix will keep record of the propagation of the virus through the network.
-\item \textit{State Matrix} The State matrix $T(t) = 1 \times |V| $ will keep at every time t the state of the game and denote which node at time t is infected with the virus.
-At time $t=0$ the State Matrix will be the null matrix.
-\end{description}
-De eerste infectie is de attack vector * Graph matrix . 
-\end{description}
-
-\section{Conclusion}
-The final section of the chapter gives an overview of the important results
-of this chapter. This implies that the introductory chapter and the
-concluding chapter don't need a conclusion.
-
 %%% Local Variables: 

### 2015-05-07 doc/template/chap-3.tex
extra toevoegingen

@@
 
-In this section we discuss the extentions that can be made on the FlipIt game.
+In this section we discuss the extentions that can be made on the FlipIt game. \\
 
@@
 In the AND model the attacker only controls the system if he controls all the resources of the system, whereas in the OR model the attacker only needs to compromise one resource to be in control of the entire system.
+\\
 %The difference with FlipThem and this paper is that we introduce a Graph Model in the beginning.\\
 Another addition of Laszka et al.
-to the game of FlipIt is extending the game to also consider non-targeted attacks by non-strategic players.
+to the game of FlipIt [mitigating covert compromises] is extending the game to also consider non-targeted attacks by non-strategic players.
 In this game the defender tries to maintain control over the resource that is subjected to both targeted and non-targeted attacks.
-
-A last important addition from Laszka et al.
-to the game of FlipIt is   The moves made by the attacker are still covert but the moves made by the defender are known to the attacker.
+Non-targeted attacks can include phishing, while targeted attacks may include threats delivered through zero day attack vulnerabilities.
+\\
+One of the last important addition from Laszka et al.
+[Mitigation of Targeted and] is to consider a game where the moves made by the attacker are still covert but the moves made by the defender are known to the attacker.
 This means that the attacker can base his attacks on the defender's moves.
@@
 The conclusion of this paper is that the optimal strategy for the defender is moving periodically.
+\\
 
-on FlipIt is done by Pham\cite{GameTheorApprCostBenefitAnalyses} [\todo{citatie needed voor Are We Compromised?}].
-Beside the action Flip their is another action Test.
+Other authors used the FlipIt game to apply it on a specific scenario. To be able to use the FlipIt game, modifications where required for the FlipIt model.
+One of the scenarios by Pham\cite{GameTheorApprCostBenefitAnalyses} [\todo{citatie needed voor Are We Compromised?}] was to find out whether a resource was compromised or not by the attacker.
+This could be verified by the defender, who has an extra move "test" beside the flip move.
 The basic idea is to test with an extra action if the resource has been compromised or not.
-This action involves also an extra cost.
-This model is useful if somebody wants to know for example if his or her password has been compromised or wants to assess the periodic security of a system.
-In \cite{MitigationCovert} \cite{MitigationNonTargeted} Laszka et al.
-they also consider non targeted attacks by non-strategic players and \todo{verder aanvullen}.
+This move involves also an extra cost.
+This model is useful if somebody wants to know for example if his or her password has been compromised.
 \\
+
+Finally researchers also have investigated the behavioural of humans when playing FlipIt.
+A Nochenson and Grossklags [A behavioural investigation of the FlipIt game]  investigate how people really act when given temporal decisions.
+[Risk-seeking in a continuous game of timing] Reitter et al.
+they observe continuous games, 20-seconds FlipIt game..
 
@@
 
-Company networks are targeted.
-It costs a lot.
-They want to defend their company networks.
-No loss of data, integrity and confidentiality.
-Many ways to attack a company.
-Viruses, Trojans, worms, DOS, ..
-Hard to protect against every attacker.
-
 %%% Local Variables: 

### 2015-05-07 doc/template/chap-10.tex
begin introductie next

@@
-\chapter{FlipIt game with virus propagation}
-\label{cha:1}
+\chapter{Introduction}
+\label{cha:10}
 %\documentclass[10pt]{article}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%			Introduction Chapter 1				%%%%%%
+%%%%%			Introduction Chapter 10			%%%%%%
 %%%%%												%%%%%%
@@
 
- \subsection{Actions of the attacker}
-A virus has different kind of ways of making his way through a company network.
-We will describe the different ways of how the virus can propagate.
-For start we will say that the virus or worm will be dropped on Node i and that it has k numbers of neighbours.
-\begin{enumerate}
-\item Node i is infected and will spread the virus or worm to every k neighbours and will stop infecting the neighbours in the next step
-\item Node i is infected and will spread the virus or worm to every k neighbours and will keep on spreading the virus to the same neighbours in every next step
-\item Node i is infected and will spread the virus to only one of the k neighbours and will stop infecting another neighbour in the next step
-\item Node i is infected and will spread the virus to only one of the k neighbours and in the next step it will infect another one of the k neighbours 
-\end{enumerate}
+\section{Introduction}
 
-In the game that will be modelled in the paper we will use the settings of the first spreading method.
-We will not use method 2 because this kind of propagation will float the network.
-Because we use the settings of a mail system and contact in a mailing list the method of 3 and 4 are not used.
-\\
-In the first method the node that has been infected can be again infected.
-If one of the neighbours infects the node again the node will infect his neighbours again.
-By using this spreading method we have three distinct states in which a node can be situated.
-An \textit{infected state}, a \textit{clean state} and a \textit{spreading state}.
-An infected state means that the node is infected and will not spread the virus to its neighbours, a clean state means that the node is not infected on that moment and a spreading state means that the node is infected and that it will spread the virus or worm to its neighbours in the next step.
-We can argument this kind of propagation through a mail worm. \todo{voorbeeld geven van zo een worm}
-%Another propagation method is that the virus works as a token. It will propagate to only one neighbour and continue to spread.
+%  Situation
+%  Complication
+%  Question
+%  Answer
 
-The Attacker itself has two different ways of attacking the company network.
-It will only infected one node of the network and will wait for the virus to spread itself through the network.
-We will model two ways of attacks of an Attacker:
-\begin{enumerate}
-\item The attacker drops the virus on a random node on the network
-\item The attacker drops the virus on a targeted node on the network
-\end{enumerate}
-The attacker in this game will put a virus or worm on one of the nodes in the network.
-(This will happen at random.) The attacker does not know on which node the virus will be dropped.
-We will use this randomness because \todo{feit uit security rapport symantec} most viruses are spread via a usb stick or a shared resource.
-If we use this spreading method where we have a targeted attack the attacker will have more information about the network.
-\\
+% woorden die in introductie moeten voorkomen: Game theory, cybersecurity, APT, FlipIt, virus propagation
 
-The attacker can choose at which rate it will drop a virus on one of the nodes on the network.
-The cost of dropping a virus will be the same.
-It will not increase.
-If it will increase this means that the attacker will eventually drop out of the game because it becomes to expensive.\\
-The attacker is in control over the game if it manages to infect a subset of all the resources of the company network.
+% -- Situation -- %
+In this era where digitalization becomes prominent in every aspect of our lives, where technology is growing fast and where business are always under attack, security becomes an issue of increasing complexity.
+Without security, their is no protection to keep somebody out of a system.
+It is the same as leaving the door of your house open for everybody to come in.
+It is important to keep a system secure.
+\todo{waarom?
+:  lekken van informatie, DOSS attacks, ..} A hacker will be a person that seeks exploits or weaknesses in a system or network, to be able to gain access.
+Many of those attacks have a different cause.
+Some of the attacks by a hacker can be benign, others can be harmful.
+Their are variant ways to break into a system.
 
-\subsection{Actions of the defender}
-The attacker wants to protect all the nodes of his network.
-It can do so by getting back control over the resources.
-We will assume that the defender of the network has knowledge over his own network.
-Which is convenient in the real world because a company has to know how his infrastructure looks like.\\
+% -- Complication -- %
+It is difficult to protect a system or network against all this variant threats. 
+Game theory is more and more used to model Cyber Security.
+Game theory analyses in this case a game where the interactions are between an attacker and a defender of a system, where both players have to make decisions.
+Both players want to ..
+%-- Question -- %
+In this paper we want to focus on situations where a computer network is attacked by APT. These threats will infect the network by means of a virus that will propagate through the network.
 
-The defender has two possible ways of defending its network:
-\begin{enumerate}
-\item The defender flips all the nodes of his network
-\item The defender will flip a subset of the nodes of his network
-\end{enumerate}
+% -- Anwer -- %
+We propose an addition to the basic FlipIt model to model a scenario where the moves by the attacker will not be instantaneous. Next we analyse what the new Nash equilibria will be and ..
 
-The cost of flipping all the nods of the network will be greater than the cost of flipping a subset of nodes.
-We make this assumption because otherwise it will be beneficial for the defender to always flip all the nodes in the network.\\
-
-We will also make the assumption that as a defender flips a node the node can get infected again.
-A flip will not be  correlated to a patch but to a clean-up.
-\todo{waarom geen patch, wormen kunnen veranderen gaandeweg}
-\todo{andere mogelijkheid:} Another setting of the game can be that the flip of the defender is equal to a patch and that the resource cannot be infected any more.
-But with this case we deviate from the flipIt game, because the attacker cannot flip the resource any more.
-Unless we work with different virusses every time the attacker flips.
-We start with the less complex game of flipping is equal to a clean-up.
-
-\subsection{Strategies of both players}
-We explained what the actions of each player are.
-
-\section{Formal definition Game}
-
-In this section we provide a formal definition of the game and the notation that we will use throughout the paper.
-
-\begin{description}
-% ---- PLAYERS ---- %
-\item \textit{Players}  There are two players in the game, one is the defender and the other one is the attacker. They are respectively identified by 0 and 1.
-
-% ---- TIME ---- %
-\item \textit{Time}  The game starts at $t=0$ and continuous indefinitely as $t \rightarrow \infty$. The game is a continuous game.
-
-% ---- GRAPH ---- %
-\item \textit{Graph} We represent the company network as a Graph $G = < V,E>$.
-G is an ordered pair where V denotes the set of resources or nodes in the network and E denotes the set of connections or links, which are a two-element subset of V.
-We use the notations resources and nodes interleaving in this paper.\\
-We have N resources in the network. $N \in $  \todo{aanvullen}. This means we can denote the resources by:
-\begin{center}
-$V \in {V_{0}, V_{1}, V_{2}, ... , V_{N} }$
-\end{center}
-The set E of connections indicates if there is a link between two resources.
-We see the links as bidirectional so the total graph is undirected.
-If there is a link between resource $V_{n}$ and $V_{n+1}$ then there is also a link between $V_{n+1}$ and $V_{n}$.
-
-% ---- GAME STATE ---- %
-\item \textit{Game State} There is also a time-dependent variable that represents the state of the game.
-$C=C(t)$ is either 0 if the game is under control by the defender and 1 if the Game is under control by the attacker.
-\\
-We start at $t=0$ with the defender who has control over the game.
-We do this because we assume that the defender will only put the network online without having a virus or worm in it.
-The Attacker can gain control over the game when it compromises a subset \textit{s} of the resources.
-The subset \textit{s} is a minimum of 1 resource and a maximum of all the resources N.
-\\
-\todo{deze variabele nodig ja of nee ?
-JA} We can also define the state of each resource by $C^{A}_{N}$ and $C^{D}_{N}$.
-If $C^{A}_{N} = 1$ then this means that the attacker has control over the resource, and 0 otherwise.
-For $C^{D}_{N}$ it is visa versa, $C^{D}_{N} = 1 - C^{A}_{N}$.\\
-
-% ---- MOVES ---- %
-\item \textit{Moves} Both players can make a move in the game.
-Moves done in a finite numbers of time in any finite time interval.
-Both players can play at any time they want, they can also play at the same time.
-If this happens the one that has control over the resource will keep having control over the resource.
-This makes the game fully symmetric \todo{beter uitleggen}. The sequence of move times are denoted by the following infinite sequence:
-\begin{center}
-$t=t_{1},t_{2},t_{3},..$
-\end{center}
-Two move times can be the same because we allow players to move at the same time.
-We can also denote the infinite sequence of times when player \textit{i}  moves. We write this as :
-\begin{center}
-$t=t_{i,1},t_{i,2},t_{i,3},..$ with \textit{i} $ \in $ $\lbrace 0,1 \rbrace$
-\end{center}
-The sequences $t_{1}$ and $t_{0}$ are disjoint subsets of the sequent t. 
-We can also denote who made the \textit{k}th move by defining a sequence \textit{p} that denotes the sequence of who played:
-\begin{center}
-$p=p_{1},p_{2},p_{3}, .. $ with $p_{k}$ $\in$ $\lbrace 0,1 \rbrace$
-\end{center}
-
-% ---- NUMVER OF MOVES ---- %
-\item \textit{Number of moves}  $n_{i}(t$ denotes the number of moves made by player \textit{i} up to and including time t. This means that 
-\begin{center}
-$n(t)=n_{1}(t) + n_{0}(t)$
-\end{center}
-is the sum of the number of moves made by the defender and the attacker up to and including time t.
-
-% ---- AVERAGE MOVE RATE ---- %
-\item \textit{Average move rate} We denote $\alpha_{i}(t)$ as the average move rate by player i:
-\begin{center}
-$\alpha_{i}(t) = n_{i}(t)/t$ with $t > 0$ and \textit{i} $ \in $ $\lbrace 0,1 \rbrace$
-\end{center}
-
-% ---- PERIOD ---- %
-\item \textit{Period} We can also define the period in terms of the average move rate:
-\begin{center}
-$\delta_{i}=1/\alpha_{i}$
-\end{center}
-
-% ---- WHO PLAYED LAST ---- %
-\item \textit{Who played last} We know who played last by taking the modulo with the period.
-$Z_{i}$ represents the time since the last flip of player i.
-We can also denote the time since the last flip of player i on resource r by $Z_{i}^{N}$.
-For a non adaptive game, period deterministic: At time $t=n$ is $Z_{i} = n mod i$.  \todo{er kan nog steeds tegelijk geflipt zijn maar dan hebben ze wel geflipt}.
-
-% ---- COST ---- %
-\item \textit{Cost} The cost is an important property of the game.
-In FlipIt for every player the cost of a move is denoted by $k_{i}$.
-These costs can be very different for every player.
-In this game we denote the players flipping cost for resource $V_{N}$ by $c_{i}^{V_{N}}$.
-\\
-For the defender the cost will be either the cost of flipping every resource or the cost of flipping a subgroup of the resources.\\
-For the attacker the cost will be the cost of dropping a virus on a node. The spreading of the virus will not imply an extra cost.
-
-% ---- UTILITY/ GAIN ---- %
-\item \textit{Utility} In FlipIt the Gain definition is the utility function.
-The Gain denotes the total time a player i has gained control over a resource.
-\todo{nu gain van een resource, moet voor verschillende resources zijn}
-The Gain $G_{i}$ denotes players \textit{i} total gain of a game, which is the total time the player has gained control over a subset of resources thus controlling the game.
-This is denoted by the following:\\
-\begin{center}
-$G_{i}(t) = \int^{t}_{0} C_{i}(x) dx$ 
-\end{center}
- If we sum up the total Gain of the attacker and the defender we end up with the time:
-\begin{center}
-$G_{1}(t) + G_{0}(t) = t$
-\end{center}
-
-% ---- AVERAGE GAIM RATE ---- %
-\item \textit{Average gain rate} The average gain rate for player i is defined as
-\begin{center}
-$\gamma_{i}(t)= G_{i}(t)/t$
-\end{center}
-
-\subsection{Formal definition}
-\begin{description}
-\item \textit{Graph Matrix} We represent the graph of the network through a matrix $ A = |V| \times |V|$.
-The (i,j)-entry of the matrix A will have a 1 if there is a connection between node $V_{i}$ and node $V_{j}$.
-If we are working with an undirected graph the matrix will be symmetric.
-\item \textit{Attack Vector} We denote $X = 1 \times |V|$ as the attack vector.
-It will be a vector with only zeros.
-The attacker will place a virus on a node V.
-This will be denoted by the Vth entry in the vector that is changed by a 1.
-\item \textit{Reset vector} The reset vector will make sure that the right entries in the matrix become zero. If the defender flips every node every time it flips then the attack vector will be 0.
-\item \textit{Cummulative Matrix} This matrix will keep record of the propagation of the virus through the network.
-\item \textit{State Matrix} The State matrix $T(t) = 1 \times |V| $ will keep at every time t the state of the game and denote which node at time t is infected with the virus.
-At time $t=0$ the State Matrix will be the null matrix.
-\end{description}
-De eerste infectie is de attack vector * Graph matrix . 
-\end{description}
-
-\section{Conclusion}
-The final section of the chapter gives an overview of the important results
-of this chapter. This implies that the introductory chapter and the
-concluding chapter don't need a conclusion.
-
-%
-%In the following paragraph an introduction to game theory is given based on the work of \cite{leyton2008essentials} and \cite{Coursera}.
-For a more detailed and full introduction to game theory, the reader is referred to \cite{leyton2008essentials}.
-%%------------------------------------------------%
-%%            Intro Game Theory 					 %
-%%------------------------------------------------%
-%\section{Intro Game Theory}
-%\label{Cha:1:Intro.Game.Theory}
-%
-%%begin over dat gametheorie handig is in de economie
-%Game theory studies the interaction between independent and self-interested agents.
-It is a mathematical way of modelling the interactions between two or more agents where the outcomes depend on what everybody does and how it should be structured to lead to good outcomes.
-For this reason it is very important for economics and also for politics, biology, computer science, philosophy and a variety of other disciplines.
-\\
-%%Every agent has different levels of happiness for the different outcomes.
-%%self interested meaning
-%
-%One of the assumptions underlying game theory is that the players of the game, the agents, are independent and self-interested.
-This does not necessarily mean that they want to harm other agents or that they only care about themselves.
-%%utility function meaning 
-%Instead it means that each agent has preferences about the states of the world he likes.
-These preferences are mapped to natural numbers and are called the utility function.
-The numbers are interpreted as a mathematical measure to tell you how much an agent likes or dislikes the states of the world.
-\\
-%It also explains the impact of uncertainty.
-When an agent is uncertain about a distribution of outcomes, his utility will describe the expected value of the utility function with respect to the probability of the distribution of the outcomes.
-For example: with 0.7 probability it will be 7 degrees outside and with 0.3 probability it will be 10 degrees.
-The agent can have a different opinion about that distribution versus another distribution.
-(\todo{uitleggen aan de hand van een voorbeeld}).\\
-%%Cooperative and non cooperative games
-%In a decision game theoretic approach an agent will try to act in such a way to maximise his expected or average utility function.
-It becomes more complicated when two or more agents want to maximise their utility and whose actions can affect each other utilities.
-This kind of games are referred to as non cooperative game theory, where the basic modelling unit is the group of agents.
-The individualistic approach, where the basic modelling is only one agent, is referred as cooperative game theory.
-%
-%There are two standard representations for games. The first one is the Normal Form. The second one is the Extensive Form.
-%
-%In the following lists a couple of terms that will be used throughout the paper.
-%\begin{description}
-%\item \textit{Players}: players are referred as the ones who are the decision makers. It can be a person, a company or an animal.
-%\item \textit{Actions}: actions are what the player can do. 
-%\item \textit{Outcomes}:  
-%\item \textit{Utility function}: the utility function is the mapping of the level of happiness of an agent about the state of the world to natural numbers.
-%\item \textit{Strategies}: A strategy is the combination of different actions. A pure strategy is only one action.
-%\end{description}
-%
-%A game in game theory consists of multiple agents and every agent has a set of actions that he can play. 
-%
-%
-%
-%%Nash equilibrium
-%
-%
-%
-%% --------------- example of a game -----------------%
-%
-%
-%%------------------------------------------------%
-%%            Intro about virusses				 %
-%%------------------------------------------------%
-%\section{Virusses}
-%
-%Many network security threats today are spread over the Internet. The most common include:
-%
-%Viruses, worms, and Trojan horses
-%Spyware and adware
-%Zero-day attacks, also called zero-hour attacks
-%Hacker attacks
-%Denial of service attacks
-%Data interception and theft
-%Identity theft
-%
-%%http://www.ists.dartmouth.edu/library/258.pdf Email Virus Propagation Modeling and Analysis
-%%Cliff C. Zou∗, Don Towsley†, Weibo Gong∗
-%%∗Department of Electrical & Computer Engineering
-%%†Department of Computer Science
-%%Univ. Massachusetts, Amherst
-%%Technical Report: TR-CSE-03-04
-%
-%Computer virus through mail. 
-%Though virus spreading through email is an old technique, it is still effective and is widely used by
-%current viruses and worms. Sending viruses through email has some advantages that are attractive to
-%virus writers:
-% Sending viruses through email does not require any security holes in computer operating systems
-%or software.
-% Almost everyone who uses computers uses email service.
-% A large number of users have little knowledge of email viruses and trust most email they receive,
-%especially email from their friends [28][29].
-% Email are private properties like post office letters. Thus correspondent laws or policies are required
-%to permit checking email content for detecting viruses before end users receive email [18].
-%
-%Send a email with malicious attachment.
-Only again infected if attachment again opened.
-Thus this is the action of attacking every neighbour node + also can attack again the node where the virus was coming from.
-%There are also email viruses were the malicious program is hidden in the txt and the attachment does not need to be opened. 
-%
-%%http://www.cisco.com/web/offer/gist_ty2_asset/Cisco_2014_ASR.pdf p49
-%%http://repo.hackerzvoice.net/depot_madchat/vxdevl/papers/avers/2004-35.pdf
-%%http://www.mcafee.com/us/resources/white-papers/foundstone/wp-managing-malware-outbreak.pdf
-%
-%
-%
-%\subsection{Malware}
-%%Does a company network faces lot of malware? what is the cost ?
-%Relevant researches:
-%\begin{itemize}
-%%http://ants.iis.sinica.edu.tw/3BkMJ9lTeWXTSrrvNoKNFDxRm3zFwRR/17/04483668.pdf
-%\item How Viruses and worm can be detected. Difference between UDP en TCP worm propagation. Difference for the propagation speed. 
-%\end{itemize}
-%
-%
-%
-%Purpose thesis: model a worm propagation with adaptations of Flip-It.
-Flip-It cannot address the evaluation of individual nodes.
-Flip-It with multiple resources has not addressed the fact that a virus does not need to compromise the whole network.
-A subpart of the network can already cause problems for the company.
-Data leackages.
-%So the virus can propagate 
-%
-%
-%=======
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%			Conclusion Chapter 1				%%%%%%
-%%%%%%												%%%%%%
-%%%%%%												%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%\section{Conclusion}
-%\label{Cha:1:Conclusion}
-%The final section of the chapter gives an overview of the important results
-%of this chapter. This implies that the introductory chapter and the
-%concluding chapter don't need a conclusion.
-%
-%
 %%%% Local Variables: 
@@
 %%%% End:
-
-%\end{document}

### 2015-05-07 doc/template/intro.tex
latex file compileert niet met chap 5
begin introductie next

@@
 (We live in an era) In this era where digitalization becomes prominent in every aspect of our lives, where technology is growing fast and where business are always under attack, security becomes an issue of increasing complexity.
-Since 2009, the number of reported security attacks has increased 66\%, year over year.
+Since 2009, the number of reported security attacks on companies has increased 66\%, year over year.
 \todo{security report van pwc}.
@@
 Defending a network of a company is not an easy job.
-To prevent intruders it can make use of firewalls, routers, IDS systems, virus scans, and other defence mechanisms.
+Malicious people will try to  To prevent intruders it can make use of firewalls, routers, IDS systems, virus scans, and other defence mechanisms.
 Unfortunately technology is growing fast and attacks are getting more sophisticated and the causes of these attacks can be very different.
@@
 In this paper we model a company network through multiple shared resources and a flip from the attacker that drops a virus that will spread itself autonomously. We show that ...
+
+%\begin{minipage}{0.5\textwidth}
+%\includegraphics{Images/killchain.jpg}
+%\end{minipage} \hfill
+%\begin{minipage}{0.45\textwidth}
+%\begin{itemize}
+%\item *Rectangle
+%\item *Color: blue
+%\end{itemize}
+%\end{minipage}
+
 %http://www.pwc.com/gx/en/consulting-services/information-security-survey/index.jhtml

### 2015-05-07 doc/template/chap-10.tex
verder intro aangevuld

@@
 Their are variant ways to break into a system.
+Virusses, worms, spyware and other malware are nr 2 for the top external threats [security report kaspersky 2014].
+Furthermore these kind of threat also causes the greatest percentage in loss of data.
+These threats will infect the network by means of a virus that will propagate through the network.
 
 % -- Complication -- %
-It is difficult to protect a system or network against all this variant threats. 
+It is difficult to protect a system or a network against APT. 
 Game theory is more and more used to model Cyber Security.
@@
 Both players want to ..
+Researchers at RSA made a game theoretic framework to model targeted attacks.
+Specific for a scenario where a system or network is taken over completely by an attacker repeatedly and this attack is not immediately detected by the defender of the system or network.
+This game theoretic game "flipit" is a two players game where the attacker and the defender are competing to get control over a shared resource.
+Both players do not know who is currently in control of the resource until they move.
+In FlipIt every move gives them immediately control over the resource.
+But what if the attacker moves and it takes a while before the attacker gets full control over the resource?
+FlipIt does not take into account that a move may not be instantaneous, but has a certain delay.
+Take for example as a resource a network with different nodes ( laptops, datacenters).
+The defender drops a virus on one of the nodes and wait till this virus infects the whole network.
+The defender will only be in control of the resource if the whole network is infected.
+This an lead us to the following resaurch questions:
+Gaat er een specieke grootte zijn van een delay waarbij de attacker al weet dat hem niet meer moet gaan spelen ? ( is niet gelijk aan de grootte van de periode van de attacker)
+%Specific research question:
+%-         Is it possible to incorporate the notion of delay in the game-theoretical analysis of the Flip-It game ?
+%
+%-        Can we calculate a Nash Equilibrium for a FlipIt Game with delay ?
+%
+%-        Does this allow us to determine an optimal defense strategy against an attacker ?
+When working with a network and a delay we can .. graph model en uitleggen hoe we de graph kunnen maken zodat de delay altijd zo groot mogelijk gaat zijn.
+
 %-- Question -- %
-In this paper we want to focus on situations where a computer network is attacked by APT. These threats will infect the network by means of a virus that will propagate through the network.
+In this paper we want to focus on situations where a computer network is attacked by APT.
 
@@
 We propose an addition to the basic FlipIt model to model a scenario where the moves by the attacker will not be instantaneous. Next we analyse what the new Nash equilibria will be and ..
+
+%In this paper we want to focus on situations where a computer network is attacked by an APT. These threats will infect the network by means of a virus that will propagate through the network. 
+%ð  Dat is geen vraag.
+%
+% 
+%Kijk naar de opdracht van je thesis. De vraag zou bv. kunnen zijn:
+%General research question: Does the application of game theory help in defining a good defense (or attacker) strategy ?
+%This question is to large, so we reduce the question to …
+%Specific research question:
+%-         Is it possible to incorporate the notion of delay in the game-theoretical analysis of the Flip-It game ?
+%
+%-        Can we calculate a Nash Equilibrium for a FlipIt Game with delay ?
+%
+%-        Does this allow us to determine an optimal defense strategy against an attacker ?
+%
+% 
+%Je matrix-berekening beantwoordt de vraag:
+%-        How can we calcuate the expected duration for a node’s infection/the entire network infection ?
+%
+%-        Can we calculate this node per node ?
+%
+% 
+%Complication
+%Wat je schrijft is veel te algemeen. Maak het heel concreet en toegespitst op jouw specifieke bijdrage ?
+%ð  De complicatie is dat de FlipIt game geen rekening houdt met de delay
 

### 2015-05-07 doc/template/chap-9.tex
ales even updaten, related work
verder intro aangevuld

@@
 This corresponds to a periodic defender strategy.
-To simplify the analysis in a first time, a periodic attacker strategy is assumed as well.
+\todo{bedrijven hebben niet het geld en de mankracht om zich daar mee bezig te houden en dan is periodisch verdedigen het handigste, bedrijven willen alles zo efficient en met de minste kost doen} To simplify the analysis in a first time, a periodic attacker strategy is assumed as well.
 Further research can investigate the effect of relaxing this assumption.\\
@@
 \end{equation}
+
+\textbf{Case 2:} $\delta_{A} \leq \delta_{D} $ (The attacker plays at least as fast as the defender.) \\
+
+Let $r = \dfrac{\delta_{D}}{ \delta_{A} }$.
+The intervals between two consecutive attacker's moves have length $\delta_{A}$.
+Consider a given attackers move interval.
+The probability over the attacker's phase selection that the defender moves in this interval is $\dfrac{\delta_{D}}{ \delta_{A} } = (1/r)$.
+Given that the defender moves within the interval, he moves exactly once within the interval (since$\delta_{A} \leq \delta_{D} $) and his move is distributed uniformly at random.
+\\
 

### 2015-05-08 doc/template/chap-10.tex
committen

@@
 % -- Situation -- %
-In this era where digitalization becomes prominent in every aspect of our lives, where technology is growing fast and where business are always under attack, security becomes an issue of increasing complexity.
-Without security, their is no protection to keep somebody out of a system.
-It is the same as leaving the door of your house open for everybody to come in.
-It is important to keep a system secure.
+In this era where digitalization becomes prominent in every aspect of our lives, where technology is growing fast and where businesses are always under attack, security becomes an issue of increasing complexity.
+Without security, there is no protection to keep somebody out of a system.
+It is the same as leaving the door of your house wide open for everyone to come in.
+Why is it so important to keep a system secure?
 \todo{waarom?
-:  lekken van informatie, DOSS attacks, ..} A hacker will be a person that seeks exploits or weaknesses in a system or network, to be able to gain access.
+:  lekken van informatie: je hebt info over klanten en moet hun privacy beschermen, je eigen data is geld waard voor concurrenten, DOSS attacks: je will je service-level agreements kunnen nakomen, ..} A hacker will be a person that seeks exploits or weaknesses in a system or network in order to gain access.
 Many of those attacks have a different cause.
 Some of the attacks by a hacker can be benign, others can be harmful.
-Their are variant ways to break into a system.
-Virusses, worms, spyware and other malware are nr 2 for the top external threats [security report kaspersky 2014].
-Furthermore these kind of threat also causes the greatest percentage in loss of data.
+There are various ways to break into a system.
+Virusses, worms, spyware and other malware are the number two of the top external threats [security report kaspersky 2014].
+Furthermore these kind of threats also causes the greatest percentage in loss of data.
 These threats will infect the network by means of a virus that will propagate through the network.
+\todo{APT uitleggen ? en dat het moeilijk is om een systeem te beschermen tegen APT ==> overgang naar de volgende paragraaf maken.}
 
 % -- Complication -- %
-It is difficult to protect a system or a network against APT. 
-Game theory is more and more used to model Cyber Security.
-Game theory analyses in this case a game where the interactions are between an attacker and a defender of a system, where both players have to make decisions.
-Both players want to ..
+Since it is so difficult to protect a system or a network against APT, researcher have been looking for effective ways to predict in advance which defense strategy might be the better one. 
+Game theory is gaining more and more interest as an effective technique to  model and study Cyber Security.
+Game theory analyses the security problem as a game where the players are an attacker and a defender of a system, and where both players have to make decisions.
+In particular, both players will aim for the strategy that results in a maximal benefit for them.
 Researchers at RSA made a game theoretic framework to model targeted attacks.
-Specific for a scenario where a system or network is taken over completely by an attacker repeatedly and this attack is not immediately detected by the defender of the system or network.
-This game theoretic game "flipit" is a two players game where the attacker and the defender are competing to get control over a shared resource.
+They study the specific scenario where a system or network is repeatedly taken over completely by an attacker and this attack is not immediately detected by the defender of the system or network.
+In game theory, such a game is known as ''FlipI''.
+This is a two players game where the attacker and the defender are competing to get control over a shared resource.
 Both players do not know who is currently in control of the resource until they move.
@@
 FlipIt does not take into account that a move may not be instantaneous, but has a certain delay.
-Take for example as a resource a network with different nodes ( laptops, datacenters).
-The defender drops a virus on one of the nodes and wait till this virus infects the whole network.
-The defender will only be in control of the resource if the whole network is infected.
-This an lead us to the following resaurch questions:
-Gaat er een specieke grootte zijn van een delay waarbij de attacker al weet dat hem niet meer moet gaan spelen ? ( is niet gelijk aan de grootte van de periode van de attacker)
+Consider for example a network with different nodes ( laptops, datacenters) as a resource.
+The attacker drops a virus on one of the nodes and then wait till this virus infects the whole network.
+The attacker will only be in control of the resource once the whole network is infected.
+
+The game theoretical approach of the FlipIt does not take such delay into account. This an lead us to the following research questions:
+- Is it possible to incorporate the notion of delay in the game-theoretical analysis of the Flip-It game ?
+
 %Specific research question:
@@
 %
-%-        Does this allow us to determine an optimal defense strategy against an attacker ?
+%-        
+- Does this allow us to determine an optimal defense strategy against an attacker ?
+for example: Gaat er een specieke grootte zijn van een delay waarbij de attacker al weet dat hem niet meer moet gaan spelen ?
+( is niet gelijk aan de grootte van de periode van de attacker)
 When working with a network and a delay we can .. graph model en uitleggen hoe we de graph kunnen maken zodat de delay altijd zo groot mogelijk gaat zijn.
-
-%-- Question -- %
-In this paper we want to focus on situations where a computer network is attacked by APT.
 

### 2015-05-10 doc/template/chap-5.tex
ales even updaten, related work
FlipIt strategien

@@
 
-In the following paragraph an introduction to game theory is given based on the work of 
-{leyton2008essentials} 
-and 
-{Coursera}. 
-For a more detailed and full introduction to game theory, the reader is referred to 
-{leyton2008essentials}. 
+Gametheory is a mathematical study to analyse interactions between independent and self-interested agents.
+To get an understanding of the most important concepts of game theory, a short introduction based on the work of
+\citep{leyton2008essentials} and \todo{Coursera} is given in section 2.1 [] \todo{ref}. For a more detailed and full introduction to game theory, the reader is referred to 
+\todo{leyton2008essentials}.  In section 2.2 [] \todo{ref} an overview of the FlipIt game is given with the definitions and concepts that will be used throughout the paper.
+The last section [] \todo{ref} will cover the extensions and additions already made on FlipIt.
+
 %------------------------------------------------%
@@
 %------------------------------------------------%
-\section{Intro Game Theory}
+\section{A brief introduction in Game Theory}
 \label{Cha:1:Intro.Game.Theory}
@@
 It is a mathematical way of modelling the interactions between two or more agents where the outcomes depend on what everybody does and how it should be structured to lead to good outcomes.
-For this reason it is very important for economics and also for politics, biology, computer science, philosophy and a variety of other disciplines.
+For this reason it can be very useful in economics and also in other branches as politics, biology, computer science, philosophy and a variety of other disciplines.
 \\
@@
 It becomes more complicated when two or more agents want to maximise their utility and whose actions can affect each other utilities.
-This kind of games are referred to as non cooperative game theory, where the basic modelling unit is the group of agents.
+This kind of games are referred to as non-cooperative game theory, where the basic modelling unit is the group of agents.
 The individualistic approach, where the basic modelling is only one agent, is referred as cooperative game theory.
 
-There are two standard representations for games. The first one is the Normal Form. The second one is the Extensive Form.
+%There are two standard representations for games. The first one is the Normal Form. The second one is the Extensive Form.
 
@@
 \begin{description}
-\item \textit{Players}: players are referred as the ones who are the decision makers. It can be a person, a company or an animal. 
-\item \textit{Actions}: actions are what the player can do. 
-\item \textit{Outcomes}:  
-\item \textit{Utility function}: the utility function is the mapping of the level of happiness of an agent about the state of the world to natural numbers.
+\item \textit{Players}: Players are referred as the ones who are the decision makers. It can be a person, a company or an animal.  (they will act rational )
+\item \textit{Actions}: Every player has actions that he or he can do. 
 \item \textit{Strategies}: A strategy is the combination of different actions. A pure strategy is only one action.
+\item \textit{Utility function}: The utility function is the mapping of the level of happiness of an agent about the state of the world to natural numbers.
+
 \end{description}
@@
 Over John Nash is een prachtige film
-\todo{Best response ook uitleggen?}
-\todo{Voorbeeld ook nog uitleggen?}
-One of the solution concepts in Game Theory for non-cooperative games is a Nash Equilibrium.
+\subsection{Best response and Nash Equilibrium}
+One of the solution concepts in Game Theory for non-cooperative games is a Nash Equilibrium that we will use in this paper.
 A Nash Equilibrium is a subset of outcomes that can be interesting to analyse a game.
@@
 In general we can say that a Nash Equilibrium is a stable strategy profile: each player is considered to know the equilibrium strategies of the other players and no player would want to change his own strategy if he knows the strategies of the other players.
+
+Formal defenition of a Nash Equilibrium:
+A strategy profile $s = (s1, . . . , sn)$ is a Nash equilibrium
+if, for all agents $i$, $si$ is a best response to $s-i$ .
+''Intuitively, a Nash equilibrium is a stable strategy profile: no agent would want to change
+his strategy if he knew what strategies the other agents were following.
+Wecan divide Nash equilibria into two categories, strict and weak, depending on whether
+or not every agent's strategy constitutes a unique best response to the other agents' strategies.''
+
+\todo{Nash beter uitleggen nog met best response erbij}
+
+POSTCONDITION: Uitgelegd: Strategien, acties, strategien, spelers, rationeel, Nash, best response
 %In game theory, the Nash equilibrium is a solution concept of a non-cooperative game involving two or more players, in which each player is assumed to know the equilibrium strategies of the other players, and no player has anything to gain by changing only their own strategy.[1] If each player has chosen a strategy and no player can benefit by changing strategies while the other players keep theirs unchanged, then the current set of strategy choices and the corresponding payoffs constitutes a Nash equilibrium.
@@
 
+\section{The FlipIt game}
+In this section, we introduce the ''stealthy takeover '' game \flip{FlipIt} \cite{FlipIt}.
+\flip{FlipIt} is a game introduced by van Dijk et al.
+First we explain the framework of FlipIt and introduce the most important formules that will be used throughout the paper.
+To understand how to model a FlipIt game with virus propagation it is important to get familiar with the concepts of the normal FlipIt game and it's notations.
+\\
+
+\flip{FlipIt} is a two-players game with a shared (single) resource that the players want to control as long as possible.
+The shared resource can be a password, a network or a secret key depending on the setting being modelled.
+In the rest of the paper we name the two players the attacker, denoted by the subscript \textit{A} and the defender, denoted by subscript \textit{D}.
+
+The game begins at $t=0$ and continuous indefinitely ($t \rightarrow \infty $).
+The time in the game can be viewed as being continuous, but a discrete time can also be viewed.
+To get control over the resource, the players \textit{i} can flip the resource at any given time.
+A flip will be regarded as a move from a player \textit{i}.
+Each move will imply a certain cost $k_{i}$ and the cost can vary for each player.
+Both players will try to minimize their cost.
+By adding a cost, it will prevent players to move to frequently.
+\\
+
+The unique feature of \flip{FlipIt} is that every move will happen in a stealthy way, meaning that the player has no clue that (his adversary) the other player has flipped the resource.
+For instance, the defender will not find out if the resource has already been compromised by the attacker, but he can only potentially know it after he flips the resource himself.
+The goal of the player is to maximize the time that he or she has control over the resource while minimizing total cost of the moves.
+A move can also result in a "wasted move", called a flop.
+It may happen that the resource was already under control by the defender.
+If the defender moves when he or she has already control over the resource, he or she would have wasted a move since it does not result in a change of ownership and a cost is involved.
+\\
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[scale=0.8]{Images/DefFlipit}
+\caption{A representation of a FlipIt game where both players are playing periodically and discrete.
+Every move or flip is indicated by a blue or orange circle.
+The attacker is the orange colour and plays with a period of $\delta_{A}=4$.
+The defender is the blue colour and plays with a period of $\delta_{D}=3$.The blue and orange rectangles represent the amount of time one of the players is in control of the resource.}
+\label{fig:FLipItDefault}
+\end{figure}
+\todo{verwijzen naar de figuur \ref{fig:FLipItDefault}}
+
+The state of the resource is denoted as a time independent variable $C=C_{i}(t)$. 
+$C_{D}(t)$ is either 1 if the game is under control by the defender and 0 if the game is under control by the attacker. For $C^{A}(t)$ it is visa versa, $C^{A}(t)= 1 - C^{D}(t)$.\\ \\
+The game starts with defender being in control of the game, $C_{D}(0)= 1$.
+
+The players receive a benefit equal to the time of units that they were in possession of the resource minus the cost of making their moves. The cost of a player \textit{i} is denoted by $k_{i}$. 
+The total gain of player \textit{i} is equal to the total amount of time that a player \textit{i} has owned the resource from the beginning of the game up to time \textit{t}.
+It is expressed as follows:
+\begin{equation}\label{first}
+G_{i}(t) = \int_0^1 \! C_{i}(x) dx.
+\end{equation}
+If we add up the gain of the defender and the gain of the attacker it should sum up to t:
+\begin{equation}\label{first}
+G_{D}(t) + G_{A}(t) = t
+\end{equation}
+The average gain of player \textit{i} is defined as:
+\begin{equation}\label{first}
+\gamma_{i}(t) = G_{i}(t)/t.
+\end{equation}
+And thus for all $t > 0$ :
+\begin{equation}\label{first}
+\gamma_{D}(t) + \gamma_{A}(t) = 1
+\end{equation}
+Let $\beta_{i}(t)$ denote player's \textit{i} average benefit upto time \textit{t}:
+\begin{equation}\label{first}
+\beta_{i}(t) = \gamma_{i}(t) - k_{i}\alpha_{i}.
+\end{equation}
+This is equal to the fraction of time the resource has been owned by player \textit{i}, minus the cost of making the moves.
+~$ \alpha_{i}$ defines the average move rate by player \textit{i} up to time \textit{t}.
+In a give game, the asymptotic benefit rate or simply benefit will be defined as the lim inf of the average benefit because time t will increase to infinity and the average benefit may not have limiting values.
+\[ \beta_{i}(t)  = \lim_{t \to \infty} inf \beta_{i}(t)  \]
+\\
+
+\subsubsection{strategies}
+Because the players move in a stealthy way, there are different types of feedback that a player can get while moving.
+These types of feedback can be divided into two groups of strategies.
+The non-adaptive strategies and the adaptive strategies.
+These are described in table \ref{table:Strategies}.
+\\
+
+If there is no need for feedback for both of the players, we say that we have a non-adaptive strategy.
+Because the player does not receive any feedback during the game it will play in the same manner against every opponent.
+They are not dependent on the opponents movements.
+This means that they can already generate the time sequence for all the moves in advance.
+But they can depend on some randomness because the non-adaptive strategies can be randomised.
+In this paper we will focus in the beginning on the non-adaptive strategies.
+Reasons behind this is that a player (defender or attacker) rarely knows what the strategies are of his opponent.
+An interesting subclass of the non-adaptive strategies is one where the time intervals between two consecutive moves are generated by a renewal process.
+Example of such a renewal strategy is the periodic strategy where the time between two consecutive moves of the players are a fixed interval.
+An exponential strategy is a renewal strategy in which the interval between two consecutive moves is exponentially distributed.
+\\
+
+-- Ik kan dit ook pas uitleggen als ik mijn FlipIt spel met virus propagatie uit de doeken doe -- \\
+We elaborate more about the periodic strategy because we will use this one to model our first extension.
+For more details about the other strategies of FlipIt we refer the reader to [the paper of FlipIt].
+
+\textit{periodic strategy}: A Periodic strategy is a non-adaptive renewal strategy where the time intervals between consecutive moves are a fixed period, denoted by $\delta$.
+Moreover it has a random phase, that is chosen uniformly and random in the interval $[0,\delta]$ for the first move.
+The average rate of play is denoted by $\alpha = \dfrac{1}{\delta}$.
+\\
+
+In the category Adaptive strategy there are two sub classes of strategies.
+The first one is the Last move (LM).
+In this class whenever a player flips it will find out the exact time that the opponent played the last time.
+In the second class the Full History (FH), whenever a player flips it will find out the whole history of the opponents move.
+If the opponent player plays with a renewal strategy the sub classes FH and LM collapse.
+\\
+ I%n this subsection we elaborate about the strategy of FlipIt . 
+%A renewal strategy is a non-adaptive strategy \\
+
+The game can be extended by the amount of information that a player receives. It can also be possible for a player to get information at the start of the game. Both interesting cases are:
+\begin{itemize}
+\item Rate-of-play (RP): The player finds out the exact rate of play of the opponent.
+\item Knowledge-of-strategy (KS): The player finds out the complete information of the strategy that the opponent is playing.
+\end{itemize}
+
+ \begin{table}
+ \centering
+ \begin{tabular}{ l | c  }
+  \textbf{Categories} & \textbf{Classes of Strategies} \\
+  \hline Non-adaptive (NA) & Exponential \\
+  & Periodic \\
+  & Renewal \\
+  & General non-adaptive \\
+  \hline Adaptive (AD) & Last move (LM) \\
+  & Full History (FH) \\  
+\end{tabular}
+ \caption{Hierarchy of Classes of strategies in FlipIt}
+ \label{table:Strategies}
+ \end{table}
+
+\subsubsection{Imporant results}
+
+\begin{itemize}
+\item periodische spellen domineren de andere renewal strategies
+\item periodische strategien wel slecht tegen attacker die laatste move weet
+\item als de defender snel speelt dan forceert die de attacker om niet meer mee te spelen
+\item iedereen die feedback krijgt heeft meer benefit dan zonder die feedback
+\end{itemize}
+
+\section{Extensions on FlipIt}
+
+In this section we discuss the extensions that can be made on the FlipIt game. \\
+
+There a various possible ways to extend \flip{FlipIt}. 
+Laszka et al.
+made a lot of additions and extensions on the original game of FlipIt.
+For instance Laszka et al.
+extended the basic \flip{FlipIt} game to multiple resources.
+The incentive is that for compromising a system in a real case it needs more than just taking over just one resource.
+An example is that one resource can be gaining access to a system and breaking the password of the system is another resource.
+The model is called FlipThem \cite{FlipThem}.
+They use two ways to flip the multiple resources: the AND and the OR control model.
+In the AND model the attacker only controls the system if he controls all the resources of the system, whereas in the OR model the attacker only needs to compromise one resource to be in control of the entire system.
+\\
+%The difference with FlipThem and this paper is that we introduce a Graph Model in the beginning.\\
+Another addition of Laszka et al.
+to the game of FlipIt [mitigating covert compromises] is extending the game to also consider non-targeted attacks by non-strategic players.
+In this game the defender tries to maintain control over the resource that is subjected to both targeted and non-targeted attacks.
+Non-targeted attacks can include phishing, while targeted attacks may include threats delivered through zero day attack vulnerabilities.
+\\
+One of the last important addition from Laszka et al.
+[Mitigation of Targeted and] is to consider a game where the moves made by the attacker are still covert but the moves made by the defender are known to the attacker.
+This means that the attacker can base his attacks on the defender's moves.
+Both the targeted and non-targeted attacks don't succeed immediately.
+For the targeted attack the time till it succeeds is given by an exponential distributed random variable with a known rate.
+The non-targeted attacks are modelled as a single attacker and the time till it succeeds is given by a Poisson process.\todo{misschien meer uitleggen}.
+The conclusion of this paper is that the optimal strategy for the defender is moving periodically.
+\\
+
+Other authors used the FlipIt game to apply it on a specific scenario. To be able to use the FlipIt game, modifications where required for the FlipIt model.
+One of the scenarios by Pham\cite{GameTheorApprCostBenefitAnalyses} [\todo{citatie needed voor Are We Compromised?}] was to find out whether a resource was compromised or not by the attacker.
+This could be verified by the defender, who has an extra move "test" beside the flip move.
+The basic idea is to test with an extra action if the resource has been compromised or not.
+This move involves also an extra cost.
+This model is useful if somebody wants to know for example if his or her password has been compromised.
+\\
+
+Finally researchers also have investigated the behavioural of humans when playing FlipIt.
+A Nochenson and Grossklags [A behavioural investigation of the FlipIt game]  investigate how people really act when given temporal decisions.
+[Risk-seeking in a continuous game of timing] Reitter et al.
+they observe continuous games, 20-seconds FlipIt game..
+
 % --------------- example of a game -----------------%
@@
 %------------------------------------------------%
-\section{Virusses}
-
-Stealth
-Regin's developers put considerable effort into making it highly inconspicuous.
-Its low key nature means it can potentially be used in espionage campaigns lasting several years.
-Even when its presence is detected, it is very difficult to ascertain what it is doing.
-Symantec was only able to analyze the payloads after it decrypted sample files.
-
-It has several ''stealth'' features.
-These include anti-forensics capabilities, a custom-built encrypted virtual file system (EVFS), and alternative encryption in the form of a variant of RC5, which isn't commonly used.
-Regin uses multiple sophisticated means to covertly communicate with the attacker including via ICMP/ping, embedding commands in HTTP cookies, and custom TCP and UDP protocols
-%http://www.symantec.com/connect/blogs/regin-top-tier-espionage-tool-enables-stealthy-surveillance
-Ways of defending a network:
-\begin{itemize}
-\item Self-defending networks: The next generation of network security
-\item Honeynet games: a game theoretic approach to defending network monitors
-
-\end{itemize}
-Many network security threats today are spread over the Internet. The most common include:
-
-Viruses, worms, and Trojan horses
-Spyware and adware
-Zero-day attacks, also called zero-hour attacks
-Hacker attacks
-Denial of service attacks
-Data interception and theft
-Identity theft
-
-%http://www.ists.dartmouth.edu/library/258.pdf Email Virus Propagation Modeling and Analysis
-%Cliff C. Zou∗, Don Towsley†, Weibo Gong∗
-%∗Department of Electrical & Computer Engineering
-%†Department of Computer Science
-%Univ. Massachusetts, Amherst
-%Technical Report: TR-CSE-03-04
-
-Computer virus through mail. 
-Though virus spreading through email is an old technique, it is still effective and is widely used by
-current viruses and worms. Sending viruses through email has some advantages that are attractive to
-virus writers:
- Sending viruses through email does not require any security holes in computer operating systems
-or software.
- Almost everyone who uses computers uses email service.
- A large number of users have little knowledge of email viruses and trust most email they receive,
-especially email from their friends [28][29].
- Email are private properties like post office letters. Thus correspondent laws or policies are required
-to permit checking email content for detecting viruses before end users receive email [18].
-
-Send a email with malicious attachment.
-Only again infected if attachment again opened.
-Thus this is the action of attacking every neighbour node + also can attack again the node where the virus was coming from.
-There are also email viruses were the malicious program is hidden in the txt and the attachment does not need to be opened.
-
-Spy vs Spy: Aldrich Ames was a CIA Counter-Intelligence officer.
-He was also a spy feeding valuable intelligence to the Soviets and compromising US intelligence operations in the Soviet Union.
-He operated for 9 years before the CIA recognized that they had a spy and began an investigation and determined that he was the leak.
-This strategic situation is the same one faced by computer networks, drug cartels, intelligence agencies and guerrilla networks.
-
-All such organisations have a reasonable expectation that trusted personal/systems will eventually be recruited/captured by enemy organisations.
-Therefore such organisations must consume valuable resources to discover such betrayals and thereby regain secrecy.
-The question is then given the possible threats how often and at what cost should they spend resources on investigations/spy hunts/virus scans.
-This is where flipIt comes in.
-
-FLIPIT: The Game of ''Stealthy Takeover:'' FlipIt was created to model these sorts of strategic situations and to study the best courses of action.
-Specifically flipIt was motivated by the recent interest in and success of Advanced Persistent Threats, or APT.
-
-The basic idea is that given the current experience that perfect protection of trusted resources is unattainable, lets think about how we can optimally manage compromises of the our most trusted systems.
-
-Rules
-
-Two players, player X (blue) and player Y (red) attempt to maintain control over a shared resource.
-At anytime in the game each player is allowed to play 'flip'.
-The only way a player can learn the state of the game (who is in control) is when they play flip.
-If a player is in control of the resource and they play flip they remain in control of the resource.
-If a player is not in control of the resource and they play flip they gain control of the resource.
-Players gain points for the length of time they control the resource.
-Players lose points every time they play flip.
-This reflects the situation that the CIA is placed in with regard to moles/enemy spies.
-They don't know if they have been compromised.
-They can perform an investigation and determine if they have been compromised, also catching the spy in the act, but this action is very expensive.
-That is, the CIA has to trade off between remaining''mole free'' (a good) and investigations (an expense).
-
-Winning: How do you win a fair game of flipIt against intelligent adaptive human adversaries? I'm not sure.
-
-In the real world what is the best move given that the other players can secretly capture/corrupt your most trusted personal/systems? Rives suggests in his talk that you:
-Be prepared to deal with repeated total failure (loss of control).
-Play fast! Aim to make opponent drop out!
-Arrange game so that your moves cost much less than your opponent's!
-%http://www.cisco.com/web/offer/gist_ty2_asset/Cisco_2014_ASR.pdf p49
-%http://repo.hackerzvoice.net/depot_madchat/vxdevl/papers/avers/2004-35.pdf
-%http://www.mcafee.com/us/resources/white-papers/foundstone/wp-managing-malware-outbreak.pdf
-
-\subsection{Malware}
-%Does a company network faces lot of malware? what is the cost ?
-Relevant researches:
-\begin{itemize}
-%http://ants.iis.sinica.edu.tw/3BkMJ9lTeWXTSrrvNoKNFDxRm3zFwRR/17/04483668.pdf
-\item How Viruses and worm can be detected. Difference between UDP en TCP worm propagation
-\end{itemize}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%			Conclusion Chapter 1				%%%%%%
-%%%%%												%%%%%%
-%%%%%												%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusion}
-\label{Cha:1:Conclusion}
-The final section of the chapter gives an overview of the important results
-of this chapter. This implies that the introductory chapter and the
-concluding chapter don't need a conclusion.
-
-%%% Local Variables: 
-%%% mode: latex
-%%% TeX-master: "thesis"
-%%% End:
-
-%\end{document}

### 2015-05-10 doc/template/chap-9.tex
FlipIt strategien

@@
-\chapter{Formula}
+\chapter{FlipIt game with virus propagation}
 \label{cha:9}
@@
 
+\section{Introduction}
+
+\section{FlipIt game with virus propagation}
 %------------------------------------------------%
@@
 %------------------------------------------------%
+\subsection{Actions of the attacker}
+A virus has different kind of ways of making his way through a company network.
+We will describe the different ways of how the virus can propagate.
+For start we will say that the virus or worm will be dropped on Node i and that it has k numbers of neighbours.
+\begin{enumerate}
+\item Node i is infected and will spread the virus or worm to every k neighbours and will stop infecting the neighbours in the next step
+\item Node i is infected and will spread the virus or worm to every k neighbours and will keep on spreading the virus to the same neighbours in every next step
+\item Node i is infected and will spread the virus to only one of the k neighbours and will stop infecting another neighbour in the next step
+\item Node i is infected and will spread the virus to only one of the k neighbours and in the next step it will infect another one of the k neighbours 
+\end{enumerate}
 
+In the game that will be modelled in the paper we will use the settings of the first spreading method.
+We will not use method 2 because this kind of propagation will float the network.
+Because we use the settings of a mail system and contact in a mailing list the method of 3 and 4 are not used.
+\\
+In the first method the node that has been infected can be again infected.
+If one of the neighbours infects the node again the node will infect his neighbours again.
+By using this spreading method we have three distinct states in which a node can be situated.
+An \textit{infected state}, a \textit{clean state} and a \textit{spreading state}.
+An infected state means that the node is infected and will not spread the virus to its neighbours, a clean state means that the node is not infected on that moment and a spreading state means that the node is infected and that it will spread the virus or worm to its neighbours in the next step.
+We can argument this kind of propagation through a mail worm. \todo{voorbeeld geven van zo een worm}
+%Another propagation method is that the virus works as a token. It will propagate to only one neighbour and continue to spread.
+
+The Attacker itself has two different ways of attacking the company network.
+It will only infected one node of the network and will wait for the virus to spread itself through the network.
+We will model two ways of attacks of an Attacker:
+\begin{enumerate}
+\item The attacker drops the virus on a random node on the network
+\item The attacker drops the virus on a targeted node on the network
+\end{enumerate}
+The attacker in this game will put a virus or worm on one of the nodes in the network.
+(This will happen at random.) The attacker does not know on which node the virus will be dropped.
+We will use this randomness because \todo{feit uit security rapport symantec} most viruses are spread via a usb stick or a shared resource.
+If we use this spreading method where we have a targeted attack the attacker will have more information about the network.
+\\
+
+The attacker can choose at which rate it will drop a virus on one of the nodes on the network.
+The cost of dropping a virus will be the same.
+It will not increase.
+If it will increase this means that the attacker will eventually drop out of the game because it becomes to expensive.\\
+The attacker is in control over the game if it manages to infect a subset of all the resources of the company network.
+
+\subsection{Actions of the defender}
+The attacker wants to protect all the nodes of his network.
+It can do so by getting back control over the resources.
+We will assume that the defender of the network has knowledge over his own network.
+Which is convenient in the real world because a company has to know how his infrastructure looks like.\\
+
+The defender has two possible ways of defending its network:
+\begin{enumerate}
+\item The defender flips all the nodes of his network
+\item The defender will flip a subset of the nodes of his network
+\end{enumerate}
+
+The cost of flipping all the nods of the network will be greater than the cost of flipping a subset of nodes.
+We make this assumption because otherwise it will be beneficial for the defender to always flip all the nodes in the network.\\
+
+We will also make the assumption that as a defender flips a node the node can get infected again.
+A flip will not be  correlated to a patch but to a clean-up.
+\todo{waarom geen patch, wormen kunnen veranderen gaandeweg}
+\todo{andere mogelijkheid:} Another setting of the game can be that the flip of the defender is equal to a patch and that the resource cannot be infected any more.
+But with this case we deviate from the flipIt game, because the attacker cannot flip the resource any more.
+Unless we work with different virusses every time the attacker flips.
+We start with the less complex game of flipping is equal to a clean-up.
+
+\subsection{Strategies of both players}
+We explained what the actions of each player are.
+
+\section{Formal definition Game}
+
+In this section we provide a formal definition of the game and the notation that we will use throughout the paper. 
 \section*{Playing periodically with virus propagation}
@@
 
+\section{Simulation}
 %nce the defender has played, it takes 1 time until the attacker plays and the defender loses control again.

### 2015-05-10 doc/template/intro.tex
FlipIt strategien

@@
 
+\section{Virusses}
+
+Stealth
+Regin's developers put considerable effort into making it highly inconspicuous.
+Its low key nature means it can potentially be used in espionage campaigns lasting several years.
+Even when its presence is detected, it is very difficult to ascertain what it is doing.
+Symantec was only able to analyze the payloads after it decrypted sample files.
+
+It has several ''stealth'' features.
+These include anti-forensics capabilities, a custom-built encrypted virtual file system (EVFS), and alternative encryption in the form of a variant of RC5, which isn't commonly used.
+Regin uses multiple sophisticated means to covertly communicate with the attacker including via ICMP/ping, embedding commands in HTTP cookies, and custom TCP and UDP protocols
+%http://www.symantec.com/connect/blogs/regin-top-tier-espionage-tool-enables-stealthy-surveillance
+Ways of defending a network:
+\begin{itemize}
+\item Self-defending networks: The next generation of network security
+\item Honeynet games: a game theoretic approach to defending network monitors
+
+\end{itemize}
+Many network security threats today are spread over the Internet. The most common include:
+
+Viruses, worms, and Trojan horses
+Spyware and adware
+Zero-day attacks, also called zero-hour attacks
+Hacker attacks
+Denial of service attacks
+Data interception and theft
+Identity theft
+
+%http://www.ists.dartmouth.edu/library/258.pdf Email Virus Propagation Modeling and Analysis
+%Cliff C. Zou∗, Don Towsley†, Weibo Gong∗
+%∗Department of Electrical & Computer Engineering
+%†Department of Computer Science
+%Univ. Massachusetts, Amherst
+%Technical Report: TR-CSE-03-04
+
+Computer virus through mail. 
+Though virus spreading through email is an old technique, it is still effective and is widely used by
+current viruses and worms. Sending viruses through email has some advantages that are attractive to
+virus writers:
+ Sending viruses through email does not require any security holes in computer operating systems
+or software.
+ Almost everyone who uses computers uses email service.
+ A large number of users have little knowledge of email viruses and trust most email they receive,
+especially email from their friends [28][29].
+ Email are private properties like post office letters. Thus correspondent laws or policies are required
+to permit checking email content for detecting viruses before end users receive email [18].
+
+Send a email with malicious attachment.
+Only again infected if attachment again opened.
+Thus this is the action of attacking every neighbour node + also can attack again the node where the virus was coming from.
+There are also email viruses were the malicious program is hidden in the txt and the attachment does not need to be opened.
+
+Spy vs Spy: Aldrich Ames was a CIA Counter-Intelligence officer.
+He was also a spy feeding valuable intelligence to the Soviets and compromising US intelligence operations in the Soviet Union.
+He operated for 9 years before the CIA recognized that they had a spy and began an investigation and determined that he was the leak.
+This strategic situation is the same one faced by computer networks, drug cartels, intelligence agencies and guerrilla networks.
+
+All such organisations have a reasonable expectation that trusted personal/systems will eventually be recruited/captured by enemy organisations.
+Therefore such organisations must consume valuable resources to discover such betrayals and thereby regain secrecy.
+The question is then given the possible threats how often and at what cost should they spend resources on investigations/spy hunts/virus scans.
+This is where flipIt comes in.
+
+FLIPIT: The Game of ''Stealthy Takeover:'' FlipIt was created to model these sorts of strategic situations and to study the best courses of action.
+Specifically flipIt was motivated by the recent interest in and success of Advanced Persistent Threats, or APT.
+
+The basic idea is that given the current experience that perfect protection of trusted resources is unattainable, lets think about how we can optimally manage compromises of the our most trusted systems.
+
+Rules
+
+Two players, player X (blue) and player Y (red) attempt to maintain control over a shared resource.
+At anytime in the game each player is allowed to play 'flip'.
+The only way a player can learn the state of the game (who is in control) is when they play flip.
+If a player is in control of the resource and they play flip they remain in control of the resource.
+If a player is not in control of the resource and they play flip they gain control of the resource.
+Players gain points for the length of time they control the resource.
+Players lose points every time they play flip.
+This reflects the situation that the CIA is placed in with regard to moles/enemy spies.
+They don't know if they have been compromised.
+They can perform an investigation and determine if they have been compromised, also catching the spy in the act, but this action is very expensive.
+That is, the CIA has to trade off between remaining''mole free'' (a good) and investigations (an expense).
+
+Winning: How do you win a fair game of flipIt against intelligent adaptive human adversaries? I'm not sure.
+
+In the real world what is the best move given that the other players can secretly capture/corrupt your most trusted personal/systems? Rives suggests in his talk that you:
+Be prepared to deal with repeated total failure (loss of control).
+Play fast! Aim to make opponent drop out!
+Arrange game so that your moves cost much less than your opponent's!
+%http://www.cisco.com/web/offer/gist_ty2_asset/Cisco_2014_ASR.pdf p49
+%http://repo.hackerzvoice.net/depot_madchat/vxdevl/papers/avers/2004-35.pdf
+%http://www.mcafee.com/us/resources/white-papers/foundstone/wp-managing-malware-outbreak.pdf
+
+\subsection{Malware}
+%Does a company network faces lot of malware? what is the cost ?
+Relevant researches:
+\begin{itemize}
+%http://ants.iis.sinica.edu.tw/3BkMJ9lTeWXTSrrvNoKNFDxRm3zFwRR/17/04483668.pdf
+\item How Viruses and worm can be detected. Difference between UDP en TCP worm propagation
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%			Conclusion Chapter 1				%%%%%%
+%%%%%												%%%%%%
+%%%%%												%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusion}
+\label{Cha:1:Conclusion}
+The final section of the chapter gives an overview of the important results
+of this chapter. This implies that the introductory chapter and the
+concluding chapter don't need a conclusion.
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "thesis"
+%%% End:
+
+%\end{document}
+
 %\begin{minipage}{0.5\textwidth}

### 2015-05-10 doc/template/chap-9.tex
verdere aanvullingen aan deel delay

@@
 \section{FlipIt game with virus propagation}
+
+Motivatie voor het veranderen van FLipIt naar een FlipIt met viruspropagatie:
+
+\section{Explaining difference between FlipIt with and without virus propagation}
+
+In chapter 2 the FlipIt game was explained.
+This chapter starts from the specific case of a non-adaptive continuous FlipIt game where both players play a periodic strategy with a random phase.
+This choice is motivated by the assumption that in the practical situation of most organisations, the defence strategy is to periodically defend the network.
+This corresponds to a periodic defender strategy.
+\todo{bedrijven hebben niet het geld en de mankracht om zich daar mee bezig te houden en dan is periodisch verdedigen het handigste, bedrijven willen alles zo efficient en met de minste kost doen} To simplify the analysis in a first time, a periodic attacker strategy is assumed as well.
+Further research can investigate the effect of relaxing this assumption.\\
+
+A FlipIt game consists of a single resource. To represent the security problem, the game now defines its single resource as a computer network with multiple
+nodes. One of the players, the defender, will try to defend his network. The defender
+will do this by flipping all the nodes of the network (i.e. the entire resource) in every move he plays. The
+attacker, the other player, will try to infect all the nodes in the network. The attacker
+will do this by flipping the node in the graph that can infect all the nodes in the
+shortest time possible.
+After dropping a virus on the first node, it takes a while for the virus to infect the entire network.
+However, since the original FlipIt game works with a single resource that is always flipped entirely, the assumption is made that the attacker is considered to have gained the control over the resource only when all the nodes of the network have been infected, i.e.
+the entire resource has been flipped.
+
+After dropping a virus on the first resource, it takes a while for the virus to infect
+the entire network. The time that it takes for the virus to infect every node will be
+denoted as parameter d. If we want to measure how long it takes for the virus to
+infect all the nodes in the network, we have to calculate the shortest path from the
+first infected node to the farthest node. This can be measured by a method that we
+will explain in section [].
+Assume that an attacker attacks at time t, then only at time t + d he gains control over the entire network.
+If the defender flips the network before the period d has elapsed (so, somewhere between t and t+ d), then the attacker will never gain control over the entire network.
+Using this parameter d, a FlipIt game with virus propagation
+can be modelled.
+
 %------------------------------------------------%
@@
 
-\subsection{Strategies of both players}
-We explained what the actions of each player are.
-
 \section{Formal definition Game}
@@
 In this section we provide a formal definition of the game and the notation that we will use throughout the paper.
+
 \section*{Playing periodically with virus propagation}
@@
 In the last section we calculate the Nash equilibrium for the benefit formula.
-
-\section{Explaining difference between FlipIt with and without virus propagation}
-zelfde als in vorige chapter:
-
-In chapter 2 the FlipIt game was explained.
-This chapter starts from the specific case of a non-adaptive continuous FlipIt game where both players play a periodic strategy with a random phase.
-This choice is motivated by the assumption that in the practical situation of most organisations, the defence strategy is to periodically defend the network.
-This corresponds to a periodic defender strategy.
-\todo{bedrijven hebben niet het geld en de mankracht om zich daar mee bezig te houden en dan is periodisch verdedigen het handigste, bedrijven willen alles zo efficient en met de minste kost doen} To simplify the analysis in a first time, a periodic attacker strategy is assumed as well.
-Further research can investigate the effect of relaxing this assumption.\\
-
-A FlipIt game consists of a single resource. To represent the security problem, the game now defines its single resource as a computer network with multiple
-nodes. One of the players, the defender, will try to defend his network. The defender
-will do this by flipping all the nodes of the network (i.e. the entire resource) in every move he plays. The
-attacker, the other player, will try to infect all the nodes in the network. The attacker
-will do this by flipping the node in the graph that can infect all the nodes in the
-shortest time possible.
-After dropping a virus on the first node, it takes a while for the virus to infect the entire network.
-However, since the original FlipIt game works with a single resource that is always flipped entirely, the assumption is made that the attacker is considered to have gained the control over the resource only when all the nodes of the network have been infected, i.e.
-the entire resource has been flipped.
-
-After dropping a virus on the first resource, it takes a while for the virus to infect
-the entire network. The time that it takes for the virus to infect every node will be
-denoted as parameter d. If we want to measure how long it takes for the virus to
-infect all the nodes in the network, we have to calculate the shortest path from the
-first infected node to the farthest node. This can be measured by a method that we
-will explain in section [].
-Assume that an attacker attacks at time t, then only at time t + d he gains control over the entire network.
-If the defender flips the network before the period d has elapsed (so, somewhere between t and t+ d), then the attacker will never gain control over the entire network.
-Using this parameter d, a FlipIt game with virus propagation
-can be modelled.
 

### 2015-05-13 doc/template/chap-10.tex
FlipIt strategien
Case 2 geschreven

@@
 
-% woorden die in introductie moeten voorkomen: Game theory, cybersecurity, APT, FlipIt, virus propagation
+% woorden die in introductie moeten voorkomen: Game theory, cybersecurity, APT, FlipIt, virus propagation, network security
 
+\subsubsection{Situation}
 % -- Situation -- %
@@
 Why is it so important to keep a system secure?
-\todo{waarom?
-:  lekken van informatie: je hebt info over klanten en moet hun privacy beschermen, je eigen data is geld waard voor concurrenten, DOSS attacks: je will je service-level agreements kunnen nakomen, ..} A hacker will be a person that seeks exploits or weaknesses in a system or network in order to gain access.
+\\
+
+Businesses can have confidential information on clients.
+Through data leakage, confidential information can be lost and possibly used by the competitor.
+Businesses wants to meet their service-level agreements.
+They will protect themselves against disruption that can be caused by DOSS attacks.
+Ultimately, system and network security helps protecting a business's reputation, which is one of its most important assets.
+%\todo{waarom?
+:  lekken van informatie: je hebt info over klanten en moet hun privacy beschermen, je eigen data is geld waard voor concurrenten, DOSS attacks: je will je service-level agreements kunnen nakomen, ..}
+ A hacker will be a person that seeks exploits or weaknesses in a system or network in order to gain access.
 Many of those attacks have a different cause.
@@
 There are various ways to break into a system.
-Virusses, worms, spyware and other malware are the number two of the top external threats [security report kaspersky 2014].
+Viruses, worms, spyware and other malware are the number two of the top external threats that a business faces [security report kaspersky 2014].
+(number one is Spam).
 Furthermore these kind of threats also causes the greatest percentage in loss of data.
 These threats will infect the network by means of a virus that will propagate through the network.
+Most of the attacks are Advanced Persistent Threats (APT).\\
+
+An APT is a targeted cyber attack that will keep trying to break into the targeted system.
+It will try to stay unnoticed for as long as possible.
+Bruce Schneier describes an APT as something different and stronger than a conventional hacker: ''\textit{A conventional hacker or criminal isn't interested in any particular target.
+He wants a thousand credit card numbers for fraud, or to break into an account and turn it into a zombie, or whatever.
+Security against this sort of attacker is relative; as long as you're more secure than almost everyone else, the attackers will go after other people, not you.
+An APT is different; it's an attacker who - for whatever reason - wants to attack you.
+Against this sort of attacker, the absolute level of your security is what's important.
+It doesn't matter how secure you are compared to your peers; all that matters is whether you're secure enough to keep him out}'' - Bruce Schneier: APT is a Useful Buzzword.
+
+-- verder uitwerken -- \\
+\textit{Stealthy key characteristic of APT. \\
+Voorbeeld APT Equastion group. \\
+Site van Kaspersky lab\\}
 \todo{APT uitleggen ? en dat het moeilijk is om een systeem te beschermen tegen APT ==> overgang naar de volgende paragraaf maken.}
 
+\subsubsection{Complication}
 % -- Complication -- %
-Since it is so difficult to protect a system or a network against APT, researcher have been looking for effective ways to predict in advance which defense strategy might be the better one. 
+Since it is so difficult to protect a system or a network against APT, researcher have been looking for effective ways to predict in advance which defence strategy might be the better one. 
 Game theory is gaining more and more interest as an effective technique to  model and study Cyber Security.
@@
 They study the specific scenario where a system or network is repeatedly taken over completely by an attacker and this attack is not immediately detected by the defender of the system or network.
-In game theory, such a game is known as ''FlipI''.
+In game theory, such a game is known as ''FlipIt''.
 This is a two players game where the attacker and the defender are competing to get control over a shared resource.
@@
 The attacker will only be in control of the resource once the whole network is infected.
+\\
 
-The game theoretical approach of the FlipIt does not take such delay into account. This an lead us to the following research questions:
-- Is it possible to incorporate the notion of delay in the game-theoretical analysis of the Flip-It game ?
-
-%Specific research question:
-%-         Is it possible to incorporate the notion of delay in the game-theoretical analysis of the Flip-It game ?
-%
-%-        Can we calculate a Nash Equilibrium for a FlipIt Game with delay ?
-%
-%-        
-- Does this allow us to determine an optimal defense strategy against an attacker ?
+\subsubsection{Research questions}
+The game theoretical approach of the FlipIt does not take such delay into account. 
+This an lead us to the following research questions:
+\begin{itemize}
+\item Is it possible to incorporate the notion of delay in the game-theoretical analysis of the Flip-It game ?
+\item Is it possible to incorporate the notion of delay in the game-theoretical analysis of the Flip-It game ?
+\item Does this allow us to determine an optimal defense strategy against an attacker ?
 for example: Gaat er een specieke grootte zijn van een delay waarbij de attacker al weet dat hem niet meer moet gaan spelen ?
 ( is niet gelijk aan de grootte van de periode van de attacker)
-When working with a network and a delay we can .. graph model en uitleggen hoe we de graph kunnen maken zodat de delay altijd zo groot mogelijk gaat zijn.
+\end{itemize}
+
+Resarch questions when working with a network and a delay: .. graph model en uitleggen hoe we de graph kunnen maken zodat de delay altijd zo groot mogelijk gaat zijn. 
+\begin{itemize}
+\item How can we calculate the expected duration for a node's infection/the entire network infection ?
+\item Can we calculate this node per node ?
+\end{itemize}
+%-        How can we calcuate the expected duration for a node’s infection/the entire network infection ?
+%
+%-        Can we calculate this node per node ?
 
 % -- Anwer -- %
-We propose an addition to the basic FlipIt model to model a scenario where the moves by the attacker will not be instantaneous. Next we analyse what the new Nash equilibria will be and ..
+\subsubsection{Contributions}
+We propose an addition to the basic FlipIt model to model a scenario where the moves by the attacker will not be instantaneous. Next we analyse what the new Nash equilibria will be and .. \\
+
+\subsubsection{Overview of the thesis}
+--opbouw van de thesis uitleggen-- \\
+
+The organisation of this paper is the following.
+In chapter 2 [] \todo{ref}  a brief introduction to Gametheory is introduced to get familiar with the game theoretic concepts that will be further used in the paper.
+In the same chapter the FlipIt framework  is summarized and the most important conclusions together with the the related work done on FlipIt and the difference with this paper.
+In chapter 3  [] \todo{ref} , we first introduce the adaptations made on FlipIt to model a FlipIt game with virus propagation.
+After that formulas are derived to model a FlipIt game with a virus propagation for a specific case where players play a periodic strategy with a random phase.
+This chapter ends with simulations where conclusions can be derived.
+Next in Chapter 4 [] \todo{ref}, we given an overview of the various ways in which a virus can propagate.
+We present a method to calculate the speed of the propagation of a virus in a network and how this network can be established to reduce the spreading of a virus.
+Finally, in Chapter 5 [] \todo{ref}, we discuss the main results and complications and provide directions for further research.
 

### 2015-05-13 doc/template/chap-9.tex
Case 2 geschreven

@@
 
-\section{Benefit for FlipIt game with virus propagation}
+\section{Benefit for a periodic FlipIt game with virus propagation}
+We start by formalizing the non-adaptive continuous basic FlipIt game where players use a periodic strategy with a random phase.
+Next we split the formalisation for the benefit formula for a FlipIt game with virus propagation in two cases.
+First the benefit formula is derived for the basis case and then it is calculated for the case with a virus propagation.
+One case is where the defender plays at least as fast as the attacker, the other case is where the attacker plays at least as fast as the defender.
+This subdivision is also used in the FlipIt paper.
+\\
 
-Periodic Game with delay for the attacker:
+ A Periodic strategy is a non-adaptive renewal strategy where the time intervals between consecutive moves are a fixed period, denoted by $\delta$.
+Moreover it has a random phase, that is chosen uniformly and random in the interval $[0,\delta]$ for the first move.
+The average rate of play of a player is denoted by $\alpha_{i} = \dfrac{1}{\delta_{i}}$.
+\\
+~~\\
 
-\textbf{Case 1:} $\delta_{D} \leq \delta_{A} $(The defender plays at least as fast as the attacker.) \\
+\subsection*{\textbf{Case 1:} $\delta_{D} \leq \delta_{A} $ (The defender plays at least as fast as the attacker.) }
 
@@
 
-The expected period of attacker control within the interval would be r/2, without considering the delay. \\
+The expected period of attacker control within the interval would be r/2, without considering the delay by a virus. Therefore the benefit without this virus can be expressed as follows:
+\begin{equation}\label{first}
+\beta_{D}(\alpha_{D},\alpha_{A}) =1 -  \dfrac {r} {2} - k_{D} \alpha_{D} = 1 - \dfrac {\delta_{D}} {2\delta_{A}} - k_{D} \alpha_{D} 
+\end{equation}
+\begin{equation}\label{first}
+\beta_{A}(\alpha_{D},\alpha_{A}) =\dfrac {r} {2} - k_{A} \alpha_{A} = \dfrac {\delta_{D}} {2\delta_{A}} - k_{A} \alpha_{A}  
+\end{equation}\\
 
-However, because of the delay, the maximal time of control is reduced to $\delta_{D}-d$.
+However, because of the delay caused by a virus, the maximal time of control is reduced to $\delta_{D}-d$.
 There is a probability of \textit{r} that the attacker will move in the interval of the defender.
@@
 \end{equation}
+~~\\
 
-\textbf{Case 2:} $\delta_{A} \leq \delta_{D} $ (The attacker plays at least as fast as the defender.) \\
+\subsection*{\textbf{Case 2:} $\delta_{A} \leq \delta_{D} $ (The attacker plays at least as fast as the defender.) }
 
-Let $r = \dfrac{\delta_{D}}{ \delta_{A} }$.
+First let $r = \dfrac{\delta_{D}}{ \delta_{A} }$.
 The intervals between two consecutive attacker's moves have length $\delta_{A}$.
@@
 
-\section{something}
-\label{Cha:9:inrto}
-
-Periodic Game with delay for the attacker:
-
-\textbf{Case 1:} $\delta_{D} \leq \delta_{A} $(The defender plays at least as fast as the attacker.) \\
-
-Let $r = \dfrac{\delta_{D}}{ \delta_{A} }$.
-The intervals between two consecutive defender's moves have length $\delta_{D}$.
-Consider a given defender move interval.
-The probability over the attacker's phase selection that the attacker moves in this interval is r.
-Given that the attacker moves within the interval, he moves exactly once within the interval (since $\delta_{D} \leq \delta_{A} $) and his move is distributed uniformly at random.
-\\
-
-The expected period of attacker control within the interval would be r/2, without considering the delay. \\
-
-However, because of the delay, the maximal time of control is reduced to $\delta_{D}-d$.
-If we consider a duration of $\delta_{D} \cdot \delta_{A}$ the attacker will play $\delta_{D}$ times.
-If the attacker plays soon enough it will get a gain of $\dfrac{\delta_{D}-d}{2}$ in $\delta_{D}-d$ of the cases.
-In \textit{d} cases it will receive a gain of zero.
-This is the case were the duration of the delay causes the defender to play before the attacker can get control over the resource.
-So the gain of the attacker can be expressed as follows:
+A similar analysis as in case 1 for a FlipIt game without virus propagation yields the following benefits:
 \begin{equation}\label{first}
-Gain =  \dfrac{\delta_{D}-d}{2} \cdot (\delta_{D}-d) + 0 \cdot d = \dfrac{\delta_{D}-d}{2} \cdot (\delta_{D}-d)
-\end{equation}
-
-The benefit of the attacker can be expressed as follows
-
-\begin{equation}\label{first}
-\beta_{A}(\alpha_{D},\alpha_{A}) = \dfrac { (\delta_{D}-d) ^{2}} {2 \cdot \delta_{D}  \delta_{A}} + k_{A} \cdot \alpha_{A}
+\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac {1} {2r} - k_{D} \alpha_{D} = \dfrac {\delta_{A}} {2\delta_{D}} - k_{D} \alpha_{D} 
 \end{equation}
 \begin{equation}\label{first}
-\beta_{A}(\alpha_{D},\alpha_{A}) = \dfrac { \delta_{D}} {2 \cdot \delta_{A}} + k_{A} \cdot \alpha_{A} + \dfrac{d}{\delta_{A}} + \dfrac{d^{2}}{2 \cdot \delta_{A} \delta_{D}}
-\end{equation}
-\\
+\beta_{A}(\alpha_{D},\alpha_{A}) =1 - \dfrac {1} {2r} - k_{A} \alpha_{A} = 1- \dfrac {\delta_{A}} {2\delta_{D}} - k_{A} \alpha_{A}  
+\end{equation}\\
 
-The benefit of the defender is then:
+For the case with a virus we consider two cases, \textbf{Case a} and \textbf{Case b}, to compute the benefit of both players.  \\
 
+\subsubsection*{\textbf{Case a:} $d + \delta_{A} \leq \delta_{D}$}
+
+If $d + \delta_{A} \leq \delta_{D}$ that means that in this timespan for the attacker, the defender will never move twice.
+$\dfrac{delta_{A}}{\delta_{D}}$ the defender will move during the interval of the attacker.
+When this happens the defender will end with beeing in control at the end of the interval.
+In the next interval the attacker will move in the beginning but will not gain control after the delay of \textit{d}.
+This means that if the defender moved in the interval prior to this one, the delay is the minimum gain of the defender.
+To calculate the gain rate of the defender we don't look at one interval of the defender.
+We know that if the defender moves within the interval of the defender it will receive a minimum of $\dfrac{\delta_{A}}{2}$ gain.
+In the next interval the defender will receive a minimum gain of $d$.
+It is also possible that the defender will move in this interval after the delay, adding a gain of $\dfrac{\delta_{A}}{2}$, but the defender will never play during the delay if in the interval prior to this, the defender already has moved.
+
+This means that every time the defender plays it will get a benefit of $\dfrac{\delta_{A}}{2}$ and the next interval will always receive a gain of $d$.
+To derive the benefit formula for the defender we can calculate the probability that the defender will move during an interval of the attacker and multiply it by the gain of the interval plus the gain that it will give to the next interval.
+This yields in the following formula:
 \begin{equation}\label{first}
-\beta_{D}(\alpha_{D},\alpha_{A}) = 1 - \dfrac { (\delta_{D}-d) ^{2}} {2 \cdot \delta_{D} \cdot \delta_{A}} + k_{D} \cdot \alpha_{D}
+\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta{A}}{\delta_{D}} \cdot \dfrac{(d+\dfrac{\delta_{A}}{2})}{\delta_{A}} - k_{D} \alpha_{D} 
 \end{equation}
 \begin{equation}\label{first}
-\beta_{D}(\alpha_{D},\alpha_{A}) = 1 - \dfrac { \delta_{D}} {2 \cdot \delta_{A}} + k_{D} \cdot \alpha_{D} - \dfrac{d}{\delta_{A}}- \dfrac{d^{2}}{2 \cdot \delta_{A} \delta_{D}}
-\end{equation}
+\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - k_{D} \alpha_{D} 
+\end{equation}\\
 
-\textbf{Case 2:} $\delta_{A} \leq \delta_{D} $ (The attacker plays at least as fast as the defender.) \\
+The benefit for the attacker will be as follows:
+\begin{equation}\label{first}
+\beta_{A}(\alpha_{D},\alpha_{A}) = 1 -\dfrac{\delta{A}}{2\delta_{D}} - \dfrac{d}{\delta_{D}} - k_{A} \alpha_{A} 
+\end{equation}\\
 
-Let $r = \dfrac{\delta_{D}}{ \delta_{A} }$.
-The intervals between two consecutive attacker's moves have length $\delta_{A}$.
-Consider a given attackers move interval.
-The probability over the attacker's phase selection that the defender moves in this interval is $\dfrac{\delta_{D}}{ \delta_{A} } = (1/r)$.
-Given that the defender moves within the interval, he moves exactly once within the interval (since$\delta_{A} \leq \delta_{D} $) and his move is distributed uniformly at random.
-\\
+\begin{figure}[hbtp]
+\caption{Case 2 where d + deltaA < deltaD}
+\centering
+\includegraphics[scale=0.7]{Images/FlipItCase2.png}
+\end{figure}
 
-If we consider a duration of $\delta_{A} \cdot \delta_{D} $ there is a probability of $\dfrac{\delta_{A} } {\delta_{D}} $ that the defender moves within the interval of the attacker.
-The defender will then receive an average gain of $\dfrac{\delta_{A}} {2} $.
-There is $1- \dfrac{\delta_{A} } {\delta_{D}} $ probability that the defender will not move in the interval of the attacker and so the defender will receive no gain.
-The benefit can be expressed as follows when the  defender plays $\delta_{D}$ times during a duration of $\delta_{A} \cdot \delta_{D}$:
-\begin{equation}\label{first}
-\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac { 1} {\delta_{A} \delta_{D}} \cdot \delta_{D} \cdot [ \dfrac{\delta_{A}}{\delta_{D}} \cdot \dfrac{\delta_{A}}{2}+[ 1-\dfrac{\delta_{A}}{ \delta_{D}}] \cdot 0 ] + k_{D} \cdot \alpha_{D}
-\end{equation}
-\begin{equation}\label{first}
-\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac {\delta_{A} }{2 \cdot \delta_{D}} + k_{D} \cdot \alpha_{D} $$same as the FlipIt solution$$
-\end{equation}
+It is crucial that the $d + \delta_{1}$ is not greater than $ \delta_{0}$.
+If it is greater that means that the defender can move during a delay after an interval of the attacker where the defender already has moved.
+If we calculate the benefit formula in the same manner, to much gain is added.
+Namely the overlap during the delay.
 
-However, because of the delay, the maximal time of control of the defender is increased by d.
-In other words, the defender has some benefit time of d before the attacker really gains control over the resource, meaning that the attacker gains control only after $\delta_{A}+d$ instead of after $\delta_{A}$.
-So, when the defender plays, with a probability of $\dfrac{\delta_{A} } {\delta_{D}} $, the expected gain of the defender's control in this interval would be more than half of the period $\delta_{A} $: it is $\dfrac{\delta_{A} + d}{2}$.
-There is $1- \dfrac{\delta_{A} } {\delta_{D}} $ probability that the defender will not move in the interval of the attacker but because of the delay the defender will receive a gain of d.
-So the benefit of the defender can be expressed as:
-\begin{equation}\label{first}
-\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac { 1} {\delta_{A} \delta_{D}} \cdot \delta_{D} \cdot [ \dfrac{\delta_{A} }{\delta_{D}} \cdot \dfrac{\delta_{A} + d}{2}+[ 1-\dfrac{\delta_{A}}{ \delta_{D}}] \cdot d ] + k_{D} \cdot \alpha_{D}
-\end{equation}
-\begin{equation}\label{first}
-\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac {\delta_{A} - d}{2 \cdot \delta_{D}} + \dfrac{d}{\delta_{A}}  + k_{D} \cdot \alpha_{D}
-\end{equation}
-
-The benefit of the attacker is expressed as follows:
-\begin{equation}\label{first}
-\beta_{A}(\alpha_{D},\alpha_{A}) = 1- [ \dfrac {\delta_{A} - d}{2 \cdot \delta_{D}} + \dfrac{d}{\delta_{A}}  ] + k_{A} \cdot \alpha_{A}
-\end{equation}
+\subsubsection*{\textbf{Case b:} $d + \delta_{1} > \delta_{D}$}
 

### 2015-05-14 IEEE Paper/paper/paper.tex
bib file in orde !

%\newcommand{\CLASSINPUTbaselinestretch}{1.0} % baselinestretch
%\newcommand{\CLASSINPUTinnersidemargin}{1in} % inner side margin
%\newcommand{\CLASSINPUToutersidemargin}{1in} % outer side margin
%\newcommand{\CLASSINPUTtoptextmargin}{1in}   % top text margin
%\newcommand{\CLASSINPUTbottomtextmargin}{1in}% bottom text margin

\documentclass[journal,a4paper]{IEEEtran}

\usepackage[pdftex]{graphicx}  
\usepackage[cmex10]{amsmath}
\usepackage[dutch]{babel}

\usepackage{float}
\usepackage{caption}
\usepackage[caption = false,subrefformat=parens,labelformat=parens]{subfig}
\usepackage{amsmath}
\usepackage{epstopdf}
\usepackage{textcomp}
\usepackage{gensymb}
\usepackage[titlenumbered,algoruled, linesnumbered]{algorithm2e}
\usepackage{algpseudocode}
\usepackage{booktabs}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\graphicspath{{C:/Users/Jeroen/Dropbox/Thesis/latex/figures/}}

\newcommand\MYhyperrefoptions{bookmarks=true,bookmarksnumbered=true,
pdfpagemode={UseOutlines},plainpages=false,pdfpagelabels=true,
colorlinks=true,linkcolor={black},citecolor={black},pagecolor={black},
urlcolor={black},
pdftitle={Ontwerp en implementatie van navigatie-assistentie voor rolstoelgebruikers met oogbesturing},
pdfauthor={Jeroen Van Hemelen}}

\begin{document}

\title{Ontwerp en implementatie van navigatie-assistentie voor rolstoelgebruikers met oogbesturing}
\author{Jeroen Van Hemelen}
\maketitle

\IEEEcompsoctitleabstractindextext{%
\begin{abstract}
Dit werk beschrijft de ontwikkeling en implementatie van een systeem op basis van oogcontrole dat de gebruiker efficiënter helpt navigeren. Dit wordt geïmplementeerd door de omgeving te verdelen in een discrete set mogelijke doelen, en de intentie van de gebruiker te schatten. Deze informatie wordt door een Maximum Likelihood, Maximum A Posteriori of Partially Observable Markov Decision Proces algoritme gebruikt om naar het juiste doel te navigeren. Uit de resultaten blijkt dat gebruiker met deze methode tot 40\% minder signalen moet geven. De POMDP en ML methode hebben specifieke zwakheden: de POMDP methode werkt niet goed indien de rolstoel grote acties mag ondernemen, en de ML methode mag niet gecombineerd worden met een continu gebruikersmodel.
\end{abstract}
}

\IEEEdisplaynotcompsoctitleabstractindextext
\IEEEpeerreviewmaketitle

\section{Inleiding}
\label{sec:inleiding}

\IEEEPARstart{E}{lektrische} rolstoelen worden door mindervaliden over heel de wereld gebruikt om zichzelf te verplaatsen. Doorgaans worden dergelijke rolstoelen bestuurd met een joystick, maar dit is voor sommige mensen geen optie. Ze beschikken niet over de benodigde handcontrole om zich veilig te verplaatsen met een dergelijke rolstoel. Deze mensen dienen van alternatieve controlemethodes gebruik te maken, of ze worden rondgereden door een verzorger.

\cite{Barea2003}

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,references}

\end{document}

### 2015-05-14 doc/template/chap-5.tex
Case 2 geschreven
bib file in orde !

@@
 To get an understanding of the most important concepts of game theory, a short introduction based on the work of
-\citep{leyton2008essentials} and \todo{Coursera} is given in section 2.1 [] \todo{ref}. For a more detailed and full introduction to game theory, the reader is referred to 
-\todo{leyton2008essentials}.  In section 2.2 [] \todo{ref} an overview of the FlipIt game is given with the definitions and concepts that will be used throughout the paper.
+\cite{leyton2008essentials} and \todo{Coursera} is given in section 2.1 [] \todo{ref}. For a more detailed and full introduction to game theory, the reader is referred to 
+\cite{MitigationCovert}.  In section 2.2 [] \todo{ref} an overview of the FlipIt game is given with the definitions and concepts that will be used throughout the paper.
 The last section [] \todo{ref} will cover the extensions and additions already made on FlipIt.
@@
 %There are two standard representations for games. The first one is the Normal Form. The second one is the Extensive Form.
-
+\subsubsection{List of terms}
 In the following list a couple of terms that will be used throughout the paper.

### 2015-05-15 IEEE Paper/paper/paper.tex
paper

@@
 \usepackage[utf8]{inputenc}
-\graphicspath{{C:/Users/Jeroen/Dropbox/Thesis/latex/figures/}}
 
@@
 urlcolor={black},
-pdftitle={Ontwerp en implementatie van navigatie-assistentie voor rolstoelgebruikers met oogbesturing},
-pdfauthor={Jeroen Van Hemelen}}
+pdftitle={Flip the virus:\\ A gametheoretic approach to cybersecurity},
+pdfauthor={Sophie Marien}}
 
@@
 
-\title{Ontwerp en implementatie van navigatie-assistentie voor rolstoelgebruikers met oogbesturing}
-\author{Jeroen Van Hemelen}
+\title{Flip the virus:\\ A gametheoretic approach to cybersecurity}
+\author{Sophie Marien}
 \maketitle
@@
 \begin{abstract}
-Dit werk beschrijft de ontwikkeling en implementatie van een systeem op basis van oogcontrole dat de gebruiker efficiënter helpt navigeren.
-Dit wordt geïmplementeerd door de omgeving te verdelen in een discrete set mogelijke doelen, en de intentie van de gebruiker te schatten.
-Deze informatie wordt door een Maximum Likelihood, Maximum A Posteriori of Partially Observable Markov Decision Proces algoritme gebruikt om naar het juiste doel te navigeren.
-Uit de resultaten blijkt dat gebruiker met deze methode tot 40\% minder signalen moet geven.
-De POMDP en ML methode hebben specifieke zwakheden: de POMDP methode werkt niet goed indien de rolstoel grote acties mag ondernemen, en de ML methode mag niet gecombineerd worden met een continu gebruikersmodel.
+Recently the attacks on Belgacom and other high profile targeted attacks shows us that even
+the most secure companies can still be compromised. It also shows that these attacks are
+not immediately detected. FlipIt is a framework that can model these stealthy takeovers and is
+proposed by a group of researchers at RSA. It is a 2-players game composed of a single attacker,
+a single defender and a single shared resource. These players will try to gain control over the
+shared resource and they do this in a stealthy way. In this paper we try to adapt FlipIt in a
+way that we can use it to model the game of defending a company network that is attacked by a virus. The FlipIt formulas are adapted for this virus propagation. Through analytic results we can provide
+useful information for the defender to defend his network.
 \end{abstract}
@@
 \label{sec:inleiding}
+\IEEEPARstart{I}{n this era} where digitalization becomes prominent in every aspect of our lives, where technology is growing fast and where businesses are always under attack, security becomes an issue of increasing complexity.
+Without security, there is no protection to keep somebody out of a system.
+It is the same as leaving the door of your house wide open for everyone to come in.
+Why is it so important to keep a system secure?
+\\
 
-\IEEEPARstart{E}{lektrische} rolstoelen worden door mindervaliden over heel de wereld gebruikt om zichzelf te verplaatsen.
-Doorgaans worden dergelijke rolstoelen bestuurd met een joystick, maar dit is voor sommige mensen geen optie.
-Ze beschikken niet over de benodigde handcontrole om zich veilig te verplaatsen met een dergelijke rolstoel.
-Deze mensen dienen van alternatieve controlemethodes gebruik te maken, of ze worden rondgereden door een verzorger.
+Businesses can have confidential information on clients.
+Through data leakage, confidential information can be lost and possibly used by the competitor.
+Businesses wants to meet their service-level agreements.
+They will protect themselves against disruption that can be caused by DOSS attacks.
+Ultimately, system and network security helps protecting a business's reputation, which is one of its most important assets.
+%\todo{waarom?
+:  lekken van informatie: je hebt info over klanten en moet hun privacy beschermen, je eigen data is geld waard voor concurrenten, DOSS attacks: je will je service-level agreements kunnen nakomen, ..} \\
+ A hacker will be a person that seeks exploits or weaknesses in a system or network in order to gain access.
+Many of those attacks have a different cause.
+Some of the attacks by a hacker can be benign, others can be harmful.
+There are various ways to break into a system.
+Viruses, worms, spyware and other malware are the number two of the top external threats that a business faces, according to the security report kaspersky 2014 \cite{kaspersky} (number one is Spam).
+Furthermore these kind of threats also causes the greatest percentage in loss of data.
+These threats will infect the network by means of a virus that will propagate through the network.
+Most of the attacks are Advanced Persistent Threats (APT).
+\\
 
-\cite{Barea2003}
+An APT is a targeted cyber attack that targets organisations in a stealthy way and stays undetected for a long period.
+This makes it so hard to protect against an APT.
+Bruce Schneier describes an APT as something different and stronger than a conventional hacker: ''\textit{A conventional hacker or criminal isn't interested in any particular target.
+He wants a thousand credit card numbers for fraud, or to break into an account and turn it into a zombie, or whatever.
+Security against this sort of attacker is relative; as long as you're more secure than almost everyone else, the attackers will go after other people, not you.
+An APT is different; it's an attacker who - for whatever reason - wants to attack you.
+Against this sort of attacker, the absolute level of your security is what's important.
+It doesn't matter how secure you are compared to your peers; all that matters is whether you're secure enough to keep him out}'' - Bruce Schneier: APT is a Useful Buzzword.\\
+
+\section{Stealthy takeover model}
+Since it is so difficult to protect a system or a network against APT, researcher have been looking for effective ways to predict in advance which defence strategy might be the better one. 
+Game theory is gaining more and more interest as an effective technique to  model and study Cyber Security.
+Game theory analyses the security problem as a game where the players are an attacker and a defender of a system, and where both players have to make decisions.
+In particular, both players will aim for the strategy that results in a maximal benefit for them.
+Researchers at RSA made a game theoretic framework to model targeted attacks.
+They study the specific scenario where a system or network is repeatedly taken over completely by an attacker and this attack is not immediately detected by the defender of the system or network.
+In game theory, such a game is known as ''FlipIt'' \cite{FlipIt}.
+This is a two players game where the attacker and the defender are competing to get control over a shared resource.
+Both players do not know who is currently in control of the resource until they move.
+In FlipIt every move gives them immediately control over the resource.
+But what if the attacker moves and it takes a while before the attacker gets full control over the resource?
+FlipIt does not take into account that a move may not be instantaneous, but has a certain delay.
+Consider for example a network with different nodes ( laptops, datacenters) as a resource.
+The attacker drops a virus on one of the nodes and then wait till this virus infects the whole network.
+The attacker will only be in control of the resource once the whole network is infected.
+\\
+
+\begin{figure}[hbtp]
+\caption{FlipIt game with a periodic strategy. The defender plays with a period of deltaA = .. The attacker plays with a period $\delta_{A}$ }
+\centering
+\includegraphics[scale=0.5]{../../doc/template/Images/DefFlipIt.png}
+\end{figure}
+
+\section{FlipIt with virus propagation}
+A FlipIt game consists of a single resource. To represent the security problem, the game now defines its single resource as a computer network with multiple
+nodes. One of the players, the defender, will try to defend his network. The defender
+will do this by flipping all the nodes of the network (i.e. the entire resource) in every move he plays. The
+attacker, the other player, will try to infect all the nodes in the network. The attacker
+will do this by flipping the node in the graph that can infect all the nodes in the
+shortest time possible.
+After dropping a virus on the first node, it takes a while for the virus to infect the entire network.
+However, since the original FlipIt game works with a single resource that is always flipped entirely, the assumption is made that the attacker is considered to have gained the control over the resource only when all the nodes of the network have been infected, i.e.
+the entire resource has been flipped.\\
+
+After dropping a virus on the first resource, it takes a while for the virus to infect
+the entire network. The time that it takes for the virus to infect every node will be
+denoted as parameter d. If we want to measure how long it takes for the virus to
+infect all the nodes in the network, we have to calculate the shortest path from the
+first infected node to the farthest node. This can be measured by a method that we
+will explain in section [].
+Assume that an attacker attacks at time t, then only at time t + d he gains control over the entire network.
+If the defender flips the network before the period d has elapsed (so, somewhere between t and t+ d), then the attacker will never gain control over the entire network.
+Using this parameter d, a FlipIt game with virus propagation
+can be modelled. \\
+
+We start by formalizing the non-adaptive continuous basic FlipIt game where players use a periodic strategy with a random phase.
+his choice is motivated by the assumption that in the practical situation of most organisations, the defence strategy is to periodically defend the network.
+This corresponds to a periodic defender strategy.
+To simplify the analysis in a first time, a periodic attacker strategy is assumed as well.
+Further research can investigate the effect of relaxing this assumption.
+\\
+
+Next we split the formalisation for the benefit formula for a FlipIt game with virus propagation in two cases.
+First the benefit formula is derived for the basis case and then it is calculated for the case with a virus propagation.
+The first case is where the defender plays at least as fast as the attacker, the second case is where the attacker plays at least as fast as the defender.
+This subdivision is also used in the FlipIt paper.
+\\
+
+ A Periodic strategy is a non-adaptive renewal strategy where the time intervals between consecutive moves are a fixed period, denoted by $\delta$.
+Moreover it has a random phase, that is chosen uniformly and random in the interval $[0,\delta]$ for the first move.
+The average rate of play of a player is denoted by $\alpha_{i} = \dfrac{1}{\delta_{i}}$.
+\\
+~~\\
+
+\subsection*{\textbf{Case 1:} $\delta_{D} \leq \delta_{A} $ (The defender plays at least as fast as the attacker.) }
+
+Let $r = \dfrac{\delta_{D}}{ \delta_{A} }$.
+The intervals between two consecutive defender's moves have length $\delta_{D}$.
+Consider a given defender move interval.
+The probability over the attacker's phase selection that the attacker moves in this interval is r.
+Given that the attacker moves within the interval, he moves exactly once within the interval (since $\delta_{D} \leq \delta_{A} $) and his move is distributed uniformly at random.
+\\
+
+The expected period of attacker control within the interval would be r/2, without considering the delay by a virus. Therefore the benefit without this virus can be expressed as follows:
+\begin{equation}\label{first}
+\beta_{D}(\alpha_{D},\alpha_{A}) =1 -  \dfrac {r} {2} - k_{D} \alpha_{D} = 1 - \dfrac {\delta_{D}} {2\delta_{A}} - k_{D} \alpha_{D} 
+\end{equation}
+\begin{equation}\label{first}
+\beta_{A}(\alpha_{D},\alpha_{A}) =\dfrac {r} {2} - k_{A} \alpha_{A} = \dfrac {\delta_{D}} {2\delta_{A}} - k_{A} \alpha_{A}  
+\end{equation}\\
+
+However, because of the delay caused by a virus, the maximal time of control is reduced to $\delta_{D}-d$.
+There is a probability of \textit{r} that the attacker will move in the interval of the defender.
+The attacker has to play soon enough to gain control, meaning that the attacker has to play during the period of $\delta_{D}-d$ during the interval of the defender.
+There is $\dfrac{\delta_{D}-d}{\delta_{D}}$ probability that the attacker will move soon enough which gives the attacker a gain of $\dfrac{\delta_{D}-d}{2}$.
+If the attacker moves after the period of $\delta_{D}-d$, the gain of the attacker will be zero.
+The average gain rate of the attacker can be expressed as follows if we look at one interval of the defender:
+\begin{equation}\label{first}
+\beta_{A}(\alpha_{D},\alpha_{A}) = \dfrac {1}{\delta_{D}} [ \dfrac{\delta_{D}}{\delta_{A}} \cdot \dfrac{\delta_{D}-d}{\delta_{D}} \cdot \dfrac{\delta_{D}-d}{2} + \dfrac{\delta_{D}}{\delta_{A}} \cdot \dfrac{d}{\delta_{D}} \cdot 0 ]
+\end{equation}
+
+To complete the formula to derive the benefit function, the cost of moving is added. In the second formula we can see the formula of the original FlipIt game.
+\begin{equation}\label{first}
+\beta_{A}(\alpha_{D},\alpha_{A}) = \dfrac { (\delta_{D}-d) ^{2}} {2 \cdot \delta_{D}  \delta_{A}} - k_{A} \alpha_{A}
+\end{equation}
+\begin{equation}\label{first}
+\beta_{A}(\alpha_{D},\alpha_{A}) = \dfrac { \delta_{D}} {2 \cdot \delta_{A}} - k_{A} \alpha_{A} + \dfrac{d}{\delta_{A}} + \dfrac{d^{2}}{2 \cdot \delta_{A} \delta_{D}}
+\end{equation}
+
+ The benefit of the defender is expressed as follows:
+ \begin{equation}\label{first}
+\beta_{D}(\alpha_{D},\alpha_{A}) = 1 - \dfrac { (\delta_{D}-d) ^{2}} {2 \cdot \delta_{D}  \delta_{A}} - k_{D} \alpha_{D}
+\end{equation}
+~~\\
+
+\subsection*{\textbf{Case 2:} $\delta_{A} \leq \delta_{D} $ (The attacker plays at least as fast as the defender.) }
+
+First let $r = \dfrac{\delta_{D}}{ \delta_{A} }$.
+The intervals between two consecutive attacker's moves have length $\delta_{A}$.
+Consider a given attackers move interval.
+The probability over the attacker's phase selection that the defender moves in this interval is $\dfrac{\delta_{D}}{ \delta_{A} } = (1/r)$.
+Given that the defender moves within the interval, he moves exactly once within the interval (since$\delta_{A} \leq \delta_{D} $) and his move is distributed uniformly at random.
+\\
+
+A similar analysis as in case 1 for a FlipIt game without virus propagation yields the following benefits:
+\begin{equation}\label{first}
+\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac {1} {2r} - k_{D} \alpha_{D} = \dfrac {\delta_{A}} {2\delta_{D}} - k_{D} \alpha_{D} 
+\end{equation}
+\begin{equation}\label{first}
+\beta_{A}(\alpha_{D},\alpha_{A}) =1 - \dfrac {1} {2r} - k_{A} \alpha_{A} = 1- \dfrac {\delta_{A}} {2\delta_{D}} - k_{A} \alpha_{A}  
+\end{equation}\\
+
+For the case with a virus we consider two cases, Case a and Case b, to compute the benefit of both players.  \\
+
+\subsubsection*{\textbf{Case a:} $d + \delta_{A} \leq \delta_{D}$}
+
+Let $d + \delta_{A} \leq \delta_{D}$.
+If $d + \delta_{A}$  is a timespan for the attacker, the defender will never move twice in this timespan.
+$\dfrac{delta_{A}}{\delta_{D}}$ the defender will move during the interval of the attacker.
+When this happens the defender will end with being in control at the end of the interval.
+In the next interval the attacker will move in the beginning but will not gain control after the delay of \textit{d}.
+This means that if the defender moved in the interval prior to this one, the delay is the minimum gain of the defender.
+To calculate the gain rate of the defender we don't look at one interval of the defender.
+We know that if the defender moves within the interval of the defender, he will receive a minimum of $\dfrac{\delta_{A}}{2}$ gain.
+In the next interval the defender will receive a gain of $d$.
+
+This means that every time the defender plays, he will get a gain of $\dfrac{\delta_{A}}{2}$ and the next interval will always receive a gain of $d$.
+To derive the benefit formula for the defender we can calculate the probability that the defender will move during an interval of the attacker and multiply it by the gain rate of the defender plus the gain rate that it will give to the next interval.
+This yields in the following formula:
+\begin{equation}\label{first}
+\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta{A}}{\delta_{D}} \cdot \dfrac{(d+\dfrac{\delta_{A}}{2})}{\delta_{A}} - k_{D} \alpha_{D} 
+\end{equation}
+\begin{equation}\label{first}
+\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - k_{D} \alpha_{D} 
+\end{equation}\\
+
+The benefit for the attacker will be as follows:
+\begin{equation}\label{first}
+\beta_{A}(\alpha_{D},\alpha_{A}) = 1 -\dfrac{\delta{A}}{2\delta_{D}} - \dfrac{d}{\delta_{D}} - k_{A} \alpha_{A} 
+\end{equation}\\
+
+\begin{figure}[hbtp]
+\caption{Case 2 where d + deltaA < deltaD}
+\centering
+\includegraphics[scale=0.7]{../../doc/template/Images/FlipItCase2.png}
+\end{figure}
+
+It is crucial that the $d + \delta_{1}$ is not greater than $ \delta_{0}$.
+If it is greater that means that the defender can move during a delay after an interval of the attacker where the defender already has moved.
+If we calculate the benefit formula in the same manner, to much gain is added.
+Namely the overlap during the delay.
+
+\subsubsection*{\textbf{Case b:} $d + \delta_{1} > \delta_{D}$}
 

### 2015-05-15 doc/template/chap-5.tex
paper

@@
 To get an understanding of the most important concepts of game theory, a short introduction based on the work of
-\cite{leyton2008essentials} and \todo{Coursera} is given in section 2.1 [] \todo{ref}. For a more detailed and full introduction to game theory, the reader is referred to 
-\cite{MitigationCovert}.  In section 2.2 [] \todo{ref} an overview of the FlipIt game is given with the definitions and concepts that will be used throughout the paper.
-The last section [] \todo{ref} will cover the extensions and additions already made on FlipIt.
+\cite{leyton2008essentials} and \cite{Coursera} is given in section \ref{Cha:1:Intro.Game.Theory}  . For a more detailed and full introduction to game theory, the reader is referred to 
+\cite{leyton2008essentials}.  In section \ref{ch:FlipItGame} an overview of the FlipIt game is given with the definitions and concepts that will be used throughout the paper. 
+The last section \ref{ch:extendedWork} will cover the extensions and additions already made on FlipIt.
 
@@
 \section{The FlipIt game}
+\label{ch:FlipItGame}
 In this section, we introduce the ''stealthy takeover '' game \flip{FlipIt} \cite{FlipIt}.
 \flip{FlipIt} is a game introduced by van Dijk et al.
-First we explain the framework of FlipIt and introduce the most important formules that will be used throughout the paper.
+First we explain the framework of FlipIt and introduce the most important formulas that will be used throughout the paper.
 To understand how to model a FlipIt game with virus propagation it is important to get familiar with the concepts of the normal FlipIt game and it's notations.
@@
 \end{equation}
-The average gain of player \textit{i} is defined as:
+The average gain rate of player \textit{i} is defined as:
 \begin{equation}\label{first}
@@
 -- Ik kan dit ook pas uitleggen als ik mijn FlipIt spel met virus propagatie uit de doeken doe -- \\
-We elaborate more about the periodic strategy because we will use this one to model our first extension.
-For more details about the other strategies of FlipIt we refer the reader to [the paper of FlipIt].
+We elaborate more about the periodic strategy because we will use this one to model our first extension. For more details about the other strategies of FlipIt we refer the reader to \citep{FlipIt}.
 
@@
 \section{Extensions on FlipIt}
+\label{ch:extendedWork}
 
@@
 Another addition of Laszka et al.
-to the game of FlipIt [mitigating covert compromises] is extending the game to also consider non-targeted attacks by non-strategic players.
+to the game of FlipIt \cite{MitigationCovert} is extending the game to also consider non-targeted attacks by non-strategic players.
 In this game the defender tries to maintain control over the resource that is subjected to both targeted and non-targeted attacks.
@@
 One of the last important addition from Laszka et al.
-[Mitigation of Targeted and] is to consider a game where the moves made by the attacker are still covert but the moves made by the defender are known to the attacker.
+\cite{MitigationNonTargeted} is to consider a game where the moves made by the attacker are still covert but the moves made by the defender are known to the attacker.
 This means that the attacker can base his attacks on the defender's moves.
@@
 Other authors used the FlipIt game to apply it on a specific scenario. To be able to use the FlipIt game, modifications where required for the FlipIt model.
-One of the scenarios by Pham\cite{GameTheorApprCostBenefitAnalyses} [\todo{citatie needed voor Are We Compromised?}] was to find out whether a resource was compromised or not by the attacker.
+One of the scenarios by Pham \cite{compromised} was to find out whether a resource was compromised or not by the attacker.
 This could be verified by the defender, who has an extra move "test" beside the flip move.
@@
 Finally researchers also have investigated the behavioural of humans when playing FlipIt.
-A Nochenson and Grossklags [A behavioural investigation of the FlipIt game]  investigate how people really act when given temporal decisions.
-[Risk-seeking in a continuous game of timing] Reitter et al.
-they observe continuous games, 20-seconds FlipIt game..
+A Nochenson and Grossklags \cite{nochenson2013behavioral}  investigate how people really act when given temporal decisions.
+Reitter et al.
+\cite{reitter2013risk} observed continuous games, 20-seconds FlipIt game..
 

### 2015-05-15 IEEE Paper/paper/paper.tex
paper nog meer aanpassen

@@
 \usepackage[utf8]{inputenc}
+\usepackage{hyperref}
 
@@
 
-\section{Inleiding}
+\section{Introduction}
 \label{sec:inleiding}
@@
 \section{Stealthy takeover model}
-Since it is so difficult to protect a system or a network against APT, researcher have been looking for effective ways to predict in advance which defence strategy might be the better one. 
+Since it is so difficult to protect a system or a network against APT's, researcher have been looking for effective ways to predict in advance which defence strategy might be the better one. 
 Game theory is gaining more and more interest as an effective technique to  model and study Cyber Security.
@@
 
+\section{The FlipIt game}
+\label{ch:FlipItGame}
+FlipIt is a game introduced by van Dijk et al.
+First we explain the framework of FlipIt and introduce the most important formulas that will be used throughout the paper.
+To understand how to model a FlipIt game with virus propagation it is important to get familiar with the concepts of the normal FlipIt game and it's notations.
+\\
+
+FlipIt is a two-players game with a shared (single) resource that the players want to control as long as possible.
+The shared resource can be a password, a network or a secret key depending on the setting being modelled.
+In the rest of the paper we name the two players the attacker, denoted by the subscript \textit{A} and the defender, denoted by subscript \textit{D}.
+
+The game begins at $t=0$ and continuous indefinitely ($t \rightarrow \infty $).
+The time in the game can be viewed as being continuous, but a discrete time can also be viewed.
+To get control over the resource, the players \textit{i} can flip the resource at any given time.
+A flip will be regarded as a move from a player \textit{i}.
+Each move will imply a certain cost $k_{i}$ and the cost can vary for each player.
+Both players will try to minimize their cost.
+By adding a cost, it will prevent players to move to frequently.
+\\
+
+The unique feature of FlipIt is that every move will happen in a stealthy way, meaning that the player has no clue that (his adversary) the other player has flipped the resource.
+For instance, the defender will not find out if the resource has already been compromised by the attacker, but he can only potentially know it after he flips the resource himself.
+The goal of the player is to maximize the time that he or she has control over the resource while minimizing total cost of the moves.
+A move can also result in a "wasted move", called a flop.
+It may happen that the resource was already under control by the defender.
+If the defender moves when he or she has already control over the resource, he or she would have wasted a move since it does not result in a change of ownership and a cost is involved.
+\\
+
 \begin{figure}[hbtp]
-\caption{FlipIt game with a periodic strategy. The defender plays with a period of deltaA = .. The attacker plays with a period $\delta_{A}$ }
 \centering
-\includegraphics[scale=0.5]{../../doc/template/Images/DefFlipIt.png}
+\includegraphics[scale=0.5]{../../doc/template/Images/DefFlipit}
+\caption{A representation of a FlipIt game where both players are playing periodically and discrete.
+Every move or flip is indicated by a blue or orange circle.
+The attacker is the orange colour and plays with a period of $\delta_{A}=4$.
+The defender is the blue colour and plays with a period of $\delta_{D}=3$.The blue and orange rectangles represent the amount of time one of the players is in control of the resource.}
+\label{fig:FLipItDefault}
 \end{figure}
+
+The state of the resource is denoted as a time independent variable $C=C_{i}(t)$. 
+$C_{D}(t)$ is either 1 if the game is under control by the defender and 0 if the game is under control by the attacker. For $C^{A}(t)$ it is visa versa, $C^{A}(t)= 1 - C^{D}(t)$.\\ \\
+The game starts with defender being in control of the game, $C_{D}(0)= 1$.
+
+The players receive a benefit equal to the time of units that they were in possession of the resource minus the cost of making their moves. The cost of a player \textit{i} is denoted by $k_{i}$. 
+The total gain of player \textit{i} is equal to the total amount of time that a player \textit{i} has owned the resource from the beginning of the game up to time \textit{t}.
+It is expressed as follows:
+\begin{equation}\label{first}
+G_{i}(t) = \int_0^1 \! C_{i}(x) dx.
+\end{equation}
+If we add up the gain of the defender and the gain of the attacker it should sum up to t:
+\begin{equation}\label{first}
+G_{D}(t) + G_{A}(t) = t
+\end{equation}
+The average gain rate of player \textit{i} is defined as:
+\begin{equation}\label{first}
+\gamma_{i}(t) = G_{i}(t)/t.
+\end{equation}
+And thus for all $t > 0$ :
+\begin{equation}\label{first}
+\gamma_{D}(t) + \gamma_{A}(t) = 1
+\end{equation}
+Let $\beta_{i}(t)$ denote player's \textit{i} average benefit upto time \textit{t}:
+\begin{equation}\label{first}
+\beta_{i}(t) = \gamma_{i}(t) - k_{i}\alpha_{i}.
+\end{equation}
+This is equal to the fraction of time the resource has been owned by player \textit{i}, minus the cost of making the moves.
+~$ \alpha_{i}$ defines the average move rate by player \textit{i} up to time \textit{t}.
+In a give game, the asymptotic benefit rate or simply benefit will be defined as the lim inf of the average benefit because time t will increase to infinity and the average benefit may not have limiting values.
+\[ \beta_{i}(t)  = \lim_{t \to \infty} inf \beta_{i}(t)  \]
+\\
+
+\subsubsection{strategies}
+Because the players move in a stealthy way, there are different types of feedback that a player can get while moving.
+These types of feedback can be divided into two groups of strategies.
+The non-adaptive strategies and the adaptive strategies.
+These are described in table \ref{table:Strategies}.
+\\
+
+If there is no need for feedback for both of the players, we say that we have a non-adaptive strategy.
+Because the player does not receive any feedback during the game it will play in the same manner against every opponent.
+They are not dependent on the opponents movements.
+This means that they can already generate the time sequence for all the moves in advance.
+But they can depend on some randomness because the non-adaptive strategies can be randomised.
+In this paper we will focus in the beginning on the non-adaptive strategies.
+Reasons behind this is that a player (defender or attacker) rarely knows what the strategies are of his opponent.
+An interesting subclass of the non-adaptive strategies is one where the time intervals between two consecutive moves are generated by a renewal process.
+Example of such a renewal strategy is the periodic strategy where the time between two consecutive moves of the players are a fixed interval.
+An exponential strategy is a renewal strategy in which the interval between two consecutive moves is exponentially distributed.
+\\
+
+In the category Adaptive strategy there are two sub classes of strategies.
+The first one is the Last move (LM).
+In this class whenever a player flips it will find out the exact time that the opponent played the last time.
+In the second class the Full History (FH), whenever a player flips it will find out the whole history of the opponents move.
+If the opponent player plays with a renewal strategy the sub classes FH and LM collapse.
+\\
+ %n this subsection we elaborate about the strategy of FlipIt . 
+%A renewal strategy is a non-adaptive strategy \\
+
+The game can be extended by the amount of information that a player receives. It can also be possible for a player to get information at the start of the game. Both interesting cases are:
+\begin{itemize}
+\item Rate-of-play (RP): The player finds out the exact rate of play of the opponent.
+\item Knowledge-of-strategy (KS): The player finds out the complete information of the strategy that the opponent is playing.
+\end{itemize}
+
+ \begin{table}
+ \centering
+ \begin{tabular}{ l | c  }
+  \textbf{Categories} & \textbf{Classes of Strategies} \\
+  \hline Non-adaptive (NA) & Exponential \\
+  & Periodic \\
+  & Renewal \\
+  & General non-adaptive \\
+  \hline Adaptive (AD) & Last move (LM) \\
+  & Full History (FH) \\  
+\end{tabular}
+ \caption{Hierarchy of Classes of strategies in FlipIt}
+ \label{table:Strategies}
+ \end{table}
+
+\subsubsection{Imporant results}
+
+\begin{itemize}
+\item periodische spellen domineren de andere renewal strategies
+\item periodische strategien wel slecht tegen attacker die laatste move weet
+\item als de defender snel speelt dan forceert die de attacker om niet meer mee te spelen
+\item iedereen die feedback krijgt heeft meer benefit dan zonder die feedback
+\end{itemize}
 
@@
 infect all the nodes in the network, we have to calculate the shortest path from the
-first infected node to the farthest node. This can be measured by a method that we
-will explain in section [].
+first infected node to the farthest node.
 Assume that an attacker attacks at time t, then only at time t + d he gains control over the entire network.
@@
 We start by formalizing the non-adaptive continuous basic FlipIt game where players use a periodic strategy with a random phase.
-his choice is motivated by the assumption that in the practical situation of most organisations, the defence strategy is to periodically defend the network.
+This choice is motivated by the assumption that in the practical situation of most organisations, the defence strategy is to periodically defend the network.
 This corresponds to a periodic defender strategy.
@@
 If the attacker moves after the period of $\delta_{D}-d$, the gain of the attacker will be zero.
-The average gain rate of the attacker can be expressed as follows if we look at one interval of the defender:
+The average gain rate of the attacker can be expressed as follows if we look at one interval of the defender and average over all the phases:
 \begin{equation}\label{first}
-\beta_{A}(\alpha_{D},\alpha_{A}) = \dfrac {1}{\delta_{D}} [ \dfrac{\delta_{D}}{\delta_{A}} \cdot \dfrac{\delta_{D}-d}{\delta_{D}} \cdot \dfrac{\delta_{D}-d}{2} + \dfrac{\delta_{D}}{\delta_{A}} \cdot \dfrac{d}{\delta_{D}} \cdot 0 ]
+\gamma_{A}(\alpha_{D},\alpha_{A}) = \dfrac {1}{\delta_{D}} [ \dfrac{\delta_{D}}{\delta_{A}} \cdot \dfrac{\delta_{D}-d}{\delta_{D}} \cdot \dfrac{\delta_{D}-d}{2} + \dfrac{\delta_{D}}{\delta_{A}} \cdot \dfrac{d}{\delta_{D}} \cdot 0 ]
 \end{equation}
@@
 \subsubsection*{\textbf{Case a:} $d + \delta_{A} \leq \delta_{D}$}
+~~\\
+If $d + \delta_{A}$  is a timespan for the attacker, the defender will never move twice during this timespan because $d + \delta_{A} \leq \delta_{D}$.
+With a probability of $\dfrac{\delta_{A}}{\delta_{D}} $ the defender move will during the interval of the attacker.
+When this happens the defender will end with being in control at the end of the interval.
+In the next interval the attacker will move in the beginning but will not gain control because of the delay, see figure \ref{fig:case2} cases (1) and (2).
+This means that the defender will keep the control over the resource in the next interval over a period of the delay, namely \textit{d}.
+Because $d + \delta_{A} \leq \delta_{D}$ the next move of the defender in this second interval will not be during the delay.
 
-Let $d + \delta_{A} \leq \delta_{D}$.
-If $d + \delta_{A}$  is a timespan for the attacker, the defender will never move twice in this timespan.
-$\dfrac{delta_{A}}{\delta_{D}}$ the defender will move during the interval of the attacker.
-When this happens the defender will end with being in control at the end of the interval.
-In the next interval the attacker will move in the beginning but will not gain control after the delay of \textit{d}.
-This means that if the defender moved in the interval prior to this one, the delay is the minimum gain of the defender.
-To calculate the gain rate of the defender we don't look at one interval of the defender.
+%To calculate the gain rate we only need to calculate the number of moves that the defender moves in the interval of the attacker.
 We know that if the defender moves within the interval of the defender, he will receive a minimum of $\dfrac{\delta_{A}}{2}$ gain.
@@
 
-This means that every time the defender plays, he will get a gain of $\dfrac{\delta_{A}}{2}$ and the next interval will always receive a gain of $d$.
-To derive the benefit formula for the defender we can calculate the probability that the defender will move during an interval of the attacker and multiply it by the gain rate of the defender plus the gain rate that it will give to the next interval.
-This yields in the following formula:
+Every time the defender plays, he will get a gain rate of $\dfrac{\delta_{A}}{2\delta_{A}}$ and the next interval will always receive a gain rate of $\dfrac{d}{\delta_{A}}$.
+To derive the total gain formula for the defender we can calculate the probability that the defender will move during an interval of the attacker and multiply it by the gain rate of the defender plus the gain rate that it will give to the next interval.
+
 \begin{equation}\label{first}
-\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta{A}}{\delta_{D}} \cdot \dfrac{(d+\dfrac{\delta_{A}}{2})}{\delta_{A}} - k_{D} \alpha_{D} 
+\gamma_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta_{A}}{\delta_{D}} \cdot \dfrac{(d+\dfrac{\delta_{A}}{2})}{\delta_{A}} 
 \end{equation}
 \begin{equation}\label{first}
-\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - k_{D} \alpha_{D} 
+\gamma_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} 
+\end{equation}\\
+This yields in the following benefit formula:
+\begin{equation}\label{first}
+\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - k_{D} \alpha_{D} 
 \end{equation}\\
@@
 \begin{equation}\label{first}
-\beta_{A}(\alpha_{D},\alpha_{A}) = 1 -\dfrac{\delta{A}}{2\delta_{D}} - \dfrac{d}{\delta_{D}} - k_{A} \alpha_{A} 
+\beta_{A}(\alpha_{D},\alpha_{A}) = 1 -\dfrac{\delta_{A}}{2\delta_{D}} - \dfrac{d}{\delta_{D}} - k_{A} \alpha_{A} 
 \end{equation}\\
@@
 \centering
-\includegraphics[scale=0.7]{../../doc/template/Images/FlipItCase2.png}
+\includegraphics[scale=0.5]{../../doc/template/Images/FlipItCase2.png}
+\label{fig:case2}
 \end{figure}
@@
 Namely the overlap during the delay.
+\\
+~~\\
 
-\subsubsection*{\textbf{Case b:} $d + \delta_{1} > \delta_{D}$}
+\subsubsection*{\textbf{Case b:} $d + \delta_{A} > \delta_{D}$}
+~~~\\
+
+\section{Simulations and results}
+
+\section{Related work on FlipIt}
+\label{ch:extendedWork}
+
+There a various possible ways to extend FlipIt. 
+Laszka et al.
+made a lot of additions and extensions on the original game of FlipIt.
+For instance Laszka et al.
+extended the basic FlipIt game to multiple resources.
+The incentive is that for compromising a system in a real case it needs more than just taking over just one resource.
+An example is that one resource can be gaining access to a system and breaking the password of the system is another resource.
+The model is called FlipThem \cite{FlipThem}
+.
+They use two ways to flip the multiple resources: the AND and the OR control model.
+In the AND model the attacker only controls the system if he controls all the resources of the system, whereas in the OR model the attacker only needs to compromise one resource to be in control of the entire system.
+\\
+
+Another addition of Laszka et al. to the game of FlipIt \cite{MitigationCovert} 
+is extending the game to also consider non-targeted attacks by non-strategic players.
+In this game the defender tries to maintain control over the resource that is subjected to both targeted and non-targeted attacks.
+Non-targeted attacks can include phishing, while targeted attacks may include threats delivered through zero day attack vulnerabilities.
+\\
+One of the last important addition from Laszka et al.
+\cite{MitigationNonTargeted} is to consider a game where the moves made by the attacker are still covert but the moves made by the defender are known to the attacker.
+This means that the attacker can base his attacks on the defender's moves.
+Both the targeted and non-targeted attacks don't succeed immediately.
+For the targeted attack the time till it succeeds is given by an exponential distributed random variable with a known rate.
+The non-targeted attacks are modelled as a single attacker and the time till it succeeds is given by a Poisson process.
+The conclusion of this paper is that the optimal strategy for the defender is moving periodically.
+\\
+
+Other authors used the FlipIt game to apply it on a specific scenario. To be able to use the FlipIt game, modifications where required for the FlipIt model.
+One of the scenarios by Pham \cite{compromised} was to find out whether a resource was compromised or not by the attacker.
+This could be verified by the defender, who has an extra move "test" beside the flip move.
+The basic idea is to test with an extra action if the resource has been compromised or not.
+This move involves also an extra cost.
+This model is useful if somebody wants to know for example if his or her password has been compromised.
+\\
+
+Finally researchers also have investigated the behavioural of humans when playing FlipIt.
+A Nochenson and Grossklags \cite{nochenson2013behavioral}  investigate how people really act when given temporal decisions.
+Reitter et al.
+\cite{reitter2013risk} observed continuous games, 20-seconds FlipIt game..
 

### 2015-05-17 IEEE Paper/paper/paper.tex
Paper aanpassen 1

@@
 It is the same as leaving the door of your house wide open for everyone to come in.
-Why is it so important to keep a system secure?
 \\
 
-Businesses can have confidential information on clients.
-Through data leakage, confidential information can be lost and possibly used by the competitor.
-Businesses wants to meet their service-level agreements.
-They will protect themselves against disruption that can be caused by DOSS attacks.
+Why is it so important to keep a system secure?
+Many businesses store confidential information on clients, which can be lost and possible be abused by competitors through data leakage.
+Also, disruption caused by DOSS attacks, may restuls in businesses failing to meet their service-level agreements.
 Ultimately, system and network security helps protecting a business's reputation, which is one of its most important assets.
-%\todo{waarom?
-:  lekken van informatie: je hebt info over klanten en moet hun privacy beschermen, je eigen data is geld waard voor concurrenten, DOSS attacks: je will je service-level agreements kunnen nakomen, ..} \\
- A hacker will be a person that seeks exploits or weaknesses in a system or network in order to gain access.
+\\
+% A hacker will be a person that seeks exploits or weaknesses in a system or network in order to gain access.
 Many of those attacks have a different cause.
@@
 
-An APT is a targeted cyber attack that targets organisations in a stealthy way and stays undetected for a long period.
-This makes it so hard to protect against an APT.
-Bruce Schneier describes an APT as something different and stronger than a conventional hacker: ''\textit{A conventional hacker or criminal isn't interested in any particular target.
+A particular kind of frequently occurring threaths are Advanced Persistent Threats (APT).
+An APT is a targeted cyber attack that targets organisations in a stealthy way and that can stay undetected for a long period.
+This makes it so hard to protect a network or a system against an APT.
+Bruce Schneier describes an APT as something different and stronger than a conventional threat: ''\textit{A conventional hacker or criminal isn't interested in any particular target.
 He wants a thousand credit card numbers for fraud, or to break into an account and turn it into a zombie, or whatever.
@@
 Against this sort of attacker, the absolute level of your security is what's important.
-It doesn't matter how secure you are compared to your peers; all that matters is whether you're secure enough to keep him out}'' - Bruce Schneier: APT is a Useful Buzzword.\\
+It doesn't matter how secure you are compared to your peers; all that matters is whether you're secure enough to keep him out}'' - Bruce Schneier: APT is a Useful Buzzword [].\\
+% hier nog de structuur van de paper inleiden.
 
 \section{Stealthy takeover model}
-Since it is so difficult to protect a system or a network against APT's, researcher have been looking for effective ways to predict in advance which defence strategy might be the better one. 
-Game theory is gaining more and more interest as an effective technique to  model and study Cyber Security.
+Since it is so difficult to protect a system or a network against APT's, researchers have been looking for effective ways to predict in advance which defence strategy might be the better one. 
+Game theory is gaining increasing interest as an effective technique to model and study Cyber Security.
 Game theory analyses the security problem as a game where the players are an attacker and a defender of a system, and where both players have to make decisions.
@@
 FlipIt is a game introduced by van Dijk et al.
-First we explain the framework of FlipIt and introduce the most important formulas that will be used throughout the paper.
-To understand how to model a FlipIt game with virus propagation it is important to get familiar with the concepts of the normal FlipIt game and it's notations.
+To understand how to model a FlipIt game with virus propagation it is important to get familiar with the concepts of the normal FlipIt game and its notations.
+Therefore, we first explain the framework of FlipIt and introduce the most important formulas that will be used throughout the paper.
 \\
 
-FlipIt is a two-players game with a shared (single) resource that the players want to control as long as possible.
+FlipIt is a two-players game with a shared single resource that the players want to control as long as possible.
 The shared resource can be a password, a network or a secret key depending on the setting being modelled.
-In the rest of the paper we name the two players the attacker, denoted by the subscript \textit{A} and the defender, denoted by subscript \textit{D}.
+In the remainder of the paper we name the two players the attacker, denoted by the subscript \textit{A} and the defender, denoted by subscript \textit{D}.
 
-The game begins at $t=0$ and continuous indefinitely ($t \rightarrow \infty $).
-The time in the game can be viewed as being continuous, but a discrete time can also be viewed.
-To get control over the resource, the players \textit{i} can flip the resource at any given time.
+The game begins at $t=0$ and continues indefinitely ($t \rightarrow \infty $).
+The time in the game is assumed as being continuous, but a discrete time could also be considered.
+To get control over the resource, the players $i$, with $i \in \{A,D\}$, can flip the resource at any given time.
 A flip will be regarded as a move from a player \textit{i}.
@@
 Both players will try to minimize their cost.
-By adding a cost, it will prevent players to move to frequently.
+Adding a cost will prevent players to move too frequently.
 \\
 
-The unique feature of FlipIt is that every move will happen in a stealthy way, meaning that the player has no clue that (his adversary) the other player has flipped the resource.
-For instance, the defender will not find out if the resource has already been compromised by the attacker, but he can only potentially know it after he flips the resource himself.
-The goal of the player is to maximize the time that he or she has control over the resource while minimizing total cost of the moves.
+The unique feature of FlipIt is that every move will happen in a stealthy way, meaning that the player has no clue that the other player (his adversary) has flipped the resource.
+For instance, the defender will not find out if the resource has already been compromised by the attacker until he flips the resource himself.
+The goal of the player is to maximize the time that he or she has control over the resource while minimizing the total cost of the moves.
 A move can also result in a "wasted move", called a flop.
-It may happen that the resource was already under control by the defender.
-If the defender moves when he or she has already control over the resource, he or she would have wasted a move since it does not result in a change of ownership and a cost is involved.
+It may happen that the resource was already under control by the player.
+If the player moves when he or she has already control over the resource, he or she would have wasted a move since it does not result in a change of ownership, so the cost is wasted.
 \\
@@
 \includegraphics[scale=0.5]{../../doc/template/Images/DefFlipit}
-\caption{A representation of a FlipIt game where both players are playing periodically and discrete.
+\caption{A representation of a FlipIt game where both players are playing periodically and at discrete time intervals.
 Every move or flip is indicated by a blue or orange circle.
-The attacker is the orange colour and plays with a period of $\delta_{A}=4$.
-The defender is the blue colour and plays with a period of $\delta_{D}=3$.The blue and orange rectangles represent the amount of time one of the players is in control of the resource.}
+The attacker is represented in orange and plays with a period of $\delta_{A}=4$.
+The defender is represented in blue and plays with a period of $\delta_{D}=3$.
+The blue and orange rectangles represent the amount of time the respective player is in control of the resource.}
 \label{fig:FLipItDefault}
@@
 
-The state of the resource is denoted as a time independent variable $C=C_{i}(t)$. 
-$C_{D}(t)$ is either 1 if the game is under control by the defender and 0 if the game is under control by the attacker. For $C^{A}(t)$ it is visa versa, $C^{A}(t)= 1 - C^{D}(t)$.\\ \\
-The game starts with defender being in control of the game, $C_{D}(0)= 1$.
+The state of the resource is denoted as a time-dependent variable $C=C_{i}(t)$. 
+$C_{D}(t)$ is 1 if the game is under control by the defender and 0 if the game is under control by the attacker.
+Reversely, $C_{A}(t)$ will be 1 if the game is under control by the attacker and 0 if under controle by the defender.
+So, $C_{A}(t)= 1 - C_{D}(t)$.
+The game starts with the defender being in control: $C_{D}(0)= 1$. \\
 
-The players receive a benefit equal to the time of units that they were in possession of the resource minus the cost of making their moves. The cost of a player \textit{i} is denoted by $k_{i}$. 
+The players receive a benefit equal to the time units they were in possession of the resource minus the cost of making their moves. The cost of a player \textit{i} is denoted by $k_{i}$. 
 The total gain of player \textit{i} is equal to the total amount of time that a player \textit{i} has owned the resource from the beginning of the game up to time \textit{t}.
@@
 \begin{equation}\label{first}
-G_{i}(t) = \int_0^1 \! C_{i}(x) dx.
+G_{i}(t) = \int_0^t \! C_{i}(x) dx.
 \end{equation}
@@
 ~$ \alpha_{i}$ defines the average move rate by player \textit{i} up to time \textit{t}.
-In a give game, the asymptotic benefit rate or simply benefit will be defined as the lim inf of the average benefit because time t will increase to infinity and the average benefit may not have limiting values.
-\[ \beta_{i}(t)  = \lim_{t \to \infty} inf \beta_{i}(t)  \]
+In a given game, the asymptotic benefit rate (or simply benefit) will be defined as the lim inf of the average benefit because time t will increase to infinity and the average benefit may not have limiting values.
+\begin{equation}
+\beta_{i}(t)  = \lim_{t \to \infty} inf \beta_{i}(t) 
+\end{equation}
 \\
@@
 
-If there is no need for feedback for both of the players, we say that we have a non-adaptive strategy.
-Because the player does not receive any feedback during the game it will play in the same manner against every opponent.
-They are not dependent on the opponents movements.
-This means that they can already generate the time sequence for all the moves in advance.
-But they can depend on some randomness because the non-adaptive strategies can be randomised.
-In this paper we will focus in the beginning on the non-adaptive strategies.
-Reasons behind this is that a player (defender or attacker) rarely knows what the strategies are of his opponent.
-An interesting subclass of the non-adaptive strategies is one where the time intervals between two consecutive moves are generated by a renewal process.
-Example of such a renewal strategy is the periodic strategy where the time between two consecutive moves of the players are a fixed interval.
+If there is no feedback for neither of the players, we have a non-adaptive strategy.
+Because a player does not receive any feedback during the game he will play in the same manner against every opponent.
+The strategy is called non-adaptive because the playing strategy is not dependent on the opponents movements.
+An interesting subclass of the non-adaptive strategies is the one where the time intervals between two consecutive moves are generated by a renewal process.
+An example of such renewal strategy is the periodic strategy where the time between two consecutive moves of the players are a fixed interval.
 An exponential strategy is a renewal strategy in which the interval between two consecutive moves is exponentially distributed.
 \\
-
-In the category Adaptive strategy there are two sub classes of strategies.
-The first one is the Last move (LM).
-In this class whenever a player flips it will find out the exact time that the opponent played the last time.
-In the second class the Full History (FH), whenever a player flips it will find out the whole history of the opponents move.
-If the opponent player plays with a renewal strategy the sub classes FH and LM collapse.
+In case there is feedback, a player can adapt his strategy to the information received about the opponent's moves.
+Depending on the amount of information received, two subclasses of adaptive strategies can be identified.
+The Last Move (LM) strategies represent the class where whenever a player flips he will find out the exact time that the opponent played the last time.
+In the second class, called Full History (FH), whenever a player flips he will find out the whole history of the opponent's move.
 \\
- %n this subsection we elaborate about the strategy of FlipIt . 
-%A renewal strategy is a non-adaptive strategy \\
-
-The game can be extended by the amount of information that a player receives. It can also be possible for a player to get information at the start of the game. Both interesting cases are:
-\begin{itemize}
-\item Rate-of-play (RP): The player finds out the exact rate of play of the opponent.
-\item Knowledge-of-strategy (KS): The player finds out the complete information of the strategy that the opponent is playing.
-\end{itemize}
+In this paper we will focus on the non-adaptive strategies.
+This choice is motivated by the fact that in a security game a player (defender or attacker) rarely has information about the moves (last move or full history) of his opponent.
+\\
 
@@
   \textbf{Categories} & \textbf{Classes of Strategies} \\
-  \hline Non-adaptive (NA) & Exponential \\
-  & Periodic \\
-  & Renewal \\
+  \hline Non-adaptive (NA) & Renewal \\
+  &  ~~~~ Periodic \\
+  &  ~~~~ Exponential \\
   & General non-adaptive \\
@@
 
-\subsubsection{Imporant results}
-
+The study of the different strategies by means of FlipIt framework allows to derive a number of interesting results:  
 \begin{itemize}
-\item periodische spellen domineren de andere renewal strategies
-\item periodische strategien wel slecht tegen attacker die laatste move weet
-\item als de defender snel speelt dan forceert die de attacker om niet meer mee te spelen
-\item iedereen die feedback krijgt heeft meer benefit dan zonder die feedback
+\item periodic games dominate the other renewal strategies, meaning that it is always advantagous to play periodically against an opponent with a renewal strategy;
+\item periodic games are disadvantagous against players following a Last Move adaptive strategy;
+\item if the defender plays with a periodic rate that is fast enough he'll force the attacker to drop out;
+\item any amount of feedback about the opponent received during the game, benefits to a player.
 \end{itemize}
@@
 will do this by flipping the node in the graph that can infect all the nodes in the
-shortest time possible.
+shortest possible time.
 After dropping a virus on the first node, it takes a while for the virus to infect the entire network.
-However, since the original FlipIt game works with a single resource that is always flipped entirely, the assumption is made that the attacker is considered to have gained the control over the resource only when all the nodes of the network have been infected, i.e.
-the entire resource has been flipped.\\
+However, since the original FlipIt game works with a single resource that is always flipped entirely, the assumption is made that the attacker is considered to gain immediate full control over the resource when the network has been infected, even it is only one node that has been infected.\\
 
-After dropping a virus on the first resource, it takes a while for the virus to infect
-the entire network. The time that it takes for the virus to infect every node will be
-denoted as parameter d. If we want to measure how long it takes for the virus to
+In reality however, after dropping a virus on the first node, it takes a while for the virus to infect
+the entire network.
+So, the assumption that the attacker has full control over the resource as soon as a node has been infected, is not realistic.
+The attacker has only control of the network once all or a sufficient large number of nodes are infected.
+The time that it takes for the virus to infect every node (or a sufficient number of nodes) will be
+denoted as an infection-delay variable \textit{d} (called 'delay' for short in the remainder of this paper). If we want to measure how long it takes for the virus to
 infect all the nodes in the network, we have to calculate the shortest path from the
 first infected node to the farthest node.
-Assume that an attacker attacks at time t, then only at time t + d he gains control over the entire network.
-If the defender flips the network before the period d has elapsed (so, somewhere between t and t+ d), then the attacker will never gain control over the entire network.
-Using this parameter d, a FlipIt game with virus propagation
-can be modelled. \\
+Rather than denoting the time needed for infecting \textit{all} the nodes, the variable $d$ can also be used to denote the time needed to infect \textit{a sufficient large number} of nodes.
 
-We start by formalizing the non-adaptive continuous basic FlipIt game where players use a periodic strategy with a random phase.
-This choice is motivated by the assumption that in the practical situation of most organisations, the defence strategy is to periodically defend the network.
-This corresponds to a periodic defender strategy.
-To simplify the analysis in a first time, a periodic attacker strategy is assumed as well.
-Further research can investigate the effect of relaxing this assumption.
+Assume that an attacker attacks at time \textit{t}, he doesn't get immediate control over the resource, but he only gains control at time \textit{t + d}, with $d$ denoting the time needed to infect a sufficiently large number (or all) nodes.
+If the defender flips the network before the period $d$ has elapsed (so, somewhere between $t$ and $t + d$), then the attacker will never gain full control over the resource.
+This implies that the mathematical formulas for gain and benefit need to be adapted to the fact that the attacker loses part of its benefit because of this delay.
+In the remainder of this paper, we will adapt the formalization of the FlipIt game using the variable $d$.
 \\
 
-Next we split the formalisation for the benefit formula for a FlipIt game with virus propagation in two cases.
-First the benefit formula is derived for the basis case and then it is calculated for the case with a virus propagation.
+The formalization starts from the model of the non-adaptive continuous basic FlipIt game where players use a periodic strategy with a random phase.
+This choice is motivated by the assumption that in most organisations, the defence strategy is to periodically defend the network.
+This corresponds to a periodic defender strategy.
+A periodic attacker strategy is assumed as well, as this also corresponds to a common real life strategy.
+%nog eventueel verder te motiveren
+Further research can investigate the effect of relaxing this assumption. \\
+
+Similarly as in [], we split the formalization in two cases.
 The first case is where the defender plays at least as fast as the attacker, the second case is where the attacker plays at least as fast as the defender.
-This subdivision is also used in the FlipIt paper.
+For each of these cases, first the benefit formula of the basic case without delay is presented, and then the delay is introduced.
 \\
 
+\subsection{Formalization the benefit formula including the infection-delay}
  A Periodic strategy is a non-adaptive renewal strategy where the time intervals between consecutive moves are a fixed period, denoted by $\delta$.
@@
 
-The expected period of attacker control within the interval would be r/2, without considering the delay by a virus. Therefore the benefit without this virus can be expressed as follows:
-\begin{equation}\label{first}
-\beta_{D}(\alpha_{D},\alpha_{A}) =1 -  \dfrac {r} {2} - k_{D} \alpha_{D} = 1 - \dfrac {\delta_{D}} {2\delta_{A}} - k_{D} \alpha_{D} 
-\end{equation}
+The expected period of attacker control within the interval would be r/2, without considering the delay by a virus.
+Therefore the benefit for the attacker, without considering the delay, can be expressed as follows:
+
 \begin{equation}\label{first}
@@
 
-However, because of the delay caused by a virus, the maximal time of control is reduced to $\delta_{D}-d$.
+Correspondeingly, the benefit for the defender can be expressed as:
+\begin{equation}\label{first}
+\beta_{D}(\alpha_{D},\alpha_{A}) =1 -  \dfrac {r} {2} - k_{D} \alpha_{D} = 1 - \dfrac {\delta_{D}} {2\delta_{A}} - k_{D} \alpha_{D} 
+\end{equation}
+
+However, because of the delay required for virus propagation, the maximal time of control is reduced to $\delta_{D}-d$.
 There is a probability of \textit{r} that the attacker will move in the interval of the defender.
-The attacker has to play soon enough to gain control, meaning that the attacker has to play during the period of $\delta_{D}-d$ during the interval of the defender.
-There is $\dfrac{\delta_{D}-d}{\delta_{D}}$ probability that the attacker will move soon enough which gives the attacker a gain of $\dfrac{\delta_{D}-d}{2}$.
+However, the gain will not be half of the interval.
+Indeed, the attacker has to play soon enough to gain control, meaning that the attacker has to play during the period of $\delta_{D}-d$ during the interval of the defender.
+The probability that the attacker plays soon enough is $\dfrac{\delta_{D}-d}{\delta_{D}}$ and this will give the attacker an avarage gain of $\dfrac{\delta_{D}-d}{2}$.
 If the attacker moves after the period of $\delta_{D}-d$, the gain of the attacker will be zero.
-The average gain rate of the attacker can be expressed as follows if we look at one interval of the defender and average over all the phases:
+The probability that this happens is  $\dfrac{d}{\delta_{D}}$.
+The average gain rate of the attacker can then be expressed as follows if we look at one interval of the defender:
 \begin{equation}\label{first}
@@
 
-To complete the formula to derive the benefit function, the cost of moving is added. In the second formula we can see the formula of the original FlipIt game.
+To derive the benefit, the cost of moving is substracted from the average gain. 
 \begin{equation}\label{first}
@@
 ~~\\
+We can easily see that when $d$=0, we obtain the formula of the original FlipIt game.\\
 

### 2015-05-17 IEEE Paper/paper/paper.tex
Paper aanpassen 2

@@
 \begin{abstract}
-Recently the attacks on Belgacom and other high profile targeted attacks shows us that even
-the most secure companies can still be compromised. It also shows that these attacks are
-not immediately detected. FlipIt is a framework that can model these stealthy takeovers and is
-proposed by a group of researchers at RSA. It is a 2-players game composed of a single attacker,
-a single defender and a single shared resource. These players will try to gain control over the
-shared resource and they do this in a stealthy way. In this paper we try to adapt FlipIt in a
-way that we can use it to model the game of defending a company network that is attacked by a virus. The FlipIt formulas are adapted for this virus propagation. Through analytic results we can provide
-useful information for the defender to defend his network.
+Recently, high profile targeted attacks such as the attack on Belgacom (a major Belgian telcom), have demonstrated that even the most secure
+companies can still be compromised, and that moreover such attacks can go undetected for a while. 
+FlipIt has been proposed by a group of
+researchers at RSA to model such stealthy takeovers. It is a 2-players game composed of a single
+attacker, a single defender and a single shared resource. The
+players will try to gain control over the shared resource and they
+do this in a stealthy way. FlipIt does however not take into account that a move may not be instantaneous, but has a certain delay. In this paper we adapt FlipIt
+such that we can use it to model the game of defending a
+company network that is attacked by a virus. The FlipIt formulas
+are adapted such as to take the delay for virus propagation into account. 
 \end{abstract}
@@
 It doesn't matter how secure you are compared to your peers; all that matters is whether you're secure enough to keep him out}'' - Bruce Schneier: APT is a Useful Buzzword [].\\
-% hier nog de structuur van de paper inleiden.
 
-\section{Stealthy takeover model}
 Since it is so difficult to protect a system or a network against APT's, researchers have been looking for effective ways to predict in advance which defence strategy might be the better one. 
@@
 \\
+
+This paper proposes the adaptation of the FlipIt formulas as presented in \cite{FlipIt} such as to take the delay for virus propagation into account.
+In the next section we first present the original FlipIt game.
+Then section [] presents the FlipIt game with virus propagation.
+Section [] presents some related work.
+Section [] concludes the paper and presents avenues for further research.
 
@@
 \begin{equation}\label{first}
-\beta_{A}(\alpha_{D},\alpha_{A}) = \dfrac { \delta_{D}} {2 \cdot \delta_{A}} - k_{A} \alpha_{A} + \dfrac{d}{\delta_{A}} + \dfrac{d^{2}}{2 \cdot \delta_{A} \delta_{D}}
+\beta_{A}(\alpha_{D},\alpha_{A}) = \dfrac { \delta_{D}} {2 \cdot \delta_{A}} - k_{A} \alpha_{A} - (\dfrac{d}{\delta_{A}} - \dfrac{d^{2}}{2 \cdot \delta_{A} \delta_{D}})
 \end{equation}
@@
 Consider a given attackers move interval.
-The probability over the attacker's phase selection that the defender moves in this interval is $\dfrac{\delta_{D}}{ \delta_{A} } = (1/r)$.
-Given that the defender moves within the interval, he moves exactly once within the interval (since$\delta_{A} \leq \delta_{D} $) and his move is distributed uniformly at random.
+The probability over the attacker's phase selection that the defender moves in this interval is $\dfrac{\delta_{A}}{ \delta_{D} } = (1/r)$.
+Given that the defender moves within the interval of the attacker, he moves exactly once within this interval (since $\delta_{A} \leq \delta_{D} $) and his move is distributed uniformly at random.
 \\
@@
 A similar analysis as in case 1 for a FlipIt game without virus propagation yields the following benefits:
+
 \begin{equation}\label{first}
@@
 
-For the case with a virus we consider two cases, Case a and Case b, to compute the benefit of both players.  \\
+For the case with a virus we consider two cases, Case a and Case b, depending on whether the delay is shorter or longer than the difference between the attacker's and the defender's period.  \\
 
@@
 ~~\\
-If $d + \delta_{A}$  is a timespan for the attacker, the defender will never move twice during this timespan because $d + \delta_{A} \leq \delta_{D}$.
-With a probability of $\dfrac{\delta_{A}}{\delta_{D}} $ the defender move will during the interval of the attacker.
+Consider a timespan $\delta_{A} + d$, representing the attacker's interval followed by the delay period in his next interval.
+The defender will never move twice during this timespan because $\delta_{A} + d \leq \delta_{D}$.
+The defender will move during the interval of the attacker with a probability of $\dfrac{\delta_{A}}{\delta_{D}} $.
 When this happens the defender will end with being in control at the end of the interval.
-In the next interval the attacker will move in the beginning but will not gain control because of the delay, see figure \ref{fig:case2} cases (1) and (2).
+In the next interval the attacker will have to regain control, meaning that during the delay, the defender stays in control, see figure \ref{fig:case2} cases (1) and (2).
 This means that the defender will keep the control over the resource in the next interval over a period of the delay, namely \textit{d}.
-Because $d + \delta_{A} \leq \delta_{D}$ the next move of the defender in this second interval will not be during the delay.
+Because $d + \delta_{A} \leq \delta_{D}$ the next move of the defender in this second interval will never occur during the delay, meaning that the entire delay can be considered as an extra benefit resulting of a play in the previous interval.
+So, every time the defender plays, he will get an average gain of $\dfrac{\delta_{A}}{2}$ in the interval where he plays and in the next interval will always receive a extra gain of $d$, yielding a total average gain per interval of
+$\dfrac{(d+\dfrac{\delta_{A}2})}{\delta_{A}}$
 
-%To calculate the gain rate we only need to calculate the number of moves that the defender moves in the interval of the attacker.
-We know that if the defender moves within the interval of the defender, he will receive a minimum of $\dfrac{\delta_{A}}{2}$ gain.
-In the next interval the defender will receive a gain of $d$.
-
-Every time the defender plays, he will get a gain rate of $\dfrac{\delta_{A}}{2\delta_{A}}$ and the next interval will always receive a gain rate of $\dfrac{d}{\delta_{A}}$.
-To derive the total gain formula for the defender we can calculate the probability that the defender will move during an interval of the attacker and multiply it by the gain rate of the defender plus the gain rate that it will give to the next interval.
+The total gain  rate of the defender is then the probability that the defender will move during an interval of the attacker multiplied by the total average gain per interval:
 
@@
 
-It is crucial that the $d + \delta_{1}$ is not greater than $ \delta_{0}$.
-If it is greater that means that the defender can move during a delay after an interval of the attacker where the defender already has moved.
-If we calculate the benefit formula in the same manner, to much gain is added.
-Namely the overlap during the delay.
+It is crucial that $ \delta_{D}$ is at least as large as $d + \delta_{A}$.
+If not, this would mean that the defender can move during the delay in the interval following the interval where the defender already moved.
+This would mean that there can be an overlap between the average gain of $\dfrac{\delta_{a}}{2}$ and the delay.
+The above benefit formula would then include to much gain for the defender: the potential overlap during the delay would be counted twice.
 \\
-~~\\
-
+To obtain the formula in case of a too long delay, we therefore need to substract this overlapping gain from the above formula.  The overlapping gain is equal to 
+~~ \\
 \subsubsection*{\textbf{Case b:} $d + \delta_{A} > \delta_{D}$}
 ~~~\\
-
-\section{Simulations and results}
 
@@
 
-There a various possible ways to extend FlipIt. 
+Various possible ways to extend FlipIt have already been proposed. 
 Laszka et al.
-made a lot of additions and extensions on the original game of FlipIt.
+made a lot of additions and extensions to the original game of FlipIt.
 For instance Laszka et al.
 extended the basic FlipIt game to multiple resources.
-The incentive is that for compromising a system in a real case it needs more than just taking over just one resource.
-An example is that one resource can be gaining access to a system and breaking the password of the system is another resource.
-The model is called FlipThem \cite{FlipThem}
-.
-They use two ways to flip the multiple resources: the AND and the OR control model.
+The rationale is that for compromising a system in real life, more than just one resource needs to be taken over.
+An example is that gaining access to deeper layers of a system may require breaking several passwords.
+The model is called FlipThem \cite{FlipThem}.
+Laszka et al.
+also use two ways to flip the multiple resources: the AND and the OR control model.
 In the AND model the attacker only controls the system if he controls all the resources of the system, whereas in the OR model the attacker only needs to compromise one resource to be in control of the entire system.
@@
 \\
-One of the last important addition from Laszka et al.
+One of the last important additions from Laszka et al.
 \cite{MitigationNonTargeted} is to consider a game where the moves made by the attacker are still covert but the moves made by the defender are known to the attacker.
@@
 Both the targeted and non-targeted attacks don't succeed immediately.
-For the targeted attack the time till it succeeds is given by an exponential distributed random variable with a known rate.
+[verschil uitleggen] For the targeted attack the time till it succeeds is given by an exponential distributed random variable with a known rate.
 The non-targeted attacks are modelled as a single attacker and the time till it succeeds is given by a Poisson process.
@@
 The basic idea is to test with an extra action if the resource has been compromised or not.
-This move involves also an extra cost.
-This model is useful if somebody wants to know for example if his or her password has been compromised.
-\\
+This move involves also an extra cost.\\
 
-Finally researchers also have investigated the behavioural of humans when playing FlipIt.
-A Nochenson and Grossklags \cite{nochenson2013behavioral}  investigate how people really act when given temporal decisions.
+Finally researchers also have investigated the behaviour of humans playing FlipIt.
+A.
+Nochenson and Grossklags \cite{nochenson2013behavioral}  investigate how people really act when given temporal decisions.
 Reitter et al.
-\cite{reitter2013risk} observed continuous games, 20-seconds FlipIt game..
+\cite{reitter2013risk} observed continuous games, 20-seconds FlipIt game.
+
+\section{Conclusions and Furhter Research}
+\label{ch:conclusion}
+In this paper we presented an adaptation of the FlipIt game to the situation of virus propagation, such as to take the delay for network infection into account.
+We discerned two cases.
+In the case the defender plays faster than the attacker, the attacker simply looses the delay.
+In the case the attacker plays faster, each time the defender plays in an interval, he will gain the extra time of the delay.
+The delay is therefore always detrimental to the benefit of the attacker.
+The formula for the second case is a slight overestimation of the defender's benefit in case the defender plays faster than the attacker period extende with the delay.
+When the defender plays in an interval, he will gain extra time in the next interval, since the attacker has to regain control.
+However, not the entire delay will be extra time, as the defender may move in again sooner.
+Futher research needs to be performed to calculate the exact overestimation in this case.\\
+Also the impact of the delay on nash equilibria and the determination of optimal defender and attacher strategies needs further investigation.
 

### 2015-05-17 IEEE Paper/paper/paper.tex
Formule !!

@@
 \usepackage[cmex10]{amsmath}
-\usepackage[dutch]{babel}
 
 \usepackage{float}
+%maakt caption van figuren groter
 \usepackage{caption}
-\usepackage[caption = false,subrefformat=parens,labelformat=parens]{subfig}
 \usepackage{amsmath}
@@
 \usepackage[titlenumbered,algoruled, linesnumbered]{algorithm2e}
-\usepackage{algpseudocode}
 \usepackage{booktabs}
@@
 
-A particular kind of frequently occurring threaths are Advanced Persistent Threats (APT).
+A particular kind of frequently occurring threats are Advanced Persistent Threats (APT).
 An APT is a targeted cyber attack that targets organisations in a stealthy way and that can stay undetected for a long period.
@@
 Against this sort of attacker, the absolute level of your security is what's important.
-It doesn't matter how secure you are compared to your peers; all that matters is whether you're secure enough to keep him out}'' - Bruce Schneier: APT is a Useful Buzzword [].\\
+It doesn't matter how secure you are compared to your peers; all that matters is whether you're secure enough to keep him out}'' - Bruce Schneier \cite{APTBruce}.\\
 
@@
 In the next section we first present the original FlipIt game.
-Then section [] presents the FlipIt game with virus propagation.
-Section [] presents some related work.
-Section [] concludes the paper and presents avenues for further research.
+Then section \ref{ch:FlipItGame} presents the FlipIt game with virus propagation.
+Section \ref{ch:extendedWork} presents some related work.
+Section \ref{ch:conclusion} concludes the paper and presents avenues for further research.
 
@@
 $C_{D}(t)$ is 1 if the game is under control by the defender and 0 if the game is under control by the attacker.
-Reversely, $C_{A}(t)$ will be 1 if the game is under control by the attacker and 0 if under controle by the defender.
+Reversely, $C_{A}(t)$ will be 1 if the game is under control by the attacker and 0 if under control by the defender.
 So, $C_{A}(t)= 1 - C_{D}(t)$.
@@
   \hline Non-adaptive (NA) & Renewal \\
-  &  ~~~~ Periodic \\
-  &  ~~~~ Exponential \\
+  & - Periodic \\
+  & ~~~ - Exponential \\
   & General non-adaptive \\
@@
 \begin{itemize}
-\item periodic games dominate the other renewal strategies, meaning that it is always advantagous to play periodically against an opponent with a renewal strategy;
-\item periodic games are disadvantagous against players following a Last Move adaptive strategy;
+\item periodic games dominate the other renewal strategies, meaning that it is always advantageous to play periodically against an opponent with a renewal strategy;
+\item periodic games are disadvantageous against players following a Last Move adaptive strategy;
 \item if the defender plays with a periodic rate that is fast enough he'll force the attacker to drop out;
@@
 
-Similarly as in [], we split the formalization in two cases.
+Similarly as in \cite{FlipIt}, we split the formalization in two cases.
 The first case is where the defender plays at least as fast as the attacker, the second case is where the attacker plays at least as fast as the defender.
@@
 
-Correspondeingly, the benefit for the defender can be expressed as:
+Correspondingly, the benefit for the defender can be expressed as:
 \begin{equation}\label{first}
@@
 Indeed, the attacker has to play soon enough to gain control, meaning that the attacker has to play during the period of $\delta_{D}-d$ during the interval of the defender.
-The probability that the attacker plays soon enough is $\dfrac{\delta_{D}-d}{\delta_{D}}$ and this will give the attacker an avarage gain of $\dfrac{\delta_{D}-d}{2}$.
+The probability that the attacker plays soon enough is $\dfrac{\delta_{D}-d}{\delta_{D}}$ and this will give the attacker an average gain of $\dfrac{\delta_{D}-d}{2}$.
 If the attacker moves after the period of $\delta_{D}-d$, the gain of the attacker will be zero.
@@
 
-To derive the benefit, the cost of moving is substracted from the average gain. 
+To derive the benefit, the cost of moving is subtracted from the average gain. 
 \begin{equation}\label{first}
@@
 \begin{equation}\label{first}
-\beta_{A}(\alpha_{D},\alpha_{A}) = \dfrac { \delta_{D}} {2 \cdot \delta_{A}} - k_{A} \alpha_{A} - (\dfrac{d}{\delta_{A}} - \dfrac{d^{2}}{2 \cdot \delta_{A} \delta_{D}})
+\beta_{A}(\alpha_{D},\alpha_{A}) = \dfrac { \delta_{D}} {2 \cdot \delta_{A}} - k_{A} \alpha_{A} - ( \dfrac{d^{2}}{2 \cdot \delta_{A} \delta_{D}} - \dfrac{d}{\delta_{A}} )
 \end{equation}
@@
 So, every time the defender plays, he will get an average gain of $\dfrac{\delta_{A}}{2}$ in the interval where he plays and in the next interval will always receive a extra gain of $d$, yielding a total average gain per interval of
-$\dfrac{(d+\dfrac{\delta_{A}2})}{\delta_{A}}$
+$\dfrac{(d+\dfrac{\delta_{A}}{2})}{\delta_{A}}$
 
@@
 If not, this would mean that the defender can move during the delay in the interval following the interval where the defender already moved.
-This would mean that there can be an overlap between the average gain of $\dfrac{\delta_{a}}{2}$ and the delay.
+This would mean that there can be an overlap between the average gain of $\dfrac{\delta_{A}}{2}$ and the delay.
 The above benefit formula would then include to much gain for the defender: the potential overlap during the delay would be counted twice.
 \\
-To obtain the formula in case of a too long delay, we therefore need to substract this overlapping gain from the above formula.  The overlapping gain is equal to 
+
 ~~ \\
-\subsubsection*{\textbf{Case b:} $d + \delta_{A} > \delta_{D}$}
+\subsubsection*{\textbf{Case b:} $d + \delta_{A} \geq \delta_{D}$}
 ~~~\\
+
+To obtain the formula in case of a too long delay, we therefore need to subtract this overlapping gain from the above formula. 
+Since $\delta_{D} \geq \delta_{A}$, if the defender enters the interval immediately after the attacker has played, then the defender cannot have played in the previous interval.
+In that case, there is no overlap.
+So the problem of the overlap only appears if the defenders enters late enough and thus only the last part of the delay is subject to overlap.
+The larger the difference between the interval of the defender and the attacker, the smaller the risk of overlap.
+Concretely, only the last part of lengt $d - (\delta_{D} - \delta_{A})$ is subject to overlap.
+Hence, the probability of overlap is $\dfrac{ d - (\delta_{D} - \delta_{A})}{\delta_{D}}$ and the gain will be half of this interval:  $\dfrac{ d - (\delta_{D} - \delta_{A})}{2}$.
+The gain rate to be subtracted is therefore:\\
+
+\begin{equation}\label{first}
+\dfrac{1} {\delta_{A}} \cdot \dfrac{d - (\delta_{D} - \delta_{A})}{\delta_{D}} \cdot \dfrac{d - (\delta_{D} - \delta_{A})}{\delta_{D}}
+\end{equation}
+
+The total gain  rate of the defender is obtained by subtracting this term from the gain rate of case a:
+ \begin{equation}\label{first}
+\gamma_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta_{A}}{\delta_{D}} \cdot \dfrac{(d+\dfrac{\delta_{A}}{2})}{\delta_{A}} - \dfrac{(d - (\delta_{D} - \delta_{A}))^{2}}{2 \delta_{D} \delta_{A}}
+\end{equation}
+\begin{equation}\label{first}
+\gamma_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - \dfrac{(d - (\delta_{D} - \delta_{A}))^{2}}{2 \delta_{D} \delta_{A}}
+\end{equation}\\
+This yields in the following benefit formula:
+\begin{equation}\label{first}
+\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - k_{D} \alpha_{D} - \dfrac{(d - (\delta_{D} - \delta_{A}))^{2}}{2 \delta_{D} \delta_{A}}
+\end{equation}\\
+
+The benefit for the attacker will be as follows:
+\begin{equation}\label{first}
+\beta_{A}(\alpha_{D},\alpha_{A}) = 1 -\dfrac{\delta_{A}}{2\delta_{D}} - \dfrac{d}{\delta_{D}} - k_{A} \alpha_{A} + \dfrac{(d - (\delta_{D} - \delta_{A}))^{2}}{2 \delta_{D} \delta_{A}}
+\end{equation}\\
 
@@
 Both the targeted and non-targeted attacks don't succeed immediately.
-[verschil uitleggen] For the targeted attack the time till it succeeds is given by an exponential distributed random variable with a known rate.
+For the targeted attack the time till it succeeds is given by an exponential distributed random variable with a known rate.
 The non-targeted attacks are modelled as a single attacker and the time till it succeeds is given by a Poisson process.
 The conclusion of this paper is that the optimal strategy for the defender is moving periodically.
+The difference with this paper is that the delay in this paper is dependent on the number of nodes that have to be flipped in a network.
+This cannot be modelled with the framework in the Laszka et al.
+paper because the delay is chosen as an exponential distributed random variable.
+Another difference is that in the case of the Laszka et al.
+paper, the moves of the defender are considered not stealthy and so the attacker knows when the defender plays.
 \\
@@
 Reitter et al.
-\cite{reitter2013risk} observed continuous games, 20-seconds FlipIt game.
+\cite{reitter2013risk} extended the work of A.
+Nochenson and Grossklags to include various visual presentation modalities for the available feedback during the investigation.
 
-\section{Conclusions and Furhter Research}
+\section{Conclusions and Further Research}
 \label{ch:conclusion}
@@
 The delay is therefore always detrimental to the benefit of the attacker.
-The formula for the second case is a slight overestimation of the defender's benefit in case the defender plays faster than the attacker period extende with the delay.
+The formula for the second case is a slight overestimation of the defender's benefit in case the defender plays faster than the attacker period extended with the delay.
 When the defender plays in an interval, he will gain extra time in the next interval, since the attacker has to regain control.
 However, not the entire delay will be extra time, as the defender may move in again sooner.
-Futher research needs to be performed to calculate the exact overestimation in this case.\\
-Also the impact of the delay on nash equilibria and the determination of optimal defender and attacher strategies needs further investigation.
+Further research needs to be performed to calculate the exact overestimation in this case.\\
+Also the impact of the delay on Nash equilibria and the determination of optimal defender and attacher strategies needs further investigation.
 

### 2015-05-17 IEEE Paper/paper/paper.tex
Conclusie aangepast

@@
 In the case the defender plays faster than the attacker, the attacker simply looses the delay.
-In the case the attacker plays faster, each time the defender plays in an interval, he will gain the extra time of the delay.
+In the case the attacker plays faster, each time the defender plays in an interval, he will gain extra time of the delay.
 The delay is therefore always detrimental to the benefit of the attacker.
-The formula for the second case is a slight overestimation of the defender's benefit in case the defender plays faster than the attacker period extended with the delay.
-When the defender plays in an interval, he will gain extra time in the next interval, since the attacker has to regain control.
-However, not the entire delay will be extra time, as the defender may move in again sooner.
-Further research needs to be performed to calculate the exact overestimation in this case.\\
-Also the impact of the delay on Nash equilibria and the determination of optimal defender and attacher strategies needs further investigation.
+This demonstrates that the FlipIt game can be adapted to a game with virus propagation.
+Further research needs to be performed to calculate the impact of the delay on Nash equilibria and the determination of optimal defender and attacker strategies.
 

### 2015-05-18 IEEE Paper/paper/paper.tex
IEEE artikel finaal

@@
 Many businesses store confidential information on clients, which can be lost and possible be abused by competitors through data leakage.
-Also, disruption caused by DOSS attacks, may restuls in businesses failing to meet their service-level agreements.
+Also, disruption caused by DOSS attacks, may results in businesses failing to meet their service-level agreements.
 Ultimately, system and network security helps protecting a business's reputation, which is one of its most important assets.
@@
 In the next section we first present the original FlipIt game.
-Then section \ref{ch:FlipItGame} presents the FlipIt game with virus propagation.
+Then section \ref{ch:flipitvirus} presents the FlipIt game with virus propagation.
 Section \ref{ch:extendedWork} presents some related work.
@@
 \section{FlipIt with virus propagation}
+\label{ch:flipitvirus}
 A FlipIt game consists of a single resource. To represent the security problem, the game now defines its single resource as a computer network with multiple
@@
 After dropping a virus on the first node, it takes a while for the virus to infect the entire network.
-However, since the original FlipIt game works with a single resource that is always flipped entirely, the assumption is made that the attacker is considered to gain immediate full control over the resource when the network has been infected, even it is only one node that has been infected.\\
+Since the original FlipIt game works with a single resource that is always flipped entirely, the assumption is made that the attacker is considered to gain immediate full control over the resource when the network has been infected, even it is only one node that has been infected.\\
 
@@
 So, the assumption that the attacker has full control over the resource as soon as a node has been infected, is not realistic.
-The attacker has only control of the network once all or a sufficient large number of nodes are infected.
+The attacker has only control of the network once all or a sufficient number of nodes are infected.
 The time that it takes for the virus to infect every node (or a sufficient number of nodes) will be
@@
 first infected node to the farthest node.
-Rather than denoting the time needed for infecting \textit{all} the nodes, the variable $d$ can also be used to denote the time needed to infect \textit{a sufficient large number} of nodes.
+Rather than denoting the time needed for infecting \textit{all} the nodes, the variable $d$ can also be used to denote the time needed to infect \textit{a sufficient number} of nodes.
 
-Assume that an attacker attacks at time \textit{t}, he doesn't get immediate control over the resource, but he only gains control at time \textit{t + d}, with $d$ denoting the time needed to infect a sufficiently large number (or all) nodes.
+Assume that an attacker attacks at time \textit{t}, he doesn't get immediate control over the resource, but he only gains control at time \textit{t + d}, with $d$ denoting the time needed to infect a sufficiently number (or all) nodes.
 If the defender flips the network before the period $d$ has elapsed (so, somewhere between $t$ and $t + d$), then the attacker will never gain full control over the resource.
@@
 This corresponds to a periodic defender strategy.
-A periodic attacker strategy is assumed as well, as this also corresponds to a common real life strategy.
+A periodic attacker strategy is assumed as well, to be able to compare the results with the periodic strategy of the FlipIt game in \cite{FlipIt}.
+%, as this also corresponds to a common real life strategy.
 %nog eventueel verder te motiveren
@@
 
-However, because of the delay required for virus propagation, the maximal time of control is reduced to $\delta_{D}-d$.
+\begin{figure}[hbtp]
+\caption{The first FlipIt game is one without virus propagation. The second one is with virus propagation and \textit{d} = 1. The delay is denoted with an arrow.}
+\centering
+\includegraphics[scale=0.4]{../../doc/template/Images/FLipItCase1.pdf}
+\label{fig:delaycase1}
+\end{figure}
+
+However, because of the delay required for virus propagation, the maximal time of control is reduced to $\delta_{D}-d$ , see figure \ref{fig:delaycase1}.
 There is a probability of \textit{r} that the attacker will move in the interval of the defender.
@@
 The larger the difference between the interval of the defender and the attacker, the smaller the risk of overlap.
-Concretely, only the last part of lengt $d - (\delta_{D} - \delta_{A})$ is subject to overlap.
+Concretely, only the last part of length $d - (\delta_{D} - \delta_{A})$ is subject to overlap.
 Hence, the probability of overlap is $\dfrac{ d - (\delta_{D} - \delta_{A})}{\delta_{D}}$ and the gain will be half of this interval:  $\dfrac{ d - (\delta_{D} - \delta_{A})}{2}$.

### 2015-05-19 doc/template/chap-10.tex
alles opslagen

@@
 % -- Situation -- %
+
 In this era where digitalization becomes prominent in every aspect of our lives, where technology is growing fast and where businesses are always under attack, security becomes an issue of increasing complexity.
@@
 It is the same as leaving the door of your house wide open for everyone to come in.
-Why is it so important to keep a system secure?
 \\
-
-Businesses can have confidential information on clients.
+%Businesses can have confidential information on clients.
 Through data leakage, confidential information can be lost and possibly used by the competitor.
@@
 Ultimately, system and network security helps protecting a business's reputation, which is one of its most important assets.
+
+Why is it so important to keep a system secure?
+Many businesses store confidential information on clients, which can be lost and possible be abused by competitors through data leakage.
+Also, disruption caused by DOSS attacks, may results in businesses failing to meet their service-level agreements.
+Ultimately, system and network security helps protecting a business's reputation, which is one of its most important assets.
+\\
 %\todo{waarom?
 :  lekken van informatie: je hebt info over klanten en moet hun privacy beschermen, je eigen data is geld waard voor concurrenten, DOSS attacks: je will je service-level agreements kunnen nakomen, ..}
- A hacker will be a person that seeks exploits or weaknesses in a system or network in order to gain access.
+% A hacker will be a person that seeks exploits or weaknesses in a system or network in order to gain access.
 Many of those attacks have a different cause.
@@
 
-An APT is a targeted cyber attack that will keep trying to break into the targeted system.
-It will try to stay unnoticed for as long as possible.
-Bruce Schneier describes an APT as something different and stronger than a conventional hacker: ''\textit{A conventional hacker or criminal isn't interested in any particular target.
+A particular kind of frequently occurring threats are Advanced Persistent Threats (APT).
+An APT is a targeted cyber attack that targets organisations in a stealthy way and that can stay undetected for a long period.
+This makes it so hard to protect a network or a system against an APT.
+Bruce Schneier describes an APT as something different and stronger than a conventional threat: ''\textit{A conventional hacker or criminal isn't interested in any particular target.
 He wants a thousand credit card numbers for fraud, or to break into an account and turn it into a zombie, or whatever.
@@
 Against this sort of attacker, the absolute level of your security is what's important.
-It doesn't matter how secure you are compared to your peers; all that matters is whether you're secure enough to keep him out}'' - Bruce Schneier: APT is a Useful Buzzword.
-
--- verder uitwerken -- \\
-\textit{Stealthy key characteristic of APT. \\
-Voorbeeld APT Equastion group. \\
-Site van Kaspersky lab\\}
-\todo{APT uitleggen ? en dat het moeilijk is om een systeem te beschermen tegen APT ==> overgang naar de volgende paragraaf maken.}
+It doesn't matter how secure you are compared to your peers; all that matters is whether you're secure enough to keep him out}'' - Bruce Schneier \cite{APTBruce}.\\
 
 \subsubsection{Complication}
-% -- Complication -- %
-Since it is so difficult to protect a system or a network against APT, researcher have been looking for effective ways to predict in advance which defence strategy might be the better one. 
-Game theory is gaining more and more interest as an effective technique to  model and study Cyber Security.
+Since it is so difficult to protect a system or a network against APT's, researchers have been looking for effective ways to predict in advance which defence strategy might be the better one. 
+Game theory is gaining increasing interest as an effective technique to model and study Cyber Security.
 Game theory analyses the security problem as a game where the players are an attacker and a defender of a system, and where both players have to make decisions.
@@
 They study the specific scenario where a system or network is repeatedly taken over completely by an attacker and this attack is not immediately detected by the defender of the system or network.
-In game theory, such a game is known as ''FlipIt''.
+In game theory, such a game is known as ''FlipIt'' \cite{FlipIt}.
 This is a two players game where the attacker and the defender are competing to get control over a shared resource.
@@
 \subsubsection{Research questions}
+
+%This paper proposes the adaptation of the FlipIt formulas as presented in \cite{FlipIt} such as to take the delay for virus propagation into account.
+In the next section we first present the original FlipIt game.
+Then section \ref{ch:flipitvirus} presents the FlipIt game with virus propagation.
+Section \ref{ch:extendedWork} presents some related work.
+Section \ref{ch:conclusion} concludes the paper and presents avenues for further research.
+
 The game theoretical approach of the FlipIt does not take such delay into account. 
@@
 \begin{itemize}
-\item Is it possible to incorporate the notion of delay in the game-theoretical analysis of the Flip-It game ?
 \item Is it possible to incorporate the notion of delay in the game-theoretical analysis of the Flip-It game ?
@@
 \end{itemize}
-%-        How can we calcuate the expected duration for a node’s infection/the entire network infection ?
-%
-%-        Can we calculate this node per node ?
 
-% -- Anwer -- %
 \subsubsection{Contributions}
@@
 The organisation of this paper is the following.
-In chapter 2 [] \todo{ref}  a brief introduction to Gametheory is introduced to get familiar with the game theoretic concepts that will be further used in the paper.
+In chapter \ref{Chapter1:Intro.Game.Theory}  a brief introduction to Gametheory is introduced to get familiar with the game theoretic concepts that will be further used in the paper.
 In the same chapter the FlipIt framework is summarized and the most important conclusions together with the the related work done on FlipIt and the difference with this paper.
-In chapter 3  [] \todo{ref} , we first introduce the adaptations made on FlipIt to model a FlipIt game with virus propagation.
+In chapter \ref{chapter2:FlipIt with virus propagation} , we first introduce the adaptations made on FlipIt to model a FlipIt game with virus propagation.
 After that formulas are derived to model a FlipIt game with a virus propagation for a specific case where players play a periodic strategy with a random phase.
 This chapter ends with simulations where conclusions can be derived.
-Next in Chapter 4 [] \todo{ref}, we given an overview of the various ways in which a virus can propagate.
+Next in Chapter \ref{chapter4: Virus propagation} we given an overview of the various ways in which a virus can propagate.
 We present a method to calculate the speed of the propagation of a virus in a network and how this network can be established to reduce the spreading of a virus.
-Finally, in Chapter 5 [] \todo{ref}, we discuss the main results and complications and provide directions for further research.
+Finally, in Chapter \ref{chapter5:conclusion} , we discuss the main results and complications and provide directions for further research.
 

### 2015-05-19 doc/template/chap-5.tex
alles opslagen

@@
 \chapter{Intoduction to GameTheory}
-\label{cha:1}
+\label{Chapter1:Intro.Game.Theory}
 %\documentclass[10pt]{article}
@@
 To get an understanding of the most important concepts of game theory, a short introduction based on the work of
-\cite{leyton2008essentials} and \cite{Coursera} is given in section \ref{Cha:1:Intro.Game.Theory}  . For a more detailed and full introduction to game theory, the reader is referred to 
-\cite{leyton2008essentials}.  In section \ref{ch:FlipItGame} an overview of the FlipIt game is given with the definitions and concepts that will be used throughout the paper. 
-The last section \ref{ch:extendedWork} will cover the extensions and additions already made on FlipIt.
+\cite{leyton2008essentials} and \cite{Coursera} is given in section \ref{Cha1:briefintro}  . For a more detailed and full introduction to game theory, the reader is referred to 
+\cite{leyton2008essentials}.  In section \ref{cha1:FlipItGame} an overview of the FlipIt game is given with the definitions and concepts that will be used throughout the paper. 
+The last section \ref{ch1:extendedWork} will cover the extensions and additions already made on FlipIt.
 
@@
 \section{A brief introduction in Game Theory}
-\label{Cha:1:Intro.Game.Theory}
+\label{Cha1:briefintro}
 
@@
 It is a mathematical way of modelling the interactions between two or more agents where the outcomes depend on what everybody does and how it should be structured to lead to good outcomes.
-For this reason it can be very useful in economics and also in other branches as politics, biology, computer science, philosophy and a variety of other disciplines.
+It has therefore important applications in many area's such as economics, politics, biology, computer science, philosophy and a variety of other disciplines.
 \\
@@
 These preferences are mapped to natural numbers and are called the utility function.
-The numbers are interpreted as a mathematical measure to tell you how much an agent likes or dislikes the states of the world.
+The numbers are interpreted as a mathematical measure that tells how much an agent likes or dislikes the states of the world.
 \\
-It also explains the impact of uncertainty.
-When an agent is uncertain about a distribution of outcomes, his utility will describe the expected value of the utility function with respect to the probability of the distribution of the outcomes.
-For example: with 0.7 probability it will be 7 degrees outside and 0.3 probability it will be 10 degrees.
-The agent can have a different opinion about that distribution versus another distribution.
-(\todo{uitleggen aan de hand van een voorbeeld}).\\
-\todo{players rationeel en max outcomes}
 %Cooperative and non cooperative games
 In a Decision Game Theoretic Approach an agent will try to act in such a way to maximise his expected or average utility function.
-It becomes more complicated when two or more agents want to maximise their utility and whose actions can affect each other utilities.
+It becomes more complicated when two or more agents want to maximise their utility and when actions of the agents can affect each other's utilities.
 This kind of games are referred to as non-cooperative game theory, where the basic modelling unit is the group of agents.
@@
 A Nash Equilibrium is a subset of outcomes that can be interesting to analyse a game.
-For a Nash Equilibrium each player has a consist list of actions and each player's action maximizes his or her payoff given the actions of the other players.
+For a Nash Equilibrium each player has a consist list of actions and each player's action maximizes his or her pay-off given the actions of the other players.
 Nobody has the incentive to change his or her action if an equilibrium profile is played.
@@
 
-Formal defenition of a Nash Equilibrium:
+Formal definition of a Nash Equilibrium:
 A strategy profile $s = (s1, . . . , sn)$ is a Nash equilibrium
@@
 his strategy if he knew what strategies the other agents were following.
-Wecan divide Nash equilibria into two categories, strict and weak, depending on whether
+We can divide Nash equilibria into two categories, strict and weak, depending on whether
 or not every agent's strategy constitutes a unique best response to the other agents' strategies.''
@@
 \section{The FlipIt game}
-\label{ch:FlipItGame}
-In this section, we introduce the ''stealthy takeover '' game \flip{FlipIt} \cite{FlipIt}.
-\flip{FlipIt} is a game introduced by van Dijk et al.
-First we explain the framework of FlipIt and introduce the most important formulas that will be used throughout the paper.
-To understand how to model a FlipIt game with virus propagation it is important to get familiar with the concepts of the normal FlipIt game and it's notations.
+\label{cha1:FlipItGame}
+FlipIt is a game introduced by van Dijk et al.
+To understand how to model a FlipIt game with virus propagation it is important to get familiar with the concepts of the normal FlipIt game and its notations.
+Therefore, we first explain the framework of FlipIt and introduce the most important formulas that will be used throughout the paper.
 \\
 
-\flip{FlipIt} is a two-players game with a shared (single) resource that the players want to control as long as possible.
+FlipIt is a two-players game with a shared single resource that the players want to control as long as possible.
 The shared resource can be a password, a network or a secret key depending on the setting being modelled.
-In the rest of the paper we name the two players the attacker, denoted by the subscript \textit{A} and the defender, denoted by subscript \textit{D}.
+In the remainder of the paper we name the two players the attacker, denoted by the subscript \textit{A} and the defender, denoted by subscript \textit{D}.
 
-The game begins at $t=0$ and continuous indefinitely ($t \rightarrow \infty $).
-The time in the game can be viewed as being continuous, but a discrete time can also be viewed.
-To get control over the resource, the players \textit{i} can flip the resource at any given time.
+The game begins at $t=0$ and continues indefinitely ($t \rightarrow \infty $).
+The time in the game is assumed as being continuous, but a discrete time could also be considered.
+To get control over the resource, the players $i$, with $i \in \{A,D\}$, can flip the resource at any given time.
 A flip will be regarded as a move from a player \textit{i}.
@@
 Both players will try to minimize their cost.
-By adding a cost, it will prevent players to move to frequently.
+Adding a cost will prevent players to move too frequently.
 \\
 
-The unique feature of \flip{FlipIt} is that every move will happen in a stealthy way, meaning that the player has no clue that (his adversary) the other player has flipped the resource.
-For instance, the defender will not find out if the resource has already been compromised by the attacker, but he can only potentially know it after he flips the resource himself.
-The goal of the player is to maximize the time that he or she has control over the resource while minimizing total cost of the moves.
+The unique feature of FlipIt is that every move will happen in a stealthy way, meaning that the player has no clue that the other player (his adversary) has flipped the resource.
+For instance, the defender will not find out if the resource has already been compromised by the attacker until he flips the resource himself.
+The goal of the player is to maximize the time that he or she has control over the resource while minimizing the total cost of the moves.
 A move can also result in a "wasted move", called a flop.
-It may happen that the resource was already under control by the defender.
-If the defender moves when he or she has already control over the resource, he or she would have wasted a move since it does not result in a change of ownership and a cost is involved.
+It may happen that the resource was already under control by the player.
+If the player moves when he or she has already control over the resource, he or she would have wasted a move since it does not result in a change of ownership, so the cost is wasted.
 \\
@@
 \centering
-\includegraphics[scale=0.8]{Images/DefFlipit}
-\caption{A representation of a FlipIt game where both players are playing periodically and discrete.
+\includegraphics[scale=0.5]{../../doc/template/Images/DefFlipit}
+\caption{A representation of a FlipIt game where both players are playing periodically and at discrete time intervals.
 Every move or flip is indicated by a blue or orange circle.
-The attacker is the orange colour and plays with a period of $\delta_{A}=4$.
-The defender is the blue colour and plays with a period of $\delta_{D}=3$.The blue and orange rectangles represent the amount of time one of the players is in control of the resource.}
+The attacker is represented in orange and plays with a period of $\delta_{A}=4$.
+The defender is represented in blue and plays with a period of $\delta_{D}=3$.
+The blue and orange rectangles represent the amount of time the respective player is in control of the resource.}
 \label{fig:FLipItDefault}
 \end{figure}
-\todo{verwijzen naar de figuur \ref{fig:FLipItDefault}}
 
-The state of the resource is denoted as a time independent variable $C=C_{i}(t)$. 
-$C_{D}(t)$ is either 1 if the game is under control by the defender and 0 if the game is under control by the attacker. For $C^{A}(t)$ it is visa versa, $C^{A}(t)= 1 - C^{D}(t)$.\\ \\
-The game starts with defender being in control of the game, $C_{D}(0)= 1$.
+The state of the resource is denoted as a time-dependent variable $C=C_{i}(t)$. 
+$C_{D}(t)$ is 1 if the game is under control by the defender and 0 if the game is under control by the attacker.
+Reversely, $C_{A}(t)$ will be 1 if the game is under control by the attacker and 0 if under control by the defender.
+So, $C_{A}(t)= 1 - C_{D}(t)$.
+The game starts with the defender being in control: $C_{D}(0)= 1$. \\
 
-The players receive a benefit equal to the time of units that they were in possession of the resource minus the cost of making their moves. The cost of a player \textit{i} is denoted by $k_{i}$. 
+The players receive a benefit equal to the time units they were in possession of the resource minus the cost of making their moves. The cost of a player \textit{i} is denoted by $k_{i}$. 
 The total gain of player \textit{i} is equal to the total amount of time that a player \textit{i} has owned the resource from the beginning of the game up to time \textit{t}.
@@
 
-If there is no need for feedback for both of the players, we say that we have a non-adaptive strategy.
-Because the player does not receive any feedback during the game it will play in the same manner against every opponent.
-They are not dependent on the opponents movements.
-This means that they can already generate the time sequence for all the moves in advance.
-But they can depend on some randomness because the non-adaptive strategies can be randomised.
-In this paper we will focus in the beginning on the non-adaptive strategies.
-Reasons behind this is that a player (defender or attacker) rarely knows what the strategies are of his opponent.
-An interesting subclass of the non-adaptive strategies is one where the time intervals between two consecutive moves are generated by a renewal process.
-Example of such a renewal strategy is the periodic strategy where the time between two consecutive moves of the players are a fixed interval.
+If there is no feedback for neither of the players, we have a non-adaptive strategy.
+Because a player does not receive any feedback during the game he will play in the same manner against every opponent.
+The strategy is called non-adaptive because the playing strategy is not dependent on the opponents movements.
+An interesting subclass of the non-adaptive strategies is the one where the time intervals between two consecutive moves are generated by a renewal process.
+An example of such renewal strategy is the periodic strategy where the time between two consecutive moves of the players are a fixed interval.
 An exponential strategy is a renewal strategy in which the interval between two consecutive moves is exponentially distributed.
 \\
-
--- Ik kan dit ook pas uitleggen als ik mijn FlipIt spel met virus propagatie uit de doeken doe -- \\
-We elaborate more about the periodic strategy because we will use this one to model our first extension. For more details about the other strategies of FlipIt we refer the reader to \citep{FlipIt}.
-
-\textit{periodic strategy}: A Periodic strategy is a non-adaptive renewal strategy where the time intervals between consecutive moves are a fixed period, denoted by $\delta$.
-Moreover it has a random phase, that is chosen uniformly and random in the interval $[0,\delta]$ for the first move.
-The average rate of play is denoted by $\alpha = \dfrac{1}{\delta}$.
+In case there is feedback, a player can adapt his strategy to the information received about the opponent's moves.
+Depending on the amount of information received, two subclasses of adaptive strategies can be identified.
+The Last Move (LM) strategies represent the class where whenever a player flips he will find out the exact time that the opponent played the last time.
+In the second class, called Full History (FH), whenever a player flips he will find out the whole history of the opponent's move.
 \\
-
-In the category Adaptive strategy there are two sub classes of strategies.
-The first one is the Last move (LM).
-In this class whenever a player flips it will find out the exact time that the opponent played the last time.
-In the second class the Full History (FH), whenever a player flips it will find out the whole history of the opponents move.
-If the opponent player plays with a renewal strategy the sub classes FH and LM collapse.
+In this paper we will focus on the non-adaptive strategies.
+This choice is motivated by the fact that in a security game a player (defender or attacker) rarely has information about the moves (last move or full history) of his opponent.
 \\
- I%n this subsection we elaborate about the strategy of FlipIt . 
-%A renewal strategy is a non-adaptive strategy \\
-
-The game can be extended by the amount of information that a player receives. It can also be possible for a player to get information at the start of the game. Both interesting cases are:
-\begin{itemize}
-\item Rate-of-play (RP): The player finds out the exact rate of play of the opponent.
-\item Knowledge-of-strategy (KS): The player finds out the complete information of the strategy that the opponent is playing.
-\end{itemize}
 
@@
   \textbf{Categories} & \textbf{Classes of Strategies} \\
-  \hline Non-adaptive (NA) & Exponential \\
-  & Periodic \\
-  & Renewal \\
+  \hline Non-adaptive (NA) & Renewal \\
+  & - Periodic \\
+  & ~~~ - Exponential \\
   & General non-adaptive \\
@@
 
-\subsubsection{Imporant results}
-
+\subsubsection{Results of the FlipIt game}
+The study of the different strategies by means of FlipIt framework allows to derive a number of interesting results:  
 \begin{itemize}
-\item periodische spellen domineren de andere renewal strategies
-\item periodische strategien wel slecht tegen attacker die laatste move weet
-\item als de defender snel speelt dan forceert die de attacker om niet meer mee te spelen
-\item iedereen die feedback krijgt heeft meer benefit dan zonder die feedback
+\item periodic games dominate the other renewal strategies, meaning that it is always advantageous to play periodically against an opponent with a renewal strategy;
+\item periodic games are disadvantageous against players following a Last Move adaptive strategy;
+\item if the defender plays with a periodic rate that is fast enough he'll force the attacker to drop out;
+\item any amount of feedback about the opponent received during the game, benefits to a player.
 \end{itemize}
@@
 \section{Extensions on FlipIt}
-\label{ch:extendedWork}
+\label{ch1:extendedWork}
 
-In this section we discuss the extensions that can be made on the FlipIt game. \\
-
-There a various possible ways to extend \flip{FlipIt}. 
+Various possible ways to extend FlipIt have already been proposed. 
 Laszka et al.
-made a lot of additions and extensions on the original game of FlipIt.
+made a lot of additions and extensions to the original game of FlipIt.
 For instance Laszka et al.
-extended the basic \flip{FlipIt} game to multiple resources.
-The incentive is that for compromising a system in a real case it needs more than just taking over just one resource.
-An example is that one resource can be gaining access to a system and breaking the password of the system is another resource.
+extended the basic FlipIt game to multiple resources.
+The rationale is that for compromising a system in real life, more than just one resource needs to be taken over.
+An example is that gaining access to deeper layers of a system may require breaking several passwords.
 The model is called FlipThem \cite{FlipThem}.
-They use two ways to flip the multiple resources: the AND and the OR control model.
+Laszka et al.
+also use two ways to flip the multiple resources: the AND and the OR control model.
 In the AND model the attacker only controls the system if he controls all the resources of the system, whereas in the OR model the attacker only needs to compromise one resource to be in control of the entire system.
 \\
-%The difference with FlipThem and this paper is that we introduce a Graph Model in the beginning.\\
-Another addition of Laszka et al.
-to the game of FlipIt \cite{MitigationCovert} is extending the game to also consider non-targeted attacks by non-strategic players.
+
+Another addition of Laszka et al. to the game of FlipIt \cite{MitigationCovert} 
+is extending the game to also consider non-targeted attacks by non-strategic players.
 In this game the defender tries to maintain control over the resource that is subjected to both targeted and non-targeted attacks.
@@
 For the targeted attack the time till it succeeds is given by an exponential distributed random variable with a known rate.
-The non-targeted attacks are modelled as a single attacker and the time till it succeeds is given by a Poisson process.\todo{misschien meer uitleggen}.
+The non-targeted attacks are modelled as a single attacker and the time till it succeeds is given by a Poisson process.
 The conclusion of this paper is that the optimal strategy for the defender is moving periodically.
+The difference with this paper is that the delay in this paper is dependent on the number of nodes that have to be flipped in a network.
+This cannot be modelled with the framework in the Laszka et al.
+paper because the delay is chosen as an exponential distributed random variable.
+Another difference is that in the case of the Laszka et al.
+paper, the moves of the defender are considered not stealthy and so the attacker knows when the defender plays.
 \\
@@
 The basic idea is to test with an extra action if the resource has been compromised or not.
-This move involves also an extra cost.
-This model is useful if somebody wants to know for example if his or her password has been compromised.
-\\
+This move involves also an extra cost.\\
 
-Finally researchers also have investigated the behavioural of humans when playing FlipIt.
-A Nochenson and Grossklags \cite{nochenson2013behavioral}  investigate how people really act when given temporal decisions.
+Finally researchers also have investigated the behaviour of humans playing FlipIt.
+A.
+Nochenson and Grossklags \cite{nochenson2013behavioral}  investigate how people really act when given temporal decisions.
 Reitter et al.
-\cite{reitter2013risk} observed continuous games, 20-seconds FlipIt game..
+\cite{reitter2013risk} extended the work of A.
+Nochenson and Grossklags to include various visual presentation modalities for the available feedback during the investigation.
 

### 2015-05-19 doc/template/chap-9.tex
little change
paper nog meer aanpassen
alles opslagen

@@
 \chapter{FlipIt game with virus propagation}
-\label{cha:9}
+\label{chapter2:FlipIt with virus propagation}
 %\documentclass[10pt]{article}
@@
 \section{Introduction}
+\label{Ch2:Intro}
 
@@
 
-In chapter 2 the FlipIt game was explained.
-This chapter starts from the specific case of a non-adaptive continuous FlipIt game where both players play a periodic strategy with a random phase.
-This choice is motivated by the assumption that in the practical situation of most organisations, the defence strategy is to periodically defend the network.
-This corresponds to a periodic defender strategy.
-\todo{bedrijven hebben niet het geld en de mankracht om zich daar mee bezig te houden en dan is periodisch verdedigen het handigste, bedrijven willen alles zo efficient en met de minste kost doen} To simplify the analysis in a first time, a periodic attacker strategy is assumed as well.
-Further research can investigate the effect of relaxing this assumption.\\
-
 A FlipIt game consists of a single resource. To represent the security problem, the game now defines its single resource as a computer network with multiple
@@
 will do this by flipping the node in the graph that can infect all the nodes in the
-shortest time possible.
+shortest possible time.
 After dropping a virus on the first node, it takes a while for the virus to infect the entire network.
-However, since the original FlipIt game works with a single resource that is always flipped entirely, the assumption is made that the attacker is considered to have gained the control over the resource only when all the nodes of the network have been infected, i.e.
-the entire resource has been flipped.
+Since the original FlipIt game works with a single resource that is always flipped entirely, the assumption is made that the attacker is considered to gain immediate full control over the resource when the network has been infected, even it is only one node that has been infected.\\
 
-After dropping a virus on the first resource, it takes a while for the virus to infect
-the entire network. The time that it takes for the virus to infect every node will be
-denoted as parameter d. If we want to measure how long it takes for the virus to
+In reality however, after dropping a virus on the first node, it takes a while for the virus to infect
+the entire network.
+So, the assumption that the attacker has full control over the resource as soon as a node has been infected, is not realistic.
+The attacker has only control of the network once all or a sufficient number of nodes are infected.
+The time that it takes for the virus to infect every node (or a sufficient number of nodes) will be
+denoted as an infection-delay variable \textit{d} (called 'delay' for short in the remainder of this paper). If we want to measure how long it takes for the virus to
 infect all the nodes in the network, we have to calculate the shortest path from the
-first infected node to the farthest node. This can be measured by a method that we
-will explain in section [].
-Assume that an attacker attacks at time t, then only at time t + d he gains control over the entire network.
-If the defender flips the network before the period d has elapsed (so, somewhere between t and t+ d), then the attacker will never gain control over the entire network.
-Using this parameter d, a FlipIt game with virus propagation
-can be modelled.
+first infected node to the farthest node.
+Rather than denoting the time needed for infecting \textit{all} the nodes, the variable $d$ can also be used to denote the time needed to infect \textit{a sufficient number} of nodes.
+
+Assume that an attacker attacks at time \textit{t}, he doesn't get immediate control over the resource, but he only gains control at time \textit{t + d}, with $d$ denoting the time needed to infect a sufficiently number (or all) nodes.
+If the defender flips the network before the period $d$ has elapsed (so, somewhere between $t$ and $t + d$), then the attacker will never gain full control over the resource.
+This implies that the mathematical formulas for gain and benefit need to be adapted to the fact that the attacker loses part of its benefit because of this delay.
+In the remainder of this paper, we will adapt the formalization of the FlipIt game using the variable $d$.
+\\
 
@@
 
-\section{Benefit for a periodic FlipIt game with virus propagation}
-We start by formalizing the non-adaptive continuous basic FlipIt game where players use a periodic strategy with a random phase.
-Next we split the formalisation for the benefit formula for a FlipIt game with virus propagation in two cases.
-First the benefit formula is derived for the basis case and then it is calculated for the case with a virus propagation.
-One case is where the defender plays at least as fast as the attacker, the other case is where the attacker plays at least as fast as the defender.
-This subdivision is also used in the FlipIt paper.
+The formalization starts from the model of the non-adaptive continuous basic FlipIt game where players use a periodic strategy with a random phase.
+This choice is motivated by the assumption that in most organisations, the defence strategy is to periodically defend the network.
+This corresponds to a periodic defender strategy.
+A periodic attacker strategy is assumed as well, to be able to compare the results with the periodic strategy of the FlipIt game in \cite{FlipIt}.
+%, as this also corresponds to a common real life strategy.
+%nog eventueel verder te motiveren
+Further research can investigate the effect of relaxing this assumption. \\
+
+Similarly as in \cite{FlipIt}, we split the formalization in two cases.
+The first case is where the defender plays at least as fast as the attacker, the second case is where the attacker plays at least as fast as the defender.
+For each of these cases, first the benefit formula of the basic case without delay is presented, and then the delay is introduced.
 \\
 
+\subsection{Formalization the benefit formula including the infection-delay}
  A Periodic strategy is a non-adaptive renewal strategy where the time intervals between consecutive moves are a fixed period, denoted by $\delta$.
@@
 
-However, because of the delay caused by a virus, the maximal time of control is reduced to $\delta_{D}-d$.
-There is a probability of \textit{r} that the attacker will move in the interval of the defender.
-The attacker has to play soon enough to gain control, meaning that the attacker has to play during the period of $\delta_{D}-d$ during the interval of the defender.
-There is $\dfrac{\delta_{D}-d}{\delta_{D}}$ probability that the attacker will move soon enough which gives the attacker a gain of $\dfrac{\delta_{D}-d}{2}$.
-If the attacker moves after the period of $\delta_{D}-d$, the gain of the attacker will be zero.
-The average gain rate of the attacker can be expressed as follows if we look at one interval of the defender:
+Correspondingly, the benefit for the defender can be expressed as:
 \begin{equation}\label{first}
-\beta_{A}(\alpha_{D},\alpha_{A}) = \dfrac {1}{\delta_{D}} [ \dfrac{\delta_{D}}{\delta_{A}} \cdot \dfrac{\delta_{D}-d}{\delta_{D}} \cdot \dfrac{\delta_{D}-d}{2} + \dfrac{\delta_{D}}{\delta_{A}} \cdot \dfrac{d}{\delta_{D}} \cdot 0 ]
+\beta_{D}(\alpha_{D},\alpha_{A}) =1 -  \dfrac {r} {2} - k_{D} \alpha_{D} = 1 - \dfrac {\delta_{D}} {2\delta_{A}} - k_{D} \alpha_{D} 
 \end{equation}
 
-To complete the formula to derive the benefit function, the cost of moving is added. In the second formula we can see the formula of the original FlipIt game.
+\begin{figure}[hbtp]
+\caption{The first FlipIt game is one without virus propagation. The second one is with virus propagation and \textit{d} = 1. The delay is denoted with an arrow.}
+\centering
+\includegraphics[scale=0.4]{../../doc/template/Images/FLipItCase1.pdf}
+\label{fig:delaycase1}
+\end{figure}
+
+However, because of the delay required for virus propagation, the maximal time of control is reduced to $\delta_{D}-d$ , see figure \ref{fig:delaycase1}.
+There is a probability of \textit{r} that the attacker will move in the interval of the defender.
+However, the gain will not be half of the interval.
+Indeed, the attacker has to play soon enough to gain control, meaning that the attacker has to play during the period of $\delta_{D}-d$ during the interval of the defender.
+The probability that the attacker plays soon enough is $\dfrac{\delta_{D}-d}{\delta_{D}}$ and this will give the attacker an average gain of $\dfrac{\delta_{D}-d}{2}$.
+If the attacker moves after the period of $\delta_{D}-d$, the gain of the attacker will be zero.
+The probability that this happens is  $\dfrac{d}{\delta_{D}}$.
+The average gain rate of the attacker can then be expressed as follows if we look at one interval of the defender:
+\begin{equation}\label{first}
+\gamma_{A}(\alpha_{D},\alpha_{A}) = \dfrac {1}{\delta_{D}} [ \dfrac{\delta_{D}}{\delta_{A}} \cdot \dfrac{\delta_{D}-d}{\delta_{D}} \cdot \dfrac{\delta_{D}-d}{2} + \dfrac{\delta_{D}}{\delta_{A}} \cdot \dfrac{d}{\delta_{D}} \cdot 0 ]
+\end{equation}
+
+To derive the benefit, the cost of moving is subtracted from the average gain. 
 \begin{equation}\label{first}
@@
 \begin{equation}\label{first}
-\beta_{A}(\alpha_{D},\alpha_{A}) = \dfrac { \delta_{D}} {2 \cdot \delta_{A}} - k_{A} \alpha_{A} + \dfrac{d}{\delta_{A}} + \dfrac{d^{2}}{2 \cdot \delta_{A} \delta_{D}}
+\beta_{A}(\alpha_{D},\alpha_{A}) = \dfrac { \delta_{D}} {2 \cdot \delta_{A}} - k_{A} \alpha_{A} - ( \dfrac{d^{2}}{2 \cdot \delta_{A} \delta_{D}} - \dfrac{d}{\delta_{A}} )
 \end{equation}
@@
 
-For the case with a virus we consider two cases, \textbf{Case a} and \textbf{Case b}, to compute the benefit of both players.  \\
+For the case with a virus we consider two cases, Case a and Case b, depending on whether the delay is shorter or longer than the difference between the attacker's and the defender's period.  \\
 
 \subsubsection*{\textbf{Case a:} $d + \delta_{A} \leq \delta_{D}$}
+~~\\
+Consider a timespan $\delta_{A} + d$, representing the attacker's interval followed by the delay period in his next interval.
+The defender will never move twice during this timespan because $\delta_{A} + d \leq \delta_{D}$.
+The defender will move during the interval of the attacker with a probability of $\dfrac{\delta_{A}}{\delta_{D}} $.
+When this happens the defender will end with being in control at the end of the interval.
+In the next interval the attacker will have to regain control, meaning that during the delay, the defender stays in control, see figure \ref{fig:case2} cases (1) and (2).
+This means that the defender will keep the control over the resource in the next interval over a period of the delay, namely \textit{d}.
+Because $d + \delta_{A} \leq \delta_{D}$ the next move of the defender in this second interval will never occur during the delay, meaning that the entire delay can be considered as an extra benefit resulting of a play in the previous interval.
+So, every time the defender plays, he will get an average gain of $\dfrac{\delta_{A}}{2}$ in the interval where he plays and in the next interval will always receive a extra gain of $d$, yielding a total average gain per interval of
+$\dfrac{(d+\dfrac{\delta_{A}}{2})}{\delta_{A}}$
 
-If $d + \delta_{A} \leq \delta_{D}$ that means that in this timespan for the attacker, the defender will never move twice.
-$\dfrac{delta_{A}}{\delta_{D}}$ the defender will move during the interval of the attacker.
-When this happens the defender will end with beeing in control at the end of the interval.
-In the next interval the attacker will move in the beginning but will not gain control after the delay of \textit{d}.
-This means that if the defender moved in the interval prior to this one, the delay is the minimum gain of the defender.
-To calculate the gain rate of the defender we don't look at one interval of the defender.
-We know that if the defender moves within the interval of the defender it will receive a minimum of $\dfrac{\delta_{A}}{2}$ gain.
-In the next interval the defender will receive a minimum gain of $d$.
-It is also possible that the defender will move in this interval after the delay, adding a gain of $\dfrac{\delta_{A}}{2}$, but the defender will never play during the delay if in the interval prior to this, the defender already has moved.
+The total gain  rate of the defender is then the probability that the defender will move during an interval of the attacker multiplied by the total average gain per interval:
 
-This means that every time the defender plays it will get a benefit of $\dfrac{\delta_{A}}{2}$ and the next interval will always receive a gain of $d$.
-To derive the benefit formula for the defender we can calculate the probability that the defender will move during an interval of the attacker and multiply it by the gain of the interval plus the gain that it will give to the next interval.
-This yields in the following formula:
 \begin{equation}\label{first}
-\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta{A}}{\delta_{D}} \cdot \dfrac{(d+\dfrac{\delta_{A}}{2})}{\delta_{A}} - k_{D} \alpha_{D} 
+\gamma_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta_{A}}{\delta_{D}} \cdot \dfrac{(d+\dfrac{\delta_{A}}{2})}{\delta_{A}} 
 \end{equation}
 \begin{equation}\label{first}
-\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - k_{D} \alpha_{D} 
+\gamma_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} 
+\end{equation}\\
+This yields in the following benefit formula:
+\begin{equation}\label{first}
+\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - k_{D} \alpha_{D} 
 \end{equation}\\
@@
 \centering
-\includegraphics[scale=0.7]{Images/FlipItCase2.png}
+\includegraphics[scale=0.5]{../../doc/template/Images/FlipItCase2.png}
+\label{fig:case2}
 \end{figure}
 
-It is crucial that the $d + \delta_{1}$ is not greater than $ \delta_{0}$.
-If it is greater that means that the defender can move during a delay after an interval of the attacker where the defender already has moved.
-If we calculate the benefit formula in the same manner, to much gain is added.
-Namely the overlap during the delay.
+It is crucial that $ \delta_{D}$ is at least as large as $d + \delta_{A}$.
+If not, this would mean that the defender can move during the delay in the interval following the interval where the defender already moved.
+This would mean that there can be an overlap between the average gain of $\dfrac{\delta_{A}}{2}$ and the delay.
+The above benefit formula would then include to much gain for the defender: the potential overlap during the delay would be counted twice.
+\\
 
-\subsubsection*{\textbf{Case b:} $d + \delta_{1} > \delta_{D}$}
+~~ \\
+\subsubsection*{\textbf{Case b:} $d + \delta_{A} \geq \delta_{D}$}
+~~~\\
+
+To obtain the formula in case of a too long delay, we therefore need to subtract this overlapping gain from the above formula. 
+Since $\delta_{D} \geq \delta_{A}$, if the defender enters the interval immediately after the attacker has played, then the defender cannot have played in the previous interval.
+In that case, there is no overlap.
+So the problem of the overlap only appears if the defenders enters late enough and thus only the last part of the delay is subject to overlap.
+The larger the difference between the interval of the defender and the attacker, the smaller the risk of overlap.
+Concretely, only the last part of length $d - (\delta_{D} - \delta_{A})$ is subject to overlap.
+Hence, the probability of overlap is $\dfrac{ d - (\delta_{D} - \delta_{A})}{\delta_{D}}$ and the gain will be half of this interval:  $\dfrac{ d - (\delta_{D} - \delta_{A})}{2}$.
+The gain rate to be subtracted is therefore:\\
+
+\begin{equation}\label{first}
+\dfrac{1} {\delta_{A}} \cdot \dfrac{d - (\delta_{D} - \delta_{A})}{\delta_{D}} \cdot \dfrac{d - (\delta_{D} - \delta_{A})}{\delta_{D}}
+\end{equation}
+
+The total gain  rate of the defender is obtained by subtracting this term from the gain rate of case a:
+ \begin{equation}\label{first}
+\gamma_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta_{A}}{\delta_{D}} \cdot \dfrac{(d+\dfrac{\delta_{A}}{2})}{\delta_{A}} - \dfrac{(d - (\delta_{D} - \delta_{A}))^{2}}{2 \delta_{D} \delta_{A}}
+\end{equation}
+\begin{equation}\label{first}
+\gamma_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - \dfrac{(d - (\delta_{D} - \delta_{A}))^{2}}{2 \delta_{D} \delta_{A}}
+\end{equation}\\
+This yields in the following benefit formula:
+\begin{equation}\label{first}
+\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - k_{D} \alpha_{D} - \dfrac{(d - (\delta_{D} - \delta_{A}))^{2}}{2 \delta_{D} \delta_{A}}
+\end{equation}\\
+
+The benefit for the attacker will be as follows:
+\begin{equation}\label{first}
+\beta_{A}(\alpha_{D},\alpha_{A}) = 1 -\dfrac{\delta_{A}}{2\delta_{D}} - \dfrac{d}{\delta_{D}} - k_{A} \alpha_{A} + \dfrac{(d - (\delta_{D} - \delta_{A}))^{2}}{2 \delta_{D} \delta_{A}}
+\end{equation}\\
 

### 2015-05-22 IEEE Paper/paper/paper.tex
Toevoegen nieuw hoofdstuk

@@
 \title{Flip the virus:\\ A gametheoretic approach to cybersecurity}
-\author{Sophie Marien}
+\author{Sophie Marien\\
+DistriNet KULeuven \\
+sophie.m.marien@gmail.com}
 \maketitle
@@
 Many businesses store confidential information on clients, which can be lost and possible be abused by competitors through data leakage.
-Also, disruption caused by DOSS attacks, may results in businesses failing to meet their service-level agreements.
+Also, disruption caused by DDoS attacks, may results in businesses failing to meet their service-level agreements.
 Ultimately, system and network security helps protecting a business's reputation, which is one of its most important assets.

### 2015-05-22 doc/template/intro.tex
Toevoegen nieuw hoofdstuk

@@
 \subsection{Conclusions}
-The "I love you" virus is an example of a virus that spreads quickly.
-This virus propagates via mail systems.
-If someone opens an email with "I love you" virus in annex this virus spreads itself by sending a mail itself to everyone in your contact list.
-So the virus can multiply rapidly and eventually a business network shut down by the heavy traffic.
-In this example, there is a need human interaction to spread the virus to do.
-If no one opens the virus can not spread the mail.
-Unfortunately, there are viruses that can spread without human interaction.
-These viruses are referred to as worms.
-A worm is also a computer program that replicates itself to spread to other computers so.
-Via a computer network, copies of the worm forwarded without an intermediary is used for.
-The worm will use vulnerabilities to infect other computers.
-Most worms are designed to spread out and just try not to make any changes to the systems that they pass.
-These worms can still inflict damage by increased network traffic they generate.
-Worms that contain Harm damage a program to install a backdoor or a rootkit on the infected computers.
-Backdoors and rootkits ensure that future use can be made of the infected computers.
-The Stuxnetworm is a very famous worm.
-Initially this worm spread via infected USB sticks and from then it could spread through the Internet to other computers.
-The purpose of the Stuxnetworm was broken to run the centrifuges in nuclear reactors.
-Many reactors have been infected.
-From the standpoint of the defender, it is very important to respond as quickly as possible so that the worm can not spread quickly.
 

### 2015-05-24 doc/template/chap-9.tex

@@
 
-\section{Simulation}
-%nce the defender has played, it takes 1 time until the attacker plays and the defender loses control again.
-Therefere, the expected period of attacker control within the interval would be 1/2.1/0, without considering the delay.
+\section{Nash Equilibria}
+-- rechtstreeks uit FlipIt paper --\\
+As a second step, we are interested in finding Nash equilibria, points
+for which neither player will increase his benefit by changing his rate of play. More
+formally, a Nash equilibrium for the periodic game is a point $(\alpha_{0}^{*},\alpha_{1}^{*})$ such that
+the defender's benefit $\beta_{0}(\alpha_{0},\alpha_{1}^{*}) $is maximized at $\alpha_{0}= \alpha_{0}^{*}$ and the attacker's benefit
+$\beta_{1}(\alpha_{0}^{*},\alpha_{1}) $ is maximized at $\alpha_{1}= \alpha_{1}^{*}$ .
+To begin with, some useful notation. We denote by opt0($\alpha_{1}$) the set of values (rates
+of play $\alpha_{0}$) that optimize the benefit of the defender for a fixed rate of play $\alpha_{1}$ of the
+attacker. Similarly, we denote by opt1($\alpha_{0}$) the set of values (rates of play $\alpha_{1}$) that optimize
+the benefit of the attacker for a fixed rate of play $\alpha_{0}$ of the defender. The following
+theorem specifies Nash equilibria for the periodic game and is proven in Appendix A. \\
 
+Nash equilibria are points whith the property that neither player benefits by deviating in isolaition form equilibrium.
+We can compute Nash Equilibria for the periodic game as an intersection points of curvest opt0 and opt1.
+To determine opt0(a0) we need to compute the derivate of  $\beta_{0}(\alpha_{0},\alpha_{1}) $ for a fixed $\alpha_{1}$.
+We consider two cases:\\
+-----
+
+\subsection{case: $\delta_{D} \leq \delta_{A} $ or $\alpha_{D} \geq \alpha_{A}$}
+
+Since 
+\begin{equation}\label{formdelta}
+\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} = \frac{\partial [1 - \dfrac { (\delta_{D}-d) ^{2}} {2 \cdot \delta_{D}  \delta_{A}]} - k_{D} \alpha_{D}}{\partial \alpha_{D}}
+\end{equation}
+
+If we change every $\delta = \dfrac{1}{\alpha}$ in formule \ref{formdelta} we get the next formula that is easier to take the derivate of:
+\begin{equation}
+\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} =\frac{\partial [1 - \dfrac { (1/\alpha_{D}-d) ^{2} \cdot \alpha_{D} \alpha_{A}} {2} - k_{D} \alpha_{D}]}{\partial \alpha_{D}}
+\end{equation}
+Which gives the following result:
+\begin{equation}
+\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} = \alpha_{A} ( \dfrac{1}{\alpha_{D}^{2}} - \dfrac{d^{2}}{2}) - k_{D}
+\end{equation}
+
+it follows that if equal to zero $\alpha_{0}=\sqrt{\dfrac{\alpha_{A}}{2k_{D} + d^{2}\alpha_{A}}}$
+
+\subsection{case 2: $\delta_{D} \geq \delta_{A} $ or $\alpha_{D} \leq \alpha_{A}$}
+
+\textbf{case A}\\
+
+\begin{equation}
+\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} = \frac{\partial [ \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - k_{D} \alpha_{D}] }{\partial \alpha_{D}}
+\end{equation}
+\begin{equation}
+\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} = \frac{\partial [ \dfrac{\alpha_{D}}{2\alpha_{A}} + d\alpha_{D} - k_{D} \alpha_{D}] }{\partial \alpha_{D}}
+\end{equation}
+\begin{equation}
+\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} = \dfrac{1}{2\delta_{A}} + d - k_{D}
+\end{equation}
+
+Dus increasing als $\alpha_{A} < \dfrac{1}{2(k_{D} -d)}$ and decreasing if $\alpha_{A} > \dfrac{1}{2(k_{D} -d)}$
+
+\textbf{Case b} \\
+
+\begin{equation}
+\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} = \frac{\partial [ \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - k_{D} \alpha_{D} - \dfrac{(d - (\delta_{D} - \delta_{A}))^{2}}{2 \delta_{D} \delta_{A}}] }{\partial \alpha_{D}}
+\end{equation}
+
+Uitgewerkt geeft de benefit formule de volgende formule omgezet naar alpha
+\begin{equation}
+\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} = \frac{\partial [ 1 - \dfrac{\alpha_{A}}{2\alpha_{D}} - k_{D}\alpha_{D} + \dfrac{d \alpha_{A}}{2} + \dfrac{d \alpha_{D}}{2} - \dfrac{d^{2} \alpha_{A} \alpha_{D}}{2} ] }{\partial \alpha_{D}}
+\end{equation}
+
+\begin{equation}
+\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} = \dfrac{\alpha_{A}}{2\alpha_{D}^{2}} - k_{D} + \dfrac{d}{2} - \dfrac{d^{2} \alpha_{A}}{2}
+\end{equation}
+
+Als ge dat gelijk stelt aan 0 krijgt ge:
+\begin{equation}
+\alpha_{D} = \sqrt{\dfrac{\alpha_{A}}{2k_{D} -d + d^{2}\alpha_{A}}}
+\end{equation}
 %%% Local Variables: 

### 2015-05-24 doc/template/chap-10.tex
verbeteringen van paper

@@
 Many businesses store confidential information on clients, which can be lost and possible be abused by competitors through data leakage.
-Also, disruption caused by DOSS attacks, may results in businesses failing to meet their service-level agreements.
+Also, disruption caused by distributed DoS attacks, may results in businesses failing to meet their service-level agreements.
 Ultimately, system and network security helps protecting a business's reputation, which is one of its most important assets.
@@
 
-A particular kind of frequently occurring threats are Advanced Persistent Threats (APT).
-An APT is a targeted cyber attack that targets organisations in a stealthy way and that can stay undetected for a long period.
+A particular kind of threats are Advanced Persistent Threats (APT).
+\todo{misschien meer zeggen over APTs en threats) An APT is a targeted cyber attack that is designed to penetrate a network or a system in a stealthy way and that can stay undetected for a long period.
 This makes it so hard to protect a network or a system against an APT.
-Bruce Schneier describes an APT as something different and stronger than a conventional threat: ''\textit{A conventional hacker or criminal isn't interested in any particular target.
+Bruce Schneier describes an APT as something different and stronger than a conventional threat: ''\textit{A conventional hacker or criminal is not interested in any particular target.
 He wants a thousand credit card numbers for fraud, or to break into an account and turn it into a zombie, or whatever.
@@
 Against this sort of attacker, the absolute level of your security is what's important.
-It doesn't matter how secure you are compared to your peers; all that matters is whether you're secure enough to keep him out}'' - Bruce Schneier \cite{APTBruce}.\\
+It does not matter how secure you are compared to your peers; all that matters is whether you're secure enough to keep him out}'' - Bruce Schneier \cite{APTBruce}.\\
 
@@
 In particular, both players will aim for the strategy that results in a maximal benefit for them.
-Researchers at RSA made a game theoretic framework to model targeted attacks.
-They study the specific scenario where a system or network is repeatedly taken over completely by an attacker and this attack is not immediately detected by the defender of the system or network.
+Researchers at RSA, van Dijk et al,  made a game theoretic model of targeted attacks.
+They study the specific scenario where a system or network is repeatedly taken over completely by an without being immediately detected by the defender of the system or network.
 In game theory, such a game is known as ''FlipIt'' \cite{FlipIt}.
@@
 Both players do not know who is currently in control of the resource until they move.
-In FlipIt every move gives them immediately control over the resource.
+In FlipIt every move gives them immediate control over the resource.
 But what if the attacker moves and it takes a while before the attacker gets full control over the resource?

### 2015-05-24 doc/template/chap-5.tex
verbeteringen van paper

@@
 The game begins at $t=0$ and continues indefinitely ($t \rightarrow \infty $).
-The time in the game is assumed as being continuous, but a discrete time could also be considered.
+The time in the game is assumed as being continuous.
 To get control over the resource, the players $i$, with $i \in \{A,D\}$, can flip the resource at any given time.
-A flip will be regarded as a move from a player \textit{i}.
-Each move will imply a certain cost $k_{i}$ and the cost can vary for each player.
-Both players will try to minimize their cost.
-Adding a cost will prevent players to move too frequently.
-\\
+%A flip will be regarded as a move from a player \textit{i}. 
+Each move implies a certain cost $k_{i}$ and can vary for each player. Both players try to minimize their cost. Adding a cost prevents players to move too frequently. \\
 
-The unique feature of FlipIt is that every move will happen in a stealthy way, meaning that the player has no clue that the other player (his adversary) has flipped the resource.
-For instance, the defender will not find out if the resource has already been compromised by the attacker until he flips the resource himself.
+The unique feature of FlipIt is that every move happens in a stealthy way, meaning that the player has no clue that the other player (his adversary) has flipped the resource.
+For instance, the defender does not find out if the resource has been compromised by the attacker until he flips the resource himself.
 The goal of the player is to maximize the time that he or she has control over the resource while minimizing the total cost of the moves.
@@
 $C_{D}(t)$ is 1 if the game is under control by the defender and 0 if the game is under control by the attacker.
-Reversely, $C_{A}(t)$ will be 1 if the game is under control by the attacker and 0 if under control by the defender.
+Reversely, $C_{A}(t)$ is 1 if the game is under control by the attacker and 0 if under control by the defender.
 So, $C_{A}(t)= 1 - C_{D}(t)$.
@@
 \end{equation}
-If we add up the gain of the defender and the gain of the attacker it should sum up to t:
+If we add up the gain of the defender and the gain of the attacker it should sum up to \textit{t}:
 \begin{equation}\label{first}
@@
 ~$ \alpha_{i}$ defines the average move rate by player \textit{i} up to time \textit{t}.
-In a given game, the asymptotic benefit rate (or simply benefit) will be defined as the lim inf of the average benefit because time t will increase to infinity and the average benefit may not have limiting values.
+In a given game, the asymptotic benefit rate (or simply benefit) will be defined as the lim inf of the average benefit because time\textit{ t} will increase to infinity and the average benefit may not have limiting values.
 \begin{equation}
@@
 
-If there is no feedback for neither of the players, we have a non-adaptive strategy.
+If there is no feedback for either player, we have a non-adaptive strategy.
 Because a player does not receive any feedback during the game he will play in the same manner against every opponent.
@@
 \\
-In this paper we will focus on the non-adaptive strategies.
+In this paper we restrict ourselves to periodic strategies.
 This choice is motivated by the fact that in a security game a player (defender or attacker) rarely has information about the moves (last move or full history) of his opponent.
@@
 This means that the attacker can base his attacks on the defender's moves.
-Both the targeted and non-targeted attacks don't succeed immediately.
+Both the targeted and non-targeted attacks do not succeed immediately.
 For the targeted attack the time till it succeeds is given by an exponential distributed random variable with a known rate.

### 2015-05-25 doc/template/chap-10.tex
toevoegingen aan de tekst

@@
 A particular kind of threats are Advanced Persistent Threats (APT).
-\todo{misschien meer zeggen over APTs en threats) An APT is a targeted cyber attack that is designed to penetrate a network or a system in a stealthy way and that can stay undetected for a long period.
+\todo{misschien meer zeggen over APTs en threats} An APT is a targeted cyber attack that is designed to penetrate a network or a system in a stealthy way and that can stay undetected for a long period.
 This makes it so hard to protect a network or a system against an APT.

### 2015-05-25 doc/template/chap-5.tex
commit

@@
 %There are two standard representations for games. The first one is the Normal Form. The second one is the Extensive Form.
-\subsubsection{List of terms}
-In the following list a couple of terms that will be used throughout the paper.
-\begin{description}
-\item \textit{Players}: Players are referred as the ones who are the decision makers. It can be a person, a company or an animal.  (they will act rational )
-\item \textit{Actions}: Every player has actions that he or he can do. 
-\item \textit{Strategies}: A strategy is the combination of different actions. A pure strategy is only one action.
-\item \textit{Utility function}: The utility function is the mapping of the level of happiness of an agent about the state of the world to natural numbers.
-
-\end{description}
-
-A game in game theory consists of multiple agents and every agent has a set of actions that he can play.
-
-\todo{strategien en acties definieren}
 
@@
 A Nash Equilibrium is a subset of outcomes that can be interesting to analyse a game.
+To define this concept we first introduce the concept of best response.
+The best response for a player is the action of a player that maximizes it's pay-off for any given action of the other player.
+We define $BR_{i}$ as the best response function for player \textit{i}.
+The best response for player \textit{1} is given by : $a_{1} = BR_{1}(a_{2})$.
+A
 For a Nash Equilibrium each player has a consist list of actions and each player's action maximizes his or her pay-off given the actions of the other players.
 Nobody has the incentive to change his or her action if an equilibrium profile is played.
-In general we can say that a Nash Equilibrium is a stable strategy profile: each player is considered to know the equilibrium strategies of the other players and no player would want to change his own strategy if he knows the strategies of the other players.
-
-Formal definition of a Nash Equilibrium:
-A strategy profile $s = (s1, . . . , sn)$ is a Nash equilibrium
-if, for all agents $i$, $si$ is a best response to $s-i$ .
-''Intuitively, a Nash equilibrium is a stable strategy profile: no agent would want to change
-his strategy if he knew what strategies the other agents were following.
-We can divide Nash equilibria into two categories, strict and weak, depending on whether
-or not every agent's strategy constitutes a unique best response to the other agents' strategies.''
-
-\todo{Nash beter uitleggen nog met best response erbij}
+We have a Nash Equilibrium for the pair $(a_{1}^{*},a_{2}^{*})$ where $a_{1}^{*} = BR_{1}(a_{2}^{*})$ and $a_{2}^{*} = BR_{2}(a_{1}^{*})$\\
 
@@
 
+\subsubsection{List of terms}
+In the following list a couple of terms that will be used throughout the paper.
+\begin{description}
+\item \textit{Players}: Players are referred as the ones who are the decision makers. It can be a person, a company or an animal.  (they will act rational )
+\item \textit{Actions}: Every player has actions that he or he can do. 
+\item \textit{Strategies}: A strategy is the combination of different actions. A pure strategy is only one action.
+\item \textit{Utility function}: The utility function is the mapping of the level of happiness of an agent about the state of the world to natural numbers.
+
+\end{description}
+
+A game in game theory consists of multiple agents and every agent has a set of actions that he can play.
 \section{The FlipIt game}

### 2015-06-04 doc/template/chap-9.tex
abstract

@@
 
-%------------------------------------------------%
-%            Intro Game Theory 					 %
-%------------------------------------------------%
-\subsection{Actions of the attacker}
-A virus has different kind of ways of making his way through a company network.
+%%------------------------------------------------%
+%%            Intro Game Theory 					 %
+%%------------------------------------------------%
+%\subsection{Actions of the attacker}
+%A virus has different kind of ways of making his way through a company network.
 We will describe the different ways of how the virus can propagate.
 For start we will say that the virus or worm will be dropped on Node i and that it has k numbers of neighbours.
-\begin{enumerate}
-\item Node i is infected and will spread the virus or worm to every k neighbours and will stop infecting the neighbours in the next step
-\item Node i is infected and will spread the virus or worm to every k neighbours and will keep on spreading the virus to the same neighbours in every next step
-\item Node i is infected and will spread the virus to only one of the k neighbours and will stop infecting another neighbour in the next step
-\item Node i is infected and will spread the virus to only one of the k neighbours and in the next step it will infect another one of the k neighbours 
-\end{enumerate}
-
-In the game that will be modelled in the paper we will use the settings of the first spreading method.
+%\begin{enumerate}
+%\item Node i is infected and will spread the virus or worm to every k neighbours and will stop infecting the neighbours in the next step
+%\item Node i is infected and will spread the virus or worm to every k neighbours and will keep on spreading the virus to the same neighbours in every next step
+%\item Node i is infected and will spread the virus to only one of the k neighbours and will stop infecting another neighbour in the next step
+%\item Node i is infected and will spread the virus to only one of the k neighbours and in the next step it will infect another one of the k neighbours 
+%\end{enumerate}
+%
+%In the game that will be modelled in the paper we will use the settings of the first spreading method.
 We will not use method 2 because this kind of propagation will float the network.
@@
 \\
-In the first method the node that has been infected can be again infected.
+%In the first method the node that has been infected can be again infected.
 If one of the neighbours infects the node again the node will infect his neighbours again.
@@
 An infected state means that the node is infected and will not spread the virus to its neighbours, a clean state means that the node is not infected on that moment and a spreading state means that the node is infected and that it will spread the virus or worm to its neighbours in the next step.
-We can argument this kind of propagation through a mail worm. \todo{voorbeeld geven van zo een worm}
-%Another propagation method is that the virus works as a token. It will propagate to only one neighbour and continue to spread.
-
-The Attacker itself has two different ways of attacking the company network.
+%We can argument this kind of propagation through a mail worm. \todo{voorbeeld geven van zo een worm}
+%%Another propagation method is that the virus works as a token. It will propagate to only one neighbour and continue to spread. 
+%
+%The Attacker itself has two different ways of attacking the company network.
 It will only infected one node of the network and will wait for the virus to spread itself through the network.
 We will model two ways of attacks of an Attacker:
-\begin{enumerate}
-\item The attacker drops the virus on a random node on the network
-\item The attacker drops the virus on a targeted node on the network
-\end{enumerate}
-The attacker in this game will put a virus or worm on one of the nodes in the network.
+%\begin{enumerate}
+%\item The attacker drops the virus on a random node on the network
+%\item The attacker drops the virus on a targeted node on the network
+%\end{enumerate}
+%The attacker in this game will put a virus or worm on one of the nodes in the network.
 (This will happen at random.) The attacker does not know on which node the virus will be dropped.
@@
 \\
-
-The attacker can choose at which rate it will drop a virus on one of the nodes on the network.
+%
+%The attacker can choose at which rate it will drop a virus on one of the nodes on the network.
 The cost of dropping a virus will be the same.
@@
 If it will increase this means that the attacker will eventually drop out of the game because it becomes to expensive.\\
-The attacker is in control over the game if it manages to infect a subset of all the resources of the company network.
-
-\subsection{Actions of the defender}
-The attacker wants to protect all the nodes of his network.
+%The attacker is in control over the game if it manages to infect a subset of all the resources of the company network.
+%
+%
+%\subsection{Actions of the defender}
+%The attacker wants to protect all the nodes of his network.
 It can do so by getting back control over the resources.
@@
 Which is convenient in the real world because a company has to know how his infrastructure looks like.\\
-
-The defender has two possible ways of defending its network:
-\begin{enumerate}
-\item The defender flips all the nodes of his network
-\item The defender will flip a subset of the nodes of his network
-\end{enumerate}
-
-The cost of flipping all the nods of the network will be greater than the cost of flipping a subset of nodes.
+%
+%The defender has two possible ways of defending its network:
+%\begin{enumerate}
+%\item The defender flips all the nodes of his network
+%\item The defender will flip a subset of the nodes of his network
+%\end{enumerate}
+%
+%The cost of flipping all the nods of the network will be greater than the cost of flipping a subset of nodes.
 We make this assumption because otherwise it will be beneficial for the defender to always flip all the nodes in the network.\\
-
-We will also make the assumption that as a defender flips a node the node can get infected again.
+%
+%We will also make the assumption that as a defender flips a node the node can get infected again.
 A flip will not be  correlated to a patch but to a clean-up.
 \todo{waarom geen patch, wormen kunnen veranderen gaandeweg}
-\todo{andere mogelijkheid:} Another setting of the game can be that the flip of the defender is equal to a patch and that the resource cannot be infected any more.
+%\todo{andere mogelijkheid:} Another setting of the game can be that the flip of the defender is equal to a patch and that the resource cannot be infected any more.
 But with this case we deviate from the flipIt game, because the attacker cannot flip the resource any more.
@@
 In this section we provide a formal definition of the game and the notation that we will use throughout the paper.
+
+\begin{description}
+\item $\delta_{i}$
+\item 
+\end{description}
 

### 2015-06-07 doc/template/chap-10.tex
aanvullingen tekst

@@
 \begin{itemize}
-\item Is it possible to incorporate the notion of delay in the game-theoretical analysis of the Flip-It game ?
-\item Does this allow us to determine an optimal defense strategy against an attacker ?
-for example: Gaat er een specieke grootte zijn van een delay waarbij de attacker al weet dat hem niet meer moet gaan spelen ?
-( is niet gelijk aan de grootte van de periode van de attacker)
-\end{itemize}
-
-Resarch questions when working with a network and a delay: .. graph model en uitleggen hoe we de graph kunnen maken zodat de delay altijd zo groot mogelijk gaat zijn. 
-\begin{itemize}
-\item How can we calculate the expected duration for a node's infection/the entire network infection ?
-\item Can we calculate this node per node ?
+\item How can we incorporate the notion of delay in the game-theoretical analysis of the Flip-It game for a periodic strategy ?
+\item Is there an optimal defense strategy against an attacker ? 
 \end{itemize}

### 2015-06-07 doc/template/chap-5.tex
abstract
aanvullingen tekst

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{What is cyber security?}
 
+Cyber security is a part of security that focuses on securing computer controlled devices such as computers and smartphones, as well as public and private computer networks, including the entire Internet.
+A private network can be an enterprise network that is protected from the public network such as the internet.
+The purpose of security is to give certainty that data will not be removed without authorization (Confidentiality) that the data is always accessible (Availability), and that data can not be read or altered by someone who does not have the authorization (Integrity).
+These are the 3 key attributes of security, also known as the CIA triad.
+\\
+
+In order to keep a system secure, it is important to mitigate the possible threats.
+A threat refers to anything that can breach the security and that can cause a possible harm.
+The some of the most common threats are spam, malware, spoofing, phishing and  DDoS attacks.
+The biggest threat are caused by malware. 
+This is a piece of malicious program that is designed to penetrate unprotected systems or computers, and getting there to sensitive information.
+Malware can be divided into different subclasses.
+These subclasses are not mutually exclusive:
+
+\begin{description}
+\item Virus: This is a malicious piece of code that replicates and tries to spread to infect other systems or files.
+The "I love you" virus is an example of a virus that quickly spreads.
+This virus propagates via mail systems.
+If someone opens an email with "I love you" with the virus in the annex this virus spreads itself by sending a mail to everyone in your contact list.
+So the virus can multiply rapidly and eventually a business network might shut down by the heavy traffic.
+In this example, there is a need for human interaction to spread the virus.
+If no one opens the mail the virus can not infect other systems.
+Unfortunately, there are also viruses that can spread without human interaction.
+These viruses are referred to as worms.
+A worm is also a computer program that replicates itself to spread to other computers.
+Via a computer network, copies of the worm can be forwarded without an intermediary.
+The worm will use vulnerabilities of the system to infect other computers.
+The Stuxnetworm is a very famous worm.
+Initially this worm spread via infected USB sticks and from then it could spread through the Internet to other computers.
+The purpose of the Stuxnetworm was to harm the centrifuges in nuclear reactors.
+Many reactors have been infected.
+From the standpoint of the defender, it is very important to respond as quickly as possible so that the worm can not spread.
+\item Rootkit:
+\item Trojan:
+\item Backdoors:
+\end{description}
+
+A security report of 2014 reveales that 61\% of the attacks on companies are caused by malware. A company can take different measures to defend himself against malware. 
+The threats caused by malware can be devided into 3 big groups: known threats (70\%), unknown threats (29\%) and advanced threats (1\%) [\todo{toevoegen referentie naar kaspersky APT report}].
+
+The known threats easy to defend against. Unkown also. Then the remaining 1\% are the advanced threats. These threats are also known as APT. APT's are ... 
+In this paper we focus on targeted attacks.
+Targeted attacks are malware designed to attack exclusively an organisation.
+Some examples of the biggest most rare APTs  to get an understanding of what APTs are capable of and how long they can stay unseen.
+[\todo{site kaspersky apt}]
+
+An APT is a persistent targeted attack that tries to penetrate a network, harm and stays unseen for a long period of time.
+The motive of an APT is mostly cyber espionage, stealing sensitive data, sabotage or other ideological attacks.
+APT stands for Advanced persistant threat.
+Advanced stands for the fact that these attacks are well funded and that the attacker itself has a great expertise.
+These APT are not easy to protect against.
+Not all APTs are technical advanced.
+The attacker can also try to exploit existing vulnerabilities and hope that his target has not yet secured himself against these vulnerabilities.
+Persistent is the fact that the attacker keeps on trying to break in to his victim.
+He will not give up.
+The attack can be over various years and different steps.
+The threats stands for the fact that an APT is ..
+
+\subsubsection{Equation}
+Equation is a complex cyberattack platform where the first known sample is from 2002, but it was only discovered 12 years later in 2014.
+This APT propagates through usb drives, cd or physical media.
+It will search for exploits and will self-replicate itself to spread the infection.
+The purpose of this virus is to steal data and cyberespionage.
+
+\subsubsection{Regin}
+
+\subsubsection{Flame}
+Way of propagation through USB drives, LAN spreading. purpose cyber espionage.
+
+\subsubsection{Black energy}
+ purpose cyber espionage and DDoS, data wiping. prop usb lan
+
+According to [\todo{APT report} ]  the damage of one successful target attack can exceed over 2.54 million dollar.
+a company needs a defense mechanism to defend itself against apt.
+Prevention more important than ..
+een van de manieren om dat te doen is via game theorie.
+
+Security belangrijk. verschillende malware. uitleggen malware. verwijzen naar kaspersky security raport. APT grote threats. voorbeelden van APT's. waarom gametheory gebruiken.
+
+\section{Intro game theory}
 Gametheory is a mathematical study to analyse interactions between independent and self-interested agents.
@@
 We have a Nash Equilibrium for the pair $(a_{1}^{*},a_{2}^{*})$ where $a_{1}^{*} = BR_{1}(a_{2}^{*})$ and $a_{2}^{*} = BR_{2}(a_{1}^{*})$\\
+
+ \todo{optimal strategy uitleggen}
 
@@
 
-The state of the resource is denoted as a time-dependent variable $C=C_{i}(t)$. 
+We denote the state of the resource as a time-dependent variable $C=C_{i}(t)$. 
 $C_{D}(t)$ is 1 if the game is under control by the defender and 0 if the game is under control by the attacker.

### 2015-06-07 doc/template/chap-9.tex
aanvullingen tekst

@@
 \begin{description}
-\item $\delta_{i}$
-\item 
+\item $i$: Defines the player. Different as in the FlipIt paper where the defender is denoted by the subscript 0 and the attacker by the subscript 1.
+\item $\delta_{i}$: The length of the interval between two consecutive moves of player \textit{i}. 
+\item $\alpha_{i}$: The average flip rate of player \textit{i}, given by $\alpha_{i}=1/\alpha_{i}$.
+\item $k_{i}$: The cost of player \textit{i}'s moves.
+\item $d$: The delay caused by the virus propagation.
+\item $G_{i}(t)$: The total gain of player \textit{i} denotes the amount of time player \textit{i} is in control over the resource up to time \textit{t}.
+\item $\gamma_{i}$: The average gain rate of player \textit{i} defined as $G_{i}(t)/t$
+\item $\beta_{i}$:  The average benefit rate up to time \textit{t} defined as  $\beta_{i} = \gamma_{i} -k_{i} \alpha_{i} $.
+\item $opt_{i}$: The optimum function.
 \end{description}
+
+\todo{tekening toevoegen waarop de delay staat aangeduid}
 
@@
 
+Intuitively we could assume that $d$ can never be bigger than $\delta_{A}$ because then the benefit for the defender would always be 1.
+This is not always true.
+It is only true if $d$ is bigger than $\delta_{D}$, because then the defender will always be in control.
+For this we only calculate the formulas for the cases where $d$ is smaller than $\delta_{D}$.
+We can already conclude that it is no use for the attacker to play when the delay is bigger than $\delta_{D}$.
+\\
+
 \subsection{Formalization the benefit formula including the infection-delay}
@@
 
-\begin{equation}\label{first}
+\begin{equation}
 \beta_{A}(\alpha_{D},\alpha_{A}) =\dfrac {r} {2} - k_{A} \alpha_{A} = \dfrac {\delta_{D}} {2\delta_{A}} - k_{A} \alpha_{A}  
@@
 Correspondingly, the benefit for the defender can be expressed as:
-\begin{equation}\label{first}
+\begin{equation}
 \beta_{D}(\alpha_{D},\alpha_{A}) =1 -  \dfrac {r} {2} - k_{D} \alpha_{D} = 1 - \dfrac {\delta_{D}} {2\delta_{A}} - k_{D} \alpha_{D} 
@@
 The average gain rate of the attacker can then be expressed as follows if we look at one interval of the defender:
-\begin{equation}\label{first}
+\begin{equation}
 \gamma_{A}(\alpha_{D},\alpha_{A}) = \dfrac {1}{\delta_{D}} [ \dfrac{\delta_{D}}{\delta_{A}} \cdot \dfrac{\delta_{D}-d}{\delta_{D}} \cdot \dfrac{\delta_{D}-d}{2} + \dfrac{\delta_{D}}{\delta_{A}} \cdot \dfrac{d}{\delta_{D}} \cdot 0 ]
@@
 To derive the benefit, the cost of moving is subtracted from the average gain. 
-\begin{equation}\label{first}
+\begin{equation}
 \beta_{A}(\alpha_{D},\alpha_{A}) = \dfrac { (\delta_{D}-d) ^{2}} {2 \cdot \delta_{D}  \delta_{A}} - k_{A} \alpha_{A}
 \end{equation}
-\begin{equation}\label{first}
+\begin{equation}
 \beta_{A}(\alpha_{D},\alpha_{A}) = \dfrac { \delta_{D}} {2 \cdot \delta_{A}} - k_{A} \alpha_{A} - ( \dfrac{d^{2}}{2 \cdot \delta_{A} \delta_{D}} - \dfrac{d}{\delta_{A}} )
@@
  The benefit of the defender is expressed as follows:
- \begin{equation}\label{first}
+\begin{equation}
 \beta_{D}(\alpha_{D},\alpha_{A}) = 1 - \dfrac { (\delta_{D}-d) ^{2}} {2 \cdot \delta_{D}  \delta_{A}} - k_{D} \alpha_{D}
@@
 
-\begin{equation}\label{first}
+\begin{equation}
 \beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac {1} {2r} - k_{D} \alpha_{D} = \dfrac {\delta_{A}} {2\delta_{D}} - k_{D} \alpha_{D} 
 \end{equation}
-\begin{equation}\label{first}
+\begin{equation}
 \beta_{A}(\alpha_{D},\alpha_{A}) =1 - \dfrac {1} {2r} - k_{A} \alpha_{A} = 1- \dfrac {\delta_{A}} {2\delta_{D}} - k_{A} \alpha_{A}  
 \end{equation}\\
+
+An intuitive solution for the case with a virus would be to subtract the benefit of the attacker received in each interval with the delay similarly as in case 1.
+This would give us the following formula if we derive it in the same way:
+\begin{equation}
+\beta_{A}=\dfrac{(\delta_{A} - d)^2}{2\delta_{A}\delta_{D}} - \dfrac{k_{A}}{\delta_{A}}
+\end{equation}
+
+But this results in an overestimation.
+How closer $\delta_{A}/\delta_{D}$ is equal to one, the better the approximation.
+If $\delta_{A}/\delta_{D} = 1$ the result is correct.
+This formula does not take into account that if the attacker was in control in the previous interval, the delay should not be subtracted, because the delay will not be in control of the defender.
+This means that we have to look at what happens in the previous interval.
+\\
+
+From the defender we know that his moves are instantaneous.
+It is easier to calculate the benefit of the defender in this case.
+Because the defender moves slower than the attacker we know that if the defender moves during the interval of the attacker, he only moves once within this interval.
+The defender will move during the interval of the attacker with a probability of $\dfrac{\delta_{A}}{\delta_{D}} $.
+When this happens the defender will end with being in control at the end of the interval.
+In the next interval the attacker will have to regain control, meaning that during the delay, the defender stays in control, see figure \ref{fig:case2} cases (1) and (2).
+The defender will keep the control over the resource in the next interval over a period of the delay, namely \textit{d}.
+\\
+
+Consider a timespan $\delta_{A} + d$, representing the attacker's interval followed by the delay period in his next interval.
+The defender will never move twice during this timespan because $\delta_{A} + d \leq \delta_{D}$.
+Because $d + \delta_{A} \leq \delta_{D}$ the next move of the defender in this second interval will never occur during the delay, meaning that the entire delay can be considered as an extra benefit resulting of a play in the previous interval.
+So, every time the defender plays, he will get an average gain of $\dfrac{\delta_{A}}{2}$ in the interval where he plays and in the next interval will always receive a extra gain of $d$, yielding a total average gain per interval of
+$\dfrac{(d+\dfrac{\delta_{A}}{2})}{\delta_{A}}$
 
@@
 \subsubsection*{\textbf{Case a:} $d + \delta_{A} \leq \delta_{D}$}
-~~\\
-Consider a timespan $\delta_{A} + d$, representing the attacker's interval followed by the delay period in his next interval.
-The defender will never move twice during this timespan because $\delta_{A} + d \leq \delta_{D}$.
-The defender will move during the interval of the attacker with a probability of $\dfrac{\delta_{A}}{\delta_{D}} $.
-When this happens the defender will end with being in control at the end of the interval.
-In the next interval the attacker will have to regain control, meaning that during the delay, the defender stays in control, see figure \ref{fig:case2} cases (1) and (2).
-This means that the defender will keep the control over the resource in the next interval over a period of the delay, namely \textit{d}.
-Because $d + \delta_{A} \leq \delta_{D}$ the next move of the defender in this second interval will never occur during the delay, meaning that the entire delay can be considered as an extra benefit resulting of a play in the previous interval.
-So, every time the defender plays, he will get an average gain of $\dfrac{\delta_{A}}{2}$ in the interval where he plays and in the next interval will always receive a extra gain of $d$, yielding a total average gain per interval of
-$\dfrac{(d+\dfrac{\delta_{A}}{2})}{\delta_{A}}$
 

### 2015-06-08 doc/template/optimum.tex
aanvullen nash schrijven

\chapter{FlipIt game with virus propagation}
\label{chapter2:FlipIt with virus propagation}
%\documentclass[10pt]{article}
%\begin{document}

The optimum functions will be piecewise functions. There will be different optimum functions depending on $k_{D}$ and $d$. Two cases are considered.
\subsubsection{$k_{D} \leq d$}

\begin{description}
\item if $\delta_{A} < 2(k_{D} - d)$ \\
For case 2.A it is decreasing. Because $k_{D} \leq d$ the term $2(k_{D} - d)$ will always be negative so $\delta_{A}$ will also be negative. This means for case 1 and case 2.b that $\delta_{D} = 0$ and so the defender will not play.
\item $\delta_{A} = 2(k_{D} - d)$ \\
$\delta_{A} = 0$ so the attacker will not play. For case 1 and case 2.b the defender will also not play.
\item $\delta_{A} > 2(k_{D} - d)$ \\
For case 2.b it is decreasing. Case 2.a it is increasing. For case 1 $2(k_{D} - d)$ is always negative so the minimum is 0. This means that $\delta_{D} \in [0, \sqrt{d^{2} + 2\delta_{A}k_{D}}]$
\end{description}

\subsubsection{$k_{D} > d$}
\begin{description}
\item if $\delta_{A} < 2(k_{D} - d)$ \\
For case 2.A it is increasing. 
\item $\delta_{A} = 2(k_{D} - d)$ \\
\item $\delta_{A} > 2(k_{D} - d)$ \\
In every case the function is decreasing so the defender will not play.
\end{description}

From this analyses we can compute $opt_{D}(\delta_{A})$ for two different cases as:
 \begin{displaymath}
  opt_{D}(\delta_{A}) = \left\{
     \begin{array}{lr}
       0, & \delta_{A} < 2(k_{D} - d)\\
       0, & \delta_{A} = 2(k_{D} - d) \\
       \big[0,\sqrt{d^{2} + 2\delta_{A}k_{D}}\big], & \delta_{A} > 2(k_{D} - d)
     \end{array}
   \right.
\end{displaymath}

For case :
 \begin{displaymath}
  opt_{D}(\delta_{A}) = \left\{
     \begin{array}{lr}
       0, & \delta_{A} < 2(k_{D} - d)\\
       0, & \delta_{A} = 2(k_{D} - d) \\
       \big[0,\sqrt{d^{2} + 2\delta_{A}k_{D}}\big], & \delta_{A} > 2(k_{D} - d)
     \end{array}
   \right.
\end{displaymath}
-------------------------------------------\\
We still consider the case where $d < \delta_{D}$. \\
To determine the Nash equilibria we also need to determine $opt_{1}(\delta_{D})$ by computing the derivative of $\beta_{A}(\delta_{D},\delta_{A})$ for a fixed $\delta_{D}$. We consider 2 cases: \\

\textbf{Case 1: $\delta_{A} \geq \delta_{D}$}\\

Since 
\begin{equation*}
\beta_{A}(\delta_{D},\delta_{A}) =\dfrac{\delta_{D}}{2\delta_{A}} - \dfrac{k_{A}}{\delta_{A}} + \dfrac{d^{2}}{2\delta_{D}\delta_{A}^{2}} - \dfrac{d}{\delta_{A}}
\end{equation*}
the derivative is:
\begin{equation*}
\dfrac{\partial \beta_{A}(\delta_{D},\delta_{A})}{\partial \delta_{A}} = -\dfrac{\delta_{D}}{2\delta_{A}^{2}} + \dfrac{k_{A}}{\delta_{A}^{2}} - \dfrac{d^{2}}{2\delta_{D}\delta_{A}^{2}} + \dfrac{d}{\delta_{A}^{2}}
\end{equation*}
it follows that $\beta_{A}(\delta_{D},\cdot)$ is increasing if $2k_{A} < (\delta_{D} - d)^{2} / \delta_{D}$ and decreasing if $2k_{A} > (\delta_{D} - d)^{2} / \delta_{D}$. \\

\textbf{Case 2: $\delta_{A} \leq \delta_{D}$ } \\

\textbf{A}: $\delta_{A} \leq d + \delta_{A} \leq \delta_{D}$ \\
Since 
\begin{equation*}
\beta_{A}(\delta_{D},\delta_{A}) =1- \dfrac{\delta_{A}}{2\delta_{D}} - \dfrac{k_{A}}{\delta_{A}} - \dfrac{d}{\delta_{D}}
\end{equation*}
the derivative is:
\begin{equation*}
\dfrac{\partial \beta_{A}(\delta_{D},\delta_{A})}{\partial \delta_{A}} = \dfrac{-1}{2\delta_{D}} + \dfrac{k_{A}}{\delta_{A}^{2}}
\end{equation*}
it follows that $\beta_{A}(\delta_{D},\cdot)$ is increasing on $[0,\sqrt{2k_{A}\delta_{D}}]$ and decreasing on $[\sqrt{2k_{A}\delta_{D}}, \infty]$ and thus has a maximum on $\delta_{A} = maximum \{\delta_{D}, \sqrt{2k_{A}\delta_{D}} \} $. The maximum between $\delta_{D}$ and $ \sqrt{2k_{A}\delta_{D}}$ is needed because $\delta_{A} $ cannot exceed $\delta_{D}$ in this case. \\

\textbf{B}: $\delta_{A}  \leq \delta_{D} \leq d + \delta_{A} $ \\

Since 
\begin{equation*}
\beta_{A}(\delta_{D},\delta_{A}) = 1 - \dfrac{\delta_{A}}{2\delta_{D}} - \dfrac{d}{\delta_{A}} - \dfrac{k_{A}}{\delta_{A}} + \dfrac{(d-(\delta_{D}-\delta_{A})^{2}}{2\delta_{D}\delta_{A}} 
\end{equation*}
the derivative is:
\begin{equation*}
\dfrac{\partial \beta_{A}(\delta_{D},\delta_{A})}{\partial \delta_{A}} = -\dfrac{\delta_{D}}{2\delta_{A}^{2}} + \dfrac{k_{A}}{\delta_{A}^{2}} - \dfrac{d^{2}}{2\delta_{D}\delta_{A}^{2}} + \dfrac{d}{\delta_{A}^{2}}
\end{equation*}
it follows that $\beta_{A}(\delta_{D},\cdot)$ is increasing if $2k_{A} < (\delta_{D} - d)^{2} / \delta_{D}$ and decreasing if $2k_{A} > (\delta_{D} - d)^{2} / \delta_{D}$. This is the same result as in case 1.\\

\section{Extensions on FlipIt}
\label{ch1:extendedWork}

Various possible ways to extend FlipIt have already been proposed. 
Laszka et al. made a lot of additions and extensions to the original game of FlipIt. For instance Laszka et al. extended the basic FlipIt game to multiple resources. The rationale is that for compromising a system in real life, more than just one resource needs to be taken over. An example is that gaining access to deeper layers of a system may require breaking several passwords. The model is called FlipThem \cite{FlipThem}. Laszka et al. also use two ways to flip the multiple resources: the AND and the OR control model. In the AND model the attacker only controls the system if he controls all the resources of the system, whereas in the OR model the attacker only needs to compromise one resource to be in control of the entire system. \\

Another addition of Laszka et al. to the game of FlipIt \cite{MitigationCovert} 
is extending the game to also consider non-targeted attacks by non-strategic players. In this game the defender tries to maintain control over the resource that is subjected to both targeted and non-targeted attacks. Non-targeted attacks can include phishing, while targeted attacks may include threats delivered through zero day attack vulnerabilities. \\
One of the last important additions from Laszka et al. \cite{MitigationNonTargeted} is to consider a game with targeted and non-targeted attacks where the moves made by the attacker do not succeed immediately. This is similar to this paper but it has still some major differences. First the moves by the attacker are still covert but the moves made by the defender are known to the attacker. This means that the attacker knows when the defender plays and can change its strategy depending on the moves of the defender. Our motivation for a defender with stealthy moves is that there is not always an intelligent individual that is behind an APT. Some APTs don't know if the computer is already been recovered. There purpose is to spread. Not to check if they have already infected. \todo{beter verwoorden}. The second difference is that even though both the targeted and non-targeted attacks do not succeed immediately, the delay is determined differently. For the targeted attack the time till it succeeds is given by an exponential distributed random variable with a known rate. The non-targeted attacks are modelled as a single attacker and the time till it succeeds is given by a Poisson process. In our paper the delay is given by one parameter, that can be the result of any virus propagation model. The third and last difference is that the paper of Laska has multiple attackers and they try to find the best strategy of the defender against both targeted and non-targeted attacks. The conclusion of this paper is that the optimal strategy for the defender is moving periodically. \\

FlipIt has also been applied to several cases in system security. Reseachers explored different applications of FlipIt for real-world problems, like password reset policies, VM refresh, cloud auditing and key rotation \cite{ApplyingFlipit}. \\
Other authors used the FlipIt game to apply it on a specific scenario. To be able to use the FlipIt game, modifications where required for the FlipIt model.
One of the scenarios by Pham \cite{compromised} was to find out whether a resource was compromised or not by the attacker. This could be verified by the defender, who has an extra move "test" beside the flip move. The basic idea is to test with an extra action if the resource has been compromised or not. This move involves also an extra cost.\\
A three-player game has also been investigated where the flipit framework of two players is extended by another player. This player represents an insider that trades value information with the attacker \cite{fengstealthy}.\\

Finally researchers also have investigated the behaviour of humans playing FlipIt. A. Nochenson and Grossklags \cite{nochenson2013behavioral}  investigate how people really act when given temporal decisions. They found out that the results improves over time but that they are dependent on gender, age, and other individual difference variables. The result also shows that the participants perform generally better when they have more information about the strategy of the opponent which is a computerized player. Reitter et al. \cite{reitter2013risk} extended the work of A. Nochenson and Grossklags to include various visual presentation modalities for the available feedback during the investigation.\\

### 2015-06-08 doc/template/chap-5.tex

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+This chapter provides the reader with an introduction to the general context of the work presented in this paper.
+Section 2.1 introduces the reader into the basic concepts of cyber security and the kind of cyber security threaths that are in the scope for this work.
+Section 2.2 then introduces the reader into the main principles of game theory.
+Subsequently, section 2.3 introduces the reader to the FlipIt game, the specific game that will be used to model cyber security attacks of periodic nature and including a delay.
+Finally, section 2.4 gives an overview of the related work, and how the research presented in this thesis is positioned compared to existing results.
+
 \section{What is cyber security?}
 
-Cyber security is a part of security that focuses on securing computer controlled devices such as computers and smartphones, as well as public and private computer networks, including the entire Internet.
-A private network can be an enterprise network that is protected from the public network such as the internet.
-The purpose of security is to give certainty that data will not be removed without authorization (Confidentiality) that the data is always accessible (Availability), and that data can not be read or altered by someone who does not have the authorization (Integrity).
-These are the 3 key attributes of security, also known as the CIA triad.
+Before the digitalization of documents, information was kept on paper and the security of this information was ensured by administrative and physical means.
+For example, you needed a key to access documents stored in a room full of cabinets where the files were kept.
+In today's digital era more and more information is kept in a digital format, stored on a computer.
+As digitalization progressed, the need for ensuring the security of digital information arose and automated tools where developed for protecting files stored on a computer.
+The generic name to protect data stored on a computer controlled device such as computers and smartphones, as well as public and private computer networks, including the entire Internet is called computer security.
+\\
+
+Security is a general term that encompasses several dimensions. More specifically, computer security has tree key objectives that are fundamental to computer security:
+\begin{description}
+\item Confidentiality: This assures that the confidential of private data is not disclosed or made available to users that do not have the authorization. 
+\item Integrity: This assures that data can not be altered by an unauthorized individual.
+\item Availability: This assures that data is always accessible and that the service is not denied to authorized individuals.
+\end{description}
+
+ the purpose of security is to give certainty that data will not be removed without authorization (Confidentiality) that the data is always accessible (Availability), and that data can not be read or altered by someone who does not have the authorization (Integrity).
+
+ These are the 3 key attributes of security, also known as the CIA triad. The tree concepts are fundamental security objectives for the securing of data, information and computing services. \\
+
+''Cyber security is the process of applying security tools to ensure confidentiality, integrity, and availability of data.
+Cybersecurity attempts to ensure the protection of assets, which includes data, desktops, servers, buildings, and most importantly, humans.
+The goal of cyber security is to protect data both in transit and at rest.
+Countermeasures can be put in place in order to increase the security of data.
+Some of these measures include, but are not limited to, access control, awareness training, audit and accountability, risk assessment, penetration testing, vulnerability management, and security assessment and authorization.'' \todo{tekst overgenomen van wikipedia} \\
+
+In the context of cyber security, the terms ' threat and attack' are often used interchangeably, referring to more or less the same thing.
+They however have a slightly different meaning.
+A threat refers to anything that can breach the security and that can cause a possible harm.
+It is a possible danger that can exploit a vulnerability.
+An attack is an assault on computer security that comes from a threat.
+It is an intelligent act that tries deliberately to breach security through vulnerabilities.
 \\
@@
 In order to keep a system secure, it is important to mitigate the possible threats.
-A threat refers to anything that can breach the security and that can cause a possible harm.
-The some of the most common threats are spam, malware, spoofing, phishing and  DDoS attacks.
-The biggest threat are caused by malware. 
+The possible threats can take many forms, the most commen being spam, malware, spoofing, phishing and  DDoS attacks.
+A security report of 2014 reveals that 61\% of the attacks on companies are caused by malware.
 This is a piece of malicious program that is designed to penetrate unprotected systems or computers, and getting there to sensitive information.
-Malware can be divided into different subclasses.
-These subclasses are not mutually exclusive:
+Malware can be divided into different subclasses that are not mutually exclusive:
 
@@
 From the standpoint of the defender, it is very important to respond as quickly as possible so that the worm can not spread.
-\item Rootkit:
-\item Trojan:
-\item Backdoors:
+\item Rootkit: A rootkit is designed to be stealthy and hide a set of programs installed on a system from methods of detection.
+The programs have administrator or root privileges to that system, which makes it possible to access the functions and services of the operating system, change files, take control over the monitor processes and send and receive network traffic.
+A rootkit can make changes to the system to keep itself concealed from detection.
+\item Trojan: This is a malicious program that disguises itself as something normal, so that users won't be suspicious of installing it.  
+\item Backdoors: Also know as trapdoor, is a whole in the system or program that allows access 
 \end{description}
 
-A security report of 2014 reveales that 61\% of the attacks on companies are caused by malware. A company can take different measures to defend himself against malware. 
-The threats caused by malware can be devided into 3 big groups: known threats (70\%), unknown threats (29\%) and advanced threats (1\%) [\todo{toevoegen referentie naar kaspersky APT report}].
+A company can take different measures to defend itself against malware.
+The threats caused by malware can be divided into three main categories: known threats (70\%), unknown threats (29\%) and advanced threats (1\%) as proposed by Kasperksy [\todo{toevoegen referentie naar kaspersky APT report}].
 
-The known threats easy to defend against. Unkown also. Then the remaining 1\% are the advanced threats. These threats are also known as APT. APT's are ... 
-In this paper we focus on targeted attacks.
-Targeted attacks are malware designed to attack exclusively an organisation.
-Some examples of the biggest most rare APTs  to get an understanding of what APTs are capable of and how long they can stay unseen.
-[\todo{site kaspersky apt}]
+The known threats are the easiest to  defend oneself against.
+Standard malware protection tools like firewalls and virus scanners can keep these kind of malware out of the system.
+Installing protection against unkown is also relatively easy.
+Then the remaining 1\% are the advanced threats, also known as APT.
+APT's are ...
+In this paper we focus on targeted attacks. Targeted attacks are malware designed to attack exclusively an organisation.
 
-An APT is a persistent targeted attack that tries to penetrate a network, harm and stays unseen for a long period of time.
-The motive of an APT is mostly cyber espionage, stealing sensitive data, sabotage or other ideological attacks.
-APT stands for Advanced persistant threat.
-Advanced stands for the fact that these attacks are well funded and that the attacker itself has a great expertise.
-These APT are not easy to protect against.
-Not all APTs are technical advanced.
-The attacker can also try to exploit existing vulnerabilities and hope that his target has not yet secured himself against these vulnerabilities.
-Persistent is the fact that the attacker keeps on trying to break in to his victim.
+An APT is a persistent targeted attack that tries to penetrate a network to harm it while staying unseen for a long period of time.
+The motive of an APT is mostly cyber espionage, stealing sensitive data, sabotage or some other kind of ideological attacks.
+Advanced Persistent Threat are called  'Advanced' for the fact that these attacks are well funded and that (usually) the attacker itself needs a great expertise to successfully penetrate a network.
+Not all APTs are technical advanced though.
+The attacker can also try to exploit existing vulnerabilities simply based in the hope that his target has not yet secured himself against these vulnerabilities.
+'Persistent' refers to the fact that the attacker keeps on trying to break in to his victim.
 He will not give up.
 The attack can be over various years and different steps.
-The threats stands for the fact that an APT is ..
+The threats stands for the fact that an APT is attempting to break the protection of valuable data the leakage of which outside the secure environment may severely damage the organisation owning the data.
+
+Some examples of the biggest most rare APTs are listed to get an understanding of what APTs are capable of and how long they can stay unseen. [\todo{site kaspersky apt}]
 

### 2015-06-08 doc/template/chap-9.tex

@@
 \end{equation}\\
-
-\section{Nash Equilibria}
--- rechtstreeks uit FlipIt paper --\\
-As a second step, we are interested in finding Nash equilibria, points
-for which neither player will increase his benefit by changing his rate of play. More
-formally, a Nash equilibrium for the periodic game is a point $(\alpha_{0}^{*},\alpha_{1}^{*})$ such that
-the defender's benefit $\beta_{0}(\alpha_{0},\alpha_{1}^{*}) $is maximized at $\alpha_{0}= \alpha_{0}^{*}$ and the attacker's benefit
-$\beta_{1}(\alpha_{0}^{*},\alpha_{1}) $ is maximized at $\alpha_{1}= \alpha_{1}^{*}$ .
-To begin with, some useful notation. We denote by opt0($\alpha_{1}$) the set of values (rates
-of play $\alpha_{0}$) that optimize the benefit of the defender for a fixed rate of play $\alpha_{1}$ of the
-attacker. Similarly, we denote by opt1($\alpha_{0}$) the set of values (rates of play $\alpha_{1}$) that optimize
-the benefit of the attacker for a fixed rate of play $\alpha_{0}$ of the defender. The following
-theorem specifies Nash equilibria for the periodic game and is proven in Appendix A. \\
-
-Nash equilibria are points whith the property that neither player benefits by deviating in isolaition form equilibrium.
-We can compute Nash Equilibria for the periodic game as an intersection points of curvest opt0 and opt1.
-To determine opt0(a0) we need to compute the derivate of  $\beta_{0}(\alpha_{0},\alpha_{1}) $ for a fixed $\alpha_{1}$.
-We consider two cases:\\
------
-
-\subsection{case: $\delta_{D} \leq \delta_{A} $ or $\alpha_{D} \geq \alpha_{A}$}
-
-Since 
-\begin{equation}\label{formdelta}
-\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} = \frac{\partial [1 - \dfrac { (\delta_{D}-d) ^{2}} {2 \cdot \delta_{D}  \delta_{A}]} - k_{D} \alpha_{D}}{\partial \alpha_{D}}
-\end{equation}
-
-If we change every $\delta = \dfrac{1}{\alpha}$ in formule \ref{formdelta} we get the next formula that is easier to take the derivate of:
-\begin{equation}
-\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} =\frac{\partial [1 - \dfrac { (1/\alpha_{D}-d) ^{2} \cdot \alpha_{D} \alpha_{A}} {2} - k_{D} \alpha_{D}]}{\partial \alpha_{D}}
-\end{equation}
-Which gives the following result:
-\begin{equation}
-\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} = \alpha_{A} ( \dfrac{1}{\alpha_{D}^{2}} - \dfrac{d^{2}}{2}) - k_{D}
-\end{equation}
-
-it follows that if equal to zero $\alpha_{0}=\sqrt{\dfrac{\alpha_{A}}{2k_{D} + d^{2}\alpha_{A}}}$
-
-\subsection{case 2: $\delta_{D} \geq \delta_{A} $ or $\alpha_{D} \leq \alpha_{A}$}
-
-\textbf{case A}\\
-
-\begin{equation}
-\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} = \frac{\partial [ \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - k_{D} \alpha_{D}] }{\partial \alpha_{D}}
-\end{equation}
-\begin{equation}
-\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} = \frac{\partial [ \dfrac{\alpha_{D}}{2\alpha_{A}} + d\alpha_{D} - k_{D} \alpha_{D}] }{\partial \alpha_{D}}
-\end{equation}
-\begin{equation}
-\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} = \dfrac{1}{2\delta_{A}} + d - k_{D}
-\end{equation}
-
-Dus increasing als $\alpha_{A} < \dfrac{1}{2(k_{D} -d)}$ and decreasing if $\alpha_{A} > \dfrac{1}{2(k_{D} -d)}$
-
-\textbf{Case b} \\
-
-\begin{equation}
-\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} = \frac{\partial [ \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - k_{D} \alpha_{D} - \dfrac{(d - (\delta_{D} - \delta_{A}))^{2}}{2 \delta_{D} \delta_{A}}] }{\partial \alpha_{D}}
-\end{equation}
-
-Uitgewerkt geeft de benefit formule de volgende formule omgezet naar alpha
-\begin{equation}
-\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} = \frac{\partial [ 1 - \dfrac{\alpha_{A}}{2\alpha_{D}} - k_{D}\alpha_{D} + \dfrac{d \alpha_{A}}{2} + \dfrac{d \alpha_{D}}{2} - \dfrac{d^{2} \alpha_{A} \alpha_{D}}{2} ] }{\partial \alpha_{D}}
-\end{equation}
-
-\begin{equation}
-\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} = \dfrac{\alpha_{A}}{2\alpha_{D}^{2}} - k_{D} + \dfrac{d}{2} - \dfrac{d^{2} \alpha_{A}}{2}
-\end{equation}
-
-Als ge dat gelijk stelt aan 0 krijgt ge:
-\begin{equation}
-\alpha_{D} = \sqrt{\dfrac{\alpha_{A}}{2k_{D} -d + d^{2}\alpha_{A}}}
-\end{equation}
-%%% Local Variables: 
-%%% mode: latex
-%%% TeX-master: "thesis"
-%%% End:
-
-%\end{document}

### 2015-06-08 doc/template/nash.tex

\chapter{Nash Equilibria}
\label{chapter:Nash}
%\documentclass[10pt]{article}
%\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%			Introduction Chapter 1				%%%%%%
%%%%%												%%%%%%
%%%%%												%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Nash Equilibria}
-- rechtstreeks uit FlipIt paper --\\
As a second step, we are interested in finding Nash equilibria, points
for which neither player will increase his benefit by changing his rate of play. More
formally, a Nash equilibrium for the periodic game is a point $(\alpha_{0}^{*},\alpha_{1}^{*})$ such that
the defender's benefit $\beta_{0}(\alpha_{0},\alpha_{1}^{*}) $is maximized at $\alpha_{0}= \alpha_{0}^{*}$ and the attacker's benefit
$\beta_{1}(\alpha_{0}^{*},\alpha_{1}) $ is maximized at $\alpha_{1}= \alpha_{1}^{*}$ .
To begin with, some useful notation. We denote by opt0($\alpha_{1}$) the set of values (rates
of play $\alpha_{0}$) that optimize the benefit of the defender for a fixed rate of play $\alpha_{1}$ of the
attacker. Similarly, we denote by opt1($\alpha_{0}$) the set of values (rates of play $\alpha_{1}$) that optimize
the benefit of the attacker for a fixed rate of play $\alpha_{0}$ of the defender. The following
theorem specifies Nash equilibria for the periodic game and is proven in Appendix A. \\

Nash equilibria are points whith the property that neither player benefits by deviating in isolaition form equilibrium. We can compute Nash Equilibria for the periodic game as an intersection points of curvest opt0 and opt1. To determine opt0(a0) we need to compute the derivate of  $\beta_{0}(\alpha_{0},\alpha_{1}) $ for a fixed $\alpha_{1}$. We consider two cases:\\
-----

\textbf{Case: $\delta_{D} \leq \delta_{A} $\\

\begin{equation}
\beta_{D}(\alpha_{D},\alpha_{A}) = 1 - \dfrac{\delta_{D}}{2\delta_{A}} - \dfrac{k_{D}}{\delta_{D}} - \dfrac{d^{2}}{2\delta_{D}\delta_{A}} + \dfrac{d}{\delta_{A}}
\end{equation}
If we take the partial derivative we get the following result:
\begin{equation}\label{formdelta}
\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} = - \dfrac{1}{2\delta_{A}} + \dfrac{k_{D}}{\delta_{D}^{2}} + \dfrac{d^{2}}{2\delta_{D}^{2}\delta_{A}}
\end{equation}

Als we de formule \ref{formdelta} gelijkstellen aan 0 krijgen we:

\begin{equation}
\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} =0 ~~~~~~ =>~~~~~~ \delta_{D} = \sqrt{2\delta_{A}k_{D} + d^{2}}
\end{equation}

The function increases on $[0, \sqrt{2\delta_{A}k_{D} + d^{2}}]$ and is decreasing on $[\sqrt{2\delta_{A}k_{D} + d^{2}}, \infty]$ So we got a maximum on $\delta_{D} = minimum \{ \delta_{A}, \sqrt{2\delta_{A}k_{D} + d^{2}} \} $ \\
~~\\

\textbf{Case 2.A: $\delta_{D} \geq d+\delta_{A} \geq \delta_{A} $ }\\

\begin{equation}
 \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - \dfrac{k_{D}}{\delta_{D}}
\end{equation}
\begin{equation}
\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} = -\dfrac{\delta_{A}}{2\delta_{D}^{2}} - \dfrac{d}{\delta_{D}^{2}} + \dfrac{k_{D}}{\delta_{D}^{2}}
\end{equation}

Als we de formule \ref{formdelta} gelijkstellen aan 0 krijgen we:
\begin{equation}
\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} =0 ~~~~~~ =>~~~~~~ \delta_{A} = 2(k_{D}-d) = dk_{D} - 2d
\end{equation}
\begin{description}
\item If $k_{D} < d$ 
\begin{description}
\item decreasing $ \delta_{A} < 2(k_{D} -d)$
\item increasing  $\delta_{A} > 2(k_{D} -d)$ \todo{equal}
\end{description}
\item If $k_{D} > d$ 
\begin{description}
\item increasing $ \delta_{A} < 2(k_{D} -d)$
\item decreasing  $\delta_{A} > 2(k_{D} -d)$ \todo{equal}
\end{description}
\end{description}
~~\\

\textbf{Case 2.B: $d+\delta_{A} \geq \delta_{D} \geq  \delta_{A} $} \\

\begin{equation}
\beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - \dfrac{k_{D}}{\delta_{D}} - \dfrac{(d-(\delta_{D} - \delta_{A}))^{2}}{d\delta_{D}\delta_{A}}
\end{equation}

\begin{equation}
\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} =  - \dfrac{1}{2\delta_{D}} + \dfrac{k_{D}}{\delta_{D}^{2}} + \dfrac{d^{2}}{2\delta_{D}^{2}\delta_{A}}
\end{equation}

Als we de formule \ref{formdelta} gelijkstellen aan 0 krijgen we:

\begin{equation}
\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} =0 ~~~~~~ =>~~~~~~ \delta_{D} = \sqrt{2\delta_{A}k_{D} + d^{2}}
\end{equation}

The function increases on $[0, \sqrt{2\delta_{A}k_{D} + d^{2}}]$ and is decreasing on $[\sqrt{2\delta_{A}k_{D} + d^{2}}, \infty]$ So we got a maximum on $\delta_{D} = minimum \{ \delta_{A}, \sqrt{2\delta_{A}k_{D} + d^{2}} \} $ \\

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "thesis"
%%% End:

%\end{document}

### 2015-06-08 doc/template/chap-5.tex
beetje overal verbeteren

@@
 One of the last important additions from Laszka et al.
-\cite{MitigationNonTargeted} is to consider a game where the moves made by the attacker are still covert but the moves made by the defender are known to the attacker.
-This means that the attacker can base his attacks on the defender's moves.
-Both the targeted and non-targeted attacks do not succeed immediately.
+\cite{MitigationNonTargeted} is to consider a game with targeted and non-targeted attacks where the moves made by the attacker do not succeed immediately.
+This is similar to this paper but it has still some major differences.
+First the moves by the attacker are still covert but the moves made by the defender are known to the attacker.
+This means that the attacker knows when the defender plays and can change its strategy depending on the moves of the defender.
+Our motivation for a defender with stealthy moves is that there is not always an intelligent individual that is behind an APT.
+Some APTs don't know if the computer is already been recovered.
+There purpose is to spread.
+Not to check if they have already infected.
+\todo{beter verwoorden}.
+The second difference is that even though both the targeted and non-targeted attacks do not succeed immediately, the delay is determined differently.
 For the targeted attack the time till it succeeds is given by an exponential distributed random variable with a known rate.
 The non-targeted attacks are modelled as a single attacker and the time till it succeeds is given by a Poisson process.
+In our paper the delay is given by one parameter, that can be the result of any virus propagation model.
+The third and last difference is that the paper of Laska has multiple attackers and they try to find the best strategy of the defender against both targeted and non-targeted attacks.
 The conclusion of this paper is that the optimal strategy for the defender is moving periodically.
-The difference with this paper is that the delay in this paper is dependent on the number of nodes that have to be flipped in a network.
-This cannot be modelled with the framework in the Laszka et al.
-paper because the delay is chosen as an exponential distributed random variable.
-Another difference is that in the case of the Laszka et al.
-paper, the moves of the defender are considered not stealthy and so the attacker knows when the defender plays.
 \\
 
+FlipIt has also been applied to several cases in system security.
+Reseachers explored different applications of FlipIt for real-world problems, like password reset policies, VM refresh, cloud auditing and key rotation \cite{ApplyingFlipit}.
+\\
 Other authors used the FlipIt game to apply it on a specific scenario. To be able to use the FlipIt game, modifications where required for the FlipIt model.
@@
 This move involves also an extra cost.\\
+A three-player game has also been investigated where the flipit framework of two players is extended by another player.
+This player represents an insider that trades value information with the attacker \cite{fengstealthy}.\\
 
@@
 Nochenson and Grossklags \cite{nochenson2013behavioral}  investigate how people really act when given temporal decisions.
+They found out that the results improves over time but that they are dependent on gender, age, and other individual difference variables.
+The result also shows that the participants perform generally better when they have more information about the strategy of the opponent which is a computerized player.
 Reitter et al.
 \cite{reitter2013risk} extended the work of A.
-Nochenson and Grossklags to include various visual presentation modalities for the available feedback during the investigation.
+Nochenson and Grossklags to include various visual presentation modalities for the available feedback during the investigation.\\
 

### 2015-06-08 doc/template/chap-9.tex
beetje overal verbeteren

@@
 \label{Ch2:Intro}
+The game of FlipIt with virus propagation considers attackers where their moves are not instantaneous.
+This can be motivated by an example of a virus.
+A virus can be dropped on a network but it only compromises the whole network if every node in the network is infected.
+The basic FlipIt game does not take this into account.
+In this chapter the FlipIt game with virus propagation is explained and how it can be modelled.
+First the main differences with the basic FlipIt game are discussed in section \ref{ch2:diffFlip}.
+Next in section \ref{ch2:formal} the formal definition of the game with virus propagation is given.
+At last, in section \ref{ch2:periodicvirus} a periodic strategy game is considered and the formalization of the formulas with virus propagation are determined.\\
 
+This chapter explains how to model a FlipIt game with a virus propagation that infects a network.
+The first section explains the difference between a normal FlipIt game and a FlipIt game with virus propagation.
+The next section derives a formula to calculate the benefit for a FlipIt game with a virus propagation.
+In the last section we calculate the Nash equilibrium for the benefit formula.
 \section{FlipIt game with virus propagation}
 
-Motivatie voor het veranderen van FLipIt naar een FlipIt met viruspropagatie:
+Motivation
 
 \section{Explaining difference between FlipIt with and without virus propagation}
+\label{ch2:diffFlip}
 
@@
 will do this by flipping all the nodes of the network (i.e. the entire resource) in every move he plays. The
-attacker, the other player, will try to infect all the nodes in the network. The attacker
+attacker on the other hand will try to infect all the nodes in the network. The attacker
 will do this by flipping the node in the graph that can infect all the nodes in the
@@
 
-\section{Formal definition Game}
-
-In this section we provide a formal definition of the game and the notation that we will use throughout the paper.
-
-\begin{description}
-\item $i$: Defines the player. Different as in the FlipIt paper where the defender is denoted by the subscript 0 and the attacker by the subscript 1.
-\item $\delta_{i}$: The length of the interval between two consecutive moves of player \textit{i}. 
-\item $\alpha_{i}$: The average flip rate of player \textit{i}, given by $\alpha_{i}=1/\alpha_{i}$.
-\item $k_{i}$: The cost of player \textit{i}'s moves.
-\item $d$: The delay caused by the virus propagation.
-\item $G_{i}(t)$: The total gain of player \textit{i} denotes the amount of time player \textit{i} is in control over the resource up to time \textit{t}.
-\item $\gamma_{i}$: The average gain rate of player \textit{i} defined as $G_{i}(t)/t$
-\item $\beta_{i}$:  The average benefit rate up to time \textit{t} defined as  $\beta_{i} = \gamma_{i} -k_{i} \alpha_{i} $.
-\item $opt_{i}$: The optimum function.
-\end{description}
-
-\todo{tekening toevoegen waarop de delay staat aangeduid}
-
-\section*{Playing periodically with virus propagation}
-
-This chapter explains how to model a FlipIt game with a virus propagation that infects a network.
-The first section explains the difference between a normal FlipIt game and a FlipIt game with virus propagation.
-The next section derives a formula to calculate the benefit for a FlipIt game with a virus propagation.
-In the last section we calculate the Nash equilibrium for the benefit formula.
+\section{Playing periodically with virus propagation}
+\label{ch2:periodicvirus}
 
@@
 \subsection{Formalization the benefit formula including the infection-delay}
+
  A Periodic strategy is a non-adaptive renewal strategy where the time intervals between consecutive moves are a fixed period, denoted by $\delta$.
@@
 The average rate of play of a player is denoted by $\alpha_{i} = \dfrac{1}{\delta_{i}}$.
-\\
+List of symbols that will be used throughout the paper.
+Figure \ref{FlipItDelay} represents a couple of the symbols for clarification.\\
 ~~\\
+\begin{figure}[hbtp]
+\caption{Formalization of a FlipIt game with delay: A representation of a FlipIt game where both players are playing periodically.
+Every move or flip is indicated by a blue or orange circle.
+The defender is represented in blue and plays with a period of $\delta_{D}$.
+The flip of the attacker is represented by a white circle, but because there is a delay d, the attacker only controls the resource after time d represented by an orange circle.
+The attacker plays with a period of $\delta_{A}$.
+The blue and orange rectangles represent the amount of time the respective player is in control of the resource.}
+\centering
+\includegraphics[scale=0.7]{Images/DefFlip.png}
+\label{FlipItDelay}
+\end{figure}
+
+\begin{description}
+\item $i$: Defines the player. Different as in the FlipIt paper where the defender is denoted by the subscript 0 and the attacker by the subscript 1.
+\item $\delta_{i}$: The length of the interval between two consecutive moves of player \textit{i}. 
+\item $\alpha_{i}$: The average flip rate of player \textit{i}, given by $\alpha_{i}=1/\delta_{i}$.
+\item $k_{i}$: The cost of player \textit{i}'s moves.
+\item $d$: The delay caused by the virus propagation.
+\item $G_{i}(t)$: The total gain of player \textit{i} denotes the amount of time player \textit{i} is in control over the resource up to time \textit{t}.
+\item $\gamma_{i}$: The average gain rate of player \textit{i} defined as $G_{i}(t)/t$
+\item $\beta_{i}$:  The average benefit rate up to time \textit{t} defined as  $\beta_{i} = \gamma_{i} -k_{i} \alpha_{i} $.
+\item $opt_{i}$: The optimum function.
+\end{description}
 
@@
 
-\begin{equation}
+\begin{equation*}
 \beta_{A}(\alpha_{D},\alpha_{A}) =\dfrac {r} {2} - k_{A} \alpha_{A} = \dfrac {\delta_{D}} {2\delta_{A}} - k_{A} \alpha_{A}  
-\end{equation}\\
+\end{equation*}\\
 
 Correspondingly, the benefit for the defender can be expressed as:
-\begin{equation}
+\begin{equation*}
 \beta_{D}(\alpha_{D},\alpha_{A}) =1 -  \dfrac {r} {2} - k_{D} \alpha_{D} = 1 - \dfrac {\delta_{D}} {2\delta_{A}} - k_{D} \alpha_{D} 
-\end{equation}
+\end{equation*}
 
 \begin{figure}[hbtp]
-\caption{The first FlipIt game is one without virus propagation. The second one is with virus propagation and \textit{d} = 1. The delay is denoted with an arrow.}
+\caption{Case 1: Difference between a basic FlipIt game and a FlipIt game with a delay.
+(1) is the FlipIt game without a delay and (2) is with a delay.
+The delay is denoted with an arrow.
+The attacker is only in control when the circle becomes orange.}
 \centering
-\includegraphics[scale=0.4]{../../doc/template/Images/FLipItCase1.pdf}
+\includegraphics[scale=0.6]{../../doc/template/Images/DiffDelayCase1.png}
 \label{fig:delaycase1}
@@
 The average gain rate of the attacker can then be expressed as follows if we look at one interval of the defender:
-\begin{equation}
+\begin{equation*}
 \gamma_{A}(\alpha_{D},\alpha_{A}) = \dfrac {1}{\delta_{D}} [ \dfrac{\delta_{D}}{\delta_{A}} \cdot \dfrac{\delta_{D}-d}{\delta_{D}} \cdot \dfrac{\delta_{D}-d}{2} + \dfrac{\delta_{D}}{\delta_{A}} \cdot \dfrac{d}{\delta_{D}} \cdot 0 ]
-\end{equation}
+\end{equation*}
 
 To derive the benefit, the cost of moving is subtracted from the average gain. 
-\begin{equation}
+\begin{equation*}
 \beta_{A}(\alpha_{D},\alpha_{A}) = \dfrac { (\delta_{D}-d) ^{2}} {2 \cdot \delta_{D}  \delta_{A}} - k_{A} \alpha_{A}
-\end{equation}
-\begin{equation}
+\end{equation*}
+\begin{equation*}
 \beta_{A}(\alpha_{D},\alpha_{A}) = \dfrac { \delta_{D}} {2 \cdot \delta_{A}} - k_{A} \alpha_{A} - ( \dfrac{d^{2}}{2 \cdot \delta_{A} \delta_{D}} - \dfrac{d}{\delta_{A}} )
-\end{equation}
+\end{equation*}
 
  The benefit of the defender is expressed as follows:
-\begin{equation}
+\begin{equation*}
 \beta_{D}(\alpha_{D},\alpha_{A}) = 1 - \dfrac { (\delta_{D}-d) ^{2}} {2 \cdot \delta_{D}  \delta_{A}} - k_{D} \alpha_{D}
-\end{equation}
+\end{equation*}
 ~~\\
@@
 
-\begin{equation}
+\begin{equation*}
 \beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac {1} {2r} - k_{D} \alpha_{D} = \dfrac {\delta_{A}} {2\delta_{D}} - k_{D} \alpha_{D} 
-\end{equation}
-\begin{equation}
+\end{equation*}
+\begin{equation*}
 \beta_{A}(\alpha_{D},\alpha_{A}) =1 - \dfrac {1} {2r} - k_{A} \alpha_{A} = 1- \dfrac {\delta_{A}} {2\delta_{D}} - k_{A} \alpha_{A}  
-\end{equation}\\
+\end{equation*}\\
 
@@
 This would give us the following formula if we derive it in the same way:
-\begin{equation}
+\begin{equation*}
 \beta_{A}=\dfrac{(\delta_{A} - d)^2}{2\delta_{A}\delta_{D}} - \dfrac{k_{A}}{\delta_{A}}
-\end{equation}
+\end{equation*}
 
@@
 
-\begin{equation}\label{first}
+\begin{equation*}\label{first}
 \gamma_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta_{A}}{\delta_{D}} \cdot \dfrac{(d+\dfrac{\delta_{A}}{2})}{\delta_{A}} 
-\end{equation}
-\begin{equation}\label{first}
+\end{equation*}
+\begin{equation*}\label{first}
 \gamma_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} 
-\end{equation}\\
+\end{equation*}\\
 This yields in the following benefit formula:
-\begin{equation}\label{first}
+\begin{equation*}\label{first}
 \beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - k_{D} \alpha_{D} 
-\end{equation}\\
+\end{equation*}\\
 
 The benefit for the attacker will be as follows:
-\begin{equation}\label{first}
+\begin{equation*}\label{first}
 \beta_{A}(\alpha_{D},\alpha_{A}) = 1 -\dfrac{\delta_{A}}{2\delta_{D}} - \dfrac{d}{\delta_{D}} - k_{A} \alpha_{A} 
-\end{equation}\\
+\end{equation*}\\
 
@@
 
-\begin{equation}\label{first}
+\begin{equation*}
 \dfrac{1} {\delta_{A}} \cdot \dfrac{d - (\delta_{D} - \delta_{A})}{\delta_{D}} \cdot \dfrac{d - (\delta_{D} - \delta_{A})}{\delta_{D}}
-\end{equation}
+\end{equation*}
 
 The total gain  rate of the defender is obtained by subtracting this term from the gain rate of case a:
- \begin{equation}\label{first}
+ \begin{equation*}
 \gamma_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta_{A}}{\delta_{D}} \cdot \dfrac{(d+\dfrac{\delta_{A}}{2})}{\delta_{A}} - \dfrac{(d - (\delta_{D} - \delta_{A}))^{2}}{2 \delta_{D} \delta_{A}}
-\end{equation}
-\begin{equation}\label{first}
+\end{equation*}
+\begin{equation*}
 \gamma_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - \dfrac{(d - (\delta_{D} - \delta_{A}))^{2}}{2 \delta_{D} \delta_{A}}
-\end{equation}\\
+\end{equation*}\\
 This yields in the following benefit formula:
-\begin{equation}\label{first}
+\begin{equation*}
 \beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - k_{D} \alpha_{D} - \dfrac{(d - (\delta_{D} - \delta_{A}))^{2}}{2 \delta_{D} \delta_{A}}
-\end{equation}\\
+\end{equation*}\\
 
 The benefit for the attacker will be as follows:
-\begin{equation}\label{first}
+\begin{equation*}
 \beta_{A}(\alpha_{D},\alpha_{A}) = 1 -\dfrac{\delta_{A}}{2\delta_{D}} - \dfrac{d}{\delta_{D}} - k_{A} \alpha_{A} + \dfrac{(d - (\delta_{D} - \delta_{A}))^{2}}{2 \delta_{D} \delta_{A}}
-\end{equation}\\
+\end{equation*}\\

### 2015-06-08 doc/template/nash.tex
beetje overal verbeteren

@@
 
-\textbf{Case: $\delta_{D} \leq \delta_{A} $\\
+\textbf{Case: $\delta_{D} \leq \delta_{A} $}\\
 
@@
 \begin{equation}
- \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - \dfrac{k_{D}}{\delta_{D}}
+\dfrac{t \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - \dfrac{k_{D}}{\delta_{D}}
 \end{equation}
@@
 \begin{equation}
-\beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - \dfrac{k_{D}}{\delta_{D}} - \dfrac{(d-(\delta_{D} - \delta_{A}))^{2}}{d\delta_{D}\delta_{A}}
+\dfrac{\beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - \dfrac{k_{D}}{\delta_{D}} - \dfrac{(d-(\delta_{D} - \delta_{A}))^{2}}{d\delta_{D}\delta_{A}}
 \end{equation}
@@
 
+The optimum functions will be piecewise functions. There will be different optimum functions depending on $k_{D}$ and $d$. Two cases are considered.
+\subsubsection{$k_{D} \leq d$}
+
+\begin{description}
+\item if $\delta_{A} < 2(k_{D} - d)$ \\
+For case 2.A it is decreasing.
+Because $k_{D} \leq d$ the term $2(k_{D} - d)$ will always be negative so $\delta_{A}$ will also be negative.
+This means for case 1 and case 2.b that $\delta_{D} = 0$ and so the defender will not play.
+\item $\delta_{A} = 2(k_{D} - d)$ \\
+$\delta_{A} = 0$ so the attacker will not play. For case 1 and case 2.b the defender will also not play.
+\item $\delta_{A} > 2(k_{D} - d)$ \\
+For case 2.b it is decreasing. Case 2.a it is increasing. For case 1 $2(k_{D} - d)$ is always negative so the minimum is 0. This means that $\delta_{D} \in [0, \sqrt{d^{2} + 2\delta_{A}k_{D}}]$
+\end{description}
+
+\subsection{Best responses}
+\subsubsection{$k_{D} > d$}
+\begin{description}
+\item if $\delta_{A} < 2(k_{D} - d)$ \\
+For case 2.A it is increasing. 
+\item $\delta_{A} = 2(k_{D} - d)$ \\
+\item $\delta_{A} > 2(k_{D} - d)$ \\
+In every case the function is decreasing so the defender will not play.
+\end{description}
+
+From this analyses we can compute $opt_{D}(\delta_{A})$ for two different cases as:
+ \begin{displaymath}
+  opt_{D}(\delta_{A}) = \left\{
+     \begin{array}{lr}
+       0, & \delta_{A} < 2(k_{D} - d)\\
+       0, & \delta_{A} = 2(k_{D} - d) \\
+       \big[0,\sqrt{d^{2} + 2\delta_{A}k_{D}}\big], & \delta_{A} > 2(k_{D} - d)
+     \end{array}
+   \right.
+\end{displaymath}
+
+For case :
+ \begin{displaymath}
+  opt_{D}(\delta_{A}) = \left\{
+     \begin{array}{lr}
+       0, & \delta_{A} < 2(k_{D} - d)\\
+       0, & \delta_{A} = 2(k_{D} - d) \\
+       \big[0,\sqrt{d^{2} + 2\delta_{A}k_{D}}\big], & \delta_{A} > 2(k_{D} - d)
+     \end{array}
+   \right.
+\end{displaymath}
+-------------------------------------------\\
+
+\subsection{betaA}
+We still consider the case where $d < \delta_{D}$. \\
+To determine the Nash equilibria we also need to determine $opt_{1}(\delta_{D})$ by computing the derivative of $\beta_{A}(\delta_{D},\delta_{A})$ for a fixed $\delta_{D}$. We consider 2 cases: \\
+
+\textbf{Case 1: $\delta_{A} \geq \delta_{D}$}\\
+
+Since 
+\begin{equation*}
+\beta_{A}(\delta_{D},\delta_{A}) =\dfrac{\delta_{D}}{2\delta_{A}} - \dfrac{k_{A}}{\delta_{A}} + \dfrac{d^{2}}{2\delta_{D}\delta_{A}^{2}} - \dfrac{d}{\delta_{A}}
+\end{equation*}
+the derivative is:
+\begin{equation*}
+\dfrac{\partial \beta_{A}(\delta_{D},\delta_{A})}{\partial \delta_{A}} = -\dfrac{\delta_{D}}{2\delta_{A}^{2}} + \dfrac{k_{A}}{\delta_{A}^{2}} - \dfrac{d^{2}}{2\delta_{D}\delta_{A}^{2}} + \dfrac{d}{\delta_{A}^{2}}
+\end{equation*}
+it follows that $\beta_{A}(\delta_{D},\cdot)$ is increasing if $2k_{A} < (\delta_{D} - d)^{2} / \delta_{D}$ and decreasing if $2k_{A} > (\delta_{D} - d)^{2} / \delta_{D}$. \\
+
+\textbf{Case 2: $\delta_{A} \leq \delta_{D}$ } \\
+
+\textbf{A}: $\delta_{A} \leq d + \delta_{A} \leq \delta_{D}$ \\
+Since 
+\begin{equation*}
+\beta_{A}(\delta_{D},\delta_{A}) =1- \dfrac{\delta_{A}}{2\delta_{D}} - \dfrac{k_{A}}{\delta_{A}} - \dfrac{d}{\delta_{D}}
+\end{equation*}
+the derivative is:
+\begin{equation*}
+\dfrac{\partial \beta_{A}(\delta_{D},\delta_{A})}{\partial \delta_{A}} = \dfrac{-1}{2\delta_{D}} + \dfrac{k_{A}}{\delta_{A}^{2}}
+\end{equation*}
+it follows that $\beta_{A}(\delta_{D},\cdot)$ is increasing on $[0,\sqrt{2k_{A}\delta_{D}}]$ and decreasing on $[\sqrt{2k_{A}\delta_{D}}, \infty]$ and thus has a maximum on $\delta_{A} = maximum \{\delta_{D}, \sqrt{2k_{A}\delta_{D}} \} $.
+The maximum between $\delta_{D}$ and $ \sqrt{2k_{A}\delta_{D}}$ is needed because $\delta_{A} $ cannot exceed $\delta_{D}$ in this case.
+\\
+
+\textbf{B}: $\delta_{A}  \leq \delta_{D} \leq d + \delta_{A} $ \\
+
+Since 
+\begin{equation*}
+\beta_{A}(\delta_{D},\delta_{A}) = 1 - \dfrac{\delta_{A}}{2\delta_{D}} - \dfrac{d}{\delta_{A}} - \dfrac{k_{A}}{\delta_{A}} + \dfrac{(d-(\delta_{D}-\delta_{A})^{2}}{2\delta_{D}\delta_{A}} 
+\end{equation*}
+the derivative is:
+\begin{equation*}
+\dfrac{\partial \beta_{A}(\delta_{D},\delta_{A})}{\partial \delta_{A}} = -\dfrac{\delta_{D}}{2\delta_{A}^{2}} + \dfrac{k_{A}}{\delta_{A}^{2}} - \dfrac{d^{2}}{2\delta_{D}\delta_{A}^{2}} + \dfrac{d}{\delta_{A}^{2}}
+\end{equation*}
+it follows that $\beta_{A}(\delta_{D},\cdot)$ is increasing if $2k_{A} < (\delta_{D} - d)^{2} / \delta_{D}$ and decreasing if $2k_{A} > (\delta_{D} - d)^{2} / \delta_{D}$.
+This is the same result as in case 1.\\
 %%% Local Variables: 

### 2015-06-08 doc/template/optimum.tex
beetje overal verbeteren

@@
 -------------------------------------------\\
-We still consider the case where $d < \delta_{D}$. \\
-To determine the Nash equilibria we also need to determine $opt_{1}(\delta_{D})$ by computing the derivative of $\beta_{A}(\delta_{D},\delta_{A})$ for a fixed $\delta_{D}$. We consider 2 cases: \\
-
-\textbf{Case 1: $\delta_{A} \geq \delta_{D}$}\\
-
-Since 
-\begin{equation*}
-\beta_{A}(\delta_{D},\delta_{A}) =\dfrac{\delta_{D}}{2\delta_{A}} - \dfrac{k_{A}}{\delta_{A}} + \dfrac{d^{2}}{2\delta_{D}\delta_{A}^{2}} - \dfrac{d}{\delta_{A}}
-\end{equation*}
-the derivative is:
-\begin{equation*}
-\dfrac{\partial \beta_{A}(\delta_{D},\delta_{A})}{\partial \delta_{A}} = -\dfrac{\delta_{D}}{2\delta_{A}^{2}} + \dfrac{k_{A}}{\delta_{A}^{2}} - \dfrac{d^{2}}{2\delta_{D}\delta_{A}^{2}} + \dfrac{d}{\delta_{A}^{2}}
-\end{equation*}
-it follows that $\beta_{A}(\delta_{D},\cdot)$ is increasing if $2k_{A} < (\delta_{D} - d)^{2} / \delta_{D}$ and decreasing if $2k_{A} > (\delta_{D} - d)^{2} / \delta_{D}$. \\
-
-\textbf{Case 2: $\delta_{A} \leq \delta_{D}$ } \\
-
-\textbf{A}: $\delta_{A} \leq d + \delta_{A} \leq \delta_{D}$ \\
-Since 
-\begin{equation*}
-\beta_{A}(\delta_{D},\delta_{A}) =1- \dfrac{\delta_{A}}{2\delta_{D}} - \dfrac{k_{A}}{\delta_{A}} - \dfrac{d}{\delta_{D}}
-\end{equation*}
-the derivative is:
-\begin{equation*}
-\dfrac{\partial \beta_{A}(\delta_{D},\delta_{A})}{\partial \delta_{A}} = \dfrac{-1}{2\delta_{D}} + \dfrac{k_{A}}{\delta_{A}^{2}}
-\end{equation*}
-it follows that $\beta_{A}(\delta_{D},\cdot)$ is increasing on $[0,\sqrt{2k_{A}\delta_{D}}]$ and decreasing on $[\sqrt{2k_{A}\delta_{D}}, \infty]$ and thus has a maximum on $\delta_{A} = maximum \{\delta_{D}, \sqrt{2k_{A}\delta_{D}} \} $.
-The maximum between $\delta_{D}$ and $ \sqrt{2k_{A}\delta_{D}}$ is needed because $\delta_{A} $ cannot exceed $\delta_{D}$ in this case.
-\\
-
-\textbf{B}: $\delta_{A}  \leq \delta_{D} \leq d + \delta_{A} $ \\
-
-Since 
-\begin{equation*}
-\beta_{A}(\delta_{D},\delta_{A}) = 1 - \dfrac{\delta_{A}}{2\delta_{D}} - \dfrac{d}{\delta_{A}} - \dfrac{k_{A}}{\delta_{A}} + \dfrac{(d-(\delta_{D}-\delta_{A})^{2}}{2\delta_{D}\delta_{A}} 
-\end{equation*}
-the derivative is:
-\begin{equation*}
-\dfrac{\partial \beta_{A}(\delta_{D},\delta_{A})}{\partial \delta_{A}} = -\dfrac{\delta_{D}}{2\delta_{A}^{2}} + \dfrac{k_{A}}{\delta_{A}^{2}} - \dfrac{d^{2}}{2\delta_{D}\delta_{A}^{2}} + \dfrac{d}{\delta_{A}^{2}}
-\end{equation*}
-it follows that $\beta_{A}(\delta_{D},\cdot)$ is increasing if $2k_{A} < (\delta_{D} - d)^{2} / \delta_{D}$ and decreasing if $2k_{A} > (\delta_{D} - d)^{2} / \delta_{D}$.
-This is the same result as in case 1.\\
-
-\section{Extensions on FlipIt}
-\label{ch1:extendedWork}
-
-Various possible ways to extend FlipIt have already been proposed. 
-Laszka et al.
-made a lot of additions and extensions to the original game of FlipIt.
-For instance Laszka et al.
-extended the basic FlipIt game to multiple resources.
-The rationale is that for compromising a system in real life, more than just one resource needs to be taken over.
-An example is that gaining access to deeper layers of a system may require breaking several passwords.
-The model is called FlipThem \cite{FlipThem}.
-Laszka et al.
-also use two ways to flip the multiple resources: the AND and the OR control model.
-In the AND model the attacker only controls the system if he controls all the resources of the system, whereas in the OR model the attacker only needs to compromise one resource to be in control of the entire system.
-\\
-
-Another addition of Laszka et al. to the game of FlipIt \cite{MitigationCovert} 
-is extending the game to also consider non-targeted attacks by non-strategic players.
-In this game the defender tries to maintain control over the resource that is subjected to both targeted and non-targeted attacks.
-Non-targeted attacks can include phishing, while targeted attacks may include threats delivered through zero day attack vulnerabilities.
-\\
-One of the last important additions from Laszka et al.
-\cite{MitigationNonTargeted} is to consider a game with targeted and non-targeted attacks where the moves made by the attacker do not succeed immediately.
-This is similar to this paper but it has still some major differences.
-First the moves by the attacker are still covert but the moves made by the defender are known to the attacker.
-This means that the attacker knows when the defender plays and can change its strategy depending on the moves of the defender.
-Our motivation for a defender with stealthy moves is that there is not always an intelligent individual that is behind an APT.
-Some APTs don't know if the computer is already been recovered.
-There purpose is to spread.
-Not to check if they have already infected.
-\todo{beter verwoorden}.
-The second difference is that even though both the targeted and non-targeted attacks do not succeed immediately, the delay is determined differently.
-For the targeted attack the time till it succeeds is given by an exponential distributed random variable with a known rate.
-The non-targeted attacks are modelled as a single attacker and the time till it succeeds is given by a Poisson process.
-In our paper the delay is given by one parameter, that can be the result of any virus propagation model.
-The third and last difference is that the paper of Laska has multiple attackers and they try to find the best strategy of the defender against both targeted and non-targeted attacks.
-The conclusion of this paper is that the optimal strategy for the defender is moving periodically.
-\\
-
-FlipIt has also been applied to several cases in system security.
-Reseachers explored different applications of FlipIt for real-world problems, like password reset policies, VM refresh, cloud auditing and key rotation \cite{ApplyingFlipit}.
-\\
-Other authors used the FlipIt game to apply it on a specific scenario. To be able to use the FlipIt game, modifications where required for the FlipIt model.
-One of the scenarios by Pham \cite{compromised} was to find out whether a resource was compromised or not by the attacker.
-This could be verified by the defender, who has an extra move "test" beside the flip move.
-The basic idea is to test with an extra action if the resource has been compromised or not.
-This move involves also an extra cost.\\
-A three-player game has also been investigated where the flipit framework of two players is extended by another player.
-This player represents an insider that trades value information with the attacker \cite{fengstealthy}.\\
-
-Finally researchers also have investigated the behaviour of humans playing FlipIt.
-A.
-Nochenson and Grossklags \cite{nochenson2013behavioral}  investigate how people really act when given temporal decisions.
-They found out that the results improves over time but that they are dependent on gender, age, and other individual difference variables.
-The result also shows that the participants perform generally better when they have more information about the strategy of the opponent which is a computerized player.
-Reitter et al.
-\cite{reitter2013risk} extended the work of A.
-Nochenson and Grossklags to include various visual presentation modalities for the available feedback during the investigation.\\

### 2015-06-10 doc/template/chap-10.tex
Tekst introduction bijgewerkt

@@
 A particular kind of threats are Advanced Persistent Threats (APT).
-\todo{misschien meer zeggen over APTs en threats} An APT is a targeted cyber attack that is designed to penetrate a network or a system in a stealthy way and that can stay undetected for a long period.
+An APT is a multi-faceted, continuous and targeted cyber attack that is designed to penetrate a network or a system in a stealthy way and that can stay undetected for a long period.
+It is something different and stronger than a conventional threat.
+A conventional threat will not attack any particular target.
+An APT is persistent and will keep on trying to attack its victim.
+It operates silently and stealthily, to prevent detection.
 This makes it so hard to protect a network or a system against an APT.
-Bruce Schneier describes an APT as something different and stronger than a conventional threat: ''\textit{A conventional hacker or criminal is not interested in any particular target.
+
+%Bruce Schneier describes an APT as something different and stronger than a conventional threat: ''\textit{A conventional hacker or criminal is not interested in any particular target.
 He wants a thousand credit card numbers for fraud, or to break into an account and turn it into a zombie, or whatever.
@@
 Since it is so difficult to protect a system or a network against APT's, researchers have been looking for effective ways to predict in advance which defence strategy might be the better one. 
-Game theory is gaining increasing interest as an effective technique to model and study Cyber Security.
-Game theory analyses the security problem as a game where the players are an attacker and a defender of a system, and where both players have to make decisions.
-In particular, both players will aim for the strategy that results in a maximal benefit for them.
+Game theory is gaining increasing interest as an effective technique to model and study cyber security problems.
+It is common to model cyber security games as a game with two players, an attacker and a defender.
+However, there are games that have more players \cite{fengstealthy}.
+In this paper we focus on the game with two players.
+The actions available to the attacker and the defender correspond to attacks on the system and defensive measures that protect the system.
+In particular, both players will execute the strategy that results in a maximal benefit for them.
 Researchers at RSA, van Dijk et al,  made a game theoretic model of targeted attacks.
@@
 This is a two players game where the attacker and the defender are competing to get control over a shared resource.
-Both players do not know who is currently in control of the resource until they move.
+Neither player knows who is currently in control of the resource until they move.
 In FlipIt every move gives them immediate control over the resource.
@@
 
-The game theoretical approach of the FlipIt does not take such delay into account. 
-This an lead us to the following research questions:
+This paper adapts the model presented in \cite{FlipIt} so as to take the delay for virus propagation into account.
+This leads us to the following research questions:
 \begin{itemize}
@@
 \subsubsection{Overview of the thesis}
---opbouw van de thesis uitleggen-- \\
 
 The organisation of this paper is the following.
-In chapter \ref{Chapter1:Intro.Game.Theory}  a brief introduction to Gametheory is introduced to get familiar with the game theoretic concepts that will be further used in the paper.
+In chapter \ref{Chapter1:Intro.Game.Theory}  a brief introduction to cyber security and gametheory is introduced to get familiar with the kind of threats that are in the scope of this work and the game theoretic concepts that will be further used in the paper.
 In the same chapter the FlipIt framework is summarized and the most important conclusions together with the the related work done on FlipIt and the difference with this paper.
@@
 After that formulas are derived to model a FlipIt game with a virus propagation for a specific case where players play a periodic strategy with a random phase.
-This chapter ends with simulations where conclusions can be derived.
+\\
+In Chapter \ref{chapter:Nash} the formules are used to determine if there is a nash equilibrium. \todo{aanpassen naar chapter met nash equilibria}
 Next in Chapter \ref{chapter4: Virus propagation} we given an overview of the various ways in which a virus can propagate.

### 2015-06-10 doc/template/chap-5.tex
Tekst introduction bijgewerkt

@@
 The generic name to protect data stored on a computer controlled device such as computers and smartphones, as well as public and private computer networks, including the entire Internet is called computer security.
-\\
+Security is a general term that encompasses several dimensions. More specifically, computer security has tree key objectives that are fundamental to computer security:
 
-Security is a general term that encompasses several dimensions. More specifically, computer security has tree key objectives that are fundamental to computer security:
-\begin{description}
-\item Confidentiality: This assures that the confidential of private data is not disclosed or made available to users that do not have the authorization. 
-\item Integrity: This assures that data can not be altered by an unauthorized individual.
-\item Availability: This assures that data is always accessible and that the service is not denied to authorized individuals.
-\end{description}
+~~\\
+ \textit{Confidentiality}: This assures that the confidential of private data is not disclosed or made available to users that do not have the authorization. \\
+ \textit{Integrity:} This assures that data can not be altered by an unauthorized individual.\\
+ \textit{Availability:} This assures that data is always accessible and that the service is not denied to authorized individuals.\\
 
- the purpose of security is to give certainty that data will not be removed without authorization (Confidentiality) that the data is always accessible (Availability), and that data can not be read or altered by someone who does not have the authorization (Integrity).
+ %The purpose of security is to give certainty that data will not be removed without authorization (Confidentiality) that the data is always accessible (Availability), and that data can not be read or altered by someone who does not have the authorization (Integrity).
+ These are the 3 key attributes of security, also known as the CIA triad. The tree concepts are fundamental security objectives for the securing of data, information and computing services. \\
+Cyber security is the process of applying security tools to ensure confidentiality, integrity, and availability of data.
+It is an attempt to protect websites, servers, data, applications, operating systems and all assets that need protection in a computer system.
+Some of these tools may include detection, identification or removal tools.
+A detection tool will determine if an infection has taken place and will trace the threat.
+An identification tool will try to identify the threat to be able to know how to remove it.
+The removal tool will once the threat has been identified, remove the threat from the system so that it cannot spread any further.
 
- These are the 3 key attributes of security, also known as the CIA triad. The tree concepts are fundamental security objectives for the securing of data, information and computing services. \\
-
-''Cyber security is the process of applying security tools to ensure confidentiality, integrity, and availability of data.
-Cybersecurity attempts to ensure the protection of assets, which includes data, desktops, servers, buildings, and most importantly, humans.
+%Cybersecurity attempts to ensure the protection of assets, which includes data, desktops, servers, buildings, and most importantly, humans.
 The goal of cyber security is to protect data both in transit and at rest.
@@
 
+\subsubsection{Threats to computer systems}
+
+In order to keep a system secure and to be able to apply security tools, it is important to mitigate the possible threats.
+The possible threats can take many forms, the most common being spam, malware, spoofing, phishing and  DDoS attacks.
 In the context of cyber security, the terms ' threat and attack' are often used interchangeably, referring to more or less the same thing.
@@
 
-In order to keep a system secure, it is important to mitigate the possible threats.
-The possible threats can take many forms, the most commen being spam, malware, spoofing, phishing and  DDoS attacks.
-A security report of 2014 reveals that 61\% of the attacks on companies are caused by malware.
-This is a piece of malicious program that is designed to penetrate unprotected systems or computers, and getting there to sensitive information.
-Malware can be divided into different subclasses that are not mutually exclusive:
+The most noteworthy group and biggest group of threats to computer systems is malware. 
+This is a piece of malicious software that is designed to penetrate unprotected systems or computers, with the intent to get sensitive information, destroy data, or compromise the confidentiality, integrity or availability of the data or applications of the victim.
+A security report of 2014 \cite{SurveyKaspersky} reveals that 61\% of the attacks on companies are caused by malware.
+For this reason this section will examine the categories of malware threats.
+Different types of malware exist such as virusses, worms, flooders, rootkits, bots, spyware, adware and many more.
+This broad range of different types can be classified into two main categories, the first one based on the propagation method that is used and the second one for the payload or the variety of actions that the malware performs.
+Propagation methods include virusses, worms and trojans.
+Payload includes flooders, rootkits, bots, spyware and adware.
+In this paper we focus on the category of propagation methods and not on the actions that the malware performs.
+We give a brief explanation of the three main propagation methods of malware.
+\todo{opsplitsing uit het boek van network security genomen}
 
 \begin{description}
-\item Virus: This is a malicious piece of code that replicates and tries to spread to infect other systems or files.
-The "I love you" virus is an example of a virus that quickly spreads.
-This virus propagates via mail systems.
-If someone opens an email with "I love you" with the virus in the annex this virus spreads itself by sending a mail to everyone in your contact list.
+\item \textit{Virus}: This is a malicious piece of code that replicates itself and tries to spread in order  to infect other systems or files.
+A typical virus will attach itself to a program, or an executable content on a computer.
+The ''I love you'' virus is an example of a virus where the virus attached itself as an executable to a mail.
+To propagate it used the mail systems.
+If someone opens an email with the "I love you"  in the annex, the virus will spread itself by sending a mail to everyone in the victim's contact list.
 So the virus can multiply rapidly and eventually a business network might shut down by the heavy traffic.
 In this example, there is a need for human interaction to spread the virus.
-If no one opens the mail the virus can not infect other systems.
-Unfortunately, there are also viruses that can spread without human interaction.
-These viruses are referred to as worms.
-A worm is also a computer program that replicates itself to spread to other computers.
-Via a computer network, copies of the worm can be forwarded without an intermediary.
+If no one opens the mail the virus can not spread itself and infect other systems.
+
+\item \textit{Worm:} A worm is a virus that can spread without human interaction.
+A worm is a computer program that replicates itself in order to spread to other hosts on a network.
+Copies of the worm can be forwarded via a computer network without an intermediary.
 The worm will use vulnerabilities of the system to infect other computers.
-The Stuxnetworm is a very famous worm.
-Initially this worm spread via infected USB sticks and from then it could spread through the Internet to other computers.
-The purpose of the Stuxnetworm was to harm the centrifuges in nuclear reactors.
-Many reactors have been infected.
-From the standpoint of the defender, it is very important to respond as quickly as possible so that the worm can not spread.
-\item Rootkit: A rootkit is designed to be stealthy and hide a set of programs installed on a system from methods of detection.
-The programs have administrator or root privileges to that system, which makes it possible to access the functions and services of the operating system, change files, take control over the monitor processes and send and receive network traffic.
-A rootkit can make changes to the system to keep itself concealed from detection.
-\item Trojan: This is a malicious program that disguises itself as something normal, so that users won't be suspicious of installing it.  
-\item Backdoors: Also know as trapdoor, is a whole in the system or program that allows access 
+The Stuxnetworm is a very prominent example of a worm.
+Initially this worm was spread via infected USB sticks and from there it could spread itself without any further human intervention through the Internet to other hosts on the network.
+The purpose of the Stuxnetworm was to harm the centrifuges in nuclear reactors and many reactors have been infected.
+
+\item \textit{Trojan:} This is a malicious program that disguises itself as something normal and useful, so that users won't be suspicious of installing it, but it has a malicious function hidden inside that can avoid security measures and cause harm.
+A notable trojan horse is Koobface, that targeted users of facebook, skype, yahoo, gmail and AOL mail.
+To spread itself the worm sended a mail or friend request with a message that directed the recipients to another third party website.
+This site tried to put the reciepient into downloading an update of Adobe Flash Player.
+Once downloaded and executed, Koobface could infect the host.
+
 \end{description}
@@
 A company can take different measures to defend itself against malware.
-The threats caused by malware can be divided into three main categories: known threats (70\%), unknown threats (29\%) and advanced threats (1\%) as proposed by Kasperksy [\todo{toevoegen referentie naar kaspersky APT report}].
+The threats caused by malware can be divided into three main categories: known threats (70\%), unknown threats (29\%) and advanced threats (1\%) as proposed by Kasperksy \cite{APTKaspersky}.
 
@@
 Standard malware protection tools like firewalls and virus scanners can keep these kind of malware out of the system.
-Installing protection against unkown is also relatively easy.
-Then the remaining 1\% are the advanced threats, also known as APT.
-APT's are ...
-In this paper we focus on targeted attacks. Targeted attacks are malware designed to attack exclusively an organisation.
+Installing protection against unkown is also relatively easy, but this will need tools that go beyond the standard methods like dynamic whitelisting.
+The remaining 1\% are the advanced threats, also known as APT.
 
-An APT is a persistent targeted attack that tries to penetrate a network to harm it while staying unseen for a long period of time.
+\subsubsection{Advanced threats to computer systems}
+%\item Backdoors: Also know as trapdoor, is a whole in the system or program that allows access 
+%\item Rootkit: A rootkit is designed to be stealthy and hide a set of programs installed on a system from methods of detection.
+The programs have administrator or root privileges to that system, which makes it possible to access the functions and services of the operating system, change files, take control over the monitor processes and send and receive network traffic.
+A rootkit can make changes to the system to keep itself concealed from detection.
+
+ An APT is a persistent targeted attack that tries to penetrate a network to cause harm while staying unseen for a long period of time.
 The motive of an APT is mostly cyber espionage, stealing sensitive data, sabotage or some other kind of ideological attacks.
@@
 The attacker can also try to exploit existing vulnerabilities simply based in the hope that his target has not yet secured himself against these vulnerabilities.
-'Persistent' refers to the fact that the attacker keeps on trying to break in to his victim.
-He will not give up.
-The attack can be over various years and different steps.
-The threats stands for the fact that an APT is attempting to break the protection of valuable data the leakage of which outside the secure environment may severely damage the organisation owning the data.
-
-Some examples of the biggest most rare APTs are listed to get an understanding of what APTs are capable of and how long they can stay unseen. [\todo{site kaspersky apt}]
-
-\subsubsection{Equation}
-Equation is a complex cyberattack platform where the first known sample is from 2002, but it was only discovered 12 years later in 2014.
+'Persistent' refers to the fact that the attacker keeps on trying to attack his target.
+The attack can be over various years and different steps can be taken.
+The 'threats' stands for the fact that an intelligent act tries deliberately to breach security through vulnerabilities which may severely damage the target organisation.
+An APT can be a mix of different types of malware and will use different kinds of propagation methods. \\
+%
+%Some examples of the biggest most rare APTs are listed to get an understanding of what APTs are capable of and how long they can stay unseen. [\todo{site kaspersky apt}] 
+%
+%\subsubsection{Equation}
+%Equation is a complex cyberattack platform where the first known sample is from 2002, but it was only discovered 12 years later in 2014.
 This APT propagates through usb drives, cd or physical media.
@@
 The purpose of this virus is to steal data and cyberespionage.
+%
+%\subsubsection{Regin}
+%
+%\subsubsection{Flame}
+%Way of propagation through USB drives, LAN spreading. purpose cyber espionage.
+%
+%\subsubsection{Black energy}
+% purpose cyber espionage and DDoS, data wiping. prop usb lan
 
-\subsubsection{Regin}
+According to a security survey of Kaspersky \cite{SurveyKaspersky}  the damage of one successful target attack against a large company can exceed over 2.54 million dollar.
+A company needs a defence mechanism to defend itself against APTs.
+As previously said, detection and identification tools will not work against APTs.
+The removal tool will only work if the threat has been identified.
+To mitigate these kind of attacks there is a need for another security countermeasures.
+Here is where the researchers of RSA fall in.
+They have investigated through game theory how to model targeted attacks and what the best defend strategy is according to the information that is available.
 
-\subsubsection{Flame}
-Way of propagation through USB drives, LAN spreading. purpose cyber espionage.
-
-\subsubsection{Black energy}
- purpose cyber espionage and DDoS, data wiping. prop usb lan
-
-According to [\todo{APT report} ]  the damage of one successful target attack can exceed over 2.54 million dollar.
-a company needs a defense mechanism to defend itself against apt.
-Prevention more important than ..
-een van de manieren om dat te doen is via game theorie.
-
-Security belangrijk. verschillende malware. uitleggen malware. verwijzen naar kaspersky security raport. APT grote threats. voorbeelden van APT's. waarom gametheory gebruiken.
-
-\section{Intro game theory}
+\section{A brief introduction in Game Theory}
+\label{Cha1:briefintro}
 Gametheory is a mathematical study to analyse interactions between independent and self-interested agents.
 To get an understanding of the most important concepts of game theory, a short introduction based on the work of
-\cite{leyton2008essentials} and \cite{Coursera} is given in section \ref{Cha1:briefintro}  . For a more detailed and full introduction to game theory, the reader is referred to 
-\cite{leyton2008essentials}.  In section \ref{cha1:FlipItGame} an overview of the FlipIt game is given with the definitions and concepts that will be used throughout the paper. 
-The last section \ref{ch1:extendedWork} will cover the extensions and additions already made on FlipIt.
-
+\cite{leyton2008essentials} and \cite{Coursera} is given in this section. For a more detailed and full introduction to game theory, the reader is referred to 
+\cite{leyton2008essentials}.  \\
+%In section \ref{cha1:FlipItGame} an overview of the FlipIt game is given with the definitions and concepts that will be used throughout the paper. 
+%The last section \ref{ch1:extendedWork} will cover the extensions and additions already made on FlipIt. \\
 %------------------------------------------------%
@@
 %------------------------------------------------%
-\section{A brief introduction in Game Theory}
-\label{Cha1:briefintro}
 
@@
 Over John Nash is een prachtige film
-\subsection{Best response and Nash Equilibrium}
-One of the solution concepts in Game Theory for non-cooperative games is a Nash Equilibrium that we will use in this paper.
+\subsubsection{Best response and Nash Equilibrium}
+One of the solution concepts in Game Theory for non-cooperative games is a Nash Equilibrium that will be used in this paper.
 A Nash Equilibrium is a subset of outcomes that can be interesting to analyse a game.
@@
 The best response for a player is the action of a player that maximizes it's pay-off for any given action of the other player.
-We define $BR_{i}$ as the best response function for player \textit{i}.
-The best response for player \textit{1} is given by : $a_{1} = BR_{1}(a_{2})$.
+We define $Opt_{i}$ as the best response function for player \textit{i}.
+The best response for player \textit{1} is given by : $a_{1} = Opt_{1}(a_{2})$.
 A
@@
 Nobody has the incentive to change his or her action if an equilibrium profile is played.
-We have a Nash Equilibrium for the pair $(a_{1}^{*},a_{2}^{*})$ where $a_{1}^{*} = BR_{1}(a_{2}^{*})$ and $a_{2}^{*} = BR_{2}(a_{1}^{*})$\\
+We have a Nash Equilibrium for the pair $(a_{1}^{*},a_{2}^{*})$ where $a_{1}^{*} = Opt_{1}(a_{2}^{*})$ and $a_{2}^{*} = Opt_{2}(a_{1}^{*})$\\
 
- \todo{optimal strategy uitleggen}
+\todo{optimal function same as best response ?}
 
-POSTCONDITION: Uitgelegd: Strategien, acties, strategien, spelers, rationeel, Nash, best response
 %In game theory, the Nash equilibrium is a solution concept of a non-cooperative game involving two or more players, in which each player is assumed to know the equilibrium strategies of the other players, and no player has anything to gain by changing only their own strategy.[1] If each player has chosen a strategy and no player can benefit by changing strategies while the other players keep theirs unchanged, then the current set of strategy choices and the corresponding payoffs constitutes a Nash equilibrium.
@@
 \item \textit{Players}: Players are referred as the ones who are the decision makers. It can be a person, a company or an animal.  (they will act rational )
-\item \textit{Actions}: Every player has actions that he or he can do. 
+\item \textit{Actions}: Every player has actions that he or she can do. 
 \item \textit{Strategies}: A strategy is the combination of different actions. A pure strategy is only one action.
-\item \textit{Utility function}: The utility function is the mapping of the level of happiness of an agent about the state of the world to natural numbers.
+\item \textit{Utility}: The utility function or also known as the pay-of is the mapping of the level of happiness of an agent about the state of the world to natural numbers.
 
@@
 \centering
-\includegraphics[scale=0.5]{../../doc/template/Images/DefFlipit}
-\caption{A representation of a FlipIt game where both players are playing periodically and at discrete time intervals.
+\includegraphics[scale=0.7]{../../doc/template/Images/FLipi.png}
+\caption{A representation of a FlipIt game where both players are playing periodically.
 Every move or flip is indicated by a blue or orange circle.

### 2015-06-10 doc/template/nash.tex

@@
 theorem specifies Nash equilibria for the periodic game and is proven in Appendix A. \\
+\begin{figure}[hbtp]
+\caption{This figure shows the three subcategories of each case.
+'A' stands for Case 2.A: $\delta_{D} \geq d+\delta_{A} \geq \delta_{A}$ and 'B' stands for Case 2.B: $d+\delta_{A} \geq \delta_{D} \geq  \delta_{A} $ }
+\centering
+\includegraphics[scale=0.4]{Images/bestresp.png}
+\label{grafiekbestr}
+\end{figure}
 
-Nash equilibria are points whith the property that neither player benefits by deviating in isolaition form equilibrium.
+\subsection{Determining the piecewise functions $opt_{D}(\delta_{A})$}
+
+Nash equilibria are points with the property that neither player benefits by deviating in isolation form equilibrium.
 We can compute Nash Equilibria for the periodic game as an intersection points of curvest opt0 and opt1.
-To determine opt0(a0) we need to compute the derivate of  $\beta_{0}(\alpha_{0},\alpha_{1}) $ for a fixed $\alpha_{1}$.
+To determine opt0(a0) we need to compute the derivative of  $\beta_{0}(\alpha_{0},\alpha_{1}) $ for a fixed $\alpha_{1}$.
 We consider two cases:\\
------
 
-\textbf{Case: $\delta_{D} \leq \delta_{A} $}\\
+\subsection*{Case 1: $\delta_{D} \leq \delta_{A} $}
+The benefit formula obtained previously for this case is as follows:
+\begin{equation*}
+\beta_{D}(\alpha_{D},\alpha_{A}) = 1 - \dfrac{\delta_{D}}{2\delta_{A}} - \dfrac{k_{D}}{\delta_{D}} - \dfrac{d^{2}}{2\delta_{D}\delta_{A}} + \dfrac{d}{\delta_{A}}
+\end{equation*}
+If we take the partial derivative for a fixed $\delta_{1}$ we get the following result:
+\begin{equation*}\label{formdelta}
+\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} = - \dfrac{1}{2\delta_{A}} + \dfrac{k_{D}}{\delta_{D}^{2}} + \dfrac{d^{2}}{2\delta_{D}^{2}\delta_{A}}
+\end{equation*}
 
-\begin{equation}
-\beta_{D}(\alpha_{D},\alpha_{A}) = 1 - \dfrac{\delta_{D}}{2\delta_{A}} - \dfrac{k_{D}}{\delta_{D}} - \dfrac{d^{2}}{2\delta_{D}\delta_{A}} + \dfrac{d}{\delta_{A}}
-\end{equation}
-If we take the partial derivative we get the following result:
-\begin{equation}\label{formdelta}
-\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} = - \dfrac{1}{2\delta_{A}} + \dfrac{k_{D}}{\delta_{D}^{2}} + \dfrac{d^{2}}{2\delta_{D}^{2}\delta_{A}}
-\end{equation}
+The stationary points (maximum, minimum) can be found by setting the first derivative equal to zero and finding the roots of the resulting equation:
+\begin{equation*}
+\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} =0 ~~~~~~ =>~~~~~~ \delta_{D} = \sqrt{2\delta_{A}k_{D} + d^{2}}
+\end{equation*}
 
-Als we de formule \ref{formdelta} gelijkstellen aan 0 krijgen we:
-
-\begin{equation}
-\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} =0 ~~~~~~ =>~~~~~~ \delta_{D} = \sqrt{2\delta_{A}k_{D} + d^{2}}
-\end{equation}
-
-The function increases on $[0, \sqrt{2\delta_{A}k_{D} + d^{2}}]$ and is decreasing on $[\sqrt{2\delta_{A}k_{D} + d^{2}}, \infty]$ So we got a maximum on $\delta_{D} = minimum \{ \delta_{A}, \sqrt{2\delta_{A}k_{D} + d^{2}} \} $ \\
+This leads to the following deduction: The function increases on $[0, \sqrt{2\delta_{A}k_{D} + d^{2}}]$ and is decreasing on $[\sqrt{2\delta_{A}k_{D} + d^{2}}, \infty]$ So we got a maximum on $\delta_{D} = min \{ \delta_{A}, \sqrt{2\delta_{A}k_{D} + d^{2}} \} $.
+The minimum of the two values is needed because $\delta_{D}$ cannot be bigger than $\delta_{A}$.
+\\
 ~~\\
 
-\textbf{Case 2.A: $\delta_{D} \geq d+\delta_{A} \geq \delta_{A} $ }\\
+\subsection*{Case 2.A: $\delta_{D} \geq d+\delta_{A} \geq \delta_{A} $ }
 
-\begin{equation}
+The benefit formula obtained previously for this case is as follows:
+\begin{equation*}
 \dfrac{t \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - \dfrac{k_{D}}{\delta_{D}}
-\end{equation}
-\begin{equation}
+\end{equation*}
+
+The derivative of the above formula for a fixed $\delta_{A}$ results in the following:
+\begin{equation*}
 \frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} = -\dfrac{\delta_{A}}{2\delta_{D}^{2}} - \dfrac{d}{\delta_{D}^{2}} + \dfrac{k_{D}}{\delta_{D}^{2}}
-\end{equation}
+\end{equation*}
+The obtain the stationary points the first derivative is set equal to zero and the roots of the resulting equation are found:
+\begin{equation*}
+\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} =0 ~~~~~~ =>~~~~~~ \delta_{A} = 2(k_{D}-d) = dk_{D} - 2d
+\end{equation*}
 
-Als we de formule \ref{formdelta} gelijkstellen aan 0 krijgen we:
-\begin{equation}
-\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} =0 ~~~~~~ =>~~~~~~ \delta_{A} = 2(k_{D}-d) = dk_{D} - 2d
-\end{equation}
+This leads to the following deduction:
 \begin{description}
-\item If $k_{D} < d$ 
+\item If $k_{D} \leq d$ 
 \begin{description}
 \item decreasing $ \delta_{A} < 2(k_{D} -d)$
-\item increasing  $\delta_{A} > 2(k_{D} -d)$ \todo{equal}
+\item increasing  $\delta_{A} > 2(k_{D} -d)$ 
 \end{description}
@@
 \item increasing $ \delta_{A} < 2(k_{D} -d)$
-\item decreasing  $\delta_{A} > 2(k_{D} -d)$ \todo{equal}
+\item decreasing  $\delta_{A} > 2(k_{D} -d)$ 
 \end{description}
@@
 
-\textbf{Case 2.B: $d+\delta_{A} \geq \delta_{D} \geq  \delta_{A} $} \\
+\subsection*{Case 2.B: $d+\delta_{A} \geq \delta_{D} \geq  \delta_{A} $}
 
-\begin{equation}
+The benefit formula obtained previously for this case is as follows:
+\begin{equation*}
 \dfrac{\beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - \dfrac{k_{D}}{\delta_{D}} - \dfrac{(d-(\delta_{D} - \delta_{A}))^{2}}{d\delta_{D}\delta_{A}}
-\end{equation}
+\end{equation*}
 
-\begin{equation}
+The derivative of the above formula for a fixed $\delta_{A}$ results in the following:
+\begin{equation*}
 \frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} =  - \dfrac{1}{2\delta_{D}} + \dfrac{k_{D}}{\delta_{D}^{2}} + \dfrac{d^{2}}{2\delta_{D}^{2}\delta_{A}}
-\end{equation}
+\end{equation*}
 
-Als we de formule \ref{formdelta} gelijkstellen aan 0 krijgen we:
+The stationary points (maximum, minimum) can be found by setting the first derivative equal to zero and finding the roots of the resulting equation:
 
-\begin{equation}
+\begin{equation*}
 \frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} =0 ~~~~~~ =>~~~~~~ \delta_{D} = \sqrt{2\delta_{A}k_{D} + d^{2}}
-\end{equation}
+\end{equation*}
 
-The function increases on $[0, \sqrt{2\delta_{A}k_{D} + d^{2}}]$ and is decreasing on $[\sqrt{2\delta_{A}k_{D} + d^{2}}, \infty]$ So we got a maximum on $\delta_{D} = minimum \{ \delta_{A}, \sqrt{2\delta_{A}k_{D} + d^{2}} \} $ \\
+This leads to the following deduction: The function increases on $[0, \sqrt{2\delta_{A}k_{D} + d^{2}}]$ and is decreasing on $[\sqrt{2\delta_{A}k_{D} + d^{2}}, \infty]$ So we got a maximum on $\delta_{D} = minimum \{ \delta_{A}, \sqrt{2\delta_{A}k_{D} + d^{2}} \} $ \\
 
-The optimum functions will be piecewise functions. There will be different optimum functions depending on $k_{D}$ and $d$. Two cases are considered.
-\subsubsection{$k_{D} \leq d$}
+\subsubsection{Best responses}
+The optimum functions will be piecewise functions.
+There will be different optimum functions depending on $k_{D}$ and $d$.
+We distinguish for these two cases, three sub cases for different values of $\delta_{A}$.
 
-\begin{description}
+\subsection*{$k_{D} \leq d$}
+Because $k_{D} \leq d$, the term $2(k_{D} - d)$ will always be negative. We point out that $\delta_{A}$ and $\delta_{D}$ are positive rates. 
+\begin{itemize}
 \item if $\delta_{A} < 2(k_{D} - d)$ \\
-For case 2.A it is decreasing.
-Because $k_{D} \leq d$ the term $2(k_{D} - d)$ will always be negative so $\delta_{A}$ will also be negative.
-This means for case 1 and case 2.b that $\delta_{D} = 0$ and so the defender will not play.
+This means that $\delta_{A}$ has to be negative which is not possible. For case the defender will not play.
 \item $\delta_{A} = 2(k_{D} - d)$ \\
-$\delta_{A} = 0$ so the attacker will not play. For case 1 and case 2.b the defender will also not play.
+$\delta_{A}$ is negative or equal to 0 so the attacker will not play. For case 1 and case 2.b the defender will also not play.
 \item $\delta_{A} > 2(k_{D} - d)$ \\
-For case 2.b it is decreasing. Case 2.a it is increasing. For case 1 $2(k_{D} - d)$ is always negative so the minimum is 0. This means that $\delta_{D} \in [0, \sqrt{d^{2} + 2\delta_{A}k_{D}}]$
-\end{description}
+Case 2.a it is increasing for every value $\delta_{A} \in [0,\infty]$.  For case 1 together with case 2.b the optimal benefit is achieved at rate $\delta_{D} = \sqrt{d^{2} + 2\delta_{A}k_{D}}$.
+\end{itemize}
 
-\subsection{Best responses}
 \subsubsection{$k_{D} > d$}
-\begin{description}
+Because $k_{D} > d$, the term $2(k_{D} - d)$ will always be positive. We point out that $\delta_{A}$ and $\delta_{D}$ are positive rates.
+\begin{itemize}
 \item if $\delta_{A} < 2(k_{D} - d)$ \\
-For case 2.A it is increasing. 
+From case 2.a it follows that the benefit of the defender increases.
+From case 1 and case 2.b together the optimal benefit of the defender is achieved at rate $\delta_{D} = \sqrt{d^{2} + 2\delta_{A}k_{D}}$.
 \item $\delta_{A} = 2(k_{D} - d)$ \\
+From case 2.a it follows that $\beta_{D}(\delta_{D},\delta_{A})=0$, for all $\delta_{A} \in [0,2(k_{D} - d)]$.
+From case 1 and case 2.b together the optimal benefit for the defender is achieved for all rates $\delta_{D} \in [0, \sqrt{d^{2} + 2\delta_{A}k_{D}}]$.
 \item $\delta_{A} > 2(k_{D} - d)$ \\
-In every case the function is decreasing so the defender will not play.
-\end{description}
+For case 2.a the benefit is decreasing. From case 1 and case 2.b the best strategy for the defender is not playing at all. 
+\end{itemize}
+\todo{dat laatste nog eens nakijken}
 
@@
        0, & \delta_{A} = 2(k_{D} - d) \\
-       \big[0,\sqrt{d^{2} + 2\delta_{A}k_{D}}\big], & \delta_{A} > 2(k_{D} - d)
+       \sqrt{d^{2} + 2\delta_{A}k_{D}}, & \delta_{A} > 2(k_{D} - d)
      \end{array}
@@
      \begin{array}{lr}
-       0, & \delta_{A} < 2(k_{D} - d)\\
-       0, & \delta_{A} = 2(k_{D} - d) \\
-       \big[0,\sqrt{d^{2} + 2\delta_{A}k_{D}}\big], & \delta_{A} > 2(k_{D} - d)
+       \sqrt{d^{2} + 2\delta_{A}k_{D}}, & \delta_{A} < 2(k_{D} - d)\\
+       \big[0,\sqrt{d^{2} + 2\delta_{A}k_{D}}\big], & \delta_{A} = 2(k_{D} - d) \\
+       0, & \delta_{A} > 2(k_{D} - d)
      \end{array}
@@
 \end{displaymath}
--------------------------------------------\\
 
-\subsection{betaA}
+%*****************************************************************
+%
+% optimum functies voor beta A
+%
+%*********************************************************************
+\subsection{Determining the piecewise functions $opt_{A}(\delta_{D})$}
 We still consider the case where $d < \delta_{D}$. \\
-To determine the Nash equilibria we also need to determine $opt_{1}(\delta_{D})$ by computing the derivative of $\beta_{A}(\delta_{D},\delta_{A})$ for a fixed $\delta_{D}$. We consider 2 cases: \\
+To determine the Nash equilibria we also need to determine $opt_{A}(\delta_{D})$ by computing the derivative of $\beta_{A}(\delta_{D},\delta_{A})$ for a fixed $\delta_{D}$. We consider 2 cases: \\
 
-\textbf{Case 1: $\delta_{A} \geq \delta_{D}$}\\
+\subsection*{Case 1: $\delta_{A} \geq \delta_{D}$}
 
-Since 
+The benefit formula obtained previously for this case is as follows:
 \begin{equation*}
@@
 \end{equation*}
-the derivative is:
+The derivative for a fixed $\delta_{D}$ is as follows:
 \begin{equation*}
@@
 \end{equation*}
-it follows that $\beta_{A}(\delta_{D},\cdot)$ is increasing if $2k_{A} < (\delta_{D} - d)^{2} / \delta_{D}$ and decreasing if $2k_{A} > (\delta_{D} - d)^{2} / \delta_{D}$. \\
+The stationary points (maximum, minimum) can be found by setting the first derivative equal to zero and finding the roots of the resulting equation:
+\begin{equation*}
+\frac{\partial \beta_{A}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} =0 ~~~~~~ =>~~~~~~ 2k_{A} = \dfrac{(\delta_{D}-d)^{2}}{\delta_{D}}
+\end{equation*}
 
-\textbf{Case 2: $\delta_{A} \leq \delta_{D}$ } \\
+It follows that $\beta_{A}(\delta_{D},\cdot)$ is increasing if $2k_{A} < (\delta_{D} - d)^{2} / \delta_{D}$ and decreasing if $2k_{A} > (\delta_{D} - d)^{2} / \delta_{D}$. \\
 
-\textbf{A}: $\delta_{A} \leq d + \delta_{A} \leq \delta_{D}$ \\
-Since 
+\subsection*{Case 2.A: $\delta_{D} \geq d+\delta_{A} \geq \delta_{A} $ }
+The benefit formula obtained previously for this case is as follows:
 \begin{equation*}
@@
 \end{equation*}
-the derivative is:
+The derivative for a fixed $\delta_{D}$ is as follows:
 \begin{equation*}
 \dfrac{\partial \beta_{A}(\delta_{D},\delta_{A})}{\partial \delta_{A}} = \dfrac{-1}{2\delta_{D}} + \dfrac{k_{A}}{\delta_{A}^{2}}
+\end{equation*}
+The stationary points (maximum, minimum) can be found by setting the first derivative equal to zero and finding the roots of the resulting equation:
+\begin{equation*}
+\frac{\partial \beta_{A}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} =0 ~~~~~~ =>~~~~~~ \delta_{A} = \sqrt{2\delta_{D}k_{A}}
 \end{equation*}
@@
 
-\textbf{B}: $\delta_{A}  \leq \delta_{D} \leq d + \delta_{A} $ \\
+\subsection*{Case 2.B: $d+\delta_{A} \geq \delta_{D} \geq  \delta_{A} $}
 
-Since 
+The benefit formula obtained previously for this case is as follows: 
 \begin{equation*}
@@
 \end{equation*}
+The stationary points (maximum, minimum) can be found by setting the first derivative equal to zero and finding the roots of the resulting equation:
+\begin{equation*}
+\frac{\partial \beta_{A}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} =0 ~~~~~~ =>~~~~~~ 2k_{A} = \dfrac{(\delta_{D}-d)^{2}}{\delta_{D}}
+\end{equation*}
+
 it follows that $\beta_{A}(\delta_{D},\cdot)$ is increasing if $2k_{A} < (\delta_{D} - d)^{2} / \delta_{D}$ and decreasing if $2k_{A} > (\delta_{D} - d)^{2} / \delta_{D}$.
 This is the same result as in case 1.\\
-%%% Local Variables: 
-%%% mode: latex
-%%% TeX-master: "thesis"
-%%% End:
 
-%\end{document}
+\subsubsection{Best responses}
+The optimum functions will be piecewise functions. We distinguish three cases for different values of $\delta_{d}$ and $k_{A}$.
+
+For this term $\dfrac{(\delta_{D}-d)^{2}}{\delta_{D}} $ , $d$ has to be bigger than  $\delta_{D}$ because the cost $k_{A}$ cannot be negative.
+This was an assumption that was already made, because the benefit of the defender will always be 1 if $d$ is bigger than  $\delta_{D}$.
+\begin{itemize}
+\item if $2k_{A} < \dfrac{(\delta_{D}-d)^{2}}{\delta_{D}} $ \\
+Then for case 1 and case 2.b the benefit of the defender is increasing. From case 2.a follows that the optimal benefit for the attacker is achieved at the rate $\delta_{A} = \delta_{D}$
+\item if $2k_{A} = \dfrac{(\delta_{D}-d)^{2}}{\delta_{D}} $ \\
+From case 1 and case 2.b it follows that $\beta_{D}(\delta_{D},\delta_{A})=0$, for all $\delta_{A} \in [0,\dfrac{(\delta_{D}-d)^{2}}{2k_{A}})]$.
+From case 2.a the optimal benefit for the defender is achieved for all rates $\delta_{A} \in [0, \delta_{D}]$.
+\item if $2k_{A} > \dfrac{(\delta_{D}-d)^{2}}{\delta_{D}} $ \\
+All decreasing.
+\end{itemize}

### 2015-06-10 doc/template/chap-5.tex
verbeteringen jeroen

@@
 Section 2.2 then introduces the reader into the main principles of game theory.
-Subsequently, section 2.3 introduces the reader to the FlipIt game, the specific game that will be used to model cyber security attacks of periodic nature and including a delay.
-Finally, section 2.4 gives an overview of the related work, and how the research presented in this thesis is positioned compared to existing results.
+Subsequently, section 2.3 introduces the reader to the FlipIt game: the game that will be used to model cyber security attacks of a periodic nature, including a delay.
+Finally, section 2.4 gives an overview of the related work, and how the research presented in this paper compares to existing results.
 
@@
 In today's digital era more and more information is kept in a digital format, stored on a computer.
-As digitalization progressed, the need for ensuring the security of digital information arose and automated tools where developed for protecting files stored on a computer.
-The generic name to protect data stored on a computer controlled device such as computers and smartphones, as well as public and private computer networks, including the entire Internet is called computer security.
-Security is a general term that encompasses several dimensions. More specifically, computer security has tree key objectives that are fundamental to computer security:
+As digitalization progressed, the need for ensuring the security of digital information arose and automated tools were developed to protect files stored on a computer.
 
-~~\\
- \textit{Confidentiality}: This assures that the confidential of private data is not disclosed or made available to users that do not have the authorization. \\
- \textit{Integrity:} This assures that data can not be altered by an unauthorized individual.\\
- \textit{Availability:} This assures that data is always accessible and that the service is not denied to authorized individuals.\\
+The generic term for data protection stored on a computer controlled device such as computers and smartphones, as well as public and private computer networks, including the entire Internet is called computer security.
+Security is a general term that encompasses several dimensions. More specifically, computer security has three key objectives that are fundamental to computer security:
+
+\begin{enumerate}
+\item  \textit{Confidentiality}: assuring that the confidentiality of private data is not disclosed or made available to users that do not have proper authorization.
+\item \textit{Integrity:} assuring that data cannot be altered by an unauthorized individual.
+\item \textit{Availability:} assuring that data is always accessible and that the service is not denied to authorized individuals.
+\end{enumerate}
 
  %The purpose of security is to give certainty that data will not be removed without authorization (Confidentiality) that the data is always accessible (Availability), and that data can not be read or altered by someone who does not have the authorization (Integrity).
- These are the 3 key attributes of security, also known as the CIA triad. The tree concepts are fundamental security objectives for the securing of data, information and computing services. \\
-Cyber security is the process of applying security tools to ensure confidentiality, integrity, and availability of data.
+ ~~\\
+ These key attributes are also known as the CIA triad. They are the fundamental security objectives for securing data, information and computing services. \\
+ \\
+A subset of computer security is cyber security.
+This is the process of applying security tools to ensure confidentiality, integrity, and availability of data.\todo{what is cyber security?
+In deze tekst lijkt het op computer security}
+
+%hieronder is wel okee
 It is an attempt to protect websites, servers, data, applications, operating systems and all assets that need protection in a computer system.
@@
 An identification tool will try to identify the threat to be able to know how to remove it.
-The removal tool will once the threat has been identified, remove the threat from the system so that it cannot spread any further.
+The removal tool will remove the threat from the system (once it has been identified) so that it cannot spread any further.
 
@@
 In order to keep a system secure and to be able to apply security tools, it is important to mitigate the possible threats.
-The possible threats can take many forms, the most common being spam, malware, spoofing, phishing and  DDoS attacks.
-In the context of cyber security, the terms ' threat and attack' are often used interchangeably, referring to more or less the same thing.
-They however have a slightly different meaning.
-A threat refers to anything that can breach the security and that can cause a possible harm.
+The possible threats can take many forms, the most common being: spam, malware, spoofing, phishing and  DDoS attacks.
+In the context of cyber security, the terms ' threat'  and 'attack' are often used interchangeably, referring to more or less the same thing.
+The meaning of these two terms, however, differ slightly from one another.
+The former refers to anything that can breach security and cause possible harm.
 It is a possible danger that can exploit a vulnerability.
-An attack is an assault on computer security that comes from a threat.
-It is an intelligent act that tries deliberately to breach security through vulnerabilities.
+The latter is an assault on computer security that originates from a threat.
+It is an intelligent act that deliberately tries to breach security through vulnerabilities.
 \\
 
-The most noteworthy group and biggest group of threats to computer systems is malware. 
-This is a piece of malicious software that is designed to penetrate unprotected systems or computers, with the intent to get sensitive information, destroy data, or compromise the confidentiality, integrity or availability of the data or applications of the victim.
+The most noteworthy and biggest group of threats to computer systems is malware. 
+This is a piece of malicious software that is designed to penetrate unprotected systems or computers, with the intent to retrieve sensitive information, destroy data, or compromise the confidentiality, integrity or availability of the data or applications of the victim.
 A security report of 2014 \cite{SurveyKaspersky} reveals that 61\% of the attacks on companies are caused by malware.
 For this reason this section will examine the categories of malware threats.
-Different types of malware exist such as virusses, worms, flooders, rootkits, bots, spyware, adware and many more.
+Different types of malware exist: virusses, worms, flooders, rootkits, bots, spyware, adware and many more.
 This broad range of different types can be classified into two main categories, the first one based on the propagation method that is used and the second one for the payload or the variety of actions that the malware performs.
 Propagation methods include virusses, worms and trojans.
-Payload includes flooders, rootkits, bots, spyware and adware.
-In this paper we focus on the category of propagation methods and not on the actions that the malware performs.
-We give a brief explanation of the three main propagation methods of malware.
+Payload includes e.g.
+flooders, rootkits, bots, spyware and adware.
+This paper focuses on the category of propagation methods and not on the actions that the malware performs.
+A brief explanation of the three main propagation methods of malware is given below.
 \todo{opsplitsing uit het boek van network security genomen}
+
+%\begin{table}
+%\begin{tabular}{|r|l|}
+% \hline
+% \textbf{Name}	 & \textbf{Description} \\  \hline 
+% Virus & octal \\  \hline 
+% Worm & binary \\  \hline 
+% Logic bomb
+% Trojan Horse
+% Backdoor
+% Mobile code
+% Exploits
+% Downloaders
+% Auto-rooter
+% Kit
+%Spammer programs
+%Keyloggers
+%Flooders
+%\end{tabular}
+%\caption{Types of Malicious Software, taken over from []}
+%\label{table}
+%\end{table}
+%
 
@@
 The ''I love you'' virus is an example of a virus where the virus attached itself as an executable to a mail.
-To propagate it used the mail systems.
-If someone opens an email with the "I love you"  in the annex, the virus will spread itself by sending a mail to everyone in the victim's contact list.
-So the virus can multiply rapidly and eventually a business network might shut down by the heavy traffic.
-In this example, there is a need for human interaction to spread the virus.
-If no one opens the mail the virus can not spread itself and infect other systems.
+It propagated by using the mailing systems.
+When a victim opened an email with the ''I love you'' virus in the annex, the virus spread itself by sending a mail to everyone in the victim's contact list.
+
+Using this method, a virus can multiply rapidly, possibly even causing a business network to shut down by the heavy traffic.
+A virus needs human interaction to spread.
+In the example above: if no one were to open the mail, the virus could not spread itself and infect other systems.
 
 \item \textit{Worm:} A worm is a virus that can spread without human interaction.
-A worm is a computer program that replicates itself in order to spread to other hosts on a network.
+It is a computer program that replicates itself in order to spread to other hosts on a network.
 Copies of the worm can be forwarded via a computer network without an intermediary.
@@
 The Stuxnetworm is a very prominent example of a worm.
-Initially this worm was spread via infected USB sticks and from there it could spread itself without any further human intervention through the Internet to other hosts on the network.
-The purpose of the Stuxnetworm was to harm the centrifuges in nuclear reactors and many reactors have been infected.
+Initially this worm was spread via infected USB sticks and from there on it could spread itself to other hosts on the network through the Internet, without any further human intervention.
+The purpose of the Stuxnetworm was to harm the centrifuges in nuclear reactors; many reactors have been infected.
 
@@
 A notable trojan horse is Koobface, that targeted users of facebook, skype, yahoo, gmail and AOL mail.
-To spread itself the worm sended a mail or friend request with a message that directed the recipients to another third party website.
-This site tried to put the reciepient into downloading an update of Adobe Flash Player.
+To spread itself the worm sent a mail or friend request with a message that directed the recipients to a third party website.
+This site would then convince the recipient into downloading an update of Adobe Flash Player.
 Once downloaded and executed, Koobface could infect the host.
@@
 \end{description}
-
-A company can take different measures to defend itself against malware.
-The threats caused by malware can be divided into three main categories: known threats (70\%), unknown threats (29\%) and advanced threats (1\%) as proposed by Kasperksy \cite{APTKaspersky}.
+~~\\
+The threats caused by the malware described above can be divided into three main categories: known threats (70\%), unknown threats (29\%) and advanced threats (1\%) as proposed by Kasperksy \cite{APTKaspersky}.
 
@@
 Standard malware protection tools like firewalls and virus scanners can keep these kind of malware out of the system.
-Installing protection against unkown is also relatively easy, but this will need tools that go beyond the standard methods like dynamic whitelisting.
+Installing protection against unkown threats is also relatively easy, but this requires tools that go beyond the standard methods, e.g.
+dynamic whitelisting.
 The remaining 1\% are the advanced threats, also known as APT.
+They are the most difficult to deal with.
 
@@
 
- An APT is a persistent targeted attack that tries to penetrate a network to cause harm while staying unseen for a long period of time.
-The motive of an APT is mostly cyber espionage, stealing sensitive data, sabotage or some other kind of ideological attacks.
-Advanced Persistent Threat are called  'Advanced' for the fact that these attacks are well funded and that (usually) the attacker itself needs a great expertise to successfully penetrate a network.
-Not all APTs are technical advanced though.
-The attacker can also try to exploit existing vulnerabilities simply based in the hope that his target has not yet secured himself against these vulnerabilities.
-'Persistent' refers to the fact that the attacker keeps on trying to attack his target.
-The attack can be over various years and different steps can be taken.
-The 'threats' stands for the fact that an intelligent act tries deliberately to breach security through vulnerabilities which may severely damage the target organisation.
-An APT can be a mix of different types of malware and will use different kinds of propagation methods. \\
+An Advanced Persistent Threat (referred to in the rest of this work as APT) is a persistent targeted attack that tries to penetrate a network to cause harm while staying unseen for a long period of time.
+The motive of an APT is mostly cyber espionage, stealing sensitive data, sabotage or other kinds of ideological attacks.
+APT's are 'advanced' because these attacks are well funded and (usually) the attacker needs a great amount of expertise to successfully penetrate a network.
+Not all APT's are \textit{technically} advanced though.
+The attacker can also try to exploit known vulnerabilities, in the hope that his target has not yet secured himself against them.
+\\
+'Persistent' refers to the fact that the attacker never quits trying to attack his target. The attack can be spread over several years, taking multiple steps.  \\
+An APT can be a mix of different types of malware and may use various propagation methods. \\
 %
@@
 
-According to a security survey of Kaspersky \cite{SurveyKaspersky}  the damage of one successful target attack against a large company can exceed over 2.54 million dollar.
-A company needs a defence mechanism to defend itself against APTs.
-As previously said, detection and identification tools will not work against APTs.
-The removal tool will only work if the threat has been identified.
-To mitigate these kind of attacks there is a need for another security countermeasures.
-Here is where the researchers of RSA fall in.
-They have investigated through game theory how to model targeted attacks and what the best defend strategy is according to the information that is available.
+According to a security survey of Kaspersky \cite{SurveyKaspersky}  the damage of one successful targeted attack against a large company can exceed over 2.54 million dollar.
+Companies need a defence mechanism to defend themselves against APT's.
+As previously stated, simple detection and identification tools are insufficient to protect oneself against APT's, and a removal tool will only work if the threat has been identified.
+To mitigate these kind of attacks another security countermeasure is needed.
+This is where the researchers of RSA come into the picture.
+They have modelled these targeted attacks using Game Theory, and - given a certain amount of information - determined optimal defence strategies.\todo{bekijk dit}
 
@@
 We have a Nash Equilibrium for the pair $(a_{1}^{*},a_{2}^{*})$ where $a_{1}^{*} = Opt_{1}(a_{2}^{*})$ and $a_{2}^{*} = Opt_{2}(a_{1}^{*})$\\
+
+An example to explain a nash equilibrium of a two-player game, for example, is the prison dilemma.
+In this game there are two players who are both rational and both of them have committed a crime.
+'Rational' means that they want the best for themselves and it is not their purpose to do harm to others.
+Both are locked in a separate room and they cannot tell in advance to each other what they are going to say.
+Each of them can betray the other or they can support each other and remain silent.
+If a player confesses, he will, depending on what the other does, go for three years in prison or he is free to go.
+If the player is silent, he will, depending on what the other player does, go for five years in jail or only one year.
+\\
+Figure 1 shows the various combinations of the possible actions of the two players.
+The Nash equilibrium of the game is that they both remain silent, but perfectly rational players still choose to confess both.
+This is because this is the dominant strategy.
+A dominant strategy is a strategy that is better than all other strategies of a player regardless of the opponent.
+Here it is advantageous for each player to confess.
+This gives them three years in prison instead of a year.
+Why players choose one or the other action can be explained by means of game theory.
+
+\begin{figure}[hbtp]
+\caption{The prison dilemma: P1 corresponds to the first column, P2 corresponds to the second column.}
+\centering
+\includegraphics[scale=1]{Images/PrisDilemma.jpg}
+\end{figure}
 

### 2015-06-11 doc/template/chap-9.tex
figuren toegevoegd

@@
 \label{ch2:diffFlip}
+\begin{figure}[hbtp]
+\caption{The first game is the basic FlipIt game.
+The second is a FlipIt game with a delay.
+During the first flip of the attacker, the defender moves after the delay, causing the attacker to get control over the resource.
+During the second flip of the attacker, the defender flips at time t+d, causing the defender to take control over the resource before the attacker.
+During the third and final flip of the attacker, the defender flips during time t+d, causing the attacker to never gain control over the resource.
+}
+\centering
+\includegraphics[scale=0.7]{Images/dtFlip.png}
+\label{dt}
+\end{figure}
 
@@
 If the defender flips the network before the period $d$ has elapsed (so, somewhere between $t$ and $t + d$), then the attacker will never gain full control over the resource.
+See figure \ref{dt}.
 This implies that the mathematical formulas for gain and benefit need to be adapted to the fact that the attacker loses part of its benefit because of this delay.
@@
 \begin{figure}[hbtp]
-\caption{Case 2 where d + deltaA < deltaD}
+\caption{All possible cases for the attacker and the defender in Case 2.A where $d + delta_{A} < delta_{D}$.
+As can be seen, the defender will have control during a period of \textit{d} over the resource when the defender has flipped in the previous interval.}
 \centering
-\includegraphics[scale=0.5]{../../doc/template/Images/FlipItCase2.png}
+\includegraphics[scale=0.7]{../../doc/template/Images/FlipItCase2a.png}
 \label{fig:case2}

### 2015-06-11 doc/template/chap-5.tex

@@
 
+\todo{stukje geschiedenis toevoegen}
+
 %begin over dat gametheorie handig is in de economie
@@
 
-An example to explain a nash equilibrium of a two-player game, for example, is the prison dilemma.
+An example to explain a Nash equilibrium of a two-player game, for example, is the prison dilemma.
 In this game there are two players who are both rational and both of them have committed a crime.
@@
 \\
-Figure 1 shows the various combinations of the possible actions of the two players.
-The Nash equilibrium of the game is that they both remain silent, but perfectly rational players still choose to confess both.
+
+%Because betraying a partner offers a greater reward than cooperating with him, all purely rational self-interested prisoners would betray the other, and so the only possible outcome for two purely rational prisoners is for them to betray each other.[1] The interesting part of this result is that pursuing individual reward logically leads both of the prisoners to betray, when they would get a better reward if they both cooperated.
+
+Figure \ref{Prison} shows the rewards of the possible actions of the two players.
+Because betraying the other player gives a greater reward than remaining silent, all prisoners will betray each other which leads to the outcome that both of them will betray each other.
 This is because this is the dominant strategy.
@@
 Here it is advantageous for each player to confess.
-This gives them three years in prison instead of a year.
-Why players choose one or the other action can be explained by means of game theory.
+This gives them three years in prison instead of five year or they will be free instead of being in prison for one year.
+The Nash equilibrium of the game is that they both confess even though that it would be better if they both cooperated and chose to stay silent.
 
 \begin{figure}[hbtp]
-\caption{The prison dilemma: P1 corresponds to the first column, P2 corresponds to the second column.}
+\caption{Prisoners dilemma: An example from the field of Game Theory.
+It is a game between two perfectly rational prisoners who don not know what the other one will do.
+Each can confess or stay silent.
+The outcomes can be seen in the picture.
+Prisoner 1 corresponds to each first column, Prisoner 2 corresponds to each second column.
+}
 \centering
-\includegraphics[scale=1]{Images/PrisDilemma.jpg}
+\includegraphics[scale=0.6]{Images/PrisDilemma}
+\label{Prison}
 \end{figure}

### 2015-06-11 doc/template/chap-3.tex
Kleine verbetering + virus prop model

@@
 
-Difference with FlipThem: sub part of nodes for control and strategy difference: dependant of grade of the nodes, instead of just periodic.
-- Literatuurstudie
-- Flipit
-- Game Motivation
-- Formal definition
+Some related work, al dan niet game theorie
 
@@
 
-\subsection{What can be done in further research}
-\begin{itemize}
-\item Looking for the dynamics of the spread of the virus/worm limited by the bandwidth of the network links, BPG routing failure with high volume scan traffic
-\end{itemize}
-\section{Conclusion}
-
 \section{Why Game Theory to model security problem}
@@
 
-\section{..}
-
-Flip-it. Some authors have written other papers about flipit. One of them is the [Game theoretic approach for cost-benefit analysis of malware proliferation prevention].
-
 %%% Local Variables: 

### 2015-06-11 doc/template/chap-5.tex
Kleine verbetering + virus prop model

@@
 The possible threats can take many forms, the most common being: spam, malware, spoofing, phishing and  DDoS attacks.
-In the context of cyber security, the terms ' threat'  and 'attack' are often used interchangeably, referring to more or less the same thing.
+In the context of cyber security, the terms `threat'  and `attack' are often used interchangeably, referring to more or less the same thing.
 The meaning of these two terms, however, differ slightly from one another.
@@
 A typical virus will attach itself to a program, or an executable content on a computer.
-The ''I love you'' virus is an example of a virus where the virus attached itself as an executable to a mail.
+The ``I love you'' virus is an example of a virus where the virus attached itself as an executable to a mail.
 It propagated by using the mailing systems.
-When a victim opened an email with the ''I love you'' virus in the annex, the virus spread itself by sending a mail to everyone in the victim's contact list.
+When a victim opened an email with the ``I love you'' virus in the annex, the virus spread itself by sending a mail to everyone in the victim's contact list.
 
@@
 The motive of an APT is mostly cyber espionage, stealing sensitive data, sabotage or other kinds of ideological attacks.
-APT's are 'advanced' because these attacks are well funded and (usually) the attacker needs a great amount of expertise to successfully penetrate a network.
+APT's are `advanced' because these attacks are well funded and (usually) the attacker needs a great amount of expertise to successfully penetrate a network.
 Not all APT's are \textit{technically} advanced though.
@@
 \\
-'Persistent' refers to the fact that the attacker never quits trying to attack his target. The attack can be spread over several years, taking multiple steps.  \\
+`Persistent' refers to the fact that the attacker never quits trying to attack his target. The attack can be spread over several years, taking multiple steps.  \\
 An APT can be a mix of different types of malware and may use various propagation methods. \\
@@
 In this game there are two players who are both rational and both of them have committed a crime.
-'Rational' means that they want the best for themselves and it is not their purpose to do harm to others.
+`Rational' means that they want the best for themselves and it is not their purpose to do harm to others.
 Both are locked in a separate room and they cannot tell in advance to each other what they are going to say.
@@
 Our motivation for a defender with stealthy moves is that there is not always an intelligent individual that is behind an APT.
-Some APTs don't know if the computer is already been recovered.
+Some APT's do not know if the computer is already been recovered.
 There purpose is to spread.

### 2015-06-11 doc/template/chap-9.tex
Kleine verbetering + virus prop model

@@
 
-Motivation
+Welke aanpassing en motivatie: attacker 1 speler die elke keer virus probeert te droppen.
+Defender die elke keer patcht als er virus gedropt wordt.
+Elke patch voldoet aan de meest recente patch dus alle virussen die ook gedropt zijn worden weggedaan.
+Voor future work: patch systeem aanpassen.
+
+The actions of a defender: the actions of an attacker:
 

### 2015-06-11 doc/template/chap-10.tex
Kleine verbetering + virus prop model
Aanvullingen overal.

@@
 They study the specific scenario where a system or network is repeatedly taken over completely by an without being immediately detected by the defender of the system or network.
-In game theory, such a game is known as ''FlipIt'' \cite{FlipIt}.
+In game theory, such a game is known as ``FlipIt'' \cite{FlipIt}.
 This is a two players game where the attacker and the defender are competing to get control over a shared resource.
@@
 FlipIt does not take into account that a move may not be instantaneous, but has a certain delay.
-Consider for example a network with different nodes ( laptops, datacenters) as a resource.
+Consider for example a network with different nodes (laptops, datacenters) as a resource.
 The attacker drops a virus on one of the nodes and then wait till this virus infects the whole network.
@@
 \subsubsection{Contributions}
-We propose an addition to the basic FlipIt model to model a scenario where the moves by the attacker will not be instantaneous. Next we analyse what the new Nash equilibria will be and .. \\
+The following contributions are made in this paper:
+\begin{itemize}
+\item[-] We propose an addition to the basic FlipIt model to model a scenario where the moves by the attacker will not be instantaneous.
+We extend the FlipIt game where the attacker flips with a delay.
+The attacker only comprimises the system if suffi"ent nodes in the network are infected.
+\item[-] Another derivation of the method to calculate the benefit of the basic FlipIt game is given.
+\item[-] The periodic case of FlipIt is modelled with a delay, resulting in optimum functions and nash equilibria. 
+\item[-] Based on our results we can give practical recommendations to take measures against advanced attacks. 
+\end{itemize}
+ Next we analyse what the new Nash equilibria will be and .. \\
 
+While it seems maybe trivial to extend FlipIt with a delay it is not so straightforward.
+First a formula to calculate the benefit of a basic FlipIt was derived to introduce the delay.
+To calculate the benefit for each case \todo{cases misschien uitleggen?}, the intervals where not independant anymore as in the basic FlipIt game.
+Er moest rekening gehouden worden met de vorige intervallen om de benefit te berekenen in het eigen interval.
 \subsubsection{Overview of the thesis}

### 2015-06-11 doc/template/chap-4.tex
Aanvullingen overal.

@@
-%\chapter{Intoduction to GameTheory}
-%\label{cha:1}
-\documentclass[10pt]{article}
-\begin{document}
+\chapter{Virus Propagation }
+\label{chapter4: Virus propagation}
 
-\section{Write down the settings of the game}
+\section{Methods of virus propagation}
+
+The spreading of worms has already been extensivly researched [beter schrijven].
+Because of all the different types of propagation it is hard to pick out one model that can model all of them.
+Modelling the spread of malware depends on two great factors: the method used for the propagation and the graph where the malware will spread itself.
+Viruses and worms are the types of malware that are the most researched.
+Under we list a couple of examples to model virus and worm propagation.
+In this text worm and virus will be used intermediate.
+
+\subsection{Models of graph representing the network}
+
+\subsection*{Methods of propagation}
+
+\begin{description}
+\item Random scanning: Some worms propagate by using the method of random scanning 
+IP addresses.
+Rate of terms of succeeded randwomly choosen IP adresses -> maximise good ip adresses.
+After a while the remaining nodes will be les likely to be reached.
+An example of such a worm is `Code Red' .
+\item Routing worms: they use BGP routing tables to scan the routable address space. 3 times faster than traditional worm that uses random scanning. Spyb0t or network bluepill.
+\item email worms: use the email systems to propagate. The Kak worm is a Javascript computer worm that spread itself by exploiting a bug in Outlook Express. 
+%Pikachu worm: The virus was mainly spread through Microsoft Outlook email attachments.
+The email containing the attached virus propagated through infected users by sending itself to all contacts in the user's Outlook address book.
+[5]
+\item self stopping worm: reduces speed to avoid detection. Atak worm or self stopping worm
+\end{description}
+
+The start of such a propagation can be different.
+Either it is a targeted attack or a random attack.
+For example a usb stick or other shared resources can be used to start the spreading.
+It can be given to some special person or it can just be left behind in a parking lot.
+This can still be a targeted attack if it is the parking lot of a targeted company but there is more insecurity of where the usb stick will end.
+Another way is by sending an email.
+It can also start from one computer that is connected directly to the internet.
+Downloading malicious programs.
+
+\subsection*{Models for worm propagation}
+To use the models for worm propagation in our model of the FlipIt game it is suffi"ent to know the propagation speed of the worm and define after how many time units the network is defected.
+This will be equal to the \textit{d} parameter in our model.
+Stackelberg game: first move is from attacker, defender is follower.
+
+Kind of network model: 
+Kind of model: SIS, SIR
+
+\subsubsection*{SIS}
+
+\subsubsection*{SIR}
+
+\subsubsection*{self disciplinary worms}
+Model for self disciplinary worms and counter measures ... []
+
+Popular mechanism that worms use to detect vulnerable targets by random ip scanning probing.
+Feasible due to use 32-bit addresses.
+128-bit adresses life harder for worms, except the ones that use email systems to propagate.
+two new strategies: uniformly distributed random number generator to select new target.
+:spread locally, by biasing the search space towards addresses within the same subnet or network.
+The second strategy is almost the same as what email virusses would get. For this reason we work an example out ..
+
+A method to calculate the propagation of the virus in an easy way. Google page ranking algorithm.
+
+\section{methode met matrixen}
 
@@
 
-\end{document}
+What do we need for an algorithm
+\begin{description}
+\item Graph network $G = < V, E>$
+\item Graph matrix $[A]$ which is $|V| \times |V| $
+\item Attack vector $[X]$ which is $1 \times |V|$
+\item cummulative matrix $[M]$ which is $|V| \times |V|$
+\item state matrix $[T]$  which is $|V| \times |V|$
+\item Reset vector $[R]$
+\item duration \textit{d}
+\item time \textit{n}
+\item rate $\delta _{0}$ of defender and $\delta _{1}$ of attacker
+\end{description}
+
+Initialisation algorithm:
+
+\begin{verbatim}
+initialisatie
+	d=0
+	A=basismatrix
+	M=A^{0}
+	n=0
+	\delta_{0}
+	\delta_{1}
+	X
+	R
+	controller = defender
+
+	Algorithm
+	n:= n + 1;
+	Check who is in control? ( through modulo )
+	if ( defender & controller=defender)
+				d:= d + 1;
+
+	if ( defender & controller=attacker )
+				G = X \times R  (flippen ten voordele van defender)
+				d = 0
+				controller = defender
+
+	if ( attacker & controller=defender )
+				controller=attacker
+				..
+
+	if ( attacker & contoller=attacker )
+				d:= d + 1
+				M = M x A
+				T = T + M
+				G = X x T
+
+\end{verbatim}
+
+%\end{document}

### 2015-06-11 doc/template/chap-5.tex
Aanvullingen overal.

@@
-\chapter{Intoduction to GameTheory}
+\chapter{Intoduction to the general context of the work}
 \label{Chapter1:Intro.Game.Theory}
@@
 
-The generic term for data protection stored on a computer controlled device such as computers and smartphones, as well as public and private computer networks, including the entire Internet is called computer security.
-Security is a general term that encompasses several dimensions. More specifically, computer security has three key objectives that are fundamental to computer security:
+The generic term for data protection stored on a computer controlled device such as computers and smartphones, as well as public and private computer networks, including the entire Internet is called information security.
+Security is a general term that encompasses several dimensions. More specifically, information security has three key objectives that are fundamental to information security:
 
@@
  \\
-A subset of computer security is cyber security.
-This is the process of applying security tools to ensure confidentiality, integrity, and availability of data.\todo{what is cyber security?
+Information security is divided into two main areas of security.
+One of them is cyber security also known as computer security.
+This is the process of applying security tools to ensure confidentiality, integrity, and availability of data.
+the other area is Network or Internet security..
+The difference between these two terms has been lost over time.
+\todo{what is cyber security?
 In deze tekst lijkt het op computer security}
@@
 \todo{opsplitsing uit het boek van network security genomen}
+
+%\begin{figure}[hbtp]
+%\caption{2014 survey results by Kaspersky}
+%\centering
+%\includegraphics[scale=0.5]{Images/kaspmalware.png}
+%\end{figure}
 
@@
 As previously stated, simple detection and identification tools are insufficient to protect oneself against APT's, and a removal tool will only work if the threat has been identified.
-To mitigate these kind of attacks another security countermeasure is needed.
+To mitigate these kind of attacks another security countermeasures are needed.
 This is where the researchers of RSA come into the picture.
-They have modelled these targeted attacks using Game Theory, and - given a certain amount of information - determined optimal defence strategies.\todo{bekijk dit}
+They have modelled the situation where targeted attacks are taking over a system and are not immediately detected by the defender.
 

### 2015-06-11 doc/template/chap-9.tex
Aanvullingen overal.

@@
 
-Similarly as in \cite{FlipIt}, we split the formalization in two cases.
-The first case is where the defender plays at least as fast as the attacker, the second case is where the attacker plays at least as fast as the defender.
-For each of these cases, first the benefit formula of the basic case without delay is presented, and then the delay is introduced.
-\\
-
-Intuitively we could assume that $d$ can never be bigger than $\delta_{A}$ because then the benefit for the defender would always be 1.
-This is not always true.
-It is only true if $d$ is bigger than $\delta_{D}$, because then the defender will always be in control.
-For this we only calculate the formulas for the cases where $d$ is smaller than $\delta_{D}$.
-We can already conclude that it is no use for the attacker to play when the delay is bigger than $\delta_{D}$.
-\\
-
 \subsection{Formalization the benefit formula including the infection-delay}
@@
 \end{description}
+
+Similarly as in \cite{FlipIt}, we split the formalization in two cases.
+The first case is where the defender plays at least as fast as the attacker, the second case is where the attacker plays at least as fast as the defender.
+For each of these cases, first the benefit formula of the basic case without delay is presented, and then the delay is introduced.
+\\
+
+Intuitively we could assume that $d$ can never be bigger than $\delta_{A}$ because then the benefit for the defender would always be 1.
+This is not always true.
+It is only true if $d$ is bigger than $\delta_{D}$, because then the defender will always be in control.
+For this we only calculate the formulas for the cases where $d$ is smaller than $\delta_{D}$.
+We can already conclude that it is no use for the attacker to play when the delay is bigger than $\delta_{D}$.
+see figure \ref{delaye}\\
+
+\todo{figuur juist maken}
+\begin{figure}[hbtp]
+\caption{Delay bigger than deltaA and delay bigger than deltaD}
+\centering
+\includegraphics[scale=0.7]{Images/DiffDelayCase1.png}
+\label{delaye}
+\end{figure}
 
@@
 \subsubsection*{\textbf{Case a:} $d + \delta_{A} \leq \delta_{D}$}
-
+The first case is where the delay is shorter than the difference between the attacker's and the defender's period so their is no problem of counting the delay twice.
+To determine the total gain rate of the defender we need to know what the probability is that the defender will move during an interval and what the average time is that the defender controls the resource.
+\todo{beter verwoorden}.
 The total gain  rate of the defender is then the probability that the defender will move during an interval of the attacker multiplied by the total average gain per interval:

### 2015-06-13 doc/template/chap-10.tex
versie voor mama

@@
 In this era where digitalization becomes prominent in every aspect of our lives, where technology is growing fast and where businesses are always under attack, security becomes an issue of increasing complexity.
+Security is needed to protect websites, servers, applications, data, operating systems and other assets that need protection in a computer network.
 Without security, there is no protection to keep somebody out of a system.
@@
 Also, disruption caused by distributed DoS attacks, may results in businesses failing to meet their service-level agreements.
-Ultimately, system and network security helps protecting a business's reputation, which is one of its most important assets.
+Ultimately, system and network security helps protecting a business's reputation against various kind of threats, which is one of its most important assets.
 \\
@@
 
-A particular kind of threats are Advanced Persistent Threats (APT).
-An APT is a multi-faceted, continuous and targeted cyber attack that is designed to penetrate a network or a system in a stealthy way and that can stay undetected for a long period.
+A particular kind of threat is an Advanced Persistent Threat (APT).
+An APT is a multi-faceted, continuous and targeted cyber attack that is designed to penetrate a network or a system in a stealthy way and can stay undetected for a long period.
 It is something different and stronger than a conventional threat.
@@
 Researchers at RSA, van Dijk et al,  made a game theoretic model of targeted attacks.
-They study the specific scenario where a system or network is repeatedly taken over completely by an without being immediately detected by the defender of the system or network.
+They study the specific scenario where a system or network is repeatedly taken over completely by an attacker without being immediately detected by the defender of the system or network.
 In game theory, such a game is known as ``FlipIt'' \cite{FlipIt}.
@@
 The attacker drops a virus on one of the nodes and then wait till this virus infects the whole network.
-The attacker will only be in control of the resource once the whole network is infected.
+The attacker will only be in control of the resource when a sufficient large amount of nodes of the network are infected.
 \\
@@
 
-\subsubsection{Contributions}
+\subsubsection{Contributions and results}
 The following contributions are made in this paper:
@@
  Next we analyse what the new Nash equilibria will be and .. \\
+ The modelling of FlipIt with a delay will give a guideline for both the attacker and the defender in when they should attack the network or how regurlarly should they clean the network.
 
@@
 Er moest rekening gehouden worden met de vorige intervallen om de benefit te berekenen in het eigen interval.
+
+\subsubsection{Conjunctures and open problems}
+
 \subsubsection{Overview of the thesis}

### 2015-06-13 doc/template/chap-5.tex
versie voor mama

@@
 This chapter provides the reader with an introduction to the general context of the work presented in this paper.
-Section 2.1 introduces the reader into the basic concepts of cyber security and the kind of cyber security threaths that are in the scope for this work.
+Section 2.1 introduces the reader into the basic concepts of cyber security and the kind of cyber security threats that are in the scope for this work.
 Section 2.2 then introduces the reader into the main principles of game theory.
@@
 
-\section{What is cyber security?}
+\section{What is security?}
 
@@
  \\
-Information security is divided into two main areas of security.
+Information security can be divided into two main areas of security.
 One of them is cyber security also known as computer security.
 This is the process of applying security tools to ensure confidentiality, integrity, and availability of data.
-the other area is Network or Internet security..
-The difference between these two terms has been lost over time.
-\todo{what is cyber security?
-In deze tekst lijkt het op computer security}
-
-%hieronder is wel okee
 It is an attempt to protect websites, servers, data, applications, operating systems and all assets that need protection in a computer system.
@@
 The removal tool will remove the threat from the system (once it has been identified) so that it cannot spread any further.
+\\
+The other area is Network or Internet security.
+This area of security protect data during transmission.
+There is no clear distinction between these two areas of security.
+For example a worm can be physically dropped on a computer network using a usb stick, but it can also arrive over the internet.
+Either way the internal computer security measurements have to be taken to recover from the virus.
+This paper focusses on cyber security.
+
+%hieronder is wel okee
 
@@
 For this reason this section will examine the categories of malware threats.
-Different types of malware exist: virusses, worms, flooders, rootkits, bots, spyware, adware and many more.
-This broad range of different types can be classified into two main categories, the first one based on the propagation method that is used and the second one for the payload or the variety of actions that the malware performs.
-Propagation methods include virusses, worms and trojans.
+Different types of malware exist: viruses, worms, flooders, rootkits, bots, spyware, adware and many more.
+This broad range of different types can be classified into two main categories, the first one based on the propagation method that is used and the second one for the payload or the variety of actions that the malware performs \cite{stallings2007network}.
+Propagation methods include viruses, worms and trojans.
 Payload includes e.g.
@@
 A brief explanation of the three main propagation methods of malware is given below.
-\todo{opsplitsing uit het boek van network security genomen}
 
@@
 As previously stated, simple detection and identification tools are insufficient to protect oneself against APT's, and a removal tool will only work if the threat has been identified.
-To mitigate these kind of attacks another security countermeasures are needed.
+To mitigate these kind of attacks another security countermeasure is needed.
 This is where the researchers of RSA come into the picture.
@@
 %------------------------------------------------%
-
-\todo{stukje geschiedenis toevoegen}
-
-%begin over dat gametheorie handig is in de economie
-
 Game theory studies the interaction between independent and self-interested agents.
@@
 It has therefore important applications in many area's such as economics, politics, biology, computer science, philosophy and a variety of other disciplines.
-\\
+It has come only in ascent during the Second World War with Oskar Morgenstern and John von Neumann both publishing a book on game theory, entitled "Theory of Games and Economic Behavior" (Game theory and economic behavior) in 1944.
+This book was about the mathematical analysis of a series of thinking games.
+It was a distinction made between games in which the strategies and the utility factor of the opponent has no effect on finding the best strategies such as chess and games in which it has an influence like poker.
+John Nash also played a major role in the history of game theory.
+He was one of the mathematicians who has formalized game theory.
+The Nash equilibrium was named after him.
+~~\\
 %Every agent has different levels of happiness for the different outcomes.
@@
 This means that the attacker knows when the defender plays and can change its strategy depending on the moves of the defender.
-Our motivation for a defender with stealthy moves is that there is not always an intelligent individual that is behind an APT.
-Some APT's do not know if the computer is already been recovered.
-There purpose is to spread.
-Not to check if they have already infected.
-\todo{beter verwoorden}.
+Our motivation for a defender with stealthy moves is that not every APT is given feedback to one place.
+Some ATP's are designed only to cause harm on the systems that are infected e.g.
+the Stuxnetworm that was developed to destroy nuclear reactors.
+The only way that the attacker got feedback was through the news.
+There purpose is to spread and cause harm.
+Another example are APT's that get onto isolated networks through usb stick or cd's.
+There is no way for the worm to communicate back on an isolated network.
 The second difference is that even though both the targeted and non-targeted attacks do not succeed immediately, the delay is determined differently.
@@
 This move involves also an extra cost.\\
-A three-player game has also been investigated where the flipit framework of two players is extended by another player.
+A three-player game has also been investigated where the FlipIt framework of two players is extended by another player.
 This player represents an insider that trades value information with the attacker \cite{fengstealthy}.\\
@@
 %------------------------------------------------%
-%            Intro about virusses				 %
+%            Intro about viruses				 %
 %------------------------------------------------%

### 2015-06-13 doc/template/chap-9.tex
versie voor mama

@@
-\chapter{FlipIt game with virus propagation}
+\chapter{FlipIt with propagation delay}
 \label{chapter2:FlipIt with virus propagation}
@@
 \label{Ch2:Intro}
-The game of FlipIt with virus propagation considers attackers where their moves are not instantaneous.
+The game of FlipIt with propagation delay considers attackers where their moves are not instantaneous.
 This can be motivated by an example of a virus.
@@
 The basic FlipIt game does not take this into account.
-In this chapter the FlipIt game with virus propagation is explained and how it can be modelled.
+In this chapter the FlipIt game with propagation delay is explained and how it can be modelled.
 First the main differences with the basic FlipIt game are discussed in section \ref{ch2:diffFlip}.
-Next in section \ref{ch2:formal} the formal definition of the game with virus propagation is given.
-At last, in section \ref{ch2:periodicvirus} a periodic strategy game is considered and the formalization of the formulas with virus propagation are determined.\\
+Next in section \ref{ch2:formal} the formal definition of the game with propagation delay is given.
+At last, in section \ref{ch2:periodicvirus} a periodic strategy game is considered and the formalization of the formulas with propagation delay are determined.
+The use of the word virus or worm will be used interchangeably in this paper.
+\todo{EEN van de twee intros nemen, virus term gebruiken ?}\\
 
-This chapter explains how to model a FlipIt game with a virus propagation that infects a network.
-The first section explains the difference between a normal FlipIt game and a FlipIt game with virus propagation.
-The next section derives a formula to calculate the benefit for a FlipIt game with a virus propagation.
+This chapter explains how to model a FlipIt game with a propagation delay that infects a network.
+The first section explains the difference between a normal FlipIt game and a FlipIt game with propagation delay.
+The next section derives a formula to calculate the benefit for a FlipIt game with a  propagation delay.
 In the last section we calculate the Nash equilibrium for the benefit formula.
-\section{FlipIt game with virus propagation}
-
-Welke aanpassing en motivatie: attacker 1 speler die elke keer virus probeert te droppen.
+%
+%\section{FlipIt game with virus propagation}
+%
+%Welke aanpassing en motivatie: attacker 1 speler die elke keer virus probeert te droppen.
 Defender die elke keer patcht als er virus gedropt wordt.
@@
 Voor future work: patch systeem aanpassen.
+%
+%The actions of a defender: the actions of an attacker:
 
-The actions of a defender: the actions of an attacker:
-
-\section{Explaining difference between FlipIt with and without virus propagation}
+\section{Difference between FlipIt with and without propagation delay}
 \label{ch2:diffFlip}
@@
 
-A FlipIt game consists of a single resource. To represent the security problem, the game now defines its single resource as a computer network with multiple
+The framework of FLipIt is used to model targeted attacks.
+In FlipIt when an attacker attacks, he will get immediately control over the resource.
+One of the differences is that FlipIt does not consider that moves of the attacker might not be instantaneous.
+With FlipIt with a propagation delay we model the fact that if the attacker flips, he does not get immediate control over the resource.
+This is motivated by the fact that it takes a while before a targeted attack can compromise a computer network.
+The virus or worm has to spread to infect every computer controlled device in the network.
+This propagation make take a while.
+\\
+
+Another difference is that the FlipIt game consists of a single resource.
+To represent the security problem of the propagation of a virus in a network, the game now defines its single resource as a computer network with multiple
 nodes. One of the players, the defender, will try to defend his network. The defender
@@
 Rather than denoting the time needed for infecting \textit{all} the nodes, the variable $d$ can also be used to denote the time needed to infect \textit{a sufficient number} of nodes.
-
-Assume that an attacker attacks at time \textit{t}, he doesn't get immediate control over the resource, but he only gains control at time \textit{t + d}, with $d$ denoting the time needed to infect a sufficiently number (or all) nodes.
-If the defender flips the network before the period $d$ has elapsed (so, somewhere between $t$ and $t + d$), then the attacker will never gain full control over the resource.
-See figure \ref{dt}.
-This implies that the mathematical formulas for gain and benefit need to be adapted to the fact that the attacker loses part of its benefit because of this delay.
-In the remainder of this paper, we will adapt the formalization of the FlipIt game using the variable $d$.
 \\
 
+The action of the defender consist of defending his network.
+In this context it will be defending every computer controlled device in a network.
+To make the network clean the defender can patch the systems with the latest security bulletins or reinstall the software or even wipe out the computer.
+All these different actions can imply other costs.
+The cost of patching a system will be smaller than the cost of wiping out a computer.
+\\
+The defender has only one action.
+Sending different viruses and worms to a computer network.
+Every time the attacker flips the resource, he does this by sending a new virus to the system.
+The attacker does not send the same virus again, because if the computers are patched, the virus cannot get in anymore.
+
+\subsubsection{Motivation Application}
+\subsubsection*{Wormlike ATP}
+A business is every day under attack of different threats.
+An APT that successfully enters a companies network can cause a lot of damage.
+In this example the ATP will be an attacker that sends worms with stealthy features to attack a company network.
+He will keep on doing this, which is the persistent thing of the APT.
+The defender is unaware that an APT is infecting his company network.
+It is better for a company to assume that it is under attack.
+\\
+The worm can use the computers on the network as botnets or steal sensitive data.
+By patching or installing new software the worm can be removed form the systems.
+The important assets of the defender to secure in a company network are the computers, servers, data, files, and so fort.
+In this particular example the resource to defend consists of all the computer controlled devices on the network, e.g.
+servers and computers.
+The attacker can drop the worm through a usb stick or by tricking a victim to open or download a malicious program.
+The last one can be achieved by a trojan horse.
 %%------------------------------------------------%
@@
 
-\section{Playing periodically with virus propagation}
+\section{Playing periodically with propagation delay}
 \label{ch2:periodicvirus}
@@
 This choice is motivated by the assumption that in most organisations, the defence strategy is to periodically defend the network.
+This is true for the example of patching.
+Companies like microsoft en google release at a specific moment patches.
+So companies will apply these patches when they come out.
+Microsoft security bulletins are released on the Second Tuesday of each month.
+%\source{https://technet.microsoft.com/en-us/security/bulletin/dn602597.aspx} 
 This corresponds to a periodic defender strategy.
@@
 
+Assume that an attacker attacks at time \textit{t}, he doesn't get immediate control over the resource, but he only gains control at time \textit{t + d}, with $d$ denoting the time needed to infect a sufficiently number (or all) nodes.
+If the defender flips the network before the period $d$ has elapsed (so, somewhere between $t$ and $t + d$), then the attacker will never gain full control over the resource.
+See figure \ref{dt}.
+This implies that the mathematical formulas for gain and benefit need to be adapted to the fact that the attacker loses part of its benefit because of this delay.
+In the remainder of this paper, we will adapt the formalization of the FlipIt game using the variable $d$.
+\\
+
 Similarly as in \cite{FlipIt}, we split the formalization in two cases.
@@
 
-Intuitively we could assume that $d$ can never be bigger than $\delta_{A}$ because then the benefit for the defender would always be 1.
-This is not always true.
+Intuitively we could assume that $d$ can never be bigger than $\delta_{A}$ because then the benefit for the defender would always be 1, \textit{but} this is not always true.
 It is only true if $d$ is bigger than $\delta_{D}$, because then the defender will always be in control.
@@
 see figure \ref{delaye}\\
+\todo{bovenstaande tekst beter uitleggen}
 
@@
 
-A similar analysis as in case 1 for a FlipIt game without virus propagation yields the following benefits:
+A similar analysis as in case 1 for a FlipIt game without a propagation delay yields the following benefits:
 
@@
 The above benefit formula would then include to much gain for the defender: the potential overlap during the delay would be counted twice.
-\\
+See figure \ref{overlap}\\
+\begin{figure}[hbtp]
+\caption{Cases where the delay would be counted twice.}
+\centering
+\includegraphics[scale=0.7]{Images/FlipItcase2b.png}
+\end{figure}
 

### 2015-06-13 doc/template/chap-10.tex
verbeteringen

@@
 
-%  Situation
-%  Complication
-%  Question
-%  Answer
-
-% woorden die in introductie moeten voorkomen: Game theory, cybersecurity, APT, FlipIt, virus propagation, network security
-
-\subsubsection{Situation}
-% -- Situation -- %
-
 In this era where digitalization becomes prominent in every aspect of our lives, where technology is growing fast and where businesses are always under attack, security becomes an issue of increasing complexity.
@@
 Why is it so important to keep a system secure?
-Many businesses store confidential information on clients, which can be lost and possible be abused by competitors through data leakage.
-Also, disruption caused by distributed DoS attacks, may results in businesses failing to meet their service-level agreements.
-Ultimately, system and network security helps protecting a business's reputation against various kind of threats, which is one of its most important assets.
+Many businesses store confidential information on clients, which can be lost through data leakage and possible be abused by competitors.
+Also, disruption caused by distributed denial of service (DDoS) attacks, may results in businesses failing to meet their service-level agreements.
+Ultimately, computer and network security helps protecting a business against various kind of threats.
 \\
@@
 
-\subsubsection{Complication}
-Since it is so difficult to protect a system or a network against APT's, researchers have been looking for effective ways to predict in advance which defence strategy might be the better one. 
+%\subsubsection{Complication}
+%Since it is so difficult to protect a system or a network against APT's, researchers have been looking for effective ways to predict in advance which defence strategy might be the better one. 
+There are a number of key strategies that an organisation can apply to defend themselves against APT's: awareness, whitelisting, system administration, network segregation, dynamic content checking and patch management.
+Nevertheless the combination of all these elements benefits from being complemented by other defence strategies to protect one selves against stealthy takeovers.
+A way to study the impact of stealthy takeovers and to determine practical recommendations for defenders is through game theory.\\
 Game theory is gaining increasing interest as an effective technique to model and study cyber security problems.
@@
 In this paper we focus on the game with two players.
-The actions available to the attacker and the defender correspond to attacks on the system and defensive measures that protect the system.
-In particular, both players will execute the strategy that results in a maximal benefit for them.
-Researchers at RSA, van Dijk et al,  made a game theoretic model of targeted attacks.
-They study the specific scenario where a system or network is repeatedly taken over completely by an attacker without being immediately detected by the defender of the system or network.
-In game theory, such a game is known as ``FlipIt'' \cite{FlipIt}.
-This is a two players game where the attacker and the defender are competing to get control over a shared resource.
+The actions available to the attacker and the defender correspond to respectively to the attacks on the system and taking defensive measures that protect the system.
+\\
+
+ Researchers at RSA, van Dijk et al,  presented a game-theoretic framework to model computer security scenarios called ``FlipIt'' \cite{FlipIt}.
+They study the specific scenario where a system or network is repeatedly taken over completely by an attacker and not immediately being detected by the defender of the system or network.
+This is a two-player game where the attacker and the defender are competing to get control over a shared resource.
 Neither player knows who is currently in control of the resource until they move.
 In FlipIt every move gives them immediate control over the resource.
+\\
  But what if the attacker moves and it takes a while before the attacker gets full control over the resource?
@@
 The attacker will only be in control of the resource when a sufficient large amount of nodes of the network are infected.
+In this paper we present FlipIt with propagation delay to model a game where the moves of the attacker are not instantaneous.
+A formalization for this game starts from the model of non-adaptive continuous basic FlipIt game where players use a periodic strategy with a random phase.
 \\
@@
 \item How can we incorporate the notion of delay in the game-theoretical analysis of the Flip-It game for a periodic strategy ?
-\item Is there an optimal defense strategy against an attacker ? 
+\item Does the resulting model allow an optimal defence strategy against an attacker ? 
 \end{itemize}
@@
 We extend the FlipIt game where the attacker flips with a delay.
-The attacker only comprimises the system if suffi"ent nodes in the network are infected.
+The attacker only compromises the system if suffi"ent nodes in the network are infected.
 \item[-] Another derivation of the method to calculate the benefit of the basic FlipIt game is given.
-\item[-] The periodic case of FlipIt is modelled with a delay, resulting in optimum functions and nash equilibria. 
+\item[-] The periodic case of FlipIt is modelled with a delay, resulting in optimum functions and Nash equilibria. 
 \item[-] Based on our results we can give practical recommendations to take measures against advanced attacks. 
@@
  Next we analyse what the new Nash equilibria will be and .. \\
- The modelling of FlipIt with a delay will give a guideline for both the attacker and the defender in when they should attack the network or how regurlarly should they clean the network.
+ The modelling of FlipIt with a delay will give a guideline for both the attacker and the defender in when they should attack the network or how regularly should they clean the network.
 
@@
 
+\begin{description}
+\item[-] Analyse other strategies like renewal strategies instead of only periodic strategies.
+\item[-] The defender can chose to Flip one node of the network or a subset instead of all the nodes.
+\item[-] Give the defender a delay. It takes a while before a patch is made after the discovery of a vulnerability.
+\item[-]
+\end{description}
+
 \subsubsection{Overview of the thesis}
@@
 The organisation of this paper is the following.
-In chapter \ref{Chapter1:Intro.Game.Theory}  a brief introduction to cyber security and gametheory is introduced to get familiar with the kind of threats that are in the scope of this work and the game theoretic concepts that will be further used in the paper.
-In the same chapter the FlipIt framework is summarized and the most important conclusions together with the the related work done on FlipIt and the difference with this paper.
-In chapter \ref{chapter2:FlipIt with virus propagation} , we first introduce the adaptations made on FlipIt to model a FlipIt game with virus propagation.
-After that formulas are derived to model a FlipIt game with a virus propagation for a specific case where players play a periodic strategy with a random phase.
+In chapter \ref{Chapter1:Intro.Game.Theory}  a brief introduction to cyber security and game theory is introduced to get familiar with the kind of threats that are in the scope of this work and the game theoretic concepts that will be further used in the paper.
+In the same chapter the FlipIt framework is summarized with its most important conclusions.
+The chapter concludes with related work on FlipIt and further clarifies the contributions of this paper compared to existing work.
+Chapter \ref{chapter2:FlipIt with virus propagation} , first introduces the adaptations made to the original FlipIt model to model a FlipIt game with virus propagation.
+Subsequently formulas are derived to model a FlipIt game with a virus propagation for the specific case where players play a periodic strategy with a random phase.
 \\
-In Chapter \ref{chapter:Nash} the formules are used to determine if there is a nash equilibrium. \todo{aanpassen naar chapter met nash equilibria}
-Next in Chapter \ref{chapter4: Virus propagation} we given an overview of the various ways in which a virus can propagate.
-We present a method to calculate the speed of the propagation of a virus in a network and how this network can be established to reduce the spreading of a virus.
-Finally, in Chapter \ref{chapter5:conclusion} , we discuss the main results and complications and provide directions for further research.
+In Chapter \ref{chapter:Nash} the formules are further analysed to determine if there is a nash equilibrium. \todo{aanpassen naar chapter met nash equilibria}
+In order to provide a clear perspective on delays, chapter \ref{chapter4: Virus propagation} gives an overview of the various ways in which a virus can propagate.
+It presents a method to calculate the speed of the propagation of a virus in a network and how this network can be established to reduce the spreading of a virus.
+Finally chapter \ref{chapter5:conclusion} , we discusses the main results and provides directions for further research.
 

### 2015-06-13 doc/template/chap-5.tex
verbeteringen

@@
-\chapter{Intoduction to the general context of the work}
+\chapter{General context}
 \label{Chapter1:Intro.Game.Theory}
@@
 This chapter provides the reader with an introduction to the general context of the work presented in this paper.
-Section 2.1 introduces the reader into the basic concepts of cyber security and the kind of cyber security threats that are in the scope for this work.
+Section 2.1 introduces the reader into the basic concepts of cyber security and the kind of cyber security threats that are in scope for this work.
 Section 2.2 then introduces the reader into the main principles of game theory.
@@
  \\
-Information security can be divided into two main areas of security.
+Information security can be divided into two main sub categories.
 One of them is cyber security also known as computer security.
@@
 \\
-The other area is Network or Internet security.
-This area of security protect data during transmission.
-There is no clear distinction between these two areas of security.
-For example a worm can be physically dropped on a computer network using a usb stick, but it can also arrive over the internet.
+The other category is network or internet security.
+This sub category of information security protects data during transmission.
+While cyber security and network security address different aspects, they partially overlap as well.
+For example a worm can be physically dropped on a computer network using a USB stick, but it can also arrive over the internet.
 Either way the internal computer security measurements have to be taken to recover from the virus.
-This paper focusses on cyber security.
+In this paper we focus on scenarios where a network has to be defended against attacks to ensure confidentiality, integrity and availability of data.
+The work presented in this paper therefore belongs to the domain of cyber security.
 
@@
 The most noteworthy and biggest group of threats to computer systems is malware. 
-This is a piece of malicious software that is designed to penetrate unprotected systems or computers, with the intent to retrieve sensitive information, destroy data, or compromise the confidentiality, integrity or availability of the data or applications of the victim.
+This is a piece of malicious software that is designed to penetrate unprotected or vulnerable systems or computers, with the intent to retrieve sensitive information, destroy data, or compromise the confidentiality, integrity or availability of the data or applications of the victim.
 A security report of 2014 \cite{SurveyKaspersky} reveals that 61\% of the attacks on companies are caused by malware.
@@
 Different types of malware exist: viruses, worms, flooders, rootkits, bots, spyware, adware and many more.
-This broad range of different types can be classified into two main categories, the first one based on the propagation method that is used and the second one for the payload or the variety of actions that the malware performs \cite{stallings2007network}.
+This broad range of different types can be classified into two main categories, the first one based on the propagation method that is used and the second one based on the payload or the variety of actions that the malware performs \cite{stallings2007network}.
 Propagation methods include viruses, worms and trojans.
@@
 
-\item \textit{Trojan:} This is a malicious program that disguises itself as something normal and useful, so that users won't be suspicious of installing it, but it has a malicious function hidden inside that can avoid security measures and cause harm.
-A notable trojan horse is Koobface, that targeted users of facebook, skype, yahoo, gmail and AOL mail.
+\item \textit{Trojan:} This is a malicious program that disguises itself as something normal and useful, so that users won't be suspicious of installing it, but it has a malicious function hidden inside that can circumvent security measures and cause harm.
+A notable trojan horse is Koobface, that targeted users of Facebook, Skype, Yahoo, Gmail and AOL mail.
 To spread itself the worm sent a mail or friend request with a message that directed the recipients to a third party website.
@@
 ~~\\
-The threats caused by the malware described above can be divided into three main categories: known threats (70\%), unknown threats (29\%) and advanced threats (1\%) as proposed by Kasperksy \cite{APTKaspersky}.
+As proposed by Kasperksy \cite{APTKaspersky} ,threats caused by the malware described above can be divided into three main categories: known threats (70\%), unknown threats (29\%) and advanced threats (1\%).
 
@@
 %------------------------------------------------%
-Game theory studies the interaction between independent and self-interested agents.
-It is a mathematical way of modelling the interactions between two or more agents where the outcomes depend on what everybody does and how it should be structured to lead to good outcomes.
-It has therefore important applications in many area's such as economics, politics, biology, computer science, philosophy and a variety of other disciplines.
-It has come only in ascent during the Second World War with Oskar Morgenstern and John von Neumann both publishing a book on game theory, entitled "Theory of Games and Economic Behavior" (Game theory and economic behavior) in 1944.
+Game theory is a mathematical way of modelling the interactions between two or more agents where the outcomes depend on what everybody does and how it should be structured to lead to good outcomes.
+Game theory has therefore important applications in many area's such as economics, politics, biology, computer science, philosophy and a variety of other disciplines.
+It has come only in ascent during the Second World War with Oskar Morgenstern and John von Neumann both publishing a book on game theory, entitled "Theory of Games and Economic Behavior" (Game theory and economic behaviour) in 1944.
 This book was about the mathematical analysis of a series of thinking games.
-It was a distinction made between games in which the strategies and the utility factor of the opponent has no effect on finding the best strategies such as chess and games in which it has an influence like poker.
+The book described a series of games making a distinction between games in which the strategies and the utility factor of the opponent have no effect on finding the best strategies such (e.g.
+chess) and games in which this has an influence (e.g.
+poker).
 John Nash also played a major role in the history of game theory.
@@
 This kind of games are referred to as non-cooperative game theory, where the basic modelling unit is the group of agents.
-The individualistic approach, where the basic modelling is only one agent, is referred as cooperative game theory.
+The individualistic approach, where the basic modelling is only one agent, is referred to as cooperative game theory.
 
@@
 Nobody has the incentive to change his or her action if an equilibrium profile is played.
-We have a Nash Equilibrium for the pair $(a_{1}^{*},a_{2}^{*})$ where $a_{1}^{*} = Opt_{1}(a_{2}^{*})$ and $a_{2}^{*} = Opt_{2}(a_{1}^{*})$\\
+We have a Nash Equilibrium for the pair $(a_{1}^{*},a_{2}^{*})$ when $a_{1}^{*} = Opt_{1}(a_{2}^{*})$ and $a_{2}^{*} = Opt_{2}(a_{1}^{*})$\\
 
-An example to explain a Nash equilibrium of a two-player game, for example, is the prison dilemma.
+An example to explain a Nash equilibrium of a two-player game is the prison dilemma.
 In this game there are two players who are both rational and both of them have committed a crime.
@@
 Figure \ref{Prison} shows the rewards of the possible actions of the two players.
-Because betraying the other player gives a greater reward than remaining silent, all prisoners will betray each other which leads to the outcome that both of them will betray each other.
-This is because this is the dominant strategy.
-A dominant strategy is a strategy that is better than all other strategies of a player regardless of the opponent.
-Here it is advantageous for each player to confess.
-This gives them three years in prison instead of five year or they will be free instead of being in prison for one year.
-The Nash equilibrium of the game is that they both confess even though that it would be better if they both cooperated and chose to stay silent.
+The figure shows that it is advantageous for each player to confess.
+If prisoner 1 talks and the other prisoner remains silent, prisoner 1 will be free.
+If prisoner 1 remains silent and the other prisoner talks, prisoner 1 gets 5 years in prison.
+If the prisoners cooperate and talk, both of them get three years in prison.
+If they both remain silent they will both get one year in prison.
+So this means that talking is the dominant strategy.
+A dominant strategy is a strategy that is better than all other strategies regardless of what the opponent does.
+The Nash equilibrium of the game is that they both confess even though it would be better if both prisoners cooperated and chose to stay silent.
 
 \begin{figure}[hbtp]
+\centering
+\includegraphics[scale=0.6]{Images/PrisDilemma}
 \caption{Prisoners dilemma: An example from the field of Game Theory.
@@
 Each can confess or stay silent.
-The outcomes can be seen in the picture.
-Prisoner 1 corresponds to each first column, Prisoner 2 corresponds to each second column.
+The outcomes are shown below each prisoner.
 }
-\centering
-\includegraphics[scale=0.6]{Images/PrisDilemma}
 \label{Prison}
@@
 
-\subsubsection{List of terms}
-In the following list a couple of terms that will be used throughout the paper.
-\begin{description}
-\item \textit{Players}: Players are referred as the ones who are the decision makers. It can be a person, a company or an animal.  (they will act rational )
-\item \textit{Actions}: Every player has actions that he or she can do. 
-\item \textit{Strategies}: A strategy is the combination of different actions. A pure strategy is only one action.
-\item \textit{Utility}: The utility function or also known as the pay-of is the mapping of the level of happiness of an agent about the state of the world to natural numbers.
-
-\end{description}
-
-A game in game theory consists of multiple agents and every agent has a set of actions that he can play.
+%\subsubsection{List of terms}
+%The following is a list with the main terms of game theory that will be used throughout the paper.
+%\begin{description}
+%\item \textit{Players}: Players are referred as the ones who are the decision makers. It can be a person, a company or an animal.  (they are assumed to act rational )
+%\item \textit{Actions}: Every player has actions that he or she can do. 
+%\item \textit{Strategies}: A strategy is the combination of different actions. A pure strategy consists of only one action.
+%\item \textit{Utility}: The utility function or also known as the pay-off is the mapping of the level of happiness of an agent about the state of the world to natural numbers.
+%
+%\end{description}
+%
+%A game in game theory consists of multiple agents and every agent has a set of actions that he can play.
 \section{The FlipIt game}
@@
 
-FlipIt is a two-players game with a shared single resource that the players want to control as long as possible.
+FlipIt is a two-players game with a shared single resource that the players want to control as much as possible, see \ref{fig:FLipItDefault}.
 The shared resource can be a password, a network or a secret key depending on the setting being modelled.
@@
 The game begins at $t=0$ and continues indefinitely ($t \rightarrow \infty $).
-The time in the game is assumed as being continuous.
+The time in the game is assumed to be continuous.
 To get control over the resource, the players $i$, with $i \in \{A,D\}$, can flip the resource at any given time.
@@
 
-The unique feature of FlipIt is that every move happens in a stealthy way, meaning that the player has no clue that the other player (his adversary) has flipped the resource.
-For instance, the defender does not find out if the resource has been compromised by the attacker until he flips the resource himself.
+The unique feature of FlipIt is that every move happens in a stealthy way, meaning that the player has no clue whether the other player (his adversary) has flipped the resource or not.
+For instance, the defender does not know if the resource has been compromised by the attacker until he flips the resource himself.
 The goal of the player is to maximize the time that he or she has control over the resource while minimizing the total cost of the moves.
@@
 It may happen that the resource was already under control by the player.
-If the player moves when he or she has already control over the resource, he or she would have wasted a move since it does not result in a change of ownership, so the cost is wasted.
+If the player moves when he or she has already control over the resource, he or she would have wasted a move since this does not result in a change of ownership, so the cost is wasted.
 \\
@@
 \centering
-\includegraphics[scale=0.7]{../../doc/template/Images/FLipi.png}
+\includegraphics[scale=0.7]{../../doc/template/Images/DefFlipItt.png}
 \caption{A representation of a FlipIt game where both players are playing periodically.
 Every move or flip is indicated by a blue or orange circle.
-The attacker is represented in orange and plays with a period of $\delta_{A}=4$.
-The defender is represented in blue and plays with a period of $\delta_{D}=3$.
+The attacker is represented in orange and the defender is represented in blue.
 The blue and orange rectangles represent the amount of time the respective player is in control of the resource.}
@@
 \end{equation}
+~$ \alpha_{i}$ defines the average move rate by player \textit{i} up to time \textit{t}.
 Let $\beta_{i}(t)$ denote player's \textit{i} average benefit upto time \textit{t}:
@@
 This is equal to the fraction of time the resource has been owned by player \textit{i}, minus the cost of making the moves.
-~$ \alpha_{i}$ defines the average move rate by player \textit{i} up to time \textit{t}.
 In a given game, the asymptotic benefit rate (or simply benefit) will be defined as the lim inf of the average benefit because time\textit{ t} will increase to infinity and the average benefit may not have limiting values.
@@
 
-\subsubsection{strategies}
+\subsubsection{Strategies}
 Because the players move in a stealthy way, there are different types of feedback that a player can get while moving.
@@
 The non-adaptive strategies and the adaptive strategies.
-These are described in table \ref{table:Strategies}.
-\\
+These are described in table \ref{table:Strategies}.\\
 
@@
 \\
+
 In case there is feedback, a player can adapt his strategy to the information received about the opponent's moves.
@@
 \\
+
 In this paper we restrict ourselves to periodic strategies.
@@
 
-FlipIt has also been applied to several cases in system security.
+FlipIt also has been applied to several cases in computer security.
 Reseachers explored different applications of FlipIt for real-world problems, like password reset policies, VM refresh, cloud auditing and key rotation \cite{ApplyingFlipit}.
 \\
-Other authors used the FlipIt game to apply it on a specific scenario. To be able to use the FlipIt game, modifications where required for the FlipIt model.
+Other authors used the FlipIt game to apply it to a specific scenario. To be able to use the FlipIt game, modifications where required for the FlipIt model.
 One of the scenarios by Pham \cite{compromised} was to find out whether a resource was compromised or not by the attacker.
@@
 The basic idea is to test with an extra action if the resource has been compromised or not.
-This move involves also an extra cost.\\
+This move also involves an extra cost.\\
 A three-player game has also been investigated where the FlipIt framework of two players is extended by another player.
@@
 Nochenson and Grossklags \cite{nochenson2013behavioral}  investigate how people really act when given temporal decisions.
-They found out that the results improves over time but that they are dependent on gender, age, and other individual difference variables.
+They found out that the results improve over time but that they are dependent on gender, age, and other individual difference variables.
 The result also shows that the participants perform generally better when they have more information about the strategy of the opponent which is a computerized player.

### 2015-06-14 doc/template/chap-9.tex
changes na  het nalezen

@@
 \label{Ch2:Intro}
-The game of FlipIt with propagation delay considers attackers where their moves are not instantaneous.
-This can be motivated by an example of a virus.
-A virus can be dropped on a network but it only compromises the whole network if every node in the network is infected.
-The basic FlipIt game does not take this into account.
-In this chapter the FlipIt game with propagation delay is explained and how it can be modelled.
-First the main differences with the basic FlipIt game are discussed in section \ref{ch2:diffFlip}.
-Next in section \ref{ch2:formal} the formal definition of the game with propagation delay is given.
-At last, in section \ref{ch2:periodicvirus} a periodic strategy game is considered and the formalization of the formulas with propagation delay are determined.
-The use of the word virus or worm will be used interchangeably in this paper.
-\todo{EEN van de twee intros nemen, virus term gebruiken ?}\\
-
-This chapter explains how to model a FlipIt game with a propagation delay that infects a network.
-The first section explains the difference between a normal FlipIt game and a FlipIt game with propagation delay.
-The next section derives a formula to calculate the benefit for a FlipIt game with a  propagation delay.
-In the last section we calculate the Nash equilibrium for the benefit formula.
+The game of FlipIt with propagation delay considers a game where the moves of the attacker are not instantaneous.
+This corresponds to the case of APT's where attacks are performed by means of malware, such as viruses, worms and trojan horses and the malware needs time to propagate.
+A virus, for example, can be dropped on a network but it only compromises the whole network if every node in the network is infected.
+So there is a certain delay between the moment of attack and the moment the attacker has control over the resource, called the propagation delay.
+The basic FlipIt game does not take this propagation delay into account.
+In this chapter explains how the FlipIt game with propagation delay can be modelled.
+Section \ref{ch2:diffFlip} explains the difference between a basic FlipIt game and a FlipIt game with propagation delay.
+Next section \ref{ch2:periodicvirus} derives a formula to calculate the benefit for a FlipIt game with propagation delay.
+Section \todo{nash} presents the Nash equilibrium for the benefit formula.
+The chapter concludes in section \todo{section toevoegen} with lessons learned en practical implications for cyber security.
 %
@@
 \label{ch2:diffFlip}
+
+\textbf{Section:Required adaptations to the FlipIt game}
+
 \begin{figure}[hbtp]
+\centering
+\includegraphics[scale=0.7]{Images/dtFlip.png}
 \caption{The first game is the basic FlipIt game.
@@
 }
-\centering
-\includegraphics[scale=0.7]{Images/dtFlip.png}
 \label{dt}
@@
 
-The framework of FLipIt is used to model targeted attacks.
-In FlipIt when an attacker attacks, he will get immediately control over the resource.
-One of the differences is that FlipIt does not consider that moves of the attacker might not be instantaneous.
-With FlipIt with a propagation delay we model the fact that if the attacker flips, he does not get immediate control over the resource.
-This is motivated by the fact that it takes a while before a targeted attack can compromise a computer network.
-The virus or worm has to spread to infect every computer controlled device in the network.
-This propagation make take a while.
-\\
-
+\subsubsection{Only one action per player}
 Another difference is that the FlipIt game consists of a single resource.
@@
 
+The action of the defender consist of defending his network.
+In this context it will be defending every computer controlled device in a network.
+To make the network clean the defender can patch the systems with the latest security bulletins or reinstall the software or even wipe out the computer.
+All these different actions can imply other costs.
+The cost of patching a system will be smaller than the cost of wiping out a computer.
+\\
+The defender has only one action.
+Sending different viruses and worms to a computer network.
+Every time the attacker flips the resource, he does this by sending a new virus to the system.
+The attacker does not send the same virus again, because if the computers are patched, the virus cannot get in any more.
+\subsubsection{Immediate effect of the move}
+The framework of FLipIt is used to model targeted attacks.
+In FlipIt when an attacker attacks, he will get immediately control over the resource.
+One of the differences is that FlipIt does not consider that moves of the attacker might not be instantaneous.
+With FlipIt with a propagation delay we model the fact that if the attacker flips, he does not get immediate control over the resource.
+This is motivated by the fact that it takes a while before a targeted attack can compromise a computer network.
+The virus or worm has to spread to infect every computer controlled device in the network.
+This propagation make take a while.
+\\
 In reality however, after dropping a virus on the first node, it takes a while for the virus to infect
@@
 \\
-
-The action of the defender consist of defending his network.
-In this context it will be defending every computer controlled device in a network.
-To make the network clean the defender can patch the systems with the latest security bulletins or reinstall the software or even wipe out the computer.
-All these different actions can imply other costs.
-The cost of patching a system will be smaller than the cost of wiping out a computer.
-\\
-The defender has only one action.
-Sending different viruses and worms to a computer network.
-Every time the attacker flips the resource, he does this by sending a new virus to the system.
-The attacker does not send the same virus again, because if the computers are patched, the virus cannot get in anymore.
+\subsubsection{Stealth character of the move}
+\subsubsection{Cost associated with the move}
+\subsubsection{Renewal strategies}
+The formalization starts from the model of the non-adaptive continuous basic FlipIt game where players use a periodic strategy with a random phase.
+This choice is motivated by the assumption that in most organisations, the defence strategy is to periodically defend the network.
+This is true for the example of patching.
+Companies like Microsoft en Google release at a specific moment patches.
+So companies will apply these patches when they come out.
+Microsoft security bulletins are released on the Second Tuesday of each month.
+%\source{https://technet.microsoft.com/en-us/security/bulletin/dn602597.aspx} 
+This corresponds to a periodic defender strategy.
+A periodic attacker strategy is assumed as well, to be able to compare the results with the periodic strategy of the FlipIt game in \cite{FlipIt}.
+%, as this also corresponds to a common real life strategy.
+%nog eventueel verder te motiveren
+Further research can investigate the effect of relaxing this assumption. \\
 
@@
 \label{ch2:periodicvirus}
-
-The formalization starts from the model of the non-adaptive continuous basic FlipIt game where players use a periodic strategy with a random phase.
-This choice is motivated by the assumption that in most organisations, the defence strategy is to periodically defend the network.
-This is true for the example of patching.
-Companies like microsoft en google release at a specific moment patches.
-So companies will apply these patches when they come out.
-Microsoft security bulletins are released on the Second Tuesday of each month.
-%\source{https://technet.microsoft.com/en-us/security/bulletin/dn602597.aspx} 
-This corresponds to a periodic defender strategy.
-A periodic attacker strategy is assumed as well, to be able to compare the results with the periodic strategy of the FlipIt game in \cite{FlipIt}.
-%, as this also corresponds to a common real life strategy.
-%nog eventueel verder te motiveren
-Further research can investigate the effect of relaxing this assumption. \\
 

### 2015-06-14 doc/template/chap-9.tex
verbeteringen van thesis verder

@@
 \end{figure}
-
-\subsubsection{Only one action per player}
-Another difference is that the FlipIt game consists of a single resource.
-To represent the security problem of the propagation of a virus in a network, the game now defines its single resource as a computer network with multiple
+\subsubsection{Single resource}
+The basic FlipIt game consists of a single resource.
+To represent the security problem of the propagation of an APT in a network, the game now defines its single resource as a computer network with multiple
 nodes. One of the players, the defender, will try to defend his network. The defender
@@
 Since the original FlipIt game works with a single resource that is always flipped entirely, the assumption is made that the attacker is considered to gain immediate full control over the resource when the network has been infected, even it is only one node that has been infected.\\
+\subsubsection{Only one action per player}
 
@@
 \subsubsection{Stealth character of the move}
+Moves in the basic FlipIt game are stealthy or covert and not immediately detected by the other payer.
+The attacker attacker's moves are stealthy because we want to model a scenario where a computer network is attacker by an APT.
+The main characteristic of an APT is that the attack is stealthy.
+Depending on how the attacker has set up his attack, and depending on the network if it is connected with the Internet or not, the moves of the defender are stealthy or not stealthy.
+For example, if the attacker launches an APT only to harm the system and not to receive feedback, the moves of the defender are stealthy.
+If the attacker launches an APT to steal sensitive information, the moves of the defender are non stealthy because the attacker can see that he doesnt receive information anymore.
+But the defender can also setup a honey pot, making the attacker believe that he is stealing sensitive information, but the information in the honeypot is all fake.
+In thesis we assume that both moves of the attacker and the defender are stealthy.
+-- kunnen enerzijds niet stealthy zijn als the apt informatie steelt depending how he setup his attacker.
+depending on network connected.
+in this a thesis assumed, non stealthy is..
 \subsubsection{Cost associated with the move}
@@
 The formalization starts from the model of the non-adaptive continuous basic FlipIt game where players use a periodic strategy with a random phase.
-This choice is motivated by the assumption that in most organisations, the defence strategy is to periodically defend the network.
+The basic FlipIt game concluded that periodic strategy is the dominant strategy against all renewal strategies.
+\todo{zelfde paper conclusie}.
+Furthermore, this choice is motivated by the assumption that in most organisations, the defence strategy is to periodically defend the network.
 This is true for the example of patching.
@@
 Microsoft security bulletins are released on the Second Tuesday of each month.
-%\source{https://technet.microsoft.com/en-us/security/bulletin/dn602597.aspx} 
+\cite{MicroPatch}
+
 This corresponds to a periodic defender strategy.
@@
 The average rate of play of a player is denoted by $\alpha_{i} = \dfrac{1}{\delta_{i}}$.
-List of symbols that will be used throughout the paper.
-Figure \ref{FlipItDelay} represents a couple of the symbols for clarification.\\
+Stated below a list of symbols that will be used throughout the paper.
+Figure \ref{FlipItDelay} represents the main symbols for clarification.\\
 ~~\\
+\todo{kleuren van de tekening aanpassen zodat ze zichtbaar zijn in een zwartwit paper}
 \begin{figure}[hbtp]
+\centering
+\includegraphics[scale=0.7]{Images/DefFlip.png}
 \caption{Formalization of a FlipIt game with delay: A representation of a FlipIt game where both players are playing periodically.
-Every move or flip is indicated by a blue or orange circle.
-The defender is represented in blue and plays with a period of $\delta_{D}$.
-The flip of the attacker is represented by a white circle, but because there is a delay d, the attacker only controls the resource after time d represented by an orange circle.
+Every move or flip is indicated by a blue or orange circle, respectively dark gray and light grey.
+The defender is represented in blue (dark grey) and plays with a period of $\delta_{D}$.
+The flip of the attacker is represented by a white circle, but because there is a delay d, the attacker only controls the resource after time d represented by an orange circle (light grey).
 The attacker plays with a period of $\delta_{A}$.
 The blue and orange rectangles represent the amount of time the respective player is in control of the resource.}
-\centering
-\includegraphics[scale=0.7]{Images/DefFlip.png}
 \label{FlipItDelay}
@@
 \begin{description}
-\item $i$: Defines the player. Different as in the FlipIt paper where the defender is denoted by the subscript 0 and the attacker by the subscript 1.
+\item $i$: Denotes the player.
+`D' denotes the defender, and `A' denotes the attacker which differs form the notation in \citep{FlipIt}, where the defender is denoted by the subscript 0 and the attacker by the subscript 1.
 \item $\delta_{i}$: The length of the interval between two consecutive moves of player \textit{i}. 
@@
 \item $d$: The delay caused by the virus propagation.
-\item $G_{i}(t)$: The total gain of player \textit{i} denotes the amount of time player \textit{i} is in control over the resource up to time \textit{t}.
-\item $\gamma_{i}$: The average gain rate of player \textit{i} defined as $G_{i}(t)/t$
-\item $\beta_{i}$:  The average benefit rate up to time \textit{t} defined as  $\beta_{i} = \gamma_{i} -k_{i} \alpha_{i} $.
-\item $opt_{i}$: The optimum function.
+\item $G_{i}(t)$: The total gain of player \textit{i}, which is the amount of time player \textit{i} is in control over the resource up to time \textit{t}.
+\item $\gamma_{i}$: The average gain rate of player \textit{i}, defined as $G_{i}(t)/t$
+\item $\beta_{i}$:  The average benefit rate up to time \textit{t}, defined as  $\beta_{i} = \gamma_{i} -k_{i} \alpha_{i} $.
+\item $opt_{i}$: The optimum function for player \textit{i}. \todo{optimum function equal or not with best response ?}
 \end{description}
 
-Assume that an attacker attacks at time \textit{t}, he doesn't get immediate control over the resource, but he only gains control at time \textit{t + d}, with $d$ denoting the time needed to infect a sufficiently number (or all) nodes.
+The adaptation of the FlipIt model starts from the assumption that when an attacker attacks at time \textit{t}, he doesn't get immediate control over the resource, but he only gains control at time \textit{t + d}, with $d$ denoting the time needed to infect a sufficient number of (or all) nodes.
 If the defender flips the network before the period $d$ has elapsed (so, somewhere between $t$ and $t + d$), then the attacker will never gain full control over the resource.
-See figure \ref{dt}.
+(See figure \ref{dt} at point 4).
 This implies that the mathematical formulas for gain and benefit need to be adapted to the fact that the attacker loses part of its benefit because of this delay.
-In the remainder of this paper, we will adapt the formalization of the FlipIt game using the variable $d$.
+In the remainder of this paper, we will adapt the formalization of the FlipIt game using the delay variable $d$.
 \\
@@
 The first case is where the defender plays at least as fast as the attacker, the second case is where the attacker plays at least as fast as the defender.
-For each of these cases, first the benefit formula of the basic case without delay is presented, and then the delay is introduced.
+For each of these cases, first the benefit formula of the basic case without delay is presented, and subsequently the delay is introduced.
 \\
 
-Intuitively we could assume that $d$ can never be bigger than $\delta_{A}$ because then the benefit for the defender would always be 1, \textit{but} this is not always true.
-It is only true if $d$ is bigger than $\delta_{D}$, because then the defender will always be in control.
-For this we only calculate the formulas for the cases where $d$ is smaller than $\delta_{D}$.
+Intuitively we could assume that $d$ can never be bigger than $\delta_{A}$ because then the benefit for the defender would always be 1, but this is not always true.
+Assume for example that an attacker plays with an interval of 3 units, the delay is equal to 4 units and that the defender only plays every 8 units.
+This situation is represented in figure \todo{figuur maken en refereren}.
+Since the delay is shorter that the period of the defender, the attacker takes control of the resource once the delay has elapsed, until the defender plays.
+However, if the delay is larger that the period of the defender, then the defender will always be in control.
+This situation is represented in figure \todo{figuur maken en refereren}, where the attacker plays with an interval of 5 units, the delay is equal to 6 units and the defender plays with an interval of 4 units.
+From this we can conclude that it only makes sence to calculate the formulas for the cases where $d$ is smaller than $\delta_{D}$.
 We can already conclude that it is no use for the attacker to play when the delay is bigger than $\delta_{D}$.
-see figure \ref{delaye}\\
+(see figure \todo{refereren naar de juiste figuur die nog aangemaakt moet worden})\\
 \todo{bovenstaande tekst beter uitleggen}
@@
 
+In the original model of FLipIt \cite{FlipIt}, the authors express the formulas in terms of $\alpha_{D}$ and $\alpha_{A}$.
+However, when introducing the delay, some formulas become much simpler when expressed in terms of $\delta_{D}$ and $\delta_{A}$.
+Therefore in the remainder of this paper, we will formulate the model using both $\alpha_{i}$ and $\delta_{i}$, depending on what result is the simplest formula.
+
 \subsection*{\textbf{Case 1:} $\delta_{D} \leq \delta_{A} $ (The defender plays at least as fast as the attacker.) }
@@
 The probability over the attacker's phase selection that the attacker moves in this interval is r.
-Given that the attacker moves within the interval, he moves exactly once within the interval (since $\delta_{D} \leq \delta_{A} $) and his move is distributed uniformly at random.
+Given that the attacker moves within the interval, he moves exactly once within the interval (since $\delta_{D} \leq \delta_{A} $) and his move is distributed uniformly at random within this interval.
 \\
@@
 \begin{equation*}
-\beta_{A}(\alpha_{D},\alpha_{A}) =\dfrac {r} {2} - k_{A} \alpha_{A} = \dfrac {\delta_{D}} {2\delta_{A}} - k_{A} \alpha_{A}  
+\beta_{A}(\delta_{D},\delta_{A}) =\dfrac {r} {2} - k_{A} \alpha_{A} = \dfrac {\delta_{D}} {2\delta_{A}} - k_{A} \alpha_{A}  
 \end{equation*}\\
@@
 \begin{equation*}
-\beta_{D}(\alpha_{D},\alpha_{A}) =1 -  \dfrac {r} {2} - k_{D} \alpha_{D} = 1 - \dfrac {\delta_{D}} {2\delta_{A}} - k_{D} \alpha_{D} 
+\beta_{D}(\delta_{D},\delta_{A}) =1 -  \dfrac {r} {2} - k_{D} \alpha_{D} = 1 - \dfrac {\delta_{D}} {2\delta_{A}} - k_{D} \alpha_{D} 
 \end{equation*}
@@
 However, because of the delay required for virus propagation, the maximal time of control is reduced to $\delta_{D}-d$ , see figure \ref{fig:delaycase1}.
-There is a probability of \textit{r} that the attacker will move in the interval of the defender.
-However, the gain will not be half of the interval.
-Indeed, the attacker has to play soon enough to gain control, meaning that the attacker has to play during the period of $\delta_{D}-d$ during the interval of the defender.
+While the probability of \textit{r} that the attacker will move in the interval of the defender is still \textit{r}, the gain will not be half of the interval.
+Indeed, if the attacker plays after $\delta_{D}-d$, given the delay \textit{d}, he will never get in control in that interval (see figure \todo{figure 3.5}).
 The probability that the attacker plays soon enough is $\dfrac{\delta_{D}-d}{\delta_{D}}$ and this will give the attacker an average gain of $\dfrac{\delta_{D}-d}{2}$.
@@
 The probability that this happens is  $\dfrac{d}{\delta_{D}}$.
-The average gain rate of the attacker can then be expressed as follows if we look at one interval of the defender:
+Looking at one interval of the defender, the average gain rate of the attacker can thus be expressed as follows:
 \begin{equation*}
-\gamma_{A}(\alpha_{D},\alpha_{A}) = \dfrac {1}{\delta_{D}} [ \dfrac{\delta_{D}}{\delta_{A}} \cdot \dfrac{\delta_{D}-d}{\delta_{D}} \cdot \dfrac{\delta_{D}-d}{2} + \dfrac{\delta_{D}}{\delta_{A}} \cdot \dfrac{d}{\delta_{D}} \cdot 0 ]
+\gamma_{A}(\delta_{D},\delta_{A}) = \dfrac {1}{\delta_{D}} [ \dfrac{\delta_{D}}{\delta_{A}} \cdot \big[ \dfrac{\delta_{D}-d}{\delta_{D}} \cdot \dfrac{\delta_{D}-d}{2} + \dfrac{d}{\delta_{D}} \cdot 0 \big] ]
+\end{equation*}
+Thus over the entire game, the average gain rate is:
+\begin{equation*}
+\gamma_{A}(\delta_{D},\delta_{A}) = \dfrac{(\delta_{D} -d)^{2}}{2\delta_{D}\delta_{A}}
 \end{equation*}
@@
 \begin{equation*}
-\beta_{A}(\alpha_{D},\alpha_{A}) = \dfrac { (\delta_{D}-d) ^{2}} {2 \cdot \delta_{D}  \delta_{A}} - k_{A} \alpha_{A}
-\end{equation*}
-\begin{equation*}
-\beta_{A}(\alpha_{D},\alpha_{A}) = \dfrac { \delta_{D}} {2 \cdot \delta_{A}} - k_{A} \alpha_{A} - ( \dfrac{d^{2}}{2 \cdot \delta_{A} \delta_{D}} - \dfrac{d}{\delta_{A}} )
+\beta_{A}(\delta_{D},\delta_{A}) = \dfrac { (\delta_{D}-d) ^{2}} {2 \delta_{D}  \delta_{A}} - \dfrac{k_{A}}{\delta_{A}}
 \end{equation*}
 
- The benefit of the defender is expressed as follows:
+ The benefit of the defender is then as follows:
 \begin{equation*}
-\beta_{D}(\alpha_{D},\alpha_{A}) = 1 - \dfrac { (\delta_{D}-d) ^{2}} {2 \cdot \delta_{D}  \delta_{A}} - k_{D} \alpha_{D}
+\beta_{D}(\delta_{D},\delta_{A}) = 1 - \dfrac { (\delta_{D}-d) ^{2}} {2 \cdot \delta_{D}  \delta_{A}} - \dfrac{k_{D}}{ \delta_{D}}
 \end{equation*}
@@
 \begin{equation*}
-\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac {1} {2r} - k_{D} \alpha_{D} = \dfrac {\delta_{A}} {2\delta_{D}} - k_{D} \alpha_{D} 
+\beta_{D}(\delta_{D},\delta_{A}) = \dfrac {1} {2r} - k_{D} \alpha_{D} = \dfrac {\delta_{A}} {2\delta_{D}} - \dfrac{k_{D} }{\delta_{D}} 
 \end{equation*}
 \begin{equation*}
-\beta_{A}(\alpha_{D},\alpha_{A}) =1 - \dfrac {1} {2r} - k_{A} \alpha_{A} = 1- \dfrac {\delta_{A}} {2\delta_{D}} - k_{A} \alpha_{A}  
+\beta_{A}(\delta_{D},\delta_{A}) =1 - \dfrac {1} {2r} - k_{A} \alpha_{A} = 1- \dfrac {\delta_{A}} {2\delta_{D}} - \dfrac{k_{A}}{ \delta_{A}}  
 \end{equation*}\\
@@
 An intuitive solution for the case with a virus would be to subtract the benefit of the attacker received in each interval with the delay similarly as in case 1.
-This would give us the following formula if we derive it in the same way:
+This would yield the following formula:
 \begin{equation*}
-\beta_{A}=\dfrac{(\delta_{A} - d)^2}{2\delta_{A}\delta_{D}} - \dfrac{k_{A}}{\delta_{A}}
+\beta_{A}(\delta_{D},\delta_{A})=\dfrac{(\delta_{A} - d)^2}{2\delta_{A}\delta_{D}} - \dfrac{k_{D}}{\delta_{A}}
 \end{equation*}
-
-But this results in an overestimation.
-How closer $\delta_{A}/\delta_{D}$ is equal to one, the better the approximation.
+\todo{formule nog eens nakijken}
+This however results in an overestimation.
+By simulation the game, it can be notices that the closer $\delta_{A}/\delta_{D}$ is equal to one, the better the approximation.
 If $\delta_{A}/\delta_{D} = 1$ the result is correct.
-This formula does not take into account that if the attacker was in control in the previous interval, the delay should not be subtracted, because the delay will not be in control of the defender.
-This means that we have to look at what happens in the previous interval.
+The reason this formula overestimates the benefit of the attacker is that it assumes that the defender is always in control during the delay.
+However, if the attacker was in control in the previous interval, then he continuous to be in control during the period of the delay, see figure \todo{figuur weer toevoegen}.
+This means that to find the correct formula, what happens in the previous interval needs to be taken into account.
 \\
 
-From the defender we know that his moves are instantaneous.
-It is easier to calculate the benefit of the defender in this case.
+We know that the defender's moves are instantaneous.
+Therefore, it is easier in this case to calculate the benefit of the defender.
 Because the defender moves slower than the attacker we know that if the defender moves during the interval of the attacker, he only moves once within this interval.
 The defender will move during the interval of the attacker with a probability of $\dfrac{\delta_{A}}{\delta_{D}} $.
-When this happens the defender will end with being in control at the end of the interval.
-In the next interval the attacker will have to regain control, meaning that during the delay, the defender stays in control, see figure \ref{fig:case2} cases (1) and (2).
+When this happens, the defender will be in control for the remainder of this interval.
+In the next interval the attacker will have to regain control, meaning that during the delay, the defender stays in control, see figure \ref{fig:case2} cases (1) and \todo{through4} (2).
 The defender will keep the control over the resource in the next interval over a period of the delay, namely \textit{d}.
@@
 Consider a timespan $\delta_{A} + d$, representing the attacker's interval followed by the delay period in his next interval.
-The defender will never move twice during this timespan because $\delta_{A} + d \leq \delta_{D}$.
+If we assume that $\delta_{A}+d<\delta_{D}$, we can infer that the defender will never move twice during this timespan.
 Because $d + \delta_{A} \leq \delta_{D}$ the next move of the defender in this second interval will never occur during the delay, meaning that the entire delay can be considered as an extra benefit resulting of a play in the previous interval.

### 2015-06-14 doc/template/chap-9.tex
alpha -> delta

@@
 So, every time the defender plays, he will get an average gain of $\dfrac{\delta_{A}}{2}$ in the interval where he plays and in the next interval will always receive a extra gain of $d$, yielding a total average gain per interval of
-$\dfrac{(d+\dfrac{\delta_{A}}{2})}{\delta_{A}}$
+$\dfrac{(d+\dfrac{\delta_{A}}{2})}{\delta_{A}}$ \todo{in deze tekst moeten nog een aantal zaken veranderen}
 
@@
 \subsubsection*{\textbf{Case a:} $d + \delta_{A} \leq \delta_{D}$}
-The first case is where the delay is shorter than the difference between the attacker's and the defender's period so their is no problem of counting the delay twice.
+The first case is where the defender's period is larger than the sum of the delay and the attacker's period.
+In that case there is no problem of counting the delay twice.
 To determine the total gain rate of the defender we need to know what the probability is that the defender will move during an interval and what the average time is that the defender controls the resource.
-\todo{beter verwoorden}.
+Given that if the defender moves he will always benefit entirely from a period of delay in the next interval of the attacker, his total gain is $\dfrac{(d+\dfrac{\delta_{A}}{2})}{\delta_{A}}$ in one interval.
 The total gain  rate of the defender is then the probability that the defender will move during an interval of the attacker multiplied by the total average gain per interval:
@@
 \begin{equation*}\label{first}
-\gamma_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta_{A}}{\delta_{D}} \cdot \dfrac{(d+\dfrac{\delta_{A}}{2})}{\delta_{A}} 
+\gamma_{D}(\delta_{D},\delta_{A}) = \dfrac{\delta_{A}}{\delta_{D}} \cdot \dfrac{(d+\dfrac{\delta_{A}}{2})}{\delta_{A}} 
 \end{equation*}
 \begin{equation*}\label{first}
-\gamma_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} 
+\gamma_{D}(\delta_{D},\delta_{A}) = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} 
 \end{equation*}\\
@@
 \begin{equation*}\label{first}
-\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - k_{D} \alpha_{D} 
+\beta_{D}(\delta_{D},\delta_{A}) = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - k_{D} \alpha_{D} 
 \end{equation*}\\
@@
 \begin{equation*}\label{first}
-\beta_{A}(\alpha_{D},\alpha_{A}) = 1 -\dfrac{\delta_{A}}{2\delta_{D}} - \dfrac{d}{\delta_{D}} - k_{A} \alpha_{A} 
+\beta_{A}(\delta_{D},\delta_{A}) = 1 -\dfrac{\delta_{A}}{2\delta_{D}} - \dfrac{d}{\delta_{D}} - \dfrac{k_{A}}{ \delta_{A}} 
 \end{equation*}\\
@@
 \begin{figure}[hbtp]
-\caption{All possible cases for the attacker and the defender in Case 2.A where $d + delta_{A} < delta_{D}$.
-As can be seen, the defender will have control during a period of \textit{d} over the resource when the defender has flipped in the previous interval.}
+\caption{All possible cases for the attacker and the defender in Case 2.A where $d + \delta_{A} < \delta_{D}$.
+As can be seen in case (1) to (4), the defender will have control during a period of \textit{d} over the resource in the next interval when the defender has flipped in the previous interval.}
 \centering
@@
 If not, this would mean that the defender can move during the delay in the interval following the interval where the defender already moved.
-This would mean that there can be an overlap between the average gain of $\dfrac{\delta_{A}}{2}$ and the delay.
+This would mean that there can be an overlap between the average gain of $\dfrac{\delta_{A}}{2}+d$ and the delay.
 The above benefit formula would then include to much gain for the defender: the potential overlap during the delay would be counted twice.
-See figure \ref{overlap}\\
+See figure \ref{overlap}.\\
 \begin{figure}[hbtp]
@@
 \includegraphics[scale=0.7]{Images/FlipItcase2b.png}
+\label{overlap}
 \end{figure}
@@
  \begin{equation*}
-\gamma_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta_{A}}{\delta_{D}} \cdot \dfrac{(d+\dfrac{\delta_{A}}{2})}{\delta_{A}} - \dfrac{(d - (\delta_{D} - \delta_{A}))^{2}}{2 \delta_{D} \delta_{A}}
+\gamma_{D}(\delta_{D},\delta_{A}) = \dfrac{\delta_{A}}{\delta_{D}} \cdot \dfrac{(d+\dfrac{\delta_{A}}{2})}{\delta_{A}} - \dfrac{(d - (\delta_{D} - \delta_{A}))^{2}}{2 \delta_{D} \delta_{A}}
 \end{equation*}
 \begin{equation*}
-\gamma_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - \dfrac{(d - (\delta_{D} - \delta_{A}))^{2}}{2 \delta_{D} \delta_{A}}
+\gamma_{D}(\delta_{D},\delta_{A}) = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - \dfrac{(d - (\delta_{D} - \delta_{A}))^{2}}{2 \delta_{D} \delta_{A}}
 \end{equation*}\\
@@
 \begin{equation*}
-\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - k_{D} \alpha_{D} - \dfrac{(d - (\delta_{D} - \delta_{A}))^{2}}{2 \delta_{D} \delta_{A}}
+\beta_{D}(\delta_{D},\delta_{A}) = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - \dfrac{k_{D} }{\delta_{D}} - \dfrac{(d - (\delta_{D} - \delta_{A}))^{2}}{2 \delta_{D} \delta_{A}}
 \end{equation*}\\
@@
 \begin{equation*}
-\beta_{A}(\alpha_{D},\alpha_{A}) = 1 -\dfrac{\delta_{A}}{2\delta_{D}} - \dfrac{d}{\delta_{D}} - k_{A} \alpha_{A} + \dfrac{(d - (\delta_{D} - \delta_{A}))^{2}}{2 \delta_{D} \delta_{A}}
+\beta_{A}(\delta_{D},\delta_{A}) = 1 -\dfrac{\delta_{A}}{2\delta_{D}} - \dfrac{d}{\delta_{D}} - \dfrac{k_{A}}{ \delta_{A}} + \dfrac{(d - (\delta_{D} - \delta_{A}))^{2}}{2 \delta_{D} \delta_{A}}
 \end{equation*}\\

### 2015-06-15 doc/template/chap-10.tex
nog verdere verbeteringen

@@
 It is common to model cyber security games as a game with two players, an attacker and a defender.
-However, there are games that have more players \cite{fengstealthy}.
+However, there are games that have more players, for example when there is a third party involved \cite{fengstealthy}.
 In this paper we focus on the game with two players.
@@
 
- Researchers at RSA, van Dijk et al,  presented a game-theoretic framework to model computer security scenarios called ``FlipIt'' \cite{FlipIt}.
+Many security games bridging game theory and cyber security have already been investigated and finding a new game is not convenient.
+In this paper we choose to extend another quite relatively new paper where the assumption of stealthiness is fairly unique and the paper had some interesting results.\\
+ The paper is from researchers at RSA, van Dijk et al,  who presented a game-theoretic framework to model computer security scenarios called ``FlipIt'' \cite{FlipIt}.
 They study the specific scenario where a system or network is repeatedly taken over completely by an attacker and not immediately being detected by the defender of the system or network.
@@
 Neither player knows who is currently in control of the resource until they move.
-In FlipIt every move gives them immediate control over the resource.
+In FlipIt every move involves a cost and gives them immediate control over the resource.
+The attacker will try to maximise the time that he has control over the network, while the defender will try to maximise the time that the network is free of malware.
 \\
@@
 \begin{itemize}
-\item How can we incorporate the notion of delay in the game-theoretical analysis of the Flip-It game for a periodic strategy ?
-\item Does the resulting model allow an optimal defence strategy against an attacker ? 
+\item How can we incorporate the notion of delay in the game-theoretical analysis of the Flip-It game for a periodic strategy?
+\item Does the resulting model allow an optimal defence strategy against an attacker? 
 \end{itemize}
@@
 We extend the FlipIt game where the attacker flips with a delay.
-The attacker only compromises the system if suffi"ent nodes in the network are infected.
-\item[-] Another derivation of the method to calculate the benefit of the basic FlipIt game is given.
+The attacker only compromises the system if sufficient nodes in the network are infected.
 \item[-] The periodic case of FlipIt is modelled with a delay, resulting in optimum functions and Nash equilibria. 
 \item[-] Based on our results we can give practical recommendations to take measures against advanced attacks.
+The modelling of FlipIt with a propagation delay will give a guideline for both the attacker and the defender in when they should attack the network or how regularly should they clean the network.
+\item[--] Related work is added about different techniques to calculate propagation delays of malware depending on the graph. 
 \end{itemize}
- Next we analyse what the new Nash equilibria will be and .. \\
- The modelling of FlipIt with a delay will give a guideline for both the attacker and the defender in when they should attack the network or how regularly should they clean the network.
 
-While it seems maybe trivial to extend FlipIt with a delay it is not so straightforward.
-First a formula to calculate the benefit of a basic FlipIt was derived to introduce the delay.
-To calculate the benefit for each case \todo{cases misschien uitleggen?}, the intervals where not independant anymore as in the basic FlipIt game.
-Er moest rekening gehouden worden met de vorige intervallen om de benefit te berekenen in het eigen interval.
+While it seems maybe trivial to extend FlipIt with a propagation delay it is not so straightforward.
+Even though the basic model of FlipIt looks simple, it's mathematical treatment is not.
+In the paper of Laznka et all.
+\citep{FlipThem} it seems that even a small extension adds to the already significant mathematical complexity.
+While the FlipIt game is quite symmetric, by adding a propagation delay the mathematical complexity rises.
+First a formula to calculate the benefit of a basic FlipIt was derived to introduce the propagation delay.
+To calculate the benefit for each case depending on the periods of the attacker, defender and the delay, the intervals where not independent any more as in the basic FlipIt game.
+The previous intervals had to be taken into account to calculate the benefit of the players in its own interval.
 
@@
 \item[-] Give the defender a delay. It takes a while before a patch is made after the discovery of a vulnerability.
-\item[-]
 \end{description}

### 2015-06-15 doc/template/chap-5.tex
nog verdere verbeteringen

@@
 While cyber security and network security address different aspects, they partially overlap as well.
-For example a worm can be physically dropped on a computer network using a USB stick, but it can also arrive over the internet.
+For example a virus can be physically dropped on a computer network using a USB stick, but it can also arrive over the internet.
 Either way the internal computer security measurements have to be taken to recover from the virus.
@@
 `Persistent' refers to the fact that the attacker never quits trying to attack his target. The attack can be spread over several years, taking multiple steps.  \\
-An APT can be a mix of different types of malware and may use various propagation methods. \\
+An APT can be a mix of different types of malware and may use various propagation methods like the way a virus, worm or trojan horse propagates. \\
 %
@@
 To mitigate these kind of attacks another security countermeasure is needed.
-This is where the researchers of RSA come into the picture.
-They have modelled the situation where targeted attacks are taking over a system and are not immediately detected by the defender.
+\\
+%This is where the researchers of RSA come into the picture. \todo{deze paper gaat zien of de game theoretische model kan leiden tot resultaten die kunnen helpen in het verdedigen tegen APTs}
+In this paper we will propose an appropriate game theoretical modelling of defending a network against APTs and analyse it in order to draw the necessary strategic conclusions to mitigate these kind of attacks.
 
@@
 %Cooperative and non cooperative games
-In a Decision Game Theoretic Approach an agent will try to act in such a way to maximise his expected or average utility function.
+In a decision game theoretic approach an agent will try to act in such a way to maximise his expected or average utility function.
 It becomes more complicated when two or more agents want to maximise their utility and when actions of the agents can affect each other's utilities.
@@
 We define $Opt_{i}$ as the best response function for player \textit{i}.
-The best response for player \textit{1} is given by : $a_{1} = Opt_{1}(a_{2})$.
+The best response for player \textit{1} is: $a_{1} = Opt_{1}(a_{2})$, given that $a_{2}$ is an action of player 2.
 A
-For a Nash Equilibrium each player has a consist list of actions and each player's action maximizes his or her pay-off given the actions of the other players.
+For a Nash Equilibrium each player has a list of actions and each player's action maximizes his or her pay-off given the actions of the other players.
 Nobody has the incentive to change his or her action if an equilibrium profile is played.
-We have a Nash Equilibrium for the pair $(a_{1}^{*},a_{2}^{*})$ when $a_{1}^{*} = Opt_{1}(a_{2}^{*})$ and $a_{2}^{*} = Opt_{2}(a_{1}^{*})$\\
+We have a Nash Equilibrium for the pair $(a_{1}^{*},a_{2}^{*})$ when $a_{1}^{*} = Opt_{1}(a_{2}^{*})$ and $a_{2}^{*} = Opt_{2}(a_{1}^{*})$.\\
 
@@
 If prisoner 1 talks and the other prisoner remains silent, prisoner 1 will be free.
-If prisoner 1 remains silent and the other prisoner talks, prisoner 1 gets 5 years in prison.
+If prisoner 1 remains silent and the other prisoner talks, prisoner 1 gets five years in prison.
 If the prisoners cooperate and talk, both of them get three years in prison.
@@
 A dominant strategy is a strategy that is better than all other strategies regardless of what the opponent does.
-The Nash equilibrium of the game is that they both confess even though it would be better if both prisoners cooperated and chose to stay silent.
+The Nash equilibrium of the game is that they both confess even though it would be better if both prisoners cooperated and choose to stay silent.
 
@@
 Each can confess or stay silent.
-The outcomes are shown below each prisoner.
+The resulting sentence is shown below each prisoner.
 }
@@
 \end{figure}
-
-\todo{optimal function same as best response ?}
 
@@
 
-FlipIt is a two-players game with a shared single resource that the players want to control as much as possible, see \ref{fig:FLipItDefault}.
+FlipIt is a two-players game with a shared single resource that the players want to control as much as possible (figure \ref{fig:FLipItDefault}).
 The shared resource can be a password, a network or a secret key depending on the setting being modelled.
@@
 
-In this paper we restrict ourselves to periodic strategies.
-This choice is motivated by the fact that in a security game a player (defender or attacker) rarely has information about the moves (last move or full history) of his opponent.
-\\
-
  \begin{table}
@@
 
+ In this paper we restrict ourselves to non-adaptive periodic strategies.
+This choice is motivated by the fact that in a security game a player (defender or attacker) rarely has information about the moves (last move or full history) of his opponent.
+The results also indicate that the periodic strategy might be the dominant strategy.
+\\
+\todo{dit stuk tekst hierboven meer motiveren}
+
 \section{Extensions on FlipIt}
@@
 \cite{MitigationNonTargeted} is to consider a game with targeted and non-targeted attacks where the moves made by the attacker do not succeed immediately.
-This is similar to this paper but it has still some major differences.
-First the moves by the attacker are still covert but the moves made by the defender are known to the attacker.
+This approach is similar to what will be done in this paper, but nevertheless has some major difference.
+In Laszka's paper, firstly the moves by the attacker are still covert but the moves made by the defender are known to the attacker.
 This means that the attacker knows when the defender plays and can change its strategy depending on the moves of the defender.
-Our motivation for a defender with stealthy moves is that not every APT is given feedback to one place.
+Our motivation for a defender with stealthy moves is that it can not be assumed that every attacker can receive feedback from an APT.
 Some ATP's are designed only to cause harm on the systems that are infected e.g.
 the Stuxnetworm that was developed to destroy nuclear reactors.
-The only way that the attacker got feedback was through the news.
-There purpose is to spread and cause harm.
-Another example are APT's that get onto isolated networks through usb stick or cd's.
-There is no way for the worm to communicate back on an isolated network.
+The Stuxnetworm was resident on a isolated network, meaning that there was no way to connect to the internet.
+Another example are APT's that get into honeypots.
+The attacker thinks that he is getting sensitive information, but actually this information is fake \todo{zien dat deze tekst goed geschreven is}.
 The second difference is that even though both the targeted and non-targeted attacks do not succeed immediately, the delay is determined differently.

### 2015-06-15 doc/template/chap-9.tex
nog verdere verbeteringen

--- 2015-06-14 doc/template/chap-9.tex
@@
 The basic FlipIt game does not take this propagation delay into account.
-In this chapter explains how the FlipIt game with propagation delay can be modelled.
+This chapter explains how the FlipIt game with propagation delay can be modelled.
 Section \ref{ch2:diffFlip} explains the difference between a basic FlipIt game and a FlipIt game with propagation delay.
-Next section \ref{ch2:periodicvirus} derives a formula to calculate the benefit for a FlipIt game with propagation delay.
-Section \todo{nash} presents the Nash equilibrium for the benefit formula.
-The chapter concludes in section \todo{section toevoegen} with lessons learned en practical implications for cyber security.
+The last section \ref{ch2:periodicvirus} derives a formula to calculate the benefit for a FlipIt game with propagation delay.
+%Section \todo{nash} presents the Nash equilibrium for the benefit formula. The chapter concludes in section \todo{section toevoegen} with lessons learned en practical implications for cyber security.
 %

### 2015-06-15 doc/template/chap-9.tex
nash aanvullen
figuren zitten aanpassen

@@
 
-\begin{figure}[hbtp]
-\centering
-\includegraphics[scale=0.7]{Images/dtFlip.png}
-\caption{The first game is the basic FlipIt game.
-The second is a FlipIt game with a delay.
-During the first flip of the attacker, the defender moves after the delay, causing the attacker to get control over the resource.
-During the second flip of the attacker, the defender flips at time t+d, causing the defender to take control over the resource before the attacker.
-During the third and final flip of the attacker, the defender flips during time t+d, causing the attacker to never gain control over the resource.
-}
-\label{dt}
-\end{figure}
 \subsubsection{Single resource}
@@
 \centering
-\includegraphics[scale=0.7]{Images/DefFlip.png}
+\includegraphics[scale=0.7]{Images/DefFlip.pdf}
 \caption{Formalization of a FlipIt game with delay: A representation of a FlipIt game where both players are playing periodically.
@@
 If the defender flips the network before the period $d$ has elapsed (so, somewhere between $t$ and $t + d$), then the attacker will never gain full control over the resource.
-(See figure \ref{dt} at point 4).
+(See figure \ref{dt} at point 3 and 4).
 This implies that the mathematical formulas for gain and benefit need to be adapted to the fact that the attacker loses part of its benefit because of this delay.
@@
 \\
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[scale=0.7]{Images/Delayuitgelegd.pdf}
+\caption{The first game is the basic FlipIt game.
+The second is a FlipIt game with a delay.
+During the first flip of the attacker, the defender moves after the delay, causing the attacker to get control over the resource.
+During the second flip of the attacker, the defender flips at time t+d, causing the defender to take control over the resource before the attacker.
+During the third and final flip of the attacker, the defender flips during time t+d, causing the attacker to never gain control over the resource.
+}
+\label{dt}
+\end{figure}
 
@@
 
-Intuitively we could assume that $d$ can never be bigger than $\delta_{A}$ because then the benefit for the defender would always be 1, but this is not always true.
+Intuitively, we could assume that $d$ can never be bigger than $\delta_{A}$ because then the benefit for the defender would always be 1, but this is not always true.
 Assume for example that an attacker plays with an interval of 3 units, the delay is equal to 4 units and that the defender only plays every 8 units.
-This situation is represented in figure \todo{figuur maken en refereren}.
+This situation is represented in figure \ref{langedelay}.
 Since the delay is shorter that the period of the defender, the attacker takes control of the resource once the delay has elapsed, until the defender plays.
 However, if the delay is larger that the period of the defender, then the defender will always be in control.
-This situation is represented in figure \todo{figuur maken en refereren}, where the attacker plays with an interval of 5 units, the delay is equal to 6 units and the defender plays with an interval of 4 units.
-From this we can conclude that it only makes sence to calculate the formulas for the cases where $d$ is smaller than $\delta_{D}$.
+This situation is represented in figure \ref{langeredelay}, where the attacker plays with an interval of 5 units, the delay is equal to 6 units and the defender plays with an interval of 4 units.
+From this we can conclude that it only makes sense to calculate the formulas for the cases where $d$ is smaller than $\delta_{D}$.
 We can already conclude that it is no use for the attacker to play when the delay is bigger than $\delta_{D}$.
-(see figure \todo{refereren naar de juiste figuur die nog aangemaakt moet worden})\\
-\todo{bovenstaande tekst beter uitleggen}
 
-\todo{figuur juist maken}
 \begin{figure}[hbtp]
-\caption{Delay bigger than deltaA and delay bigger than deltaD}
 \centering
-\includegraphics[scale=0.7]{Images/DiffDelayCase1.png}
-\label{delaye}
+\includegraphics[scale=0.7]{Images/FlipItCase1delay.pdf} 
+\caption{FlipIt with delay propagation where $\delta_{D} > d > \delta_{A}$.   }
+\label{langedelay}
 \end{figure}
 
-In the original model of FLipIt \cite{FlipIt}, the authors express the formulas in terms of $\alpha_{D}$ and $\alpha_{A}$.
+\begin{figure}[hbtp]
+\centering
+\includegraphics[scale=0.7]{Images/FlipItCase1delaytobig.pdf} 
+\caption{FlipIt with delay propagation where $ d > \delta_{D}$ }
+\label{langeredelay}
+\end{figure}
+
+In the original model of FlipIt \cite{FlipIt}, the authors express the formulas in terms of $\alpha_{D}$ and $\alpha_{A}$.
 However, when introducing the delay, some formulas become much simpler when expressed in terms of $\delta_{D}$ and $\delta_{A}$.
@@
 \begin{figure}[hbtp]
+\centering
+\includegraphics[scale=0.7]{../../doc/template/Images/DiffDelayCase1.pdf}
 \caption{Case 1: Difference between a basic FlipIt game and a FlipIt game with a delay.
-(1) is the FlipIt game without a delay and (2) is with a delay.
+Case (1) is the FlipIt game without a propagation delay and case (2) is with a propagation delay.
 The delay is denoted with an arrow.
-The attacker is only in control when the circle becomes orange.}
-\centering
-\includegraphics[scale=0.6]{../../doc/template/Images/DiffDelayCase1.png}
+The attacker is only in control when the circle becomes orange (light grey).}
 \label{fig:delaycase1}
@@
 However, because of the delay required for virus propagation, the maximal time of control is reduced to $\delta_{D}-d$ , see figure \ref{fig:delaycase1}.
-While the probability of \textit{r} that the attacker will move in the interval of the defender is still \textit{r}, the gain will not be half of the interval.
+While the probability that the attacker will move in the interval of the defender is still \textit{r}, the gain will not be half of the interval.
 Indeed, if the attacker plays after $\delta_{D}-d$, given the delay \textit{d}, he will never get in control in that interval (see figure \todo{figure 3.5}).
@@
 \centering
-\includegraphics[scale=0.7]{../../doc/template/Images/FlipItCase2a.png}
+\includegraphics[scale=0.7]{../../doc/template/Images/FlipIt2.pdf}
 \label{fig:case2}
@@
 \centering
-\includegraphics[scale=0.7]{Images/FlipItcase2b.png}
+\includegraphics[scale=0.7]{Images/flipcase2nb.pdf}
 \end{figure}

### 2015-06-15 doc/template/chap-10.tex
aanpassingen jeroen

@@
 Why is it so important to keep a system secure?
-Many businesses store confidential information on clients, which can be lost through data leakage and possible be abused by competitors.
+Many businesses store confidential information on clients, which can be lost through data leakage and can possibly be abused by competitors.
 Also, disruption caused by distributed denial of service (DDoS) attacks, may results in businesses failing to meet their service-level agreements.
@@
 A particular kind of threat is an Advanced Persistent Threat (APT).
-An APT is a multi-faceted, continuous and targeted cyber attack that is designed to penetrate a network or a system in a stealthy way and can stay undetected for a long period.
-It is something different and stronger than a conventional threat.
+An APT is a multi-faceted, continuous and targeted cyber attack that is designed to penetrate a network or a system in a stealthy way and can stay undetected for a long period of time.
+It is different and stronger than a conventional threat.
 A conventional threat will not attack any particular target.
@@
 %Since it is so difficult to protect a system or a network against APT's, researchers have been looking for effective ways to predict in advance which defence strategy might be the better one. 
-There are a number of key strategies that an organisation can apply to defend themselves against APT's: awareness, whitelisting, system administration, network segregation, dynamic content checking and patch management.
-Nevertheless the combination of all these elements benefits from being complemented by other defence strategies to protect one selves against stealthy takeovers.
-A way to study the impact of stealthy takeovers and to determine practical recommendations for defenders is through game theory.\\
+There are a number of key strategies an organisation can apply to defend themselves against APT's: awareness, whitelisting, system administration, network segregation, dynamic content checking and patch management.
+Nevertheless the combination of all these elements benefit from being complemented by other defence strategies to protect one selves against stealthy takeovers.
+One possible way to study the impact of stealthy takeovers and to determine practical recommendations for defenders is through game theory.\\
 Game theory is gaining increasing interest as an effective technique to model and study cyber security problems.
-It is common to model cyber security games as a game with two players, an attacker and a defender.
-However, there are games that have more players, for example when there is a third party involved \cite{fengstealthy}.
-In this paper we focus on the game with two players.
-The actions available to the attacker and the defender correspond to respectively to the attacks on the system and taking defensive measures that protect the system.
+It is common to model cyber security problems as a game with two players, an attacker and a defender.
+There are, however, games that have more players e.g.
+when a third party is involved \cite{fengstealthy}.
+This paper focusses on a game with two players.
+The actions available to the attacker and the defender correspond respectively to the attacks on the system and taking defensive measures that protect the system.
 \\
 
-Many security games bridging game theory and cyber security have already been investigated and finding a new game is not convenient.
-In this paper we choose to extend another quite relatively new paper where the assumption of stealthiness is fairly unique and the paper had some interesting results.\\
+Many security games that bridge game theory and cyber security have already been investigated, so finding a new game can be challenging.
+This paper builds on a relatively new paper where the assumption of stealthiness is fairly unique, giving some interesting results.\\
  The paper is from researchers at RSA, van Dijk et al,  who presented a game-theoretic framework to model computer security scenarios called ``FlipIt'' \cite{FlipIt}.
-They study the specific scenario where a system or network is repeatedly taken over completely by an attacker and not immediately being detected by the defender of the system or network.
-This is a two-player game where the attacker and the defender are competing to get control over a shared resource.
+They study the specific scenario where a system or network is repeatedly taken over completely by an attacker.
+This take-over is not immediately detected by the defender.
+It is a two-player game where the attacker and the defender are competing to get control over a shared resource.
 Neither player knows who is currently in control of the resource until they move.
 In FlipIt every move involves a cost and gives them immediate control over the resource.
-The attacker will try to maximise the time that he has control over the network, while the defender will try to maximise the time that the network is free of malware.
+The attacker will try to maximise the time that he controls the network, while the defender will try to maximise the time that the network is free of malware.
 \\
- But what if the attacker moves and it takes a while before the attacker gets full control over the resource?
+ But what if the attacker moves and it takes some time before the attacker gets full control over the resource?
 FlipIt does not take into account that a move may not be instantaneous, but has a certain delay.
 Consider for example a network with different nodes (laptops, datacenters) as a resource.
-The attacker drops a virus on one of the nodes and then wait till this virus infects the whole network.
+The attacker drops a virus on one of the nodes and waits until this virus infects the whole network.
 The attacker will only be in control of the resource when a sufficient large amount of nodes of the network are infected.
-In this paper we present FlipIt with propagation delay to model a game where the moves of the attacker are not instantaneous.
-A formalization for this game starts from the model of non-adaptive continuous basic FlipIt game where players use a periodic strategy with a random phase.
+In this paper we present an adaptation of FlipIt to model a game where the moves of the attacker are not instantaneous.
+The formalization for this game starts from the model of non-adaptive continuous basic FlipIt game where players use a periodic strategy with a random phase.
 \\
@@
 \item[-] We propose an addition to the basic FlipIt model to model a scenario where the moves by the attacker will not be instantaneous.
-We extend the FlipIt game where the attacker flips with a delay.
+We extend the FlipIt game to a game wherein the attacker flips with a delay.
 The attacker only compromises the system if sufficient nodes in the network are infected.
@@
 \item[-] Based on our results we can give practical recommendations to take measures against advanced attacks.
-The modelling of FlipIt with a propagation delay will give a guideline for both the attacker and the defender in when they should attack the network or how regularly should they clean the network.
-\item[--] Related work is added about different techniques to calculate propagation delays of malware depending on the graph. 
+The modelling of FlipIt with a propagation delay will give a guideline for both the attacker and the defender on how to respectively plan attacks or network clean-ups.
+\item[-] This work includes an overview of different techniques used to calculate the propagation delay of malware (depending on the network layout). \todo{en het wordt ook toegepast op de formules}
 \end{itemize}
 
-While it seems maybe trivial to extend FlipIt with a propagation delay it is not so straightforward.
-Even though the basic model of FlipIt looks simple, it's mathematical treatment is not.
+While it may seem trivial to extend the basic FlipIt model with a propagation delay, it's mathematical treatment is not.
 In the paper of Laznka et all.
@@
 While the FlipIt game is quite symmetric, by adding a propagation delay the mathematical complexity rises.
-First a formula to calculate the benefit of a basic FlipIt was derived to introduce the propagation delay.
+% First a formula to calculate the benefit of a basic FlipIt was derived to introduce the propagation delay.
 To calculate the benefit for each case depending on the periods of the attacker, defender and the delay, the intervals where not independent any more as in the basic FlipIt game.

### 2015-06-15 doc/template/chap-9.tex
aanpassingen jeroen

@@
 While the probability that the attacker will move in the interval of the defender is still \textit{r}, the gain will not be half of the interval.
-Indeed, if the attacker plays after $\delta_{D}-d$, given the delay \textit{d}, he will never get in control in that interval (see figure \todo{figure 3.5}).
-The probability that the attacker plays soon enough is $\dfrac{\delta_{D}-d}{\delta_{D}}$ and this will give the attacker an average gain of $\dfrac{\delta_{D}-d}{2}$.
+Indeed, if the attacker plays after $\delta_{D}-d$, given the delay \textit{d}, he will never get in control in that interval (see figure \ref{tijdens interval}).
+\begin{figure}[hbtp]
+\centering
+\includegraphics[scale=0.7]{../../doc/template/Images/delaydtijdens.pdf}
+\caption{Attacker playing to late. If the attacker enters the defender's interval after $\delta_{D} -d$, he can not get in control in that interval.}
+\label{tijdens interval}
+\end{figure}The probability that the attacker plays soon enough is $\dfrac{\delta_{D}-d}{\delta_{D}}$ and this will give the attacker an average gain of $\dfrac{\delta_{D}-d}{2}$.
 If the attacker moves after the period of $\delta_{D}-d$, the gain of the attacker will be zero.
@@
 Looking at one interval of the defender, the average gain rate of the attacker can thus be expressed as follows:
+
 \begin{equation*}
@@
 The reason this formula overestimates the benefit of the attacker is that it assumes that the defender is always in control during the delay.
-However, if the attacker was in control in the previous interval, then he continuous to be in control during the period of the delay, see figure \todo{figuur weer toevoegen}.
+However, if the attacker was in control in the previous interval, then he continuous to be in control during the period of the delay, see figure \ref{fig:case2}.
 This means that to find the correct formula, what happens in the previous interval needs to be taken into account.
@@
 When this happens, the defender will be in control for the remainder of this interval.
-In the next interval the attacker will have to regain control, meaning that during the delay, the defender stays in control, see figure \ref{fig:case2} cases (1) and \todo{through4} (2).
+In the next interval the attacker will have to regain control, meaning that during the delay, the defender stays in control, see figure \ref{fig:case2} cases (1) to (4).
 The defender will keep the control over the resource in the next interval over a period of the delay, namely \textit{d}.
@@
 For the case with a virus we consider two cases, Case a and Case b, depending on whether the delay is shorter or longer than the difference between the attacker's and the defender's period.  \\
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[scale=0.7]{../../doc/template/Images/FlipIt2.pdf}
+\caption{All possible cases for the attacker and the defender in Case 2.A where $d + \delta_{A} < \delta_{D}$.
+As can be seen in cases (1) to (4), the defender will have control during a period of \textit{d} over the resource in the next interval when the defender has flipped in the previous interval.}
+\label{fig:case2}
+\end{figure}
 
@@
 
-\begin{figure}[hbtp]
-\caption{All possible cases for the attacker and the defender in Case 2.A where $d + delta_{A} < delta_{D}$.
-As can be seen, the defender will have control during a period of \textit{d} over the resource when the defender has flipped in the previous interval.}
-\centering
-\includegraphics[scale=0.7]{../../doc/template/Images/FlipIt2.pdf}
-\label{fig:case2}
-\end{figure}
-
 It is crucial that $ \delta_{D}$ is at least as large as $d + \delta_{A}$.
 If not, this would mean that the defender can move during the delay in the interval following the interval where the defender already moved.
-This would mean that there can be an overlap between the average gain of $\dfrac{\delta_{A}}{2}$ and the delay.
+This would mean that there can be an overlap between the average gain of $\dfrac{\delta_{A}}{2} +d$ and the delay.
 The above benefit formula would then include to much gain for the defender: the potential overlap during the delay would be counted twice.
-See figure \ref{overlap}\\
+See figure \ref{countedtwice}\\
+
 \begin{figure}[hbtp]
+\centering
 \caption{Cases where the delay would be counted twice.}
-\centering
 \includegraphics[scale=0.7]{Images/flipcase2nb.pdf}
+\label{countedtwice}
 \end{figure}

### 2015-06-15 doc/template/chap-9.tex
Figuren en nog alpha beta die verloren gegaan zijn

@@
 \begin{equation*}\label{first}
-\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - k_{D} \alpha_{D} 
+\beta_{D}(\delta_{D},\delta_{A}) = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - \dfrac{k_{D}}{ \delta_{D}}
 \end{equation*}\\
@@
 \begin{equation*}\label{first}
-\beta_{A}(\alpha_{D},\alpha_{A}) = 1 -\dfrac{\delta_{A}}{2\delta_{D}} - \dfrac{d}{\delta_{D}} - k_{A} \alpha_{A} 
+\beta_{A}(\delta_{D},\delta_{A}) = 1 -\dfrac{\delta_{A}}{2\delta_{D}} - \dfrac{d}{\delta_{D}} - \dfrac{k_{A}}{ \delta_{A}}
 \end{equation*}\\
@@
 \begin{equation*}
-\beta_{D}(\alpha_{D},\alpha_{A}) = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - k_{D} \alpha_{D} - \dfrac{(d - (\delta_{D} - \delta_{A}))^{2}}{2 \delta_{D} \delta_{A}}
+\beta_{D}(\delta_{D},\delta_{A}) = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - \dfrac{k_{D}}{ \delta_{D}} - \dfrac{(d - (\delta_{D} - \delta_{A}))^{2}}{2 \delta_{D} \delta_{A}}
 \end{equation*}\\
@@
 \begin{equation*}
-\beta_{A}(\alpha_{D},\alpha_{A}) = 1 -\dfrac{\delta_{A}}{2\delta_{D}} - \dfrac{d}{\delta_{D}} - k_{A} \alpha_{A} + \dfrac{(d - (\delta_{D} - \delta_{A}))^{2}}{2 \delta_{D} \delta_{A}}
+\beta_{A}(\delta_{D},\delta_{A}) = 1 -\dfrac{\delta_{A}}{2\delta_{D}} - \dfrac{d}{\delta_{D}} - \dfrac{k_{A}}{ \delta_{A}} + \dfrac{(d - (\delta_{D} - \delta_{A}))^{2}}{2 \delta_{D} \delta_{A}}
 \end{equation*}\\
+
+%~~\\
+%\subsection{Conclusion of this chapter}
+%\begin{description}
+%\item[-]
+%\item[-]
+%\item[-]
+%\end{description}

### 2015-06-16 doc/template/chap-5.tex
extra verbeteringen en aanvullingen

@@
 The Stuxnetworm was resident on a isolated network, meaning that there was no way to connect to the internet.
-Another example are APT's that get into honeypots.
-The attacker thinks that he is getting sensitive information, but actually this information is fake \todo{zien dat deze tekst goed geschreven is}.
+Another example is by the use of honeypots.
+Honeypots emulate services or creates multiple instances of real operating systems and can pretend to have sensitive information.
+Honeypots do not detect all malicious attacks so the attacker can stay unnoticed and can still provide information as feedback.
+This is a way how the defender can also be stealthy by providing  false information through the honeypot.
 The second difference is that even though both the targeted and non-targeted attacks do not succeed immediately, the delay is determined differently.

### 2015-06-16 doc/template/chap-9.tex
extra verbeteringen en aanvullingen

@@
 
-\textbf{Section:Required adaptations to the FlipIt game}
+%\textbf{Section:Required adaptations to the FlipIt game}
+
+Below a list of required adaptations that should be made to the FlipIt game to model a game with propagation delay or motivations to keep the assumptions made by FlipIt.
 
@@
 The basic FlipIt game consists of a single resource.
-To represent the security problem of the propagation of an APT in a network, the game now defines its single resource as a computer network with multiple
-nodes. One of the players, the defender, will try to defend his network. The defender
+To represent the security problem of the propagation of an APT in a network, the adapted game defines its single resource as a computer network with multiple
+nodes. One of the players, the defender, will also try to defend his network. The defender
 will do this by flipping all the nodes of the network (i.e. the entire resource) in every move he plays. The
-attacker on the other hand will try to infect all the nodes in the network. The attacker
-will do this by flipping the node in the graph that can infect all the nodes in the
-shortest possible time.
-After dropping a virus on the first node, it takes a while for the virus to infect the entire network.
-Since the original FlipIt game works with a single resource that is always flipped entirely, the assumption is made that the attacker is considered to gain immediate full control over the resource when the network has been infected, even it is only one node that has been infected.\\
+attacker on the other hand will try to infect all the nodes in the network.
+For an attacker on the other hand, it is quite impossible to infect all computers in a network at once.
+Rather, the attacker will try to identify the node in the graph that can infect all the nodes in the
+shortest possible time. Then, after dropping a virus on this first node, it takes a while for the virus to infect the entire network. 
+%Since the original FlipIt game works with a single resource that is always flipped entirely, the assumption is made that the attacker is considered to gain immediate full control over the resource when the network has been infected, even it is only one node that has been infected.\\
 \subsubsection{Only one action per player}
@@
 in this a thesis assumed, non stealthy is..
+\\
+Our motivation for a defender with stealthy moves is that it can not be assumed that every attacker can receive feedback from an APT.
+Some ATP's are designed only to cause harm on the systems that are infected e.g.
+the Stuxnetworm that was developed to destroy nuclear reactors.
+The Stuxnetworm was resident on a isolated network, meaning that there was no way to connect to the internet.
+Another example is by the use of honeypots.
+Honeypots emulate services or creates multiple instances of real operating systems and can pretend to have sensitive information.
+Honeypots do not detect all malicious attacks so the attacker can stay unnoticed and can still provide information as feedback.
+This is a way how the defender can also be stealthy by providing  false information through the honeypot.
 \subsubsection{Cost associated with the move}

### 2015-06-17 doc/template/chap-9.tex
Kleine veranderingen aan diff flip en flipvirus

@@
 To represent the security problem of the propagation of an APT in a network, the adapted game defines its single resource as a computer network with multiple
-nodes. One of the players, the defender, will also try to defend his network. The defender
-will do this by flipping all the nodes of the network (i.e. the entire resource) in every move he plays. The
-attacker on the other hand will try to infect all the nodes in the network.
-For an attacker on the other hand, it is quite impossible to infect all computers in a network at once.
-Rather, the attacker will try to identify the node in the graph that can infect all the nodes in the
-shortest possible time. Then, after dropping a virus on this first node, it takes a while for the virus to infect the entire network. 
+nodes. One of the players, the defender, will also try to defend his network.
+% The defender will do this by flipping all the nodes of the network (i.e. the entire resource) in every move he plays. 
+The defender will do this by flipping the nodes of the network.
+The attacker on the other hand will try to infect all the nodes in the network. 
+%For an attacker on the other hand, it is quite impossible to infect all computers in a network at once. Rather, t
+The attacker will try to identify the node in the graph that can infect all the nodes in the shortest possible time.
+Then, after dropping a virus on this first node, it takes a while for the virus to infect the entire network.
 %Since the original FlipIt game works with a single resource that is always flipped entirely, the assumption is made that the attacker is considered to gain immediate full control over the resource when the network has been infected, even it is only one node that has been infected.\\
@@
 The action of the defender consist of defending his network.
-In this context it will be defending every computer controlled device in a network.
-To make the network clean the defender can patch the systems with the latest security bulletins or reinstall the software or even wipe out the computer.
-All these different actions can imply other costs.
-The cost of patching a system will be smaller than the cost of wiping out a computer.
-\\
+The network consists of multiple nodes.
+The defender flips all the nodes of the network in every move he plays.
+The action of the defender can be extended to the action where the defender only flips a subset of the nodes of the network.
+This extention has already been investigated in the context of placing anti-virus systems in a graph.
+This problem is known to be NP-hard.
+[\todo{referentie}].
 The defender has only one action.
@@
 The attacker does not send the same virus again, because if the computers are patched, the virus cannot get in any more.
+Sending a virus in this context is equal to flipping a node in the network.
+The node can be a targeted node or a random node.
+If the attacker has knowledge about the graph, the attacker can choose to target the node that can infect all the nodes in the shortest time possible.
 \subsubsection{Immediate effect of the move}
@@
 \\
+The moves of the defender are immediate.
+It is assumed that if a defender tries to clean the network that this will happen without a delay.
+The defender can de-plug the computer from the network, or clean.
 \subsubsection{Stealth character of the move}
@@
 For example, if the attacker launches an APT only to harm the system and not to receive feedback, the moves of the defender are stealthy.
-If the attacker launches an APT to steal sensitive information, the moves of the defender are non stealthy because the attacker can see that he doesnt receive information anymore.
-But the defender can also setup a honey pot, making the attacker believe that he is stealing sensitive information, but the information in the honeypot is all fake.
+If the attacker launches an APT to steal sensitive information, the moves of the defender are non stealthy because the attacker can see that he doesn not receive information any more.
+But the defender can also set-up a honey pot, making the attacker believe that he is stealing sensitive information, but the information in the honeypot is all fake.
 In thesis we assume that both moves of the attacker and the defender are stealthy.
@@
 \subsubsection{Cost associated with the move}
+ The defender will be defending every computer controlled device in a network.
+To make the network clean the defender can patch the systems with the latest security bulletins or reinstall the software or even wipe out the computer.
+All these different actions can imply other costs.
+The cost of patching a system will be smaller than the cost of wiping out a computer.
+\\
+\todo{cost of the attacker ook toevoegen?}
 \subsubsection{Renewal strategies}
@@
 
-\subsubsection{Motivation Application}
-\subsubsection*{Wormlike ATP}
+\subsubsection{Motivation Application: Wormlike ATP}
 A business is every day under attack of different threats.

### 2015-06-20 Thesisfollowup/ThesisFollowUpSophieMarien.tex
worm propagatie afgeschreven + related work propagatie

@@
 
+Thesis follow up vragen van Maart en April terug beantwoord. 
 \section{Vergelijk wat je gedaan hebt met het opgestelde werkplan. Hoever sta je? Moet het werkplan aangepast worden?}
@@
 \section{Evalueer de tijd die je in je thesis gestoken hebt in de voorbije 4 weken: veel te weinig / te weinig / voldoende / ruim voldoende / meer dan verwacht.}
-De voorbije weken heb ik minder dan voldoende aan mijn thesis gewerkt. Ik liep vast op een aantal zaken en ik ben een week op Athens cursus geweest. 
+De voorbije weken heb ik meer dan verwacht aan mijn thesis gewerkt.  
 \section{Evalueer de vorderingen van je thesis t.o.v.

### 2015-06-20 doc/template/chap-4.tex
versie voor mama
worm propagatie afgeschreven + related work propagatie

@@
-\chapter{Virus Propagation }
-\label{chapter4: Virus propagation}
+\chapter{Worm Propagation }
+\label{chapter4: Worm propagation}
 
-\section{Methods of virus propagation}
-
-The spreading of worms has already been extensivly researched [beter schrijven].
+The spreading of malware has already been extensively researched.
 Because of all the different types of propagation it is hard to pick out one model that can model all of them.
@@
 Viruses and worms are the types of malware that are the most researched.
-Under we list a couple of examples to model virus and worm propagation.
-In this text worm and virus will be used intermediate.
+\\
+In this chapter we focus on the propagation models of worms.
+The propagation of a worm is not depending on the human interaction as it is with viruses.
+In section [] we present an overview of the most used propagation methods by worms.
+Next, in section [] different kind of graphs to model a computer network are introduced.
+Section [] lists all the kind of models to determine the time a worm needs to infect a sufficient number of nodes.
+These results can be used for the completion of parameter \textit{d}.
+Finally we introduce an easy method to calculate the delay of the propagation of a worm.
+\subsection{Introduction to this chapter}
 
-\subsection{Models of graph representing the network}
-
+Advanced Persistent Threads propagation koppelen met die van een worm.
+A virus infects one node on the network and has to wait for human interaction to spread. So the spreading speed depends on the human interaction. 
 \subsection*{Methods of propagation}
@@
 A method to calculate the propagation of the virus in an easy way. Google page ranking algorithm.
+
+\todo{toch eigen methode introduceren}
+
+\section{Weetjes}
+\begin{description}
+\item Crouching Yeti is hardly a sophisticated campaign.
+For example, the attackers used no zero-day exploits, only exploits that are widely available on the Internet.
+But that did not prevent the campaign from staying under the radar for several years.
+\url{http://www.kaspersky.com/internet-security-center/threats/crouching-yeti-energetic-bear-malware-threat}
+\end{description}
 

### 2015-06-20 doc/template/chap-5.tex
worm propagatie afgeschreven + related work propagatie

@@
 \begin{itemize}
-\item periodic games dominate the other renewal strategies, meaning that it is always advantageous to play periodically against an opponent with a renewal strategy;
+\item periodic strategy strongly dominates the other renewal strategies if the opponent has a periodic or non-arithmetic renewal strategy, meaning that it is always advantageous for the opponent to play periodically against a player with a non-adaptive strategy;
 \item periodic games are disadvantageous against players following a Last Move adaptive strategy;
-\item if the defender plays with a periodic rate that is fast enough he'll force the attacker to drop out;
+\item if a defender facing an LM attacker and with a cost much lower than the attacker, he can play with a periodic rate that is fast enough he'll force the attacker to drop out;
+\item a defender facing an LM attacker has to play with a randomized strategy e.g. exponential strategy;
 \item any amount of feedback about the opponent received during the game, benefits to a player.
@@
 
- In this paper we restrict ourselves to non-adaptive periodic strategies.
+% In this paper we restrict ourselves to non-adaptive periodic strategies.
 This choice is motivated by the fact that in a security game a player (defender or attacker) rarely has information about the moves (last move or full history) of his opponent.
-The results also indicate that the periodic strategy might be the dominant strategy.
+The results \cite{FlipIt} also indicate that the periodic strategy might be the dominant strategy.
 \\
-\todo{dit stuk tekst hierboven meer motiveren}
 

### 2015-06-20 doc/template/chap-9.tex
worm propagatie afgeschreven + related work propagatie

@@
 
-\subsubsection{Single resource}
+\subsection{Single resource}
 The basic FlipIt game consists of a single resource.
 To represent the security problem of the propagation of an APT in a network, the adapted game defines its single resource as a computer network with multiple
-nodes. One of the players, the defender, will also try to defend his network.
+nodes. One of the players, the defender, will try to defend his network.
 % The defender will do this by flipping all the nodes of the network (i.e. the entire resource) in every move he plays. 
@@
 %For an attacker on the other hand, it is quite impossible to infect all computers in a network at once. Rather, t
-The attacker will try to identify the node in the graph that can infect all the nodes in the shortest possible time.
+%The attacker will try to identify the node in the graph that can infect all the nodes in the shortest possible time.
 Then, after dropping a virus on this first node, it takes a while for the virus to infect the entire network.
+\\
+The attacker will do this by dropping a virus on a node on the network. The virus will then spread itself and infect other nodes in the network. \\
+By defining the resource as a network with multiple nodes it is possible to extend the actions of the defender and the attacker to actions depending on which node is being flipped.
+These actions will be explained in the following section.
 %Since the original FlipIt game works with a single resource that is always flipped entirely, the assumption is made that the attacker is considered to gain immediate full control over the resource when the network has been infected, even it is only one node that has been infected.\\
-\subsubsection{Only one action per player}
+\subsection{Only one action per player}
 
 The action of the defender consist of defending his network.
-The network consists of multiple nodes.
-The defender flips all the nodes of the network in every move he plays.
-The action of the defender can be extended to the action where the defender only flips a subset of the nodes of the network.
-This extention has already been investigated in the context of placing anti-virus systems in a graph.
+The network is composed of multiple nodes.
+The defender can choose to flip one node, a subset of nodes or all nodes of the network.
+In this paper, the defender flips all the nodes of the network in every move he plays.
+The action of the defender can be extended to the action where the defender only flips a subset of the nodes or a single node of the network.
+This extension has already been investigated in the context of placing anti-virus systems in a graph.
 This problem is known to be NP-hard.
 [\todo{referentie}].
+\\
 The defender has only one action.
 Sending different viruses and worms to a computer network.
-Every time the attacker flips the resource, he does this by sending a new virus to the system.
-The attacker does not send the same virus again, because if the computers are patched, the virus cannot get in any more.
-Sending a virus in this context is equal to flipping a node in the network.
+Every time the attacker flips the resource, he does this by sending a new kind of malware e.g worm to the system.
+The attacker does not send the same malware again, because if the computers are patched, the malware cannot get in any more.
+Sending malware in this context is equal to flipping a node in the network.
 The node can be a targeted node or a random node.
 If the attacker has knowledge about the graph, the attacker can choose to target the node that can infect all the nodes in the shortest time possible.
-\subsubsection{Immediate effect of the move}
+\subsection{Immediate effect of the move}
 The framework of FLipIt is used to model targeted attacks.
@@
 This is motivated by the fact that it takes a while before a targeted attack can compromise a computer network.
-The virus or worm has to spread to infect every computer controlled device in the network.
+The virus or worm has to spread to infect sufficient computer controlled device in the network.
 This propagation make take a while.
 \\
-In reality however, after dropping a virus on the first node, it takes a while for the virus to infect
-the entire network.
+%In reality however, after dropping a virus on the first node, it takes a while for the virus to infect
+%the entire network.
 So, the assumption that the attacker has full control over the resource as soon as a node has been infected, is not realistic.
@@
 It is assumed that if a defender tries to clean the network that this will happen without a delay.
-The defender can de-plug the computer from the network, or clean.
-\subsubsection{Stealth character of the move}
+The defender can de-plug the computer from the network, or just clean a pc.
+\subsection{Stealth character of the move}
 Moves in the basic FlipIt game are stealthy or covert and not immediately detected by the other payer.
@@
 For example, if the attacker launches an APT only to harm the system and not to receive feedback, the moves of the defender are stealthy.
-If the attacker launches an APT to steal sensitive information, the moves of the defender are non stealthy because the attacker can see that he doesn not receive information any more.
-But the defender can also set-up a honey pot, making the attacker believe that he is stealing sensitive information, but the information in the honeypot is all fake.
-In thesis we assume that both moves of the attacker and the defender are stealthy.
--- kunnen enerzijds niet stealthy zijn als the apt informatie steelt depending how he setup his attacker.
-depending on network connected.
-in this a thesis assumed, non stealthy is..
+If the attacker launches an APT to steal sensitive information, the moves of the defender are non stealthy because if the defender takes actions the attacker can see that he does not receive information any more.
 \\
+In paper we assume that both moves of the attacker and the defender are stealthy.
 Our motivation for a defender with stealthy moves is that it can not be assumed that every attacker can receive feedback from an APT.
-Some ATP's are designed only to cause harm on the systems that are infected e.g.
-the Stuxnetworm that was developed to destroy nuclear reactors.
-The Stuxnetworm was resident on a isolated network, meaning that there was no way to connect to the internet.
-Another example is by the use of honeypots.
+Some APT's can be resident on a isolated network (e.g.
+Stuxnetworm), meaning that there was no way to connect to the internet.
+%Another example is by the use of honeypots. 
+Even when the APT is lauched to steal sensitive information the defender can set-up a honey pot, making the attacker believe that he is stealing sensitive information, but the information in the honeypot is all fake.
 Honeypots emulate services or creates multiple instances of real operating systems and can pretend to have sensitive information.
 Honeypots do not detect all malicious attacks so the attacker can stay unnoticed and can still provide information as feedback.
-This is a way how the defender can also be stealthy by providing  false information through the honeypot.
-\subsubsection{Cost associated with the move}
+This is a way how the defender can also be stealthy by providing  false information through the honeypot.\\
+
+%  The case where the moves of the defender are non stealthy has already been investigated in a paper of Laszka \citep{MitigationNonTargeted}.
+
+% Some ATP's are designed only to cause harm on the systems that are infected e.g. the Stuxnetworm that was developed to destroy nuclear reactors.
+
+\subsection{Cost associated with the move}
  The defender will be defending every computer controlled device in a network.
@@
 \\
-\todo{cost of the attacker ook toevoegen?}
-\subsubsection{Renewal strategies}
-The formalization starts from the model of the non-adaptive continuous basic FlipIt game where players use a periodic strategy with a random phase.
-The basic FlipIt game concluded that periodic strategy is the dominant strategy against all renewal strategies.
-\todo{zelfde paper conclusie}.
-Furthermore, this choice is motivated by the assumption that in most organisations, the defence strategy is to periodically defend the network.
+The cost of the defender depend on the complexity of the vulnerabilities exploited and how difficult it was to program the malware.
+\subsection{Strategies}
+The formalization starts from the model of the non-adaptive continuous basic FlipIt game where players use a periodic strategy with a random phase. \\
+
+\textbf{No feedback (non-adaptive)}\\
+Non-adaptive strategies are taken because of the assumption that both players will receive no feedback during the game.
+This is motivated by the fact that the identity of the attackers and their attack strategy is rarely known to the defenders.
+In the case of the attacker, if an attacker wants to have feedback it may be disadvantageous if he wants to stay undetected.
+\\
+
+\textbf{Periodic strategy}\\
+The choice of periodic strategy is motivated by the assumption that in most organisations, the defence strategy is to periodically defend the network.
 This is true for the example of patching.
@@
 Microsoft security bulletins are released on the Second Tuesday of each month.
-\cite{MicroPatch}
+\cite{MicroPatch} \\
 
-This corresponds to a periodic defender strategy.
-A periodic attacker strategy is assumed as well, to be able to compare the results with the periodic strategy of the FlipIt game in \cite{FlipIt}.
-%, as this also corresponds to a common real life strategy.
-%nog eventueel verder te motiveren
+In the basic FlipIt game \citep{FlipIt} the authors concluded that periodic strategy is the dominant strategy against all other renewal strategies if the opponent uses a periodic or non-arithmetic renewal strategy.
+This means that a good choice for the attacker is also to play a periodic strategy.
+Furthermore, a periodic attacker strategy is assumed as well, to be able to compare the results with the periodic strategy of the FlipIt game in \cite{FlipIt}.
+
+The paper from Laszka \citep{MitigationNonTargeted}, which was similar to this one, also concluded that periodic is the dominant strategy against all other non-adaptive strategies.
+%This corresponds to a periodic defender strategy.   %, as this also corresponds to a common real life strategy. %nog eventueel verder te motiveren
 Further research can investigate the effect of relaxing this assumption. \\
+
+%---Periodic is probably the strategy most widely used in practice as most systems require passwords, cryptographic
+%keys, etc. to be changed at regular intervals, for example, every thirty days or every three months.
+%In [13], it was shown that the periodic strategy strongly dominates all other renewal strategies if the other
+%player uses a periodic or non-arithmetic renewal strategy. Thus, the periodic strategy is a good choice for
+%an attacker who plays against a non-adaptive (NA) defender.---
 
@@
 servers and computers.
-The attacker can drop the worm through a usb stick or by tricking a victim to open or download a malicious program.
+The attacker can drop the worm through a USB stick or by tricking a victim to open or download a malicious program.
 The last one can be achieved by a trojan horse.
@@
 ~~\\
-\todo{kleuren van de tekening aanpassen zodat ze zichtbaar zijn in een zwartwit paper}
+
 \begin{figure}[hbtp]

### 2015-06-21 Thesisfollowup/ThesisFollowUpSophieMarien.tex
hoofdstuk worm propagatie aangevuld

@@
 Thesis follow up vragen van Maart en April terug beantwoord. 
-\section{Vergelijk wat je gedaan hebt met het opgestelde werkplan. Hoever sta je? Moet het werkplan aangepast worden?}
-Ik moet een tandje bijsteken en al verder beginnen schrijven. Ik was vergeten in mijn werkplan bij te steken dat ik een week op Athens ging gaan. 
-\section{Schrijf een eerste versie van een hoofstuk over een afgewerkte bijdrage van je masterproef.}
-Dit is meegegeven in bijlage. Ik heb ineens heel mijn voorlopige thesistekst in bijlage gestoken. 
-\section{Geef een korte inhoud van het tweede artikel dat je gaat schrijven. (Dit is niet van toepassing voor de studenten van de Master Toegepaste Informatica)}
-Het tweede artikel gaat gaan over mijn implementatie van FlipIt op een bedrijfsnetwerk.
-Eerst bespreek ik hoe ik FlipIt aanpas.
-Daarna stel ik een methode voor via matrices om de propagatie van een virus te beschrijven.
-De methode die gebruikt wordt is gelijkaardig aan die van Google Page Ranking.
-Via deze functie kan ik de totale controle berekenen van een speler over een spel.
-In het artikel leg ik de methode uit aan de hand van een default spel.
-Ook geef ik een eerste setting van een spel en de bijhorende benefit formules waaruit ik al dan niet een equilibrium kan bepalen.
+\section{Hoever sta je? Moet het werkplan aangepast worden?}
+ Momenteel moet ik mijn tekst verfijnen van het hoofdstuk over worm propagatie en de related work.
+De nash equilibria moeten ook nog berekent worden.
+Na de examens ga ik terug verder werken aan het berekenen van de Nash equilibria.
+Ondertussen werk ik elke avond een deel van de tekst af.
+De hoofdstukken tot en met het Nash Equilibria zijn al volledig uitgeschreven.
+\section{Wat zijn de belangrijkste elementen uit het besluit dat je gaat schrijven in je tekst?}
+In mijn besluit ga ik, als ze bestaan, kunnen vaststellen wat de nash equilibria zijn en hier conclusies kunnen uittrekken voor de verdedigings-strategie"en van de defender. 
+\section{Verfijn de outline van je thesistekst. Geef per hoofdstuk de subsecties, en per subsectie het kernidee en een schatting van het aantal bladzijdes van die subsectie.}
+
+\begin{enumerate}
+\item \textbf{Introduction} 
+\begin{itemize}
+\item Research questions
+\item Contributions and results
+\item Conjunctures and open problems
+\item Overview thesis
+\end{itemize}
+\item \textbf{General context}
+\begin{itemize}
+\item[2.1] What is security
+\item[2.2] A brief introduction to game theory
+\item[2.3] The FlipIt game
+\item[2.4] Extensions on FlipIt
+\end{itemize}
+\item \textbf{FlipIt with propagation delay}
+\begin{itemize}
+\item[3.1] Introduction
+\item[3.2] Difference between FlipIt with and without propagation delay
+\begin{itemize}
+\item Single resource
+\item Only one action per player
+\item Immediate effect of the move
+\item Stealth character of the moves
+\item Cost associated with the move
+\item Strategies
+\item Application: Wormlike ATP
+\end{itemize}
+\item[3.3] Playing periodically with propagation delay
+\end{itemize}
+\item \textbf{Nash Equilibria}
+\begin{itemize}
+\item[4.1] Determining piecewise functions $opt_{D}(\delta_{A}$
+\item[4.2] Determining piecewise functions $opt_{A}(\delta_{D}$
+\item[4.3] Determing the Nash equilibria
+\end{itemize}
+\item\textbf{ Worm propagation}
+\begin{itemize}
+\item[5.1] Methods of worm propagation
+\item[5.2] Graph models
+\item[5.3] Models for worm propagation
+\item[5.4] Method for calculating propagation delay through matrices
+\end{itemize}
+\item \textbf{Related Work}
+\item \textbf{Conclusion}
+\end{enumerate}
+
 \section{Beschrijf in 10 lijnen de demo die je bij je verdediging zult geven.}
@@
 \section{Wat is er afgesproken met je begeleider en/of promotor om op te leveren als major milestone die het praktische werk van je thesis afsluit?}
-Dit is nog niet afgesproken.
+Dit is besproken met de begeleiders maar er is niets afgesproken omdat op voorhand niet kan gezegd worden of de analyse van het FlipIt spel voldoende zal zijn omdat de indruk kan ontstaan dat dit resultaat eenvoudig te verkrijgen was.
+Er is dus niet afgesproken wat ik moet inleveren om mijn thesis af te sluiten.
+%Als ik hier niet mee akkoord dan moet ik dit ook verduidelijken, en moet ik mijn standpunt verdedigen (in mijn tekst, en op de verdediging). 
+Voorlopig werk ik verder af wat ik zelf al voor ogen had; namelijk het vinden van de nash equilibria en het schrijven van een related work over worm propagatie. \\
+Mijn promotor stelde het volgende: Mijn masterproef is in de eerste plaats mijn verantwoordelijkheid, ik ben verantwoordelijk voor wat ik indien en wat ik doe.
+De rol van mijn begeleiders en promotor is hierbij advies en richtlijnen geven.
+Als ik het hiermee niet eens ben dan zorg ik voor de nodige argumenten en stel een sterke tekst op samen met een goede verdediging van mijn werk. 
 \section{Evalueer de tijd die je in je thesis gestoken hebt in de voorbije 4 weken: veel te weinig / te weinig / voldoende / ruim voldoende / meer dan verwacht.}
 De voorbije weken heb ik meer dan verwacht aan mijn thesis gewerkt.
+Momenteel ben ik al aan het aantal uren gekomen dat ik aan mijn thesis moet werken.
+Hierbij zullen waarschijnlijk nog eens 150 uren bijkomen.
 \section{Evalueer de vorderingen van je thesis t.o.v.
 het bereiken van de vooropgestelde doelstelling: ver achter op het schema / achter op het schema / goed op schema / vooruit op het schema / ver vooruit op het schema.}
-Ik loop een beetje achter op schema want ik had toch graag nu al een aantal formuleringen bepaald en een model uitgewerkt.
+Ik zit achter op schema want anders had ik al in eerste zit kunnen afgeven.
+De komende week probeer ik mijn Nash equilibria hoofdstuk af te werken en daarna focus ik mij op het schrijven van de hoofdstukken.
 \section{Stuur de ingevulde timesheet mee als bijlage.}

### 2015-06-21 doc/template/chap-4.tex
hoofdstuk worm propagatie aangevuld

@@
 Finally we introduce an easy method to calculate the delay of the propagation of a worm.
-\subsection{Introduction to this chapter}
 
+This related work is based on the following paper: [importance journal], 
 Advanced Persistent Threads propagation koppelen met die van een worm.
-A virus infects one node on the network and has to wait for human interaction to spread. So the spreading speed depends on the human interaction. 
-\subsection*{Methods of propagation}
+
+\subsection{Methods of propagation}
+
+There exist two kinds of APT's in the context of propagation.
+There are APT's that only launch an attack on just one target.
+The mechanism they use to propagate is the dropper mechanism.
+The dropper is the initial attack vector that compromises the system.
+The other kind of APT's target multiple victims.
+They also use a dropper mechanism and have an additional mechanism for self-propagation.
+It can be either the virus propagation or the worm like propagation.
+A virus infects one node on the network and has to wait for human interaction to spread.
+So the spreading speed depends on the human interaction.
+If an APT uses a worm propagation method it will spread by itself after it has been dropped on the network.
+Because it is not easy to incorporate the human factor in a spreading method we will only examine the worm propagation models for our APT.
+\\
+
+Below is a list of several common propagation methods:
 
 \begin{description}
-\item Random scanning: Some worms propagate by using the method of random scanning 
-IP addresses.
-Rate of terms of succeeded randwomly choosen IP adresses -> maximise good ip adresses.
-After a while the remaining nodes will be les likely to be reached.
-An example of such a worm is `Code Red' .
-\item Routing worms: they use BGP routing tables to scan the routable address space. 3 times faster than traditional worm that uses random scanning. Spyb0t or network bluepill.
-\item email worms: use the email systems to propagate. The Kak worm is a Javascript computer worm that spread itself by exploiting a bug in Outlook Express. 
+\item Scanning methods: A worm can use distinct ways of scanning: random, localized, topological or hit-list scanning.
+\begin{itemize}
+\item Random scanning: Some worms propagate by using the method of random targeting IP addresses.
+Rate of terms of succeeded randomly chosen IP addresses is very low.
+An example of such a worm is `Code Red' ans 'Slammer'.
+\item Localized scanning: --Localized scanning,
+used by the Code Red II and Nimda worms, preferentially
+scans for hosts in the ''local'' address space.--
+\item Topological scanning: --Topological scanning, used by the Morris worm, relies
+on the ''address'' information contained in the victim
+machines to locate new targets.--
+\item Hit-list scanning: With the method of hit-list scanning, before hand a short-list of vulnerable systems are made. --to speed up
+the spread of worms at the initial stage. This list consists
+of potentially vulnerable machines that are gathered
+beforehand and targeted first when the worm
+is released. An extreme case for the hit-list-scanning
+worms is a flash worm, which gathers all vulnerable
+machines into the list.--
+\end{itemize}   
+\item Routing worms: they use BGP routing tables to scan the routable address space.
+3 times faster than traditional worm that uses random scanning.
+Spyb0t or network Bluepill.
+--a ''routing worm'' [20].
+Zou et al.
+designed two types
+of routing worms [20]. One type, based on Class-A
+(x.0.0.0/8) address allocations, is thus called 'ClassA
+routing worms.'' Such worms can reduce the scanning
+space to 45.3\% of the entire IPv4 address space.
+The other type, based on BGP routing tables, is thus
+called''BGP routing worms'' Such worms can reduce
+the scanning space to only about 28.6\% of the entire
+IPv4 address space.--
+\item email worms: use the email systems to propagate. The Kak worm is a Javascript computer worm that spread itself by exploiting a bug in Outlook Express.  GhostNet
+\item DNS random scanning: --Another strategy that a worm
+can potentially employ is DNS random scanning [5],
+in which a worm uses the DNS infrastructure to locate
+likely targets by guessing DNS names instead of
+IP addresses. Such a worm in the IPv6 Internet is
+shown to exhibit a propagation speed comparable to
+that of an IPv4 random-scanning worm--
 %Pikachu worm: The virus was mainly spread through Microsoft Outlook email attachments.
@@
 \item self stopping worm: reduces speed to avoid detection. Atak worm or self stopping worm
+\item network sharing Shamoon
+\item shared files or spreading through SMB (shared message block or  Common Internet File System (CIFS)) Flame
+\item zero-day vulnerabilities
 \end{description}
 
+--dit stuk tekst overbodig--
 The start of such a propagation can be different.
@@
 
-\subsection*{Models for worm propagation}
+\subsection{Graph models}
+For some of the spreading methods, the graph of the network matters.
+It is important to have the right topology for the right method.
+Email worms need a topology that represent a social network, BGP routing worms need a topology on network level.
+--iets zeggen dat het ook belangrijk is voor de defender om zijn netwerk zo aan te passen dat het virus moeilijker kan verspreiden. Network segregation mss aanhalen ?--
+
+\begin{description}
+\item Power law
+\item Small world topology
+\item random graph
+\item ..
+\end{description}
+
+\subsection{Models for worm propagation}
 To use the models for worm propagation in our model of the FlipIt game it is suffi"ent to know the propagation speed of the worm and define after how many time units the network is defected.
@@
 
-What do we need for an algorithm
-\begin{description}
-\item Graph network $G = < V, E>$
-\item Graph matrix $[A]$ which is $|V| \times |V| $
-\item Attack vector $[X]$ which is $1 \times |V|$
-\item cummulative matrix $[M]$ which is $|V| \times |V|$
-\item state matrix $[T]$  which is $|V| \times |V|$
-\item Reset vector $[R]$
-\item duration \textit{d}
-\item time \textit{n}
-\item rate $\delta _{0}$ of defender and $\delta _{1}$ of attacker
-\end{description}
-
-Initialisation algorithm:
-
-\begin{verbatim}
-initialisatie
-	d=0
-	A=basismatrix
-	M=A^{0}
-	n=0
-	\delta_{0}
-	\delta_{1}
-	X
-	R
-	controller = defender
-
-	Algorithm
-	n:= n + 1;
-	Check who is in control? ( through modulo )
-	if ( defender & controller=defender)
-				d:= d + 1;
-
-	if ( defender & controller=attacker )
-				G = X \times R  (flippen ten voordele van defender)
-				d = 0
-				controller = defender
-
-	if ( attacker & controller=defender )
-				controller=attacker
-				..
-
-	if ( attacker & contoller=attacker )
-				d:= d + 1
-				M = M x A
-				T = T + M
-				G = X x T
-
-\end{verbatim}
+%What do we need for an algorithm
+%\begin{description}
+%\item Graph network $G = < V, E>$
+%\item Graph matrix $[A]$ which is $|V| \times |V| $
+%\item Attack vector $[X]$ which is $1 \times |V|$
+%\item cummulative matrix $[M]$ which is $|V| \times |V|$
+%\item state matrix $[T]$  which is $|V| \times |V|$
+%\item Reset vector $[R]$
+%\item duration \textit{d}
+%\item time \textit{n}
+%\item rate $\delta _{0}$ of defender and $\delta _{1}$ of attacker
+%\end{description}
+%
+%
+%
+%Initialisation algorithm:
+%
+%
+%\begin{verbatim}
+%initialisatie
+%	d=0
+%	A=basismatrix
+%	M=A^{0}
+%	n=0
+%	\delta_{0}
+%	\delta_{1}
+%	X
+%	R
+%	controller = defender
+%	
+%	
+%
+%	Algorithm
+%	n:= n + 1;
+%	Check who is in control? ( through modulo )
+%	if ( defender & controller=defender)
+%				d:= d + 1;
+%	
+%	if ( defender & controller=attacker )
+%				G = X \times R  (flippen ten voordele van defender)
+%				d = 0
+%				controller = defender
+%				
+%	if ( attacker & controller=defender )
+%				controller=attacker
+%				..
+%				
+%	if ( attacker & contoller=attacker )
+%				d:= d + 1
+%				M = M x A
+%				T = T + M
+%				G = X x T
+%				
+%		
+%\end{verbatim}
 

### 2015-06-21 doc/template/chap-9.tex
hoofdstuk worm propagatie aangevuld

@@
 This problem is known to be NP-hard.
-[\todo{referentie}].
+\cite{NPhard}.
 \\
@@
 %Another example is by the use of honeypots. 
-Even when the APT is lauched to steal sensitive information the defender can set-up a honey pot, making the attacker believe that he is stealing sensitive information, but the information in the honeypot is all fake.
+Even when the APT is launched to steal sensitive information the defender can set-up a honey pot, making the attacker believe that he is stealing sensitive information, but the information in the honeypot is all fake.
 Honeypots emulate services or creates multiple instances of real operating systems and can pretend to have sensitive information.

### 2015-06-22 Thesisfollowup/ThesisFollowUpSophieMarien.tex
nelezen verbeteringen

@@
 \section{Hoever sta je? Moet het werkplan aangepast worden?}
- Momenteel moet ik mijn tekst verfijnen van het hoofdstuk over worm propagatie en de related work.
-De nash equilibria moeten ook nog berekent worden.
-Na de examens ga ik terug verder werken aan het berekenen van de Nash equilibria.
-Ondertussen werk ik elke avond een deel van de tekst af.
-De hoofdstukken tot en met het Nash Equilibria zijn al volledig uitgeschreven.
+Het werkplan werd aangepast in functie van indiening in Augustus.
+Doelstellingen voor juni zijn: afwerken van de tekst van hoofdstukken 1,2, 3 en 5 en finaliseren van de formules voor de Nash evenwichten.
+Momenteel moet ik nog de tekst verfijnen van hoofdstuk 5 over worm propagatie en de related work.
+De formules voor de nash equilibria zijn al deels ontwikkelde, maar moeten nog verder uitgewerkt worden.
+De week van 22 juni is gereserveerd voor de examens.
+Vanaf 27 juni ga ik terug verder werken aan het berekenen van de Nash equilibria.
+Tussen 6 en 31 juli wordt de tekst door vrienden nagelezen op tikfouten, structuur en verstaanbaarheid.
+In augustus wordt de tekst van hoofdstukken 4 en 6 afgewerkt
 \section{Wat zijn de belangrijkste elementen uit het besluit dat je gaat schrijven in je tekst?}
-In mijn besluit ga ik, als ze bestaan, kunnen vaststellen wat de nash equilibria zijn en hier conclusies kunnen uittrekken voor de verdedigings-strategie"en van de defender. 
+\begin{itemize}
+\item vaststellen wat de nash equilibria zijn
+\item conclusies voor de verdedigingsstrategie"en van de defender
+\item conclusies voor de aanvalsstrategie"en van de attacker
+\item Ide"en voor verder onderzoek. 
+\end{itemize}
+
 \section{Verfijn de outline van je thesistekst. Geef per hoofdstuk de subsecties, en per subsectie het kernidee en een schatting van het aantal bladzijdes van die subsectie.}
@@
 \begin{enumerate}
-\item \textbf{Introduction} 
+\item \textbf{Introduction}(4 blz., final draft afgewerkt) 
 \begin{itemize}
@@
 \end{itemize}
-\item \textbf{General context}
+\item \textbf{General context}(11 blz., final draft afgewerkt)
 \begin{itemize}
@@
 \item[2.3] The FlipIt game
-\item[2.4] Extensions on FlipIt
+\item[2.4] Related Work on Extensions to FlipIt
 \end{itemize}
-\item \textbf{FlipIt with propagation delay}
+\item \textbf{FlipIt with propagation delay} (13 blz., final draft afgewerkt)
 \begin{itemize}
@@
 \end{itemize}
-\item[3.3] Playing periodically with propagation delay
+\item[3.3] Formalization of the periodic game with propagation delay
 \end{itemize}
-\item \textbf{Nash Equilibria}
+\item \textbf{Nash Equilibria}(formules deels uitgewerkt, geschatte lenge: 12 blz.)
 \begin{itemize}
-\item[4.1] Determining piecewise functions $opt_{D}(\delta_{A}$
-\item[4.2] Determining piecewise functions $opt_{A}(\delta_{D}$
-\item[4.3] Determing the Nash equilibria
+\item[4.1] Determining piecewise functions $opt_{D}(\delta_{A})$
+\item[4.2] Determining piecewise functions $opt_{A}(\delta_{D})$
+\item[4.3] Determining the Nash equilibria
 \end{itemize}
-\item\textbf{ Worm propagation}
+\item\textbf{Models for the Delay}(8 blz., final draft afgewerkt)
 \begin{itemize}
@@
 \end{itemize}
-\item \textbf{Related Work}
-\item \textbf{Conclusion}
+\item \textbf{Related Work: Alternative approaches to the security problem}  
+\begin{itemize}
+\item (optioneel hoofdstuk nog uit te werken in functie van beschikbare tijd, geschatte lengte: 2 blz.
+\item In dit hoofdstuk zou ik alle literatuur willen bespreken die ik aan het begin van mijn thesis heb gelezen met het oog op het afbakenen van het onderwerp.
+Dat zijn een hele hoop papers, maar er moet nog een goed stramien gevonden worden voor het structureren van de bespreking.
+\end{itemize}
+\item \textbf{Conclusion}(nog uit te werken, geschatte lengte: 2 blz.) 
 \end{enumerate}
@@
 \section{Beschrijf in 10 lijnen de demo die je bij je verdediging zult geven.}
-Ik heb een theoretische thesis dus kan ik moeilijk een demo geven. Ik zal enkele voorbeelden uitwerken om mijn werkingsmethode uit te leggen.
+Ik heb een theoretische thesis en dus kan ik geen live demo geven. Ik ben van plan om mijn werkingsmethode uit te leggen aan de hand van enkele voorbeelden die ik in figuren zal uitwerken.
 \section{Wat is er afgesproken met je begeleider en/of promotor om op te leveren als major milestone die het praktische werk van je thesis afsluit?}
-Dit is besproken met de begeleiders maar er is niets afgesproken omdat op voorhand niet kan gezegd worden of de analyse van het FlipIt spel voldoende zal zijn omdat de indruk kan ontstaan dat dit resultaat eenvoudig te verkrijgen was.
-Er is dus niet afgesproken wat ik moet inleveren om mijn thesis af te sluiten.
+Dit werd op de laatste vergadering besproken met de begeleiders.
+Ik stuurde na de vergadering een samenvattend verslag van de bespreking, maar de begeleiders antwoordden dat er op voorhand niet kan gezegd worden of de analyse van het FlipIt spel voldoende zal zijn omdat de indruk kan ontstaan dat dit resultaat eenvoudig te verkrijgen was.
+Er is dus niets afgesproken wat ik moet inleveren om mijn thesis af te sluiten.
 %Als ik hier niet mee akkoord dan moet ik dit ook verduidelijken, en moet ik mijn standpunt verdedigen (in mijn tekst, en op de verdediging). 
 Voorlopig werk ik verder af wat ik zelf al voor ogen had; namelijk het vinden van de nash equilibria en het schrijven van een related work over worm propagatie. \\
-Mijn promotor stelde het volgende: Mijn masterproef is in de eerste plaats mijn verantwoordelijkheid, ik ben verantwoordelijk voor wat ik indien en wat ik doe.
+Ik stelde dezelfde vraag aan mijn promotor en kreeg het volgende antwoord: Mijn masterproef is in de eerste plaats mijn verantwoordelijkheid, ik ben verantwoordelijk voor wat ik indien en wat ik doe.
 De rol van mijn begeleiders en promotor is hierbij advies en richtlijnen geven.
 Als ik het hiermee niet eens ben dan zorg ik voor de nodige argumenten en stel een sterke tekst op samen met een goede verdediging van mijn werk. 
+Besluit: Er is dus vanuit promotor en werkleider geen concrete afspraak over wat ik moet inleveren om mijn thesis af te sluiten.
+Ik ben zelf van oordeel dat het uitwerken van de periodische strategie volstaat om tegemoet te komen aan de doelstelling van mijn masterproef en zal dit beargumenteren in de tekst van mijn masterthesis.
 \section{Evalueer de tijd die je in je thesis gestoken hebt in de voorbije 4 weken: veel te weinig / te weinig / voldoende / ruim voldoende / meer dan verwacht.}
-De voorbije weken heb ik meer dan verwacht aan mijn thesis gewerkt.
-Momenteel ben ik al aan het aantal uren gekomen dat ik aan mijn thesis moet werken.
-Hierbij zullen waarschijnlijk nog eens 150 uren bijkomen.
+De voorbije weken heb ik meer dan verwacht aan mijn thesis gewerkt. Momenteel heb  ik al 765 uren gewerkt. Hier zullen naar schatting nog eens 150 uren bijkomen. 
 \section{Evalueer de vorderingen van je thesis t.o.v.
 het bereiken van de vooropgestelde doelstelling: ver achter op het schema / achter op het schema / goed op schema / vooruit op het schema / ver vooruit op het schema.}
-Ik zit achter op schema want anders had ik al in eerste zit kunnen afgeven.
-De komende week probeer ik mijn Nash equilibria hoofdstuk af te werken en daarna focus ik mij op het schrijven van de hoofdstukken.
+Het geplande werk voor juni is uitgevoerd volgens schema. Tot 28 juni werk ik de tekst van hoofdstuk 5 verder af. In de week van 29 juni probeer ik mijn Nash equilibria hoofdstuk af te werken. 
 \section{Stuur de ingevulde timesheet mee als bijlage.}

### 2015-06-22 doc/template/chap-4.tex
nelezen verbeteringen

@@
-\chapter{Worm Propagation }
+\chapter{Models for the Delay }
 \label{chapter4: Worm propagation}
 
+\todo{nog uitleggen wat dit hoofdstuk hier staat te doen.
+The formalisation of the FlipIt game with delay relies on some value "d" that represents the time needed to infect a sufficient number of nodes in a network after the initial infection.
+This chapter provides the reader with more insight on how to calculate the value of this parameter d.
+}
 The  spreading of malware has already been extensively researched.
-Because of all the different types of propagation it is hard to pick out one model that can model all of them.
-Modelling the spread of malware depends on two great factors: the method used for the propagation and the graph where the malware will spread itself.
+Because of the many different types of propagation it is hard to defined a single model that can model all of them.
+Modelling the spread of malware depends on two key factors: the method used for the propagation and the graph where the malware will spread itself.
+\\
 Viruses and worms are the types of malware that are the most researched.
-\\
-In this chapter we focus on the propagation models of worms.
-The propagation of a worm is not depending on the human interaction as it is with viruses.
-In section [] we present an overview of the most used propagation methods by worms.
-Next, in section [] different kind of graphs to model a computer network are introduced.
-Section [] lists all the kind of models to determine the time a worm needs to infect a sufficient number of nodes.
+Since the spreading of viruses requires human interaction, their propagation delay is depending on (hard to predict) human behaviour.
+Worms on the other hand, spread without human intervention, and are their propagation is therefore easier to model.
+Since the purpose of this chapter is only to illustrate how a delay can be calculated, we will limit the presentation to propagation models of worms .
+Section []  presents an overview of the most frequently used propagation methods by worms.
+Next, section [] introduces  different kind of graphs to model a computer network are introduced.
+Section [] presents    different kinds of models to determine the time a worm needs to infect a sufficient number of nodes.
 These results can be used for the completion of parameter \textit{d}.
@@
 
-\subsection{Methods of propagation}
+\subsection {Methods of propagation}
 
-There exist two kinds of APT's in the context of propagation.
-There are APT's that only launch an attack on just one target.
-The mechanism they use to propagate is the dropper mechanism.
-The dropper is the initial attack vector that compromises the system.
-The other kind of APT's target multiple victims.
-They also use a dropper mechanism and have an additional mechanism for self-propagation.
-It can be either the virus propagation or the worm like propagation.
+In the context of malware propagation, there exists two kinds of APT's .
+First, there are APTs that launch an attack on just one target node of a network.
+The mechanism these APT's use to propagate malware is the dropper mechanism.
+The dropper is the initial attack vector that compromises the single targeted node of the system.
+The second kind of APTs target multiple victims.
+These APTs also use a dropper mechanism but have an additional mechanism for self-propagation in order to propagate themselves to the multiple targeted nodes of the system.
+This mechanism can either be a virus propagation or a worm propagation.
 A virus infects one node on the network and has to wait for human interaction to spread.
-So the spreading speed depends on the human interaction.
+So the spreading speed depends on the human interaction, and modeling virus propagation therefore requires modeling human behavior.
 If an APT uses a worm propagation method it will spread by itself after it has been dropped on the network.
-Because it is not easy to incorporate the human factor in a spreading method we will only examine the worm propagation models for our APT.
+Given the additional complexity of incorporating the human factor in a spreading method and that this chapter is merely meant as illustration on calculating the delay for use in the game theoretic approach, we will limit the presentation to worm propagation models for APT's.
 \\
 
-Below is a list of several common propagation methods:
-
+Infection by worms start with dropping the work on the network.
+Different dropping mechanisms can be used, whereby the initial attack can be random or targeted at a specific node.
+Frequently used mechanisms are USB sticks (given to a specific person or left behind for pick up by a random person), email, or malicious software.
+More important for determining our delay is the propagation strategy.
+Propagation is performed by determining the next nodes to spread the worm to.
+The following are common propagation methods:
 \begin{description}
-\item Scanning methods: A worm can use distinct ways of scanning: random, localized, topological or hit-list scanning.
+\item Scanning methods: A worm tries to guess (scan) the address of potential target nodes to infect. It can use distinct ways of scanning: random, localized, topological or hit-list scanning.
 \begin{itemize}
 \item Random scanning: Some worms propagate by using the method of random targeting IP addresses.
-Rate of terms of succeeded randomly chosen IP addresses is very low.
-An example of such a worm is `Code Red' ans 'Slammer'.
+The rate of success for randomly chosen IP addresses is very low.
+Example of such worms are `Code Red' and 'Slammer '.
 \item Localized scanning: --Localized scanning,
@@
 machines to locate new targets.--
-\item Hit-list scanning: With the method of hit-list scanning, before hand a short-list of vulnerable systems are made. --to speed up
+\item Hit-list scanning: With the method of hit-list scanning, a short-list of vulnerable systems is made beforehand. --to speed up
 the spread of worms at the initial stage. This list consists
@@
 --dit stuk tekst overbodig--
-The start of such a propagation can be different.
-Either it is a targeted attack or a random attack.
-For example a usb stick or other shared resources can be used to start the spreading.
-It can be given to some special person or it can just be left behind in a parking lot.
-This can still be a targeted attack if it is the parking lot of a targeted company but there is more insecurity of where the usb stick will end.
-Another way is by sending an email.
-It can also start from one computer that is connected directly to the internet.
-Downloading malicious programs.
 
@@
 We model the network through an undirected Graph $G = < V, E> $ where $|V|$ denotes the number of resources in the network and $|E|$ the number of connections.
-We can convert this to a adjacent matrix where we can represent which vertices of the graph are neighbours of other vertices.
+We can convert this to an adjacency matrix which represents which vertices of the graph are neighbors of other vertices.
 \\
-For our graph we have an $|V| \times |V|$ matrix with on every entry $a_{ij}$ a 1 as value if there is a connection between node $V_{i}$ and $V_{j}$ and with zeros its diagonal.
-Because our graph is undirected we have a symmetric matrix.
-
+The graph is represented as a $|V| \times |V|$ matrix with for every entry $a_{ij}$ a 1 as value if there is a connection between node $V_{i}$ and $V_{j}$ and 0 otherwise, and with 0's for every $a_{ii}$.
+Because the graph is undirected we have a symmetric matrix.
+Adjacency matrices have many interesting applications, amongst which calculating the paths between vertices:
 \textit{"If \textit{A} is the adjacency matrix of the directed or undirected graph \textit{G}, then the matrix $A^{n}$ (i.e., the matrix product of n copies of \textit{A}) has an interesting interpretation: the entry in row i and column j gives the number of (directed or undirected) walks of length n from vertex i to vertex j.
@@
 %\] 
-The adjacent matrix becomes this matrix $[A]$: \\
+Assuming a network like in Fig, the corresponding adjacency matrix is the  matrix $[A]$: \\
 
@@
 
-Matrix $A \times A = A^{2}$ becomes the matrix with the number of paths with 2 steps from $N_{i}$ to $N_{j}$: We denote this matrix as matrix \textit{[B]}\\
+In matrix $A \times A = A^{2}$, each entry represents the number of 2 step paths from $N_{i}$ to $N_{j}$: We denote this matrix as matrix \textit{[B]}\\
 
@@
 
-Matrix $A^{2} \times A = A^{3}$ becomes the matrix with the number of paths with 3 steps from $N_{i}$ to $N_{j}$: We denote this matrix as matrix \textit{[C]}\\
+Likewise, in matrix $A^{2} \times A = A^{3}$ each entry represents the number of paths with 3 steps from $N_{i}$ to $N_{j}$: We denote this matrix as matrix \textit{[C]}\\
 
@@
 }$
-\\
 
-So for $A^{N}$ every $a_{ij}$ entry gives the number of paths with N steps from $N_{i}$ to $N_{j}$.\\
+~~\\
+So, in $A^{N}$ every $a_{ij}$ entry gives the number of paths with N steps from $N_{i}$ to $N_{j}$.\\
 
-With this knowledge we can calculate in how many steps a node is infected.
-$A$ calculates which nodes are infected after 1 step, $A^{N}$ calculates which nodes are infected in N steps..
-So if we want to know how many nodes are infected after 3 steps we have to add every matrix $(A + A^{2} + A^{3}) $ and see which entry is a non zero entry.
+Using this knowledge we can calculate in how many steps a node is infected: 
+The infection vector or start vector I0 of lenght $|V|$ indicates which node is infected and which not.
+Multiplying the infection vector with $A$ results in a vector I1, which represents which nodes are infected after 1 step.
+Multiplying the start vector I0 with $A^{N}$ calculates which nodes are infected in N steps: each non zero entry represents an node infected after N steps.
+In the context of the FlipIt game, it is safe to assume that once a node is infected, it stays infected until the defender Flips the node.
+This means that after d steps, it is assumed that all nodes that could be reached in less than d steps from the node where the worm was dropped, are infected.
+In order to know how many nodes are infected after (for example) at most 3 steps, we have to consider nodes that are infected initially (step 1), or after 2 steps, or after 3 steps.
+Calculating the sum of the three matrices $(A + A^{2} + A^{3}) $ results in a matrix that indicates for each node, the number of paths of length 1, 2 or 3 from i to j.
+Multiplying the start vector with this matrix, results in a vector that indicates which nodes will be infected in at most 3 steps .
+This technique can be applied to calculate the state of the network after any number of steps.
+Determining the length of the delay boils down to determining which configuration of infected nodes is considered as corresponding to the attacker having flipped the resource.
+It may be that all nodes need to be flipped, a sufficient amount of nodes, or a (set of) particular nodes.
+--Hier nog een beetje verder uitschrijven hoe je exact de d dan moet bepalen, gegeven dat je niet weet in welke knoop het virus gedropt wordt--
 

### 2015-06-22 doc/template/chap-9.tex
nelezen verbeteringen

@@
 The time that it takes for the virus to infect every node (or a sufficient number of nodes) will be
-denoted as an infection-delay variable \textit{d} (called 'delay' for short in the remainder of this paper). If we want to measure how long it takes for the virus to
+denoted as an propagation delay variable \textit{d} (called 'delay' for short in the remainder of this paper). If we want to measure how long it takes for the virus to
 infect all the nodes in the network, we have to calculate the shortest path from the
@@
 
-\section{Playing periodically with propagation delay}
+\section{Formalization of the periodic game with propagation delay}
 \label{ch2:periodicvirus}
-
-\subsection{Formalization the benefit formula including the infection-delay}
 

### 2015-06-22 doc/template/chap-4.tex
verdere aanpassingen

@@
-\chapter{Models for the Delay }
+\chapter{Models for the Delay}
 \label{chapter4: Worm propagation}
@@
 
-\subsection {Methods of propagation}
+\section{Methods of propagation}
 
@@
 
-\subsection{Graph models}
+\section{Graph models}
 For some of the spreading methods, the graph of the network matters.
@@
 
-\subsection{Models for worm propagation}
+\section{Models for worm propagation}
 To use the models for worm propagation in our model of the FlipIt game it is suffi"ent to know the propagation speed of the worm and define after how many time units the network is defected.
@@
 
-\section{Weetjes}
+\subsection{Weetjes}
 \begin{description}

### 2015-06-22 doc/template/chap-9.tex
verdere aanpassingen

@@
 The last section \ref{ch2:periodicvirus} derives a formula to calculate the benefit for a FlipIt game with propagation delay.
+In the next Chapter, this benefit formula will be used to determine through nash equilibria what the best defence and attack strategies are.
 %Section \todo{nash} presents the Nash equilibrium for the benefit formula. The chapter concludes in section \todo{section toevoegen} with lessons learned en practical implications for cyber security.
@@
 By defining the resource as a network with multiple nodes it is possible to extend the actions of the defender and the attacker to actions depending on which node is being flipped.
-These actions will be explained in the following section.
+These actions will be explained in the following subsection.
 %Since the original FlipIt game works with a single resource that is always flipped entirely, the assumption is made that the attacker is considered to gain immediate full control over the resource when the network has been infected, even it is only one node that has been infected.\\
-\subsection{Only one action per player}
+\subsection{Actions of the players}
 
@@
 \cite{NPhard}.
-\\
+\\ \todo{niet volledig juist, defender systemen geplaatst en dan gezien hoe ver de worm kan geraken}
 The defender has only one action.
@@
 \subsection{Immediate effect of the move}
-The framework of FLipIt is used to model targeted attacks.
+The framework of FlipIt is used to model targeted attacks.
 In FlipIt when an attacker attacks, he will get immediately control over the resource.
@@
 \subsection{Stealth character of the move}
-Moves in the basic FlipIt game are stealthy or covert and not immediately detected by the other payer.
-The attacker attacker's moves are stealthy because we want to model a scenario where a computer network is attacker by an APT.
+Moves in the basic FlipIt game are stealthy or covert and not immediately detected by the other player.
+The attacker's moves are stealthy because we want to model a scenario where a computer network is attacked by an APT.
 The main characteristic of an APT is that the attack is stealthy.
@@
 \\
-In paper we assume that both moves of the attacker and the defender are stealthy.
+In this paper we assume that both moves of the attacker and the defender are stealthy.
 Our motivation for a defender with stealthy moves is that it can not be assumed that every attacker can receive feedback from an APT.
@@
 Honeypots do not detect all malicious attacks so the attacker can stay unnoticed and can still provide information as feedback.
-This is a way how the defender can also be stealthy by providing  false information through the honeypot.\\
-
-%  The case where the moves of the defender are non stealthy has already been investigated in a paper of Laszka \citep{MitigationNonTargeted}.
-
-% Some ATP's are designed only to cause harm on the systems that are infected e.g. the Stuxnetworm that was developed to destroy nuclear reactors.
+This is a way how the defender can also be stealthy by providing  false information through the honeypot.
+Furhtermore, the case where the moves of the defender are non stealthy has already been investigated by Laszka \citep{MitigationNonTargeted}.
 
@@
 \\
-The cost of the defender depend on the complexity of the vulnerabilities exploited and how difficult it was to program the malware.
+The cost of the attacker depend on the complexity of the vulnerabilities exploited and how difficult it was to program the malware.
 \subsection{Strategies}
@@
 
-In the basic FlipIt game \citep{FlipIt} the authors concluded that periodic strategy is the dominant strategy against all other renewal strategies if the opponent uses a periodic or non-arithmetic renewal strategy.
+In the basic FlipIt game \citep{FlipIt} the authors concluded that a periodic strategy is the dominant strategy against all other renewal strategies if the opponent uses a periodic or non-arithmetic renewal strategy.
 This means that a good choice for the attacker is also to play a periodic strategy.
@@
 
-The paper from Laszka \citep{MitigationNonTargeted}, which was similar to this one, also concluded that periodic is the dominant strategy against all other non-adaptive strategies.
+The paper from Laszka \citep{MitigationNonTargeted}, which was similar to this one, also concluded that a periodic strategy is the dominant strategy against all other non-adaptive strategies.
 %This corresponds to a periodic defender strategy.   %, as this also corresponds to a common real life strategy. %nog eventueel verder te motiveren

### 2015-06-22 doc/template/chap-4.tex
worm propagation methods

@@
 
-\todo{nog uitleggen wat dit hoofdstuk hier staat te doen.
-The formalisation of the FlipIt game with delay relies on some value "d" that represents the time needed to infect a sufficient number of nodes in a network after the initial infection.
-This chapter provides the reader with more insight on how to calculate the value of this parameter d.
-}
+The formalisation of the FlipIt game with delay relies on some value  \textit{d} that represents the time needed to infect a sufficient number of nodes in a network after the initial infection.
+This chapter provides the reader with more insight on how to calculate the value of this parameter \textit{d}.
+\\
+
 The  spreading of malware has already been extensively researched.
@@
 Since the purpose of this chapter is only to illustrate how a delay can be calculated, we will limit the presentation to propagation models of worms .
-Section []  presents an overview of the most frequently used propagation methods by worms.
-Next, section [] introduces  different kind of graphs to model a computer network are introduced.
-Section [] presents    different kinds of models to determine the time a worm needs to infect a sufficient number of nodes.
+Section \ref{methodsofpropagation} presents an overview of the most frequently used propagation methods by worms.
+These propagation methods will be covered by different kinds of models in section \ref{modelsforpropagation} that presents models to determine the time a worm needs to infect a sufficient number of nodes.
 These results can be used for the completion of parameter \textit{d}.
-Finally we introduce an easy method to calculate the delay of the propagation of a worm.
+Finally, in section \ref{eigenmatrixmethode},  we introduce an easy method to calculate the delay of the propagation of a worm.
+Different kind of graphs to model a computer network are introduced.
 
-This related work is based on the following paper: [importance journal], 
-Advanced Persistent Threads propagation koppelen met die van een worm.
+This chapter is based on the following papers: \cite{importantjournal}, \cite{OnWorms2005survey}, \cite{GameTheorApprCostBenefitAnalyses} and \cite{SecurelistAPT}.
 
 \section{Methods of propagation}
+\label{methodsofpropagation}
 
 In the context of malware propagation, there exists two kinds of APT's .
-First, there are APTs that launch an attack on just one target node of a network.
+First, there are APT's that launch an attack on just one target node of a network.
 The mechanism these APT's use to propagate malware is the dropper mechanism.
 The dropper is the initial attack vector that compromises the single targeted node of the system.
-The second kind of APTs target multiple victims.
-These APTs also use a dropper mechanism but have an additional mechanism for self-propagation in order to propagate themselves to the multiple targeted nodes of the system.
+The second kind of APT's target multiple victims.
+These APT's also use a dropper mechanism but have an additional mechanism for self-propagation in order to propagate themselves to the multiple targeted nodes of the system.
 This mechanism can either be a virus propagation or a worm propagation.
 A virus infects one node on the network and has to wait for human interaction to spread.
-So the spreading speed depends on the human interaction, and modeling virus propagation therefore requires modeling human behavior.
+So the spreading speed depends on the human interaction, and modelling virus propagation therefore requires modelling human behaviour.
 If an APT uses a worm propagation method it will spread by itself after it has been dropped on the network.
@@
 Different dropping mechanisms can be used, whereby the initial attack can be random or targeted at a specific node.
-Frequently used mechanisms are USB sticks (given to a specific person or left behind for pick up by a random person), email, or malicious software.
+Frequently used mechanisms are USB sticks (given to a specific person or left behind for pick up by a random person), email, or malicious software through phising or trojan horses.
 More important for determining our delay is the propagation strategy.
@@
 \begin{description}
-\item Scanning methods: A worm tries to guess (scan) the address of potential target nodes to infect. It can use distinct ways of scanning: random, localized, topological or hit-list scanning.
-\begin{itemize}
-\item Random scanning: Some worms propagate by using the method of random targeting IP addresses.
-The rate of success for randomly chosen IP addresses is very low.
-Example of such worms are `Code Red' and 'Slammer '.
-\item Localized scanning: --Localized scanning,
-used by the Code Red II and Nimda worms, preferentially
-scans for hosts in the ''local'' address space.--
-\item Topological scanning: --Topological scanning, used by the Morris worm, relies
-on the ''address'' information contained in the victim
-machines to locate new targets.--
-\item Hit-list scanning: With the method of hit-list scanning, a short-list of vulnerable systems is made beforehand. --to speed up
-the spread of worms at the initial stage. This list consists
+%\item Scanning methods: A worm tries to guess (scan) the address of potential target nodes to infect. It can use distinct ways of scanning.
+%% random, localized, topological or hit-list scanning.
+%\begin{itemize}
+\item \textbf{Selective Random scanning:} Some worms propagate by using the method of selective random targeting IP addresses.
+The worm selects randomly a part of the selected address space instead of scanning the whole address space.
+The reserved address blocks and the unassigned address are excluded from the address space.
+The rate of success for randomly chosen IP addresses is very low, but it is easy to implement.
+Example of such worms are `Code Red' and `Slammer'.
+\item \textbf{Localized scanning:} A worm that uses localized scanning will scan for hosts in the local address space. This method is used by the `Code Red II' and `Nimda' worm.
+\item \textbf{Topological scanning:} Address information stored in the victim machines are used to locate new targets. This method used by the `Morris' worm.
+\item \textbf{Hit-list scanning:} With the method of hit-list scanning, a short-list of vulnerable systems is made beforehand.  This is done to speed up
+the spread of worms at the initial stage. --This list consists
 of potentially vulnerable machines that are gathered
@@
 machines into the list.--
-\end{itemize}   
-\item Routing worms: they use BGP routing tables to scan the routable address space.
-3 times faster than traditional worm that uses random scanning.
-Spyb0t or network Bluepill.
+\item \textbf{Sequential scanning:}
+\item \textbf{DNS random scanning:} Another strategy is a kind of strategy in which the DNS infrastructure is used to locate a new target address.
+The IP address table from a DNS server is acquired from DNS records.
+The speed of a DNS scanning worm in the IPv6 internet is comparable to the speed of an IPv4 random scanning worm.
+There are some difficulties with this scanning technique.
+First it is difficult to obtain the whole address space from the DNS records.
+Secondly, the worm has to carry the database, which can slow down the propagation if the database is to big.
+Last, the IP addresses stored in the address table are only hosts with public domain names.
+\item \textbf{Routable scanning:} they use BGP routing tables to scan the routable address space.
+This method is three times faster than a traditional worm that uses random scanning.
+Examples are `Spyb0t' or `network Bluepill'.
 --a ''routing worm'' [20].
@@
 IPv4 address space.--
-\item email worms: use the email systems to propagate. The Kak worm is a Javascript computer worm that spread itself by exploiting a bug in Outlook Express.  GhostNet
-\item DNS random scanning: --Another strategy that a worm
-can potentially employ is DNS random scanning [5],
-in which a worm uses the DNS infrastructure to locate
-likely targets by guessing DNS names instead of
-IP addresses. Such a worm in the IPv6 Internet is
-shown to exhibit a propagation speed comparable to
-that of an IPv4 random-scanning worm--
+%\end{itemize}
+
+\item \textbf{Email worms:} use the email systems to propagate. The Kak worm is a Javascript computer worm that spread itself by exploiting a bug in Outlook Express.  Another example is `GhostNet'
+
 %Pikachu worm: The virus was mainly spread through Microsoft Outlook email attachments.
@@
 [5]
-\item self stopping worm: reduces speed to avoid detection. Atak worm or self stopping worm
-\item network sharing Shamoon
-\item shared files or spreading through SMB (shared message block or  Common Internet File System (CIFS)) Flame
-\item zero-day vulnerabilities
+\item \textbf{self stopping worm:} reduces speed to avoid detection. Atak worm or self stopping worm \cite{VirusChangePropSpeed}.
+\item \textbf{network sharing} `Shamoon'
+\item \textbf{shared files or spreading through SMB} (shared message block or  Common Internet File System (CIFS)) Flame
+\item \textbf{Zero-day vulnerabilities} \begin{description}
+\item Crouching Yeti is hardly a sophisticated campaign.
+For example, the attackers used no zero-day exploits, only exploits that are widely available on the Internet.
+But that did not prevent the campaign from staying under the radar for several years.
+\url{http://www.kaspersky.com/internet-security-center/threats/crouching-yeti-energetic-bear-malware-threat}
 \end{description}
-
---dit stuk tekst overbodig--
-
-\section{Graph models}
-For some of the spreading methods, the graph of the network matters.
-It is important to have the right topology for the right method.
-Email worms need a topology that represent a social network, BGP routing worms need a topology on network level.
---iets zeggen dat het ook belangrijk is voor de defender om zijn netwerk zo aan te passen dat het virus moeilijker kan verspreiden. Network segregation mss aanhalen ?--
-
-\begin{description}
-\item Power law
-\item Small world topology
-\item random graph
-\item ..
 \end{description}
@@
 \section{Models for worm propagation}
+\label{modelsforpropagation}
 To use the models for worm propagation in our model of the FlipIt game it is suffi"ent to know the propagation speed of the worm and define after how many time units the network is defected.
@@
 
+\subsubsection{Simple Epidemic model}
+
+\subsubsection{Kermack-Mckendrick model}
+
 \subsubsection*{SIS}
@@
 \subsubsection*{SIR}
+
+\subsubsection{Two-Factor model}
 
@@
 
-\subsection{Weetjes}
+\section{methode met matrixen}
+\label{eigenmatrixmethode}
+
+or some of the spreading methods, the graph of the network matters.
+It is important to have the right topology for the right method.
+Email worms need a topology that represent a social network, BGP routing worms need a topology on network level.
+--iets zeggen dat het ook belangrijk is voor de defender om zijn netwerk zo aan te passen dat het virus moeilijker kan verspreiden. Network segregation mss aanhalen ?--
+
 \begin{description}
-\item Crouching Yeti is hardly a sophisticated campaign.
-For example, the attackers used no zero-day exploits, only exploits that are widely available on the Internet.
-But that did not prevent the campaign from staying under the radar for several years.
-\url{http://www.kaspersky.com/internet-security-center/threats/crouching-yeti-energetic-bear-malware-threat}
+\item Power law
+\item Small world topology
+\item random graph
+\item ..
 \end{description}
 
-\section{methode met matrixen}
-
+The usefulness of this method is that it is graph model independent. The topology of the network has to be in the form of a square matrix. Each connection is determined by the entries on the matrix. 
 %$N_{0}$ denotes the initially infected resources at the beginning of the virus propagation. 
@@
 %So the chance that a resource is infected after time t is the chance that the resource is infected by all the previous infections and that the defender has not flipped the resource.
-source: \url{http://en.wikipedia.org/wiki/Adjacency_matrix} \\
-We model the network through an undirected Graph $G = < V, E> $ where $|V|$ denotes the number of resources in the network and $|E|$ the number of connections.
-We can convert this to an adjacency matrix which represents which vertices of the graph are neighbors of other vertices.
+
+The computer network can be modelled by an undirected Graph $G = < V, E> $ where $|V|$ denotes the number of resources in the network and $|E|$ the number of connections.
+We can convert this to an adjacency matrix which represents which vertices of the graph are neighbours of other vertices.
 \\
@@
 Because the graph is undirected we have a symmetric matrix.
+\\
+
 Adjacency matrices have many interesting applications, amongst which calculating the paths between vertices:
-\textit{"If \textit{A} is the adjacency matrix of the directed or undirected graph \textit{G}, then the matrix $A^{n}$ (i.e., the matrix product of n copies of \textit{A}) has an interesting interpretation: the entry in row i and column j gives the number of (directed or undirected) walks of length n from vertex i to vertex j.
-If n is the smallest nonnegative integer, such that for all i ,j , the (i,j)-entry of $A^{n} > 0$, then n is the distance between vertex i and vertex j."} [Wikipedia]
+\textit{``If \textit{A} is the adjacency matrix of the directed or undirected graph \textit{G}, then the matrix $A^{n}$ (i.e., the matrix product of \textit{n} copies of \textit{A}) has an interesting interpretation: the entry in row \textit{i} and column \textit{j} gives the number of (directed or undirected) walks of length \textit{n} from vertex \textit{i} to vertex \textit{j}.
+If n is the smallest nonnegative integer, such that for all i ,j , the (i,j)-entry of $A^{n} > 0$, then n is the distance between vertex i and vertex \textit{j}.''}  source: \cite{wikimatrix} .\\
 
@@
 
-Using this knowledge we can calculate in how many steps a node is infected: 
+Using this knowledge we can calculate in how many steps a node is infected: \\
 The infection vector or start vector I0 of lenght $|V|$ indicates which node is infected and which not.
@@
 This technique can be applied to calculate the state of the network after any number of steps.
+\\
+
 Determining the length of the delay boils down to determining which configuration of infected nodes is considered as corresponding to the attacker having flipped the resource.
-It may be that all nodes need to be flipped, a sufficient amount of nodes, or a (set of) particular nodes.
---Hier nog een beetje verder uitschrijven hoe je exact de d dan moet bepalen, gegeven dat je niet weet in welke knoop het virus gedropt wordt--
+It may be that all nodes need to be flipped, a sufficient amount of nodes, or a (set of) particular nodes.\\
+
+--Hier nog een beetje verder uitschrijven hoe je exact de d dan moet bepalen, gegeven dat je niet weet in welke knoop het virus gedropt wordt-- \\
 

### 2015-06-23 doc/template/chap-9.tex
Jeroen nalees

@@
 \label{Ch2:Intro}
-The game of FlipIt with propagation delay considers a game where the moves of the attacker are not instantaneous.
-This corresponds to the case of APT's where attacks are performed by means of malware, such as viruses, worms and trojan horses and the malware needs time to propagate.
+The FlipIt game with propagation delay considers a game where the moves of the attacker are not instantaneous.
+This corresponds to APT's which use malware - viruses, worms or trojan horses - to perform attacks, as the malware needs time to propagate.
 A virus, for example, can be dropped on a network but it only compromises the whole network if every node in the network is infected.
@@
 The last section \ref{ch2:periodicvirus} derives a formula to calculate the benefit for a FlipIt game with propagation delay.
-In the next Chapter, this benefit formula will be used to determine through nash equilibria what the best defence and attack strategies are.
+In the next Chapter, this benefit formula will be used to determine what the best defence and attack strategies are.
 %Section \todo{nash} presents the Nash equilibrium for the benefit formula. The chapter concludes in section \todo{section toevoegen} with lessons learned en practical implications for cyber security.
@@
 
-Below a list of required adaptations that should be made to the FlipIt game to model a game with propagation delay or motivations to keep the assumptions made by FlipIt.
+The following paragraphs will list the required adaptations to model the basic FlipIt game with propagation delay.
 
@@
 The attacker will do this by dropping a virus on a node on the network. The virus will then spread itself and infect other nodes in the network. \\
-By defining the resource as a network with multiple nodes it is possible to extend the actions of the defender and the attacker to actions depending on which node is being flipped.
+By defining the resource as a network with multiple nodes it is possible to increase the number of possible actions by the defender or attacker.
 These actions will be explained in the following subsection.
@@
 
-The action of the defender consist of defending his network.
 The network is composed of multiple nodes.
 The defender can choose to flip one node, a subset of nodes or all nodes of the network.
-In this paper, the defender flips all the nodes of the network in every move he plays.
-The action of the defender can be extended to the action where the defender only flips a subset of the nodes or a single node of the network.
+In this paper, the defender flips all the nodes in the network every time he plays.
+This action can be extended to flipping only a subset of the nodes or a single node of the network.
 This extension has already been investigated in the context of placing anti-virus systems in a graph.
@@
 \\ \todo{niet volledig juist, defender systemen geplaatst en dan gezien hoe ver de worm kan geraken}
-The defender has only one action.
-Sending different viruses and worms to a computer network.
+The attacker only has one action: sending different kinds of malware to a computer network.
 Every time the attacker flips the resource, he does this by sending a new kind of malware e.g worm to the system.
-The attacker does not send the same malware again, because if the computers are patched, the malware cannot get in any more.
-Sending malware in this context is equal to flipping a node in the network.
+The attacker does not send the same malware again, because if the computers are patched, the malware will be unable to penetrate the system.
 The node can be a targeted node or a random node.
-If the attacker has knowledge about the graph, the attacker can choose to target the node that can infect all the nodes in the shortest time possible.
+If the attacker has knowledge about the topology of the network, the attacker can choose to target a specific node.
+Most likely, this will be the node that can infect all other nodes in a minimum timespan.
 \subsection{Immediate effect of the move}
-The framework of FlipIt is used to model targeted attacks.
+%The framework of FlipIt is used to model targeted attacks.
 In FlipIt when an attacker attacks, he will get immediately control over the resource.
 One of the differences is that FlipIt does not consider that moves of the attacker might not be instantaneous.
-With FlipIt with a propagation delay we model the fact that if the attacker flips, he does not get immediate control over the resource.
+%With FlipIt with a propagation delay we model the fact that if the attacker flips, he does not get immediate control over the resource.
 This is motivated by the fact that it takes a while before a targeted attack can compromise a computer network.
@@
 The attacker has only control of the network once all or a sufficient number of nodes are infected.
-The time that it takes for the virus to infect every node (or a sufficient number of nodes) will be
-denoted as an propagation delay variable \textit{d} (called 'delay' for short in the remainder of this paper). If we want to measure how long it takes for the virus to
-infect all the nodes in the network, we have to calculate the shortest path from the
+The time that it takes for a piece of malware to infect every node (or a sufficient number of nodes) will be
+denoted as a propagation delay variable \textit{d} (called 'delay' for short in the remainder of this paper).
+If we want to measure how long it takes for the malware to infect all the nodes in the network, we have to calculate the shortest path from the
 first infected node to the farthest node.
@@
 It is assumed that if a defender tries to clean the network that this will happen without a delay.
-The defender can de-plug the computer from the network, or just clean a pc.
+The defender can de-plug the computer from the network, push a security update to all pc's, or even format a pc.
 \subsection{Stealth character of the move}
@@
 The main characteristic of an APT is that the attack is stealthy.
-Depending on how the attacker has set up his attack, and depending on the network if it is connected with the Internet or not, the moves of the defender are stealthy or not stealthy.
+Depending on how the attacker has set up his attack, and depending on whether or not the network is connected with the Internet, the moves of the defender are stealthy or not stealthy.
 For example, if the attacker launches an APT only to harm the system and not to receive feedback, the moves of the defender are stealthy.
@@
 Some APT's can be resident on a isolated network (e.g.
-Stuxnetworm), meaning that there was no way to connect to the internet.
+Stuxnetworm), meaning that there is no way to connect to the internet.
 %Another example is by the use of honeypots. 
@@
 Honeypots do not detect all malicious attacks so the attacker can stay unnoticed and can still provide information as feedback.
-This is a way how the defender can also be stealthy by providing  false information through the honeypot.
-Furhtermore, the case where the moves of the defender are non stealthy has already been investigated by Laszka \citep{MitigationNonTargeted}.
+Furthermore, the case where the moves of the defender are non stealthy has already been investigated by Laszka \citep{MitigationNonTargeted}.
 
 \subsection{Cost associated with the move}
- The defender will be defending every computer controlled device in a network.
-To make the network clean the defender can patch the systems with the latest security bulletins or reinstall the software or even wipe out the computer.
+ The defender will defend every computer-controlled device in a network.
+In order to clean the network, the defender can patch the systems with the latest security bulletins, reinstall the software or even format the computer.
 All these different actions can imply other costs.
-The cost of patching a system will be smaller than the cost of wiping out a computer.
+The cost of patching a system, for example, will most likely be smaller than replacing a computer.
 \\
-The cost of the attacker depend on the complexity of the vulnerabilities exploited and how difficult it was to program the malware.
+The cost of the attacker depends on the complexity of the vulnerabilities exploited and how difficult it was to program the malware.
 \subsection{Strategies}
-The formalization starts from the model of the non-adaptive continuous basic FlipIt game where players use a periodic strategy with a random phase. \\
+The formal definition of the adapted FlipIt game starts from the model of the non-adaptive continuous basic FlipIt game where players use a periodic strategy with a random phase. \\
 
 \textbf{No feedback (non-adaptive)}\\
-Non-adaptive strategies are taken because of the assumption that both players will receive no feedback during the game.
+Non-adaptive strategies are chosen because of the assumption that both players will receive no feedback during the game.
 This is motivated by the fact that the identity of the attackers and their attack strategy is rarely known to the defenders.
 In the case of the attacker, if an attacker wants to have feedback it may be disadvantageous if he wants to stay undetected.
-\\
+\todo{meer uitleg over waarom het een nadeel voor de attacker is om geen info terug te vragen}\\
 
@@
 This is true for the example of patching.
-Companies like Microsoft en Google release at a specific moment patches.
-So companies will apply these patches when they come out.
+Companies like Microsoft and Google release their security patches at fixed intervals, so companies will apply these patches as they come out.
 Microsoft security bulletins are released on the Second Tuesday of each month.
@@
 
-In the basic FlipIt game \citep{FlipIt} the authors concluded that a periodic strategy is the dominant strategy against all other renewal strategies if the opponent uses a periodic or non-arithmetic renewal strategy.
-This means that a good choice for the attacker is also to play a periodic strategy.
-Furthermore, a periodic attacker strategy is assumed as well, to be able to compare the results with the periodic strategy of the FlipIt game in \cite{FlipIt}.
+In the basic FlipIt game \citep{FlipIt} the authors\todo{blabla et.
+al} concluded that a periodic strategy is the dominant strategy against all other renewal strategies if the opponent uses a periodic or non-arithmetic renewal strategy.
+So regardless of whether the defender chooses to play periodically or not, it's a good choice for the attacker to play periodically as well.
+\\
 
-The paper from Laszka \citep{MitigationNonTargeted}, which was similar to this one, also concluded that a periodic strategy is the dominant strategy against all other non-adaptive strategies.
+The paper from Laszka \citep{MitigationNonTargeted}, which is similar to this one, also concluded that a periodic strategy is the dominant strategy against all other non-adaptive strategies.
 %This corresponds to a periodic defender strategy.   %, as this also corresponds to a common real life strategy. %nog eventueel verder te motiveren
@@
 
-\subsubsection{Motivation Application: Wormlike ATP}
-A business is every day under attack of different threats.
-An APT that successfully enters a companies network can cause a lot of damage.
+%\subsubsection{Real world example: Wormlike ATP}
+%Businesses are under constant attack of different threats.
+If an APT manages to successfully enters a companies network can cause a lot of damage.
 In this example the ATP will be an attacker that sends worms with stealthy features to attack a company network.
@@
 \\
-The worm can use the computers on the network as botnets or steal sensitive data.
+%The worm can use the computers on the network as botnets or steal sensitive data.
 By patching or installing new software the worm can be removed form the systems.

### 2015-06-26 doc/template/chap-4.tex
vanalles

@@
 The worm selects randomly a part of the selected address space instead of scanning the whole address space.
-The reserved address blocks and the unassigned address are excluded from the address space.
+The reserved address blocks and the unassigned addresses are excluded from the address space.
 The rate of success for randomly chosen IP addresses is very low, but it is easy to implement.
@@
 \item \textbf{Topological scanning:} Address information stored in the victim machines are used to locate new targets. This method used by the `Morris' worm.
-\item \textbf{Hit-list scanning:} With the method of hit-list scanning, a short-list of vulnerable systems is made beforehand.  This is done to speed up
-the spread of worms at the initial stage. --This list consists
-of potentially vulnerable machines that are gathered
-beforehand and targeted first when the worm
-is released. An extreme case for the hit-list-scanning
-worms is a flash worm, which gathers all vulnerable
-machines into the list.--
+%\item \textbf{Hit-list scanning:} With the method of hit-list scanning, a short-list of vulnerable systems is made beforehand.
+This is done to speed up the spread of worms at the initial stage.
+--This list consists of potentially vulnerable machines that are gathered beforehand and targeted first when the worm is released.
+An extreme case for the hit-list-scanning worms is a flash worm, which gathers all vulnerable machines into the list.--
 \item \textbf{Sequential scanning:}
+With sequential scanning, the worm will scan the IP addresses sequentially.
+This means that once a vulnerable host is compromised, it will look for IP addresses that are near to this host.
+For example, the address of the host is A, the next addresses that the worm will scan are A+1, or A-1.
+This method is used by the `Blaster worm'.
 \item \textbf{DNS random scanning:} Another strategy is a kind of strategy in which the DNS infrastructure is used to locate a new target address.
@@
 Last, the IP addresses stored in the address table are only hosts with public domain names.
-\item \textbf{Routable scanning:} they use BGP routing tables to scan the routable address space.
+\item \textbf{Routable scanning:} The worms using routable scanning acquire target IP addresses based on the routing information in a network.
+Through  the BGP routing tables they can scan the routable address space.
 This method is three times faster than a traditional worm that uses random scanning.
 Examples are `Spyb0t' or `network Bluepill'.
---a ''routing worm'' [20].
+%--a ''routing worm'' [20].
 Zou et al.
-designed two types
-of routing worms [20]. One type, based on Class-A
-(x.0.0.0/8) address allocations, is thus called 'ClassA
-routing worms.'' Such worms can reduce the scanning
-space to 45.3\% of the entire IPv4 address space.
-The other type, based on BGP routing tables, is thus
-called''BGP routing worms'' Such worms can reduce
-the scanning space to only about 28.6\% of the entire
-IPv4 address space.--
+designed two types of routing worms [20].
+One type, based on Class-A (x.0.0.0/8) address allocations, is thus called 'ClassA routing worms.'' Such worms can reduce the scanning space to 45.3\% of the entire IPv4 address space.
+The other type, based on BGP routing tables, is thus called''BGP routing worms'' Such worms can reduce the scanning space to only about 28.6\% of the entire IPv4 address space.--
 %\end{itemize}
 
-\item \textbf{Email worms:} use the email systems to propagate. The Kak worm is a Javascript computer worm that spread itself by exploiting a bug in Outlook Express.  Another example is `GhostNet'
+\item \textbf{Email and other client application worms:} An email worms uses the email systems to find email addresses to propagate.
+Other client applications can include: Internet Relay Chat (IRC), Instant Messenger (IM), and a variety of peer-to-peer
+file sharing systems, which have been used by worms to propagate in a similar way as email worms.
+For example, the Kak worm is a Javascript computer worm that spread itself by exploiting a bug in Outlook Express.
+% Another example is `GhostNet'
 
@@
 [5]
-\item \textbf{self stopping worm:} reduces speed to avoid detection. Atak worm or self stopping worm \cite{VirusChangePropSpeed}.
-\item \textbf{network sharing} `Shamoon'
-\item \textbf{shared files or spreading through SMB} (shared message block or  Common Internet File System (CIFS)) Flame
-\item \textbf{Zero-day vulnerabilities} \begin{description}
-\item Crouching Yeti is hardly a sophisticated campaign.
+%\item \textbf{self stopping worm:} reduces speed to avoid detection. Atak worm or self stopping worm \cite{VirusChangePropSpeed}.
+%\item \textbf{network sharing} `Shamoon'
+%\item \textbf{shared files or spreading through SMB} (shared message block or  Common Internet File System (CIFS)) Flame
+%\item \textbf{Zero-day vulnerabilities} \begin{description}
+%\item Crouching Yeti is hardly a sophisticated campaign.
 For example, the attackers used no zero-day exploits, only exploits that are widely available on the Internet.
@@
 \end{description}
-\end{description}
 
@@
 \label{modelsforpropagation}
-To use the models for worm propagation in our model of the FlipIt game it is suffi"ent to know the propagation speed of the worm and define after how many time units the network is defected.
+To use the models for worm propagation in our model of the FlipIt game it is sufficient to know the propagation speed of the worm and define after how many time units the network is defected.
 This will be equal to the \textit{d} parameter in our model.
@@
 Feasible due to use 32-bit addresses.
-128-bit adresses life harder for worms, except the ones that use email systems to propagate.
+128-bit addresses life harder for worms, except the ones that use email systems to propagate.
 two new strategies: uniformly distributed random number generator to select new target.
@@
 
+All of the above models for modelling the mechanism of the spreading of a worm depend on the kind of topology of the network.
+In this chapter we propose a method to calculate the spread of the worm in way where the topology of the network is easily integrated.
+The method will give an approximation of how fast a worm can infect a network.
+The method is based on the method for Google Page Ranking \todo{ref} [].
 or some of the spreading methods, the graph of the network matters.
@@
 
-The usefulness of this method is that it is graph model independent. The topology of the network has to be in the form of a square matrix. Each connection is determined by the entries on the matrix. 
+%The usefulness of this method is that it is graph model independent. The topology of the network has to be in the form of a square matrix. Each connection is determined by the entries on the matrix. 
 %$N_{0}$ denotes the initially infected resources at the beginning of the virus propagation. 
@@
 \textit{``If \textit{A} is the adjacency matrix of the directed or undirected graph \textit{G}, then the matrix $A^{n}$ (i.e., the matrix product of \textit{n} copies of \textit{A}) has an interesting interpretation: the entry in row \textit{i} and column \textit{j} gives the number of (directed or undirected) walks of length \textit{n} from vertex \textit{i} to vertex \textit{j}.
-If n is the smallest nonnegative integer, such that for all i ,j , the (i,j)-entry of $A^{n} > 0$, then n is the distance between vertex i and vertex \textit{j}.''}  source: \cite{wikimatrix} .\\
+If \textit{n} is the smallest nonnegative integer, such that for all i ,j , the (i,j)-entry of $A^{n} > 0$, then n is the distance between vertex i and vertex \textit{j}.''}  source: \cite{wikimatrix} .\\
 
@@
 %%% End: 
+\begin{figure}
+\centering
 \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2.8cm,
@@
 \end{tikzpicture}
-\\
+\caption{Network with 6 nodes. The arrows represent the connections between the nodes. The start point is were the worm has been dropped, here node 1.}
+\label{netwerkfiguur}
+\end{figure}
+
 %\[
@@
 %\] 
-Assuming a network like in Fig, the corresponding adjacency matrix is the  matrix $[A]$: \\
+Assuming a network like in figure \ref{netwerkfiguur}, the corresponding adjacency matrix is the  matrix $[A]$: \\
 
@@
 Using this knowledge we can calculate in how many steps a node is infected: \\
-The infection vector or start vector I0 of lenght $|V|$ indicates which node is infected and which not.
-Multiplying the infection vector with $A$ results in a vector I1, which represents which nodes are infected after 1 step.
-Multiplying the start vector I0 with $A^{N}$ calculates which nodes are infected in N steps: each non zero entry represents an node infected after N steps.
+The infection vector or start vector $I_{0}$ of lenght $|V|$ indicates which node is infected and which not.
+Multiplying the infection vector with $A$ results in a vector $I_{1}$, which represents which nodes are infected after 1 step.
+Multiplying the start vector $I_{0}$ with $A^{N}$ calculates which nodes are infected in \textit{N} steps: each non zero entry represents an node infected after \textit{N} steps.
+\\
 In the context of the FlipIt game, it is safe to assume that once a node is infected, it stays infected until the defender Flips the node.
-This means that after d steps, it is assumed that all nodes that could be reached in less than d steps from the node where the worm was dropped, are infected.
+This means that after d steps, it is assumed that all nodes that could be reached in less than \textit{d} steps from the node where the worm was dropped, are infected.
 In order to know how many nodes are infected after (for example) at most 3 steps, we have to consider nodes that are infected initially (step 1), or after 2 steps, or after 3 steps.
-Calculating the sum of the three matrices $(A + A^{2} + A^{3}) $ results in a matrix that indicates for each node, the number of paths of length 1, 2 or 3 from i to j.
+Calculating the sum of the three matrices $(A + A^{2} + A^{3}) $ results in a matrix that indicates for each node, the number of paths of length 1, 2 or 3 from \textit{i} to \textit{ j}.
 Multiplying the start vector with this matrix, results in a vector that indicates which nodes will be infected in at most 3 steps .

### 2015-06-26 doc/template/chap-9.tex
vanalles

@@
 Stuxnetworm), meaning that there is no way to connect to the internet.
+\\
 %Another example is by the use of honeypots. 
@@
 In the case of the attacker, if an attacker wants to have feedback it may be disadvantageous if he wants to stay undetected.
-\todo{meer uitleg over waarom het een nadeel voor de attacker is om geen info terug te vragen}\\
+If the attacker wants to have feedback, the malware needs to make a connection with the attacker to provide this feedback.
+For example, this can cause more network traffic for the defender to detect.
+\\
 
@@
 
-In the basic FlipIt game \citep{FlipIt} the authors\todo{blabla et.
-al} concluded that a periodic strategy is the dominant strategy against all other renewal strategies if the opponent uses a periodic or non-arithmetic renewal strategy.
+In the basic FlipIt game \citep{FlipIt} the authors van Dijk et al.
+concluded that a periodic strategy is the dominant strategy against all other renewal strategies if the opponent uses a periodic or non-arithmetic renewal strategy.
 So regardless of whether the defender chooses to play periodically or not, it's a good choice for the attacker to play periodically as well.

### 2015-06-28 doc/template/chap-4.tex
delay methods aanvullen

@@
 Different dropping mechanisms can be used, whereby the initial attack can be random or targeted at a specific node.
-Frequently used mechanisms are USB sticks (given to a specific person or left behind for pick up by a random person), email, or malicious software through phising or trojan horses.
+Frequently used mechanisms are USB sticks (given to a specific person or left behind for pick up by a random person), email, or malicious software through fishing or trojan horses.
 More important for determining our delay is the propagation strategy.
@@
 
-\item \textbf{Email and other client application worms:} An email worms uses the email systems to find email addresses to propagate.
+\item \textbf{Topology-based Wroms} Email and other client application worms:An email worms uses the email systems to find email addresses to propagate.
 Other client applications can include: Internet Relay Chat (IRC), Instant Messenger (IM), and a variety of peer-to-peer
@@
 
-Kind of network model: 
-Kind of model: SIS, SIR
+SI, SIR, SIS assume $\beta$ to be constant.
 
+Continious-time model
 \subsubsection{Simple Epidemic model}
+%Codered paper gebruikt voor model
+A Simple epidemic model is SI model. Nodes in the network can be either susceptible or infected.
+Once a node is infected it will stay infected. 
+The simple epidemic model is considered to be of a fixed size. The model for a fixed population is as follows:
+\begin{equation}
+\dfrac{d I(t)}{dt} = \beta I(t)[N-I(t)]
+\end{equation} 
+where I(t) is the number of infected hosts at time t; $\beta$ is the propagation rate; and N is the number of nodes in the network.
+In the beginning, $t=0$, I(0) hosts are infected.
+All the other nodes, $N - I(0)$, in the network are susceptible.
+The solution of this equation is the following logistic curve:
+\begin{equation}
+I = \dfrac{e^{\beta(t-T)}}{1+e^{\beta(t-T)}}
+\end{equation}
+where T is a time parameter representing the point of maximum increase in the
+growth.\\
 
-\subsubsection{Kermack-Mckendrick model}
+\begin{figure}[hbtp]
+\centering
+\includegraphics[scale=0.5]{Images/SEMmodel.png}
+\caption{Simple Epidemic Model. The x-coordinate is the propagation time and the y-coordinate is the infected percentage of the whole network. Original image from \cite{OnWorms2005survey}}
+\label{SEMmodel}
+\end{figure}
 
+Figuur \ref{SEMmodel} laat het SEM model zien.
+This Simple epidemic model has been used in various papers \cite{OwnInternetSI} \cite{CodeRed} to model random scanning worm such as Code Red and Slammer. \\
+For our paper we have to approximate the propagation speed en Infect gelijkstellen aan het aantal genoeg nodes die geinfecteerd moeten zijn door de APT om de controle te hebben.
+Dan weten we hoelang het duurt.
+In het begin zullen er ook geen geinfecteerde hosts zijn, dus de vergelijking kan opgelost worden.\\
+Er wordt geen rekening gehouden met de topology ? nog uitzoeken.
+\subsubsection{Kermack-Mckendrick model: SIR}
+In the Kermack-Mckendrick model, also known as SIR model, nodes can have three states: susceptible, infected or removed.
+Once a node of the network has been recovered from a worm, the node will stay in the removed mode and never becomes infected again.
+These nodes are not able to infect other nodes and can no longer be infected.
+\\
+Let I(t) be the number of infectious hosts at time \textit{t}, R(t) be the number of removed hosts at time \textit{t} and J(t) is the number of infected hosts by time \textit{i}, regardless the fact that a node can be in a removed state.
+\begin{equation}
+J(t) = I(t) + R(t).
+\end{equation} 
+The Kermack-McKendrick model can be represented as follows:
+\begin{equation}
+\begin{Bmatrix} \dfrac{d J(t)}{dt} = \beta J(t) \big[N- J(t) \\
+\dfrac{d R(t)}{dt} = \gamma I(t) \\
+J(t) = I(t) + R(t) = N - S(t)
+ \end{Bmatrix}
+\end{equation}
+Parameter $\beta$ is again the rate of infection and $\gamma$ is the rate of removal of infected hosts. S(T) is the number of susceptible hosts at time t.
+\begin{figure}[hbtp]
+\centering
+\includegraphics[scale=0.5]{Images/KMmodel.png}
+\caption{KM Model.
+The x-coordinate is the propagation time and the y-coordinate is the infected percentage of the whole network.
+Original image from \cite{OnWorms2005survey}.
+$N=10000, \beta = 1/10000000$.}
+\label{KMmodel}
+\end{figure}
+
+%codered%
+--Define $\rho = \gamma/\beta$ to be the relative removal rate [3]. One
+interesting result coming out of this model is
+dI(t)
+dt > 0 if and only if S(t) > p. (6)
+Since there is no new susceptible host to be generated,
+the number of susceptible hosts S(t) is a monotonically decreasing
+function of time t. If S(t0) < p, then S(t) < p and
+dI(t)/dt < 0 for all future time t>t0. In other words, if the
+initial number of susceptible hosts is smaller than some critical
+value, S(0) < p, there will be no epidemic and outbreak
+[15].
+The Kermack-Mckendrick model improves the classical
+simple epidemic model by considering that some infectious
+hosts either recover or die after some time. However, this
+model is still not suitable for modeling Internet worm propagation.
+First, in the Internet, cleaning, patching, and filtering
+countermeasures against worms will remove both susceptible
+hosts and infectious hosts from circulation, but KermackMckendrick
+model only accounts for the removal of infectious
+hosts. Second, this model assumes the infection rate
+to be constant, which isn't true for a rampantly spreading
+Internet worm such as the Code Red worm--
 \subsubsection*{SIS}
 
-\subsubsection*{SIR}
-
 \subsubsection{Two-Factor model}
-
+%distributed worm simulation
+--Zou et
+al. [15] present a ''two-factor'' model that extends SIR
+epidemiological model to capture the effects of human
+countermeasures and the congestion due to the worm
+spread. Shen et al. [16] provide a discrete-time worm
+model that considers patching and cleaning effect and can
+model worms with local scanning techniques. All these
+approaches abstract specifics of the Internet topology,
+change in the size of vulnerable population as the worm
+spreads, and the effect of the individual host and network
+defenses (except for patching, in case of [16]) on the
+spread--
 \subsubsection*{self disciplinary worms}
@@
 
-Popular mechanism that worms use to detect vulnerable targets by random ip scanning probing.
+Popular mechanism that worms use to detect vulnerable targets by random IP scanning probing.
 Feasible due to use 32-bit addresses.

### 2015-06-28 doc/template/chap-10.tex
Nog verdere aanvullingen

@@
 The organisation of this paper is the following.
-In chapter \ref{Chapter1:Intro.Game.Theory}  a brief introduction to cyber security and game theory is introduced to get familiar with the kind of threats that are in the scope of this work and the game theoretic concepts that will be further used in the paper.
+An introduction to cyber security and game theory is given in chapter \ref{Chapter1:Intro.Game.Theory} to get familiar with the kind of threats that are in the scope of this work and the game theoretic concepts that will be further used in the paper.
 In the same chapter the FlipIt framework is summarized with its most important conclusions.
@@
 \\
-In Chapter \ref{chapter:Nash} the formules are further analysed to determine if there is a nash equilibrium. \todo{aanpassen naar chapter met nash equilibria}
-In order to provide a clear perspective on delays, chapter \ref{chapter4: Virus propagation} gives an overview of the various ways in which a virus can propagate.
-It presents a method to calculate the speed of the propagation of a virus in a network and how this network can be established to reduce the spreading of a virus.
+In Chapter \ref{chapter:Nash} the formulas are further analysed to determine if there is a nash equilibrium. 
+In order to provide a clear perspective on delays, chapter \ref{chapter4: Worm propagation} gives an overview of the various methods of propagation and worm propagation models.
+It presents a method to calculate the speed of the propagation of a worm in a network independent of the topology of the network.
 Finally chapter \ref{chapter5:conclusion} , we discusses the main results and provides directions for further research.

### 2015-06-28 doc/template/chap-4.tex
Nog verdere aanvullingen

@@
 Worms on the other hand, spread without human intervention, and are their propagation is therefore easier to model.
-Since the purpose of this chapter is only to illustrate how a delay can be calculated, we will limit the presentation to propagation models of worms .
-Section \ref{methodsofpropagation} presents an overview of the most frequently used propagation methods by worms.
+Since the purpose of this chapter is only to illustrate how a delay can be calculated, we will limit the presentation to propagation models of worms.
+\\
+
+The overview of this chapter will be as follow: Section \ref{methodsofpropagation} presents an overview of the most frequently used propagation methods by worms.
 These propagation methods will be covered by different kinds of models in section \ref{modelsforpropagation} that presents models to determine the time a worm needs to infect a sufficient number of nodes.
@@
 
+The following papers cover the most common propagation models: \cite{wang2014modeling}, \cite{OnWorms2005survey}, \cite{xiang2009propagation} and \cite{serazzi2004computer}.
+In this section we only list the models that are the most interesting from the perspective of calculating the delay.
 SI, SIR, SIS assume $\beta$ to be constant.
 
+Tree model figure \ref{tree}
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.5]{Images/TreeModel.png} 
+\caption{Taxonomy of worm modeling. Original tree from \cite{wang2014modeling}}.
+\label{ShapeUp}
+\end{figure}
 Continious-time model
@@
 Figuur \ref{SEMmodel} laat het SEM model zien.
-This Simple epidemic model has been used in various papers \cite{OwnInternetSI} \cite{CodeRed} to model random scanning worm such as Code Red and Slammer. \\
+This Simple epidemic model has been used in various papers \cite{OwnInternetSI}, \cite{CodeRed} to model random scanning worm such as Code Red and Slammer. \\
 For our paper we have to approximate the propagation speed en Infect gelijkstellen aan het aantal genoeg nodes die geinfecteerd moeten zijn door de APT om de controle te hebben.

### 2015-06-28 doc/template/nash.tex
figuren toegevoegd
versie voor mama
alpha -> delta
nash aanvullen
Nog verdere aanvullingen

@@
 \begin{figure}[hbtp]
+\centering
+\includegraphics[scale=0.4]{Images/bestresp.png}
 \caption{This figure shows the three subcategories of each case.
 'A' stands for Case 2.A: $\delta_{D} \geq d+\delta_{A} \geq \delta_{A}$ and 'B' stands for Case 2.B: $d+\delta_{A} \geq \delta_{D} \geq  \delta_{A} $ }
-\centering
-\includegraphics[scale=0.4]{Images/bestresp.png}
 \label{grafiekbestr}
@@
 
-Nash equilibria are points with the property that neither player benefits by deviating in isolation form equilibrium.
-We can compute Nash Equilibria for the periodic game as an intersection points of curvest opt0 and opt1.
-To determine opt0(a0) we need to compute the derivative of  $\beta_{0}(\alpha_{0},\alpha_{1}) $ for a fixed $\alpha_{1}$.
+Nash equilibria are points with the property that neither player benefits by deviating in isolation from the equilibrium.
+We can compute Nash Equilibria for the periodic game as an intersection point of curves $opt_{D}$ and $opt_{A}$.
+To determine $opt_{D}(\alpha_{A})$ we need to compute the derivative of  $\beta_{D}(\alpha_{D},\alpha_{A}) $ for a fixed $\alpha_{A}$.
 We consider two cases:\\
@@
 \subsection*{Case 1: $\delta_{D} \leq \delta_{A} $}
-The benefit formula obtained previously for this case is as follows:
+The benefit formula obtained in the previous chapter for this case is as follows:
 \begin{equation*}
-\beta_{D}(\alpha_{D},\alpha_{A}) = 1 - \dfrac{\delta_{D}}{2\delta_{A}} - \dfrac{k_{D}}{\delta_{D}} - \dfrac{d^{2}}{2\delta_{D}\delta_{A}} + \dfrac{d}{\delta_{A}}
+\beta_{D}(\delta_{D},\delta_{A}) = 1 - \dfrac{\delta_{D}}{2\delta_{A}} - \dfrac{k_{D}}{\delta_{D}} - \dfrac{d^{2}}{2\delta_{D}\delta_{A}} + \dfrac{d}{\delta_{A}}
 \end{equation*}
-If we take the partial derivative for a fixed $\delta_{1}$ we get the following result:
+If we take the partial derivative for a fixed $\delta_{A}$ we get the following result:
 \begin{equation*}\label{formdelta}
-\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} = - \dfrac{1}{2\delta_{A}} + \dfrac{k_{D}}{\delta_{D}^{2}} + \dfrac{d^{2}}{2\delta_{D}^{2}\delta_{A}}
+\frac{\partial \beta_{D}(\delta_{D},\delta_{A})}{\partial \delta_{D}} = - \dfrac{1}{2\delta_{A}} + \dfrac{k_{D}}{\delta_{D}^{2}} + \dfrac{d^{2}}{2\delta_{D}^{2}\delta_{A}}
 \end{equation*}
@@
 \begin{equation*}
-\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} =0 ~~~~~~ =>~~~~~~ \delta_{D} = \sqrt{2\delta_{A}k_{D} + d^{2}}
+\frac{\partial \beta_{D}(\delta_{D},\delta_{A})}{\partial \delta_{D}} =0 ~~~~~~ =>~~~~~~ \delta_{D} = \sqrt{2\delta_{A}k_{D} + d^{2}}
 \end{equation*}
 
-This leads to the following deduction: The function increases on $[0, \sqrt{2\delta_{A}k_{D} + d^{2}}]$ and is decreasing on $[\sqrt{2\delta_{A}k_{D} + d^{2}}, \infty]$ So we got a maximum on $\delta_{D} = min \{ \delta_{A}, \sqrt{2\delta_{A}k_{D} + d^{2}} \} $.
-The minimum of the two values is needed because $\delta_{D}$ cannot be bigger than $\delta_{A}$.
+This leads to the following deduction given the sign of the coefficient of $\delta_{D}^{2}$: The function increases on $[0, \sqrt{2\delta_{A}k_{D} + d^{2}}]$ and is decreasing on $[\sqrt{2\delta_{A}k_{D} + d^{2}}, \infty]$.
+So we have a maximum at $\delta_{D} = min \{ \delta_{A}, \sqrt{2\delta_{A}k_{D} + d^{2}} \} $.
+Taking the minimum of the two values is needed because $\delta_{D}$ cannot be larger than $\delta_{A}$.
 \\
@@
 
-The benefit formula obtained previously for this case is as follows:
+The benefit formula obtained in the previous chapter for this case is as follows:
 \begin{equation*}
-\dfrac{t \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - \dfrac{k_{D}}{\delta_{D}}
+\beta_{D}(\delta_{D},\delta_{A})= \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - \dfrac{k_{D}}{\delta_{D}} = \dfrac{\delta_{A} + 2 (d-k_{D})}{\delta_{D}}
 \end{equation*}
 
-The derivative of the above formula for a fixed $\delta_{A}$ results in the following:
-\begin{equation*}
-\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} = -\dfrac{\delta_{A}}{2\delta_{D}^{2}} - \dfrac{d}{\delta_{D}^{2}} + \dfrac{k_{D}}{\delta_{D}^{2}}
-\end{equation*}
-The obtain the stationary points the first derivative is set equal to zero and the roots of the resulting equation are found:
-\begin{equation*}
-\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} =0 ~~~~~~ =>~~~~~~ \delta_{A} = 2(k_{D}-d) = dk_{D} - 2d
-\end{equation*}
+Given that $\delta_{D}$ is always positive, the benefit function can be either positive or negative depending on the numerator of the above fraction.
+For $\delta_{A} + 2(d-k_{D}) > 0$ the benefit will be always positive but decreasing, see figure \ref{ShapeUp}.
+\\
+The defender will play if $ \delta_{A} > 2(k_{D} -d)$ and $k_{D} > d$. The defender will not play if $k_{D} < d$ because then $ \delta_{A}$ is negative.\\
+\begin{figure}
+\centering
+\includegraphics[scale=0.5]{Images/ShapesUp.png} 
+\caption{The benefit function is of the shape of 1/x if $\delta_{D}$ is always positive and if $\delta_{A} + 2(d-k_{D}) > 0$. }
+\label{ShapeUp}
+\end{figure}
 
+For $\delta_{A} + 2(d-k_{D}) < 0$, the benefit will always be negative so the defender will not play. See figure \ref{ShapeDown}. This is independent of the value of $k_{D}$ and $ d$.
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.5]{Images/ShapeDown.png} 
+\caption{The benefit function is of the shape of -1/x if $\delta_{D}$ is always positive and if $\delta_{A} + 2(d-k_{D}) < 0$.}
+\label{ShapeDown}
+\end{figure}
+%The derivative of the above formula for a fixed $\delta_{A}$ results in the following:
+%\begin{equation*}
+%\frac{\partial \beta_{D}(\delta_{D},\delta_{A})}{\partial \delta_{D}} = -\dfrac{\delta_{A}}{2\delta_{D}^{2}} - \dfrac{d}{\delta_{D}^{2}} + \dfrac{k_{D}}{\delta_{D}^{2}}
+%\end{equation*}
+%The obtain the stationary points the first derivative is set equal to zero and the roots of the resulting equation are found:
+%\begin{equation*}
+%\frac{\partial \beta_{D}(\delta_{D},\delta_{A})}{\partial \delta_{D}} =0 ~~~~~~ =>~~~~~~ \delta_{A} = 2(k_{D}-d) = dk_{D} - 2d
+%\end{equation*}
+%
 This leads to the following deduction:
@@
 \begin{description}
-\item decreasing $ \delta_{A} < 2(k_{D} -d)$
+\item decreasing $ \delta_{A} < 2(k_{D} -d)$, $ \delta_{A}$ is negative
 \item increasing  $\delta_{A} > 2(k_{D} -d)$ 
@@
 \begin{equation*}
-\dfrac{\beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - \dfrac{k_{D}}{\delta_{D}} - \dfrac{(d-(\delta_{D} - \delta_{A}))^{2}}{d\delta_{D}\delta_{A}}
+\dfrac{\beta_{D}(\alpha_{D},\alpha_{A})}{\partial \delta_{D}} = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - \dfrac{k_{D}}{\delta_{D}} - \dfrac{(d-(\delta_{D} - \delta_{A}))^{2}}{2\delta_{D}\delta_{A}}
 \end{equation*}
@@
 \begin{equation*}
-\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} =  - \dfrac{1}{2\delta_{D}} + \dfrac{k_{D}}{\delta_{D}^{2}} + \dfrac{d^{2}}{2\delta_{D}^{2}\delta_{A}}
+\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \delta_{D}} =  - \dfrac{1}{2\delta_{D}} + \dfrac{k_{D}}{\delta_{D}^{2}} + \dfrac{d^{2}}{2\delta_{D}^{2}\delta_{A}}
 \end{equation*}
@@
 \begin{equation*}
-\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} =0 ~~~~~~ =>~~~~~~ \delta_{D} = \sqrt{2\delta_{A}k_{D} + d^{2}}
+\frac{\partial \beta_{D}(\delta_{D},\delta_{A})}{\partial \delta_{D}} =0 ~~~~~~ =>~~~~~~ \delta_{D} = \sqrt{2\delta_{A}k_{D} + d^{2}}
 \end{equation*}
 
-This leads to the following deduction: The function increases on $[0, \sqrt{2\delta_{A}k_{D} + d^{2}}]$ and is decreasing on $[\sqrt{2\delta_{A}k_{D} + d^{2}}, \infty]$ So we got a maximum on $\delta_{D} = minimum \{ \delta_{A}, \sqrt{2\delta_{A}k_{D} + d^{2}} \} $ \\
+This leads to the following deduction: The function increases on $[0, \sqrt{2\delta_{A}k_{D} + d^{2}}]$ and is decreasing on $[\sqrt{2\delta_{A}k_{D} + d^{2}}, \infty]$.
+So there is a maximum on $\delta_{D} = minimum \{ \delta_{A}, \sqrt{2\delta_{A}k_{D} + d^{2}} \} $ \\
 
@@
 \begin{equation*}
-\frac{\partial \beta_{A}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} =0 ~~~~~~ =>~~~~~~ 2k_{A} = \dfrac{(\delta_{D}-d)^{2}}{\delta_{D}}
+\frac{\partial \beta_{A}(\delta_{D},\delta_{A})}{\partial \delta_{D}} =0 ~~~~~~ =>~~~~~~ 2k_{A} = \dfrac{(\delta_{D}-d)^{2}}{\delta_{D}}
 \end{equation*}
@@
 \begin{equation*}
-\frac{\partial \beta_{A}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} =0 ~~~~~~ =>~~~~~~ \delta_{A} = \sqrt{2\delta_{D}k_{A}}
+\frac{\partial \beta_{A}(\delta_{D},\delta_{A})}{\partial \delta_{D}} =0 ~~~~~~ =>~~~~~~ \delta_{A} = \sqrt{2\delta_{D}k_{A}}
 \end{equation*}
@@
 \begin{equation*}
-\frac{\partial \beta_{A}(\alpha_{D},\alpha_{A})}{\partial \alpha_{D}} =0 ~~~~~~ =>~~~~~~ 2k_{A} = \dfrac{(\delta_{D}-d)^{2}}{\delta_{D}}
+\frac{\partial \beta_{A}(\alpha_{D},\alpha_{A})}{\partial \delta_{D}} =0 ~~~~~~ =>~~~~~~ 2k_{A} = \dfrac{(\delta_{D}-d)^{2}}{\delta_{D}}
 \end{equation*}
@@
 \end{itemize}
+
+This brings us to the following opt function for : \todo{juiste resultaten hier nog invullen}
+From this analyses we can compute $opt_{D}(\delta_{A})$ for two different cases as:
+ \begin{displaymath}
+  opt_{D}(\delta_{A}) = \left\{
+     \begin{array}{lr}
+       0, & \delta_{A} < 2(k_{D} - d)\\
+       0, & \delta_{A} = 2(k_{D} - d) \\
+       \sqrt{d^{2} + 2\delta_{A}k_{D}}, & \delta_{A} > 2(k_{D} - d)
+     \end{array}
+   \right.
+\end{displaymath}
+
+\todo{ kA met kD vergelijken en nash evenwichten vinden}
+\todo{ een voorbeeld invullen met specifieke waarden}

### 2015-06-30 doc/template/conclusion.tex
alles opslagen
versie voor mama
conclusie

@@
 \chapter{Conclusion}
-\label{cha:conclusion}
-The final chapter contains the overall conclusion. It also contains
-suggestions for future work and industrial applications.
+\label{chapter5:conclusion}
+
+This paper presents an adaption to the basic FlipIt game by \citep{FlipIt} to model an attacker with a delay.  We have
+
+The work presented in this paper represents an initial caclulation of the strategies. 
+\subsection{Further work}
+
+\begin{description}
+\item Analysing other renewal strategies.
+\item Delay for the defender. It takes some time to make a patch if a new exploit is found.
+\item The delay a variable that can change with every Flip.
+\item ..
+\end{description}
 

### 2015-08-03 doc/template/chap-4.tex
nieuwe figuren
nog een updqte

@@
 The  spreading of malware has already been extensively researched.
-Because of the many different types of propagation it is hard to defined a single model that can model all of them.
+Because of the many different types of propagation it is hard to define a single model that can model all of them.
 Modelling the spread of malware depends on two key factors: the method used for the propagation and the graph where the malware will spread itself.
@@
 Since the spreading of viruses requires human interaction, their propagation delay is depending on (hard to predict) human behaviour.
-Worms on the other hand, spread without human intervention, and are their propagation is therefore easier to model.
+Worms on the other hand, spread without human intervention, and their propagation is therefore easier to model.
 Since the purpose of this chapter is only to illustrate how a delay can be calculated, we will limit the presentation to propagation models of worms.
@@
 Finally, in section \ref{eigenmatrixmethode},  we introduce an easy method to calculate the delay of the propagation of a worm.
-Different kind of graphs to model a computer network are introduced.
+% Different kind of graphs to model a computer network are introduced.
 
-This chapter is based on the following papers: \cite{importantjournal}, \cite{OnWorms2005survey}, \cite{GameTheorApprCostBenefitAnalyses} and \cite{SecurelistAPT}.
+%This chapter is based on the following papers: \cite{importantjournal}, \cite{OnWorms2005survey}, \cite{GameTheorApprCostBenefitAnalyses} and \cite{SecurelistAPT}.
 
@@
 
-In the context of malware propagation, there exists two kinds of APT's .
+In the context of malware propagation, there are two kinds of APT's.
 First, there are APT's that launch an attack on just one target node of a network.
@@
 These APT's also use a dropper mechanism but have an additional mechanism for self-propagation in order to propagate themselves to the multiple targeted nodes of the system.
-This mechanism can either be a virus propagation or a worm propagation.
-A virus infects one node on the network and has to wait for human interaction to spread.
+If the APT uses virus spreading, the virus infects one node on the network and has to wait for human interaction to spread.
 So the spreading speed depends on the human interaction, and modelling virus propagation therefore requires modelling human behaviour.
@@
 
-Infection by worms start with dropping the work on the network.
+Infection by worms start by dropping the worm on the network.
 Different dropping mechanisms can be used, whereby the initial attack can be random or targeted at a specific node.
@@
 file sharing systems, which have been used by worms to propagate in a similar way as email worms.
-For example, the Kak worm is a Javascript computer worm that spread itself by exploiting a bug in Outlook Express.
+For example, the `Kak worm' is a Javascript computer worm that spread itself by exploiting a bug in Outlook Express.
 % Another example is `GhostNet'
@@
 This will be equal to the \textit{d} parameter in our model.
-Stackelberg game: first move is from attacker, defender is follower.
+\\
 
-The following papers cover the most common propagation models: \cite{wang2014modeling}, \cite{OnWorms2005survey}, \cite{xiang2009propagation} and \cite{serazzi2004computer}.
-In this section we only list the models that are the most interesting from the perspective of calculating the delay.
-SI, SIR, SIS assume $\beta$ to be constant.
+The methods of propagation can be divided into different model types.
+We are only interested in models that are based on SI model type (Susceptible-Infected), because in our model every node that is infected will stay infected.
+In table \ref{tree} the taxonomy of the propagation methods are given and the type of model.
+\\
 
-Tree model figure \ref{tree}
+For more information about other models the reader is refered to the following papers with other most common propagation models: \cite{wang2014modeling}, \cite{OnWorms2005survey}, \cite{xiang2009propagation} and \cite{serazzi2004computer}.
+%In this section we only list the models that are the most interesting from the perspective of calculating the delay, namely the SI type models.
+%SI, SIR, SIS assume $\beta$ to be constant.
+
+%Tree model figure \ref{tree}. We introduce homogeneous models. Network topology is mixed.
 
@@
 \centering
-\includegraphics[scale=0.5]{Images/TreeModel.png} 
-\caption{Taxonomy of worm modeling. Original tree from \cite{wang2014modeling}}.
-\label{ShapeUp}
+\includegraphics[scale=0.4]{Images/tabel3.png} 
+\includegraphics[scale=0.55]{Images/tableworms.png}
+\caption{Taxonomy of worm modelling. Image based on taxonomy given in \cite{wang2014modeling}}.
+\label{tree}
 \end{figure}
-Continious-time model
+
+\subsection{Continious-time models}
 \subsubsection{Simple Epidemic model}
@@
 A Simple epidemic model is SI model. Nodes in the network can be either susceptible or infected.
-Once a node is infected it will stay infected. 
-The simple epidemic model is considered to be of a fixed size. The model for a fixed population is as follows:
+Once a node is infected it will stay infected. Every node has the same
+chance to be infected. The simple epidemic model is considered to be of a fixed size. The model for a fixed population is as follows:
 \begin{equation}
@@
 Internet worm such as the Code Red worm--
-\subsubsection*{SIS}
 
-\subsubsection{Two-Factor model}
-%distributed worm simulation
---Zou et
-al. [15] present a ''two-factor'' model that extends SIR
-epidemiological model to capture the effects of human
-countermeasures and the congestion due to the worm
-spread. Shen et al. [16] provide a discrete-time worm
-model that considers patching and cleaning effect and can
-model worms with local scanning techniques. All these
-approaches abstract specifics of the Internet topology,
-change in the size of vulnerable population as the worm
-spreads, and the effect of the individual host and network
-defenses (except for patching, in case of [16]) on the
-spread--
+%\subsubsection{Two-Factor model}
+%%distributed worm simulation
+%--Zou et
+%al. [15] present a ''two-factor'' model that extends SIR
+%epidemiological model to capture the effects of human
+%countermeasures and the congestion due to the worm
+%spread. Shen et al. [16] provide a discrete-time worm
+%model that considers patching and cleaning effect and can
+%model worms with local scanning techniques. All these
+%approaches abstract specifics of the Internet topology,
+%change in the size of vulnerable population as the worm
+%spreads, and the effect of the individual host and network
+%defenses (except for patching, in case of [16]) on the
+%spread--
 \subsubsection*{self disciplinary worms}
@@
 
-Popular mechanism that worms use to detect vulnerable targets by random IP scanning probing.
+%Popular mechanism that worms use to detect vulnerable targets by random IP scanning probing.
 Feasible due to use 32-bit addresses.
@@
 :spread locally, by biasing the search space towards addresses within the same subnet or network.
-The second strategy is almost the same as what email virusses would get. For this reason we work an example out ..
+%The second strategy is almost the same as what email virusses would get. For this reason we work an example out .. 
+%
+%
 
-A method to calculate the propagation of the virus in an easy way. Google page ranking algorithm.
-
-\todo{toch eigen methode introduceren}
+\subsection{Continious-time models}
+-- notities hier op overschrijven --
+\subsubsection{OSN worms model}
+\subsubsection{Email worms simulation model}
 
@@
 
-All of the above models for modelling the mechanism of the spreading of a worm depend on the kind of topology of the network.
-In this chapter we propose a method to calculate the spread of the worm in way where the topology of the network is easily integrated.
+%All of the above models for modelling the mechanism of the spreading of a worm depend on the kind of topology of the network. 
+In this chapter we propose a method to calculate the spread of the worm in way that the topology of the network can be easily integrated.
 The method will give an approximation of how fast a worm can infect a network.
-The method is based on the method for Google Page Ranking \todo{ref} [].
-or some of the spreading methods, the graph of the network matters.
+The method is based on the method for Google Page Ranking \cite{GoogleRank}.
+For some of the spreading methods, the graph of the network matters.
 It is important to have the right topology for the right method.
 Email worms need a topology that represent a social network, BGP routing worms need a topology on network level.
---iets zeggen dat het ook belangrijk is voor de defender om zijn netwerk zo aan te passen dat het virus moeilijker kan verspreiden. Network segregation mss aanhalen ?--
-
-\begin{description}
-\item Power law
-\item Small world topology
-\item random graph
-\item ..
-\end{description}
+\\
+%--iets zeggen dat het ook belangrijk is voor de defender om zijn netwerk zo aan te passen dat het virus moeilijker kan verspreiden. Network segregation mss aanhalen ?--
+%
+%\begin{description}
+%\item Power law
+%\item Small world topology
+%\item random graph
+%\item ..
+%\end{description}
 

### 2015-08-03 doc/template/nash.tex
nog een updqte

@@
 \section{Nash Equilibria}
--- rechtstreeks uit FlipIt paper --\\
+%TODO-- rechtstreeks uit FlipIt paper --\\
 As a second step, we are interested in finding Nash equilibria, points
@@
 attacker. Similarly, we denote by opt1($\alpha_{0}$) the set of values (rates of play $\alpha_{1}$) that optimize
-the benefit of the attacker for a fixed rate of play $\alpha_{0}$ of the defender. The following
-theorem specifies Nash equilibria for the periodic game and is proven in Appendix A. \\
+the benefit of the attacker for a fixed rate of play $\alpha_{0}$ of the defender. . \\
 \begin{figure}[hbtp]

### 2015-08-03 doc/template/chap-10.tex
nog een updqte
India stuff + verbetering ska

@@
 Why is it so important to keep a system secure?
-Many businesses store confidential information on clients, which can be lost through data leakage and can possibly be abused by competitors.
-Also, disruption caused by distributed denial of service (DDoS) attacks, may results in businesses failing to meet their service-level agreements.
+Many businesses store confidential information, which can be lost through data leakage and can possibly be abused by competitors.
+Also, disruption caused by distributed denial of service (DDoS) attacks, may result in businesses failing to meet their service-level agreements.
 Ultimately, computer and network security helps protecting a business against various kind of threats.
@@
 %Since it is so difficult to protect a system or a network against APT's, researchers have been looking for effective ways to predict in advance which defence strategy might be the better one. 
-There are a number of key strategies an organisation can apply to defend themselves against APT's: awareness, whitelisting, system administration, network segregation, dynamic content checking and patch management.
-Nevertheless the combination of all these elements benefit from being complemented by other defence strategies to protect one selves against stealthy takeovers.
+There are a number of key strategies an organisation can apply to defend themselve against APT's: awareness, whitelisting, system administration, network segregation, dynamic content checking and patch management.
+Nevertheless the combination of all these elements benefits from being complemented by other defence strategies to protect one selves against stealthy takeovers.
 One possible way to study the impact of stealthy takeovers and to determine practical recommendations for defenders is through game theory.\\
@@
 This paper focusses on a game with two players.
-The actions available to the attacker and the defender correspond respectively to the attacks on the system and taking defensive measures that protect the system.
+The actions available to the attacker and the defender correspond respectively to the attacks on the system and to take defensive measures that protect the system.
 \\
@@
 
-\subsubsection{Conjunctures and open problems}
-
-\begin{description}
-\item[-] Analyse other strategies like renewal strategies instead of only periodic strategies.
-\item[-] The defender can chose to Flip one node of the network or a subset instead of all the nodes.
-\item[-] Give the defender a delay. It takes a while before a patch is made after the discovery of a vulnerability.
-\end{description}
+%\subsubsection{Conjunctures and open problems}
+%
+%\begin{description}
+%\item[-] Analyse other strategies like renewal strategies instead of only periodic strategies.
+%\item[-] The defender can chose to Flip one node of the network or a subset instead of all the nodes.
+%\item[-] Give the defender a delay. It takes a while before a patch is made after the discovery of a vulnerability.
+%\end{description}
 

### 2015-08-03 doc/template/chap-5.tex
verdere aanpassingen
India stuff + verbetering ska

@@
 
-The generic term for data protection stored on a computer controlled device such as computers and smartphones, as well as public and private computer networks, including the entire Internet is called information security.
+Information security is the generic term for data protection stored on a computer controlled device such as computers and smartphones, as well as public and private computer networks, including the entire Internet.
 Security is a general term that encompasses several dimensions. More specifically, information security has three key objectives that are fundamental to information security:
@@
 Information security can be divided into two main sub categories.
-One of them is cyber security also known as computer security.
+One of them is cyber security, also known as computer security.
 This is the process of applying security tools to ensure confidentiality, integrity, and availability of data.
@@
 While cyber security and network security address different aspects, they partially overlap as well.
-For example a virus can be physically dropped on a computer network using a USB stick, but it can also arrive over the internet.
-Either way the internal computer security measurements have to be taken to recover from the virus.
+For example, a virus can be physically dropped on a computer network using a USB stick, but it can also arrive over the internet.
+Either way internal computer security measurements have to be taken to recover from the virus.
 In this paper we focus on scenarios where a network has to be defended against attacks to ensure confidentiality, integrity and availability of data.
@@
 Different types of malware exist: viruses, worms, flooders, rootkits, bots, spyware, adware and many more.
-This broad range of different types can be classified into two main categories, the first one based on the propagation method that is used and the second one based on the payload or the variety of actions that the malware performs \cite{stallings2007network}.
+This broad range of different types can be classified into two main categories.
+The first one based on the propagation method that is used and the second one based on the payload or the variety of actions that the malware performs \cite{stallings2007network}.
 Propagation methods include viruses, worms and trojans.
@@
 %This is where the researchers of RSA come into the picture. \todo{deze paper gaat zien of de game theoretische model kan leiden tot resultaten die kunnen helpen in het verdedigen tegen APTs}
-In this paper we will propose an appropriate game theoretical modelling of defending a network against APTs and analyse it in order to draw the necessary strategic conclusions to mitigate these kind of attacks.
+In this paper we will propose an appropriate game-theoretical modelling of defending a network against APTs and analyse it in order to draw the necessary strategic conclusions to mitigate these kind of attacks.
 
@@
 This book was about the mathematical analysis of a series of thinking games.
-The book described a series of games making a distinction between games in which the strategies and the utility factor of the opponent have no effect on finding the best strategies such (e.g.
+The book described a series of games making a distinction between games in which the strategies and the utility factor of the opponent have no effect on finding the best strategies (e.g.
 chess) and games in which this has an influence (e.g.
@@
 %Cooperative and non cooperative games
-In a decision game theoretic approach an agent will try to act in such a way to maximise his expected or average utility function.
+In a decision game-theoretic approach an agent will try to act in such a way to maximise his expected or average utility function.
 It becomes more complicated when two or more agents want to maximise their utility and when actions of the agents can affect each other's utilities.
@@
 \subsubsection{Best response and Nash Equilibrium}
-One of the solution concepts in Game Theory for non-cooperative games is a Nash Equilibrium that will be used in this paper.
+One of the solution concepts in Game Theory for non-cooperative games that will be used in this paper is a Nash Equilibrium.
 A Nash Equilibrium is a subset of outcomes that can be interesting to analyse a game.
@@
 A move can also result in a "wasted move", called a flop.
-It may happen that the resource was already under control by the player.
+It may happen that the resource was already under control of the player.
 If the player moves when he or she has already control over the resource, he or she would have wasted a move since this does not result in a change of ownership, so the cost is wasted.
@@
 \end{equation}
-~$ \alpha_{i}$ defines the average move rate by player \textit{i} up to time \textit{t}.
+~$ \alpha_{i}$ Defines the average move rate by player \textit{i} up to time \textit{t}.
 Let $\beta_{i}(t)$ denote player's \textit{i} average benefit upto time \textit{t}:
@@
 Depending on the amount of information received, two subclasses of adaptive strategies can be identified.
-The Last Move (LM) strategies represent the class where whenever a player flips he will find out the exact time that the opponent played the last time.
-In the second class, called Full History (FH), whenever a player flips he will find out the whole history of the opponent's move.
+The Last Move (LM) strategies represent the class where whenever a player flips he will find out the exact moment that the opponent moved the last time.
+In the second class, called Full History (FH), whenever a player flips he will find out the whole history of the opponent's moves.
 \\
@@
 \begin{itemize}
-\item periodic strategy strongly dominates the other renewal strategies if the opponent has a periodic or non-arithmetic renewal strategy, meaning that it is always advantageous for the opponent to play periodically against a player with a non-adaptive strategy;
+\item periodic strategy strongly dominates the other renewal strategies if the opponent has a periodic or non-arithmetic renewal strategy.
+This means that it is always advantageous for the opponent to play periodically against a player with a non-adaptive strategy;
 \item periodic games are disadvantageous against players following a Last Move adaptive strategy;
@@
 
-\section{Extensions on FlipIt}
+\section{Related Work on Extensions to FlipIt}
 \label{ch1:extendedWork}
@@
 \cite{MitigationNonTargeted} is to consider a game with targeted and non-targeted attacks where the moves made by the attacker do not succeed immediately.
-This approach is similar to what will be done in this paper, but nevertheless has some major difference.
+This approach is similar to what will be done in this paper, but nevertheless has some major differences.
 In Laszka's paper, firstly the moves by the attacker are still covert but the moves made by the defender are known to the attacker.
@@
 Another example is by the use of honeypots.
-Honeypots emulate services or creates multiple instances of real operating systems and can pretend to have sensitive information.
+Honeypots emulate services or create multiple instances of real operating systems and can pretend to have sensitive information.
 Honeypots do not detect all malicious attacks so the attacker can stay unnoticed and can still provide information as feedback.
@@
 One of the scenarios by Pham \cite{compromised} was to find out whether a resource was compromised or not by the attacker.
-This could be verified by the defender, who has an extra move "test" beside the flip move.
+This could be verified by the defender, who has an extra move ``test'' beside the flip move.
 The basic idea is to test with an extra action if the resource has been compromised or not.
@@
 They found out that the results improve over time but that they are dependent on gender, age, and other individual difference variables.
-The result also shows that the participants perform generally better when they have more information about the strategy of the opponent which is a computerized player.
+The result also shows that the participants perform generally better when they have more information about the strategy of the opponent, which is a computerized player.
 Reitter et al.

### 2015-08-03 doc/template/nash.tex
India stuff + verbetering ska

@@
 %TODO-- rechtstreeks uit FlipIt paper --\\
-As a second step, we are interested in finding Nash equilibria, points
-for which neither player will increase his benefit by changing his rate of play. More
-formally, a Nash equilibrium for the periodic game is a point $(\alpha_{0}^{*},\alpha_{1}^{*})$ such that
-the defender's benefit $\beta_{0}(\alpha_{0},\alpha_{1}^{*}) $is maximized at $\alpha_{0}= \alpha_{0}^{*}$ and the attacker's benefit
-$\beta_{1}(\alpha_{0}^{*},\alpha_{1}) $ is maximized at $\alpha_{1}= \alpha_{1}^{*}$ .
-To begin with, some useful notation. We denote by opt0($\alpha_{1}$) the set of values (rates
-of play $\alpha_{0}$) that optimize the benefit of the defender for a fixed rate of play $\alpha_{1}$ of the
-attacker. Similarly, we denote by opt1($\alpha_{0}$) the set of values (rates of play $\alpha_{1}$) that optimize
-the benefit of the attacker for a fixed rate of play $\alpha_{0}$ of the defender. . \\
+
+In this chapter we are interested in finding the optimal strategies.
+To calculate the optimal strategies of both players we have to find the Nash Equilibria of the game.
+First the optimal functions are derived from the formulas in the previous chapter.
+From these piecewise functions we can derive the Nash Equilibria.
+\\
+
+Nash equilibria are points with the property that neither player benefits by deviating in isolation from the equilibrium.
+We can compute Nash Equilibria for the periodic game as an intersection point of curves $opt_{D}$ and $opt_{A}$.
+\\
+%As a second step, we are interested in finding Nash equilibria, points
+%for which neither player will increase his benefit by changing his rate of play. 
+More formally, a Nash equilibrium for the periodic game is a point $(\delta_{D}^{*},\delta_{A}^{*})$ such that
+the defender's benefit $\beta_{D}(\alpha_{D},\delta_{A}^{*}) $is maximized at $\delta_{D}= \delta_{D}^{*}$ and the attacker's benefit
+$\beta_{A}(\delta_{D}^{*},\delta_{A}) $ is maximized at $\delta_{A}= \delta_{A}^{*}$.
+To begin with, some useful notation. We denote by $opt_{D}(\delta_{A}$) the set of values (rates
+of play $\delta_{D}$) that optimize the benefit of the defender for a fixed rate of play $\delta_{A}$ of the
+attacker. Similarly, we denote by $opt_{D}(\delta_{D}$) the set of values (rates of play $\delta_{A}$) that optimize
+the benefit of the attacker for a fixed rate of play $\delta_{D}$ of the defender. \\
+
+To determine $opt_{D}(\delta_{A})$ we need to compute the derivative of  $\beta_{D}(\delta_{D},\delta_{A}) $ for a fixed $\delta_{A}$.
+ We consider two cases, where case 2 is divided into two subcategories. Fig \todo{fig} shows all the cases.\\
 \begin{figure}[hbtp]
@@
 \includegraphics[scale=0.4]{Images/bestresp.png}
-\caption{This figure shows the three subcategories of each case.
+\caption{This figure shows the all the cases with their subcategories.
 'A' stands for Case 2.A: $\delta_{D} \geq d+\delta_{A} \geq \delta_{A}$ and 'B' stands for Case 2.B: $d+\delta_{A} \geq \delta_{D} \geq  \delta_{A} $ }
@@
 
-Nash equilibria are points with the property that neither player benefits by deviating in isolation from the equilibrium.
-We can compute Nash Equilibria for the periodic game as an intersection point of curves $opt_{D}$ and $opt_{A}$.
-To determine $opt_{D}(\alpha_{A})$ we need to compute the derivative of  $\beta_{D}(\alpha_{D},\alpha_{A}) $ for a fixed $\alpha_{A}$.
-We consider two cases:\\
-
 \subsection*{Case 1: $\delta_{D} \leq \delta_{A} $}
@@
 \end{equation*}
-If we take the partial derivative for a fixed $\delta_{A}$ we get the following result:
+If we take the partial derivative of this formula for a fixed $\delta_{A}$ we get the following:
 \begin{equation*}\label{formdelta}
@@
 
-The benefit formula obtained previously for this case is as follows:
+The benefit formula obtained in the previous chapter for this case is as follows:
 \begin{equation*}
@@
 
-This leads to the following deduction: The function increases on $[0, \sqrt{2\delta_{A}k_{D} + d^{2}}]$ and is decreasing on $[\sqrt{2\delta_{A}k_{D} + d^{2}}, \infty]$.
+For case 2.B this leads to the following deduction, which results in the same formula as for case 1: The function increases on $[0, \sqrt{2\delta_{A}k_{D} + d^{2}}]$ and is decreasing on $[\sqrt{2\delta_{A}k_{D} + d^{2}}, \infty]$.
 So there is a maximum on $\delta_{D} = minimum \{ \delta_{A}, \sqrt{2\delta_{A}k_{D} + d^{2}} \} $ \\
@@
 \subsection{Determining the piecewise functions $opt_{A}(\delta_{D})$}
-We still consider the case where $d < \delta_{D}$. \\
-To determine the Nash equilibria we also need to determine $opt_{A}(\delta_{D})$ by computing the derivative of $\beta_{A}(\delta_{D},\delta_{A})$ for a fixed $\delta_{D}$. We consider 2 cases: \\
+To start with we only consider the case where $d < \delta_{D}$, because if $d > \delta_{D}$ the benefit of the defender \todo{nakijken, def of att} is always 1. \\
+To determine the Nash equilibria we also need to determine $opt_{A}(\delta_{D})$ by computing the derivative of $\beta_{A}(\delta_{D},\delta_{A})$ for a fixed $\delta_{D}$.
+We consider again two cases, where case 2 is divided into two subcategories: \\
 
@@
 \end{equation*}
-it follows that $\beta_{A}(\delta_{D},\cdot)$ is increasing on $[0,\sqrt{2k_{A}\delta_{D}}]$ and decreasing on $[\sqrt{2k_{A}\delta_{D}}, \infty]$ and thus has a maximum on $\delta_{A} = maximum \{\delta_{D}, \sqrt{2k_{A}\delta_{D}} \} $.
+It follows that $\beta_{A}(\delta_{D},\cdot)$ is increasing on $[0,\sqrt{2k_{A}\delta_{D}}]$ and decreasing on $[\sqrt{2k_{A}\delta_{D}}, \infty]$ and thus has a maximum on $\delta_{A} = maximum \{\delta_{D}, \sqrt{2k_{A}\delta_{D}} \} $.
 The maximum between $\delta_{D}$ and $ \sqrt{2k_{A}\delta_{D}}$ is needed because $\delta_{A} $ cannot exceed $\delta_{D}$ in this case.
@@
 \subsubsection{Best responses}
-The optimum functions will be piecewise functions. We distinguish three cases for different values of $\delta_{d}$ and $k_{A}$.
+The optimum functions will be piecewise functions. We distinguish three cases for different values of $\delta_{D}$ and $k_{A}$.
 
@@
  \begin{displaymath}
-  opt_{D}(\delta_{A}) = \left\{
+  opt_{A}(\delta_{D}) = \left\{
      \begin{array}{lr}
-       0, & \delta_{A} < 2(k_{D} - d)\\
-       0, & \delta_{A} = 2(k_{D} - d) \\
-       \sqrt{d^{2} + 2\delta_{A}k_{D}}, & \delta_{A} > 2(k_{D} - d)
+       \delta_{D}, & \delta_{A} < 2(k_{D} - d)\\
+       \left[ 0, \delta_{D} \right],  & \delta_{A} = 2(k_{D} - d) \\
+       0 & \delta_{A} > 2(k_{D} - d)
      \end{array}
@@
 \end{displaymath}
+\\
 
+The Nash Equilibria can be found as the intersection points of the piecewise functions $opt_{A}(\delta_{D})$ and $opt_{D}(\delta_{A})$.
+For this we have to compare the function in terms of the relationship of the players move cost.
+We distinguish three cases: $k_{A} < k_{D} , ~k_{A} > k_{D} $ and $k_{A} = k_{D}$.
 \todo{ kA met kD vergelijken en nash evenwichten vinden}

### 2015-08-05 doc/template/chap-10.tex
Verbeteringen Jeroen

@@
 An APT is a multi-faceted, continuous and targeted cyber attack that is designed to penetrate a network or a system in a stealthy way and can stay undetected for a long period of time.
-It is different and stronger than a conventional threat.
+It is different and more severe than a conventional threat.
 A conventional threat will not attack any particular target.
@@
 %Since it is so difficult to protect a system or a network against APT's, researchers have been looking for effective ways to predict in advance which defence strategy might be the better one. 
-There are a number of key strategies an organisation can apply to defend themselve against APT's: awareness, whitelisting, system administration, network segregation, dynamic content checking and patch management.
-Nevertheless the combination of all these elements benefits from being complemented by other defence strategies to protect one selves against stealthy takeovers.
+There are a number of key strategies an organisation can apply to defend itself against APT's: awareness, whitelisting, system administration, network segregation, dynamic content checking and patch management.
+Nevertheless the combination of all these elements benefits from being complemented by other defence strategies to protect oneself against stealthy takeovers.
 One possible way to study the impact of stealthy takeovers and to determine practical recommendations for defenders is through game theory.\\
@@
 The attacker drops a virus on one of the nodes and waits until this virus infects the whole network.
-The attacker will only be in control of the resource when a sufficient large amount of nodes of the network are infected.
+The attacker will only be in control of the resource when a sufficiently large amount of nodes of the network are infected.
 In this paper we present an adaptation of FlipIt to model a game where the moves of the attacker are not instantaneous.
@@
 The attacker only compromises the system if sufficient nodes in the network are infected.
-\item[-] The periodic case of FlipIt is modelled with a delay, resulting in optimum functions and Nash equilibria. 
+\item[-] The periodic case of FlipIt is modelled with a delay, resulting in optimum functions and Nash equilibria. \todo{punt herschrijven en meer nadruk leggen op het vinden van de Nash equilibria}
 \item[-] Based on our results we can give practical recommendations to take measures against advanced attacks.
@@
 \item[-] This work includes an overview of different techniques used to calculate the propagation delay of malware (depending on the network layout). \todo{en het wordt ook toegepast op de formules}
+\item[-] It presents a method to calculate the speed of the propagation of a worm in a network independent of the topology of the network.
 \end{itemize}
@@
 The organisation of this paper is the following.
-An introduction to cyber security and game theory is given in chapter \ref{Chapter1:Intro.Game.Theory} to get familiar with the kind of threats that are in the scope of this work and the game theoretic concepts that will be further used in the paper.
+An introduction to cyber security and game theory is given in chapter \ref{Chapter1:Intro.Game.Theory} to allow the reader to become familiar with the kind of threats that are in the scope of this work and the game theoretic concepts that will be further used in the paper.
 In the same chapter the FlipIt framework is summarized with its most important conclusions.
 The chapter concludes with related work on FlipIt and further clarifies the contributions of this paper compared to existing work.
-Chapter \ref{chapter2:FlipIt with virus propagation} , first introduces the adaptations made to the original FlipIt model to model a FlipIt game with virus propagation.
+Chapter \ref{chapter2:FlipIt with virus propagation} first introduces the adaptations made to the original FlipIt model to model a FlipIt game with virus propagation.
 Subsequently formulas are derived to model a FlipIt game with a virus propagation for the specific case where players play a periodic strategy with a random phase.
@@
 It presents a method to calculate the speed of the propagation of a worm in a network independent of the topology of the network.
-Finally chapter \ref{chapter5:conclusion} , we discusses the main results and provides directions for further research.
+Finally chapter \ref{chapter5:conclusion} discusses the main results and provides directions for further research.
 

### 2015-08-05 doc/template/chap-5.tex
Verbeteringen Jeroen

@@
 Information security is the generic term for data protection stored on a computer controlled device such as computers and smartphones, as well as public and private computer networks, including the entire Internet.
-Security is a general term that encompasses several dimensions. More specifically, information security has three key objectives that are fundamental to information security:
+Security is a general term that encompasses several dimensions. More specifically, information security has three fundamental key objectives:
 
@@
 An identification tool will try to identify the threat to be able to know how to remove it.
-The removal tool will remove the threat from the system (once it has been identified) so that it cannot spread any further.
+A removal tool will remove the threat from the system (once it has been identified) so that it cannot spread any further.
 \\
@@
 A virus needs human interaction to spread.
-In the example above: if no one were to open the mail, the virus could not spread itself and infect other systems.
+In the example above: if no one were to open the mail, the virus would not be able to spread itself and infect other systems.
 
@@
 A notable trojan horse is Koobface, that targeted users of Facebook, Skype, Yahoo, Gmail and AOL mail.
-To spread itself the worm sent a mail or friend request with a message that directed the recipients to a third party website.
+To spread itself the trojan sent a mail or friend request with a message that directed the recipients to a third party website.
 This site would then convince the recipient into downloading an update of Adobe Flash Player.
@@
 
-An Advanced Persistent Threat (referred to in the rest of this work as APT) is a persistent targeted attack that tries to penetrate a network to cause harm while staying unseen for a long period of time.
+An Advanced Persistent Threat (referred to for the remainder of this work as APT) is a persistent targeted attack that tries to penetrate a network to cause harm while staying unseen for a long period of time.
 The motive of an APT is mostly cyber espionage, stealing sensitive data, sabotage or other kinds of ideological attacks.

### 2015-08-05 doc/template/chap-4.tex
Laatste hoofdstuk bijgewerkt

@@
 Because of the many different types of propagation it is hard to define a single model that can model all of them.
-Modelling the spread of malware depends on two key factors: the method used for the propagation and the graph where the malware will spread itself.
+Modelling the spread of malware depends on two key factors: the method used for the propagation and the graph of the network where the malware will spread itself.
 \\
@@
 %\begin{itemize}
+
 \item \textbf{Selective Random scanning:} Some worms propagate by using the method of selective random targeting IP addresses.
@@
 The rate of success for randomly chosen IP addresses is very low, but it is easy to implement.
-Example of such worms are `Code Red' and `Slammer'.
-\item \textbf{Localized scanning:} A worm that uses localized scanning will scan for hosts in the local address space. This method is used by the `Code Red II' and `Nimda' worm.
-\item \textbf{Topological scanning:} Address information stored in the victim machines are used to locate new targets. This method used by the `Morris' worm.
+Example of such worms are \textit{Code Red} and \textit{Slammer}.
+
+\item \textbf{Localized scanning:} A worm that uses localized scanning will scan for hosts in the local address space. This method is used by the \textit{Code Red II} and \textit{Nimda worm}.
+%\item \textbf{Topological scanning:} Address information stored in the victim machines are used to locate new targets. This method used by the `Morris' worm.
+
 %\item \textbf{Hit-list scanning:} With the method of hit-list scanning, a short-list of vulnerable systems is made beforehand.
@@
 An extreme case for the hit-list-scanning worms is a flash worm, which gathers all vulnerable machines into the list.--
+
 \item \textbf{Sequential scanning:}
@@
 For example, the address of the host is A, the next addresses that the worm will scan are A+1, or A-1.
-This method is used by the `Blaster worm'.
+This method is used by the \textit{Blaster worm}.
+
 \item \textbf{DNS random scanning:} Another strategy is a kind of strategy in which the DNS infrastructure is used to locate a new target address.
@@
 The speed of a DNS scanning worm in the IPv6 internet is comparable to the speed of an IPv4 random scanning worm. 
-There are some difficulties with this scanning technique.
-First it is difficult to obtain the whole address space from the DNS records.
-Secondly, the worm has to carry the database, which can slow down the propagation if the database is to big.
-Last, the IP addresses stored in the address table are only hosts with public domain names.
+%There are some difficulties with this scanning technique. 
+%First it is difficult to obtain the whole address space from the DNS records. Secondly, the worm has to carry the database, which can slow down the propagation if the database is to big. Last, t
+The IP addresses stored in the address table are only hosts with public domain names. This propagation method is used by \textit{MyDoom}.
+
 \item \textbf{Routable scanning:} The worms using routable scanning acquire target IP addresses based on the routing information in a network.
@@
 This method is three times faster than a traditional worm that uses random scanning.
-Examples are `Spyb0t' or `network Bluepill'.
+Examples are \textit{Spyb0t} or \textit{network Bluepill}.
 %--a ''routing worm'' [20].
@@
 
-\item \textbf{Topology-based Wroms} Email and other client application worms:An email worms uses the email systems to find email addresses to propagate.
+\item \textbf{Topology-based Worms} Email and other client application worms:An email worms uses the email systems to find email addresses to propagate.
 Other client applications can include: Internet Relay Chat (IRC), Instant Messenger (IM), and a variety of peer-to-peer
 file sharing systems, which have been used by worms to propagate in a similar way as email worms.
-For example, the `Kak worm' is a Javascript computer worm that spread itself by exploiting a bug in Outlook Express.
+For example, the \textit{Kak worm} is a JavaScript computer worm that spread itself by exploiting a bug in Outlook Express.
 % Another example is `GhostNet'
@@
 %Codered paper gebruikt voor model
-A Simple epidemic model is SI model. Nodes in the network can be either susceptible or infected.
-Once a node is infected it will stay infected. Every node has the same
-chance to be infected. The simple epidemic model is considered to be of a fixed size. The model for a fixed population is as follows:
+A Simple epidemic model is an SI model. Nodes in the network can be either susceptible or infected. 
+Once a node is infected it will stay infected. Every node has the same chance to be infected
+The simple epidemic model is considered to be of a fixed size. The model for a fixed population is as follows:
+
 \begin{equation}
@@
 In het begin zullen er ook geen geinfecteerde hosts zijn, dus de vergelijking kan opgelost worden.\\
-Er wordt geen rekening gehouden met de topology ? nog uitzoeken.
-\subsubsection{Kermack-Mckendrick model: SIR}
-In the Kermack-Mckendrick model, also known as SIR model, nodes can have three states: susceptible, infected or removed.
+%Er wordt geen rekening gehouden met de topology ? nog uitzoeken.
+
+\subsubsection{RCS model}
+
+RCS model stands for Random Constant Spread model and is developed by Paxson and Weaver at Stanford \citep{OwnInternetSI}.
+It is a model derived from the classical Simple Epidemic Model.
+They used this model to analyse the Code Red I worm.
+For this model it is assumed that the worm owns a good random number generator that is properly seeded.
+\\
+Let N be the total of vulnerable hosts in the network that can be potentially infected.
+It is assumed that no system is patched, shut down, deployed or disconnected, which means that the number of hosts in the system stay constant.
+The model also ignores spreads of the worm behind firewalls on private networks.
+\\
+Let K be the initial compromise rate.
+This is the rate of hosts that the worm can find and compromise per hour at the beginning of the infection.
+K is a global constant and therefore does not depend on the speed of the network or the processor speed.
+Every machine can infect only one other machine after it has been compromised.
+It cannot increase the rate that a worm can find new hosts.
+The internet topology is considered as a complete undirected graph.  The model type is SI, so the state of the host can only be infected or suspected. Once the host is compromised it stays that way.  
+Let T be the moment of the start of the infection. Variable \textit{a} is the proportion of vulnerable hosts that has been compromised. Variable \textit{t} is the time in hours. \\
+
+The formula to model the spread of the worm is as follows:
+\begin{equation}
+N da = (N a)K(1 - a)dt
+\end{equation}
+It tells how many vulnerable machines will be compromised in the next amount of time \textit{dt}, when it is known the proportion of machines \text{a} that already have been compromised.\\
+
+From this, it follows the simple differential equation:
+\begin{equation}
+\dfrac{d a}{dt} = Ka(1-a)
+\end{equation}
+With the following solution:
+\begin{equation}
+a = \dfrac{e^{K(t-T)}}{1+e^{K(t-T)}}
+\end{equation}
+\\
+
+To to apply this formula for the case of FlipIt we need to know the maximum numbers of hosts that can be infected before the total system is compromised.
+The initial compromise rate has to be approximated.
+\textit{t} is the variable in our formula and says how many time has passed.
+If a passes a certain value we know the value of t which is the same as d in our FlipIt game.
+
+%\subsubsection{Kermack-Mckendrick model: SIR}
+%In the Kermack-Mckendrick model, also known as SIR model, nodes can have three states: susceptible, infected or removed.
 Once a node of the network has been recovered from a worm, the node will stay in the removed mode and never becomes infected again.
@@
 \\
-Let I(t) be the number of infectious hosts at time \textit{t}, R(t) be the number of removed hosts at time \textit{t} and J(t) is the number of infected hosts by time \textit{i}, regardless the fact that a node can be in a removed state.
-\begin{equation}
-J(t) = I(t) + R(t).
-\end{equation} 
-The Kermack-McKendrick model can be represented as follows:
-\begin{equation}
-\begin{Bmatrix} \dfrac{d J(t)}{dt} = \beta J(t) \big[N- J(t) \\
-\dfrac{d R(t)}{dt} = \gamma I(t) \\
-J(t) = I(t) + R(t) = N - S(t)
- \end{Bmatrix}
-\end{equation}
-Parameter $\beta$ is again the rate of infection and $\gamma$ is the rate of removal of infected hosts. S(T) is the number of susceptible hosts at time t.
-\begin{figure}[hbtp]
-\centering
-\includegraphics[scale=0.5]{Images/KMmodel.png}
-\caption{KM Model.
+%Let I(t) be the number of infectious hosts at time \textit{t}, R(t) be the number of removed hosts at time \textit{t} and J(t) is the number of infected hosts by time \textit{i}, regardless the fact that a node can be in a removed state.
+%\begin{equation}
+%J(t) = I(t) + R(t).
+%\end{equation} 
+%The Kermack-McKendrick model can be represented as follows:
+%\begin{equation}
+%\begin{Bmatrix} \dfrac{d J(t)}{dt} = \beta J(t) \big[N- J(t) \\
+%\dfrac{d R(t)}{dt} = \gamma I(t) \\
+%J(t) = I(t) + R(t) = N - S(t)
+% \end{Bmatrix}
+%\end{equation}
+%Parameter $\beta$ is again the rate of infection and $\gamma$ is the rate of removal of infected hosts. S(T) is the number of susceptible hosts at time t.
+%\begin{figure}[hbtp]
+%\centering
+%\includegraphics[scale=0.5]{Images/KMmodel.png}
+%\caption{KM Model.
 The x-coordinate is the propagation time and the y-coordinate is the infected percentage of the whole network.
@@
 $N=10000, \beta = 1/10000000$.}
-\label{KMmodel}
-\end{figure}
-
-%codered%
---Define $\rho = \gamma/\beta$ to be the relative removal rate [3]. One
-interesting result coming out of this model is
-dI(t)
-dt > 0 if and only if S(t) > p. (6)
-Since there is no new susceptible host to be generated,
-the number of susceptible hosts S(t) is a monotonically decreasing
-function of time t. If S(t0) < p, then S(t) < p and
-dI(t)/dt < 0 for all future time t>t0. In other words, if the
-initial number of susceptible hosts is smaller than some critical
-value, S(0) < p, there will be no epidemic and outbreak
-[15].
-The Kermack-Mckendrick model improves the classical
-simple epidemic model by considering that some infectious
-hosts either recover or die after some time. However, this
-model is still not suitable for modeling Internet worm propagation.
-First, in the Internet, cleaning, patching, and filtering
-countermeasures against worms will remove both susceptible
-hosts and infectious hosts from circulation, but KermackMckendrick
-model only accounts for the removal of infectious
-hosts. Second, this model assumes the infection rate
-to be constant, which isn't true for a rampantly spreading
-Internet worm such as the Code Red worm--
-
-%\subsubsection{Two-Factor model}
-%%distributed worm simulation
-%--Zou et
-%al. [15] present a ''two-factor'' model that extends SIR
-%epidemiological model to capture the effects of human
-%countermeasures and the congestion due to the worm
-%spread. Shen et al. [16] provide a discrete-time worm
-%model that considers patching and cleaning effect and can
-%model worms with local scanning techniques. All these
-%approaches abstract specifics of the Internet topology,
-%change in the size of vulnerable population as the worm
-%spreads, and the effect of the individual host and network
-%defenses (except for patching, in case of [16]) on the
-%spread--
+%\label{KMmodel}
+%\end{figure}
+%
+%%codered%
+%--Define $\rho = \gamma/\beta$ to be the relative removal rate [3]. One
+%interesting result coming out of this model is
+%dI(t)
+%dt > 0 if and only if S(t) > p. (6)
+%Since there is no new susceptible host to be generated,
+%the number of susceptible hosts S(t) is a monotonically decreasing
+%function of time t. If S(t0) < p, then S(t) < p and
+%dI(t)/dt < 0 for all future time t>t0. In other words, if the
+%initial number of susceptible hosts is smaller than some critical
+%value, S(0) < p, there will be no epidemic and outbreak
+%[15].
+%The Kermack-Mckendrick model improves the classical
+%simple epidemic model by considering that some infectious
+%hosts either recover or die after some time. However, this
+%model is still not suitable for modeling Internet worm propagation.
+%First, in the Internet, cleaning, patching, and filtering
+%countermeasures against worms will remove both susceptible
+%hosts and infectious hosts from circulation, but KermackMckendrick
+%model only accounts for the removal of infectious
+%hosts. Second, this model assumes the infection rate
+%to be constant, which isn't true for a rampantly spreading
+%Internet worm such as the Code Red worm--
+%
+%%\subsubsection{Two-Factor model}
+%%%distributed worm simulation
+%%--Zou et
+%%al. [15] present a ''two-factor'' model that extends SIR
+%%epidemiological model to capture the effects of human
+%%countermeasures and the congestion due to the worm
+%%spread. Shen et al. [16] provide a discrete-time worm
+%%model that considers patching and cleaning effect and can
+%%model worms with local scanning techniques. All these
+%%approaches abstract specifics of the Internet topology,
+%%change in the size of vulnerable population as the worm
+%%spreads, and the effect of the individual host and network
+%%defenses (except for patching, in case of [16]) on the
+%%spread--
 \subsubsection*{self disciplinary worms}
@@
 %The second strategy is almost the same as what email virusses would get. For this reason we work an example out .. 
+\subsection{Discrete-time models}
+\subsubsection{Bluetooth worm model}
+\todo{ref:Modeling Propagation Dynamics of Bluetooth Worms}
+
+The Bluetooth worm model is introduced by Yan and Eidenbenz []. The model captures the behavior of the propagation of a worm that spreads through the Bluetooth protocol. 
+A Bluetooth device that is compromised can only infect neighbour devices that are in its radio range
+Let \textit{i(t)} be the average density of infected hosts in the network at time \textit{t}.
+
+Formula is as follows: 
+\begin{equation}
+i(t_{k+1})=i(t_{k}) \cdot \dfrac{\rho(t_{k})}{i'(t_{k}) + (\rho(t_{k}) - i'(t_{k}))e^{-\alpha' \cdot \rho(t_{k}) / (\rho(t_{k}) - i'(t_{k}))}}
+\end{equation}
+where $\rho$(t) and $\beta$(t) are the average device density and the pairwise infection rate at time \textit{t} respectively.
+The number of new infections out of the infection cycle is denoted by $\alpha$(t).\\
+
+The formula for a better estimation of the worm propagation
+\begin{equation}
+\alpha'=\dfrac{\rho(t_{k})-t(t_{k})}{\rho(t_{k})} \cdot \alpha(t_{k}) + \dfrac{i(t_{k})}{\rho(t_{k})} \cdot \alpha(t_{x})
+\end{equation}
+
+To apply this model on FlipIt we again set up a threshold value for i(t) and see how long it took to compromise this amount of hosts.
+The paper concluded in their work  that after setting model parameters accordingly ($\lambda_{ne} = 0.2108$ the average node degree and $J_{in}= 0.2372$ the average meeting rate of neighbours), the model predicts that the time it would take to infect 99$\%$ of the devices is slightly less than one hour.
+So in this case the delay is equal to an hour and the amount of hosts that has to be infected by the attacker before the network is compromised is 99$\%$ of the hosts in the network.
+
+\subsubsection{OSN worms model}
+OSN worms stands for Online Social Network worms.
+This model is based on worms that propagate through the use of social media, like Facebook.
+The difference between an email worm and a social network worms is that the topology of the network is different.
+
+%\subsubsection{AAWP model}
+%SIR AAWP stands for Analytic Active Worm Propagation.
+The AAWP model is a discrete time propagation model.
+This model is used for worms that propagate through uniform scanning.
+When uniform scanning is used, each IP address
 %
-%
+%Popular mechanism that worms use to detect vulnerable targets by random IP scanning probing.
+Feasible due to use 32-bit addresses.
+128-bit addresses life harder for worms, except the ones that use email systems to propagate.
+two new strategies: uniformly distributed random number generator to select new target.
+:spread locally, by biasing the search space towards addresses within the same subnet or network.
+%The second strategy is almost the same as what email virusses would get. For this reason we work an example out .. \\
 
-\subsection{Continious-time models}
--- notities hier op overschrijven --
-\subsubsection{OSN worms model}
-\subsubsection{Email worms simulation model}
+A method to calculate the propagation of the virus in an easy way. Google page ranking algorithm.
 
@@
                     semithick]
-  \tikzstyle{every state}=[fill=blue,draw=none,text=white]
+  \tikzstyle{every state}=[fill=black!60!green,draw=none,text=white]
 

### 2015-08-05 doc/template/chap-10.tex
aanpassingen van Matthias

@@
 This paper focusses on a game with two players.
-The actions available to the attacker and the defender correspond respectively to the attacks on the system and to take defensive measures that protect the system.
+The actions available to the attacker and the defender correspond respectively to the attacks on the system and the defensive measures that protect the system.
 \\
 
-Many security games that bridge game theory and cyber security have already been investigated, so finding a new game can be challenging.
+Many security games that bridge the gap between game theory and cyber security have already been investigated, so finding a new game can be challenging.
 This paper builds on a relatively new paper where the assumption of stealthiness is fairly unique, giving some interesting results.\\
@@
 
-While it may seem trivial to extend the basic FlipIt model with a propagation delay, it's mathematical treatment is not.
+While it may seem trivial to extend the basic FlipIt model with a propagation delay, its mathematical treatment is not.
 In the paper of Laznka et all.
 \citep{FlipThem} it seems that even a small extension adds to the already significant mathematical complexity.
-While the FlipIt game is quite symmetric, by adding a propagation delay the mathematical complexity rises.
+Even though the FlipIt game is quite symmetric, the mathematical complexity rises by adding a propagation delay.
 % First a formula to calculate the benefit of a basic FlipIt was derived to introduce the propagation delay.
@@
 In the same chapter the FlipIt framework is summarized with its most important conclusions.
-The chapter concludes with related work on FlipIt and further clarifies the contributions of this paper compared to existing work.
+The chapter concludes with an overview of the related work on FlipIt and further clarifies the contributions of this paper compared to existing work.
 Chapter \ref{chapter2:FlipIt with virus propagation} first introduces the adaptations made to the original FlipIt model to model a FlipIt game with virus propagation.
@@
 \\
-In Chapter \ref{chapter:Nash} the formulas are further analysed to determine if there is a nash equilibrium. 
+In Chapter \ref{chapter:Nash} the formulas are further analysed to determine if there is a Nash equilibrium. 
 In order to provide a clear perspective on delays, chapter \ref{chapter4: Worm propagation} gives an overview of the various methods of propagation and worm propagation models.
@@
 Finally chapter \ref{chapter5:conclusion} discusses the main results and provides directions for further research.
+\todo{chapters kloppen nog niet}
 

### 2015-08-05 doc/template/chap-5.tex
verbeteringen jeroen

@@
 The motive of an APT is mostly cyber espionage, stealing sensitive data, sabotage or other kinds of ideological attacks.
-APT's are `advanced' because these attacks are well funded and (usually) the attacker needs a great amount of expertise to successfully penetrate a network.
-Not all APT's are \textit{technically} advanced though.
-The attacker can also try to exploit known vulnerabilities, in the hope that his target has not yet secured himself against them.
+APT's are `advanced' because these attacks are well funded and because the attacker (usually) needs a great amount of expertise to successfully penetrate a network.
+Not all APT's are technically advanced though.
+The attacker can also try to exploit known vulnerabilities, in the hope that his target has not yet secured itself against them.
 \\
-`Persistent' refers to the fact that the attacker never quits trying to attack his target. The attack can be spread over several years, taking multiple steps.  \\
-An APT can be a mix of different types of malware and may use various propagation methods like the way a virus, worm or trojan horse propagates. \\
+`Persistent' refers to the fact that the attacker isn't trying to gain immediate results, and the attacker won't stop trying after one failed attempt.
+The attack can be spread over several years, taking multiple steps.
+\\
+An APT can be a mix of different types of malware and may use various propagation methods. \\
+% like the way a virus, worm or trojan horse propagates. \\
 %
@@
 
-According to a security survey of Kaspersky \cite{SurveyKaspersky}  the damage of one successful targeted attack against a large company can exceed over 2.54 million dollar.
-Companies need a defence mechanism to defend themselves against APT's.
+According to a security survey of Kaspersky \cite{SurveyKaspersky}  the damage of one successful targeted attack against a large company can exceed 2.5 million dollars.
+As such, companies need a defence mechanism to defend themselves against APT's.
 As previously stated, simple detection and identification tools are insufficient to protect oneself against APT's, and a removal tool will only work if the threat has been identified.
@@
 %This is where the researchers of RSA come into the picture. \todo{deze paper gaat zien of de game theoretische model kan leiden tot resultaten die kunnen helpen in het verdedigen tegen APTs}
-In this paper we will propose an appropriate game-theoretical modelling of defending a network against APTs and analyse it in order to draw the necessary strategic conclusions to mitigate these kind of attacks.
+This paper proposes an appropriate game-theoretical modelling of defending a network against APTs and will analyse it in order to draw the necessary strategic conclusions to mitigate these kind of attacks.
 
-\section{A brief introduction in Game Theory}
+\section{A Brief Introduction in Game Theory}
 \label{Cha1:briefintro}
-Gametheory is a mathematical study to analyse interactions between independent and self-interested agents.
+Game theory is a mathematical study to analyse interactions between independent and self-interested agents.
 To get an understanding of the most important concepts of game theory, a short introduction based on the work of
-\cite{leyton2008essentials} and \cite{Coursera} is given in this section. For a more detailed and full introduction to game theory, the reader is referred to 
+\cite{leyton2008essentials} and \cite{Coursera} is given in this section. For a more detailed and complete introduction to game theory, the reader is referred to 
 \cite{leyton2008essentials}.  \\
@@
 %------------------------------------------------%
-Game theory is a mathematical way of modelling the interactions between two or more agents where the outcomes depend on what everybody does and how it should be structured to lead to good outcomes.
-Game theory has therefore important applications in many area's such as economics, politics, biology, computer science, philosophy and a variety of other disciplines.
-It has come only in ascent during the Second World War with Oskar Morgenstern and John von Neumann both publishing a book on game theory, entitled "Theory of Games and Economic Behavior" (Game theory and economic behaviour) in 1944.
-This book was about the mathematical analysis of a series of thinking games.
-The book described a series of games making a distinction between games in which the strategies and the utility factor of the opponent have no effect on finding the best strategies (e.g.
-chess) and games in which this has an influence (e.g.
+Game theory is a mathematical way of modelling the interactions between two or more agents where the outcomes depend on what each agent does and how these interactions should be structured to lead to good outcomes.
+Game theory therefore has important applications in many areas such as economics, politics, biology, computer science, philosophy and a variety of other disciplines.
+It gained recognition during the Second World War, when Oskar Morgenstern and John von Neumann both published a book on game theory, titled "Theory of Games and Economic Behavior" in 1944 \cite{von1944theory}.
+This book addressed the mathematical analysis of a series of thinking games.
+A distinction was made between games in which the strategies and the utility factor of the opponent have no effect on finding the best strategies (e.g.
+chess) and games wherein this factor does have an influence (e.g.
 poker).
@@
 He was one of the mathematicians who has formalized game theory.
-The Nash equilibrium was named after him.
+The Nash equilibrium, a common solution concept of a non-cooperative game, was named after him.
 ~~\\
@@
 
-One of the assumptions underlying game theory is that the players of the game, the agents, are independent and self-interested.
+One of the underlying assumptions game theory is that the players of the game are independent and self-interested.
 This does not necessarily mean that they want to harm other agents or that they only care about themselves.
 %utility function meaning 
-Instead it means that each agent has preferences about the states of the world he likes.
-These preferences are mapped to natural numbers and are called the utility function.
+Instead it means that each agent has preferences about which states of the world he likes.
+These preferences are mapped to natural numbers and are called the utility functions.
 The numbers are interpreted as a mathematical measure that tells how much an agent likes or dislikes the states of the world.
@@
 %Cooperative and non cooperative games
-In a decision game-theoretic approach an agent will try to act in such a way to maximise his expected or average utility function.
-It becomes more complicated when two or more agents want to maximise their utility and when actions of the agents can affect each other's utilities.
-This kind of games are referred to as non-cooperative game theory, where the basic modelling unit is the group of agents.
-The individualistic approach, where the basic modelling is only one agent, is referred to as cooperative game theory.
+In a decision game-theoretic approach an agent will try to act in such a way as to maximise his expected or average utility function.
+It becomes more complicated when two or more agents want to maximise their utility and when actions of an agent can affect other agents' utilities.
+In this kind of game, the basic modelling unit is the group of agents.
+It is referred to as \textit{non-cooperative game theory}.
+In the individualistic approach the basic modelling is only one agent.
+This type of game is referred to as \textit{cooperative game theory}.
 
@@
 The best response for player \textit{1} is: $a_{1} = Opt_{1}(a_{2})$, given that $a_{2}$ is an action of player 2.
-A
 For a Nash Equilibrium each player has a list of actions and each player's action maximizes his or her pay-off given the actions of the other players.
@@
 Each of them can betray the other or they can support each other and remain silent.
-If a player confesses, he will, depending on what the other does, go for three years in prison or he is free to go.
-If the player is silent, he will, depending on what the other player does, go for five years in jail or only one year.
-\\
+If a player confesses, he will either be imprisoned for 3 years or go free, depending on what the other does.
+If the player is silent, he will either be imprisoned for 5 years or one year, depending on what the other does.
+\\ \todo{nakijken met praten, bekennen of andere veraden?}
 
@@
 \caption{Prisoners dilemma: An example from the field of Game Theory.
-It is a game between two perfectly rational prisoners who don not know what the other one will do.
-Each can confess or stay silent.
+It is a game between two perfectly rational prisoners who do not know what the other one will do.
+Each prisoner can confess or stay silent.
 The resulting sentence is shown below each prisoner.
@@
 %A flip will be regarded as a move from a player \textit{i}. 
-Each move implies a certain cost $k_{i}$ and can vary for each player. Both players try to minimize their cost. Adding a cost prevents players to move too frequently. \\
+Each move implies a certain cost $k_{i}$ and can vary for each player. Both players try to minimize their cost. Adding a cost prevents players from moving too frequently. \\
 
-The unique feature of FlipIt is that every move happens in a stealthy way, meaning that the player has no clue whether the other player (his adversary) has flipped the resource or not.
-For instance, the defender does not know if the resource has been compromised by the attacker until he flips the resource himself.
+The unique feature of FlipIt is that every move happens in a stealthy way, meaning that the player does not immediately finds out if the other player (his adversary) has flipped the resource or not.
+A player only finds out about the state of the game when he moves himself.
 The goal of the player is to maximize the time that he or she has control over the resource while minimizing the total cost of the moves.
@@
 
-The players receive a benefit equal to the time units they were in possession of the resource minus the cost of making their moves. The cost of a player \textit{i} is denoted by $k_{i}$. 
-The total gain of player \textit{i} is equal to the total amount of time that a player \textit{i} has owned the resource from the beginning of the game up to time \textit{t}.
+The players receive a benefit equal to the time units they were in possession of the resource minus the cost of making their moves. The cost for player \textit{i} is denoted by $k_{i}$. 
+The total gain for player \textit{i} is equal to the total amount of time that player \textit{i} has owned the resource from the beginning of the game up to time \textit{t}.
 It is expressed as follows:
@@
 \end{equation}
-If we add up the gain of the defender and the gain of the attacker it should sum up to \textit{t}:
+The gain of the attacker and the defender always sums up to \textit{t}:
 \begin{equation}\label{first}
@@
 \end{equation}
-~$ \alpha_{i}$ Defines the average move rate by player \textit{i} up to time \textit{t}.
-Let $\beta_{i}(t)$ denote player's \textit{i} average benefit upto time \textit{t}:
+~Let $ \alpha_{i}$ defines the average move rate by player \textit{i} up to time \textit{t} and let $\beta_{i}(t)$ denote player's \textit{i} average benefit rate upto time \textit{t}:
 \begin{equation}\label{first}
@@
 \subsubsection{Strategies}
-Because the players move in a stealthy way, there are different types of feedback that a player can get while moving.
+Because the players move in a stealthy way, there are different types of feedback a player can get by flipping the resource.
 These types of feedback can be divided into two groups of strategies.
@@
 
-If there is no feedback for either player, we have a non-adaptive strategy.
-Because a player does not receive any feedback during the game he will play in the same manner against every opponent.
-The strategy is called non-adaptive because the playing strategy is not dependent on the opponents movements.
-An interesting subclass of the non-adaptive strategies is the one where the time intervals between two consecutive moves are generated by a renewal process.
+If there is no feedback for either player, he will play in the same manner against every opponent.
+The strategy is then called \textit{non-adaptive} because the playing strategy is not dependent on the opponents movements.
+An interesting subclass of the non-adaptive strategies is the one where the time intervals between two consecutive moves are generated by a renewal process \todo{renewal beter uitleggen}.
 An example of such renewal strategy is the periodic strategy where the time between two consecutive moves of the players are a fixed interval.
@@
 
-In case there is feedback, a player can adapt his strategy to the information received about the opponent's moves.
+If the players receive feedback, a player can adapt his strategy to the information received about the opponent's moves.
+This strategy is called \textit{adaptive} strategies.
 Depending on the amount of information received, two subclasses of adaptive strategies can be identified.
-The Last Move (LM) strategies represent the class where whenever a player flips he will find out the exact moment that the opponent moved the last time.
+The Last Move (LM) strategies represent the class where, whenever a player flips, he will find out the exact moment that the opponent moved the last time.
 In the second class, called Full History (FH), whenever a player flips he will find out the whole history of the opponent's moves.
@@
 \subsubsection{Results of the FlipIt game}
-The study of the different strategies by means of FlipIt framework allows to derive a number of interesting results:  
+The study of the different strategies by means of FlipIt framework allows to derive a number of interesting results \cite{FlipIt}:  
 \begin{itemize}
@@
 The rationale is that for compromising a system in real life, more than just one resource needs to be taken over.
-An example is that gaining access to deeper layers of a system may require breaking several passwords.
-The model is called FlipThem \cite{FlipThem}.
+For example, gaining access to deeper layers of a system may require breaking several passwords.
+This model is called FlipThem \cite{FlipThem}.
 Laszka et al.
@@
 This approach is similar to what will be done in this paper, but nevertheless has some major differences.
-In Laszka's paper, firstly the moves by the attacker are still covert but the moves made by the defender are known to the attacker.
+Firstly, in Laszka's paper, the moves by the attacker are still covert but the moves made by the defender are known to the attacker.
 This means that the attacker knows when the defender plays and can change its strategy depending on the moves of the defender.
@@
 The Stuxnetworm was resident on a isolated network, meaning that there was no way to connect to the internet.
-Another example is by the use of honeypots.
+Another example is the use of honeypots.
 Honeypots emulate services or create multiple instances of real operating systems and can pretend to have sensitive information.
 Honeypots do not detect all malicious attacks so the attacker can stay unnoticed and can still provide information as feedback.
-This is a way how the defender can also be stealthy by providing  false information through the honeypot.
+By providing false information through honeypots, the defender can also remain stealthy.
 The second difference is that even though both the targeted and non-targeted attacks do not succeed immediately, the delay is determined differently.
 For the targeted attack the time till it succeeds is given by an exponential distributed random variable with a known rate.
-The non-targeted attacks are modelled as a single attacker and the time till it succeeds is given by a Poisson process.
-In our paper the delay is given by one parameter, that can be the result of any virus propagation model.
-The third and last difference is that the paper of Laska has multiple attackers and they try to find the best strategy of the defender against both targeted and non-targeted attacks.
+The non-targeted attacks are modelled as a single attacker and the time until it succeeds is given by a Poisson process.
+In this paper the delay is given by one parameter, which can be the result of any virus propagation model.
+The third and last difference is that the paper of Laska has multiple attackers who try to find the best strategy of the defender against both targeted and non-targeted attacks.
 The conclusion of this paper is that the optimal strategy for the defender is moving periodically.
@@
 \\
-Other authors used the FlipIt game to apply it to a specific scenario. To be able to use the FlipIt game, modifications where required for the FlipIt model.
+Other authors used the FlipIt game to apply it to a specific scenario. To be able to use the FlipIt game, modifications were required for the FlipIt model.
 One of the scenarios by Pham \cite{compromised} was to find out whether a resource was compromised or not by the attacker.
@@
 
-Finally researchers also have investigated the behaviour of humans playing FlipIt.
+Finally, researchers have also investigated the behaviour of humans playing FlipIt.
 A.
-Nochenson and Grossklags \cite{nochenson2013behavioral}  investigate how people really act when given temporal decisions.
+Nochenson and Grossklags \cite{nochenson2013behavioral}  investigated how people really act when given temporal decisions.
 They found out that the results improve over time but that they are dependent on gender, age, and other individual difference variables.

### 2015-08-05 doc/template/chap-4.tex
Vervolg schrijven SI

@@
 \label{modelsforpropagation}
-To use the models for worm propagation in our model of the FlipIt game it is sufficient to know the propagation speed of the worm and define after how many time units the network is defected.
+To use the models for worm propagation in our model of the FlipIt game it is sufficient to know the propagation speed of the worm and define after how many time units a sufficient number of hosts of the network is infected.
 This will be equal to the \textit{d} parameter in our model.
@@
 
-The methods of propagation can be divided into different model types.
+%The methods of propagation can be divided into different model types.
+SI, SIS and SIS...
+uitbreiden..
 We are only interested in models that are based on SI model type (Susceptible-Infected), because in our model every node that is infected will stay infected.
-In table \ref{tree} the taxonomy of the propagation methods are given and the type of model.
+
+The propagation models of diseases are applicable to the propagation of worms.
+SI, SIS and SIR.
+The SI model or also SEM (Simple Epidemic Model) is a model where the hosts in the network only have two states: Susceptible or Infected.
+Once they are infected they stay that way.
+The second model, SIS, is a model where the hosts can return to the Susceptible state, which in turn can become infected again.
+SIR is a model where the hosts can go in a Removed state.
+Once it is in a Removed state the hosts can not be infected any more.
 \\
 
-For more information about other models the reader is refered to the following papers with other most common propagation models: \cite{wang2014modeling}, \cite{OnWorms2005survey}, \cite{xiang2009propagation} and \cite{serazzi2004computer}.
+In this chapter we are only interested in the SI model.
+To apply the propagation models on FlipIt we need to know the propagation speed of the virus, without hosts becoming susceptible again.
+All hosts must stay infected.
+For this we only look at SI models.
+In table \ref{tree} the taxonomy of the propagation methods are given and the type of model.
+The General Epidemic model and Two-factor model are based on the Simple Epidemic Model with the difference that they both have extra parameters to model the Removed state.
+We discuss a couple of SI models and then introduce a method of out own.\\
+
+For more information about other models the reader is referred to the following papers with other most common propagation models: \cite{wang2014modeling}, \cite{OnWorms2005survey}, \cite{xiang2009propagation} and \cite{serazzi2004computer}.
 %In this section we only list the models that are the most interesting from the perspective of calculating the delay, namely the SI type models.
@@
 
-\subsection{Continious-time models}
+\subsection{Continuous-time models}
 \subsubsection{Simple Epidemic model}
@@
 Figuur \ref{SEMmodel} laat het SEM model zien.
-This Simple epidemic model has been used in various papers \cite{OwnInternetSI}, \cite{CodeRed} to model random scanning worm such as Code Red and Slammer. \\
+This Simple epidemic model has been used in various papers \cite{OwnInternetSI}, \cite{CodeRed} to model random scanning worm such as \textit{Code Red} and \textit{Slammer}. \\
 For our paper we have to approximate the propagation speed en Infect gelijkstellen aan het aantal genoeg nodes die geinfecteerd moeten zijn door de APT om de controle te hebben.
@@
 It is a model derived from the classical Simple Epidemic Model.
-They used this model to analyse the Code Red I worm.
+They used this model to analyse the \textit{Code Red I} worm.
 For this model it is assumed that the worm owns a good random number generator that is properly seeded.
 \\
-Let N be the total of vulnerable hosts in the network that can be potentially infected.
+Let\textit{ N} be the total of vulnerable hosts in the network that can be potentially infected.
 It is assumed that no system is patched, shut down, deployed or disconnected, which means that the number of hosts in the system stay constant.
@@
 This is the rate of hosts that the worm can find and compromise per hour at the beginning of the infection.
-K is a global constant and therefore does not depend on the speed of the network or the processor speed.
+\textit{K} is a global constant and therefore does not depend on the speed of the network or the processor speed.
 Every machine can infect only one other machine after it has been compromised.
@@
 The internet topology is considered as a complete undirected graph.  The model type is SI, so the state of the host can only be infected or suspected. Once the host is compromised it stays that way.  
-Let T be the moment of the start of the infection. Variable \textit{a} is the proportion of vulnerable hosts that has been compromised. Variable \textit{t} is the time in hours. \\
+Let \textit{T} be the moment of the start of the infection. Variable \textit{a} is the proportion of vulnerable hosts that has been compromised. Variable \textit{t} is the time in hours. \\
 
@@
 \subsubsection*{self disciplinary worms}
-Model for self disciplinary worms and counter measures ... []
+Model for self disciplinary worms and counter measures ... []  \todo{verder uitwerken ? ja of nee? }
 
@@
 
-\subsubsection{OSN worms model}
+\subsubsection{OSN worm model}
 OSN worms stands for Online Social Network worms.
@@
 
-%\subsubsection{AAWP model}
-%SIR AAWP stands for Analytic Active Worm Propagation.
-The AAWP model is a discrete time propagation model.
-This model is used for worms that propagate through uniform scanning.
+\subsubsection{AAWP model}
+SIR AAWP stands for Analytic Active Worm Propagation.
+The AAWP model is a SIR model which uses a death rate in the formula to calculate the amount of hosts that become Removed.
+Yet this model is still analysed because if the death rate is removed, this model becomes a SI mode.
+The model is different from other SI models because it includes the time that it takes to infect a host.
+A hosts cannot infect another hosts before it is completely compromised.
+The model also considers the fact that a hosts can be scanned and hit by multiple worms at the same time.
+The Epidemiological model ignores the case.
+%This model is used for worms that propagate through uniform scanning.
 When uniform scanning is used, each IP address
-%
+
 %Popular mechanism that worms use to detect vulnerable targets by random IP scanning probing.
@@
 
-\section{methode met matrixen}
+\section{Matrix worm model}
 \label{eigenmatrixmethode}

### 2015-08-05 doc/template/chap-5.tex
Vervolg schrijven SI

@@
 A typical virus will attach itself to a program, or an executable content on a computer.
-The ``I love you'' virus is an example of a virus where the virus attached itself as an executable to a mail.
+The \textit{I love you} virus is an example of a virus where the virus attached itself as an executable to a mail.
 It propagated by using the mailing systems.
-When a victim opened an email with the ``I love you'' virus in the annex, the virus spread itself by sending a mail to everyone in the victim's contact list.
+When a victim opened an email with the \textit{I love you} virus in the annex, the virus spread itself by sending a mail to everyone in the victim's contact list.
 
@@
 The worm will use vulnerabilities of the system to infect other computers.
-The Stuxnetworm is a very prominent example of a worm.
+The \textit{Stuxnetworm} is a very prominent example of a worm.
 Initially this worm was spread via infected USB sticks and from there on it could spread itself to other hosts on the network through the Internet, without any further human intervention.
-The purpose of the Stuxnetworm was to harm the centrifuges in nuclear reactors; many reactors have been infected.
+The purpose of the \textit{Stuxnetworm} was to harm the centrifuges in nuclear reactors; many reactors have been infected.
 
 \item \textit{Trojan:} This is a malicious program that disguises itself as something normal and useful, so that users won't be suspicious of installing it, but it has a malicious function hidden inside that can circumvent security measures and cause harm.
-A notable trojan horse is Koobface, that targeted users of Facebook, Skype, Yahoo, Gmail and AOL mail.
+A notable trojan horse is \textit{Koobface}, that targeted users of Facebook, Skype, Yahoo, Gmail and AOL mail.
 To spread itself the trojan sent a mail or friend request with a message that directed the recipients to a third party website.
 This site would then convince the recipient into downloading an update of Adobe Flash Player.
-Once downloaded and executed, Koobface could infect the host.
+Once downloaded and executed, \textit{Koobface }could infect the host.
 
@@
 Standard malware protection tools like firewalls and virus scanners can keep these kind of malware out of the system.
-Installing protection against unkown threats is also relatively easy, but this requires tools that go beyond the standard methods, e.g.
+Installing protection against unknown threats is also relatively easy, but this requires tools that go beyond the standard methods, e.g.
 dynamic whitelisting.
@@
 %This is where the researchers of RSA come into the picture. \todo{deze paper gaat zien of de game theoretische model kan leiden tot resultaten die kunnen helpen in het verdedigen tegen APTs}
-This paper proposes an appropriate game-theoretical modelling of defending a network against APTs and will analyse it in order to draw the necessary strategic conclusions to mitigate these kind of attacks.
+This paper proposes an appropriate game-theoretical modelling of defending a network against APT's and will analyse it in order to draw the necessary strategic conclusions to mitigate these kind of attacks.
 
@@
 \end{equation}
-~Let $ \alpha_{i}$ defines the average move rate by player \textit{i} up to time \textit{t} and let $\beta_{i}(t)$ denote player's \textit{i} average benefit rate upto time \textit{t}:
+$\alpha_{i}$ defines the average move rate by player \textit{i} up to time \textit{t} with $n_{i}(t)$ the amount of moves made by player \textit{i} up to time \textit{t}:
+\begin{equation}
+\alpha_{i}(t) = \dfrac{n_{i}(t)}{t}
+\end{equation}
+Let $\beta_{i}(t)$ denote player's \textit{i} average benefit rate up to and including time \textit{t}:
 \begin{equation}\label{first}
@@
 FlipIt also has been applied to several cases in computer security.
-Reseachers explored different applications of FlipIt for real-world problems, like password reset policies, VM refresh, cloud auditing and key rotation \cite{ApplyingFlipit}.
+Researchers explored different applications of FlipIt for real-world problems, like password reset policies, VM refresh, cloud auditing and key rotation \cite{ApplyingFlipit}.
 \\

### 2015-08-05 doc/template/chap-4.tex
commit weet ik veel

--- 2015-08-05 doc/template/chap-4.tex
@@
 
-The methods of propagation can be divided into different model types.
-We are only interested in models that are based on SI model type (Susceptible-Infected), because in our model every node that is infected will stay infected.
-In table \ref{tree} the taxonomy of the propagation methods are given and the type of model.
+Scan based or Topology based
+A node in a network can be in three different states: susceptible, infected or removed.
+A susceptible node is a node that is vulnerable to infection.
+An infected node is a node that has been compromised and can infect other nodes.
+A removed node is dead or immune, which means it cannot be infected again by worms.
+With these three states three main propagation models are proposed: SI, SIR and SIS.
+SI model is a model where a node once infected stays infected.
+In an SIR model a node can once it is infected become removed.
+In an SIS model a node can become again susceptible after it has been infected.
+Based on these three models various other models have been proposed.
+We are only interested in models where we can find the time that it takes for a worm to infect a sufficient amount of nodes.
+All models that are based on SI model type (Susceptible-Infected) are good, because in our model every node that is infected will stay infected.
+In SIR and SIS models ..
+kunnen we het ook vinden (AAWP).
 \\
 
-For more information about other models the reader is refered to the following papers with other most common propagation models: \cite{wang2014modeling}, \cite{OnWorms2005survey}, \cite{xiang2009propagation} and \cite{serazzi2004computer}.
+In this chapter we gaan dieper op sommige models and toepassen the models on the delay.
+We can divide the propagation models in continuous and discrete time models.
+Continious time models are ..
+Discrete time models are ..
+ Table \ref{tree} gives an overview of the taxonomy of the propagation methods are given and the type of model. \todo{deze figuur moet aangepast worden want niet elk model wordt uitgelegd}.\\
+
+For more information about other models the reader is refered to the following papers with other common propagation models: \cite{wang2014modeling}, \cite{OnWorms2005survey}, \cite{xiang2009propagation} and \cite{serazzi2004computer}.
 %In this section we only list the models that are the most interesting from the perspective of calculating the delay, namely the SI type models.
@@
 \label{eigenmatrixmethode}
-
+As can be seen in table \ref{tree} most propagation models are for a specific network topology.
+What we want to have is to have a method to calculate the delay that is independent of the network topology.
 %All of the above models for modelling the mechanism of the spreading of a worm depend on the kind of topology of the network. 

### 2015-08-05 doc/template/chap-4.tex

--- 2015-08-05 doc/template/chap-4.tex
@@
 
+Scan based or Topology based
+A node in a network can be in three different states: susceptible, infected or removed.
+A susceptible node is a node that is vulnerable to infection.
+An infected node is a node that has been compromised and can infect other nodes.
+A removed node is dead or immune, which means it cannot be infected again by worms.
+With these three states three main propagation models are proposed: SI, SIR and SIS.
+SI model is a model where a node once infected stays infected.
+In an SIR model a node can once it is infected become removed.
+In an SIS model a node can become again susceptible after it has been infected.
+Based on these three models various other models have been proposed.
+We are only interested in models where we can find the time that it takes for a worm to infect a sufficient amount of nodes.
+All models that are based on SI model type (Susceptible-Infected) are good, because in our model every node that is infected will stay infected.
+In SIR and SIS models ..
+kunnen we het ook vinden (AAWP).
+\\
+
+In this chapter we gaan dieper op sommige models and toepassen the models on the delay.
+We can divide the propagation models in continuous and discrete time models.
+Continious time models are ..
+Discrete time models are ..
+ Table \ref{tree} gives an overview of the taxonomy of the propagation methods are given and the type of model. \todo{deze figuur moet aangepast worden want niet elk model wordt uitgelegd}.\\
+
+For more information about other models the reader is refered to the following papers with other common propagation models: \cite{wang2014modeling}, \cite{OnWorms2005survey}, \cite{xiang2009propagation} and \cite{serazzi2004computer}.
+
 %The methods of propagation can be divided into different model types.
@@
 For more information about other models the reader is referred to the following papers with other most common propagation models: \cite{wang2014modeling}, \cite{OnWorms2005survey}, \cite{xiang2009propagation} and \cite{serazzi2004computer}.
+
 %In this section we only list the models that are the most interesting from the perspective of calculating the delay, namely the SI type models.

### 2015-08-06 doc/template/chap-4.tex
laatste chapter aangepast

@@
 The rate of success for randomly chosen IP addresses is very low, but it is easy to implement.
-Example of such worms are \textit{Code Red} and \textit{Slammer}.
+Example of such worms are \textit{Code Red} \citep{OwnInternetSI} and \textit{Slammer} \citep{moore2003inside}.
 
-\item \textbf{Localized scanning:} A worm that uses localized scanning will scan for hosts in the local address space. This method is used by the \textit{Code Red II} and \textit{Nimda worm}.
+\item \textbf{Localized scanning:} A worm that uses localized scanning will scan for hosts in the local address space.
+This method is used by the \textit{Code Red II} \citep{OwnInternetSI} and \textit{Nimda worm} \citep{OwnInternetSI}.
 %\item \textbf{Topological scanning:} Address information stored in the victim machines are used to locate new targets. This method used by the `Morris' worm.
@@
 For example, the address of the host is A, the next addresses that the worm will scan are A+1, or A-1.
-This method is used by the \textit{Blaster worm}.
+This method is used by the \textit{Blaster worm} \citep{zou2006performance}.
 
@@
 %First it is difficult to obtain the whole address space from the DNS records. Secondly, the worm has to carry the database, which can slow down the propagation if the database is to big. Last, t
-The IP addresses stored in the address table are only hosts with public domain names. This propagation method is used by \textit{MyDoom}.
+The IP addresses stored in the address table are only hosts with public domain names. This propagation method is used by \textit{MyDoom} \citep{kamra2005effect}.
 
@@
 file sharing systems, which have been used by worms to propagate in a similar way as email worms.
-For example, the \textit{Kak worm} is a JavaScript computer worm that spread itself by exploiting a bug in Outlook Express.
+For example, the \textit{Kak worm} \citep{Kakworm} is a JavaScript computer worm that spread itself by exploiting a bug in Outlook Express.
 % Another example is `GhostNet'
@@
 \label{modelsforpropagation}
-To use the models for worm propagation in our model of the FlipIt game it is sufficient to know the propagation speed of the worm and define after how many time units a sufficient number of hosts of the network is infected.
-This will be equal to the \textit{d} parameter in our model.
+To use the models for worm propagation in our model of the FlipIt game it is sufficient to know the propagation speed of the worm and define after how many time units a sufficient number of nodes in the network are infected.
+This will be equal to the \textit{d} parameter in our FlipIt model.
 \\
 
-Scan based or Topology based
+%Scan based or Topology based
 A node in a network can be in three different states: susceptible, infected or removed.
@@
 We are only interested in models where we can find the time that it takes for a worm to infect a sufficient amount of nodes.
+Table \ref{tree} gives an overview of the most common propagation methods with the given type.
+This table comes from the most recent survey \citep{wang2014modeling}, which encompasses results of older surveys.
+Moreover, the survey explicitly focuses on propagation methods, while other papers also focus on other aspects such as  detection mechanisms and containment systems.
 All models that are based on SI model type (Susceptible-Infected) are good, because in our model every node that is infected will stay infected.
-In SIR and SIS models ..
-kunnen we het ook vinden (AAWP).
+Only models Local Preference and Email worms model will not be explained because it involves human interaction.
+OSN model is done by means of simulation.
+The SIR AAWP model is also interesting to explain because the parameter that involves the R part of the model can be deleted.
 \\
-
-In this chapter we gaan dieper op sommige models and toepassen the models on the delay.
-We can divide the propagation models in continuous and discrete time models.
-Continious time models are ..
-Discrete time models are ..
- Table \ref{tree} gives an overview of the taxonomy of the propagation methods are given and the type of model. \todo{deze figuur moet aangepast worden want niet elk model wordt uitgelegd}.\\
-
-For more information about other models the reader is refered to the following papers with other common propagation models: \cite{wang2014modeling}, \cite{OnWorms2005survey}, \cite{xiang2009propagation} and \cite{serazzi2004computer}.
 
@@
 
-The propagation models of diseases are applicable to the propagation of worms.
+%The propagation models of diseases are applicable to the propagation of worms.
 SI, SIS and SIR.
@@
 
-In this chapter we are only interested in the SI model.
+%In this chapter we are only interested in the SI model.
 To apply the propagation models on FlipIt we need to know the propagation speed of the virus, without hosts becoming susceptible again.
@@
 \centering
-\includegraphics[scale=0.4]{Images/tabel3.png} 
+%\includegraphics[scale=0.4]{Images/tabel3.png} 
+%\caption{Taxonomy of worm modelling. Image based on taxonomy given in \cite{wang2014modeling}}.
+%\label{tree}
+%\end{figure}
 \includegraphics[scale=0.55]{Images/tableworms.png}
@@
 
-\subsection{Continuous-time models}
-\subsubsection{Simple Epidemic model}
+\subsection{Simple Epidemic Model}
 %Codered paper gebruikt voor model
-A Simple epidemic model is an SI model. Nodes in the network can be either susceptible or infected. 
-Once a node is infected it will stay infected. Every node has the same chance to be infected
-The simple epidemic model is considered to be of a fixed size. The model for a fixed population is as follows:
+A Simple epidemic model is another name for the general SI model. This model assumes that each node in the network can be either susceptible or infected. 
+Once a node is infected it will stay infected. Every node in the network has the same chance to be infected. 
+The simple epidemic model is considered to be of a fixed size, meaning that no nodes are added to the network or removed. The model for a fixed population is as follows:
 
@@
 \end{equation} 
-where I(t) is the number of infected hosts at time t; $\beta$ is the propagation rate; and N is the number of nodes in the network.
-In the beginning, $t=0$, I(0) hosts are infected.
+where \textit{I(t)} is the number of infected nodes at time \textit{t}, $\beta$ is the propagation rate, and \textit{N} is the number of nodes in the network.
+In the beginning, $t=0$, \textit{I(0)} nodes are infected.
 All the other nodes, $N - I(0)$, in the network are susceptible.
@@
 I = \dfrac{e^{\beta(t-T)}}{1+e^{\beta(t-T)}}
+\label{SIdelay}
 \end{equation}
-where T is a time parameter representing the point of maximum increase in the
-growth.\\
+where \textit{T} is a time parameter representing the point of maximum increase in the growth.\\
 
-\begin{figure}[hbtp]
-\centering
-\includegraphics[scale=0.5]{Images/SEMmodel.png}
-\caption{Simple Epidemic Model. The x-coordinate is the propagation time and the y-coordinate is the infected percentage of the whole network. Original image from \cite{OnWorms2005survey}}
-\label{SEMmodel}
-\end{figure}
+%\begin{figure}[hbtp]
+%\centering
+%\includegraphics[scale=0.5]{Images/SEMmodel.png}
+%\caption{Simple Epidemic Model. The x-coordinate is the propagation time and the y-coordinate is the infected percentage of the whole network. Original image from \cite{OnWorms2005survey}}
+%\label{SEMmodel}
+%\end{figure}
 
-Figuur \ref{SEMmodel} laat het SEM model zien.
-This Simple epidemic model has been used in various papers \cite{OwnInternetSI}, \cite{CodeRed} to model random scanning worm such as \textit{Code Red} and \textit{Slammer}. \\
-For our paper we have to approximate the propagation speed en Infect gelijkstellen aan het aantal genoeg nodes die geinfecteerd moeten zijn door de APT om de controle te hebben.
+%Figure \ref{SEMmodel} laat het SEM model zien.
+%This simple epidemic model has been used in various papers \cite{OwnInternetSI}, \cite{CodeRed} to model random scanning worm such as \textit{Code Red} and \textit{Slammer}. \\
+%For our paper we have to approximate the propagation speed en Infect gelijkstellen aan het aantal genoeg nodes die geinfecteerd moeten zijn door de APT om de controle te hebben.
 Dan weten we hoelang het duurt.
 In het begin zullen er ook geen geinfecteerde hosts zijn, dus de vergelijking kan opgelost worden.\\
-%Er wordt geen rekening gehouden met de topology ? nog uitzoeken.
+%%Er wordt geen rekening gehouden met de topology ? nog uitzoeken.
+To extract the delay from the formula we need the total amount of nodes in the network.
+If formula \ref{SIdelay} is equal to this amount, variable \textit{t} will be the value of the delay that we need.
+This is only applicable to nodes in a homogeneous network.
+A homogeneous network is a network where every node has approximately the same degree.
+This model is thus suitable for propagation of worms that are topology independent (e.g scan-based worms), because every node has the same chance to be infected by another node in the network.
+The \textit{Code Red} worm, which is a random scan based worm, has been analysed by this kind of model in \cite{OwnInternetSI}.
 
-\subsubsection{RCS model}
+\subsection{RCS model}
 
@@
 
-To to apply this formula for the case of FlipIt we need to know the maximum numbers of hosts that can be infected before the total system is compromised.
+Toextract form this formula the delay for the case of FlipIt with a propagation delay, we need to know the maximum numbers of hosts that can be infected before the total system is compromised.
 The initial compromise rate has to be approximated.
 \textit{t} is the variable in our formula and says how many time has passed.
-If a passes a certain value we know the value of t which is the same as d in our FlipIt game.
+If \textit{a} passes the proportion of nodes that have to be compromised before the whole system is compromised by the attacker, value \textit{t}  is the value of \textit{d} in our FlipIt game.
 
@@
 %%spread--
-\subsubsection*{self disciplinary worms}
+\subsection*{self disciplinary worms}
 Model for self disciplinary worms and counter measures ... []  \todo{verder uitwerken ? ja of nee? }
@@
 %The second strategy is almost the same as what email virusses would get. For this reason we work an example out .. 
-\subsection{Discrete-time models}
-\subsubsection{Bluetooth worm model}
-\todo{ref:Modeling Propagation Dynamics of Bluetooth Worms}
+\subsection{Bluetooth worm model}
 
-The Bluetooth worm model is introduced by Yan and Eidenbenz []. The model captures the behavior of the propagation of a worm that spreads through the Bluetooth protocol. 
+The Bluetooth worm model is introduced by Yan and Eidenbenz \citep{yan2009modeling}. The model captures the behaviour of the propagation of a worm that spreads through the Bluetooth protocol. 
 A Bluetooth device that is compromised can only infect neighbour devices that are in its radio range
@@
 
-\subsubsection{OSN worm model}
-OSN worms stands for Online Social Network worms.
+%\subsubsection{OSN worm model}
+%OSN worms stands for Online Social Network worms.
 This model is based on worms that propagate through the use of social media, like Facebook.
@@
 
-\subsubsection{AAWP model}
+\subsection{AAWP model}
 SIR AAWP stands for Analytic Active Worm Propagation.
@@
 The model also considers the fact that a hosts can be scanned and hit by multiple worms at the same time.
-The Epidemiological model ignores the case.
 %This model is used for worms that propagate through uniform scanning.
 When uniform scanning is used, each IP address
+
+The spread of the AAWP model is characterized as follows:
+\begin{equation}
+I_{t+1}=I_{t}+(N-I_{t})[1-(1-\dfrac{1}{\Omega})^{sI_{t}}]
+\end{equation}
+where $I_{t}$ is the amount of infected hosts at time \textit{t}, \textit{N} is the number of vulnerable hosts, \textit{s} is
+the scanning rate of the worm and $\Omega$ is the scanning
+space.
 
@@
 
-A method to calculate the propagation of the virus in an easy way. Google page ranking algorithm.
+%A method to calculate the propagation of the virus in an easy way. Google page ranking algorithm.
 
@@
 As can be seen in table \ref{tree} most propagation models are for a specific network topology.
+The different kinds of network topology of each system is either homogeneous, small-world network, random network or power-law network.
+A homogeneous network is a network where every node has about the same degree of connectivity.
+This means that every node can infect each node with the same opportunity.
+A small-world topology is a network where most nodes are not connected with each other but they can all reach each other in a few steps.
+Social networks are an example of a small-world network.
+A random network is a topology where each connection is chosen at random with equal probability.
+In a power law network the degrees of each node in the network follow the power law.
+That is that some of the nodes have a small degree and other have a very large degree of connection.
 What we want to have is to have a method to calculate the delay that is independent of the network topology.
+A method where we can chose in the beginning which topology we have and then calculate the delay.\\
+
 %All of the above models for modelling the mechanism of the spreading of a worm depend on the kind of topology of the network.  

### 2015-08-07 doc/template/chap-9.tex
nog een updqte
Vanalle aanpassingen

@@
  A Periodic strategy is a non-adaptive renewal strategy where the time intervals between consecutive moves are a fixed period, denoted by $\delta$.
-Moreover it has a random phase, that is chosen uniformly and random in the interval $[0,\delta]$ for the first move.
+It has a random phase, that is chosen uniformly and randomly in the interval $[0,\delta]$ for the first move.
 The average rate of play of a player is denoted by $\alpha_{i} = \dfrac{1}{\delta_{i}}$.
-Stated below a list of symbols that will be used throughout the paper.
-Figure \ref{FlipItDelay} represents the main symbols for clarification.\\
+Given below is a list of symbols that will be used throughout the paper.
+Figure \ref{FlipItDelay} clarifies the main symbols.\\
 ~~\\
@@
 \caption{Formalization of a FlipIt game with delay: A representation of a FlipIt game where both players are playing periodically.
-Every move or flip is indicated by a blue or orange circle, respectively dark gray and light grey.
+Every move or flip is indicated by a green or orange circle, respectively dark gray and light grey.
 The defender is represented in blue (dark grey) and plays with a period of $\delta_{D}$.
@@
 The attacker plays with a period of $\delta_{A}$.
-The blue and orange rectangles represent the amount of time the respective player is in control of the resource.}
+The green and orange rectangles represent the amount of time the respective player is in control of the resource.}
 \label{FlipItDelay}
@@
 \item $i$: Denotes the player.
-`D' denotes the defender, and `A' denotes the attacker which differs form the notation in \citep{FlipIt}, where the defender is denoted by the subscript 0 and the attacker by the subscript 1.
+\textit{D} denotes the defender, and \textit{A} denotes the attacker which differs form the notation in \citep{FlipIt}, where the defender is denoted by the subscript \textit{0} and the attacker by the subscript \textit{1}.
 \item $\delta_{i}$: The length of the interval between two consecutive moves of player \textit{i}. 
@@
 \item $\beta_{i}$:  The average benefit rate up to time \textit{t}, defined as  $\beta_{i} = \gamma_{i} -k_{i} \alpha_{i} $.
-\item $opt_{i}$: The optimum function for player \textit{i}. \todo{optimum function equal or not with best response ?}
+\item $opt_{i}$: The optimum function for player \textit{i}. 
 \end{description}
 
-The adaptation of the FlipIt model starts from the assumption that when an attacker attacks at time \textit{t}, he doesn't get immediate control over the resource, but he only gains control at time \textit{t + d}, with $d$ denoting the time needed to infect a sufficient number of (or all) nodes.
+The adaptation of the FlipIt model starts from the assumption that, when an attacker attacks at time \textit{t}, he doesn't get immediate control over the resource, but he only gains control at time \textit{t + d}, with $d$ denoting the time needed to infect a sufficient number of (or all) nodes.
 If the defender flips the network before the period $d$ has elapsed (so, somewhere between $t$ and $t + d$), then the attacker will never gain full control over the resource.
 (See figure \ref{dt} at point 3 and 4).
-This implies that the mathematical formulas for gain and benefit need to be adapted to the fact that the attacker loses part of its benefit because of this delay.
+This implies that the mathematical formulas for gain and benefit need to be adapted to the fact that the attacker loses part of his benefit because of this delay.
 In the remainder of this paper, we will adapt the formalization of the FlipIt game using the delay variable $d$.
@@
 \caption{The first game is the basic FlipIt game.
-The second is a FlipIt game with a delay.
+The second is the FlipIt game with a delay.
 During the first flip of the attacker, the defender moves after the delay, causing the attacker to get control over the resource.
@@
 Similarly as in \cite{FlipIt}, we split the formalization in two cases.
-The first case is where the defender plays at least as fast as the attacker, the second case is where the attacker plays at least as fast as the defender.
-For each of these cases, first the benefit formula of the basic case without delay is presented, and subsequently the delay is introduced.
+In the first case the defender plays at least as fast as the attacker, in the second case the attacker plays at least as fast as the defender.
+For each of these cases, the benefit formula of the basic case without delay is presented first, and subsequently the delay is introduced.
 \\
 
-Intuitively, we could assume that $d$ can never be bigger than $\delta_{A}$ because then the benefit for the defender would always be 1, but this is not always true.
-Assume for example that an attacker plays with an interval of 3 units, the delay is equal to 4 units and that the defender only plays every 8 units.
+Intuitively, we could assume that $d$ can never be bigger than $\delta_{A}$ because then the attacker would play again before the delay has finished.
+This would seemingly result in a gain for the defender that is always 1, but this is not always true.
+Assume for example that an attacker plays with an interval of 3 time units, that the delay is equal to 4 time units and that the defender only plays every 8 time units.
 This situation is represented in figure \ref{langedelay}.
-Since the delay is shorter that the period of the defender, the attacker takes control of the resource once the delay has elapsed, until the defender plays.
+Since the delay is shorter than the period of the defender, the attacker takes control of the resource once the delay has elapsed, until the defender plays.
 However, if the delay is larger that the period of the defender, then the defender will always be in control.
-This situation is represented in figure \ref{langeredelay}, where the attacker plays with an interval of 5 units, the delay is equal to 6 units and the defender plays with an interval of 4 units.
+This situation is represented in figure \ref{langeredelay}, where the attacker plays with an interval of 5 time units, the delay is equal to 6 time units and the defender plays with an interval of 4 time units.
 From this we can conclude that it only makes sense to calculate the formulas for the cases where $d$ is smaller than $\delta_{D}$.
-We can already conclude that it is no use for the attacker to play when the delay is bigger than $\delta_{D}$.
+We can already conclude that it is of no use for the attacker to play when the delay is bigger than $\delta_{D}$.
 
@@
 However, when introducing the delay, some formulas become much simpler when expressed in terms of $\delta_{D}$ and $\delta_{A}$.
-Therefore in the remainder of this paper, we will formulate the model using both $\alpha_{i}$ and $\delta_{i}$, depending on what result is the simplest formula.
+Therefore in the remainder of this paper, we will formulate the model using both $\alpha_{i}$ and $\delta_{i}$, depending on which variable gives the simplest representation of the formulas.
 
@@
 
-The expected period of attacker control within the interval would be r/2, without considering the delay by a virus.
+The expected period of attacker control within the interval as the moment on which the attacker gains control is uniformly distributed over the defender's interval.
+On average the attacker will gain control at time r/2.
+So the expected gain is equal to the remainder of the defender interval, i.e.
+r/2, without considering the delay by a virus.
 Therefore the benefit for the attacker, without considering the delay, can be expressed as follows:
@@
 While the probability that the attacker will move in the interval of the defender is still \textit{r}, the gain will not be half of the interval.
-Indeed, if the attacker plays after $\delta_{D}-d$, given the delay \textit{d}, he will never get in control in that interval (see figure \ref{tijdens interval}).
+Indeed, if the attacker plays after $\delta_{D}-d$, given the delay \textit{d}, he will never gain control in that interval (see figure \ref{tijdens interval}).
 \begin{figure}[hbtp]
@@
 \label{tijdens interval}
-\end{figure}The probability that the attacker plays soon enough is $\dfrac{\delta_{D}-d}{\delta_{D}}$ and this will give the attacker an average gain of $\dfrac{\delta_{D}-d}{2}$.
+\end{figure}The probability that the attacker plays early enough is $\dfrac{\delta_{D}-d}{\delta_{D}}$, giving the attacker an average gain of $\dfrac{\delta_{D}-d}{2}$ (the average remainder of the defender interval after the attacker flipped).
 If the attacker moves after the period of $\delta_{D}-d$, the gain of the attacker will be zero.
@@
 \end{equation*}
-Thus over the entire game, the average gain rate is:
+As the formula above is valid for each defender interval, the average gain rate over the entire game is:
 \begin{equation*}
@@
 
-To derive the benefit, the cost of moving is subtracted from the average gain. 
+To find the benefit, the cost of moving is subtracted from the average gain. 
 \begin{equation*}
@@
 ~~\\
-We can easily see that when $d$=0, we obtain the formula of the original FlipIt game.\\
+For this case where $d$=0, we this equals the formula of the original FlipIt game \citep{FlipIt} [p675].\\
 
@@
 \end{equation*}
-\todo{formule nog eens nakijken}
+
 This however results in an overestimation. 
-By simulation the game, it can be notices that the closer $\delta_{A}/\delta_{D}$ is equal to one, the better the approximation.
-If $\delta_{A}/\delta_{D} = 1$ the result is correct.
+%By simulation the game, it can be notices that the closer $\delta_{A}/\delta_{D}$ is equal to one, the better the approximation. If $\delta_{A}/\delta_{D} = 1$ the result is correct. 
 The reason this formula overestimates the benefit of the attacker is that it assumes that the defender is always in control during the delay.
 However, if the attacker was in control in the previous interval, then he continuous to be in control during the period of the delay, see figure \ref{fig:case2}.
-This means that to find the correct formula, what happens in the previous interval needs to be taken into account.
+This means that the average benefit formulas for this case cannot be derived from one interval only; what happens in the previous interval must be taken into account.
 \\
@@
 We know that the defender's moves are instantaneous.
-Therefore, it is easier in this case to calculate the benefit of the defender.
+Therefore, it is easier to calculate the benefit of the defender.
 Because the defender moves slower than the attacker we know that if the defender moves during the interval of the attacker, he only moves once within this interval.
+
 The defender will move during the interval of the attacker with a probability of $\dfrac{\delta_{A}}{\delta_{D}} $.
-When this happens, the defender will be in control for the remainder of this interval.
+If this happens, the defender will be in control for the remainder of this interval.
 In the next interval the attacker will have to regain control, meaning that during the delay, the defender stays in control, see figure \ref{fig:case2} cases (1) to (4).
@@
 
-For the case with a virus we consider two cases, Case a and Case b, depending on whether the delay is shorter or longer than the difference between the attacker's and the defender's period.  \\
+For the case with a delay we consider two cases, Case a and Case b, depending on whether the delay is shorter or longer than the difference between the attacker's and the defender's period.  \\
 
@@
 \subsubsection*{\textbf{Case a:} $d + \delta_{A} \leq \delta_{D}$}
-The first case is where the defender's period is larger than the sum of the delay and the attacker's period.
-In that case there is no problem of counting the delay twice.
-To determine the total gain rate of the defender we need to know what the probability is that the defender will move during an interval and what the average time is that the defender controls the resource.
-Given that if the defender moves he will always benefit entirely from a period of delay in the next interval of the attacker, his total gain is $\dfrac{(d+\dfrac{\delta_{A}}{2})}{\delta_{A}}$ in one interval.
+%The first case is where the defender's period is larger than the sum of the delay and the attacker's period. 
+In this case the delay will never be counted twice in the defender's  benefit formula.
+To determine the total gain rate of the defender we need to know the probability that the defender will move during an interval and what the average time is that the defender controls the resource.
+Given that if the defender moves he will always benefit entirely from a period of delay in the next interval of the attacker, his total gain is $\dfrac{(d+\dfrac{\delta_{A}}{2})}{\delta_{A}}$ \todo{formule valt uit de lucht} in one interval.
 The total gain  rate of the defender is then the probability that the defender will move during an interval of the attacker multiplied by the total average gain per interval:
@@
 \end{equation*}\\
-This yields in the following benefit formula:
+This yields the following benefit formula:
 \begin{equation*}\label{first}
@@
 It is crucial that $ \delta_{D}$ is at least as large as $d + \delta_{A}$.
-If not, this would mean that the defender can move during the delay in the interval following the interval where the defender already moved.
-This would mean that there can be an overlap between the average gain of $\dfrac{\delta_{A}}{2} +d$ and the delay.
+If not, the defender can move during the delay in the interval following the interval in which the defender already moved.
+This would result in an overlap between the average gain of $\dfrac{\delta_{A}}{2} +d$ and the delay.
 The above benefit formula would then include to much gain for the defender: the potential overlap during the delay would be counted twice.
@@
 To obtain the formula in case of a too long delay, we therefore need to subtract this overlapping gain from the above formula. 
-Since $\delta_{D} \geq \delta_{A}$, if the defender enters the interval immediately after the attacker has played, then the defender cannot have played in the previous interval.
+Since $\delta_{D} \geq \delta_{A}$, if the defender enters the interval immediately after the attacker has played, the defender cannot have played in the previous interval.
 In that case, there is no overlap.
@@
 Concretely, only the last part of length $d - (\delta_{D} - \delta_{A})$ is subject to overlap.
-Hence, the probability of overlap is $\dfrac{ d - (\delta_{D} - \delta_{A})}{\delta_{D}}$ and the gain will be half of this interval:  $\dfrac{ d - (\delta_{D} - \delta_{A})}{2}$.
+Hence, the probability of overlap is $\dfrac{ d - (\delta_{D} - \delta_{A})}{\delta_{D}}$ and the average gain will be half of this interval:  $\dfrac{ d - (\delta_{D} - \delta_{A})}{2}$.
 The gain rate to be subtracted is therefore:\\
@@
 \begin{equation*}
-\dfrac{1} {\delta_{A}} \cdot \dfrac{d - (\delta_{D} - \delta_{A})}{\delta_{D}} \cdot \dfrac{d - (\delta_{D} - \delta_{A})}{\delta_{D}}
+\dfrac{1} {\delta_{A}} \cdot \dfrac{d - (\delta_{D} - \delta_{A})}{\delta_{D}} \cdot \dfrac{d - (\delta_{D} - \delta_{A})}{2}
 \end{equation*}
 
-The total gain  rate of the defender is obtained by subtracting this term from the gain rate of case a:
+The total gain rate for the defender is obtained by subtracting this term from the gain rate of case a:
  \begin{equation*}
@@
 
-The benefit for the attacker will be as follows:
+Consequently, the benefit for the attacker will be:
 \begin{equation*}

### 2015-08-07 doc/template/nash.tex
Vanalles
Vanalle aanpassingen

@@
 To determine $opt_{D}(\delta_{A})$ we need to compute the derivative of  $\beta_{D}(\delta_{D},\delta_{A}) $ for a fixed $\delta_{A}$.
- We consider two cases, where case 2 is divided into two subcategories. Fig \todo{fig} shows all the cases.\\
-\begin{figure}[hbtp]
-\centering
-\includegraphics[scale=0.4]{Images/bestresp.png}
-\caption{This figure shows the all the cases with their subcategories.
+ We consider two cases, where case 2 is divided into two subcategories. 
+% Fig \todo{fig} shows all the cases.\\
+%\begin{figure}[hbtp]
+%\centering
+%\includegraphics[scale=0.4]{Images/bestresp.png}
+%\caption{This figure shows the all the cases with their subcategories.
 'A' stands for Case 2.A: $\delta_{D} \geq d+\delta_{A} \geq \delta_{A}$ and 'B' stands for Case 2.B: $d+\delta_{A} \geq \delta_{D} \geq  \delta_{A} $ }
-\label{grafiekbestr}
-\end{figure}
+%\label{grafiekbestr}
+%\end{figure}
 
@@
 
-Given that $\delta_{D}$ is always positive, the benefit function can be either positive or negative depending on the numerator of the above fraction.
+Given that $\delta_{D}$ is always positive, the benefit function can be either increasing or decreasing depending on the numerator of the above fraction.
 For $\delta_{A} + 2(d-k_{D}) > 0$ the benefit will be always positive but decreasing, see figure \ref{ShapeUp}.
 \\
-The defender will play if $ \delta_{A} > 2(k_{D} -d)$ and $k_{D} > d$. The defender will not play if $k_{D} < d$ because then $ \delta_{A}$ is negative.\\
+The defender will always play if $ \delta_{A} > 2(k_{D} -d)$ for $k_{D} < d$ or $k_{D} > d$ because $\delta_{A}$ will be positive in either case.
+Their is an edge case if $d=k_{D}$.
+The benefit will result in $\beta_{D}(\delta_{D},\delta_{A})= \dfrac{\delta_{A}}{\delta_{D}}$.4 \\
 \begin{figure}
@@
 \includegraphics[scale=0.5]{Images/ShapesUp.png} 
-\caption{The benefit function is of the shape of 1/x if $\delta_{D}$ is always positive and if $\delta_{A} + 2(d-k_{D}) > 0$. }
+\caption{The benefit function is of the shape of 1/x if $\delta_{D}$ is always decreasing and if $\delta_{A} + 2(d-k_{D}) > 0$. }
 \label{ShapeUp}
@@
 \includegraphics[scale=0.5]{Images/ShapeDown.png} 
-\caption{The benefit function is of the shape of -1/x if $\delta_{D}$ is always positive and if $\delta_{A} + 2(d-k_{D}) < 0$.}
+\caption{The benefit function is of the shape of -1/x if $\delta_{D}$ is always increasing and if $\delta_{A} + 2(d-k_{D}) < 0$.}
 \label{ShapeDown}
@@
 \begin{description}
-\item decreasing $ \delta_{A} < 2(k_{D} -d)$, $ \delta_{A}$ is negative
-\item increasing  $\delta_{A} > 2(k_{D} -d)$ 
+ \item $\beta_{D}$ will be decreasing but always positive. If we minimize $\delta_{D}$ the value of $\beta_{D}$ will be higher. 
 \end{description}
@@
 \begin{description}
-\item increasing $ \delta_{A} < 2(k_{D} -d)$
-\item decreasing  $\delta_{A} > 2(k_{D} -d)$ 
+\item if $ \delta_{A} > 2(k_{D} -d)$, \\
+$\beta_{D}$ will be decreasing but always positive. If we minimize $\delta_{D}$ the value of $\beta_{D}$ will be higher. 
+\item if  $\delta_{A} = 2(k_{D} -d)$, \\
+the benefit of the defender will be $\beta_{D}=0$.
+\item if $\delta_{A} > 2(k_{D} -d)$, \\
+$\beta_{D}$ will be increasing but always negative. In this case the defender will not play. 
 \end{description}
@@
 For case 2.B this leads to the following deduction, which results in the same formula as for case 1: The function increases on $[0, \sqrt{2\delta_{A}k_{D} + d^{2}}]$ and is decreasing on $[\sqrt{2\delta_{A}k_{D} + d^{2}}, \infty]$.
-So there is a maximum on $\delta_{D} = minimum \{ \delta_{A}, \sqrt{2\delta_{A}k_{D} + d^{2}} \} $ \\
+So there is a maximum on $\delta_{D} = minimum \{ \delta_{A}+d, \sqrt{2\delta_{A}k_{D} + d^{2}} \} $ \\
 
@@
 
-From this analyses we can compute $opt_{D}(\delta_{A})$ for two different cases as:
+From this analyses we can compute $opt_{D}(\delta_{A})$ for two different cases as: \\
+
+For case  $k_{D} \leq d$:
  \begin{displaymath}
@@
 
-For case :
+For case $k_{D} > d$ :
  \begin{displaymath}

### 2015-08-08 doc/template/nash.tex
chapter nash

@@
 \subsection*{Case 1: $\delta_{D} \leq \delta_{A} $}
-The benefit formula obtained in the previous chapter for this case is as follows:
+The benefit formula obtained in the previous chapter (formula \ref{Benfcase1:defender}) for the defender in this case is as follows:
 \begin{equation*}
-\beta_{D}(\delta_{D},\delta_{A}) = 1 - \dfrac{\delta_{D}}{2\delta_{A}} - \dfrac{k_{D}}{\delta_{D}} - \dfrac{d^{2}}{2\delta_{D}\delta_{A}} + \dfrac{d}{\delta_{A}}
+\beta_{D}(\delta_{D},\delta_{A}) = 1 - \dfrac{\delta_{D}}{2\delta_{A}} - \dfrac{d^{2}}{2\delta_{D}\delta_{A}} + \dfrac{d}{\delta_{A}}  - \dfrac{k_{D}}{\delta_{D}}
 \end{equation*}
-If we take the partial derivative of this formula for a fixed $\delta_{A}$ we get the following:
+
+To know if the function decreases or increases we take the partial derivative of this formula for a fixed $\delta_{A}$:
 \begin{equation*}\label{formdelta}
@@
 
-The benefit formula obtained in the previous chapter for this case is as follows:
+The benefit formula obtained in the previous chapter (formula \ref{benfcase2a:defender}) for the defender in this case is as follows:
 \begin{equation*}
@@
 
-Given that $\delta_{D}$ is always positive, the benefit function can be either increasing or decreasing depending on the numerator of the above fraction.
+Given that $\delta_{D}$ is always positive, the benefit function can be either increasing or decreasing depending on the numerator of the above fraction. \\
+
 For $\delta_{A} + 2(d-k_{D}) > 0$ the benefit will be always positive but decreasing, see figure \ref{ShapeUp}. 
+The defender will always play if $\delta_{A} + 2(d-k_{D}) > 0$ for $k_{D} < d$ or $k_{D} > d$ because $\delta_{A}$ will be positive in either case.
+There is an edge case if $d=k_{D}$, which results in a benefit of $\beta_{D}(\delta_{D},\delta_{A})= \dfrac{\delta_{A}}{\delta_{D}}$.
 \\
-The defender will always play if $ \delta_{A} > 2(k_{D} -d)$ for $k_{D} < d$ or $k_{D} > d$ because $\delta_{A}$ will be positive in either case.
-Their is an edge case if $d=k_{D}$.
-The benefit will result in $\beta_{D}(\delta_{D},\delta_{A})= \dfrac{\delta_{A}}{\delta_{D}}$.4 \\
 \begin{figure}
@@
 
-For $\delta_{A} + 2(d-k_{D}) < 0$, the benefit will always be negative so the defender will not play. See figure \ref{ShapeDown}. This is independent of the value of $k_{D}$ and $ d$.
+For $\delta_{A} + 2(d-k_{D}) < 0$, the benefit will always be negative so the defender will not play. See figure \ref{ShapeDown}. This is independent of the value of $k_{D}$ and $ d$. \\
 
@@
 %
-This leads to the following deduction:
-\begin{description}
-\item If $k_{D} \leq d$ 
-\begin{description}
- \item $\beta_{D}$ will be decreasing but always positive. If we minimize $\delta_{D}$ the value of $\beta_{D}$ will be higher. 
-\end{description}
-\item If $k_{D} > d$ 
-\begin{description}
-\item if $ \delta_{A} > 2(k_{D} -d)$, \\
-$\beta_{D}$ will be decreasing but always positive. If we minimize $\delta_{D}$ the value of $\beta_{D}$ will be higher. 
-\item if  $\delta_{A} = 2(k_{D} -d)$, \\
-the benefit of the defender will be $\beta_{D}=0$.
-\item if $\delta_{A} > 2(k_{D} -d)$, \\
-$\beta_{D}$ will be increasing but always negative. In this case the defender will not play. 
-\end{description}
-\end{description}
+%This leads to the following deduction:
+%\begin{description}
+%\item If $k_{D} \leq d$ 
+%\begin{description}
+% \item $\beta_{D}$ will be decreasing but always positive. If we minimize $\delta_{D}$ the value of $\beta_{D}$ will be higher. 
+%\end{description}
+%\item If $k_{D} > d$ 
+%\begin{description}
+%\item if $ \delta_{A} > 2(k_{D} -d)$, \\
+%$\beta_{D}$ will be decreasing but always positive. If we minimize $\delta_{D}$ the value of $\beta_{D}$ will be higher. 
+%\item if  $\delta_{A} = 2(k_{D} -d)$, \\
+%the benefit of the defender will be $\beta_{D}=0$.
+%\item if $\delta_{A} < 2(k_{D} -d)$, \\
+%$\beta_{D}$ will be increasing but always negative. In this case the defender will not play. 
+%\end{description}
+%\end{description}
 ~~\\
@@
 
-The benefit formula obtained in the previous chapter for this case is as follows:
+The benefit formula obtained in the previous chapter  (formula \ref{benfcase2b:defender}) for the defender in this case is as follows:
 \begin{equation*}
@@
 
-For case 2.B this leads to the following deduction, which results in the same formula as for case 1: The function increases on $[0, \sqrt{2\delta_{A}k_{D} + d^{2}}]$ and is decreasing on $[\sqrt{2\delta_{A}k_{D} + d^{2}}, \infty]$.
-So there is a maximum on $\delta_{D} = minimum \{ \delta_{A}+d, \sqrt{2\delta_{A}k_{D} + d^{2}} \} $ \\
+For case 2.B this leads to the following deduction, which results in the same formula as for case 1 but with a small difference for the value of $\delta_{D}$: The function increases on $[0, \sqrt{2\delta_{A}k_{D} + d^{2}}]$ and is decreasing on $[\sqrt{2\delta_{A}k_{D} + d^{2}}, \infty]$.
+Because $\delta_{D} \geq \delta_{A}$ there is a maximum on $\delta_{D} = maximum \{ \delta_{A}, \sqrt{2\delta_{A}k_{D} + d^{2}} \} $ or because $d+\delta_{A} \geq \delta_{D}$ there is also on $\delta_{D} = minimum \{ \delta_{A}+d, \sqrt{2\delta_{A}k_{D} + d^{2}} \} $.
+\todo{beter uitschrijven, deltaD moet tussen die twee waarden liggen}\\
 

### 2015-08-08 doc/template/chap-4.tex
Laatste blabla pagerank

@@
 The method is based on the method for Google Page Ranking \cite{GoogleRank}.
+
+\subsubsection{Google PageRank algorithm}
+The PageRank algorithm is introduced by Page and Brin in 1998 as one of the main features of the search engine Google to improve the search results.
+PageRank models the human behaviour when users surf through the net.
+It can also model \textit{random surfers}.
+A random surfer is the probability that a surfer gets bored and randomly visits another page which is not linked with the initial page.
+The probability that he will visit a random page is the PageRank of that page.
+A page will have a high PageRank if many pages will point to this page.
+This means that this page is well cited through other pages and maybe an important one to look at.
+The main idea of the PageRank algorithm is to look at the number of (important) web pages that point to a particular page.
+The ranking of this page depends than on the number of out coming links and the importance of the pages that link to this page.
+With a probability of $P$ the surfer will follow a link of the page to another page, and with a probability of $1-P$ the surfer will surf to a random page.
+The PageRank algorithm is expressed as follows:
+\begin{equation}
+PageRank(A)=P \sum \dfrac{PageRank(i)}{d_{out,i}} + (1-P) \cdot e_{A}
+\end{equation}
+where $N_{A}$ is the number of pages, $PR(i)$ is the ranking of web page $i$, $D_{out,i}$ is the number of outgoing links of page $i$, $(1-P)$ the probability that a random page is searched, and $e_{A}$ the restart value for web page A which is often uniformly distributed among all web pages.
+
+\subsubsection{Model with matrix}
+
 For some of the spreading methods, the graph of the network matters.

### 2015-08-08 doc/template/chap-4.tex
vanalles

@@
 
-The formalisation of the FlipIt game with delay relies on some value  \textit{d} that represents the time needed to infect a sufficient number of nodes in a network after the initial infection.
+The formalisation of the FlipIt game with a delay relies on some value  \textit{d} that represents the time needed to infect a sufficient number of nodes in a network after the initial infection.
 This chapter provides the reader with more insight on how to calculate the value of this parameter \textit{d}.
@@
 The  spreading of malware has already been extensively researched.
-Because of the many different types of propagation it is hard to define a single model that can model all of them.
-Modelling the spread of malware depends on two key factors: the method used for the propagation and the graph of the network where the malware will spread itself.
+Because of the many different types of propagation methods it is hard to define a single model that can model all of them.
+Modelling the spread of malware depends on two key factors: the method used for the propagation and the graph of the network in which the malware will spread itself.
 \\
 Viruses and worms are the types of malware that are the most researched.
-Since the spreading of viruses requires human interaction, their propagation delay is depending on (hard to predict) human behaviour.
+Since the spreading of viruses requires human interaction, their propagation delay depends on (hard to predict) human behaviour.
 Worms on the other hand, spread without human intervention, and their propagation is therefore easier to model.
-Since the purpose of this chapter is only to illustrate how a delay can be calculated, we will limit the presentation to propagation models of worms.
+Since the purpose of this chapter is only to illustrate how a delay can be calculated, we will limit this chapter to propagation models of worms.
 \\
 
-The overview of this chapter will be as follow: Section \ref{methodsofpropagation} presents an overview of the most frequently used propagation methods by worms.
-These propagation methods will be covered by different kinds of models in section \ref{modelsforpropagation} that presents models to determine the time a worm needs to infect a sufficient number of nodes.
+The overview of this chapter will be as follows: Section \ref{methodsofpropagation} presents an overview of the most frequently used propagation methods by worms.
+These propagation methods will be covered by different kinds of models in section \ref{modelsforpropagation} presents models to determine the time a worm needs to infect a sufficient number of nodes.
 These results can be used for the completion of parameter \textit{d}.
-Finally, in section \ref{eigenmatrixmethode},  we introduce an easy method to calculate the delay of the propagation of a worm.
+\todo{wat wordt er met de vorige zin bedoeld?} Finally, section \ref{eigenmatrixmethode}, introduces an easy method to calculate the delay of the propagation of a worm.
 % Different kind of graphs to model a computer network are introduced.
@@
 The dropper is the initial attack vector that compromises the single targeted node of the system.
-The second kind of APT's target multiple victims.
+The second kind of APT's target multiple nodes.
 These APT's also use a dropper mechanism but have an additional mechanism for self-propagation in order to propagate themselves to the multiple targeted nodes of the system.
 If the APT uses virus spreading, the virus infects one node on the network and has to wait for human interaction to spread.
-So the spreading speed depends on the human interaction, and modelling virus propagation therefore requires modelling human behaviour.
+As such the spreading speed depends on the human interaction, and modelling virus propagation therefore requires modelling human behaviour.
 If an APT uses a worm propagation method it will spread by itself after it has been dropped on the network.
-Given the additional complexity of incorporating the human factor in a spreading method and that this chapter is merely meant as illustration on calculating the delay for use in the game theoretic approach, we will limit the presentation to worm propagation models for APT's.
+Given the additional complexity of incorporating the human factor in a spreading method and that this chapter is merely meant as illustration on calculating the delay for use in the game theoretic approach, this chapter will be limited to worm propagation models for APT's.
 \\

### 2015-08-09 doc/template/chap-4.tex
Jeroen verbeteringe

@@
 Different dropping mechanisms can be used, whereby the initial attack can be random or targeted at a specific node.
-Frequently used mechanisms are USB sticks (given to a specific person or left behind for pick up by a random person), email, or malicious software through fishing or trojan horses.
-More important for determining our delay is the propagation strategy.
-Propagation is performed by determining the next nodes to spread the worm to.
+Frequently used mechanisms are USB sticks (given to a specific person or left behind to be picked up by a random person), email, or malicious software through fishing or trojan horses.
+To determine the total delay, however, the propagation strategy is of higher importance than the drop mechanism.
+Propagation is achieved by determining the next nodes to spread the worm to.
 The following are common propagation methods:
@@
 
-\item \textbf{Selective Random scanning:} Some worms propagate by using the method of selective random targeting IP addresses.
-The worm selects randomly a part of the selected address space instead of scanning the whole address space.
+\item \textbf{Selective Random scanning:} The worm randomly  selects a part of the selected IP address space instead of scanning the whole address space.
 The reserved address blocks and the unassigned addresses are excluded from the address space.
@@
 Example of such worms are \textit{Code Red} \citep{OwnInternetSI} and \textit{Slammer} \citep{moore2003inside}.
+\todo{Voorbeeeld als mensen het idee niet snappen?}2111
 

### 2015-08-09 doc/template/chap-4.tex
verbeteringen Jeroen nog

@@
 \label{modelsforpropagation}
-To use the models for worm propagation in our model of the FlipIt game it is sufficient to know the propagation speed of the worm and define after how many time units a sufficient number of nodes in the network are infected.
+To use the models for worm propagation in the model of the FlipIt game with propagation delay it is sufficient to know the propagation speed of the worm and determine after how many time units a sufficient number of nodes in the network is infected.
 This will be equal to the \textit{d} parameter in our FlipIt model.
@@
 With these three states three main propagation models are proposed: SI, SIR and SIS.
-SI model is a model where a node once infected stays infected.
-In an SIR model a node can once it is infected become removed.
-In an SIS model a node can become again susceptible after it has been infected.
+In the SI model, a node that has once been infected, stays infected.
+In the SIR model, an infected node can be removed afterwards.
+This node cannot become infected again.
+In the SIS model a node can become susceptible again after it has been infected.
 Based on these three models various other models have been proposed.
-We are only interested in models where we can find the time that it takes for a worm to infect a sufficient amount of nodes.
+\todo{referenties naar andere modellen} We are only interested in models that will allow us to find the time required for a worm to infect a sufficient amount of nodes.
 Table \ref{tree} gives an overview of the most common propagation methods with the given type.
@@
 All models that are based on SI model type (Susceptible-Infected) are good, because in our model every node that is infected will stay infected.
-Only models Local Preference and Email worms model will not be explained because it involves human interaction.
+Only the Local Preference and Email worms models will not be explained because they involve human interaction.
 OSN model is done by means of simulation.
-The SIR AAWP model is also interesting to explain because the parameter that involves the R part of the model can be deleted.
-\\
+The SIR AAWP model is also interesting because the parameter that involves the R part of the model can be deleted.
+\\ \todo{laatste zinnen zijn wazig: Het deel erboven over SIS enzo was duidelijk maar hier worden andere modellen er bij de haren bijgesleurd.
+Als we die niet bekijken, gewoon kort zeggen dat er andere modellen zijn (e.g.
+SIR AAWP, Email worms, ....) maar dat deze niet worden bekeken ofzo?}
 

### 2015-08-09 doc/template/chap-5.tex
Matthias opmerkingen

@@
 This chapter provides the reader with an introduction to the general context of the work presented in this paper.
-Section 2.1 introduces the reader into the basic concepts of cyber security and the kind of cyber security threats that are in scope for this work.
+Section 2.1 introduces the reader into the basic concepts of cyber security and the kind of cyber security threats that are in scope of this work.
 Section 2.2 then introduces the reader into the main principles of game theory.
-Subsequently, section 2.3 introduces the reader to the FlipIt game: the game that will be used to model cyber security attacks of a periodic nature, including a delay.
+Subsequently, section 2.3 introduces the reader to the FlipIt game: this game will be used to model cyber security attacks of a periodic nature, including a delay.
 Finally, section 2.4 gives an overview of the related work, and how the research presented in this paper compares to existing results.
-
+%- Remark: to the reader weg
 \section{What is security?}
@@
 
-Information security is the generic term for data protection stored on a computer controlled device such as computers and smartphones, as well as public and private computer networks, including the entire Internet.
+Information security is the generic term for protection of data stored on a computer controlled device such as computers and smartphones, as well as public and private computer networks, including the entire Internet.
 Security is a general term that encompasses several dimensions. More specifically, information security has three fundamental key objectives:
@@
  \\
-Information security can be divided into two main sub categories.
+Information security can be divided into two main subcategory.
 One of them is cyber security, also known as computer security.
@@
 
-In order to keep a system secure and to be able to apply security tools, it is important to mitigate the possible threats.
+%In order to keep a system secure and to be able to apply security tools, it is important to mitigate the possible threats. -Remark-Zegt te weinig volgens Matthias 
 The possible threats can take many forms, the most common being: spam, malware, spoofing, phishing and  DDoS attacks.
@@
 \item \textit{Virus}: This is a malicious piece of code that replicates itself and tries to spread in order  to infect other systems or files.
-A typical virus will attach itself to a program, or an executable content on a computer.
+A typical virus will attach itself to a program, or executable content on a computer.
 The \textit{I love you} virus is an example of a virus where the virus attached itself as an executable to a mail.
@@
 ~~\\
-As proposed by Kasperksy \cite{APTKaspersky} ,threats caused by the malware described above can be divided into three main categories: known threats (70\%), unknown threats (29\%) and advanced threats (1\%).
+As proposed by Kasperksy \cite{APTKaspersky}, threats caused by the malware described above can be divided into three main categories: known threats (70\%), unknown threats (29\%) and advanced threats (1\%).
 
@@
 They are the most difficult to deal with.
+
+%-Remark- APTs en niet APT's
 
@@
 The motive of an APT is mostly cyber espionage, stealing sensitive data, sabotage or other kinds of ideological attacks.
-APT's are `advanced' because these attacks are well funded and because the attacker (usually) needs a great amount of expertise to successfully penetrate a network.
-Not all APT's are technically advanced though.
+APTs are `advanced' because these attacks are well funded and because the attacker (usually) needs a great amount of expertise to successfully penetrate a network.
+Not all APTs are technically advanced though.
 The attacker can also try to exploit known vulnerabilities, in the hope that his target has not yet secured itself against them.
@@
 According to a security survey of Kaspersky \cite{SurveyKaspersky}  the damage of one successful targeted attack against a large company can exceed 2.5 million dollars.
-As such, companies need a defence mechanism to defend themselves against APT's.
-As previously stated, simple detection and identification tools are insufficient to protect oneself against APT's, and a removal tool will only work if the threat has been identified.
+As such, companies need a defence mechanism to defend themselves against APTs.
+As previously stated, simple detection and identification tools are insufficient to protect oneself against APTs, and a removal tool will only work if the threat has been identified.
 To mitigate these kind of attacks another security countermeasure is needed.
@@
 %This is where the researchers of RSA come into the picture. \todo{deze paper gaat zien of de game theoretische model kan leiden tot resultaten die kunnen helpen in het verdedigen tegen APTs}
-This paper proposes an appropriate game-theoretical modelling of defending a network against APT's and will analyse it in order to draw the necessary strategic conclusions to mitigate these kind of attacks.
+This paper proposes an appropriate game-theoretical modelling of defending a network against APTs and will analyse it in order to draw the necessary strategic conclusions to mitigate these kind of attacks.
 
@@
 \label{Cha1:briefintro}
+%-Remar- based on the work of - Matthias zegt namen erbij zetten
 Game theory is a mathematical study to analyse interactions between independent and self-interested agents.
@@
 %------------------------------------------------%
-Game theory is a mathematical way of modelling the interactions between two or more agents where the outcomes depend on what each agent does and how these interactions should be structured to lead to good outcomes.
+Game theory is a mathematical way of modelling the interactions between two or more agents where the outcomes depend on what each agent does and and the study of how these interactions should be structured to lead to good outcomes.
 Game theory therefore has important applications in many areas such as economics, politics, biology, computer science, philosophy and a variety of other disciplines.
-It gained recognition during the Second World War, when Oskar Morgenstern and John von Neumann both published a book on game theory, titled "Theory of Games and Economic Behavior" in 1944 \cite{von1944theory}.
+It gained recognition during the Second World War, when Oskar Morgenstern and John von Neumann both published a book on game theory, titled ``Theory of Games and Economic Behavior'' in 1944 \cite{von1944theory}.
 This book addressed the mathematical analysis of a series of thinking games.
@@
 
-One of the underlying assumptions game theory is that the players of the game are independent and self-interested.
+One of the assumptions underlying game theory is that the players of the game are independent and self-interested.
 This does not necessarily mean that they want to harm other agents or that they only care about themselves.
@@
 %Cooperative and non cooperative games
-In a decision game-theoretic approach an agent will try to act in such a way as to maximise his expected or average utility function.
+%-Remark- Stuk herschrijven-- Payoff introduceren -- zin van wikipedia
+%Binnen de speltheorie is een 'spel' een (model van) interactie tussen spelers die individueel beslissingen nemen.
+De spelers ontvangen een beloning (pay-off) die afhangt van hun eigen beslissingen maar ook van die van de anderen.
+De volgende typen kunnen worden onderscheiden.
+%
+%Coöperatieve spellen en niet-coöperatieve spellen[bewerken]
+%Het verschil tussen deze twee typen spellen is dat bij coöperatieve spellen (cooperative game theory) bindende afspraken tussen de spelers gemaakt kunnen worden terwijl dit niet mogelijk is bij de niet-coöperatieve spellen (non-cooperative game theory).
+
+In a decision game-theoretic approach an agent will try to act in such a way as to maximize his expected or average utility function.
 It becomes more complicated when two or more agents want to maximise their utility and when actions of an agent can affect other agents' utilities.
@@
 \subsubsection{Best response and Nash Equilibrium}
-One of the solution concepts in Game Theory for non-cooperative games that will be used in this paper is a Nash Equilibrium.
+One of the solution concepts in game theory for non-cooperative games that will be used in this paper is a Nash equilibrium.
 A Nash Equilibrium is a subset of outcomes that can be interesting to analyse a game.
 To define this concept we first introduce the concept of best response.
-The best response for a player is the action of a player that maximizes it's pay-off for any given action of the other player.
+The best response, given an action of the other player, is the action that maximizes its pay-off.
 We define $Opt_{i}$ as the best response function for player \textit{i}.
@@
 For a Nash Equilibrium each player has a list of actions and each player's action maximizes his or her pay-off given the actions of the other players.
-Nobody has the incentive to change his or her action if an equilibrium profile is played.
+Nobody has the incentive to independently or unilaterally change his or her action if an equilibrium profile is played.
 We have a Nash Equilibrium for the pair $(a_{1}^{*},a_{2}^{*})$ when $a_{1}^{*} = Opt_{1}(a_{2}^{*})$ and $a_{2}^{*} = Opt_{2}(a_{1}^{*})$.\\
+
+%-Remark- Matthias: Beter uitleggen: the best response, given an action of the other player, is the action that maximizes its pay-off.
+Anders lijk je de definitie  v.e.
+dominante strategie te geven.
+(Een actie die ongeacht de actie v.d.
+tegenstander de beste pay-off geeft) Maak onmiddelijk duidelijk dat het een functie van de acties van de tegenstander betreft.
+Zo is een dominante actie een actie a waarvoor geldt forall a2 : Opt(a2) = a.
+% -Remark- Matthias: Hier moet ‘independently’ of ‘one-sidedly’ of ‘unilateraly’ bij staan, anders beschrijf je een dominante strategie.
+Het mag voor beide beter zijn om gezamelijk te veranderen, maar niet om unilateraal te wijzigen!
+Kijk maar naar ’t prisoners dilemma: eenzijdig wisselen is slecht voor iedereen in het T/T Nash equilibrium, maar beide weizigen naar S/S zou voor beiden positief zijn!
 
@@
 Each of them can betray the other or they can support each other and remain silent.
-If a player confesses, he will either be imprisoned for 3 years or go free, depending on what the other does.
-If the player is silent, he will either be imprisoned for 5 years or one year, depending on what the other does.
+If a player confesses, he will either be imprisoned for 3 years or go free, depending on whether the other confesses.
+If the player is silent, he will either be imprisoned for 5 years or one year, depending on whether the other confesses.
 \\ \todo{nakijken met praten, bekennen of andere veraden?}
@@
 If there is no feedback for either player, he will play in the same manner against every opponent.
-The strategy is then called \textit{non-adaptive} because the playing strategy is not dependent on the opponents movements.
+The strategy is then called \textit{non-adaptive} because the playing strategy is not dependent on the opponent's movements.
 An interesting subclass of the non-adaptive strategies is the one where the time intervals between two consecutive moves are generated by a renewal process \todo{renewal beter uitleggen}.
-An example of such renewal strategy is the periodic strategy where the time between two consecutive moves of the players are a fixed interval.
+An example of such renewal strategy is the periodic strategy where the time between two consecutive moves of the players is a fixed interval.
 An exponential strategy is a renewal strategy in which the interval between two consecutive moves is exponentially distributed.
@@
 \item a defender facing an LM attacker has to play with a randomized strategy e.g. exponential strategy;
+%-Remark- Matthias: Hoe is ‘exponential strategy’ randomized? En uitgebreider spreken over de condities: een defender facing an LM attacker and comparable cost…
 \item any amount of feedback about the opponent received during the game, benefits to a player.
 \end{itemize}
+%-Remark- Matthias: Wat is ‘benefits the player’?
+Als de speler de feedback negeert benefit die info de speler niet.
+Ik zou zeggen ‘any amount of feedback about the opponent received during the game allows for better strategies/leads to a better strategy’ ofzo.
+Let op: Dit is een sterkere/straffere uitspraak, dus denk hier wel goed over na natuurlijk!
 

### 2015-08-09 doc/template/chap-5.tex
vanalles veranderd

@@
 %Het verschil tussen deze twee typen spellen is dat bij coöperatieve spellen (cooperative game theory) bindende afspraken tussen de spelers gemaakt kunnen worden terwijl dit niet mogelijk is bij de niet-coöperatieve spellen (non-cooperative game theory).
-
-In a decision game-theoretic approach an agent will try to act in such a way as to maximize his expected or average utility function.
-It becomes more complicated when two or more agents want to maximise their utility and when actions of an agent can affect other agents' utilities.
-In this kind of game, the basic modelling unit is the group of agents.
-It is referred to as \textit{non-cooperative game theory}.
+The choices of each agent produce an outcome, with respect to the utilities of the agent. A pay-off of an agent is the 
+Games can be devised into two types of games: cooperative or non-cooperative. In a non cooperative game, the basic modelling unit is the group of agents. 
+Two or more agents want to maximise their utility and their actions can be affected by other agents' utilities.
 In the individualistic approach the basic modelling is only one agent.
@@
 Each of them can betray the other or they can support each other and remain silent.
-If a player confesses, he will either be imprisoned for 3 years or go free, depending on whether the other confesses.
-If the player is silent, he will either be imprisoned for 5 years or one year, depending on whether the other confesses.
-\\ \todo{nakijken met praten, bekennen of andere veraden?}
+If a player talks, he will either be imprisoned for 3 years or go free, depending on whether the other talks.
+If the player is silent, he will either be imprisoned for 5 years or one year, depending on whether the other talks.
 
@@
 Figure \ref{Prison} shows the rewards of the possible actions of the two players.
-The figure shows that it is advantageous for each player to confess.
+The figure shows that it is advantageous for each player to talk (betraying the other).
 If prisoner 1 talks and the other prisoner remains silent, prisoner 1 will be free.
@@
 A dominant strategy is a strategy that is better than all other strategies regardless of what the opponent does.
-The Nash equilibrium of the game is that they both confess even though it would be better if both prisoners cooperated and choose to stay silent.
+The Nash equilibrium of the game is that they both talk even though it would be better if both prisoners cooperated and choose to stay silent.
 
@@
 It is a game between two perfectly rational prisoners who do not know what the other one will do.
-Each prisoner can confess or stay silent.
+Each prisoner can talk (betraying the other) or remain silent.
 The resulting sentence is shown below each prisoner.
@@
 The strategy is then called \textit{non-adaptive} because the playing strategy is not dependent on the opponent's movements.
-An interesting subclass of the non-adaptive strategies is the one where the time intervals between two consecutive moves are generated by a renewal process \todo{renewal beter uitleggen}.
+An interesting subclass of the non-adaptive strategies are the Renewal strategies where the time intervals between two consecutive moves are generated by a renewal process.
+The length between two consecutive moves are independent and identically distributed random variables chosen from a density formula \textit{f}.
+This renewal process means that the strategies are renewed after each move: the interval until the next move only depends on the current move time and not on previous history.
+\todo{check comment in latex}
+%-Remark- Vorige zin is gecopypaste van FlipIt
 An example of such renewal strategy is the periodic strategy where the time between two consecutive moves of the players is a fixed interval.
@@
 \item periodic strategy strongly dominates the other renewal strategies if the opponent has a periodic or non-arithmetic renewal strategy.
-This means that it is always advantageous for the opponent to play periodically against a player with a non-adaptive strategy;
-\item periodic games are disadvantageous against players following a Last Move adaptive strategy;
-\item if a defender facing an LM attacker and with a cost much lower than the attacker, he can play with a periodic rate that is fast enough he'll force the attacker to drop out;
-\item a defender facing an LM attacker has to play with a randomized strategy e.g. exponential strategy;
+This means that it is a good choice for the opponent to play periodically against a player with a non-adaptive strategy;
+\item periodic games are disadvantageous against players following a Last Move adaptive strategy.
+The opponent can observe the exact time of the player's next move and play immediately afterwards.
+If the costs are from the same magnitude, the opponent can keep the control over the resource with little interrupts from the other player;
+\item a player facing an LM opponent and with a cost much lower than the opponent has two options.
+The first option is to move with a periodic rate that is fast enough he'll force the opponent to drop out.
+The other option is to play with a randomized strategy, such that the opponent cannot learn any information regarding the next move of the player;
+\item the best defence strategy is to play fast, to make the opponent drop out of the game.
+To be able to move fast, the player has to make sure that the cost of moving is much less than the opponents moves.
 %-Remark- Matthias: Hoe is ‘exponential strategy’ randomized? En uitgebreider spreken over de condities: een defender facing an LM attacker and comparable cost…
-\item any amount of feedback about the opponent received during the game, benefits to a player.
+%\item any amount of feedback about the opponent received during the game, benefits to a player.
 \end{itemize}

### 2015-08-10 doc/template/chap-5.tex
verbetering naar nederlands

@@
 %- Remark: to the reader weg
-\section{What is security?}
+\section{What Is Security?}
 
@@
 %Het verschil tussen deze twee typen spellen is dat bij coöperatieve spellen (cooperative game theory) bindende afspraken tussen de spelers gemaakt kunnen worden terwijl dit niet mogelijk is bij de niet-coöperatieve spellen (non-cooperative game theory).
-The choices of each agent produce an outcome, with respect to the utilities of the agent. A pay-off of an agent is the 
+%The choices of each agent produce an outcome, with respect to the utilities of the agent. A pay-off of an agent is the 
+\todo{utility, pay-off, outcomes}
 Games can be devised into two types of games: cooperative or non-cooperative. In a non cooperative game, the basic modelling unit is the group of agents. 
@@
 Over John Nash is een prachtige film
-\subsubsection{Best response and Nash Equilibrium}
+\subsubsection{Best Response and Nash Equilibrium}
 One of the solution concepts in game theory for non-cooperative games that will be used in this paper is a Nash equilibrium.
@@
 %A game in game theory consists of multiple agents and every agent has a set of actions that he can play.
-\section{The FlipIt game}
+\section{The FlipIt Game}
 \label{cha1:FlipItGame}

### 2015-08-10 doc/template/nash.tex
Laatste blabla pagerank
Matthias opmerkingen
verbetering naar nederlands

@@
 In this chapter we are interested in finding the optimal strategies.
-To calculate the optimal strategies of both players we have to find the Nash Equilibria of the game.
+To calculate the optimal strategies of both players we have to find the Nash equilibria of the game.
 First the optimal functions are derived from the formulas in the previous chapter.
-From these piecewise functions we can derive the Nash Equilibria.
+From these piecewise functions we can derive the Nash equilibria.
 \\
@@
 Nash equilibria are points with the property that neither player benefits by deviating in isolation from the equilibrium.
-We can compute Nash Equilibria for the periodic game as an intersection point of curves $opt_{D}$ and $opt_{A}$.
+We can compute Nash equilibria for the periodic game as an intersection point of curves $opt_{D}$ and $opt_{A}$.
 \\
@@
 More formally, a Nash equilibrium for the periodic game is a point $(\delta_{D}^{*},\delta_{A}^{*})$ such that
-the defender's benefit $\beta_{D}(\alpha_{D},\delta_{A}^{*}) $is maximized at $\delta_{D}= \delta_{D}^{*}$ and the attacker's benefit
+the defender's benefit $\beta_{D}(\delta_{D},\delta_{A}^{*}) $is maximized at $\delta_{D}= \delta_{D}^{*}$ and the attacker's benefit
 $\beta_{A}(\delta_{D}^{*},\delta_{A}) $ is maximized at $\delta_{A}= \delta_{A}^{*}$.
@@
 \begin{equation*}
-\beta_{D}(\delta_{D},\delta_{A})= \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - \dfrac{k_{D}}{\delta_{D}} = \dfrac{\delta_{A} + 2 (d-k_{D})}{\delta_{D}}
+\beta_{D}(\delta_{D},\delta_{A})= \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - \dfrac{k_{D}}{\delta_{D}} = \dfrac{\delta_{A} + 2 (d-k_{D})}{2\delta_{D}}
 \end{equation*}
@@
 For $\delta_{A} + 2(d-k_{D}) > 0$ the benefit will be always positive but decreasing, see figure \ref{ShapeUp}. 
-The defender will always play if $\delta_{A} + 2(d-k_{D}) > 0$ for $k_{D} < d$ or $k_{D} > d$ because $\delta_{A}$ will be positive in either case.
+The defender will always play as fast as he can if $\delta_{A} + 2(d-k_{D}) > 0$ for $k_{D} < d$ or $k_{D} > d$ because $\delta_{A}$ will be positive in either case.
 There is an edge case if $d=k_{D}$, which results in a benefit of $\beta_{D}(\delta_{D},\delta_{A})= \dfrac{\delta_{A}}{\delta_{D}}$.
@@
 
-For $\delta_{A} + 2(d-k_{D}) < 0$, the benefit will always be negative so the defender will not play. See figure \ref{ShapeDown}. This is independent of the value of $k_{D}$ and $ d$. \\
+For $\delta_{A} + 2(d-k_{D}) < 0$, the benefit will always be increasing but negative so the defender will not play. See figure \ref{ShapeDown}.  \\
 
@@
 \begin{equation*}
-\dfrac{\beta_{D}(\alpha_{D},\alpha_{A})}{\partial \delta_{D}} = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - \dfrac{k_{D}}{\delta_{D}} - \dfrac{(d-(\delta_{D} - \delta_{A}))^{2}}{2\delta_{D}\delta_{A}}
+\beta_{D}(\delta_{D},\delta_{A}) = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - \dfrac{k_{D}}{\delta_{D}} - \dfrac{(d-(\delta_{D} - \delta_{A}))^{2}}{2\delta_{D}\delta_{A}}
 \end{equation*}
@@
 \begin{equation*}
-\frac{\partial \beta_{D}(\alpha_{D},\alpha_{A})}{\partial \delta_{D}} =  - \dfrac{1}{2\delta_{D}} + \dfrac{k_{D}}{\delta_{D}^{2}} + \dfrac{d^{2}}{2\delta_{D}^{2}\delta_{A}}
+\beta_{D}(\delta_{D},\delta_{A}) =  - \dfrac{1}{2\delta_{A}} + \dfrac{k_{D}}{\delta_{D}^{2}} + \dfrac{d^{2}}{2\delta_{D}^{2}\delta_{A}}
 \end{equation*}
@@
 \begin{equation*}
-\frac{\partial \beta_{A}(\alpha_{D},\alpha_{A})}{\partial \delta_{D}} =0 ~~~~~~ =>~~~~~~ 2k_{A} = \dfrac{(\delta_{D}-d)^{2}}{\delta_{D}}
+\frac{\partial \beta_{A}(\delta_{D},\delta_{A})}{\partial \delta_{D}} =0 ~~~~~~ =>~~~~~~ 2k_{A} = \dfrac{(\delta_{D}-d)^{2}}{\delta_{D}}
 \end{equation*}

### 2015-08-10 doc/template/nash.tex
nash stuff

@@
 \subsubsection{Best responses}
-The optimum functions will be piecewise functions.
-There will be different optimum functions depending on $k_{D}$ and $d$.
-We distinguish for these two cases, three sub cases for different values of $\delta_{A}$.
+The optimum functions will be piecewise functions. We distinguish three cases for different values of $\delta_{A}$ depending on the values of $k_{D}$ and $d$. 
+We point out that $\delta_{A}$ and $\delta_{D}$ are positive rates.
 
-\subsection*{$k_{D} \leq d$}
-Because $k_{D} \leq d$, the term $2(k_{D} - d)$ will always be negative. We point out that $\delta_{A}$ and $\delta_{D}$ are positive rates. 
-\begin{itemize}
-\item if $\delta_{A} < 2(k_{D} - d)$ \\
-This means that $\delta_{A}$ has to be negative which is not possible. For case the defender will not play.
-\item $\delta_{A} = 2(k_{D} - d)$ \\
-$\delta_{A}$ is negative or equal to 0 so the attacker will not play. For case 1 and case 2.b the defender will also not play.
-\item $\delta_{A} > 2(k_{D} - d)$ \\
-Case 2.a it is increasing for every value $\delta_{A} \in [0,\infty]$.  For case 1 together with case 2.b the optimal benefit is achieved at rate $\delta_{D} = \sqrt{d^{2} + 2\delta_{A}k_{D}}$.
-\end{itemize}
+\subsection*{$\delta_{A} < 2(k_{D} - d)$}
+From case 1 above, it follows that the benefit is increasing but it will always be non-positive.
+The defender will try not to play ($\delta_{D}= \infty$).
+From case 2.a and case 2.b it follows that from interval $[2(k_{D} -d), \sqrt{2k_{D}\delta_{A} + d^{2}}]$  (because for case 2 $\delta_{D} \geq \delta_{A}$), that the benefit is increasing.
+The maximum of case 2.a and the maximum of case 2.b can be brought together.
+It follows that the optimal benefit is achieved at the $\delta_{D}=minimum[\delta_{A} + d, \sqrt{2k_{D}\delta_{A} + d^{2}}]$, which is $\delta_{D}=\sqrt{2k_{D}\delta_{A} + d^{2}}$.
+But because of case 2.a, the defender's maximum benefit is not to play at all.
 
-\subsubsection{$k_{D} > d$}
-Because $k_{D} > d$, the term $2(k_{D} - d)$ will always be positive. We point out that $\delta_{A}$ and $\delta_{D}$ are positive rates.
-\begin{itemize}
-\item if $\delta_{A} < 2(k_{D} - d)$ \\
-From case 2.a it follows that the benefit of the defender increases.
-From case 1 and case 2.b together the optimal benefit of the defender is achieved at rate $\delta_{D} = \sqrt{d^{2} + 2\delta_{A}k_{D}}$.
-\item $\delta_{A} = 2(k_{D} - d)$ \\
-From case 2.a it follows that $\beta_{D}(\delta_{D},\delta_{A})=0$, for all $\delta_{A} \in [0,2(k_{D} - d)]$.
-From case 1 and case 2.b together the optimal benefit for the defender is achieved for all rates $\delta_{D} \in [0, \sqrt{d^{2} + 2\delta_{A}k_{D}}]$.
-\item $\delta_{A} > 2(k_{D} - d)$ \\
-For case 2.a the benefit is decreasing. From case 1 and case 2.b the best strategy for the defender is not playing at all. 
-\end{itemize}
-\todo{dat laatste nog eens nakijken}
+\subsubsection{$\delta_{A} = 2(k_{D} - d)$}
+Then for case 2.a, $\beta_{D}(\delta_{D}, \delta_{A})=0$ for  all $\delta_{D} \in [\sqrt{2k_{D}\delta_{A} + d^{2}}, \infty]$.
+For case 2.b and case 1, the benefit is increasing in the interval $[\delta_{A},\sqrt{2k_{D}\delta_{A} + d^{2}}]$.
+If we put both cases back together, we get a maximum at $\sqrt{2k_{D}\delta_{A} + d^{2}}$.
+So in this case the defender's maximum benefit is achieved for any $\delta_{D}$ in $[\sqrt{2k_{D}\delta_{A} + d^{2}}, \infty]$ with value 0.
 
-From this analyses we can compute $opt_{D}(\delta_{A})$ for two different cases as: \\
+\subsubsection{$\delta_{A} > 2(k_{D} - d)$ }
+From case 2.a above, it follows that the benefit is decreasing but positive.
+So the defender will try to play as fast as possible ($\delta_{D} \rightarrow 0$).
+For case 2.b it is increasing in the interval $[ \delta_{A} ,\sqrt{2k_{D}\delta_{A} + d^{2}}]$ and for case 1 it is increasing in the interval $[0,\delta_{A}]$.
+It follows that the defender's maximum benefit is achieved for $\delta_{D} = \sqrt{2k_{D}\delta_{A} + d^{2}} $  \\
 
-For case  $k_{D} \leq d$:
+~~\\
+From this analyses we can compute $opt_{D}(\delta_{A})$ : \\
+
  \begin{displaymath}
@@
      \begin{array}{lr}
-       0, & \delta_{A} < 2(k_{D} - d)\\
-       0, & \delta_{A} = 2(k_{D} - d) \\
-       \sqrt{d^{2} + 2\delta_{A}k_{D}}, & \delta_{A} > 2(k_{D} - d)
-     \end{array}
-   \right.
-\end{displaymath}
-
-For case $k_{D} > d$ :
- \begin{displaymath}
-  opt_{D}(\delta_{A}) = \left\{
-     \begin{array}{lr}
-       \sqrt{d^{2} + 2\delta_{A}k_{D}}, & \delta_{A} < 2(k_{D} - d)\\
-       \big[0,\sqrt{d^{2} + 2\delta_{A}k_{D}}\big], & \delta_{A} = 2(k_{D} - d) \\
-       0, & \delta_{A} > 2(k_{D} - d)
+          \infty , & \delta_{A} < 2(k_{D} - d)\\
+      \left[ \sqrt{2k_{D}\delta_{A} + d^{2}},\infty\right] , & \delta_{A} = 2(k_{D} - d) \\
+      \sqrt{2k_{D}\delta_{A} + d^{2}}, & \delta_{A} > 2(k_{D} - d)
      \end{array}
@@
 To start with we only consider the case where $d < \delta_{D}$, because if $d > \delta_{D}$ the benefit of the defender \todo{nakijken, def of att} is always 1. \\
-To determine the Nash equilibria we also need to determine $opt_{A}(\delta_{D})$ by computing the derivative of $\beta_{A}(\delta_{D},\delta_{A})$ for a fixed $\delta_{D}$.
-We consider again two cases, where case 2 is divided into two subcategories: \\
+To determine the Nash equilibria we also need to determine $opt_{A}(\delta_{D})$ by computing the derivative of $\beta_{A}(\delta_{D},\delta_{A})$ for a fixed $\delta_{D}$. We consider three cases: \\
 
@@
 
-The benefit formula obtained in the previous chapter for this case is as follows:
+The benefit formula obtained in the previous chapter \ref{Benfcase1:attacker} for this case is as follows:
 \begin{equation*}
-\beta_{A}(\delta_{D},\delta_{A}) =\dfrac{\delta_{D}}{2\delta_{A}} - \dfrac{k_{A}}{\delta_{A}} + \dfrac{d^{2}}{2\delta_{D}\delta_{A}^{2}} - \dfrac{d}{\delta_{A}}
+\beta_{A}(\delta_{D},\delta_{A}) =\dfrac{\delta_{D}}{2\delta_{A}} - \dfrac{k_{A}}{\delta_{A}} + \dfrac{d^{2}}{2\delta_{D}\delta_{A}} - \dfrac{d}{\delta_{A}}
 \end{equation*}
@@
 \begin{equation*}
-\dfrac{\partial \beta_{A}(\delta_{D},\delta_{A})}{\partial \delta_{A}} = -\dfrac{\delta_{D}}{2\delta_{A}^{2}} + \dfrac{k_{A}}{\delta_{A}^{2}} - \dfrac{d^{2}}{2\delta_{D}\delta_{A}^{2}} + \dfrac{d}{\delta_{A}^{2}}
+\dfrac{\partial \beta_{A}(\delta_{D},\delta_{A})}{\partial \delta_{A}} = -\dfrac{\delta_{D}}{2\delta_{A}^{2}} + \dfrac{k_{A}}{\delta_{A}^{2}} - \dfrac{d^{2}}{2\delta_{D}\delta_{A}^{2}} + \dfrac{d}{\delta_{A}^{2}} = \dfrac{-\delta_{D}^{2} - d^{2} + 2\delta_{D}d + 2\delta_{D}k_{A}}{2\delta_{A}^{2}\delta_{D}}
 \end{equation*}
@@
 \begin{equation*}
-\frac{\partial \beta_{A}(\delta_{D},\delta_{A})}{\partial \delta_{D}} =0 ~~~~~~ =>~~~~~~ 2k_{A} = \dfrac{(\delta_{D}-d)^{2}}{\delta_{D}}
+\frac{\partial \beta_{A}(\delta_{D},\delta_{A})}{\partial \delta_{D}} =0 ~~~~~~ =>~~~~~~  \delta_{D}= \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}
 \end{equation*}
 
-It follows that $\beta_{A}(\delta_{D},\cdot)$ is increasing if $2k_{A} < (\delta_{D} - d)^{2} / \delta_{D}$ and decreasing if $2k_{A} > (\delta_{D} - d)^{2} / \delta_{D}$. \\
+It follows that $\beta_{A}(\delta_{D},\cdot)$ is increasing if $\delta_{D}> \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}$ and decreasing if $\delta_{D} < \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}$  \\
 
 \subsection*{Case 2.A: $\delta_{D} \geq d+\delta_{A} \geq \delta_{A} $ }
-The benefit formula obtained previously for this case is as follows:
+The benefit formula obtained from the previous chapter \ref{benfcase2a:attacker} for this case is as follows:
 \begin{equation*}
@@
 \end{equation*}
-It follows that $\beta_{A}(\delta_{D},\cdot)$ is increasing on $[0,\sqrt{2k_{A}\delta_{D}}]$ and decreasing on $[\sqrt{2k_{A}\delta_{D}}, \infty]$ and thus has a maximum on $\delta_{A} = maximum \{\delta_{D}, \sqrt{2k_{A}\delta_{D}} \} $.
-The maximum between $\delta_{D}$ and $ \sqrt{2k_{A}\delta_{D}}$ is needed because $\delta_{A} $ cannot exceed $\delta_{D}$ in this case.
+It follows that $\beta_{A}(\delta_{D},\cdot)$ is increasing on $[0,\sqrt{2k_{A}\delta_{D}}]$ and decreasing on $[\sqrt{2k_{A}\delta_{D}}, \infty]$ and thus has a maximum on $\delta_{A} = minimum \{\delta_{D} -d, \sqrt{2k_{A}\delta_{D}} \} $.
+The minimum between $\delta_{D}-d$ and $ \sqrt{2k_{A}\delta_{D}}$ is needed because $\delta_{A} $ cannot exceed $\delta_{D}-d$ in this case.
 \\
@@
 
-The benefit formula obtained previously for this case is as follows: 
+The benefit formula obtained from the previous chapter \ref{benfcase2b:attacker} for this case is as follows: 
 \begin{equation*}
@@
 \begin{equation*}
-\frac{\partial \beta_{A}(\delta_{D},\delta_{A})}{\partial \delta_{D}} =0 ~~~~~~ =>~~~~~~ 2k_{A} = \dfrac{(\delta_{D}-d)^{2}}{\delta_{D}}
+\frac{\partial \beta_{A}(\delta_{D},\delta_{A})}{\partial \delta_{D}} =0 ~~~~~~ =>~~~~~~ \delta_{D}= \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}
 \end{equation*}
 
-it follows that $\beta_{A}(\delta_{D},\cdot)$ is increasing if $2k_{A} < (\delta_{D} - d)^{2} / \delta_{D}$ and decreasing if $2k_{A} > (\delta_{D} - d)^{2} / \delta_{D}$.
+It follows that $\beta_{A}(\delta_{D},\cdot)$ is increasing if $\delta_{D} > \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}$ and decreasing if $\delta_{D} < \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}$.
 This is the same result as in case 1.\\
@@
 \subsubsection{Best responses}
-The optimum functions will be piecewise functions. We distinguish three cases for different values of $\delta_{D}$ and $k_{A}$.
+The optimum functions will be piecewise functions. We distinguish three cases for different values of $\delta_{D}$ regarding $d$ and $k_{A}$.
 
-For this term $\dfrac{(\delta_{D}-d)^{2}}{\delta_{D}} $ , $d$ has to be bigger than  $\delta_{D}$ because the cost $k_{A}$ cannot be negative.
+For this term $\dfrac{(\delta_{D}-d)^{2}}{2k_{A}} $ , $d$ has to be bigger than  $\delta_{D}$ because the cost $\delta_{D}$ cannot be negative.
 This was an assumption that was already made, because the benefit of the defender will always be 1 if $d$ is bigger than  $\delta_{D}$.
-\begin{itemize}
-\item if $2k_{A} < \dfrac{(\delta_{D}-d)^{2}}{\delta_{D}} $ \\
-Then for case 1 and case 2.b the benefit of the defender is increasing. From case 2.a follows that the optimal benefit for the attacker is achieved at the rate $\delta_{A} = \delta_{D}$
-\item if $2k_{A} = \dfrac{(\delta_{D}-d)^{2}}{\delta_{D}} $ \\
-From case 1 and case 2.b it follows that $\beta_{D}(\delta_{D},\delta_{A})=0$, for all $\delta_{A} \in [0,\dfrac{(\delta_{D}-d)^{2}}{2k_{A}})]$.
-From case 2.a the optimal benefit for the defender is achieved for all rates $\delta_{A} \in [0, \delta_{D}]$.
-\item if $2k_{A} > \dfrac{(\delta_{D}-d)^{2}}{\delta_{D}} $ \\
-All decreasing.
-\end{itemize}
+
+\subsubsection*{$\delta_{D} < \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}$} 
+From case 1 and case 2.b above, it follows that the benefit is decreasing but positive, which means that the attacker will play as fast as he can.
+For case 2.a  $\delta_{D} \geq \delta_{A}$ is increasing.
+It follows that for this case the attacker's maximum benefit is achieved for $\delta_{A} = \sqrt{dk_{A}\delta_{D}$.
+. In this case the attacker's maximum benefit is reached at $\delta_{A} = \delta_{A} = \sqrt{dk_{A}\delta_{D}$  .
+\subsubsection*{$\delta_{D} = \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}$} 
+The benefit $\beta_{A}(\delta_{D},\delta_{A})=0$ for case 1 and case 2.b for all $\delta_{D}$ in $[0, \infty [$.
+For case 2.a it follows that the benefit is increasing.
+The attacker's maximum benefit is achieved at any $\delta_{A} $ in $
+\subsubsection*{$\delta_{D} > \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}$} 
+For case 1 and case 2.b the benefit is increasing but non-positive. This means that the attacker won't play ($\delta_{A} = \infty$).
 

### 2015-08-10 doc/template/nash.tex
kleine fix

@@
 For case 2.a  $\delta_{D} \geq \delta_{A}$ is increasing.
-It follows that for this case the attacker's maximum benefit is achieved for $\delta_{A} = \sqrt{dk_{A}\delta_{D}$.
-. In this case the attacker's maximum benefit is reached at $\delta_{A} = \delta_{A} = \sqrt{dk_{A}\delta_{D}$  .
+It follows that for this case the attacker's maximum benefit is achieved for $\delta_{A} = \sqrt{d k_{A}\delta_{D}}$.
+. In this case the attacker's maximum benefit is reached at $\delta_{A} = \delta_{A} = \sqrt{dk_{A}\delta_{D}}$  .
 \subsubsection*{$\delta_{D} = \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}$} 
@@
 For case 2.a it follows that the benefit is increasing.
-The attacker's maximum benefit is achieved at any $\delta_{A} $ in $
+The attacker's maximum benefit is achieved at any $\delta_{A} $ in
 \subsubsection*{$\delta_{D} > \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}$} 
-For case 1 and case 2.b the benefit is increasing but non-positive. This means that the attacker won't play ($\delta_{A} = \infty$).
+For case 1 and case 2.b the benefit is increasing but non-positive. This means that the attacker won't play ($delta_{A} = \infty$).
 

### 2015-08-10 doc/template/chap-5.tex
Figuren in orde en nog andere dingen

@@
 \centering
-\includegraphics[scale=0.7]{../../doc/template/Images/DefFlipItt.png}
+\includegraphics[scale=0.7]{../../doc/template/Images/FlipBasic}
 \caption{A representation of a FlipIt game where both players are playing periodically.
-Every move or flip is indicated by a blue or orange circle.
-The attacker is represented in orange and the defender is represented in blue.
-The blue and orange rectangles represent the amount of time the respective player is in control of the resource.}
+Every move or flip is indicated by a green (dark grey) or orange (light grey) circle.
+The attacker is represented in orange and the defender is represented in green.
+The blue and orange rectangles represent which player is in control of the resource.}
 \label{fig:FLipItDefault}

### 2015-08-10 doc/template/nash.tex
Figuren in orde en nog andere dingen

@@
 Because $\delta_{D} \geq \delta_{A}$ there is a maximum on $\delta_{D} = maximum \{ \delta_{A}, \sqrt{2\delta_{A}k_{D} + d^{2}} \} $ or because $d+\delta_{A} \geq \delta_{D}$ there is also on $\delta_{D} = minimum \{ \delta_{A}+d, \sqrt{2\delta_{A}k_{D} + d^{2}} \} $.
-\todo{beter uitschrijven, deltaD moet tussen die twee waarden liggen}\\
+%-Remark- \todo{beter uitschrijven, deltaD moet tussen die twee waarden liggen}\\
 
@@
 
-It follows that $\beta_{A}(\delta_{D},\cdot)$ is increasing if $\delta_{D}> \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}$ and decreasing if $\delta_{D} < \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}$  \\
+It follows that $\beta_{A}(\delta_{D},\cdot)$ is increasing and non-positive if $\delta_{D}> \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}$ and decreasing and positive if $\delta_{D} < \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}$  \\
 
@@
 
-It follows that $\beta_{A}(\delta_{D},\cdot)$ is increasing if $\delta_{D} > \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}$ and decreasing if $\delta_{D} < \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}$.
+It follows that $\beta_{A}(\delta_{D},\cdot)$ is increasing and non-positif if $\delta_{D} > \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}$ and decreasing and positive if $\delta_{D} < \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}$.
 This is the same result as in case 1.\\
@@
 \subsubsection*{$\delta_{D} > \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}$} 
-For case 1 and case 2.b the benefit is increasing but non-positive. This means that the attacker won't play ($delta_{A} = \infty$).
+For case 1 and case 2.b the benefit is increasing but non-positive. This means that the attacker won't play ($\delta_{A} = \infty$). 
+\\
+~~\\
+From this analyses we can compute $opt_{A}(\delta_{D})$ :
 
-This brings us to the following opt function for : \todo{juiste resultaten hier nog invullen}
-From this analyses we can compute $opt_{D}(\delta_{A})$ for two different cases as:
  \begin{displaymath}
@@
      \begin{array}{lr}
-       \delta_{D}, & \delta_{A} < 2(k_{D} - d)\\
-       \left[ 0, \delta_{D} \right],  & \delta_{A} = 2(k_{D} - d) \\
-       0 & \delta_{A} > 2(k_{D} - d)
+       \sqrt{d k_{A}\delta_{D}}, & \delta_{D} > \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}\\
+       \left[ \sqrt{d k_{A}\delta_{D}}, \infty\right],  & \delta_{D} = \dfrac{(\delta_{D}-d)^{2}}{2k_{A}} \\
+       \infty & \delta_{D} < \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}
      \end{array}

### 2015-08-11 doc/template/chap-4.tex
Figuren in orde en nog andere dingen
delay models

@@
 Example of such worms are \textit{Code Red} \citep{OwnInternetSI} and \textit{Slammer} \citep{moore2003inside}.
-\todo{Voorbeeeld als mensen het idee niet snappen?}2111
+\todo{Voorbeeeld als mensen het idee niet snappen?}
 
@@
 
+ Modelling the propagation methods described above can help us to find the time needed for a worm to infect the whole network.
+This will be equal to the \textit{d} parameter in the FlipIt model with propagation delay.
+\\
+
 \section{Models for worm propagation}
 \label{modelsforpropagation}
-To use the models for worm propagation in the model of the FlipIt game with propagation delay it is sufficient to know the propagation speed of the worm and determine after how many time units a sufficient number of nodes in the network is infected.
-This will be equal to the \textit{d} parameter in our FlipIt model.
-\\
-
+Early work on worm propagation models are based on the spread of real-world epidemic diseases. These models are based on the transition state of each node in the network.\\
 %Scan based or Topology based
@@
 In the SIS model a node can become susceptible again after it has been infected.
-Based on these three models various other models have been proposed.
-\todo{referenties naar andere modellen} We are only interested in models that will allow us to find the time required for a worm to infect a sufficient amount of nodes.
+Currently, various other models have been proposed based on these three models.
+\todo{referenties naar andere modellen} We are only interested in models that will allow us to find the delay.
+\\
+
 Table \ref{tree} gives an overview of the most common propagation methods with the given type.
@@
 The \textit{Code Red} worm, which is a random scan based worm, has been analysed by this kind of model in \cite{OwnInternetSI}.
+Another drawback is that the graph is presented as an undirected graph.
+This means that the spread can always happen in both ways, which may not be suitable for every worm propagation.
 
@@
 
-Toextract form this formula the delay for the case of FlipIt with a propagation delay, we need to know the maximum numbers of hosts that can be infected before the total system is compromised.
+To extract the delay from this formula, we need to know the maximum numbers of hosts that can be infected before the total system is compromised.
 The initial compromise rate has to be approximated.
-\textit{t} is the variable in our formula and says how many time has passed.
-If \textit{a} passes the proportion of nodes that have to be compromised before the whole system is compromised by the attacker, value \textit{t}  is the value of \textit{d} in our FlipIt game.
+\textit{t} is a variable in the above formula that says how many time has passed.
+If \textit{a} passes the proportion of nodes that have to be compromised before the whole system is compromised by the attacker, value \textit{t}  is the value of \textit{d} in the FlipIt game with propagation delay.
+A drawback for this model is that the network topology is homogeneous and that the graphical representation of the graph is an undirected graph.
 
@@
 %%spread--
-\subsection*{self disciplinary worms}
-Model for self disciplinary worms and counter measures ... []  \todo{verder uitwerken ? ja of nee? }
+%\subsection*{self disciplinary worms}
+%Model for self disciplinary worms and counter measures ... []  \todo{verder uitwerken ? ja of nee? }
 
@@
 The Bluetooth worm model is introduced by Yan and Eidenbenz \citep{yan2009modeling}. The model captures the behaviour of the propagation of a worm that spreads through the Bluetooth protocol. 
-A Bluetooth device that is compromised can only infect neighbour devices that are in its radio range
-Let \textit{i(t)} be the average density of infected hosts in the network at time \textit{t}.
-
-Formula is as follows: 
+A Bluetooth device that is compromised can only infect neighbour devices that are in its radio range.
+Let \textit{i(t)} be the average density of infected hosts in the network at time \textit{t}: 
 \begin{equation}
@@
 
-The formula for a better estimation of the worm propagation
+The formula for a better estimation of the worm propagation:
 \begin{equation}
@@
 
-To apply this model on FlipIt we again set up a threshold value for i(t) and see how long it took to compromise this amount of hosts.
-The paper concluded in their work  that after setting model parameters accordingly ($\lambda_{ne} = 0.2108$ the average node degree and $J_{in}= 0.2372$ the average meeting rate of neighbours), the model predicts that the time it would take to infect 99$\%$ of the devices is slightly less than one hour.
+To apply this model on FlipIt we again set up a threshold value for i(t) and see how long it takes to compromise this amount of hosts.
+The paper by Yan and Eidenbenz concluded in their work  that after setting model parameters accordingly ($\lambda_{ne} = 0.2108$ the average node degree and $J_{in}= 0.2372$ the average meeting rate of neighbours), the model predicts that the time it would take to infect 99$\%$ of the devices is slightly less than one hour.
 So in this case the delay is equal to an hour and the amount of hosts that has to be infected by the attacker before the network is compromised is 99$\%$ of the hosts in the network.
@@
 \subsection{AAWP model}
-SIR AAWP stands for Analytic Active Worm Propagation.
-The AAWP model is a SIR model which uses a death rate in the formula to calculate the amount of hosts that become Removed.
-Yet this model is still analysed because if the death rate is removed, this model becomes a SI mode.
+AAWP stands for Analytic Active Worm Propagation.
+The model is \cite{chen2003modeling} The AAWP model is a SIR model which uses a death rate in the formula to calculate the amount of hosts that become Removed.
+Yet this model is still analysed because if the death rate is removed, this model becomes an SI mode.
 The model is different from other SI models because it includes the time that it takes to infect a host.
@@
 the scanning rate of the worm and $\Omega$ is the scanning
-space.
+space. \\
+
+The biggest drawback of this model is that it is a discrete time model. Continuous models are more appropriate for large scale models.
 

### 2015-08-11 doc/template/chap-4.tex
models propagtion

@@
 AAWP stands for Analytic Active Worm Propagation.
-The model is \cite{chen2003modeling} The AAWP model is a SIR model which uses a death rate in the formula to calculate the amount of hosts that become Removed.
-Yet this model is still analysed because if the death rate is removed, this model becomes an SI mode.
-The model is different from other SI models because it includes the time that it takes to infect a host.
+The model is proposed in \cite{chen2003modeling} by Chen et al.
+to model the discrete behaviour of a worm.
+The AAWP model is a SIR model which uses a death and patch rate to calculate the amount of hosts that become Removed.
+Yet this model can still be useful to calculate the delay because if the death and patch rate are removed, this model becomes an SI model.
+It is different from other SI models because it includes the time that it takes to infect a host.
 A hosts cannot infect another hosts before it is completely compromised.
 The model also considers the fact that a hosts can be scanned and hit by multiple worms at the same time.
-%This model is used for worms that propagate through uniform scanning.
+\\%This model is used for worms that propagate through uniform scanning.
 When uniform scanning is used, each IP address
 
-The spread of the AAWP model is characterized as follows:
+The spread of the AAWP model with death and patch rate is characterized as follows:
+\begin{equation}
+I_{t+1}=(1-d-p)I_{t}+[(1-p)^{t})N-I_{t}][1-(1-\dfrac{1}{\Omega})^{sI_{t}}]
+\end{equation}
+where \textit{d} is the death rate, $p$ is the patch rate, $I_{t}$ is the amount of infected hosts at time \textit{t}, \textit{N} is the number of vulnerable hosts, \textit{s} is
+the scanning rate of the worm and $\Omega$ is the scanning
+space. \\
+
+The spread of the worm without the death and the patch rate is as follows:
 \begin{equation}
@@
 \end{equation}
-where $I_{t}$ is the amount of infected hosts at time \textit{t}, \textit{N} is the number of vulnerable hosts, \textit{s} is
-the scanning rate of the worm and $\Omega$ is the scanning
-space. \\
 
-The biggest drawback of this model is that it is a discrete time model. Continuous models are more appropriate for large scale models.
+The iteration procedure with death and patch rate will stop when all the nodes in the network are infected or when the number of infected nodes remains the same.
+When the death and patch rate are removed from the formula the iteration procedure will stop when all the nodes are infected.
+To extract the delay from this formula we need to know when the iteration procedure stops.
+\\
+The biggest drawback of this model is that it is a discrete time model. Continuous models are more appropriate for large scale models. The model also uses a complete undirected graph.
 
@@
 \label{eigenmatrixmethode}
-As can be seen in table \ref{tree} most propagation models are for a specific network topology.
-The different kinds of network topology of each system is either homogeneous, small-world network, random network or power-law network.
+For some of the propagation methods, the graph of the network matters.
+It is important to have the right topology for the right method.
+Email worms need a topology that represent a social network, BGP routing worms need a topology on network level.
+\\
+
+As can be seen in table \ref{tree} all propagation models are depending on a specific network topology.
+The different kinds of network topology of each system is either homogeneous, small-world network, random network or a power-law network.
 A homogeneous network is a network where every node has about the same degree of connectivity.
@@
 In a power law network the degrees of each node in the network follow the power law.
-That is that some of the nodes have a small degree and other have a very large degree of connection.
-What we want to have is to have a method to calculate the delay that is independent of the network topology.
-A method where we can chose in the beginning which topology we have and then calculate the delay.\\
+Some of the nodes have a small degree of connectivity, others have a very large degree of connectivity.\\
+
+What we want to have is a method to calculate the delay that is independent of a network topology. A method where we can chose in the beginning which topology we have and then calculate the delay.\\
 
 %All of the above models for modelling the mechanism of the spreading of a worm depend on the kind of topology of the network.  
-In this chapter we propose a method to calculate the spread of the worm in way that the topology of the network can be easily integrated.
+In this chapter we propose a method to calculate the spread of a worm in a way that the topology of the network can be easily integrated.
 The method will give an approximation of how fast a worm can infect a network.
-The method is based on the method for Google Page Ranking \cite{GoogleRank}.
+The method is based on Google Page Ranking \cite{GoogleRank} used by search engines to rank pages.
 
@@
 
-For some of the spreading methods, the graph of the network matters.
-It is important to have the right topology for the right method.
-Email worms need a topology that represent a social network, BGP routing worms need a topology on network level.
-\\
 %--iets zeggen dat het ook belangrijk is voor de defender om zijn netwerk zo aan te passen dat het virus moeilijker kan verspreiden. Network segregation mss aanhalen ?--

### 2015-08-11 doc/template/chap-4.tex
Google Matrix

@@
 In the SIS model a node can become susceptible again after it has been infected.
-Currently, various other models have been proposed based on these three models.
-\todo{referenties naar andere modellen} We are only interested in models that will allow us to find the delay.
+Currently, various other models have been proposed based on these three models (\cite{wang2014modeling}, \cite{OnWorms2005survey}, \cite{xiang2009propagation} and \cite{serazzi2004computer}).
+We are only interested in models that will allow us to find the delay.
 \\
@@
 All models that are based on SI model type (Susceptible-Infected) are good, because in our model every node that is infected will stay infected.
-Only the Local Preference and Email worms models will not be explained because they involve human interaction.
+The other propagation models, SIR and SIS, are only interesting if there is a possibility to reduce the models to an SI model.
+The following section will apply the extraction of the delay to a few models.
+These models are not meant as an exhaustive list, but rather as an illustration to the possible ways of extracting the delay out of an propagation model.
+%Only the Local Preference and Email worms models will not be explained because they involve human interaction.
 OSN model is done by means of simulation.
@@
 We discuss a couple of SI models and then introduce a method of out own.\\
-
-For more information about other models the reader is referred to the following papers with other most common propagation models: \cite{wang2014modeling}, \cite{OnWorms2005survey}, \cite{xiang2009propagation} and \cite{serazzi2004computer}.
 
@@
 The method will give an approximation of how fast a worm can infect a network.
-The method is based on Google Page Ranking \cite{GoogleRank} used by search engines to rank pages.
-
-\subsubsection{Google PageRank algorithm}
-The PageRank algorithm is introduced by Page and Brin in 1998 as one of the main features of the search engine Google to improve the search results.
-PageRank models the human behaviour when users surf through the net.
-It can also model \textit{random surfers}.
-A random surfer is the probability that a surfer gets bored and randomly visits another page which is not linked with the initial page.
-The probability that he will visit a random page is the PageRank of that page.
-A page will have a high PageRank if many pages will point to this page.
-This means that this page is well cited through other pages and maybe an important one to look at.
-The main idea of the PageRank algorithm is to look at the number of (important) web pages that point to a particular page.
-The ranking of this page depends than on the number of out coming links and the importance of the pages that link to this page.
-With a probability of $P$ the surfer will follow a link of the page to another page, and with a probability of $1-P$ the surfer will surf to a random page.
-The PageRank algorithm is expressed as follows:
-\begin{equation}
-PageRank(A)=P \sum \dfrac{PageRank(i)}{d_{out,i}} + (1-P) \cdot e_{A}
-\end{equation}
-where $N_{A}$ is the number of pages, $PR(i)$ is the ranking of web page $i$, $D_{out,i}$ is the number of outgoing links of page $i$, $(1-P)$ the probability that a random page is searched, and $e_{A}$ the restart value for web page A which is often uniformly distributed among all web pages.
+This method is capable of calculating the delay for any network topology, but it is not the most efficient way of doing this.
+For this paper it is just a proof of concept.
 
@@
 Using this knowledge we can calculate in how many steps a node is infected: \\
-The infection vector or start vector $I_{0}$ of lenght $|V|$ indicates which node is infected and which not.
+The infection vector or start vector $I_{0}$ of length $|V|$ indicates which node is infected and which not.
 Multiplying the infection vector with $A$ results in a vector $I_{1}$, which represents which nodes are infected after 1 step.
@@
 %\end{verbatim}
+\subsubsection{Google PageRank algorithm}
 
-%\end{document}
+To calculate the delay of a worm propagation we assumed that every node is even important.
+A possible relaxation of this rule can be done by the Page Rank Algorithm, which can be used to determine the importance of certain nodes.
+
+The PageRank algorithm is introduced by Page and Brin in 1998 as one of the main features of the search engine Google to improve the search results.
+PageRank models the human behaviour when users surf through the net.
+It can also model \textit{random surfers}.
+A random surfer is the probability that a surfer gets bored and randomly visits another page which is not linked with the initial page.
+The probability that he will visit a random page is the PageRank of that page.
+A page will have a high PageRank if many pages will point to this page.
+This means that this page is well cited through other pages and maybe an important one to look at.
+The main idea of the PageRank algorithm is to look at the number of (important) web pages that point to a particular page.
+The ranking of this page depends than on the number of out coming links and the importance of the pages that link to this page.
+With a probability of $P$ the surfer will follow a link of the page to another page, and with a probability of $1-P$ the surfer will surf to a random page.
+The PageRank algorithm is expressed as follows:
+\begin{equation}
+PageRank(A)=P \sum \dfrac{PageRank(i)}{d_{out,i}} + (1-P) \cdot e_{A}
+\end{equation}
+where $N_{A}$ is the number of pages, $PR(i)$ is the ranking of web page $i$, $D_{out,i}$ is the number of outgoing links of page $i$, $(1-P)$ the probability that a random page is searched, and $e_{A}$ the restart value for web page A which is often uniformly distributed among all web pages.
+\\
+
+The Google matrix is a stochastic matrix that is used to calculate the PageRank of all the web pages on the internet.
+The dominant eigenvalue of the matrix is the PageRank.
+The matrix is a graph where the edges denote the links between each page.
+Using the power method, the PageRank of each page can be calculated iteratively.
+\\
+
+The Google Matrix of a graph is defined as follow:
+\begin{equation}
+M = (1-p) \cdot A + p \cdot B ~~where~~B=\dfrac{1}{n}\cdot \begin{bmatrix}
+       1 & 1 & \ldots &1           \\[0.3em]
+       \vdots & \vdots          & \ddots & \vdots \\[0.3em]
+       1 & 1& \cdots &1
+     \end{bmatrix}
+\end{equation}
+
+where $p$ defines the damping factor and indicates the probability that a surfer quits a current page and continuous on a new random page.
+Matrix A is an transposed matrix of the graph matrix.
+The graph matrix is an $n x n$ matrix (\textit{n} defines the number of pages) that represents the graph where every entry corresponds to a non-zero entry if there is a link from one page to the other page.
+For dangling nodes (nodes with no outgoing links), the row entries are equal to $\dfrac{1}{n}$.
+Matrix A for the graph represented figure \ref{GoogleMatrix} is equal to:
+
+\begin{figure}
+\centering
+\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2cm,
+                    semithick]
+  \tikzstyle{every state}=[fill=black!60!green,draw=none,text=white]
+
+  \node[state] (A)                    {$N_1$};
+  \node[state]         (B) [above right of=A] {$N_2$};
+  \node[state]         (D) [below right of=A] {$N_3$};
+
+  \path (B) edge              node {} (A)
+        (D) edge 			  node {} (A);
+\end{tikzpicture}
+\caption{A representation of a graph with three nodes. Node 1 is a dangling node with no outgoing links.}
+\label{GoogleMatrix}
+\end{figure}
+
+$ A =
+\bordermatrix{
+         & N_1		& N_2	& N_3	 \cr
+    N_1   & 1/3		& 1		& 1			     \cr
+    N_2   & 1/3		& 0		& 0		     \cr
+    N_3   & 1/3		& 0		& 0		 \cr
+  }$
+
+~~\\
+
+The PageRank vector of a graph, with the transposition matrix A and the damping variable p, is equal to the probabilistic eigenvector of the matrix M, corresponding to the eigenvalue 1.
+This can be calculated by the use of the Power Method.
+\\
+
+We can use the Google matrix to calculate the probability that a node in a graph is infected by a worm after a certain amount op time.
+When a worm uses the topology of the graph to propagate the damping variable p must be equal to zero, as in the worm will not suddenly attack a node that is not connected.
+When the propagation of a worm is topology independent, the damping variable p can be set on 1.
+\\
+This can be useful for further research, when the defender can also flip a subset of nodes.
+It can be advantageous to flip the nodes with a higher PageRank more often than the other nodes.
+A higer PageRank means that the page has a bigger probability of being compromised by a worm.

### 2015-08-11 doc/template/chap-5.tex
Google Matrix

@@
 These preferences are mapped to natural numbers and are called the utility functions.
-The numbers are interpreted as a mathematical measure that tells how much an agent likes or dislikes the states of the world.
+The numbers are interpreted as a mathematical measure that tells how much an agent likes or dislikes the outcome of the game.
+Outcomes are the result of a specific combination of a player's strategy.
+Each combination of a player's strategy is an outcome of the game.
+\\% The utility function can be ordinal or cardinal.
+In case of a ordinal utility function only relative rankings are important.
+The utility function can also be cardinal which is important for games with mixed strategies.
 \\
+
 %Cooperative and non cooperative games
@@
 \todo{utility, pay-off, outcomes}
-Games can be devised into two types of games: cooperative or non-cooperative. In a non cooperative game, the basic modelling unit is the group of agents. 
+Games can be divided into two types of games: cooperative or non-cooperative. In a non cooperative game, the basic modelling unit is the group of agents. 
 Two or more agents want to maximise their utility and their actions can be affected by other agents' utilities.
@@
 
-The players receive a benefit equal to the time units they were in possession of the resource minus the cost of making their moves. The cost for player \textit{i} is denoted by $k_{i}$. 
-The total gain for player \textit{i} is equal to the total amount of time that player \textit{i} has owned the resource from the beginning of the game up to time \textit{t}.
-It is expressed as follows:
+The total gain for player \textit{i} is equal to the total amount of time that player \textit{i} has owned the resource from the beginning of the game up to time \textit{t}:
 \begin{equation}\label{first}
@@
 \end{equation}
-$\alpha_{i}$ defines the average move rate by player \textit{i} up to time \textit{t} with $n_{i}(t)$ the amount of moves made by player \textit{i} up to time \textit{t}:
+The players receive a benefit equal to the time units they were in possession of the resource minus the cost of making their moves. 
+\begin{equation}\label{first}
+\beta_{i}(t) = \gamma_{i}(t) - k_{i}\alpha_{i}.
+\end{equation}
+where  $\beta_{i}(t)$ denote player's \textit{i} average benefit rate, $k_{i}$ denotes the cost for player \textit{i} and $\alpha_{i}$ defines the average move rate by player \textit{i} up to time \textit{t} with $n_{i}(t)$ the amount of moves made by player \textit{i} up to time \textit{t}:
 \begin{equation}
@@
 \end{equation}
-Let $\beta_{i}(t)$ denote player's \textit{i} average benefit rate up to and including time \textit{t}:
-\begin{equation}\label{first}
-\beta_{i}(t) = \gamma_{i}(t) - k_{i}\alpha_{i}.
-\end{equation}
-This is equal to the fraction of time the resource has been owned by player \textit{i}, minus the cost of making the moves.
-In a given game, the asymptotic benefit rate (or simply benefit) will be defined as the lim inf of the average benefit because time\textit{ t} will increase to infinity and the average benefit may not have limiting values.
+In a given game, the asymptotic benefit rate (or simply benefit) will be defined as the \textit{lim inf} of the average benefit because time\textit{ t} will increase to infinity and the average benefit may not have limiting values.
 \begin{equation}

### 2015-08-11 doc/template/chap-9.tex
chapter nash
Google Matrix

@@
 Every move or flip is indicated by a green or orange circle, respectively dark gray and light grey.
-The defender is represented in blue (dark grey) and plays with a period of $\delta_{D}$.
+The defender is represented in green (dark grey) and plays with a period of $\delta_{D}$.
 The flip of the attacker is represented by a white circle, but because there is a delay d, the attacker only controls the resource after time d represented by an orange circle (light grey).
@@
 To find the benefit, the cost of moving is subtracted from the average gain. 
-\begin{equation*}
+\begin{equation}
 \beta_{A}(\delta_{D},\delta_{A}) = \dfrac { (\delta_{D}-d) ^{2}} {2 \delta_{D}  \delta_{A}} - \dfrac{k_{A}}{\delta_{A}}
-\end{equation*}
+\label{Benfcase1:attacker}
+\end{equation}
 
  The benefit of the defender is then as follows:
-\begin{equation*}
+\begin{equation}
 \beta_{D}(\delta_{D},\delta_{A}) = 1 - \dfrac { (\delta_{D}-d) ^{2}} {2 \cdot \delta_{D}  \delta_{A}} - \dfrac{k_{D}}{ \delta_{D}}
-\end{equation*}
+\label{Benfcase1:defender}
+\end{equation}
 ~~\\
@@
 
-\subsubsection*{\textbf{Case a:} $d + \delta_{A} \leq \delta_{D}$}
+\subsubsection*{\textbf{Case 2.a:} $ \delta_{D} \geq d + \delta_{A} \geq \delta_{A}$}
 %The first case is where the defender's period is larger than the sum of the delay and the attacker's period. 
@@
 To determine the total gain rate of the defender we need to know the probability that the defender will move during an interval and what the average time is that the defender controls the resource.
-Given that if the defender moves he will always benefit entirely from a period of delay in the next interval of the attacker, his total gain is $\dfrac{(d+\dfrac{\delta_{A}}{2})}{\delta_{A}}$ \todo{formule valt uit de lucht} in one interval.
+Given that if the defender moves he will always benefit entirely from a period of delay in the next interval of the attacker, his total gain is $\dfrac{(d+\dfrac{\delta_{A}}{2})}{\delta_{A}}$ in one interval (as previously calculated.
 The total gain  rate of the defender is then the probability that the defender will move during an interval of the attacker multiplied by the total average gain per interval:
@@
 This yields the following benefit formula:
-\begin{equation*}\label{first}
+\begin{equation}\label{benfcase2a:defender}
 \beta_{D}(\delta_{D},\delta_{A}) = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - \dfrac{k_{D}}{ \delta_{D}}
-\end{equation*}\\
+\end{equation}\\
 
 The benefit for the attacker will be as follows:
-\begin{equation*}\label{first}
+\begin{equation}\label{benfcase2a:attacker}
 \beta_{A}(\delta_{D},\delta_{A}) = 1 -\dfrac{\delta_{A}}{2\delta_{D}} - \dfrac{d}{\delta_{D}} - \dfrac{k_{A}}{ \delta_{A}}
-\end{equation*}\\
+\end{equation}\\
 
@@
 ~~ \\
-\subsubsection*{\textbf{Case b:} $d + \delta_{A} \geq \delta_{D}$}
+\subsubsection*{\textbf{Case 2.b:} $d + \delta_{A} \geq \delta_{D} \geq \delta_{A}$}
 ~~~\\
@@
 This yields in the following benefit formula:
-\begin{equation*}
+\begin{equation}\label{benfcase2b:defender}
 \beta_{D}(\delta_{D},\delta_{A}) = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - \dfrac{k_{D}}{ \delta_{D}} - \dfrac{(d - (\delta_{D} - \delta_{A}))^{2}}{2 \delta_{D} \delta_{A}}
-\end{equation*}\\
+\end{equation}\\
 
 Consequently, the benefit for the attacker will be:
-\begin{equation*}
+\begin{equation}\label{benfcase2b:attacker}
 \beta_{A}(\delta_{D},\delta_{A}) = 1 -\dfrac{\delta_{A}}{2\delta_{D}} - \dfrac{d}{\delta_{D}} - \dfrac{k_{A}}{ \delta_{A}} + \dfrac{(d - (\delta_{D} - \delta_{A}))^{2}}{2 \delta_{D} \delta_{A}}
-\end{equation*}\\
+\end{equation}\\
 

### 2015-08-12 doc/template/chap-4.tex
laatste van gisteren

@@
 The overview of this chapter will be as follows: Section \ref{methodsofpropagation} presents an overview of the most frequently used propagation methods by worms.
-These propagation methods will be covered by different kinds of models in section \ref{modelsforpropagation} presents models to determine the time a worm needs to infect a sufficient number of nodes.
-These results can be used for the completion of parameter \textit{d}.
-\todo{wat wordt er met de vorige zin bedoeld?} Finally, section \ref{eigenmatrixmethode}, introduces an easy method to calculate the delay of the propagation of a worm.
+These propagation methods will be covered by different kinds of models in section \ref{modelsforpropagation} illustrates a couple of models as examples to determine the propagation of a worm.
+These models can be used to extract parameter \textit{d}.
+Section \ref{eigenmatrixmethode} introduces an easy method to calculate the delay of the propagation of a worm.
+In the last section \ref{GoogleMatrix}, a method based on the PageRank algorithm of Google is used to determine which node has a higher probability of being infected by a worm.
 % Different kind of graphs to model a computer network are introduced.

### 2015-08-12 doc/template/chap-4.tex
Model propagation verder afgewerkt

@@
 
-In the context of malware propagation, there are two kinds of APT's.
-First, there are APT's that launch an attack on just one target node of a network.
-The mechanism these APT's use to propagate malware is the dropper mechanism.
+In the context of malware propagation, there are two kinds of APTs.
+First, there are APTs that launch an attack on just one target node of a network.
+The mechanism these APTs use to propagate malware is the dropper mechanism.
 The dropper is the initial attack vector that compromises the single targeted node of the system.
-The second kind of APT's target multiple nodes.
-These APT's also use a dropper mechanism but have an additional mechanism for self-propagation in order to propagate themselves to the multiple targeted nodes of the system.
+The second kind of APTs target multiple nodes.
+These APTs also use a dropper mechanism but have an additional mechanism for self-propagation in order to propagate themselves to the multiple targeted nodes of the system.
 If the APT uses virus spreading, the virus infects one node on the network and has to wait for human interaction to spread.
@@
 If an APT uses a worm propagation method it will spread by itself after it has been dropped on the network.
-Given the additional complexity of incorporating the human factor in a spreading method and that this chapter is merely meant as illustration on calculating the delay for use in the game theoretic approach, this chapter will be limited to worm propagation models for APT's.
+Given the additional complexity of incorporating the human factor in a spreading method and that this chapter is merely meant as illustration on calculating the delay for use in the game theoretic approach, this chapter will be limited to worm propagation models for APTs.
 \\
@@
 \\
-The biggest drawback of this model is that it is a discrete time model. Continuous models are more appropriate for large scale models. The model also uses a complete undirected graph.
+The biggest drawback is that the model is a discrete time model. Continuous models are more appropriate for large scale models. The model also uses a complete undirected graph.
 
@@
 
-The computer network can be modelled by an undirected Graph $G = < V, E> $ where $|V|$ denotes the number of resources in the network and $|E|$ the number of connections.
-We can convert this to an adjacency matrix which represents which vertices of the graph are neighbours of other vertices.
+A computer network can be modelled by an undirected or directed graph $G = < N, E> $ where $|N|$ denotes the number of nodes in the network and $|E|$ the number of connections.
+This graph can be converted to an  $|N| \times |N|$ adjacency matrix where the entries represent the connections between the nodes of the network.
 \\
-The graph is represented as a $|V| \times |V|$ matrix with for every entry $a_{ij}$ a 1 as value if there is a connection between node $V_{i}$ and $V_{j}$ and 0 otherwise, and with 0's for every $a_{ii}$.
-Because the graph is undirected we have a symmetric matrix.
-\\
+The matrix has a non-zero entry $a_{ij}$ if there is a connection from node $N_{i}$ to $N_{j}$. \\
 
@@
 If \textit{n} is the smallest nonnegative integer, such that for all i ,j , the (i,j)-entry of $A^{n} > 0$, then n is the distance between vertex i and vertex \textit{j}.''}  source: \cite{wikimatrix} .\\
+Using this property of an adjacency matrix it is possible to calculate the time it takes for a worm, starting on a specific node, to infect a sufficient amount of other nodes in the network.
+Every matrix $A^{n}$ has a non-zero \textit{ij}-entry , if the worm starting in node \textit{i} can reach node $j$ in \textit{n} time steps.
+If we sum up all the matrices $A^{1} + A^{2} + ~...~+ A^{n-1}+A^{n}$, every \textit{i}-row indicates which node \textit{j} is infected by node \textit{i} after time $t=n$.
 
@@
 
-  \node[initial,state] (A)                    {$N_1$};
+  \node[state] (A)                    {$N_1$};
   \node[state]         (B) [above right of=A] {$N_2$};
@@
         (B) edge              node {} (A)
-        	edge			  node {} (C)
         (C) edge              node {} (B)
-            edge 			  node {} (D)
-            edge			  node {} (E)
             edge 			  node {} (F)
         (D) edge 			  node {} (C)
-            edge              node {} (A)
-        (E) edge 			  node {} (C)
-    	(F)	edge			  node {} (C);
+        (F) edge 			  node {} (E);
 \end{tikzpicture}
-\caption{Network with 6 nodes. The arrows represent the connections between the nodes. The start point is were the worm has been dropped, here node 1.}
+\caption{Network with 6 nodes. The arrows represent the connections between the nodes. }
 \label{netwerkfiguur}
@@
 %\] 
-Assuming a network like in figure \ref{netwerkfiguur}, the corresponding adjacency matrix is the  matrix $[A]$: \\
+Assuming a network like in figure \ref{netwerkfiguur}, the corresponding adjacency matrix is the  matrix $A$: \\
 
-$
+\begin{equation}
+A =
 \bordermatrix{
@@
     N_1   & 0		& 1		& 1		& 0		& 0		& 0	     \cr
+    N_2   & 1		& 0		& 0		& 0		& 0		& 0	     \cr
+    N_3   & 0		& 0		& 0		& 1		& 0		& 0	     \cr
+    N_4   & 0		& 1		& 0		& 0		& 0		& 1	     \cr
+	N_5   & 0		& 0		& 0		& 0		& 0		& 0	     \cr
+	N_6   & 0		& 0		& 0		& 0		& 1		& 0	     \cr
+}
+\end{equation}
+
+In matrix $A \times A = A^{2}$, each entry represents the number of  paths with length 2 from $N_{i}$ to $N_{j}$:\\
+
+\begin{equation}
+ A \times A = A^{2} =
+\bordermatrix{
+         & N_1		& N_2	& N_3	& N_4 	& N_5	&N_6     \cr
+    N_1   & 1		& 0		& 0		& 1		& 0		& 0	     \cr
+    N_2   & 0		& 1		& 1		& 0		& 0		& 0	     \cr
+    N_3   & 0		& 1		& 0		& 0		& 0		& 1 	     \cr
+    N_4   & 1	    & 0		& 0		& 0		& 1		& 0	     \cr
+	N_5   & 0		& 0		& 0		& 0		& 0		& 0	     \cr
+	N_6   & 0		& 0		& 0		& 0		& 0		& 0	     \cr
+}
+\end{equation}
+
+Likewise, in matrix $A^{2} \times A = A^{3}$ each entry represents the number of paths with 3 steps from $N_{i}$ to $N_{j}$.\\
+
+\begin{equation}
+ A \times A \times A = A^{3} =
+\bordermatrix{
+         & N_1		& N_2	& N_3	& N_4 	& N_5	&N_6     \cr
+    N_1   & 0		& 2		& 1		& 0		& 0		& 0	     \cr
     N_2   & 1		& 0		& 0		& 1		& 0		& 0	     \cr
-    N_3   & 1		& 0		& 0		& 1		& 0		& 0	     \cr
-    N_4   & 0		& 1		& 1		& 0		& 1		& 1	     \cr
-	N_5   & 0		& 0		& 0		& 1		& 0		& 0	     \cr
-	N_6   & 0		& 0		& 0		& 1		& 0		& 0	     \cr
-}$
+    N_3   & 1		& 0		& 0		& 0		& 1		& 0	     \cr
+    N_4   & 0		& 1		& 1		& 0		& 0		& 0	     \cr
+	N_5   & 0		& 0		& 0		& 0		& 0		& 0	     \cr
+	N_6   & 0		& 0		& 0		& 0		& 0		& 0	     \cr
+}
+\end{equation}
+
+~~\\
+So, in $A^{n}$ every $a_{ij}$ entry gives the number of paths with \textit{n} steps from $N_{i}$ to $N_{j}$.\\
+
+Calculating the sum of the three matrices $(A + A^{2} + A^{3}) $ results in a matrix that indicates which nodes are infected after a 3 time steps:
+
+\begin{equation}
+ A + A^{2} + A^{3} = \sum A^{n} =
+\bordermatrix{
+         & N_1		& N_2	& N_3	& N_4 	& N_5	&N_6     \cr
+    N_1   & 1		& 3		& 2		& 1		& 0		& 1	     \cr
+    N_2   & 2		& 1		& 1		& 1		& 0		& 0	     \cr
+    N_3   & 1		& 1		& 0		& 1		& 1		& 1	     \cr
+    N_4   & 1		& 2		& 1		& 0		& 1		& 1	     \cr
+	N_5   & 0		& 0		& 0		& 0		& 0		& 0	     \cr
+	N_6   & 0		& 0		& 0		& 0		& 1		& 0	     \cr
+}
+\end{equation}
+
+With the formula $\sum A^{n}$ the average delay can be calculated by counting the expected delay of every node when a worm has been dropped and devide it by the number of nodes in the network. \\
+
+Matrix \textit{A} can also represent a graph where every link get a weight.
+The weight is equal to the probability that the worm spreads through that link.
+The outcome of $ \sum A^{n}$ is than equal to the probability that each node has been infected after n time steps.
+The average delay can be calculated by setting up a threshold value for the probability that a node is infected.
 \\
 
-In matrix $A \times A = A^{2}$, each entry represents the number of 2 step paths from $N_{i}$ to $N_{j}$: We denote this matrix as matrix \textit{[B]}\\
-
-$
-\bordermatrix{
-         & N_1		& N_2	& N_3	& N_4 	& N_5	&N_6     \cr
-    N_1   & 2		& 0		& 0		& 2		& 0		& 0	     \cr
-    N_2   & 0		& 2		& 2		& 0		& 1		& 1	     \cr
-    N_3   & 0		& 2		& 2		& 0		& 1		& 1	     \cr
-    N_4   & 2		& 0		& 0		& 4		& 0		& 0	     \cr
-	N_5   & 0		& 1		& 1		& 0		& 1		& 1	     \cr
-	N_6   & 0		& 1		& 1		& 0		& 1		& 1	     \cr
-}$
+\subsubsection*{Conclusions}
+The advantage of this matrix worm model in comparison with the above models is that matrix  \textit{A} can represent any network topology. Directed or undirected. \\
+With this method it is also possible to determine the shortest path in the network.
+By summing up all the matrices, the first row that has all non-zero entries determines the shortest path.
+From node \textit{i} to the last \textit{ij}-entry that became a non-zero value.
+Consequently, if a worm is dropped on node \textit{i}, this worm will compromise the network in the shortest time possible.
+A defender can use this knowledge to adapt its network configurations.
 \\
-
-Likewise, in matrix $A^{2} \times A = A^{3}$ each entry represents the number of paths with 3 steps from $N_{i}$ to $N_{j}$: We denote this matrix as matrix \textit{[C]}\\
-
-$
-\bordermatrix{
-         & N_1		& N_2	& N_3	& N_4 	& N_5	&N_6     \cr
-    N_1   & 0		& 4		& 4		& 0		& 2		& 2	     \cr
-    N_2   & 4		& 0		& 0		& 6		& 0		& 0	     \cr
-    N_3   & 4		& 0		& 0		& 6		& 0		& 0	     \cr
-    N_4   & 0		& 6		& 6		& 0		& 4		& 4	     \cr
-	N_5   & 2		& 0		& 0		& 4		& 0		& 0	     \cr
-	N_6   & 2		& 0		& 0		& 4		& 0		& 0	     \cr
-}$
-
-~~\\
-So, in $A^{N}$ every $a_{ij}$ entry gives the number of paths with N steps from $N_{i}$ to $N_{j}$.\\
-
-Using this knowledge we can calculate in how many steps a node is infected: \\
-The infection vector or start vector $I_{0}$ of length $|V|$ indicates which node is infected and which not.
+A disadvantage of this method is the cost. A very large network needs a very large (sparse) matrix. 
+%Using this knowledge we can calculate in how many steps a node is infected: \\
+%The infection vector or start vector $I_{0}$ of length $|V|$ indicates which node is infected and which not.
 Multiplying the infection vector with $A$ results in a vector $I_{1}$, which represents which nodes are infected after 1 step.
@@
 \\
-In the context of the FlipIt game, it is safe to assume that once a node is infected, it stays infected until the defender Flips the node.
+%In the context of the FlipIt game, it is safe to assume that once a node is infected, it stays infected until the defender Flips the node.
 This means that after d steps, it is assumed that all nodes that could be reached in less than \textit{d} steps from the node where the worm was dropped, are infected.
@@
 \\
-
-Determining the length of the delay boils down to determining which configuration of infected nodes is considered as corresponding to the attacker having flipped the resource.
+%
+%Determining the length of the delay boils down to determining which configuration of infected nodes is considered as corresponding to the attacker having flipped the resource.
 It may be that all nodes need to be flipped, a sufficient amount of nodes, or a (set of) particular nodes.\\
-
---Hier nog een beetje verder uitschrijven hoe je exact de d dan moet bepalen, gegeven dat je niet weet in welke knoop het virus gedropt wordt-- \\
+%
+%--Hier nog een beetje verder uitschrijven hoe je exact de d dan moet bepalen, gegeven dat je niet weet in welke knoop het virus gedropt wordt-- \\
 
@@
 %\end{verbatim}
-\subsubsection{Google PageRank algorithm}
+\section{Google PageRank algorithm}
 
 To calculate the delay of a worm propagation we assumed that every node is even important.
-A possible relaxation of this rule can be done by the Page Rank Algorithm, which can be used to determine the importance of certain nodes.
+A possible relaxation of this rule can be done by the PageRank Algorithm, which can be used to determine the importance of certain nodes.
 

### 2015-08-12 doc/template/chap-4.tex
Samenvatting nog eens beschreven

@@
 where $p$ defines the damping factor and indicates the probability that a surfer quits a current page and continuous on a new random page.
-Matrix A is an transposed matrix of the graph matrix.
-The graph matrix is an $n x n$ matrix (\textit{n} defines the number of pages) that represents the graph where every entry corresponds to a non-zero entry if there is a link from one page to the other page.
+Matrix \textit{A} is an transposed matrix of the graph matrix.
+The graph matrix is an $n \times n$ matrix (\textit{n} defines the number of pages) that represents the graph where every entry corresponds to a non-zero entry if there is a link from one page to the other page.
 For dangling nodes (nodes with no outgoing links), the row entries are equal to $\dfrac{1}{n}$.
-Matrix A for the graph represented figure \ref{GoogleMatrix} is equal to:
+Matrix \textit{A} for the graph represented figure \ref{GoogleMatrix} is equal to:
 
@@
 
-The PageRank vector of a graph, with the transposition matrix A and the damping variable p, is equal to the probabilistic eigenvector of the matrix M, corresponding to the eigenvalue 1.
+The PageRank vector of a graph, with the transposition matrix\textit{ A} and the damping variable \textit{p}, is equal to the probabilistic eigenvector of the matrix \textit{M}, corresponding to the eigenvalue 1.
 This can be calculated by the use of the Power Method.
@@
 When a worm uses the topology of the graph to propagate the damping variable p must be equal to zero, as in the worm will not suddenly attack a node that is not connected.
-When the propagation of a worm is topology independent, the damping variable p can be set on 1.
+When the propagation of a worm is topology independent, the damping variable \textit{p} can be set on 1.
 \\
@@
 It can be advantageous to flip the nodes with a higher PageRank more often than the other nodes.
-A higer PageRank means that the page has a bigger probability of being compromised by a worm.
+A higher PageRank means that the page has a bigger probability of being compromised by a worm.

### 2015-08-15 doc/template/chap-10.tex
vanalles veranderd

@@
 %Since it is so difficult to protect a system or a network against APT's, researchers have been looking for effective ways to predict in advance which defence strategy might be the better one. 
-There are a number of key strategies an organisation can apply to defend itself against APT's: awareness, whitelisting, system administration, network segregation, dynamic content checking and patch management.
+There are a number of key strategies an organisation can apply to defend itself against APTs: awareness, whitelisting, system administration, network segregation, dynamic content checking and patch management.
 Nevertheless the combination of all these elements benefits from being complemented by other defence strategies to protect oneself against stealthy takeovers.
@@
 
-This paper adapts the model presented in \cite{FlipIt} so as to take the delay for virus propagation into account.
+This paper adapts the model presented in \cite{FlipIt} so as to take the delay for APT propagation into account.
 This leads us to the following research questions:
 \begin{itemize}
-\item How can we incorporate the notion of delay in the game-theoretical analysis of the Flip-It game for a periodic strategy?
+\item How can we incorporate the notion of delay in the game-theoretical analysis of the FlipIt game for a periodic strategy?
 \item Does the resulting model allow an optimal defence strategy against an attacker? 
@@
 The modelling of FlipIt with a propagation delay will give a guideline for both the attacker and the defender on how to respectively plan attacks or network clean-ups.
-\item[-] This work includes an overview of different techniques used to calculate the propagation delay of malware (depending on the network layout). \todo{en het wordt ook toegepast op de formules}
-\item[-] It presents a method to calculate the speed of the propagation of a worm in a network independent of the topology of the network.
+\item[-] This work includes an overview of different propagation techniques of worms and models used to calculate the propagation delay (depending on the network layout).
+It also presents a method to calculate the speed of the propagation of a worm in a propagation independent of the topology of the network.
+\item[-] The Google Page Rank matrix is introduced as a proof of concept to calculate the importance of every node in the network.
 \end{itemize}
@@
 The chapter concludes with an overview of the related work on FlipIt and further clarifies the contributions of this paper compared to existing work.
-Chapter \ref{chapter2:FlipIt with virus propagation} first introduces the adaptations made to the original FlipIt model to model a FlipIt game with virus propagation.
-Subsequently formulas are derived to model a FlipIt game with a virus propagation for the specific case where players play a periodic strategy with a random phase.
+Chapter \ref{chapter2:FlipIt with virus propagation} first introduces the adaptations made to the original FlipIt game to model a FlipIt game with a propagation delay.
+Subsequently formulas are derived to model a FlipIt game with a propagation delay for the specific case where players play a periodic strategy with a random phase.
 \\
@@
 Finally chapter \ref{chapter5:conclusion} discusses the main results and provides directions for further research.
-\todo{chapters kloppen nog niet}
 

### 2015-08-15 doc/template/chap-5.tex
vanalles veranderd

@@
-\chapter{General context}
+\chapter{General Context}
 \label{Chapter1:Intro.Game.Theory}
@@
 This chapter provides the reader with an introduction to the general context of the work presented in this paper.
-Section 2.1 introduces the reader into the basic concepts of cyber security and the kind of cyber security threats that are in scope of this work.
+First, section 2.1 introduces the reader into the basic concepts of cyber security and the kind of cyber security threats that are in scope of this work.
 Section 2.2 then introduces the reader into the main principles of game theory.
@@
  \\
-Information security can be divided into two main subcategory.
+Information security can be divided into two main subcategories.
 One of them is cyber security, also known as computer security.
@@
 The other category is network or internet security.
-This sub category of information security protects data during transmission.
+This subcategory of information security protects data during transmission.
 While cyber security and network security address different aspects, they partially overlap as well.
@@
 This is a piece of malicious software that is designed to penetrate unprotected or vulnerable systems or computers, with the intent to retrieve sensitive information, destroy data, or compromise the confidentiality, integrity or availability of the data or applications of the victim.
-A security report of 2014 \cite{SurveyKaspersky} reveals that 61\% of the attacks on companies are caused by malware.
+A security report of Kaspersky in 2014 \cite{SurveyKaspersky} reveals that 61\% of the attacks on companies are caused by malware.
 For this reason this section will examine the categories of malware threats.

### 2015-08-15 doc/template/chap-9.tex
case 1 en case 2.b

@@
 
-Similarly as in \cite{FlipIt}, we split the formalization in two cases.
-In the first case the defender plays at least as fast as the attacker, in the second case the attacker plays at least as fast as the defender.
+%
+We split the formalization in three cases.
+Similarly as in \cite{FlipIt}, there is the case where the defender plays at least as fast as the attacker, and the case where the attacker plays at least as fast as the defender.
 For each of these cases, the benefit formula of the basic case without delay is presented first, and subsequently the delay is introduced.
-\\
+The third case is a special case where the period of the attacker is smaller than the delay.\\
 
-Intuitively, we could assume that $d$ can never be bigger than $\delta_{A}$ because then the attacker would play again before the delay has finished.
+In the original model of FlipIt \cite{FlipIt}, the authors express the formulas in terms of $\alpha_{D}$ and $\alpha_{A}$.
+However, when introducing the delay, some formulas become much simpler when expressed in terms of $\delta_{D}$ and $\delta_{A}$.
+Therefore in the remainder of this paper, we will formulate the model using both $\alpha_{i}$ and $\delta_{i}$, depending on which variable gives the simplest representation of the formulas.
+
+\subsection*{Case 0: $\delta_{D} \leq d$ (The defender plays with a period smaller than the delay}
+
+When the delay $d$ is bigger than $\delta_{A}$, the attacker would play again before the delay has finished.
 This would seemingly result in a gain for the defender that is always 1, but this is not always true.
@@
 This situation is represented in figure \ref{langedelay}.
-Since the delay is shorter than the period of the defender, the attacker takes control of the resource once the delay has elapsed, until the defender plays.
-However, if the delay is larger that the period of the defender, then the defender will always be in control.
+Since the delay is shorter than the period of the defender, the attacker takes control of the resource once the delay has elapsed, until the defender plays.\\
+
+However, if the delay is larger than or equal to the period of the defender, the attacker will never gain control over the resource.
+The defender will always flip before the delay is over.
 This situation is represented in figure \ref{langeredelay}, where the attacker plays with an interval of 5 time units, the delay is equal to 6 time units and the defender plays with an interval of 4 time units.
-From this we can conclude that it only makes sense to calculate the formulas for the cases where $d$ is smaller than $\delta_{D}$.
+\\
+For the attacker it does not matter if he plays faster or slower than the defender. The benefit of the defender and the attacker are as follows:
+\begin{equation}\label{case0def}
+\beta_{D}(\delta_{D},\delta_{A})=1-\dfrac{k_{D}}{\delta_{D}}
+\end{equation}
+\begin{equation}\label{case0att}
+\beta_{A}(\delta_{D},\delta_{A})=-\dfrac{k_{A}}{\delta_{A}}
+\end{equation}
+%-Remark- hier staat iets dat niet juist is.
+%From this we can conclude that it only makes sense to calculate the formulas for the cases where $d$ is smaller than $\delta_{D}$.
 We can already conclude that it is of no use for the attacker to play when the delay is bigger than $\delta_{D}$.
@@
 
-In the original model of FlipIt \cite{FlipIt}, the authors express the formulas in terms of $\alpha_{D}$ and $\alpha_{A}$.
-However, when introducing the delay, some formulas become much simpler when expressed in terms of $\delta_{D}$ and $\delta_{A}$.
-Therefore in the remainder of this paper, we will formulate the model using both $\alpha_{i}$ and $\delta_{i}$, depending on which variable gives the simplest representation of the formulas.
+For the rest of the cases we assume that $\delta_{D} \geq d$.
 
@@
 For this case where $d$=0, we this equals the formula of the original FlipIt game \citep{FlipIt} [p675].\\
+For the cornercase with case 0, when $d=\delta_{D}$, the benefit of the defender and the attacker result in the same benefits as in case 0.
 
@@
 
+\subsection{Conclusion}
+
+The benefit formulas of case 1 and case 2.b result in the same formula.
+Consequently, these two formulas can be brought together by adapting the border conditions.
+Figure \ref{Borders} shows all the cases and it can be seen that the border condition of case 1 and case 2.b can be brought together.
+This results in the following figure \ref{endborders}.
+\\
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[scale=0.7]{Images/borders.pdf}
+\caption{the representation of the cases where the benefit functions are calculated.}
+\label{Borders}
+\end{figure}
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[scale=0.7]{Images/defborders.pdf}
+\caption{The representation of the new border conditions of piece B which is a merge of case 2.b and case 1.}
+\label{endborders}
+\end{figure}
+
+The benefit functions can be represented as piecewise funtions for each case, with case 1 and case 2.b together:\\
+The benefit formula for the defender:
+\begin{displaymath}
+\beta_{D}(\delta_{D},\delta_{A})= \left\{
+\begin{array}{lll}
+	1 - \dfrac{k_{D}}{\delta_{D}}, &~~~~~ & \delta_{D} \leq d \\
+	~ & ~ \\
+   1 - \dfrac{(\delta_{D} - d)^2}{2\delta_{D}\delta_{A}} - \dfrac{k_{D}}{\delta_{D}}, & ~~~~~~& d \leq \delta_{D} \leq d + \delta_{A} \\
+   ~ & ~ \\
+   \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - \dfrac{k_{D}}{\delta_{D}} &~~~~~& \delta_{D} \geq d + \delta_{A} 
+\end{array}
+\right.
+\end{displaymath}
+~~ \\
+The benefit formula for the attacker:
+\begin{displaymath}
+\beta_{A}(\delta_{D},\delta_{A})= \left\{
+\begin{array}{lll}
+	 -\dfrac{k_{A}}{\delta_{A}}, & ~~~~~& \delta_{D} \leq d \\
+	~ & ~ \\
+   \dfrac{(\delta_{D} - d)^2}{2\delta_{D}\delta_{A}} - \dfrac{k_{A}}{\delta_{A}}, & ~~~~~~& d \leq \delta_{D} \leq d + \delta_{A} \\
+   ~ & ~ \\
+   1 - \dfrac{\delta_{A}}{2\delta_{D}} - \dfrac{d}{\delta_{D}} - \dfrac{k_{A}}{\delta_{A}} &~~~~~& \delta_{D} \geq d + \delta_{A}
+
+\end{array}
+\right.
+\end{displaymath}
+
+The piecewise functions are also continuous functions. Filling in the border cases results in the same formula.
+
 %~~\\

### 2015-08-15 doc/template/chap-4.tex
vanalles veranderd
verbetering stijn

@@
 Because of the many different types of propagation methods it is hard to define a single model that can model all of them.
-Modelling the spread of malware depends on two key factors: the method used for the propagation and the graph of the network in which the malware will spread itself.
+Modelling the spread of malware depends on two key factors: the method used for the propagation and the graph of the network in which the malware will spread.
 \\
-Viruses and worms are the types of malware that are the most researched.
+Viruses and worms are the two most researched types of malware.
  Since the spreading of viruses requires human interaction, their propagation delay depends on (hard to predict) human behaviour.
@@
 The overview of this chapter will be as follows: Section \ref{methodsofpropagation} presents an overview of the most frequently used propagation methods by worms.
-These propagation methods will be covered by different kinds of models in section \ref{modelsforpropagation} illustrates a couple of models as examples to determine the propagation of a worm.
-These models can be used to extract parameter \textit{d}.
-Section \ref{eigenmatrixmethode} introduces an easy method to calculate the delay of the propagation of a worm.
+These propagation methods will be covered by different models in section \ref{modelsforpropagation}.
+The models are used to extract parameter \textit{d}.
+Section \ref{eigenmatrixmethode} introduces an easy method to calculate the propagation delay of a worm.
 In the last section \ref{GoogleMatrix}, a method based on the PageRank algorithm of Google is used to determine which node has a higher probability of being infected by a worm.
@@
 Different dropping mechanisms can be used, whereby the initial attack can be random or targeted at a specific node.
-Frequently used mechanisms are USB sticks (given to a specific person or left behind to be picked up by a random person), email, or malicious software through fishing or trojan horses.
+Frequently used mechanisms are USB sticks (given to a specific person or left behind to be picked up by a random person), email, or malicious software through phishing or trojan horses.
 To determine the total delay, however, the propagation strategy is of higher importance than the drop mechanism.
@@
 
-\item \textbf{Topology-based Worms} Email and other client application worms:An email worms uses the email systems to find email addresses to propagate.
+\item \textbf{Topology-based Worms:} Email and other client application worms: An email worms uses the email systems to find email addresses to propagate.
 Other client applications can include: Internet Relay Chat (IRC), Instant Messenger (IM), and a variety of peer-to-peer

### 2015-08-15 doc/template/chap-4.tex
verbeteringen stijn einde

@@
 To determine the total delay, however, the propagation strategy is of higher importance than the drop mechanism.
-Propagation is achieved by determining the next nodes to spread the worm to.
+The propagation strategy is used to determine the next nodes the worm will spread to.
 The following are common propagation methods:
@@
 \todo{Voorbeeeld als mensen het idee niet snappen?}
+%-Remark- Ik weet niet wat er bedoeld wordt met de selected address space-Stijn
 
@@
 This method is used by the \textit{Blaster worm} \citep{zou2006performance}.
+%-Remark -Stijn- Hier welke IP addressen worden er juist gescand? Alle of alleen lokale, niet echt duidelijk :) kan ook zijn dat het van de implementatie afhangt natuurlijk
 
@@
 A removed node is dead or immune, which means it cannot be infected again by worms.
-With these three states three main propagation models are proposed: SI, SIR and SIS.
+With these three states three main propagation models are proposed: SI (Suspected-Infected), SIR (Suspected-Infected-Removed) and SIS (Suspected-Infected-Suspected).
 In the SI model, a node that has once been infected, stays infected.
@@
 In the SIS model a node can become susceptible again after it has been infected.
-Currently, various other models have been proposed based on these three models (\cite{wang2014modeling}, \cite{OnWorms2005survey}, \cite{xiang2009propagation} and \cite{serazzi2004computer}).
+Currently, various other models have been proposed based on these three models ( Wang et al.
+\cite{wang2014modeling}, \cite{OnWorms2005survey}, \cite{xiang2009propagation} and \cite{serazzi2004computer}).
 We are only interested in models that will allow us to find the delay.
-\\
+\\ \todo{bij referenties titels schrijven}
 
@@
 The other propagation models, SIR and SIS, are only interesting if there is a possibility to reduce the models to an SI model.
-The following section will apply the extraction of the delay to a few models.
+The following section will extract the delay parameter from some of the models.
 These models are not meant as an exhaustive list, but rather as an illustration to the possible ways of extracting the delay out of an propagation model.
@@
 \end{figure}
+\todo{tabel herschrijven of op zen mist de tekst wegdoen}
 
@@
 
+%-Remark- Stijn- prentje van de curve maar hoeft niet
 %\begin{figure}[hbtp]
@@
 
+%-Remark -Stijn- begreep degree niet van nodes en topology independent
 \subsection{RCS model}
@@
 It is a model derived from the classical Simple Epidemic Model.
-They used this model to analyse the \textit{Code Red I} worm.
+This model was used to analyse \textit{Code Red I} worm.
 For this model it is assumed that the worm owns a good random number generator that is properly seeded.
@@
 Let\textit{ N} be the total of vulnerable hosts in the network that can be potentially infected.
-It is assumed that no system is patched, shut down, deployed or disconnected, which means that the number of hosts in the system stay constant.
+It is assumed that no system is patched, shut down, deployed or disconnected.
+This means that the number of hosts in the system stay constant.
 The model also ignores spreads of the worm behind firewalls on private networks.
-\\
+\\ \todo{wat betekent de laatste zin?}
 Let K be the initial compromise rate.
@@
 Every machine can infect only one other machine after it has been compromised.
-It cannot increase the rate that a worm can find new hosts.
+The rate that a worm can find hosts can not be increased.
 The internet topology is considered as a complete undirected graph.  The model type is SI, so the state of the host can only be infected or suspected. Once the host is compromised it stays that way.  
@@
 \end{equation}
-It tells how many vulnerable machines will be compromised in the next amount of time \textit{dt}, when it is known the proportion of machines \text{a} that already have been compromised.\\
+It tells how many vulnerable machines will be compromised in the next amount of time \textit{dt}, when the proportion of the machines \textit{a} that are already compromised is known.\\
 
@@
 The initial compromise rate has to be approximated.
-\textit{t} is a variable in the above formula that says how many time has passed.
+\textit{t} is the time variable.
 If \textit{a} passes the proportion of nodes that have to be compromised before the whole system is compromised by the attacker, value \textit{t}  is the value of \textit{d} in the FlipIt game with propagation delay.
 A drawback for this model is that the network topology is homogeneous and that the graphical representation of the graph is an undirected graph.  
+\todo{voorlaatste zin is nog moeilijk en te lang}
 
@@
 
+\todo{betekenis van alpha' uitleggen}
+
 To apply this model on FlipIt we again set up a threshold value for i(t) and see how long it takes to compromise this amount of hosts.
@@
 So in this case the delay is equal to an hour and the amount of hosts that has to be infected by the attacker before the network is compromised is 99$\%$ of the hosts in the network.
+\todo{remark stijn}
+%-Remark -Stijn- waarom 99 procent en niet 100?
+gewoon een uitleg in de trant van bij 99 procent kunt ge even goed zeggen dat heel het netwerk geinfecteerd is is goed.
+Of mss is er effectief een reden in de paper genoemd voor de 99
 
@@
 the scanning rate of the worm and $\Omega$ is the scanning
-space. \\
+space. \\ \todo{scanning space uitleggen}
 
@@
 The biggest drawback is that the model is a discrete time model. Continuous models are more appropriate for large scale models. The model also uses a complete undirected graph.
+\todo{The biggest drawback of the model is that is expressed in discrete time. Ofzoiets vind de zin niet heel goed}
 
@@
 As can be seen in table \ref{tree} all propagation models are depending on a specific network topology.
-The different kinds of network topology of each system is either homogeneous, small-world network, random network or a power-law network.
+The different kinds of network topology of each model is either homogeneous, small-world network, random network or a power-law network.
 A homogeneous network is a network where every node has about the same degree of connectivity.
@@
 A random network is a topology where each connection is chosen at random with equal probability.
-In a power law network the degrees of each node in the network follow the power law.
+In a power law \todo{ref power law} network the degrees of each node in the network follow the power law.
 Some of the nodes have a small degree of connectivity, others have a very large degree of connectivity.\\
@@
 \subsubsection{Model with matrix}
+%-Remark -Stijn- Ik vind de matrices maar weinig zeggen ik snap dat ze nodig zijn maar misschien kunt ge de overeenkomstige graph erbij tekenen of een pad erin uitlichten in kleur.
+bvb wie node 1 allemaal geinfecteerd heeft.
+Maar misschien is dat ook niet nuttig.
+Ik zou mij hier niet te veel van aantrekken
 
@@
 
-Matrix \textit{A} can also represent a graph where every link get a weight.
+Matrix \textit{A} can also represent a graph where every link has a weight.
 The weight is equal to the probability that the worm spreads through that link.
@@
 The advantage of this matrix worm model in comparison with the above models is that matrix  \textit{A} can represent any network topology. Directed or undirected. \\
-With this method it is also possible to determine the shortest path in the network.
+With this method it is also possible to determine the shortest path through al the nodes in the network.
 By summing up all the matrices, the first row that has all non-zero entries determines the shortest path.
@@
 
+%-Remark stijn- 2 verschillende definities of page rank? of heb ik het niet door
 To calculate the delay of a worm propagation we assumed that every node is even important.
@@
 \end{equation}
-where $N_{A}$ is the number of pages, $PR(i)$ is the ranking of web page $i$, $D_{out,i}$ is the number of outgoing links of page $i$, $(1-P)$ the probability that a random page is searched, and $e_{A}$ the restart value for web page A which is often uniformly distributed among all web pages.
+where $N_{A}$ is the number of pages, $PageRank(i)$ is the ranking of web page $i$, $D_{out,i}$ is the number of outgoing links of page $i$, $(1-P)$ the probability that a random page is searched, and $e_{A}$ the restart value for web page A which is often uniformly distributed among all web pages.
 \\
@@
 
+\todo{graph matrix symbool geven}
 where $p$ defines the damping factor and indicates the probability that a surfer quits a current page and continuous on a new random page.

### 2015-08-15 doc/template/chap-4.tex
Laatste chapter verbeteringen

@@
 Section \ref{eigenmatrixmethode} introduces an easy method to calculate the propagation delay of a worm.
-In the last section \ref{GoogleMatrix}, a method based on the PageRank algorithm of Google is used to determine which node has a higher probability of being infected by a worm.
+In the last section \ref{googlesec:5.4}, a method based on the PageRank algorithm of Google is used to determine the probability of infection after a certain time period.%which node has a higher probability of being infected by a worm.
 % Different kind of graphs to model a computer network are introduced.
@@
 
-\item \textbf{Topology-based Worms:} Email and other client application worms: An email worms uses the email systems to find email addresses to propagate.
+\item \textbf{Topology-based Worms:} Email and other client application worms: an email worms uses the email systems to find email addresses to propagate.
 Other client applications can include: Internet Relay Chat (IRC), Instant Messenger (IM), and a variety of peer-to-peer
@@
 Moreover, the survey explicitly focuses on propagation methods, while other papers also focus on other aspects such as  detection mechanisms and containment systems.
-All models that are based on SI model type (Susceptible-Infected) are good, because in our model every node that is infected will stay infected.
+All models that are based on the SI model type are appropriate, because in our model every node that is infected will stay infected.
 The other propagation models, SIR and SIS, are only interesting if there is a possibility to reduce the models to an SI model.
@@
 %%Er wordt geen rekening gehouden met de topology ? nog uitzoeken.
-To extract the delay from the formula we need the total amount of nodes in the network.
-If formula \ref{SIdelay} is equal to this amount, variable \textit{t} will be the value of the delay that we need.
-This is only applicable to nodes in a homogeneous network.
+To extract the delay from the formula we need the total number of nodes in the network.
+If \textit{I} in formula \ref{SIdelay} equals the total number of nodes, variable \textit{t} will be the value of the delay that we need.
+A limitation of this model is that it is only applicable to nodes in a homogeneous network.
 A homogeneous network is a network where every node has approximately the same degree.
@@
 The \textit{Code Red} worm, which is a random scan based worm, has been analysed by this kind of model in \cite{OwnInternetSI}.
-Another drawback is that the graph is presented as an undirected graph.
+Another limitation is that the graph is presented as an undirected graph.
 This means that the spread can always happen in both ways, which may not be suitable for every worm propagation.
@@
 
-RCS model stands for Random Constant Spread model and is developed by Paxson and Weaver at Stanford \citep{OwnInternetSI}.
-It is a model derived from the classical Simple Epidemic Model.
-This model was used to analyse \textit{Code Red I} worm.
+The RCS model stands for Random Constant Spread model and was developed by Paxson and Weaver at Stanford \citep{OwnInternetSI}.
+It is a model derived from the classical Simple Epidemic Model and was used to analyse the \textit{Code Red I} worm.
 For this model it is assumed that the worm owns a good random number generator that is properly seeded.
 \\
+
 Let\textit{ N} be the total of vulnerable hosts in the network that can be potentially infected.
 It is assumed that no system is patched, shut down, deployed or disconnected.
-This means that the number of hosts in the system stay constant.
+This means that the number of hosts in the system stays constant.
 The model also ignores spreads of the worm behind firewalls on private networks.
-\\ \todo{wat betekent de laatste zin?}
+\todo{wat betekent de laatste zin?}
 Let K be the initial compromise rate.
-This is the rate of hosts that the worm can find and compromise per hour at the beginning of the infection.
+This is the rate per hour at which the worm can find and compromise hosts at the beginning of the infection.
 \textit{K} is a global constant and therefore does not depend on the speed of the network or the processor speed.
@@
 The rate that a worm can find hosts can not be increased.
-The internet topology is considered as a complete undirected graph.  The model type is SI, so the state of the host can only be infected or suspected. Once the host is compromised it stays that way.  
 Let \textit{T} be the moment of the start of the infection. Variable \textit{a} is the proportion of vulnerable hosts that has been compromised. Variable \textit{t} is the time in hours. \\
@@
 \textit{t} is the time variable.
-If \textit{a} passes the proportion of nodes that have to be compromised before the whole system is compromised by the attacker, value \textit{t}  is the value of \textit{d} in the FlipIt game with propagation delay.
-A drawback for this model is that the network topology is homogeneous and that the graphical representation of the graph is an undirected graph.  
-\todo{voorlaatste zin is nog moeilijk en te lang}
+If \textit{a} passes the proportion of nodes that have to be compromised before the whole system is compromised by the attacker, value \textit{t}  is the value of \textit{d} in the FlipIt game with propagation delay.\\
+
+A drawback of this model is that the network topology is assumed to be homogeneous and that the paths between the nodes are undirected. 
+%-Remark- toevoeging van mama:
+%An example would be the Internet, since the topology of the Internet is considered to be a complete undirected graph.
+The RCS model is an SI model and therefor the state of a host can only be syspected or infected.
+Once a node has been compromised it stays that way.
+Finally, the model also ignores the spreading of worms behind firewalls on private networks.
 
@@
 
-The Bluetooth worm model is introduced by Yan and Eidenbenz \citep{yan2009modeling}. The model captures the behaviour of the propagation of a worm that spreads through the Bluetooth protocol. 
+The Bluetooth worm model was introduced by Yan and Eidenbenz \citep{yan2009modeling}. The model captures the behaviour of the propagation of a worm that spreads through the Bluetooth protocol. 
 A Bluetooth device that is compromised can only infect neighbour devices that are in its radio range.
-Let \textit{i(t)} be the average density of infected hosts in the network at time \textit{t}: 
+Let \textit{i(t)} be the average density of infected hosts in the network at time $t_{0}$. Then: 
 \begin{equation}
@@
 Of mss is er effectief een reden in de paper genoemd voor de 99
+\todo{limitations schrijven}
 
@@
 AAWP stands for Analytic Active Worm Propagation.
-The model is proposed in \cite{chen2003modeling} by Chen et al.
+The model was proposed in \cite{chen2003modeling} by Chen et al.
 to model the discrete behaviour of a worm.
@@
 It is different from other SI models because it includes the time that it takes to infect a host.
-A hosts cannot infect another hosts before it is completely compromised.
-The model also considers the fact that a hosts can be scanned and hit by multiple worms at the same time.
+A host cannot infect another hosts before it is completely compromised.
+The model also considers the fact that a host can be scanned and hit by multiple worms at the same time.
 \\%This model is used for worms that propagate through uniform scanning.
@@
 To extract the delay from this formula we need to know when the iteration procedure stops.
-\\
-The biggest drawback is that the model is a discrete time model. Continuous models are more appropriate for large scale models. The model also uses a complete undirected graph.
-\todo{The biggest drawback of the model is that is expressed in discrete time. Ofzoiets vind de zin niet heel goed}
+\\ \todo{en nu?}
+The biggest limitation of the model is that is expressed in discrete time. Continuous models are more appropriate for large scale models. The model also uses a complete undirected graph.
+\todo{. Ofzoiets vind de zin niet heel goed}
 
@@
 %All of the above models for modelling the mechanism of the spreading of a worm depend on the kind of topology of the network.  
-In this chapter we propose a method to calculate the spread of a worm in a way that the topology of the network can be easily integrated.
+In this chapter we propose a method to calculate the spread of a worm in a way that the specific topology of the network can be easily integrated.
 The method will give an approximation of how fast a worm can infect a network.
@@
 
-\subsubsection{Model with matrix}
+\subsubsection{Proposal of a matrix based model}
 %-Remark -Stijn- Ik vind de matrices maar weinig zeggen ik snap dat ze nodig zijn maar misschien kunt ge de overeenkomstige graph erbij tekenen of een pad erin uitlichten in kleur.
@@
 
-\begin{equation}
+\begin{equation}\label{5:13}
  A \times A \times A = A^{3} =
@@
 
-Calculating the sum of the three matrices $(A + A^{2} + A^{3}) $ results in a matrix that indicates which nodes are infected after a 3 time steps:
+Calculating the sum of the three matrices $(A + A^{2} + A^{3}) $ results in a matrix where the non-zero entries $a_{ij}$ indicate which nodes are infected after 3 time steps if a virus is dropped on noded \textit{i}:
 
@@
 
+For this sample network, the matrix in \ref{5:13} indicates that by dropping a virus on node $N_{1}$, all but node $N_{5}$ will be infected after 3 time steps.
+If the virus is dropped on node $N_{5}$, the virus will not spread.
+\\
+
 With the formula $\sum A^{n}$ the average delay can be calculated by counting the expected delay of every node when a worm has been dropped and devide it by the number of nodes in the network. \\
@@
 The weight is equal to the probability that the worm spreads through that link.
-The outcome of $ \sum A^{n}$ is than equal to the probability that each node has been infected after n time steps.
+The outcome of $ \sum A^{n}$ is then equal to the probability that each node has been infected after n time steps.
 The average delay can be calculated by setting up a threshold value for the probability that a node is infected.
@@
 \section{Google PageRank algorithm}
+\label{googlesec:5.4}
 
@@
 \begin{equation}
-PageRank(A)=P \sum \dfrac{PageRank(i)}{d_{out,i}} + (1-P) \cdot e_{A}
+PageRank(A)=P \sum \limits{i \in N_{A}} \dfrac{PageRank(i)}{d_{out,i}} + (1-P) \cdot e_{A}
 \end{equation}
-where $N_{A}$ is the number of pages, $PageRank(i)$ is the ranking of web page $i$, $D_{out,i}$ is the number of outgoing links of page $i$, $(1-P)$ the probability that a random page is searched, and $e_{A}$ the restart value for web page A which is often uniformly distributed among all web pages.
+\todo{som juist doen}
+where $N_{A}$ is the set of pages, $PageRank(i)$ is the page rank of web page $i$, $D_{out,i}$ is the number of outgoing links of page $i$, $(1-P)$ the probability that a random page is searched, and $e_{A}$ the restart value for web page A which is often uniformly distributed among all web pages.
 \\
@@
 
-We can use the Google matrix to calculate the probability that a node in a graph is infected by a worm after a certain amount op time.
-When a worm uses the topology of the graph to propagate the damping variable p must be equal to zero, as in the worm will not suddenly attack a node that is not connected.
-When the propagation of a worm is topology independent, the damping variable \textit{p} can be set on 1.
+We can use the Google matrix to calculate the probability that a node in a graph is infected by a worm after a certain amount of time.
+When a worm uses the topology of the graph to propagate the damping variable p must be set equal to zero, as in the worm will not suddenly attack a node that is not connected.
+When the propagation of a worm is topology independent, the damping variable \textit{p} can be set to 1.
 \\
-This can be useful for further research, when the defender can also flip a subset of nodes.
+The ranking of nodes provides interesting avenues for further research for the case where the defender flips a subset of nodes (rather than the entire network).
 It can be advantageous to flip the nodes with a higher PageRank more often than the other nodes.
-A higher PageRank means that the page has a bigger probability of being compromised by a worm.
+A higher PageRank means that the page has a higher probability of being compromised by a worm.

### 2015-08-15 doc/template/chap-9.tex
verbetering stijn
Samenvatting nog verder verbeterd

@@
 \includegraphics[scale=0.7]{Images/FlipItCase1delay.pdf} 
-\caption{FlipIt with delay propagation where $\delta_{D} > d > \delta_{A}$.   }
+\caption{FlipIt with delay propagation where $\delta_{D} \geq d \geq \delta_{A}$.   }
 \label{langedelay}
@@
 \includegraphics[scale=0.7]{Images/FlipItCase1delaytobig.pdf} 
-\caption{FlipIt with delay propagation where $ d > \delta_{D}$ }
+\caption{FlipIt with delay propagation where $ d \geq \delta_{D}$ }
 \label{langeredelay}
@@
 To determine the total gain rate of the defender we need to know the probability that the defender will move during an interval and what the average time is that the defender controls the resource.
-Given that if the defender moves he will always benefit entirely from a period of delay in the next interval of the attacker, his total gain is $\dfrac{(d+\dfrac{\delta_{A}}{2})}{\delta_{A}}$ in one interval (as previously calculated.
+Given that if the defender moves he will always benefit entirely from a period of delay in the next interval of the attacker, his total gain is $\dfrac{(d+\dfrac{\delta_{A}}{2})}{\delta_{A}}$ in one interval (as previously calculated).
 The total gain  rate of the defender is then the probability that the defender will move during an interval of the attacker multiplied by the total average gain per interval:
@@
 
-\subsection{Conclusion}
+\section{Summary}
 

### 2015-08-15 doc/template/chap-10.tex
APT virus

@@
 
-This paper adapts the model presented in \cite{FlipIt} so as to take the delay for APT propagation into account.
+This paper adapts the model presented in \cite{FlipIt} so as to take the delay for virus propagation into account.
 This leads us to the following research questions:
@@
 The chapter concludes with an overview of the related work on FlipIt and further clarifies the contributions of this paper compared to existing work.
-Chapter \ref{chapter2:FlipIt with virus propagation} first introduces the adaptations made to the original FlipIt game to model a FlipIt game with a propagation delay.
+Chapter \ref{chapter2:FlipIt with virus propagation} first introduces the adaptations made to the original FlipIt game to model a FlipIt game with a virus propagation delay.
 Subsequently formulas are derived to model a FlipIt game with a propagation delay for the specific case where players play a periodic strategy with a random phase.

### 2015-08-16 doc/template/chap-10.tex
Alles

@@
 The attacker only compromises the system if sufficient nodes in the network are infected.
-\item[-] The periodic case of FlipIt is modelled with a delay, resulting in optimum functions and Nash equilibria. \todo{punt herschrijven en meer nadruk leggen op het vinden van de Nash equilibria}
+\item[-] The periodic case of FlipIt is modelled with a delay, resulting in optimum functions which lead to interesting cases. 
 \item[-] Based on our results we can give practical recommendations to take measures against advanced attacks.
@@
 It also presents a method to calculate the speed of the propagation of a worm in a propagation independent of the topology of the network.
-\item[-] The Google Page Rank matrix is introduced as a proof of concept to calculate the importance of every node in the network.
+\item[-] The Page Rank matrix is introduced as a proof of concept to calculate the importance of every node in the network.
 \end{itemize}
@@
 \\
-In Chapter \ref{chapter:Nash} the formulas are further analysed to determine if there is a Nash equilibrium. 
+In Chapter \ref{chapter:Nash} the formulas are further analysed to determine optimal strategies for the defender and the attacker.
 In order to provide a clear perspective on delays, chapter \ref{chapter4: Worm propagation} gives an overview of the various methods of propagation and worm propagation models.
 It presents a method to calculate the speed of the propagation of a worm in a network independent of the topology of the network.
+It also introduces a PageRank matrix to calculate the importance of every node in the network.
 Finally chapter \ref{chapter5:conclusion} discusses the main results and provides directions for further research.

### 2015-08-16 doc/template/chap-4.tex
Alles

@@
 These propagation methods will be covered by different models in section \ref{modelsforpropagation}.
-The models are used to extract parameter \textit{d}.
+A set of models is selected to illustrate the different possible ways to extract parameter \textit{d}.
 Section \ref{eigenmatrixmethode} introduces an easy method to calculate the propagation delay of a worm.
@@
 \item \textbf{Selective Random scanning:} The worm randomly  selects a part of the selected IP address space instead of scanning the whole address space.
+The selected IP address space is a certain IP address area that the attackers are planning to attack.
 The reserved address blocks and the unassigned addresses are excluded from the address space.
@@
 Example of such worms are \textit{Code Red} \citep{OwnInternetSI} and \textit{Slammer} \citep{moore2003inside}.
-\todo{Voorbeeeld als mensen het idee niet snappen?}
-%-Remark- Ik weet niet wat er bedoeld wordt met de selected address space-Stijn
 
@@
 Early work on worm propagation models are based on the spread of real-world epidemic diseases. These models are based on the transition state of each node in the network.\\
+
 %Scan based or Topology based
@@
 Currently, various other models have been proposed based on these three models ( Wang et al.
-\cite{wang2014modeling}, \cite{OnWorms2005survey}, \cite{xiang2009propagation} and \cite{serazzi2004computer}).
+\cite{wang2014modeling}, Qing and Wen et al.
+\cite{OnWorms2005survey}, Xiang t al.
+\cite{xiang2009propagation} and Serazzi et al.
+\cite{serazzi2004computer}).
 We are only interested in models that will allow us to find the delay.
-\\ \todo{bij referenties titels schrijven}
+\\
 
-Table \ref{tree} gives an overview of the most common propagation methods with the given type.
-This table comes from the most recent survey \citep{wang2014modeling}, which encompasses results of older surveys.
-Moreover, the survey explicitly focuses on propagation methods, while other papers also focus on other aspects such as  detection mechanisms and containment systems.
+The most recent survey on common propagation methods is the one by Wang et al.
+\citep{wang2014modeling}.
+It encompasses the results of older surveys and explicitly focuses on propagation methods, while other papers also focus on other aspects such as  detection mechanisms and containment systems.
+Table \ref{tree} provides the taxonomy on worm modelling, given in \citep{wang2014modeling} (only analytic modelling methods).
  All models that are based on the SI model type are appropriate, because in our model every node that is infected will stay infected.
@@
 %\end{figure}
-\includegraphics[scale=0.55]{Images/tableworms.png}
-\caption{Taxonomy of worm modelling. Image based on taxonomy given in \cite{wang2014modeling}}.
+\includegraphics[scale=0.55]{Images/Picture2.png}
+\caption{Taxonomy of worm modelling.Table based on only analytic worm propagation models given in \cite{wang2014modeling}. The models in green are the models illustrated in this paper.}
 \label{tree}
 \end{figure}
-\todo{tabel herschrijven of op zen mist de tekst wegdoen}
 
@@
 A limitation of this model is that it is only applicable to nodes in a homogeneous network.
-A homogeneous network is a network where every node has approximately the same degree.
+A homogeneous network is a network where every node has approximately the same degree, the number of connections the node has.
 This model is thus suitable for propagation of worms that are topology independent (e.g scan-based worms), because every node has the same chance to be infected by another node in the network.
@@
 This means that the number of hosts in the system stays constant.
-The model also ignores spreads of the worm behind firewalls on private networks.
-\todo{wat betekent de laatste zin?}
+%The model also ignores a sudden spreads of the worm behind firewalls on private networks, because this can be misleading. \todo{wat betekent de laatste zin?}
 Let K be the initial compromise rate.
@@
 where $\rho$(t) and $\beta$(t) are the average device density and the pairwise infection rate at time \textit{t} respectively.
-The number of new infections out of the infection cycle is denoted by $\alpha$(t).\\
-
-The formula for a better estimation of the worm propagation:
+The number of new infections out of the infection cycle is denoted by $\alpha$(t).
+$\alpha'$ is used for a better estimation of the worm propagation, because the growth rate of the worm can change and this can maybe result in an overestimation of the number of new onfections.
+$\alpha'$ is determined as follows:
 \begin{equation}
@@
 
-\todo{betekenis van alpha' uitleggen}
-
 To apply this model on FlipIt we again set up a threshold value for i(t) and see how long it takes to compromise this amount of hosts.
@@
 So in this case the delay is equal to an hour and the amount of hosts that has to be infected by the attacker before the network is compromised is 99$\%$ of the hosts in the network.
+The Bluetooth worm spreads rapidly once the density of the infected hosts reach 10 percent. The model shows that the spread in the beginning is very slowly at en early stage.
 \todo{remark stijn}
@@
 Of mss is er effectief een reden in de paper genoemd voor de 99
-\todo{limitations schrijven}
+The limitation of this model is that it assumes that all the hosts in the network are homogeneously mixed and that the time increases in a discrete fashion.
+The graphical representation is also a undirected graph.
 
@@
 the scanning rate of the worm and $\Omega$ is the scanning
-space. \\ \todo{scanning space uitleggen}
+space.
 
@@
 When the death and patch rate are removed from the formula the iteration procedure will stop when all the nodes are infected.
-To extract the delay from this formula we need to know when the iteration procedure stops.
-\\ \todo{en nu?}
+So when the iteration procedure stops the delay will be equal to \textit{t+1}.
+\\
+
 The biggest limitation of the model is that is expressed in discrete time. Continuous models are more appropriate for large scale models. The model also uses a complete undirected graph.
@@
 \label{eigenmatrixmethode}
-For some of the propagation methods, the graph of the network matters.
-It is important to have the right topology for the right method.
-Email worms need a topology that represent a social network, BGP routing worms need a topology on network level.
-\\
 
@@
 
-What we want to have is a method to calculate the delay that is independent of a network topology. A method where we can chose in the beginning which topology we have and then calculate the delay.\\
+However, for some of the propagation methods, the actual graph of the network matters.
+Each propagation method depends on having a right topology.
+Email worms need a topology that represent a social network, BGP routing worms need a topology on network level.
+Ideally we need a method to calculate the delay that does not assume a specific kind of network topology, but where we can use the actual topology at hand to calculate the delay. \\
 
@@
 The method will give an approximation of how fast a worm can infect a network.
-This method is capable of calculating the delay for any network topology, but it is not the most efficient way of doing this.
-For this paper it is just a proof of concept.
+This method is capable of calculating the delay for any network topology.
 
@@
 \\
-A disadvantage of this method is the cost. A very large network needs a very large (sparse) matrix. 
+While matrix calculations may seem very time consuming, in the domain of network analytics efficient algorithms are available capable of calculating network attributes per node in networks of millions of nodes.
+In this paper, we only provide a proof of concept of how to apply matrices to calculate the delay.
+Developing an efficient algorithm is considered as a potential avenue for further research.
 %Using this knowledge we can calculate in how many steps a node is infected: \\
@@
 %\end{verbatim}
-\section{Google PageRank algorithm}
+\section{PageRank algorithm}
 \label{googlesec:5.4}
@@
 \begin{equation}
-PageRank(A)=P \sum \limits{i \in N_{A}} \dfrac{PageRank(i)}{d_{out,i}} + (1-P) \cdot e_{A}
+PageRank(A)=P \sum \limits_{i \in N_{A}} \dfrac{PageRank(i)}{d_{out,i}} + (1-P) \cdot e_{A}
 \end{equation}
-\todo{som juist doen}
+
 where $N_{A}$ is the set of pages, $PageRank(i)$ is the page rank of web page $i$, $D_{out,i}$ is the number of outgoing links of page $i$, $(1-P)$ the probability that a random page is searched, and $e_{A}$ the restart value for web page A which is often uniformly distributed among all web pages.
@@
 
-The Google matrix is a stochastic matrix that is used to calculate the PageRank of all the web pages on the internet.
+\textit{The Google matrix is a stochastic matrix that is used to calculate the PageRank of all the web pages on the internet.
 The dominant eigenvalue of the matrix is the PageRank.
@@
 Using the power method, the PageRank of each page can be calculated iteratively.
-\\
+} --Moet anders verwoord worden.
+Hoe ?
+komt uit wikipedia\\
 
@@
 
-\todo{graph matrix symbool geven}
-where $p$ defines the damping factor and indicates the probability that a surfer quits a current page and continuous on a new random page.
-Matrix \textit{A} is an transposed matrix of the graph matrix.
+%Matrix \textit{A} is an transposed matrix of the graph matrix.
 The graph matrix is an $n \times n$ matrix (\textit{n} defines the number of pages) that represents the graph where every entry corresponds to a non-zero entry if there is a link from one page to the other page.
@@
 Matrix \textit{A} for the graph represented figure \ref{GoogleMatrix} is equal to:
+Matrix \textit{A} corresponds to an $n \times n$ matrix (\textit{n} defines the number of pages) that represents the graph where every entry corresponds to a non-zero entry if there is a link from one page to the other page.
+$p$ defines the damping factor and indicates the probability that a surfer quits a current page and continuous on a new random page.
+Because the surfer can go to any random page, each page has a probability of $\dfrac{1}{n}$ to be chosen.
+This corresponds to matrix \textit{B}.\\
 
@@
 
-$ A =
-\bordermatrix{
-         & N_1		& N_2	& N_3	 \cr
-    N_1   & 1/3		& 1		& 1			     \cr
-    N_2   & 1/3		& 0		& 0		     \cr
-    N_3   & 1/3		& 0		& 0		 \cr
-  }$
-
-~~\\
+%$ A =
+%\bordermatrix{
+%         & N_1		& N_2	& N_3	 \cr
+%    N_1   & 1/3		& 1		& 1			     \cr
+%    N_2   & 1/3		& 0		& 0		     \cr
+%    N_3   & 1/3		& 0		& 0		 \cr
+%  }$ 
+% 
+%~~\\
 
@@
 A higher PageRank means that the page has a higher probability of being compromised by a worm.
+
+\todo{nog een algemene conclusie schrijven ?}

### 2015-08-16 doc/template/chap-5.tex
Samenvatting nog verder verbeterd
Alles

@@
 One of the assumptions underlying game theory is that the players of the game are independent and self-interested.
-This does not necessarily mean that they want to harm other agents or that they only care about themselves.
+%This does not necessarily mean that they want to harm other agents or that they only care about themselves.
 %utility function meaning 
-Instead it means that each agent has preferences about which states of the world he likes.
-These preferences are mapped to natural numbers and are called the utility functions.
-The numbers are interpreted as a mathematical measure that tells how much an agent likes or dislikes the outcome of the game.
+This means they do not actively seek to harm other agents.
+Instead, each agent has preferences about the state of the world.
+These preferences are mapped to natural numbers in a function called the utility function.
+These utility values are a mathematical measure indicating how much an agent likes or dislikes the outcome of the game.
 Outcomes are the result of a specific combination of a player's strategy.
 Each combination of a player's strategy is an outcome of the game.
-\\% The utility function can be ordinal or cardinal.
+A rational, independent and self-interested player prefers outcomes with a higher personal utility value, with no regard for the outcome of other players.\\
+
+% The utility function can be ordinal or cardinal.
 In case of a ordinal utility function only relative rankings are important.
@@
 %The choices of each agent produce an outcome, with respect to the utilities of the agent. A pay-off of an agent is the
-\todo{utility, pay-off, outcomes}
-Games can be divided into two types of games: cooperative or non-cooperative. In a non cooperative game, the basic modelling unit is the group of agents. 
+
+Games can be divided into two types of games: cooperative or non-cooperative. In \textit{non cooperative games}, the basic modelling unit is the group of agents. 
 Two or more agents want to maximise their utility and their actions can be affected by other agents' utilities.
 In the individualistic approach the basic modelling is only one agent.
-This type of game is referred to as \textit{cooperative game theory}.
+This type of game is referred to as \textit{cooperative games}.
 
@@
 This renewal process means that the strategies are renewed after each move: the interval until the next move only depends on the current move time and not on previous history.
-\todo{check comment in latex}
 %-Remark- Vorige zin is gecopypaste van FlipIt

### 2015-08-16 doc/template/chap-9.tex
Alles

@@
 
-\section{Introduction}
-\label{Ch2:Intro}
+%\section{Introduction}
+%\label{Ch2:Intro}
 The FlipIt game with propagation delay considers a game where the moves of the attacker are not instantaneous.
@@
 The defender can choose to flip one node, a subset of nodes or all nodes of the network.
-In this paper, the defender flips all the nodes in the network every time he plays.
+The analyses in this paper is based on the assumption that the defender flips all the nodes in the network every time he plays.
 This action can be extended to flipping only a subset of the nodes or a single node of the network.
-This extension has already been investigated in the context of placing anti-virus systems in a graph.
-This problem is known to be NP-hard.
-\cite{NPhard}.
-\\ \todo{niet volledig juist, defender systemen geplaatst en dan gezien hoe ver de worm kan geraken}
+This extension has already been investigated in the context of installing anti-virus systems on specific nodes in a graph and investigating how this affects the spreading of malware.
+In chapter \ref{chapter4: Worm propagation}, when discussing future work, the use of a matrix model is discussed to determine the probability that a node is infected by malware.  \\
 The attacker only has one action: sending different kinds of malware to a computer network.
@@
 Some APT's can be resident on a isolated network (e.g.
-Stuxnetworm), meaning that there is no way to connect to the internet.
+the Stuxnetworm), meaning that there is no way to connect to the Internet.
 \\
@@
 Honeypots emulate services or creates multiple instances of real operating systems and can pretend to have sensitive information.
-Honeypots do not detect all malicious attacks so the attacker can stay unnoticed and can still provide information as feedback.
-Furthermore, the case where the moves of the defender are non stealthy has already been investigated by Laszka \citep{MitigationNonTargeted}.
+The case where the moves of the defender are non stealthy has already been investigated by Laszka \citep{MitigationNonTargeted}.
 
@@
 Companies like Microsoft and Google release their security patches at fixed intervals, so companies will apply these patches as they come out.
-Microsoft security bulletins are released on the Second Tuesday of each month.
+Microsoft security bulletins are released, for example, on the Second Tuesday of each month.
 \cite{MicroPatch} \\
@@
 The second is the FlipIt game with a delay.
-During the first flip of the attacker, the defender moves after the delay, causing the attacker to get control over the resource.
-During the second flip of the attacker, the defender flips at time t+d, causing the defender to take control over the resource before the attacker.
+During the first flip of the attacker, the defender moves after the delay (point 2), resulting in the attacker gaining control over the resource, until the defender flips.
+During the second flip of the attacker, the defender flips at time t+d, causing the defender to retain control over the resource before the attacker.
 During the third and final flip of the attacker, the defender flips during time t+d, causing the attacker to never gain control over the resource.
@@
 
-\subsection*{Case 0: $\delta_{D} \leq d$ (The defender plays with a period smaller than the delay}
+\subsection*{Case 0: $\delta_{D} \leq d$ (The defender plays with a period smaller than the delay)}
 
@@
 \\
-For the attacker it does not matter if he plays faster or slower than the defender. The benefit of the defender and the attacker are as follows:
+When $ \delta_{D} \leq d$, it does not matter for the attacker whether he plays faster or slower than the defender. For this case the benefit of the defender and the attacker are as follows:
 \begin{equation}\label{case0def}
@@
 
-For the rest of the cases we assume that $\delta_{D} \geq d$.
+For the remainder of the cases we assume that $\delta_{D} \geq d$.
 
@@
 ~~\\
-For this case where $d$=0, we this equals the formula of the original FlipIt game \citep{FlipIt} [p675].\\
-For the cornercase with case 0, when $d=\delta_{D}$, the benefit of the defender and the attacker result in the same benefits as in case 0.
+For this case where $d$=0, this equals the formula of the original FlipIt game \citep{FlipIt} [p675].\\
+For the border case, when $d=\delta_{D}$, the benefit of the defender and the attacker result in the same benefits as in case 0.
 
@@
 The reason this formula overestimates the benefit of the attacker is that it assumes that the defender is always in control during the delay.
-However, if the attacker was in control in the previous interval, then he continuous to be in control during the period of the delay, see figure \ref{fig:case2}.
+However, if the attacker was in control in the previous interval, then he continues to be in control during the period of the delay, see figure \ref{fig:case2} situation 5 to 7.
 This means that the average benefit formulas for this case cannot be derived from one interval only; what happens in the previous interval must be taken into account.
@@
 Therefore, it is easier to calculate the benefit of the defender.
-Because the defender moves slower than the attacker we know that if the defender moves during the interval of the attacker, he only moves once within this interval.
+Because the defender moves more slowly than the attacker we know that if the defender moves during the interval of the attacker, he only moves once within this interval.
 
@@
 
-For the case with a delay we consider two cases, Case a and Case b, depending on whether the delay is shorter or longer than the difference between the attacker's and the defender's period.  \\
+For the remainder of the analyses we consider two cases, Case a and Case b, depending on whether the delay is shorter or longer than the difference between the attacker's and the defender's period.
+\\
 
@@
 This would result in an overlap between the average gain of $\dfrac{\delta_{A}}{2} +d$ and the delay.
-The above benefit formula would then include to much gain for the defender: the potential overlap during the delay would be counted twice.
+The above benefit formula would then include too much gain for the defender: the potential overlap during the delay would be counted twice.
 See figure \ref{countedtwice}\\
@@
 \centering
+\includegraphics[scale=0.7]{Images/flipcase2nb.pdf}
 \caption{Cases where the delay would be counted twice.}
-\includegraphics[scale=0.7]{Images/flipcase2nb.pdf}
 \label{countedtwice}
@@
 
+We calculated the benefit for three cases: when $\delta_{D}$ is smaller than $d$. When $\delta_{A}$ is smaller than $\delta_{A}$ and when it is larger than $\delta_{A}$.
 The benefit formulas of case 1 and case 2.b result in the same formula.

### 2015-08-16 doc/template/nash.tex
Alles

@@
-\chapter{Nash Equilibria}
+\chapter{Optimal defence and attack strategies}
 \label{chapter:Nash}
@@
 
-\section{Nash Equilibria}
 %TODO-- rechtstreeks uit FlipIt paper --\\
 
-In this chapter we are interested in finding the optimal strategies.
+In this chapter we are interested in finding the optimal defence and attack strategies.
 To calculate the optimal strategies of both players we have to find the Nash equilibria of the game.
@@
 To determine $opt_{D}(\delta_{A})$ we need to compute the derivative of  $\beta_{D}(\delta_{D},\delta_{A}) $ for a fixed $\delta_{A}$.
- We consider two cases, where case 2 is divided into two subcategories. 
+ We consider three cases for each piece of the piecewise function of $\beta_{D}$.
 % Fig \todo{fig} shows all the cases.\\
@@
 
-\subsection{Determining the piecewise functions $opt_{D}(\delta_{A})$}
+\section{Determining the piecewise functions $opt_{D}(\delta_{A})$}
 
-\subsection*{Case 1: $\delta_{D} \leq \delta_{A} $}
+\subsection*{Case A: $\delta_{D} \leq d$}
+
+The benefit formula obtained in the previous chapter is as follows:
+\begin{equation}
+\beta_{D}(\delta_{D},\delta_{A}) = 1 - \dfrac{k_{D}}{\delta_{D}}
+\end{equation}
+
+The function is of the type $1-1/x$, see figure \ref{1x}. The root of the benefit function  and the root of the first derivative are as follows:
+\begin{equation}
+\beta_{D}(\delta_{D},\delta_{A}) =0  ~~~~~~ =>~~~~~~\delta_{D} = k_{D} \\
+\end{equation}
+\begin{equation}
+\dfrac{\partial \beta_{D}(\delta_{D},\delta_{A})}{\partial \delta_{D}} =0 ~~~~~~ =>~~~~~~ k_{D} = 0
+\end{equation}
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[scale=1]{Images/1x.png}
+\caption{function of type $1-1/x$}
+\label{1x}
+\end{figure}
+
+This means that if $\delta_{D} < k_{D}$  the function is negative and if $\delta_{D} > k_{D}$ the function is positive.
+If $k_{D} < 0$ the function will decrease, and if $k_{D} > 0$ the function will increase.
+
+\subsection*{Case B: $ d \leq\delta_{D} \leq \delta_{A} + d$}
 The benefit formula obtained in the previous chapter (formula \ref{Benfcase1:defender}) for the defender in this case is as follows:
@@
 This leads to the following deduction given the sign of the coefficient of $\delta_{D}^{2}$: The function increases on $[0, \sqrt{2\delta_{A}k_{D} + d^{2}}]$ and is decreasing on $[\sqrt{2\delta_{A}k_{D} + d^{2}}, \infty]$.
-So we have a maximum at $\delta_{D} = min \{ \delta_{A}, \sqrt{2\delta_{A}k_{D} + d^{2}} \} $.
+So we have a maximum at $\delta_{D} = min \{ \delta_{A} +d, \sqrt{2\delta_{A}k_{D} + d^{2}} \} $ and $\delta_{D}$ has to be greater than $d$.
 Taking the minimum of the two values is needed because $\delta_{D}$ cannot be larger than $\delta_{A}$.
@@
 
-\subsection*{Case 2.A: $\delta_{D} \geq d+\delta_{A} \geq \delta_{A} $ }
+\subsection*{Case C: $\delta_{D} \geq d+\delta_{A} $ }
 
@@
 
-\subsection*{Case 2.B: $d+\delta_{A} \geq \delta_{D} \geq  \delta_{A} $}
-
-The benefit formula obtained in the previous chapter  (formula \ref{benfcase2b:defender}) for the defender in this case is as follows:
-\begin{equation*}
-\beta_{D}(\delta_{D},\delta_{A}) = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - \dfrac{k_{D}}{\delta_{D}} - \dfrac{(d-(\delta_{D} - \delta_{A}))^{2}}{2\delta_{D}\delta_{A}}
-\end{equation*}
-
-The derivative of the above formula for a fixed $\delta_{A}$ results in the following:
-\begin{equation*}
-\beta_{D}(\delta_{D},\delta_{A}) =  - \dfrac{1}{2\delta_{A}} + \dfrac{k_{D}}{\delta_{D}^{2}} + \dfrac{d^{2}}{2\delta_{D}^{2}\delta_{A}}
-\end{equation*}
-
-The stationary points (maximum, minimum) can be found by setting the first derivative equal to zero and finding the roots of the resulting equation:
-
-\begin{equation*}
-\frac{\partial \beta_{D}(\delta_{D},\delta_{A})}{\partial \delta_{D}} =0 ~~~~~~ =>~~~~~~ \delta_{D} = \sqrt{2\delta_{A}k_{D} + d^{2}}
-\end{equation*}
-
-For case 2.B this leads to the following deduction, which results in the same formula as for case 1 but with a small difference for the value of $\delta_{D}$: The function increases on $[0, \sqrt{2\delta_{A}k_{D} + d^{2}}]$ and is decreasing on $[\sqrt{2\delta_{A}k_{D} + d^{2}}, \infty]$.
+%\subsection*{Case 2.B: $d+\delta_{A} \geq \delta_{D} \geq  \delta_{A} $} 
+%
+%The benefit formula obtained in the previous chapter  (formula \ref{benfcase2b:defender}) for the defender in this case is as follows:
+%\begin{equation*}
+%\beta_{D}(\delta_{D},\delta_{A}) = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - \dfrac{k_{D}}{\delta_{D}} - \dfrac{(d-(\delta_{D} - \delta_{A}))^{2}}{2\delta_{D}\delta_{A}}
+%\end{equation*}
+%
+%The derivative of the above formula for a fixed $\delta_{A}$ results in the following:
+%\begin{equation*}
+%\beta_{D}(\delta_{D},\delta_{A}) =  - \dfrac{1}{2\delta_{A}} + \dfrac{k_{D}}{\delta_{D}^{2}} + \dfrac{d^{2}}{2\delta_{D}^{2}\delta_{A}}
+%\end{equation*}
+%
+%
+%The stationary points (maximum, minimum) can be found by setting the first derivative equal to zero and finding the roots of the resulting equation:
+%
+%\begin{equation*}
+%\frac{\partial \beta_{D}(\delta_{D},\delta_{A})}{\partial \delta_{D}} =0 ~~~~~~ =>~~~~~~ \delta_{D} = \sqrt{2\delta_{A}k_{D} + d^{2}}
+%\end{equation*}
+%
+%
+%For case 2.B this leads to the following deduction, which results in the same formula as for case 1 but with a small difference for the value of $\delta_{D}$: The function increases on $[0, \sqrt{2\delta_{A}k_{D} + d^{2}}]$ and is decreasing on $[\sqrt{2\delta_{A}k_{D} + d^{2}}, \infty]$.
 Because $\delta_{D} \geq \delta_{A}$ there is a maximum on $\delta_{D} = maximum \{ \delta_{A}, \sqrt{2\delta_{A}k_{D} + d^{2}} \} $ or because $d+\delta_{A} \geq \delta_{D}$ there is also on $\delta_{D} = minimum \{ \delta_{A}+d, \sqrt{2\delta_{A}k_{D} + d^{2}} \} $.
@@
 
-\subsubsection{Best responses}
-The optimum functions will be piecewise functions. We distinguish three cases for different values of $\delta_{A}$ depending on the values of $k_{D}$ and $d$. 
+\section{Best responses of $\delta_{D}$}
+The optimum functions will be piecewise functions. We distinguish six cases for different values of $\delta_{A}$ depending on the values of $k_{D}$ and $d$. 
 We point out that $\delta_{A}$ and $\delta_{D}$ are positive rates.
@@
 
-\subsubsection{$\delta_{A} = 2(k_{D} - d)$}
+\subsection*{$\delta_{A} = 2(k_{D} - d)$}
 Then for case 2.a, $\beta_{D}(\delta_{D}, \delta_{A})=0$ for  all $\delta_{D} \in [\sqrt{2k_{D}\delta_{A} + d^{2}}, \infty]$.
@@
 
-\subsubsection{$\delta_{A} > 2(k_{D} - d)$ }
+\subsection*{$\delta_{A} > 2(k_{D} - d)$ }
 From case 2.a above, it follows that the benefit is decreasing but positive.
@@
 
+~~\\
+
+\subsection*{Bordercases:}
+
+\subsection*{$\delta_{A}=0$ and $k_{D}=0$}
+If $\delta_{A}=0$ it means that the attacker is playing as fast as possible.
+For the defender it does not cost anything to flip, because $k_{D}=0$.
+We know that if $\delta_{D} \leq d$ the defender always has a benefit of 1.
+This means that for this case the defender will play with a rate $\delta_{D} \in [0,d]$.
+
+-- ofwel volgende uitleg--- \\
+If $k_{D}=0$ it means that every flip of the defender is free.
+So for the cost it does not matter for the defender if he plays faster or slower.
+$\delta_{A}=0$ means that the attacker will play as fast as he can.
+\begin{itemize}
+\item if $d=0$, if we look at $opt_{D}(\delta_{A})$ it means that we are in the second piece. The defender will play with a rate of $[0,\infty]$. 
+\item if $d >0$, we have to look at the third piece of  $opt_{D}(\delta_{A})$ . Here it follows that the defender has to play with a rate of \textit{d}.
+\end{itemize}
+
+Combining those two leads to a rate for the defender equal to $[0,d]$.
+\textit{Ofwel is de oplossing d}.
+This means that the defender will never not play, which is intuitively correct if we know that it has no disadvantage to play because there is no cost involved.
+\subsection*{$\delta_{A}=0$}
+If $\delta_{A}=0$ it means that the attacker is playing as fast as possible. The defender has still a cost for every flip.
+\begin{itemize}
+\item if $k_{D} > d$, we look at the first piece of $opt_{D}(\delta_{A})$. It follows that the defender will not play (rate equal to $\infty$).
+\item if $k_{D}=d$, corresponds to the second piece: the defender will play with a rate of $[d,\infty]$.
+\item if $k_{D} < 0$, for the last case the rate of the defender corresponds to \textit{d}.
+\end{itemize}
+~~\\
+Combining the results leads to a rate for the defender equal to $[d,\infty]$. 
+\subsection*{$k_{D}=0$}
+The cost of flipping for the defender is equal to zero.
+
 %*****************************************************************
@@
 %*********************************************************************
-\subsection{Determining the piecewise functions $opt_{A}(\delta_{D})$}
-To start with we only consider the case where $d < \delta_{D}$, because if $d > \delta_{D}$ the benefit of the defender \todo{nakijken, def of att} is always 1. \\
-To determine the Nash equilibria we also need to determine $opt_{A}(\delta_{D})$ by computing the derivative of $\beta_{A}(\delta_{D},\delta_{A})$ for a fixed $\delta_{D}$. We consider three cases: \\
+\section{Determining the piecewise functions $opt_{A}(\delta_{D})$}
+%To start with we only consider the case where $d < \delta_{D}$, because if $d > \delta_{D}$ the benefit of the defender \todo{nakijken, def of att} is always 1. \\
+To determine the Nash equilibria we also need to determine $opt_{A}(\delta_{D})$ by computing the derivative of $\beta_{A}(\delta_{D},\delta_{A})$ for a fixed $\delta_{D}$.
+We consider the three cases of the piecewise function of $\beta_{A}$: \\
 
-\subsection*{Case 1: $\delta_{A} \geq \delta_{D}$}
+\subsection*{Case A: $\delta_{D} \leq d$}
+
+The benefit formula obtained in the previous chapter is as follows:
+\begin{equation}
+\beta_{D}(\delta_{D},\delta_{A}) = 1 - \dfrac{k_{D}}{\delta_{D}}
+\end{equation}
+
+The function is of the type $1-1/x$, see figure \ref{1x}. The root of the benefit function  and the root of the first derivative are as follows:
+\begin{equation}
+\beta_{D}(\delta_{D},\delta_{A}) =0  ~~~~~~ =>~~~~~~\delta_{D} = k_{D} \\
+\end{equation}
+\begin{equation}
+\dfrac{\partial \beta_{D}(\delta_{D},\delta_{A})}{\partial \delta_{D}} =0 ~~~~~~ =>~~~~~~ k_{D} = 0
+\end{equation}
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[scale=1]{Images/1x.png}
+\caption{function of type $1-1/x$}
+\label{1x}
+\end{figure}
+
+This means that if $\delta_{D} < k_{D}$  the function is negative and if $\delta_{D} > k_{D}$ the function is positive.
+If $k_{D} < 0$ the function will decrease, and if $k_{D} > 0$ the function will increase.
+
+\subsection*{Case B: $d \leq \delta_{D} \geq \delta_{A} + d$}
 
@@
 
-\subsection*{Case 2.A: $\delta_{D} \geq d+\delta_{A} \geq \delta_{A} $ }
+\subsection*{Case B: $\delta_{D} \geq d+\delta_{A} \geq \delta_{A} $ }
 The benefit formula obtained from the previous chapter \ref{benfcase2a:attacker} for this case is as follows:
@@
 
-\subsubsection{Best responses}
+\section{Best responses of $\delta_{A}$}
 The optimum functions will be piecewise functions. We distinguish three cases for different values of $\delta_{D}$ regarding $d$ and $k_{A}$.

### 2015-08-17 doc/template/chap-10.tex
Verbeteringen aanpassen

@@
 This makes it so hard to protect a network or a system against an APT.
-
-%Bruce Schneier describes an APT as something different and stronger than a conventional threat: ''\textit{A conventional hacker or criminal is not interested in any particular target.
-He wants a thousand credit card numbers for fraud, or to break into an account and turn it into a zombie, or whatever.
-Security against this sort of attacker is relative; as long as you're more secure than almost everyone else, the attackers will go after other people, not you.
-An APT is different; it's an attacker who - for whatever reason - wants to attack you.
-Against this sort of attacker, the absolute level of your security is what's important.
-It does not matter how secure you are compared to your peers; all that matters is whether you're secure enough to keep him out}'' - Bruce Schneier \cite{APTBruce}.\\
-
-%\subsubsection{Complication}
-%Since it is so difficult to protect a system or a network against APT's, researchers have been looking for effective ways to predict in advance which defence strategy might be the better one. 
+\\
 There are a number of key strategies an organisation can apply to defend itself against APTs: awareness, whitelisting, system administration, network segregation, dynamic content checking and patch management.
@@
 Neither player knows who is currently in control of the resource until they move.
-In FlipIt every move involves a cost and gives them immediate control over the resource.
+In FlipIt every move involves a cost and gives the player immediate control over the resource.
 The attacker will try to maximise the time that he controls the network, while the defender will try to maximise the time that the network is free of malware.
@@
 The attacker only compromises the system if sufficient nodes in the network are infected.
-\item[-] The periodic case of FlipIt is modelled with a delay, resulting in optimum functions which lead to interesting cases. 
-\item[-] Based on our results we can give practical recommendations to take measures against advanced attacks.
+\item[-] The periodic case of FlipIt is modelled with a delay, resulting in adapted optimum functions for the defender and the attacker.
+These adapted optimum functions yield a number of interesting cases.
+%adapted optimum functions for the defender and the attacker. These adapted optimum functions yield a number of interesting cases.
+\item[-] Based on our results we give practical recommendations to take measures against advanced attacks.
 The modelling of FlipIt with a propagation delay will give a guideline for both the attacker and the defender on how to respectively plan attacks or network clean-ups.
-\item[-] This work includes an overview of different propagation techniques of worms and models used to calculate the propagation delay (depending on the network layout).
-It also presents a method to calculate the speed of the propagation of a worm in a propagation independent of the topology of the network.
+\item[-] While the exact value of the delay for modelling the FlipIt game is not required, this paper complements the results of the FlipIt game with an overview of different propagation techniques of worms and models.
+The models are used to calculate the propagation delay (depending on the network layout).
+The paper also presents a method to calculate the speed of the propagation of a worm in a propagation independent of the topology of the network.
 \item[-] The Page Rank matrix is introduced as a proof of concept to calculate the importance of every node in the network.
@@
 While it may seem trivial to extend the basic FlipIt model with a propagation delay, its mathematical treatment is not.
-In the paper of Laznka et all.
+In the paper of Laszka et all.
 \citep{FlipThem} it seems that even a small extension adds to the already significant mathematical complexity.
-Even though the FlipIt game is quite symmetric, the mathematical complexity rises by adding a propagation delay.
+While the FlipIt game is quite symmetric, the mathematical complexity rises substantially by adding a propagation delay on the side of the attacker.
+Besides yielding more complex functions, the symmetry is also largely lost.
 % First a formula to calculate the benefit of a basic FlipIt was derived to introduce the propagation delay.
@@
 
-The organisation of this paper is the following.
+The organisation of this paper is as follows.
 An introduction to cyber security and game theory is given in chapter \ref{Chapter1:Intro.Game.Theory} to allow the reader to become familiar with the kind of threats that are in the scope of this work and the game theoretic concepts that will be further used in the paper.
@@
 It presents a method to calculate the speed of the propagation of a worm in a network independent of the topology of the network.
-It also introduces a PageRank matrix to calculate the importance of every node in the network.
+It also explains how the PageRank algorithm may be used to calculate the sensitivity to infection of every node in the network.
+%introduces a PageRank matrix to calculate the importance of every node in the network.
 Finally chapter \ref{chapter5:conclusion} discusses the main results and provides directions for further research.

### 2015-08-17 doc/template/chap-4.tex
Verbeteringen aanpassen

@@
 %\subsection*{self disciplinary worms}
-%Model for self disciplinary worms and counter measures ... []  \todo{verder uitwerken ? ja of nee? }
+%Model for self disciplinary worms and counter measures ... []  \$•${verder uitwerken ? ja of nee? }
 
@@
 So in this case the delay is equal to an hour and the amount of hosts that has to be infected by the attacker before the network is compromised is 99$\%$ of the hosts in the network.
-The Bluetooth worm spreads rapidly once the density of the infected hosts reach 10 percent. The model shows that the spread in the beginning is very slowly at en early stage.
-\todo{remark stijn}
+The Bluetooth worm spreads rapidly once the density of the infected hosts reach 10 percent. The model shows that the spread in the beginning is very slowly at en early stage.\\
+
 %-Remark -Stijn- waarom 99 procent en niet 100?
@@
 The biggest limitation of the model is that is expressed in discrete time. Continuous models are more appropriate for large scale models. The model also uses a complete undirected graph.
-\todo{. Ofzoiets vind de zin niet heel goed}
 
@@
 
-As can be seen in table \ref{tree} all propagation models are depending on a specific network topology.
-The different kinds of network topology of each model is either homogeneous, small-world network, random network or a power-law network.
-A homogeneous network is a network where every node has about the same degree of connectivity.
-This means that every node can infect each node with the same opportunity.
-A small-world topology is a network where most nodes are not connected with each other but they can all reach each other in a few steps.
-Social networks are an example of a small-world network.
-A random network is a topology where each connection is chosen at random with equal probability.
-In a power law \todo{ref power law} network the degrees of each node in the network follow the power law.
+As can be seen in table \ref{tree} all propagation models depend on a specific network topology.
+A distinction is made between four kinds of network topologies: homogeneous, non-homogeneous, random network or power-law network.
+A homogeneous network is a network where every node has about the same degree of connectivity and each connection has the same probability.
+A non-homogeneous network is a network in which not all the nodes have the same degree and the connections have a different probability.
+%A small-world topology is a network where most nodes are not connected with each other but they can all reach each other in a few steps. Social networks are an example of such a small-world network. 
+A random network is a topology in which each connection is chosen at random with equal probability.
+In a power law network the degrees of each node in the network follow the power law.
 Some of the nodes have a small degree of connectivity, others have a very large degree of connectivity.\\
@@
 However, for some of the propagation methods, the actual graph of the network matters.
-Each propagation method depends on having a right topology.
-Email worms need a topology that represent a social network, BGP routing worms need a topology on network level.
-Ideally we need a method to calculate the delay that does not assume a specific kind of network topology, but where we can use the actual topology at hand to calculate the delay. \\
+Each propagation method depends on having the right topology.
+Email worms need a topology that correspond to a social network, BGP routing worms need a topology on IP-network level.
+Ideally, the propagation delay should be calculated without assuming that the network has one of these predefined topologies, but rather using the actual topology of the network at hand. \\
 
 %All of the above models for modelling the mechanism of the spreading of a worm depend on the kind of topology of the network.  
-In this chapter we propose a method to calculate the spread of a worm in a way that the specific topology of the network can be easily integrated.
-The method will give an approximation of how fast a worm can infect a network.
-This method is capable of calculating the delay for any network topology.
+This chapter proposes a method to calculate the spread of a worm in a way that allows easy integration of the network topology.
+This method approximate how fast a worm can infect a network.
+It is capable of calculating the delay for any network topology.
 
@@
 A computer network can be modelled by an undirected or directed graph $G = < N, E> $ where $|N|$ denotes the number of nodes in the network and $|E|$ the number of connections.
-This graph can be converted to an  $|N| \times |N|$ adjacency matrix where the entries represent the connections between the nodes of the network.
+This graph can be converted to an  $|N| \times |N|$ adjacency matrix in which the entries represent the connections between the nodes of the network.
 \\
@@
 If \textit{n} is the smallest nonnegative integer, such that for all i ,j , the (i,j)-entry of $A^{n} > 0$, then n is the distance between vertex i and vertex \textit{j}.''}  source: \cite{wikimatrix} .\\
-Using this property of an adjacency matrix it is possible to calculate the time it takes for a worm, starting on a specific node, to infect a sufficient amount of other nodes in the network.
+Using this property of an adjacency matrix, it is possible to calculate the time it takes for a worm, starting on a specific node, to infect a sufficient amount of other nodes in the network.
 Every matrix $A^{n}$ has a non-zero \textit{ij}-entry , if the worm starting in node \textit{i} can reach node $j$ in \textit{n} time steps.
@@
 The outcome of $ \sum A^{n}$ is then equal to the probability that each node has been infected after n time steps.
-The average delay can be calculated by setting up a threshold value for the probability that a node is infected.
+The average delay can be calculated by setting a threshold value for the probability that a node is infected.
 \\
 
-\subsubsection*{Conclusions}
-The advantage of this matrix worm model in comparison with the above models is that matrix  \textit{A} can represent any network topology. Directed or undirected. \\
-With this method it is also possible to determine the shortest path through al the nodes in the network.
-By summing up all the matrices, the first row that has all non-zero entries determines the shortest path.
+\subsubsection*{Conclusion}
+The advantage of this matrix worm model in comparison with the above models is that matrix  \textit{A} can represent any network topology, directed or undirected. \\
+Using this method it is also possible to determine the shortest path through al the nodes in the network.
+The shortest path that is necessary to infect all nodes can be found by summing all the matrices.
+The first row that only has non-zero entries determines this path.
 From node \textit{i} to the last \textit{ij}-entry that became a non-zero value.
@@
 \\
-While matrix calculations may seem very time consuming, in the domain of network analytics efficient algorithms are available capable of calculating network attributes per node in networks of millions of nodes.
-In this paper, we only provide a proof of concept of how to apply matrices to calculate the delay.
+These matrix calculations may seem very time consuming, but in the domain of network analytics efficient algorithms are available that are capable of calculating network attributes per node in networks of millions of nodes.
+This paper only provides a proof of concept of how to apply these matrices to calculate the delay.
 Developing an efficient algorithm is considered as a potential avenue for further research.
@@
 %-Remark stijn- 2 verschillende definities of page rank? of heb ik het niet door
-To calculate the delay of a worm propagation we assumed that every node is even important.
-A possible relaxation of this rule can be done by the PageRank Algorithm, which can be used to determine the importance of certain nodes.
 
-The PageRank algorithm is introduced by Page and Brin in 1998 as one of the main features of the search engine Google to improve the search results.
+The above method gives the propagation delay with the infection starting on a certain node.
+The following method, the PageRank algorithm, will give the probability that a node is infected, with the assumption that every node in the network is equally likely to be infected as a starting point.\\
+
+The PageRank algorithm was introduced by Page and Brin in 1998 as one of the main features of the search engine Google to improve the search results.
 PageRank models the human behaviour when users surf through the net.
 It can also model \textit{random surfers}.
-A random surfer is the probability that a surfer gets bored and randomly visits another page which is not linked with the initial page.
+A random surfer represents the probability that a surfer gets bored and randomly visits another page which is not linked with the initial page.
 The probability that he will visit a random page is the PageRank of that page.
-A page will have a high PageRank if many pages will point to this page.
-This means that this page is well cited through other pages and maybe an important one to look at.
+A page will have a high PageRank if many pages point to this page.
+This means that the page is well-cited through other pages and may be important to look at.
 The main idea of the PageRank algorithm is to look at the number of (important) web pages that point to a particular page.
-The ranking of this page depends than on the number of out coming links and the importance of the pages that link to this page.
-With a probability of $P$ the surfer will follow a link of the page to another page, and with a probability of $1-P$ the surfer will surf to a random page.
+The ranking of this page then depends on the number of outgoing links and the importance of the pages that link to this page.
+With a probability of $P$, the surfer will follow a link of the page to another page, and with a probability of $1-P$ the surfer will surf to a random page.
 The PageRank algorithm is expressed as follows:
@@
 
-\textit{The Google matrix is a stochastic matrix that is used to calculate the PageRank of all the web pages on the internet.
-The dominant eigenvalue of the matrix is the PageRank.
-The matrix is a graph where the edges denote the links between each page.
-Using the power method, the PageRank of each page can be calculated iteratively.
-} --Moet anders verwoord worden.
-Hoe ?
-komt uit wikipedia\\
+The PageRank of each page on the internet is the dominant eigenvalue of the Google Matrix.
+The Google Matrix is a stochastic matrix and represents a graph where every edge denotes a link between pages.
+Through the power iteration  \ref{Arnoldi}, the PageRank of all the pages can be calculated iteratively.
+\\
 
-The Google Matrix of a graph is defined as follow:
+The Google Matrix of a graph is defined as follows:
 \begin{equation}
@@
 The PageRank vector of a graph, with the transposition matrix\textit{ A} and the damping variable \textit{p}, is equal to the probabilistic eigenvector of the matrix \textit{M}, corresponding to the eigenvalue 1.
-This can be calculated by the use of the Power Method.
+This can be calculated by the use of the power iteration.
 \\
@@
 When a worm uses the topology of the graph to propagate the damping variable p must be set equal to zero, as in the worm will not suddenly attack a node that is not connected.
-When the propagation of a worm is topology independent, the damping variable \textit{p} can be set to 1.
+If the propagation of a worm is topology independent, the damping variable \textit{p} can be set to 1.
 \\
@@
 A higher PageRank means that the page has a higher probability of being compromised by a worm.
-
-\todo{nog een algemene conclusie schrijven ?}

### 2015-08-17 doc/template/nash.tex
Verbeteringen aanpassen

@@
 
-In this chapter we are interested in finding the optimal defence and attack strategies.
-To calculate the optimal strategies of both players we have to find the Nash equilibria of the game.
-First the optimal functions are derived from the formulas in the previous chapter.
-From these piecewise functions we can derive the Nash equilibria.
-\\
+In this chapter we are interested in finding the optimal defence and attack strategies. Ultimately, the optimal strategies can allow the determination of the Nash equilibria of the game. \\
+
+%First the optimal functions are derived from the formulas in the previous chapter. From these piecewise functions we can derive the Nash equilibria. \\
 
@@
 
-To determine $opt_{D}(\delta_{A})$ we need to compute the derivative of  $\beta_{D}(\delta_{D},\delta_{A}) $ for a fixed $\delta_{A}$.
- We consider three cases for each piece of the piecewise function of $\beta_{D}$.
 % Fig \todo{fig} shows all the cases.\\
@@
 \section{Determining the piecewise functions $opt_{D}(\delta_{A})$}
+
+To determine $opt_{D}(\delta_{A})$ we need to compute the derivative of  $\beta_{D}(\delta_{D},\delta_{A}) $ for a fixed $\delta_{A}$.
+ We consider three cases for each piece of the piecewise function of $\beta_{D}$.
 
@@
 
-This means that if $\delta_{D} < k_{D}$  the function is negative and if $\delta_{D} > k_{D}$ the function is positive.
-If $k_{D} < 0$ the function will decrease, and if $k_{D} > 0$ the function will increase.
+This means that if $\delta_{D} < k_{D}$  the function is negative and the defender will therefore not play, if $\delta_{D} > k_{D}$ the function is positive.
+If $k_{D} < 0$ the function will decrease and the defender will again not play (cost cannot be negative).
+Assuming that costs are always non-negative $k_{D} > 0$, the function will increase, meaning that the slower the defender plays, the larger the benefit.
 
 \subsection*{Case B: $ d \leq\delta_{D} \leq \delta_{A} + d$}
-The benefit formula obtained in the previous chapter (formula \ref{Benfcase1:defender}) for the defender in this case is as follows:
+The defender's benefit formula obtained in the previous chapter in this case is as follows:
 \begin{equation*}
@@
 
-This leads to the following deduction given the sign of the coefficient of $\delta_{D}^{2}$: The function increases on $[0, \sqrt{2\delta_{A}k_{D} + d^{2}}]$ and is decreasing on $[\sqrt{2\delta_{A}k_{D} + d^{2}}, \infty]$.
-So we have a maximum at $\delta_{D} = min \{ \delta_{A} +d, \sqrt{2\delta_{A}k_{D} + d^{2}} \} $ and $\delta_{D}$ has to be greater than $d$.
-Taking the minimum of the two values is needed because $\delta_{D}$ cannot be larger than $\delta_{A}$.
+Given the sign of the coefficient of $\delta_{D}^{2}$, this
+leads to the following deduction: The function increases on $[0, \sqrt{2\delta_{A}k_{D} + d^{2}}]$ and is decreasing on $[\sqrt{2\delta_{A}k_{D} + d^{2}}, \infty]$.
+So we have a maximum at $\delta_{D} = min \{ \delta_{A} +d, \sqrt{2\delta_{A}k_{D} + d^{2}} \} $ and $\delta_{D}$ has to be larger than $d$.
+Taking the minimum of the two values is needed because $\delta_{D}$ cannot be larger than $\delta_{A} +d$.
 \\
@@
 
-The benefit formula obtained in the previous chapter (formula \ref{benfcase2a:defender}) for the defender in this case is as follows:
+The benefit formula for the defender obtained in the previous chapter in this case is as follows:
 \begin{equation*}
@@
 The defender will always play as fast as he can if $\delta_{A} + 2(d-k_{D}) > 0$ for $k_{D} < d$ or $k_{D} > d$ because $\delta_{A}$ will be positive in either case.
-There is an edge case if $d=k_{D}$, which results in a benefit of $\beta_{D}(\delta_{D},\delta_{A})= \dfrac{\delta_{A}}{\delta_{D}}$.
+There is an edge case if $d=k_{D}$, which results in a benefit of $\beta_{D}(\delta_{D},\delta_{A})= \dfrac{\delta_{A}}{2\delta_{D}}$.
 \\
@@
 \includegraphics[scale=0.5]{Images/ShapesUp.png} 
-\caption{The benefit function is of the shape of 1/x if $\delta_{D}$ is always decreasing and if $\delta_{A} + 2(d-k_{D}) > 0$. }
+\caption{The benefit function is of the shape of 1/x and is always decreasing and if $\delta_{A} + 2(d-k_{D}) > 0$. }
 \label{ShapeUp}
@@
 \includegraphics[scale=0.5]{Images/ShapeDown.png} 
-\caption{The benefit function is of the shape of -1/x if $\delta_{D}$ is always increasing and if $\delta_{A} + 2(d-k_{D}) < 0$.}
+\caption{The benefit function is of the shape of -1/x and is always increasing and if $\delta_{A} + 2(d-k_{D}) < 0$.}
 \label{ShapeDown}
@@
 The optimum functions will be piecewise functions. We distinguish six cases for different values of $\delta_{A}$ depending on the values of $k_{D}$ and $d$. 
-We point out that $\delta_{A}$ and $\delta_{D}$ are positive rates.
+%We point out that $\delta_{A}$ and $\delta_{D}$ are positive rates.
 
@@
 
-\subsection*{Bordercases:}
+\subsection*{Edge cases:}
 

### 2015-08-18 doc/template/conclusion.tex
Google Matrix
Conclusie geschreven enzo

--- 2015-02-18 doc/template/conclusion.tex
@@
 \chapter{Conclusion}
-\label{cha:conclusion}
-The final chapter contains the overall conclusion. It also contains
-suggestions for future work and industrial applications.
+\label{chapter5:conclusion}
 
-\section{trala}
+This paper presents an adaptation to the basic FlipIt game by \citep{FlipIt} to model an attacker with a delay.  We list a set of extensions that might be interesting for further research.
+\section{Further work}
+
+\begin{description}
+\item \textit{Nash equilibria}\\ This paper concluded with the best responds functions of the attacker and the defender.
+The next step would be to calculate the Nash equilibria.
+For these calculation multiple cases have to be examined.
+All the possible cases regarding $k_{A}, k_{D}$ and $d$.
+\item \textit{Analysing other renewal strategies} \\ By analysing other renewal strategies, it might be interesting to find out if the original result regarding periodic and renewal strategies of the basic FlipIt game still stands.
+This result was that periodic strategies strongly dominate the other renewal strategies if an opponent plays with a non-adaptive strategy.
+\item \textit{Delay for the defender}\\ This paper only assumed that the attacker had a delay.
+It could be interesting to have a defender with a delay.
+This situation could happen when it takes some time to make a patch a system or when a new exploit is found.
+\item \textit{Variable delay} \\This paper assumed that the propagation delay had a fixed value.
+This can be relaxed by giving the attacker a delay that can vary every flip.
+The delay is always negative for the attacker, so the attacker will always choose the lowest delay.
+But if the attacker always sends a new worm at every flip, it could be that the delays vary.
+So a variable delay would be more accurate to simulate a real world case.
+\item \textit{The defender flips a subset of nodes} \\ Our model assumes that if the defender flips, he flips the whole network.
+An option for the defender would be to flip a subset of nodes in the network.
+If the costs of the flipping are related to the amount of nodes that are flipped, this option might be interesting to look at.
+It could be that a certain subset of nodes is found that protects the whole network.
+The PageRank matrix explained in Chapter 5 can help with finding the right nodes.
+\\
+
+\end{description}
+
+\section{General results and conclusions}
+Our FlipIt model with worm propagation delay showed us some interesting results. 
+\begin{itemize}
+\item When the defender plays faster than the delay, the attacker will either have a negative benefit or a benefit equal to zero if the flips do not cost anything.
+It is for the defender a target to be able to play at a rate smaller or equal to the delay.
+\item If the defender can play with a cost equal to zero, the attacker will not play. The defender can play at any rate he wants, and the attacker is always disadvantaged by his delay.
+\item From a certain value for the speed of the attacker, the defender will not play any more or will play at the same rate. The same for the attacker.
+\end{itemize}
+
+As of today APTs become more and more extraordinary pieces of malicious code.
+There are APTs known to survive military-grade disk wiping and reformatting, and even after the reformatting and reinstalling the operating system are able to send sensitive data (Equation group \citep{Equation}).
+This means that even if you know that an APT is on your network, the actual practical "flip" has to exist or known by the defender.
+\\
+The best way to secure a system is to find the weakest link.
+This is mostly the employee.
+To counter this problem, you have to raise the awareness of the risks.
+Awareness is very important.
+Even if you have good security protection layers, spam filters, firewalls, a phishing mail opened by one of your employees or a USB stick is enough to infect a pc on the network and by that it can spread further.
+But a system can never be 100\% waterproof.
+If an infection has occurred and the defender knows a counter measure, the FlipIt game will help to prevent a system to be compromised.
 %%% Local Variables: 

### 2015-08-18 doc/template/nash.tex
Conclusie geschreven enzo

@@
 \subsection*{$\delta_{A} < 2(k_{D} - d)$}
-From case 1 above, it follows that the benefit is increasing but it will always be non-positive.
-The defender will try not to play ($\delta_{D}= \infty$).
-From case 2.a and case 2.b it follows that from interval $[2(k_{D} -d), \sqrt{2k_{D}\delta_{A} + d^{2}}]$  (because for case 2 $\delta_{D} \geq \delta_{A}$), that the benefit is increasing.
+From case C above, it follows that if $\delta_{A} < 2(k_{D} - d)$ the benefit is increasing but it will always be non-positive. The defender will try not to play ($\delta_{D}= \infty$).
+
+%From case C it follows that from interval $[2(k_{D} -d), \sqrt{2k_{D}\delta_{A} + d^{2}}]$  (because for case 2 $\delta_{D} \geq \delta_{A}$), that the benefit is increasing.
 The maximum of case 2.a and the maximum of case 2.b can be brought together.
@@
 \subsection*{$\delta_{A} = 2(k_{D} - d)$}
-Then for case 2.a, $\beta_{D}(\delta_{D}, \delta_{A})=0$ for  all $\delta_{D} \in [\sqrt{2k_{D}\delta_{A} + d^{2}}, \infty]$.
-For case 2.b and case 1, the benefit is increasing in the interval $[\delta_{A},\sqrt{2k_{D}\delta_{A} + d^{2}}]$.
+For case C, $\beta_{D}(\delta_{D}, \delta_{A})=0$ for  all $\delta_{D} \in [\sqrt{2k_{D}\delta_{A} + d^{2}}, \infty]$.
+For case B, the benefit is increasing in the interval $[\delta_{A},\sqrt{2k_{D}\delta_{A} + d^{2}}]$.
 If we put both cases back together, we get a maximum at $\sqrt{2k_{D}\delta_{A} + d^{2}}$.
@@
 \subsection*{$\delta_{A} > 2(k_{D} - d)$ }
-From case 2.a above, it follows that the benefit is decreasing but positive.
+For case C, it follows that the benefit is decreasing but positive.
 So the defender will try to play as fast as possible ($\delta_{D} \rightarrow 0$).
-For case 2.b it is increasing in the interval $[ \delta_{A} ,\sqrt{2k_{D}\delta_{A} + d^{2}}]$ and for case 1 it is increasing in the interval $[0,\delta_{A}]$.
+For case B it is increasing in the interval $[ 0,\sqrt{2k_{D}\delta_{A} + d^{2}}]$.
 It follows that the defender's maximum benefit is achieved for $\delta_{D} = \sqrt{2k_{D}\delta_{A} + d^{2}} $  \\
@@
 Combining those two leads to a rate for the defender equal to $[0,d]$.
-\textit{Ofwel is de oplossing d}.
-This means that the defender will never not play, which is intuitively correct if we know that it has no disadvantage to play because there is no cost involved.
+This means that the defender will never not play, which is intuitively correct if we know that the defender has no disadvantage to play because there is no cost involved.
 \subsection*{$\delta_{A}=0$}
@@
 ~~\\
+   $  \infty \& \delta_{A}=0 ~~ \& ~~ k_{D} > d$ \\
 Combining the results leads to a rate for the defender equal to $[d,\infty]$. 
@@
 
+\subsection{Conclusion}
+With the edge cases we can compute $opt_{D}(\delta_{A})$ : \\
+
+ \begin{displaymath}
+  opt_{D}(\delta_{A}) = \left\{
+     \begin{array}{lr}
+     \left[0,d\right] & \delta_{A} =0 ~~\& ~~k_{D}=0 \\
+     \left[0,d\right] & k_{D}=0\\
+     d & \delta_{A} =0 ~~ \& k_{D} \leq d\\
+          \infty , & \delta_{A} < 2(k_{D} - d)\\
+      \left[ \sqrt{2k_{D}\delta_{A} + d^{2}},\infty\right[ , & \delta_{A} = 2(k_{D} - d) \\
+      \sqrt{2k_{D}\delta_{A} + d^{2}}, & \delta_{A} > 2(k_{D} - d)
+     \end{array}
+   \right.
+\end{displaymath}
 %*****************************************************************
@@
 %To start with we only consider the case where $d < \delta_{D}$, because if $d > \delta_{D}$ the benefit of the defender \todo{nakijken, def of att} is always 1. \\
-To determine the Nash equilibria we also need to determine $opt_{A}(\delta_{D})$ by computing the derivative of $\beta_{A}(\delta_{D},\delta_{A})$ for a fixed $\delta_{D}$.
+To determine the optimal strategy of the attacker we also need to determine $opt_{A}(\delta_{D})$ by computing the derivative of $\beta_{A}(\delta_{D},\delta_{A})$ for a fixed $\delta_{D}$.
 We consider the three cases of the piecewise function of $\beta_{A}$: \\
@@
 \begin{equation}
-\beta_{D}(\delta_{D},\delta_{A}) = 1 - \dfrac{k_{D}}{\delta_{D}}
+\beta_{A}(\delta_{D},\delta_{A}) =  - \dfrac{k_{A}}{\delta_{A}}
 \end{equation}
 
-The function is of the type $1-1/x$, see figure \ref{1x}. The root of the benefit function  and the root of the first derivative are as follows:
+The function is of the type $-1/x$, see figure \ref{1xx}. The root of the benefit function is as follows:
 \begin{equation}
-\beta_{D}(\delta_{D},\delta_{A}) =0  ~~~~~~ =>~~~~~~\delta_{D} = k_{D} \\
-\end{equation}
-\begin{equation}
-\dfrac{\partial \beta_{D}(\delta_{D},\delta_{A})}{\partial \delta_{D}} =0 ~~~~~~ =>~~~~~~ k_{D} = 0
+\beta_{A}(\delta_{D},\delta_{A}) =0  ~~~~~~ =>~~~~~~k_{D} = 0 \\
 \end{equation}
@@
 \includegraphics[scale=1]{Images/1x.png}
-\caption{function of type $1-1/x$}
-\label{1x}
+\caption{function of type $-1/x$}
+\label{1xx}
 \end{figure}
 
-This means that if $\delta_{D} < k_{D}$  the function is negative and if $\delta_{D} > k_{D}$ the function is positive.
-If $k_{D} < 0$ the function will decrease, and if $k_{D} > 0$ the function will increase.
+This means that if $k_{A} < 0$ the benefit function is positive and decreasing but $k_{D}$ cannot be negative so the attacker will not play.
+If $k_{A} > 0$ the function will be negative but always increasing.
 
@@
 
-\subsection*{Case 2.B: $d+\delta_{A} \geq \delta_{D} \geq  \delta_{A} $}
-
-The benefit formula obtained from the previous chapter \ref{benfcase2b:attacker} for this case is as follows: 
-\begin{equation*}
-\beta_{A}(\delta_{D},\delta_{A}) = 1 - \dfrac{\delta_{A}}{2\delta_{D}} - \dfrac{d}{\delta_{A}} - \dfrac{k_{A}}{\delta_{A}} + \dfrac{(d-(\delta_{D}-\delta_{A})^{2}}{2\delta_{D}\delta_{A}} 
-\end{equation*}
-The derivative for a fixed $\delta_{D}$ is as follows:
-\begin{equation*}
-\dfrac{\partial \beta_{A}(\delta_{D},\delta_{A})}{\partial \delta_{A}} = -\dfrac{\delta_{D}}{2\delta_{A}^{2}} + \dfrac{k_{A}}{\delta_{A}^{2}} - \dfrac{d^{2}}{2\delta_{D}\delta_{A}^{2}} + \dfrac{d}{\delta_{A}^{2}}
-\end{equation*}
-The stationary points (maximum, minimum) can be found by setting the first derivative equal to zero and finding the roots of the resulting equation:
-\begin{equation*}
-\frac{\partial \beta_{A}(\delta_{D},\delta_{A})}{\partial \delta_{D}} =0 ~~~~~~ =>~~~~~~ \delta_{D}= \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}
-\end{equation*}
-
-It follows that $\beta_{A}(\delta_{D},\cdot)$ is increasing and non-positif if $\delta_{D} > \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}$ and decreasing and positive if $\delta_{D} < \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}$.
+%\subsection*{Case 2.B: $d+\delta_{A} \geq \delta_{D} \geq  \delta_{A} $} 
+%
+%The benefit formula obtained from the previous chapter \ref{benfcase2b:attacker} for this case is as follows: 
+%\begin{equation*}
+%\beta_{A}(\delta_{D},\delta_{A}) = 1 - \dfrac{•}{•}\delta_{A}}{2\delta_{D}} - \dfrac{d}{\delta_{A}} - \dfrac{k_{A}}{\delta_{A}} + \dfrac{(d-(\delta_{D}-\delta_{A})^{2}}{2\delta_{D}\delta_{A}} 
+%\end{equation*}
+%The derivative for a fixed $\delta_{D}$ is as follows:
+%\begin{equation*}
+%\dfrac{\partial \beta_{A}(\delta_{D},\delta_{A})}{\partial \delta_{A}} = -\dfrac{\delta_{D}}{2\delta_{A}^{2}} + \dfrac{k_{A}}{\delta_{A}^{2}} - \dfrac{d^{2}}{2\delta_{D}\delta_{A}^{2}} + \dfrac{d}{\delta_{A}^{2}}
+%\end{equation*}
+%The stationary points (maximum, minimum) can be found by setting the first derivative equal to zero and finding the roots of the resulting equation:
+%\begin{equation*}
+%\frac{\partial \beta_{A}(\delta_{D},\delta_{A})}{\partial \delta_{D}} =0 ~~~~~~ =>~~~~~~ \delta_{D}= \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}
+%\end{equation*}
+%
+%
+%It follows that $\beta_{A}(\delta_{D},\cdot)$ is increasing and non-positif if $\delta_{D} > \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}$ and decreasing and positive if $\delta_{D} < \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}$.
 This is the same result as in case 1.\\
@@
      \begin{array}{lr}
-       \sqrt{d k_{A}\delta_{D}}, & \delta_{D} > \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}\\
+            \infty & \delta_{D} < \dfrac{(\delta_{D}-d)^{2}}{2k_{A}} \\
        \left[ \sqrt{d k_{A}\delta_{D}}, \infty\right],  & \delta_{D} = \dfrac{(\delta_{D}-d)^{2}}{2k_{A}} \\
-       \infty & \delta_{D} < \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}
+       \sqrt{d k_{A}\delta_{D}}, & \delta_{D} > \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}
      \end{array}
@@
 
+\subsection*{Edge cases:}
+
+\subsection*{$\delta_{D}=0$ and $k_{A}=0$}
+If $\delta_{D}=0$, looking at the benefit formula of the attacker, it follows that the period of the defender will always be equal or smaller than the delay.
+The delay cannot be negative.
+So from case $\delta_{D} \leq d$ it follows that the attacker will always have a benefit equal to zero, because $k_{A}=0$.
+For the attacker it does not matter what he plays.
+The benefit will always be the same.
+The optimal strategy in this case for the attacker will be playing at a rate $\delta_{A} \in [0,\infty ]$.
+
+\subsection*{$k_{A}=0$}
+If the cost of playing for the attacker is zero, the attacker can play as fast as he can.
+If $\delta_{D} \leq d$, looking at the benefit formula for the attacker, it does not matter what the attacker does, his benefit will always be zero.
+He will play with a rate $\delta_{A} \in [0,\infty ]$.
+We can merge this with the previous case where $\delta_{D}=0$ and $k_{A}=0$.
+If $\delta_{D} > d$, the attacker will play.
+He will play with a rate of $\delta_{A} \in [0, \delta_{D} - d ]$.
+\subsection*{$\delta_{D}=0$}
+If $\delta_{D}=0$ it follows from the case $\delta_{D} \leq d$ that the attacker will always have a negative benefit, unless the cost is zero.
+If the cost is zero we have again the case of $\delta_{D}=0$ and $k_{A}=0$.
+But if the cost is non-zero, the optimal strategy of the defender is not moving at all.
+
+\subsection{Conclusion}
+With the edge cases we can compute $opt_{A}(\delta_{D})$ : \\
+
+ \begin{displaymath}
+  opt_{A}(\delta_{D}) = \left\{
+     \begin{array}{lr}
+     \infty & \delta_{D} = 0 \\
+     \left[0,\infty\right] & \delta_{D} \leq d ~~\& ~~k_{A}=0 \\
+          \left[0,\delta_{D} - d\right] & \delta_{D} > d ~~\& ~~k_{A}=0 \\
+                 \infty & \delta_{D} < \dfrac{(\delta_{D}-d)^{2}}{2k_{A}} \\
+       \left[ \sqrt{d k_{A}\delta_{D}}, \infty\right],  & \delta_{D} = \dfrac{(\delta_{D}-d)^{2}}{2k_{A}} \\
+       \sqrt{d k_{A}\delta_{D}}, & \delta_{D} > \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}
+     \end{array}
+   \right.
+\end{displaymath}
+\\
+
+\section{Conclusion}
 The Nash Equilibria can be found as the intersection points of the piecewise functions $opt_{A}(\delta_{D})$ and $opt_{D}(\delta_{A})$.
@@
 We distinguish three cases: $k_{A} < k_{D} , ~k_{A} > k_{D} $ and $k_{A} = k_{D}$.
-\todo{ kA met kD vergelijken en nash evenwichten vinden}
-\todo{ een voorbeeld invullen met specifieke waarden}

### 2015-08-18 doc/template/chap-4.tex
Terug vanalles

@@
 
-The formalisation of the FlipIt game with a delay relies on some value  \textit{d} that represents the time needed to infect a sufficient number of nodes in a network after the initial infection.
+The formalization of the FlipIt game with a delay relies on some value  \textit{d} that represents the time needed to infect a sufficient number of nodes in a network after the initial infection.
 This chapter provides the reader with more insight on how to calculate the value of this parameter \textit{d}.

### 2015-08-18 doc/template/chap-5.tex
Terug vanalles

@@
 To define this concept we first introduce the concept of best response.
-The best response, given an action of the other player, is the action that maximizes its pay-off.
+The best response, given an action of the other player, is the action that maximises its pay-off.
 We define $Opt_{i}$ as the best response function for player \textit{i}.
 The best response for player \textit{1} is: $a_{1} = Opt_{1}(a_{2})$, given that $a_{2}$ is an action of player 2.
-For a Nash Equilibrium each player has a list of actions and each player's action maximizes his or her pay-off given the actions of the other players.
+For a Nash Equilibrium each player has a list of actions and each player's action maximises his or her pay-off given the actions of the other players.
 Nobody has the incentive to independently or unilaterally change his or her action if an equilibrium profile is played.
@@
 A player only finds out about the state of the game when he moves himself.
-The goal of the player is to maximize the time that he or she has control over the resource while minimizing the total cost of the moves.
+The goal of the player is to maximise the time that he or she has control over the resource while minimizing the total cost of the moves.
 A move can also result in a "wasted move", called a flop.
@@
 In this paper the delay is given by one parameter, which can be the result of any virus propagation model.
-The third and last difference is that the paper of Laska has multiple attackers who try to find the best strategy of the defender against both targeted and non-targeted attacks.
+The third and last difference is that the paper of Laszka has multiple attackers who try to find the best strategy of the defender against both targeted and non-targeted attacks.
 The conclusion of this paper is that the optimal strategy for the defender is moving periodically.

### 2015-08-18 doc/template/chap-9.tex
Verbeteringen aanpassen
Terug vanalles

@@
 The FlipIt game with propagation delay considers a game where the moves of the attacker are not instantaneous.
-This corresponds to APT's which use malware - viruses, worms or trojan horses - to perform attacks, as the malware needs time to propagate.
+This corresponds to APTs which use malware - viruses, worms or trojan horses - to perform attacks, as the malware needs time to propagate.
 A virus, for example, can be dropped on a network but it only compromises the whole network if every node in the network is infected.
@@
 Our motivation for a defender with stealthy moves is that it can not be assumed that every attacker can receive feedback from an APT.
-Some APT's can be resident on a isolated network (e.g.
+Some APTs can be resident on a isolated network (e.g.
 the Stuxnetworm), meaning that there is no way to connect to the Internet.
@@
 For each of these cases, the benefit formula of the basic case without delay is presented first, and subsequently the delay is introduced.
-The third case is a special case where the period of the attacker is smaller than the delay.\\
+The third case is a special case where the period of the defender is smaller than the delay.\\
 
@@
 Consequently, these two formulas can be brought together by adapting the border conditions.
-Figure \ref{Borders} shows all the cases and it can be seen that the border condition of case 1 and case 2.b can be brought together.
-This results in the following figure \ref{endborders}.
+The first figure in figure \ref{Borders} shows all the cases and it can be seen that the border condition of case 1 and case 2.b can be brought together.
+This can be seen as the second figure in \ref{endborders}.
 \\
 
-\begin{figure}[hbtp]
-\centering
-\includegraphics[scale=0.7]{Images/borders.pdf}
-\caption{the representation of the cases where the benefit functions are calculated.}
-\label{Borders}
-\end{figure}
-
-\begin{figure}[hbtp]
-\centering
-\includegraphics[scale=0.7]{Images/defborders.pdf}
-\caption{The representation of the new border conditions of piece B which is a merge of case 2.b and case 1.}
-\label{endborders}
-\end{figure}
-
-The benefit functions can be represented as piecewise funtions for each case, with case 1 and case 2.b together:\\
-The benefit formula for the defender:
+The benefit functions can be represented as piecewise functions for each case, with case 1 and case 2.b together:\\
+The benefit formula for the defender, see figure \ref{pfb} :
 \begin{displaymath}
@@
 ~~ \\
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[scale=1]{Images/functie4.pdf} 
+\caption{The benefit function of the defender for a cost $k_{D} =1$, $d =2$ and $\delta_{D}=1.5$. The three pieces match with the three colours.}
+\label{pfb}
+\end{figure}
+
 The benefit formula for the attacker:
@@
 The piecewise functions are also continuous functions. Filling in the border cases results in the same formula.  
+\begin{figure}[hbtp]
+\centering
+\includegraphics[scale=0.7]{Images/borders.pdf}
+\caption{The first figure is the representation of the cases where the benefit functions are calculated.
+The second figure is the representation of the new border conditions of piece B which is a merge of case 2.b and case 1.}
+\label{Borders}
+\end{figure}
 

### 2015-08-18 doc/template/nash.tex
Terug vanalles

@@
 More formally, a Nash equilibrium for the periodic game is a point $(\delta_{D}^{*},\delta_{A}^{*})$ such that
-the defender's benefit $\beta_{D}(\delta_{D},\delta_{A}^{*}) $is maximized at $\delta_{D}= \delta_{D}^{*}$ and the attacker's benefit
-$\beta_{A}(\delta_{D}^{*},\delta_{A}) $ is maximized at $\delta_{A}= \delta_{A}^{*}$.
+the defender's benefit $\beta_{D}(\delta_{D},\delta_{A}^{*}) $is maximised at $\delta_{D}= \delta_{D}^{*}$ and the attacker's benefit
+$\beta_{A}(\delta_{D}^{*},\delta_{A}) $ is maximised at $\delta_{A}= \delta_{A}^{*}$.
 To begin with, some useful notation. We denote by $opt_{D}(\delta_{A}$) the set of values (rates
-of play $\delta_{D}$) that optimize the benefit of the defender for a fixed rate of play $\delta_{A}$ of the
-attacker. Similarly, we denote by $opt_{D}(\delta_{D}$) the set of values (rates of play $\delta_{A}$) that optimize
+of play $\delta_{D}$) that optimise the benefit of the defender for a fixed rate of play $\delta_{A}$ of the
+attacker. Similarly, we denote by $opt_{D}(\delta_{D}$) the set of values (rates of play $\delta_{A}$) that optimise
 the benefit of the attacker for a fixed rate of play $\delta_{D}$ of the defender. \\
@@
 The cost of flipping for the defender is equal to zero.
+If we look at the benefit function of the defender and it's derivative in figure \ref{cost0}, we can see that the benefit for case B increases until $d$ and that the benefit in case A stays the same until $d$.
+After point $d$ the benefit decreases in all the cases.
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[scale=1]{Images/cost0.pdf} 
+\caption{Benefit function and derivative for a cost $k_{D} = 0$ and with $d (=2)$ and $\delta_{D}=(1.5)$ not equal to zero.}
+\label{cost0}
+\end{figure}
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[scale=1]{Images/cost1.pdf} 
+\caption{Benefit function and derivative for a cost $k_{D} > 0$ and with $d (=2)$ and $\delta_{D}=(1.5)$ not equal to zero.}
+\label{cost1}
+\end{figure}
 

### 2015-08-18 doc/template/chap-9.tex
shizzle

@@
 \centering
-\includegraphics[scale=1]{Images/functie4.pdf} 
-\caption{The benefit function of the defender for a cost $k_{D} =1$, $d =2$ and $\delta_{D}=1.5$. The three pieces match with the three colours.}
+\includegraphics[scale=1.4]{Images/functie4.pdf} 
+\caption{The benefit function of the defender for a cost $k_{D} =1$, $d =2$ and $\delta_{A}=1.5$. The three pieces match with the three colours.}
 \label{pfb}
@@
 
-The benefit formula for the attacker:
+The benefit formula for the attacker, see figure \ref{pfa}:
 \begin{displaymath}
@@
    1 - \dfrac{\delta_{A}}{2\delta_{D}} - \dfrac{d}{\delta_{D}} - \dfrac{k_{A}}{\delta_{A}} &~~~~~& \delta_{D} \geq d + \delta_{A} 
-
 \end{array}
@@
 \end{displaymath}
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[scale=1.4]{Images/benfa.pdf} 
+\caption{The benefit function of the attacker for a cost $k_{A} =1.2$, $d =0.5$ and $\delta_{D}=5$. The three pieces match with the three colours.}
+\label{pfa}
+\end{figure}
 

### 2015-08-18 doc/template/conclusion.tex
shizzle

--- 2015-02-18 doc/template/conclusion.tex
@@
 \chapter{Conclusion}
-\label{cha:conclusion}
-The final chapter contains the overall conclusion. It also contains
-suggestions for future work and industrial applications.
+\label{chapter5:conclusion}
 
-\section{trala}
+This paper presents an adaptation to the basic FlipIt game by \citep{FlipIt} to model an attacker with a delay.
+
+\section{General results and conclusions}
+Our FlipIt model with worm propagation delay showed us some interesting results. 
+\begin{itemize}
+\item When the defender plays faster than the delay, the attacker will either have a negative benefit or a benefit equal to zero if the flips do not cost anything.
+It is for the defender a target to be able to play at a rate smaller or equal to the delay.
+\item If the defender can play with a cost equal to zero, the attacker will not play. The defender can play at any rate he wants, and the attacker is always disadvantaged by his delay.
+\item If the cost of the defender is non-zero, then, depending on the ratio between it's cost and the delay, from a certain value for the speed of the attacker, the defender will not play.
+The same thing goes for the attacker.
+This result is similar to the results obtained for the original FlipIt game.
+\end{itemize}
+
+While the first two results are intuitively clear, in the third case, the mathematical analysis is important to determine the exact ratio's at which it is no longer beneficial for the defender to play.
+\\
+In the next section we list a set of extensions that might be interesting for further research.
+\section{Further work}
+
+\begin{description}
+\item \textit{Nash equilibria}\\ This paper concluded with the best responds functions of the attacker and the defender.
+The next step would be to calculate the Nash equilibria.
+For these calculation multiple cases have to be examined, i.e.
+all the possible cases regarding the mutual relations between $k_{A}, k_{D}$ and $d$.
+\item \textit{Analysing other renewal strategies} \\ By analysing other renewal strategies, it might be interesting to find out if the original result regarding periodic versus other renewal strategies of the basic FlipIt game still stands.
+This result implied that periodic strategies strongly dominate the other renewal strategies if an opponent plays with a non-adaptive strategy.
+\item \textit{Delay for the defender}\\ This paper only assumed that the attacker had a delay.
+It could be interesting to have a defender with a delay.
+This situation could happen if it takes some time to make a patch to a system or when a new exploit is found.
+\item \textit{Variable delay} \\This paper assumed that the propagation delay has a fixed value.
+This can be relaxed by giving the attacker a delay could vary.
+The delay always negatively impacts the benefit of the attacker, so the attacker will always choose the lowest delay.
+But if the attacker always sends a new worm at every flip, it could be that the delays vary.
+In that regard, a variable delay would be more accurate to simulate a real world case.
+\item \textit{The defender flips a subset of nodes} \\ Our model assumes that if the defender flips, he flips the whole network.
+An option for the defender would be to flip a subset of nodes in the network.
+If the costs of flipping are related to the amount of nodes that are flipped, this option might be interesting to look at.
+It could be that a certain subset of nodes is found that protects the whole network.
+The PageRank matrix explained in Chapter 5 can help finding the right nodes.
+\\
+
+\end{description}
+
+As of today APTs become more and more extraordinary pieces of malicious code.
+There are APTs known to survive military-grade disk wiping and reformatting, and even after reformatting and reinstalling the operating system they are able to send sensitive data (Equation group \citep{Equation}).
+This means that even if you know that an APT is on your network, the actual practical ``flip'' has to exist and be known by the defender.
+\\
+Most often the best way to secure a system is to find the weakest link.
+This is mostly the employee.
+An effective way to counter this problem is to raise awareness about risks of infections amongst the employees.
+Even in the presence of good security protection layers, spam filters, and firewalls, the opening of a phishing mail or use of an unknown USB stick by one of your employees may be suficient to infect a PC on the network, from which the virus can spread further.
+A system can however never be 100\% waterproof.
+If an infection has occurred and the defender knows a countermeasure, the FlipIt game will help to determine the most effective pace at which countermeasures need to be taken.
 %%% Local Variables: 

### 2015-08-18 doc/template/nash.tex
shizzle

@@
 %We point out that $\delta_{A}$ and $\delta_{D}$ are positive rates. 
+Figure XXX shows on the left how the three cases can be combined in a single continuous function (shown as the thick line).
+On the right, the derivative function indicates at which points the benefit will change.
+It is interesting to note that the derivative is also continuous, meaning that the benefit function is smooth and has no rough edges or corners.
+This implies that there are no sudden changes in benefit for small changes of period.
 
@@
 \subsection*{$\delta_{A}=0$}
-If $\delta_{A}=0$ it means that the attacker is playing as fast as possible. The defender has still a cost for every flip.
+If $\delta_{A}=0$ it means that the attacker is playing as fast as possible. The defender has still a cost for every flip. This means we have to look at different values for $k_{D}$ and $d$.
 \begin{itemize}
-\item if $k_{D} > d$, we look at the first piece of $opt_{D}(\delta_{A})$. It follows that the defender will not play (rate equal to $\infty$).
-\item if $k_{D}=d$, corresponds to the second piece: the defender will play with a rate of $[d,\infty]$.
-\item if $k_{D} < 0$, for the last case the rate of the defender corresponds to \textit{d}.
+\item if $k_{D} > d$, it follows that the defender will not play (rate equal to $\infty$).
+This is already the case for the first piece in the piecewise function.
+$\delta_{A}$ will always be smaller than $2(k_{D}-d)$ if $k_{D} > d.$
+\item if $k_{D}=d$, it corresponds to the second piece: the defender will play with a rate of $[d,\infty]$.
+\item if $k_{D} < 0$, will correspond with the last case, where the rate of the defender is equal to \textit{d}.
 \end{itemize}
-~~\\
-   $  \infty \& \delta_{A}=0 ~~ \& ~~ k_{D} > d$ \\
-Combining the results leads to a rate for the defender equal to $[d,\infty]$. 
+By combining the last two pieces we have  $d ~~for~~ \delta_{A} =0 ~~ and ~~k_{D} \leq d$.\\
+
 \subsection*{$k_{D}=0$}
@@
 \label{cost0}
-\end{figure}
-
-\begin{figure}[hbtp]
-\centering
-\includegraphics[scale=1]{Images/cost1.pdf} 
-\caption{Benefit function and derivative for a cost $k_{D} > 0$ and with $d (=2)$ and $\delta_{D}=(1.5)$ not equal to zero.}
-\label{cost1}
 \end{figure}
@@
 \section{Conclusion}
+
+The piecewise functions $opt_{D}$ and $opt_{A}$ give some interesting results. \todo{resultaten nog toevoegen, zelfde als in de andere conclusie}
+
 The Nash Equilibria can be found as the intersection points of the piecewise functions $opt_{A}(\delta_{D})$ and $opt_{D}(\delta_{A})$.

### 2015-08-19 doc/template/nash.tex
even alles opslagen

@@
 \includegraphics[scale=0.5]{Images/ShapesUp.png} 
-\caption{The benefit function is of the shape of 1/x and is always decreasing and if $\delta_{A} + 2(d-k_{D}) > 0$. }
+\caption{The benefit function is of the shape of 1/x and is always decreasing if $\delta_{A} + 2(d-k_{D}) > 0$. }
 \label{ShapeUp}
@@
 \includegraphics[scale=0.5]{Images/ShapeDown.png} 
-\caption{The benefit function is of the shape of -1/x and is always increasing and if $\delta_{A} + 2(d-k_{D}) < 0$.}
+\caption{The benefit function is of the shape of -1/x and is always increasing if $\delta_{A} + 2(d-k_{D}) < 0$.}
 \label{ShapeDown}
@@
 \subsection*{$\delta_{A} < 2(k_{D} - d)$}
-From case C above, it follows that if $\delta_{A} < 2(k_{D} - d)$ the benefit is increasing but it will always be non-positive. The defender will try not to play ($\delta_{D}= \infty$).
+For case A, the function will always increase if $k_{D} > 0$.
+If $k_{D} < 0$ it means that the cost is negative and we assume a positive cost.
+The function will thus never decrease.
+$k_{D} =0$ is an edge case and will be covered later in the paragraph of edge cases.
+So for case A we can conclude that the function is always increasing.
+In the next case, case B, the benefit function will increase in the interval $[0, 2(k_{D}-d)[$.
+In case C, for $\delta_{A} < 2(k_{D} - d)$, the benefit function will always increase.
+Because for all the cases we are dealing with a negative benefit function that is increasing, the maximum benefit is achieved for $\delta_{} = \infty$.
+This can be seen in (A) in figure \ref{case1D}.
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[scale=1]{Images/case1b.pdf}
+\caption{ Illustration of the best response functions.
+The left functions are the piecewise benefit function of the defender for a certain values of $k_{D},d$ and $\delta_{A}$..
+The right functions are the derivatives of the benefit functions on the left.
+The best responses are the values where the function intersects with the x-as.
+(A): $\delta_{A} < 2(k_{D}-d)$, (B): $\delta_{A} = 2(k_{D} - d)$, (C): $\delta_{A} > 2(k_{D} - d)$}
+\label{case1D}
+\end{figure}
+
+%From case C above, it follows that if $\delta_{A} < 2(k_{D} - d)$ the benefit is increasing but it will always be non-positive. The defender will try not to play ($\delta_{D}= \infty$).
 
@@
 \subsection*{$\delta_{A} = 2(k_{D} - d)$}
-For case C, $\beta_{D}(\delta_{D}, \delta_{A})=0$ for  all $\delta_{D} \in [\sqrt{2k_{D}\delta_{A} + d^{2}}, \infty]$.
+The delay can not be a negative value. So for $\delta_{A}$, a positive period, to be equal to  $2(k_{D} - d)$, $k_{D}$ has to be positive. \\
+For case A, if $k_{D}>0$ the benefit function is always increasing.
+In case B, the value of $\delta_{A} =2(k_{D} - d)$ has to filled in in the root, $\sqrt{2k_{D}\delta_{A} + d^{2}}$.
+This gives the value $2k_{D}-d$.
+$2(k_{D}-d) < 2k_{D}-d$ so this means that $\delta_{A}$ is smaller than $ \sqrt{2k_{D}\delta_{A} + d^{2}}$, so the function is increasing.
+A maximum is achieved for $\delta_{D}$ equal to the maximum of $\left\lbrace \delta_{A}+d, \sqrt{2k_{D}\delta_{A} + d^{2}}\right\rbrace $.
+If we fill in $\delta_{A} = 2(k_{D} - d)$, the two values are the same so they are both a maximum.
+If we fill in the value of $\delta_{A} = 2(k_{D} - d)$ in case C, the benefit will always be 0.
+Case C is the case for all values $\delta_{D} \geq \delta_{A} + d$, so $\beta_{D}=0$ is true for all $\delta_{D} \in [\delta_{A}+d, \infty]$.
+The defender's maximum benefit is thus achieved for all $\delta_{D} \in [\sqrt{2k_{D}\delta_{A} + d^{2}}, \infty]$. See (B) in figure \ref{case1D}.
+
+%For case C, $\beta_{D}(\delta_{D}, \delta_{A})=0$ for  all $\delta_{D} \in [\sqrt{2k_{D}\delta_{A} + d^{2}}, \infty]$.
 For case B, the benefit is increasing in the interval $[\delta_{A},\sqrt{2k_{D}\delta_{A} + d^{2}}]$.
@@
 \subsection*{$\delta_{A} > 2(k_{D} - d)$ }
-For case C, it follows that the benefit is decreasing but positive.
+Again for case A, we assume that $k_{D}$ is positive which implies that the benefit function for the defender is always increasing.
+If $\delta_{A} > 2(k_{D} - d)$, the benefit function of the defender in case B will only increase in the interval $]2(k_{D} - d), \sqrt{2k_{D}\delta_{A} + d^{2}}$.
+The maximum is achieved for $\delta_{D}$ equal to the maximum of $\left\lbrace \delta_{A}+d, \sqrt{2k_{D}\delta_{A} + d^{2}}\right\rbrace $.
+If we fill in $\delta_{A} > 2(k_{D} - d)$ , $\sqrt{2k_{D}\delta_{A} + d^{2}}$ is the maximum value.
+In case C, for all $\delta_{A} > 2(k_{D} - d)$, the benefit function is always decreasing.
+If we put all the pieces together, the defender's maximum benefit is achieved for $\delta_{D} = \sqrt{2k_{D}\delta_{A} + d^{2}}$.
+See (C) in figure \ref{case1D}.
+
+%For case C, it follows that the benefit is decreasing but positive.
 So the defender will try to play as fast as possible ($\delta_{D} \rightarrow 0$).
@@
 \subsection*{$\delta_{A}=0$ and $k_{D}=0$}
-If $\delta_{A}=0$ it means that the attacker is playing as fast as possible.
-For the defender it does not cost anything to flip, because $k_{D}=0$.
-We know that if $\delta_{D} \leq d$ the defender always has a benefit of 1.
-This means that for this case the defender will play with a rate $\delta_{D} \in [0,d]$.
 
--- ofwel volgende uitleg--- \\
 If $k_{D}=0$ it means that every flip of the defender is free.
@@
 Combining those two leads to a rate for the defender equal to $[0,d]$.
-This means that the defender will never not play, which is intuitively correct if we know that the defender has no disadvantage to play because there is no cost involved.
+This means that the defender will never not play, which is intuitively correct if we know that the defender has no disadvantage to play because there is no cost involved.\\
+
+Or more intuitively: If $\delta_{A}=0$ it means that the attacker is playing as fast as possible.
+For the defender it does not cost anything to flip, because $k_{D}=0$.
+We know that if $\delta_{D} \leq d$ the defender always has a benefit of 1.
+This means that for this case the defender will play with a rate $\delta_{D} \in [0,d]$.\\
 \subsection*{$\delta_{A}=0$}
@@
 \item if $k_{D}=d$, it corresponds to the second piece: the defender will play with a rate of $[d,\infty]$.
-\item if $k_{D} < 0$, will correspond with the last case, where the rate of the defender is equal to \textit{d}.
+\item if $k_{D} < 0$, this corresponds with the last piece, where the rate of the defender is equal to \textit{d}.
 \end{itemize}
-By combining the last two pieces we have  $d ~~for~~ \delta_{A} =0 ~~ and ~~k_{D} \leq d$.\\
+By combining the last two pieces we have  $d ~~$for$~~ \delta_{A} =0 ~~$ and$ ~~k_{D} \leq d$.\\
 
@@
 \includegraphics[scale=1]{Images/cost0.pdf} 
-\caption{Benefit function and derivative for a cost $k_{D} = 0$ and with $d (=2)$ and $\delta_{D}=(1.5)$ not equal to zero.}
+\caption{Benefit function and derivative for a cost $k_{D} = 0$ and with $d (=2)$ and $\delta_{A}=(1.5)$ not equal to zero.}
 \label{cost0}
@@
 \subsection{Conclusion}
-With the edge cases we can compute $opt_{D}(\delta_{A})$ : \\
+Including the edge cases, $opt_{D}(\delta_{A})$ is as follows:  \\
 
@@
 
-The benefit formula obtained in the previous chapter is as follows:
+The benefit formula for the attacker for this case obtained in the previous chapter is as follows:
 \begin{equation}
@@
 
-This means that if $k_{A} < 0$ the benefit function is positive and decreasing but $k_{D}$ cannot be negative so the attacker will not play.
-If $k_{A} > 0$ the function will be negative but always increasing.
+This means that the optimal benefit is achieved for a value $\delta_{A}=0$.
 
-\subsection*{Case B: $d \leq \delta_{D} \geq \delta_{A} + d$}
-
-The benefit formula obtained in the previous chapter \ref{Benfcase1:attacker} for this case is as follows:
-\begin{equation*}
-\beta_{A}(\delta_{D},\delta_{A}) =\dfrac{\delta_{D}}{2\delta_{A}} - \dfrac{k_{A}}{\delta_{A}} + \dfrac{d^{2}}{2\delta_{D}\delta_{A}} - \dfrac{d}{\delta_{A}}
-\end{equation*}
-The derivative for a fixed $\delta_{D}$ is as follows:
-\begin{equation*}
-\dfrac{\partial \beta_{A}(\delta_{D},\delta_{A})}{\partial \delta_{A}} = -\dfrac{\delta_{D}}{2\delta_{A}^{2}} + \dfrac{k_{A}}{\delta_{A}^{2}} - \dfrac{d^{2}}{2\delta_{D}\delta_{A}^{2}} + \dfrac{d}{\delta_{A}^{2}} = \dfrac{-\delta_{D}^{2} - d^{2} + 2\delta_{D}d + 2\delta_{D}k_{A}}{2\delta_{A}^{2}\delta_{D}}
-\end{equation*}
-The stationary points (maximum, minimum) can be found by setting the first derivative equal to zero and finding the roots of the resulting equation:
-\begin{equation*}
-\frac{\partial \beta_{A}(\delta_{D},\delta_{A})}{\partial \delta_{D}} =0 ~~~~~~ =>~~~~~~  \delta_{D}= \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}
-\end{equation*}
-
-It follows that $\beta_{A}(\delta_{D},\cdot)$ is increasing and non-positive if $\delta_{D}> \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}$ and decreasing and positive if $\delta_{D} < \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}$  \\
-
-\subsection*{Case B: $\delta_{D} \geq d+\delta_{A} \geq \delta_{A} $ }
-The benefit formula obtained from the previous chapter \ref{benfcase2a:attacker} for this case is as follows:
+\subsection*{Case B: $ \delta_{A} \leq \delta_{D} -d$ }
+The benefit formula for the attacker for this case obtained from the previous chapter is as follows:
 \begin{equation*}
@@
 %\begin{equation*}
-%\beta_{A}(\delta_{D},\delta_{A}) = 1 - \dfrac{•}{•}\delta_{A}}{2\delta_{D}} - \dfrac{d}{\delta_{A}} - \dfrac{k_{A}}{\delta_{A}} + \dfrac{(d-(\delta_{D}-\delta_{A})^{2}}{2\delta_{D}\delta_{A}} 
+%\beta_{A}(\delta_{D},\delta_{A}) = 1 - \delta_{A}}{2\delta_{D}} - \dfrac{d}{\delta_{A}} - \dfrac{k_{A}}{\delta_{A}} + \dfrac{(d-(\delta_{D}-\delta_{A})^{2}}{2\delta_{D}\delta_{A}} 
 %\end{equation*}
@@
 
+\subsection*{Case C: $d \leq \delta_{D} -d \leq \delta_{A} $}
+
+The benefit formula for the attacker for this case obtained in the previous chapter is as follows:
+\begin{equation*}
+\beta_{A}(\delta_{D},\delta_{A}) =\dfrac{\delta_{D}}{2\delta_{A}} - \dfrac{k_{A}}{\delta_{A}} + \dfrac{d^{2}}{2\delta_{D}\delta_{A}} - \dfrac{d}{\delta_{A}}
+\end{equation*}
+The derivative for a fixed $\delta_{D}$ is as follows:
+\begin{equation*}
+\dfrac{\partial \beta_{A}(\delta_{D},\delta_{A})}{\partial \delta_{A}} = -\dfrac{\delta_{D}}{2\delta_{A}^{2}} + \dfrac{k_{A}}{\delta_{A}^{2}} - \dfrac{d^{2}}{2\delta_{D}\delta_{A}^{2}} + \dfrac{d}{\delta_{A}^{2}} = \dfrac{-\delta_{D}^{2} - d^{2} + 2\delta_{D}d + 2\delta_{D}k_{A}}{2\delta_{A}^{2}\delta_{D}}
+\end{equation*}
+The stationary points (maximum, minimum) can be found by setting the first derivative equal to zero and finding the roots of the resulting equation:
+\begin{equation*}
+\frac{\partial \beta_{A}(\delta_{D},\delta_{A})}{\partial \delta_{D}} =0 ~~~~~~ =>~~~~~~  \delta_{D}^{2}-\delta_{D}(2k_{A}-2d) + d^{2}= 0
+\end{equation*}
+
+The roots of this quadratic equation are $\delta_{D}=d+k_{A}\pm \sqrt{2dk_{A}+k_{A}^{2}}$.
+Because for this case $d\leq \delta_{D}$, we only look at root $\delta_{D}=d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$.
+$\sqrt{2dk_{A}+k_{A}^{2}}$ is bigger than $k_{A}$ so if we subtract it from $d+k_{A}$, we are in the part where $delta_{D}$ is smaller than d.
+\\
+It follows that $\beta_{A}(\delta_{D},\delta_{A})$ is increasing and non-positive if $\delta_{D}> d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$ and decreasing and positive if $\delta_{D} < d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$  \\
+
 \section{Best responses of $\delta_{A}$}
-The optimum functions will be piecewise functions. We distinguish three cases for different values of $\delta_{D}$ regarding $d$ and $k_{A}$.
-
-For this term $\dfrac{(\delta_{D}-d)^{2}}{2k_{A}} $ , $d$ has to be bigger than  $\delta_{D}$ because the cost $\delta_{D}$ cannot be negative.
+The optimum function of the attacker is found in the same way as the optimum function of the defender.
+The derivatives of each piece of the piecewise benefit function of the attacker is taken.
+From these derivatives the best responses are derived.
+\\
+Case A is an edge case and will be addressed in the paragraph about the edge cases.
+%The optimum functions will be piecewise functions. We distinguish three cases for different values of $\delta_{D}$ regarding $d$ and $k_{A}$. 
+%
+%
+%For this term $\dfrac{(\delta_{D}-d)^{2}}{2k_{A}} $ , $d$ has to be bigger than  $\delta_{D}$ because the cost $\delta_{D}$ cannot be negative.
 This was an assumption that was already made, because the benefit of the defender will always be 1 if $d$ is bigger than  $\delta_{D}$.
 
-\subsubsection*{$\delta_{D} < \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}$} 
-From case 1 and case 2.b above, it follows that the benefit is decreasing but positive, which means that the attacker will play as fast as he can.
+\subsubsection*{$\delta_{D} < d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$} 
+For case B, the function is increasing.
+The function has a maximum on on $\delta_{A} = minimum \{\delta_{D} -d, \sqrt{2k_{A}\delta_{D}} \} $.
+For $\delta_{D} < d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$, the minimum of the two values is $\delta_{D}-d$.\\
+For case C, if $\delta_{D} < d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$, the benefit function is increasing.
+Because we have a continuous function and the benefit is still increasing and negative in case C, the defender will try not to play.
+The defender's optimum benefit is achieved in $\delta_{A}=\infty$.
+See (A) in figure \ref{case1A}.
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[scale=1]{Images/case1A.pdf}
+\caption{ Illustration of the best response functions.
+The left functions are the piecewise benefit function of the attacker for a certain values of $k_{A},d$ and $\delta_{D}$.
+The right functions are the derivatives of the benefit functions on the left.
+The best responses are the values where the function intersects with the x-as.
+(A): $\delta_{D} > d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$, (B): $\delta_{D} = d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$, (C): $\delta_{D} < d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$}
+\label{case1A}
+\end{figure}
+
+%From case 1 and case 2.b above, it follows that the benefit is decreasing but positive, which means that the attacker will play as fast as he can.
 For case 2.a  $\delta_{D} \geq \delta_{A}$ is increasing.
 It follows that for this case the attacker's maximum benefit is achieved for $\delta_{A} = \sqrt{d k_{A}\delta_{D}}$.
-. In this case the attacker's maximum benefit is reached at $\delta_{A} = \delta_{A} = \sqrt{dk_{A}\delta_{D}}$  .
-\subsubsection*{$\delta_{D} = \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}$} 
-The benefit $\beta_{A}(\delta_{D},\delta_{A})=0$ for case 1 and case 2.b for all $\delta_{D}$ in $[0, \infty [$.
+%. In this case the attacker's maximum benefit is reached at $\delta_{A} = \delta_{A} = \sqrt{dk_{A}\delta_{D}}$  .
+\subsubsection*{$\delta_{D} = d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$} 
+For case B, the function is increasing.
+The function has a maximum on on $\delta_{A} = minimum \{\delta_{D} -d, \sqrt{2k_{A}\delta_{D}} \} $.
+For $\delta_{D} = d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$, the minimum of the two values is $\delta_{D}-d$.\\
+For case C, $\delta_{D} = d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$ is a root so the benefit function of the attacker will be equal to zero.
+Because case C is the case where $\delta_{A}$ has to be bigger or equal to $\delta_{D}-d$, $\beta_{A}=0$ is valid for all $\delta_{A} \in [\delta_{D}-d,\infty]$.\\
+Putting the pieces together gives us the optimum benefit for the attacker for all $\delta_{A} \in [\delta_{D}-d,\infty]$. See (B) in figure \ref{case1A}.
+
+%The benefit $\beta_{A}(\delta_{D},\delta_{A})=0$ for case 1 and case 2.b for all $\delta_{D}$ in $[0, \infty [$.
 For case 2.a it follows that the benefit is increasing.
 The attacker's maximum benefit is achieved at any $\delta_{A} $ in
-\subsubsection*{$\delta_{D} > \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}$} 
-For case 1 and case 2.b the benefit is increasing but non-positive. This means that the attacker won't play ($\delta_{A} = \infty$). 
-\\
-~~\\
+\subsubsection*{$\delta_{D} > d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$} 
+For case B, the function is increasing.
+The function has a maximum on on $\delta_{A} = minimum \{\delta_{D} -d, \sqrt{2k_{A}\delta_{D}} \} $.
+For $\delta_{D} > d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$, the minimum of the two values is $\delta_{D}-d$.\\
+For case C, if $\delta_{D} > d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$ the function is decreasing.
+So the optimum benefit for all the cases is when $\delta_{D} = \delta_{D}-d$.
+See (C) in figure \ref{case1A}.\\
+%For case 1 and case 2.b the benefit is increasing but non-positive. This means that the attacker won't play ($\delta_{A} = \infty$).
+
 From this analyses we can compute $opt_{A}(\delta_{D})$ :
@@
      \begin{array}{lr}
-            \infty & \delta_{D} < \dfrac{(\delta_{D}-d)^{2}}{2k_{A}} \\
-       \left[ \sqrt{d k_{A}\delta_{D}}, \infty\right],  & \delta_{D} = \dfrac{(\delta_{D}-d)^{2}}{2k_{A}} \\
-       \sqrt{d k_{A}\delta_{D}}, & \delta_{D} > \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}
+            \infty & \delta_{D} < d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}} \\
+       \left[ \delta_{D}-d, \infty\right],  & \delta_{D} = d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}} \\
+    	\delta_{D}-d, & \delta_{D} > d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}
      \end{array}
@@
 \subsection{Conclusion}
-With the edge cases we can compute $opt_{A}(\delta_{D})$ : \\
+Including the edge cases yields the following for $opt_{A}(\delta_{D})$ : \\
 

### 2015-08-19 doc/template/chap-10.tex
veranderingen opslagen

@@
 This paper builds on a relatively new paper where the assumption of stealthiness is fairly unique, giving some interesting results.\\
- The paper is from researchers at RSA, van Dijk et al,  who presented a game-theoretic framework to model computer security scenarios called ``FlipIt'' \cite{FlipIt}.
+ The paper is from researchers at RSA, van Dijk et al.,  who presented a game-theoretic framework to model computer security scenarios called ``FlipIt'' \cite{FlipIt}.
 They study the specific scenario where a system or network is repeatedly taken over completely by an attacker.
@@
 %adapted optimum functions for the defender and the attacker. These adapted optimum functions yield a number of interesting cases.
-\item[-] Based on our results we give practical recommendations to take measures against advanced attacks.
+%\item[-] Based on our results we give practical recommendations to take measures against advanced attacks.
 The modelling of FlipIt with a propagation delay will give a guideline for both the attacker and the defender on how to respectively plan attacks or network clean-ups.
@@
 While it may seem trivial to extend the basic FlipIt model with a propagation delay, its mathematical treatment is not.
-In the paper of Laszka et all.
+In the paper of Laszka et al.
 \citep{FlipThem} it seems that even a small extension adds to the already significant mathematical complexity.
@@
 The organisation of this paper is as follows.
-An introduction to cyber security and game theory is given in chapter \ref{Chapter1:Intro.Game.Theory} to allow the reader to become familiar with the kind of threats that are in the scope of this work and the game theoretic concepts that will be further used in the paper.
+An introduction to cyber security and game theory is given in chapter \ref{Chapter1:Intro.Game.Theory}.
+The chapter allows the reader to become familiar with the kind of threats that are in the scope of this work and the game theoretic concepts that will be further used in the paper.
 In the same chapter the FlipIt framework is summarized with its most important conclusions.

### 2015-08-19 doc/template/chap-5.tex
veranderingen opslagen

@@
 Section 2.2 then introduces the reader into the main principles of game theory.
-Subsequently, section 2.3 introduces the reader to the FlipIt game: this game will be used to model cyber security attacks of a periodic nature, including a delay.
+Subsequently, section 2.3 introduces the FlipIt game: this game will be used to model cyber security attacks of a periodic nature, including a delay.
 Finally, section 2.4 gives an overview of the related work, and how the research presented in this paper compares to existing results.
@@
 While cyber security and network security address different aspects, they partially overlap as well.
-For example, a virus can be physically dropped on a computer network using a USB stick, but it can also arrive over the internet.
+For example, a virus can be physically dropped on a computer network using a USB stick, but it can also arrive over the Internet.
 Either way internal computer security measurements have to be taken to recover from the virus.
@@
 dynamic whitelisting.
-The remaining 1\% are the advanced threats, also known as APT.
+The remaining 1\% are the advanced threats, also known as APTs.
 They are the most difficult to deal with.
@@
 
-Games can be divided into two types of games: cooperative or non-cooperative. In \textit{non cooperative games}, the basic modelling unit is the group of agents. 
+Games can be divided into two types of games: cooperative or non-cooperative. In non-cooperative games, the basic modelling unit is the group of agents. 
 Two or more agents want to maximise their utility and their actions can be affected by other agents' utilities.
 In the individualistic approach the basic modelling is only one agent.
-This type of game is referred to as \textit{cooperative games}.
+This type of game is referred to as cooperative games.
+The game modelled in this paper belongs to type of games that are non-cooperative.
 
@@
 Over John Nash is een prachtige film
-\subsubsection{Best Response and Nash Equilibrium}
+\subsubsection{Best response and Nash equilibrium}
 One of the solution concepts in game theory for non-cooperative games that will be used in this paper is a Nash equilibrium.
-A Nash Equilibrium is a subset of outcomes that can be interesting to analyse a game.
+A Nash equilibrium is a subset of outcomes that can be interesting to analyse a game.
 To define this concept we first introduce the concept of best response.
@@
 The best response for player \textit{1} is: $a_{1} = Opt_{1}(a_{2})$, given that $a_{2}$ is an action of player 2.
-For a Nash Equilibrium each player has a list of actions and each player's action maximises his or her pay-off given the actions of the other players.
+For a Nash equilibrium each player has a list of actions and each player's action maximises his or her pay-off given the actions of the other players.
 Nobody has the incentive to independently or unilaterally change his or her action if an equilibrium profile is played.
-We have a Nash Equilibrium for the pair $(a_{1}^{*},a_{2}^{*})$ when $a_{1}^{*} = Opt_{1}(a_{2}^{*})$ and $a_{2}^{*} = Opt_{2}(a_{1}^{*})$.\\
+We have a Nash equilibrium for the pair $(a_{1}^{*},a_{2}^{*})$ when $a_{1}^{*} = Opt_{1}(a_{2}^{*})$ and $a_{2}^{*} = Opt_{2}(a_{1}^{*})$.\\
 
@@
 Each of them can betray the other or they can support each other and remain silent.
-If a player talks, he will either be imprisoned for 3 years or go free, depending on whether the other talks.
+If a player talks (betray the other prisoner), he will either be imprisoned for 3 years or go free, depending on whether the other talks.
 If the player is silent, he will either be imprisoned for 5 years or one year, depending on whether the other talks.
@@
 Figure \ref{Prison} shows the rewards of the possible actions of the two players.
-The figure shows that it is advantageous for each player to talk (betraying the other).
+The figure shows that it is advantageous for each player to talk.
 If prisoner 1 talks and the other prisoner remains silent, prisoner 1 will be free.
@@
 It is a game between two perfectly rational prisoners who do not know what the other one will do.
-Each prisoner can talk (betraying the other) or remain silent.
+Each prisoner can talk (betray the other prisoner) or remain silent.
 The resulting sentence is shown below each prisoner.
@@
 
-FlipIt is a two-players game with a shared single resource that the players want to control as much as possible (figure \ref{fig:FLipItDefault}).
+FlipIt is a two-player game with a shared single resource that the players want to control as much as possible (figure \ref{fig:FLipItDefault}).
 The shared resource can be a password, a network or a secret key depending on the setting being modelled.
@@
 The time in the game is assumed to be continuous.
-To get control over the resource, the players $i$, with $i \in \{A,D\}$, can flip the resource at any given time.
+To get control over the resource, the players $i$, with $i \in \{A,D\}$, can move (flipping the resource) at any given time.
 %A flip will be regarded as a move from a player \textit{i}. 
@@
 The goal of the player is to maximise the time that he or she has control over the resource while minimizing the total cost of the moves.
-A move can also result in a "wasted move", called a flop.
+A move can also result in a ``wasted move'', called a flop.
 It may happen that the resource was already under control of the player.
@@
 Every move or flip is indicated by a green (dark grey) or orange (light grey) circle.
-The attacker is represented in orange and the defender is represented in green.
-The blue and orange rectangles represent which player is in control of the resource.}
+The defender is represented in green and the attacker is represented in orange.
+The green and orange rectangles represent which player is in control of the resource.}
 \label{fig:FLipItDefault}
@@
 
-If there is no feedback for either player, he will play in the same manner against every opponent.
+If there is no feedback for either player, the player will play in the same manner against every opponent.
 The strategy is then called \textit{non-adaptive} because the playing strategy is not dependent on the opponent's movements.
@@
 If the players receive feedback, a player can adapt his strategy to the information received about the opponent's moves.
-This strategy is called \textit{adaptive} strategies.
+This strategy is called an \textit{adaptive} strategy.
 Depending on the amount of information received, two subclasses of adaptive strategies can be identified.
@@
 \subsubsection{Results of the FlipIt game}
-The study of the different strategies by means of FlipIt framework allows to derive a number of interesting results \cite{FlipIt}:  
+The study of the different strategies by means of the FlipIt framework allows to derive a number of interesting results \cite{FlipIt}:  
 \begin{itemize}
-\item periodic strategy strongly dominates the other renewal strategies if the opponent has a periodic or non-arithmetic renewal strategy.
+\item a periodic strategy strongly dominates the other renewal strategies if the opponent has a periodic or non-arithmetic renewal strategy.
 This means that it is a good choice for the opponent to play periodically against a player with a non-adaptive strategy;
-\item periodic games are disadvantageous against players following a Last Move adaptive strategy.
+\item periodic games are disadvantageous against players following a LM adaptive strategy.
 The opponent can observe the exact time of the player's next move and play immediately afterwards.
@@
 Laszka et al.
-also use two ways to flip the multiple resources: the AND and the OR control model.
+uses two ways to flip the multiple resources: the AND and the OR control model.
 In the AND model the attacker only controls the system if he controls all the resources of the system, whereas in the OR model the attacker only needs to compromise one resource to be in control of the entire system.
@@
 
-Another addition of Laszka et al. to the game of FlipIt \cite{MitigationCovert} 
+Another addition of Laszka et al. \cite{MitigationCovert}  to the game of FlipIt 
 is extending the game to also consider non-targeted attacks by non-strategic players.
@@
 Our motivation for a defender with stealthy moves is that it can not be assumed that every attacker can receive feedback from an APT.
-Some ATP's are designed only to cause harm on the systems that are infected e.g.
+Some ATPs are designed only to cause harm on the systems that are infected e.g.
 the Stuxnetworm that was developed to destroy nuclear reactors.
@@
 The third and last difference is that the paper of Laszka has multiple attackers who try to find the best strategy of the defender against both targeted and non-targeted attacks.
-The conclusion of this paper is that the optimal strategy for the defender is moving periodically.
+The conclusion of Laszka's paper is that the optimal strategy for the defender is moving periodically.
 \\

### 2015-08-19 doc/template/chap-9.tex
veranderingen opslagen

@@
-\chapter{FlipIt with propagation delay}
+\chapter{FlipIt with Propagation Delay}
 \label{chapter2:FlipIt with virus propagation}
@@
 
-\section{Difference between FlipIt with and without propagation delay}
+\section{Difference between FlipIt with and without Propagation Delay}
 \label{ch2:diffFlip}
@@
 This extension has already been investigated in the context of installing anti-virus systems on specific nodes in a graph and investigating how this affects the spreading of malware.
-In chapter \ref{chapter4: Worm propagation}, when discussing future work, the use of a matrix model is discussed to determine the probability that a node is infected by malware.  \\
+In chapter \ref{chapter4: Worm propagation}, when discussing future work, the use of a matrix model is discussed to determine the probability that a node is infected by malware.
+A practical example of a flip of the defender may consist of patching, formatting, disk-wiping or unplugging a computer on the network.
+\\
 The attacker only has one action: sending different kinds of malware to a computer network.
-Every time the attacker flips the resource, he does this by sending a new kind of malware e.g worm to the system.
-The attacker does not send the same malware again, because if the computers are patched, the malware will be unable to penetrate the system.
+Every time the attacker flips the resource, he does this by sending a new kind of malware e.g virus to the system.
+The attacker does not send the same malware again, because if the action of the defender, e.g.
+consists of patching, the malware will be unable to penetrate the system.
 The node can be a targeted node or a random node.
@@
 Some APTs can be resident on a isolated network (e.g.
-the Stuxnetworm), meaning that there is no way to connect to the Internet.
+the \textit{Stuxnetworm}), meaning that there is no way to connect to the Internet.
 \\
@@
 Honeypots emulate services or creates multiple instances of real operating systems and can pretend to have sensitive information.
-The case where the moves of the defender are non stealthy has already been investigated by Laszka \citep{MitigationNonTargeted}.
+While the analysis in this paper is limited to immediate effect of the defender's moves, the case where the moves of the defender are non stealthy has already been investigated by Laszka \citep{MitigationNonTargeted}.
 
@@
 This is motivated by the fact that the identity of the attackers and their attack strategy is rarely known to the defenders.
-In the case of the attacker, if an attacker wants to have feedback it may be disadvantageous if he wants to stay undetected.
-If the attacker wants to have feedback, the malware needs to make a connection with the attacker to provide this feedback.
-For example, this can cause more network traffic for the defender to detect.
+In the case of the attacker, if an attacker wants to have feedback, receiving feedback may compromise the desire to stay undetected.
+To be able to get feedback, the malware needs to make a connection with the attacker to provide this feedback.
+This may for example cause extra network traffic, which can be detected by the defender.
 \\
@@
 
-\section{Formalization of the periodic game with propagation delay}
+\section{Formalization of the Periodic Game with Propagation Delay}
 \label{ch2:periodicvirus}
@@
 \includegraphics[scale=0.7]{Images/DefFlip.pdf}
-\caption{Formalization of a FlipIt game with delay: A representation of a FlipIt game where both players are playing periodically.
+\caption{Formalization of a FlipIt game with  propagation delay: A representation of a FlipIt game where both players are playing periodically.
 Every move or flip is indicated by a green or orange circle, respectively dark gray and light grey.
-The defender is represented in green (dark grey) and plays with a period of $\delta_{D}$.
-The flip of the attacker is represented by a white circle, but because there is a delay d, the attacker only controls the resource after time d represented by an orange circle (light grey).
+The defender is represented in green and plays with a period of $\delta_{D}$.
+The flip of the attacker is represented by a white circle, but because there is a delay d, the attacker only controls the resource after time d represented by an orange circle.
 The attacker plays with a period of $\delta_{A}$.
@@
 \item $i$: Denotes the player.
-\textit{D} denotes the defender, and \textit{A} denotes the attacker which differs form the notation in \citep{FlipIt}, where the defender is denoted by the subscript \textit{0} and the attacker by the subscript \textit{1}.
+\textit{D} denotes the defender, and \textit{A} denotes the attacker which differs from the notation in \citep{FlipIt}, where the defender is denoted by the subscript \textit{0} and the attacker by the subscript \textit{1}.
 \item $\delta_{i}$: The length of the interval between two consecutive moves of player \textit{i}. 
@@
 If the defender flips the network before the period $d$ has elapsed (so, somewhere between $t$ and $t + d$), then the attacker will never gain full control over the resource.
-(See figure \ref{dt} at point 3 and 4).
+(See figure \ref{dt} at points 3 and 4).
 This implies that the mathematical formulas for gain and benefit need to be adapted to the fact that the attacker loses part of his benefit because of this delay.
@@
 \\
-
+We adapt the same definition for players that play at the same time as in the original FlipIt game.
+The probability that this happens will be close to zero, but if it happens the moves cancel each other out and no change of control happens (see figure \ref{dt} point 3).\\
 \begin{figure}[hbtp]
@@
 During the first flip of the attacker, the defender moves after the delay (point 2), resulting in the attacker gaining control over the resource, until the defender flips.
-During the second flip of the attacker, the defender flips at time t+d, causing the defender to retain control over the resource before the attacker.
+During the second flip of the attacker, the defender flips at time t+d, causing the defender to retain control over the resource before the attacker can gain control.
 During the third and final flip of the attacker, the defender flips during time t+d, causing the attacker to never gain control over the resource.
@@
 We split the formalization in three cases.
-Similarly as in \cite{FlipIt}, there is the case where the defender plays at least as fast as the attacker, and the case where the attacker plays at least as fast as the defender.
+Similarly as in the original FlipIt game, there is the case where the defender plays at least as fast as the attacker, and the case where the attacker plays at least as fast as the defender.
 For each of these cases, the benefit formula of the basic case without delay is presented first, and subsequently the delay is introduced.
@@
 
-In the original model of FlipIt \cite{FlipIt}, the authors express the formulas in terms of $\alpha_{D}$ and $\alpha_{A}$.
+In the original model of FlipIt, the authors express the formulas in terms of $\alpha_{D}$ and $\alpha_{A}$.
 However, when introducing the delay, some formulas become much simpler when expressed in terms of $\delta_{D}$ and $\delta_{A}$.
@@
 The reason this formula overestimates the benefit of the attacker is that it assumes that the defender is always in control during the delay.
-However, if the attacker was in control in the previous interval, then he continues to be in control during the period of the delay, see figure \ref{fig:case2} situation 5 to 7.
+However, if the attacker was in control in the previous interval, then he continues to be in control during the period of the delay, see figure \ref{fig:case2} situations 5 to 7.
 This means that the average benefit formulas for this case cannot be derived from one interval only; what happens in the previous interval must be taken into account.
@@
 Because $d + \delta_{A} \leq \delta_{D}$ the next move of the defender in this second interval will never occur during the delay, meaning that the entire delay can be considered as an extra benefit resulting of a play in the previous interval.
-So, every time the defender plays, he will get an average gain of $\dfrac{\delta_{A}}{2}$ in the interval where he plays and in the next interval will always receive a extra gain of $d$, yielding a total average gain per interval of
+So, every time the defender plays, he will get an average gain of $\dfrac{\delta_{A}}{2}$ in the interval where he plays and in the next interval will always receive an extra gain of $d$, yielding a total average gain per interval of
 $\dfrac{(d+\dfrac{\delta_{A}}{2})}{\delta_{A}}$
 
-For the remainder of the analyses we consider two cases, Case a and Case b, depending on whether the delay is shorter or longer than the difference between the attacker's and the defender's period.
+For the remainder of the analyses we consider two cases, Case 2.a and Case 2.b, depending on whether the delay is shorter or longer than the difference between the attacker's and the defender's period.
 \\
@@
 \includegraphics[scale=0.7]{../../doc/template/Images/FlipIt2.pdf}
-\caption{All possible cases for the attacker and the defender in Case 2.A where $d + \delta_{A} < \delta_{D}$.
+\caption{All possible cases for the attacker and the defender in Case 2.A where $d + \delta_{A} \leq \delta_{D}$.
 As can be seen in cases (1) to (4), the defender will have control during a period of \textit{d} over the resource in the next interval when the defender has flipped in the previous interval.}
@@
 The above benefit formula would then include too much gain for the defender: the potential overlap during the delay would be counted twice.
-See figure \ref{countedtwice}\\
+See figure \ref{countedtwice}.\\
 
@@
 
-We calculated the benefit for three cases: when $\delta_{D}$ is smaller than $d$. When $\delta_{A}$ is smaller than $\delta_{A}$ and when it is larger than $\delta_{A}$.
+We calculated the benefit for three cases: when $\delta_{D}$ is smaller than $d$, when $\delta_{D}$ is smaller than $\delta_{A}$ and when $\delta_{D}$ is larger than $\delta_{A}$.
 The benefit formulas of case 1 and case 2.b result in the same formula.
@@
 The first figure in figure \ref{Borders} shows all the cases and it can be seen that the border condition of case 1 and case 2.b can be brought together.
-This can be seen as the second figure in \ref{endborders}.
+This can be seen as the second figure in \ref{Borders}.
 \\

### 2015-08-19 doc/template/conclusion.tex
veranderingen opslagen

@@
 
-This paper presents an adaptation to the basic FlipIt game by \citep{FlipIt} to model an attacker with a delay.
+This paper presented an adaptation to the basic FlipIt game by \citep{FlipIt} to model an attacker with a delay.
 
-\section{General results and conclusions}
+\section{General Results and Conclusions}
 Our FlipIt model with worm propagation delay showed us some interesting results. 
@@
 In the next section we list a set of extensions that might be interesting for further research.
-\section{Further work}
+\section{Further Work}
 
@@
 An effective way to counter this problem is to raise awareness about risks of infections amongst the employees.
-Even in the presence of good security protection layers, spam filters, and firewalls, the opening of a phishing mail or use of an unknown USB stick by one of your employees may be suficient to infect a PC on the network, from which the virus can spread further.
+Even in the presence of good security protection layers, spam filters, and firewalls, the opening of a phishing mail or use of an unknown USB stick by one of your employees may be sufficient to infect a PC on the network, from which the virus can spread further.
 A system can however never be 100\% waterproof.

### 2015-08-19 doc/template/nash.tex
veranderingen opslagen

@@
-\chapter{Optimal defence and attack strategies}
+\chapter{Optimal Defence and Attack Strategies}
 \label{chapter:Nash}
@@
 
-\section{Determining the piecewise functions $opt_{D}(\delta_{A})$}
+\section{Determining the Piecewise Functions $opt_{D}(\delta_{A})$}
 
@@
 
-\section{Best responses of $\delta_{D}$}
+\subsection{Best Responses of $\delta_{D}$}
 The optimum functions will be piecewise functions. We distinguish six cases for different values of $\delta_{A}$ depending on the values of $k_{D}$ and $d$. 
@@
 %*********************************************************************
-\section{Determining the piecewise functions $opt_{A}(\delta_{D})$}
+\section{Determining the Piecewise Functions $opt_{A}(\delta_{D})$}
 %To start with we only consider the case where $d < \delta_{D}$, because if $d > \delta_{D}$ the benefit of the defender \todo{nakijken, def of att} is always 1. \\
@@
 \\
-It follows that $\beta_{A}(\delta_{D},\delta_{A})$ is increasing and non-positive if $\delta_{D}> d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$ and decreasing and positive if $\delta_{D} < d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$  \\
+It follows that $\beta_{A}(\delta_{D},\delta_{A})$ is increasing and non-positive if $\delta_{D}> d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$ and decreasing and positive if $\delta_{D} < d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$.
+\\
 
-\section{Best responses of $\delta_{A}$}
+\subsection{Best Responses of $\delta_{A}$}
 The optimum function of the attacker is found in the same way as the optimum function of the defender.
@@
 The best responses are the values where the function intersects with the x-as.
-(A): $\delta_{D} > d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$, (B): $\delta_{D} = d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$, (C): $\delta_{D} < d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$}
+(A): $\delta_{D} > d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$, (B): $\delta_{D} = d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$, (C): $\delta_{D} < d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$}.
 \label{case1A}
@@
           \left[0,\delta_{D} - d\right] & \delta_{D} > d ~~\& ~~k_{A}=0 \\
-                 \infty & \delta_{D} < \dfrac{(\delta_{D}-d)^{2}}{2k_{A}} \\
-       \left[ \sqrt{d k_{A}\delta_{D}}, \infty\right],  & \delta_{D} = \dfrac{(\delta_{D}-d)^{2}}{2k_{A}} \\
-       \sqrt{d k_{A}\delta_{D}}, & \delta_{D} > \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}
+            \infty & \delta_{D} < d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}} \\
+       \left[ \delta_{D}-d, \infty\right],  & \delta_{D} = d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}} \\
+    	\delta_{D}-d, & \delta_{D} > d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}
      \end{array}
@@
 
-The piecewise functions $opt_{D}$ and $opt_{A}$ give some interesting results. \todo{resultaten nog toevoegen, zelfde als in de andere conclusie}
+The piecewise functions $opt_{D}$ and $opt_{A}$ give some interesting results.
+
+\begin{itemize}
+\item When the defender plays faster than the delay, the attacker will either have a negative benefit or a benefit equal to zero if the flips do not cost anything.
+It is for the defender a target to be able to play at a rate smaller or equal to the delay.
+\item If the defender can play with a cost equal to zero, the attacker will not play. The defender can play at any rate he wants, and the attacker is always disadvantaged by his delay.
+\item If the cost of the defender is non-zero, then, depending on the ratio between it's cost and the delay, from a certain value for the speed of the attacker, the defender will not play.
+The same thing goes for the attacker.
+This result is similar to the results obtained for the original FlipIt game.
+\end{itemize}
 
@@
 For this we have to compare the function in terms of the relationship of the players move cost.
-We distinguish three cases: $k_{A} < k_{D} , ~k_{A} > k_{D} $ and $k_{A} = k_{D}$.
+%We distinguish three cases: $k_{A} < k_{D} , ~k_{A} > k_{D} $ and $k_{A} = k_{D}$.

### 2015-08-19 doc/template/chap-4.tex
veranderingen opslagen
finaal

@@
 
-\section{Methods of propagation}
+\section{Methods of Propagation}
 \label{methodsofpropagation}
@@
 
-\section{Models for worm propagation}
+\section{Models for Worm Propagation}
 \label{modelsforpropagation}
@@
 
-\section{Matrix worm model}
+\section{Matrix based Worm Model}
 \label{eigenmatrixmethode}
@@
 %\end{verbatim}
-\section{PageRank algorithm}
+\section{PageRank Algorithm}
 \label{googlesec:5.4}
@@
 The Google Matrix is a stochastic matrix and represents a graph where every edge denotes a link between pages.
-Through the power iteration  \ref{Arnoldi}, the PageRank of all the pages can be calculated iteratively.
+Through the power iteration \citep{Arnoldi}, the PageRank of all the pages can be calculated iteratively.
 \\
@@
 It can be advantageous to flip the nodes with a higher PageRank more often than the other nodes.
-A higher PageRank means that the page has a higher probability of being compromised by a worm.
+A higher PageRank means that the node has a higher probability of being compromised by a worm.

### 2015-08-19 doc/template/chap-5.tex
finaal

@@
 Some ATPs are designed only to cause harm on the systems that are infected e.g.
-the Stuxnetworm that was developed to destroy nuclear reactors.
-The Stuxnetworm was resident on a isolated network, meaning that there was no way to connect to the internet.
+the \textit{Stuxnetworm} that was developed to destroy nuclear reactors.
+The \textit{Stuxnetworm} was resident on a isolated network, meaning that there was no way to connect to the internet.
 Another example is the use of honeypots.

### 2015-08-19 doc/template/chap-9.tex
finaal

@@
 \\
-We adapt the same definition for players that play at the same time as in the original FlipIt game.
-The probability that this happens will be close to zero, but if it happens the moves cancel each other out and no change of control happens (see figure \ref{dt} point 3).\\
+
+We apply the same definition for players that play at the same time as in the original FlipIt game.
+As the time in the game is continuous, the probability of this happening is close to zero, but if it happens the moves cancel each other out and no change of control happens (see figure \ref{dt} point 3).\\
 \begin{figure}[hbtp]
@@
 %
-We split the formalization in three cases.
+The formalization is split three cases.
 Similarly as in the original FlipIt game, there is the case where the defender plays at least as fast as the attacker, and the case where the attacker plays at least as fast as the defender.
@@
 
-We calculated the benefit for three cases: when $\delta_{D}$ is smaller than $d$, when $\delta_{D}$ is smaller than $\delta_{A}$ and when $\delta_{D}$ is larger than $\delta_{A}$.
+\begin{figure}[hbtp]
+\centering
+\includegraphics[scale=0.7]{Images/borders.pdf}
+\caption{The first figure is the representation of the cases where the benefit functions are calculated.
+The second figure is the representation of the new domain of piece B which is a merge of case 2.b and case 1.}
+\label{Borders}
+\end{figure}
+We calculated the benefit for three cases: when $\delta_{D}$ is smaller or equal to $d$, when $\delta_{D}$ is smaller or equal to $\delta_{A}$ and when $\delta_{D}$ is larger or equal to $\delta_{A}$.
 The benefit formulas of case 1 and case 2.b result in the same formula.
 Consequently, these two formulas can be brought together by adapting the border conditions.
-The first figure in figure \ref{Borders} shows all the cases and it can be seen that the border condition of case 1 and case 2.b can be brought together.
+The first figure in figure \ref{Borders} shows all the cases and it can be seen that the domain of case 1 and case 2.b can be brought together.
 This can be seen as the second figure in \ref{Borders}.
@@
 The benefit functions can be represented as piecewise functions for each case, with case 1 and case 2.b together:\\
-The benefit formula for the defender, see figure \ref{pfb} :
-\begin{displaymath}
+The benefit formula for the defender is as follows:
+\begin{displaymath}\label{PWFD}
 \beta_{D}(\delta_{D},\delta_{A})= \left\{
@@
 
-The benefit formula for the attacker, see figure \ref{pfa}:
+The benefit formula for the attacker is as follows: 
 \begin{displaymath}
@@
 
+To be able to show the continuity of the benefit function, piece 2 and piece 3 are switched and the boder conditions are reformulated.
+
+\begin{displaymath}\label{PWFA}
+\beta_{D}(\delta_{D},\delta_{A})= \left\{
+\begin{array}{lll}
+	1 - \dfrac{k_{D}}{\delta_{D}}, &~~~~~ & \delta_{D} \leq d \\
+	~ & ~ \\
+	\dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - \dfrac{k_{D}}{\delta_{D}} &~~~~~& \delta_{A}  \leq \delta_{D} -d \\
+	   ~ & ~ \\
+   1 - \dfrac{(\delta_{D} - d)^2}{2\delta_{D}\delta_{A}} - \dfrac{k_{D}}{\delta_{D}}, & ~~~~~~&  \delta_{D} -d \leq \delta_{A}
+
+\end{array}
+\right.
+\end{displaymath}
+~~ \\
+
 \begin{figure}[hbtp]
 \centering
-\includegraphics[scale=1.4]{Images/benfa.pdf} 
-\caption{The benefit function of the attacker for a cost $k_{A} =1.2$, $d =0.5$ and $\delta_{D}=5$. The three pieces match with the three colours.}
+\includegraphics[scale=1.4]{Images/att.pdf} 
+\caption{The benefit function of the attacker for a cost $k_{A} =0.5$, $d =3$ and $\delta_{D}=7$.}
 \label{pfa}
@@
 
-The piecewise functions are also continuous functions. Filling in the border cases results in the same formula.  
-\begin{figure}[hbtp]
-\centering
-\includegraphics[scale=0.7]{Images/borders.pdf}
-\caption{The first figure is the representation of the cases where the benefit functions are calculated.
-The second figure is the representation of the new border conditions of piece B which is a merge of case 2.b and case 1.}
-\label{Borders}
-\end{figure}
+The piecewise functions are also continuous functions.
+Filling in the border cases results in the same formula.
+This can be seen on both figures \ref{pfb} and \ref{pfa} that give respectively a visual representation of the benefit function of the defender and the benefit function of the attacker.
+The thick pieces indicate the cases for which each piece of the benefit function holds.
+For the benefit of the attacker case B and case C correspond respectively with piece 2 and piece 3 of the reformulated benefit function of the attacker.
 

### 2015-08-19 doc/template/conclusion.tex
finaal

@@
 This result is similar to the results obtained for the original FlipIt game.
+\item The optimum strategy for the defender and the attacker is depending on the value of \textit{d}.
+In the beginning a non-adaptive strategy was assumed for both players, so no player gets feedback in the game.
+None of the players have knowledge about the real value of \textit{d}, so both players have to make a guess.
+By using the model matrix based model, different kind of attacks can be simulated and an average delay can be used.
 \end{itemize}

### 2015-08-19 doc/template/nash.tex
finaal

@@
 
-The benefit formula obtained in the previous chapter is as follows:
+The benefit formula for this case corresponds with the first piece of the benefit function of the defender obtained in the previous chapter:
 \begin{equation}
@@
 \subsection*{Case B: $ d \leq\delta_{D} \leq \delta_{A} + d$}
-The defender's benefit formula obtained in the previous chapter in this case is as follows:
+The benefit formula for this case corresponds with the second piece of the benefit function of the defender obtained in the previous chapter:
 \begin{equation*}
@@
 
-The benefit formula for the defender obtained in the previous chapter in this case is as follows:
+The benefit formula for this case corresponds with the third piece of the benefit function of the defender obtained in the previous chapter:
 \begin{equation*}
@@
 
-For $\delta_{A} + 2(d-k_{D}) > 0$ the benefit will be always positive but decreasing, see figure \ref{ShapeUp}. 
-The defender will always play as fast as he can if $\delta_{A} + 2(d-k_{D}) > 0$ for $k_{D} < d$ or $k_{D} > d$ because $\delta_{A}$ will be positive in either case.
-There is an edge case if $d=k_{D}$, which results in a benefit of $\beta_{D}(\delta_{D},\delta_{A})= \dfrac{\delta_{A}}{2\delta_{D}}$.
-\\
+For $\delta_{A} + 2(d-k_{D}) > 0$ the benefit will always be positive but decreasing, see figure \ref{ShapeUp}. 
+The defender will always play as fast as he can if $\delta_{A} + 2(d-k_{D}) > 0$ for $k_{D} < d$ or $k_{D} > d$ because $\delta_{A}$ will be positive in either case. \\
 \begin{figure}
@@
 \subsection{Best Responses of $\delta_{D}$}
-The optimum functions will be piecewise functions. We distinguish six cases for different values of $\delta_{A}$ depending on the values of $k_{D}$ and $d$. 
+The optimum functions are piecewise functions. We distinguish three cases for different values of $\delta_{A}$ depending on the values of $k_{D}$ and $d$. 
 %We point out that $\delta_{A}$ and $\delta_{D}$ are positive rates. 
-Figure XXX shows on the left how the three cases can be combined in a single continuous function (shown as the thick line).
+Figure \ref{cost0} shows on the left how the three cases can be combined in a single continuous function (shown as the thick line).
 On the right, the derivative function indicates at which points the benefit will change.
@@
 This implies that there are no sudden changes in benefit for small changes of period.
-
+The points that are interesting to look at are the points of the derivative function that intersect with the x-axis ($\delta_{A}$).
+These points are the roots of the derivative function and show where the benefit function is increasing or decreasing and therefore give the value of the maximum benefit of the defender.
+\begin{figure}[hbtp]
+\centering
+\includegraphics[scale=1]{Images/cost0.pdf} 
+\caption{Benefit function and derivative for a cost $k_{D} = 0$ and with $d (=2)$ and $\delta_{A}=(1.5)$ not equal to zero.}
+\label{cost0}
+\end{figure}
 \subsection*{$\delta_{A} < 2(k_{D} - d)$}
-For case A, the function will always increase if $k_{D} > 0$.
-If $k_{D} < 0$ it means that the cost is negative and we assume a positive cost.
-The function will thus never decrease.
+For case A, the function is increasing if $k_{D} > 0$.
+If $k_{D} < 0$ the function is decreasing but a positive cost is assumed.
+The function will therefore never decrease.
 $k_{D} =0$ is an edge case and will be covered later in the paragraph of edge cases.
 So for case A we can conclude that the function is always increasing.
-In the next case, case B, the benefit function will increase in the interval $[0, 2(k_{D}-d)[$.
+In the next case, case B, the benefit function will increase in the interval $[0, 2(k_{D}-d)]$.
 In case C, for $\delta_{A} < 2(k_{D} - d)$, the benefit function will always increase.
-Because for all the cases we are dealing with a negative benefit function that is increasing, the maximum benefit is achieved for $\delta_{} = \infty$.
+Case B and case C are both increasing and both functions are negative.
+The benefit will never become positive as such, the defender will try to not to play.
+The maximum benefit is achieved for $\delta_{} = \infty$.
 This can be seen in (A) in figure \ref{case1D}.
@@
 For case A, if $k_{D}>0$ the benefit function is always increasing.
-In case B, the value of $\delta_{A} =2(k_{D} - d)$ has to filled in in the root, $\sqrt{2k_{D}\delta_{A} + d^{2}}$.
+In case B, the value of $\delta_{A} =2(k_{D} - d)$ has to be filled in the root, $\sqrt{2k_{D}\delta_{A} + d^{2}}$.
 This gives the value $2k_{D}-d$.
@@
 A maximum is achieved for $\delta_{D}$ equal to the maximum of $\left\lbrace \delta_{A}+d, \sqrt{2k_{D}\delta_{A} + d^{2}}\right\rbrace $.
-If we fill in $\delta_{A} = 2(k_{D} - d)$, the two values are the same so they are both a maximum.
+If we fill in $\delta_{A} = 2(k_{D} - d)$, the two formulas have the same result so they are both a maximum.
 If we fill in the value of $\delta_{A} = 2(k_{D} - d)$ in case C, the benefit will always be 0.
-Case C is the case for all values $\delta_{D} \geq \delta_{A} + d$, so $\beta_{D}=0$ is true for all $\delta_{D} \in [\delta_{A}+d, \infty]$.
-The defender's maximum benefit is thus achieved for all $\delta_{D} \in [\sqrt{2k_{D}\delta_{A} + d^{2}}, \infty]$. See (B) in figure \ref{case1D}.
+Case C is the case for all values $\delta_{D} -d \geq \delta_{A}$, so $\beta_{D}=0$ is true for all $\delta_{D} \in [\delta_{D}-d, \infty]$.
+The defender's maximum benefit is thus achieved for all $\delta_{D} \in [\delta_{D}-d, \infty]$. See (B) in figure \ref{case1D}.
 
@@
 After point $d$ the benefit decreases in all the cases.
-
-\begin{figure}[hbtp]
-\centering
-\includegraphics[scale=1]{Images/cost0.pdf} 
-\caption{Benefit function and derivative for a cost $k_{D} = 0$ and with $d (=2)$ and $\delta_{A}=(1.5)$ not equal to zero.}
-\label{cost0}
-\end{figure}
+So the period of the defender will be equal to $\delta_{D} \in [0,d]$.
 
@@
 
-The benefit formula for the attacker for this case obtained in the previous chapter is as follows:
+The benefit formula for this case corresponds with the first piece of the benefit function of the attacker obtained in the previous chapter:
 \begin{equation}
@@
 \subsection*{Case B: $ \delta_{A} \leq \delta_{D} -d$ }
-The benefit formula for the attacker for this case obtained from the previous chapter is as follows:
+The benefit formula for this case corresponds with the second piece of the benefit function of the attacker obtained in the previous chapter:
 \begin{equation*}
@@
 
-The benefit formula for the attacker for this case obtained in the previous chapter is as follows:
+The benefit formula for this case corresponds with the third piece of the benefit function of the attacker obtained in the previous chapter:
 \begin{equation*}
@@
 Because for this case $d\leq \delta_{D}$, we only look at root $\delta_{D}=d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$.
-$\sqrt{2dk_{A}+k_{A}^{2}}$ is bigger than $k_{A}$ so if we subtract it from $d+k_{A}$, we are in the part where $delta_{D}$ is smaller than d.
+$\sqrt{2dk_{A}+k_{A}^{2}}$ is bigger than $k_{A}$ so if we subtract it from $d+k_{A}$, we are in the part where $\delta_{D}$ is smaller than \textit{d}.
 \\
@@
 For case B, the function is increasing.
-The function has a maximum on on $\delta_{A} = minimum \{\delta_{D} -d, \sqrt{2k_{A}\delta_{D}} \} $.
+The function has a maximum on $\delta_{A} = minimum \{\delta_{D} -d, \sqrt{2k_{A}\delta_{D}} \} $.
 For $\delta_{D} < d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$, the minimum of the two values is $\delta_{D}-d$.\\
@@
 The right functions are the derivatives of the benefit functions on the left.
-The best responses are the values where the function intersects with the x-as.
-(A): $\delta_{D} > d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$, (B): $\delta_{D} = d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$, (C): $\delta_{D} < d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$}.
+The best responses are the values where the function intersects with the x-axis.
+(A): $\delta_{D} < d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$, (B): $\delta_{D} = d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$, (C): $\delta_{D} > d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$}.
 \label{case1A}
@@
 For case B, the function is increasing.
-The function has a maximum on on $\delta_{A} = minimum \{\delta_{D} -d, \sqrt{2k_{A}\delta_{D}} \} $.
+The function has a maximum on $\delta_{A} = minimum \{\delta_{D} -d, \sqrt{2k_{A}\delta_{D}} \} $.
 For $\delta_{D} > d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$, the minimum of the two values is $\delta_{D}-d$.\\
@@
 We can merge this with the previous case where $\delta_{D}=0$ and $k_{A}=0$.
-If $\delta_{D} > d$, the attacker will play.
-He will play with a rate of $\delta_{A} \in [0, \delta_{D} - d ]$.
+If $\delta_{D} > d$, the attacker will play as fast as he can.
+He will play with a rate of $\delta_{A} =0$.
 \subsection*{$\delta_{D}=0$}
@@
      \left[0,\infty\right] & \delta_{D} \leq d ~~\& ~~k_{A}=0 \\
-          \left[0,\delta_{D} - d\right] & \delta_{D} > d ~~\& ~~k_{A}=0 \\
+         0 & \delta_{D} > d ~~\& ~~k_{A}=0 \\
             \infty & \delta_{D} < d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}} \\
@@
 
-The Nash Equilibria can be found as the intersection points of the piecewise functions $opt_{A}(\delta_{D})$ and $opt_{D}(\delta_{A})$.
-For this we have to compare the function in terms of the relationship of the players move cost.
   %We distinguish three cases: $k_{A} < k_{D} , ~k_{A} > k_{D} $ and $k_{A} = k_{D}$.
